### 2018-08-30 LatexNotes/Notes.tex
Added a bunch of files to the repo including the reading list (Literature), a scan of Andys original notes, my typed up Latex notes, my Julia codes for sampling from Gibbs distributions

\documentclass[12pt]{article}

% PACKAGES
\usepackage{graphicx, amsmath, amssymb, amsfonts, mathtools, mathrsfs, color}
\usepackage{comment, enumerate, tabularx}
\usepackage{natbib, hyperref, url}
\usepackage[margin=1in]{geometry}
%\usepackage[justification=RaggedRight]{caption}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% LATEX DEFINITIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
% Derivatives
\newcommand{\pd}[2]    { \frac{\partial #1} {\partial #2} }
\newcommand{\ppd}[2]  { \frac{\partial^2 #1}{{\partial #2}^2} }
\newcommand{\pdi}[2] { {\partial_#2} #1 }
\newcommand{\td}[2] { \frac{d #1} { d #2 } }
\newcommand{\grad}{\nabla}
\newcommand {\Lap} {\grad^2}
% Vectors and operators
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\eps}{\varepsilon}

% Experimental parameters
\newcommand{\lamm}{\lambda^{-}}
\newcommand{\lamp}{\lambda^{+}}
\newcommand{\hm}{h^{-}}
\newcommand{\hp}{h^{+}}

% Theory for truncated KdV and stat mech
\newcommand{\uhat}{\hat{u}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\Real}{\text{Re}}
\newcommand{\Fspace}{\mathscr{F}_{\Lambda}}
\newcommand{\Proj}{P_{\Lambda}}
\newcommand{\sumk}{\sum_{k=1}^{\Lambda}}
\newcommand{\sumn}{\sum_{n=1}^{N}}
\newcommand{\uDir}{u_{\text{Dir}}}
\newcommand{\Gibbs}{\mathcal{G}}

% Dashed integral
\def\Xint#1{\mathchoice
   {\XXint\displaystyle\textstyle{#1}}%
   {\XXint\textstyle\scriptstyle{#1}}%
   {\XXint\scriptstyle\scriptscriptstyle{#1}}%
   {\XXint\scriptscriptstyle\scriptscriptstyle{#1}}%
   \!\int}
\def\XXint#1#2#3{{\setbox0=\hbox{$#1{#2#3}{\int}$}
     \vcenter{\hbox{$#2#3$}}\kern-.5\wd0}}
\def\ddashint{\Xint=}
\def\dashint{\Xint-}
\newcommand{\intt}{\dashint}%_0^{2 \pi}}
\newcommand{\dx}{\, dx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TITLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change}
\author{}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Mathematical Preliminaries
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Mathematical preliminaries}

\subsection{Truncated Fourier series}

We will be dealing with the truncated KdV equation (TKDV). 
Consider a real-valued function $u(x): [0, 2\pi] \to \RR$ given by a truncated Fourier series
\begin{equation}
u(x) = \sum_{k= -\Lambda}^{\Lambda} \uhat_k e^{i k x}	\qquad \text{for } 0 \le x \le 2\pi
\end{equation}
where
\begin{equation}
\uhat_k = \intt u(x) e^{-i k x} \dx
\end{equation}
Above, and throughout the paper, we use the notation
\begin{equation}
\intt \cdot \, \dx  = \frac{1}{2\pi} \int_0^{2 \pi} \cdot \, \dx
\end{equation}
We will be dealing with zero-momentum functions
\begin{equation}
M = \intt u(x) \dx = 0
\end{equation}
Let $\Fspace$ denote the space of all zero-momentum functions that can be represented by the above truncated Fourier Series (with maximum wavenumber $\Lambda$).

Since any element $u \in \Fspace$ is real valued and has zero momentum, it can be written as
\begin{equation}
u(x) = \sumk \uhat_k e^{i k x} + \uhat_k^{*} e^{-i k x} 
= 2 \Real \sumk \uhat_k e^{i k x}
\end{equation}
This implies that $\dim(\Fspace) = \Lambda$ as a complex vector space (or the dimension of the equivalent real vector space is $2\Lambda$).
For any $u \in \Fspace$, we introduce the energy
\begin{equation}
E = \frac{1}{2} \intt \Proj(u^2) \dx = \sumk \abs{\uhat_k}^2
\end{equation}
where $\Proj$ is the projection onto $\Fspace$.
In order to define the Hamiltonian for TKDV, we introduce the following components
\begin{equation}
H_3 = \frac{1}{6} \intt \Proj(u^3) \dx \qquad H_2 = \frac{1}{2} \intt \Proj(u_x^2) \dx
\end{equation}
We note that $H_3$ can be written explicitly in Fourier space as \cite{abramov2003hamiltonian} 
\begin{equation}
\label{H3dirsum}
H_3 = \frac{1}{6} \sum_{\substack{ k_1 + k_2 + k_3 = 0 \\ \abs{k_1}, \abs{k_2}, \abs{k_3} \le \Lambda}}\uhat_{k_1} \uhat_{k_2} \uhat_{k_3}
\end{equation}

\subsection{Some useful estimates/bounds for $H_2$ and $H_3$}

Since we examine the statistical mechanics of TKDV, it is valuable to have some a priori bounds and estimates for the components of the Hamiltonian, $H_2$ and $H_3$. In this section, we frequently appeal to the idea of an equipartition of energy among all the modes. For energy $E_0$, if the equipartition holds, then
\begin{equation}
\abs{u_k} = \sqrt{\frac{E_0}{\Lambda}} \qquad \text{for } k=1,\cdots,\Lambda
\end{equation}
We might expect equipartition of energy to hold, for example, in the case of long-time averages or ensemble averages (appealing to ergodicity).

\subsubsection{Expected value of $H_2$ given $E=E_0$}

First, we consider $H_2$. In Fourier space, $u_x$ is represented as
\begin{equation}
u_x(x) = 2 \Real \sumk i k \uhat_k e^{i k x}
\end{equation}
Thus, for equipartitioned energy, we have
\begin{equation}
H_2 = \sumk \abs{k \uhat_k}^2 = \frac{E_0}{\Lambda} \sumk k^2 = \frac{E_0}{6} (\Lambda+1)(2 \Lambda+1) \sim \frac{E_0 \Lambda^2}{3} \text{for } \Lambda \gg 1
\end{equation}
The above provides a useful estimate for the expected value of $H_2$ (to test the numerics for example).

\subsubsection{Upper bound on $H_3$ given $E=E_0$}

Next, consider $H_3$. By the symmetry of the cubic form, the expected value of $H_3$ is zero. However, to sample from Gibbs measures with non-vanishing inverse temperatures using an acceptance/rejection based algorithm, we will need to estimate the maximum allowed value of $H_3$ on the hypersphere of fixed energy $E = E_0$ so that the acceptance rate can be normalized. Note that the set $E=E_0$ is compact and therefore a maximum is guaranteed to exist. A reasonable guess for function that maximizes $H_3$ is something approximating a Dirac delta function, which naturally motivates the Dirichlet kernel
\begin{equation}
\label{Dkern}
\uDir(x) = \sqrt{\frac{E_0}{\Lambda}}\sumk e^{i k x}
\end{equation}
We can compute $H_3(\uDir)$ by using the direct summation formula \eqref{H3dirsum}. Since $\uhat_k = \sqrt{E_0/\Lambda}$, each term in this sum takes the value $(E_0/\Lambda)^{3/2}$, and so we simply need to count the number of terms present in this sum. A careful counting (being sure to dismiss any terms in which one of $k_1, k_2, k_3$ is zero or one of these terms exceeds $\Lambda$) gives
\begin{equation}
\label{H3Dir}
H_3(\uDir) = \frac{1}{2} E_0^{3/2} (\Lambda^{1/2} - \Lambda^{-1/2})
\end{equation}
Certainly, it must be that
\begin{equation}
\max H_3 \ge H_3(\uDir) 
\end{equation}
My conjecture is that
\begin{equation}
\max H_3 = H_3(\uDir) 
\end{equation}
or at least that this relationship holds asymptotically for $\Lambda \gg 1$.
Preliminary numerical tests seem to support this conjecture for tested values $\Lambda = 6, 8, 10, 20$. In fact, the maximum of $H_3$ encountered over about $10^6$ samples is usually significantly smaller than the value predicted in \eqref{H3Dir}. However, for $\Lambda = 4$, my tests encounter values of $H_3$ slightly larger than the predicted value.

Maximizing $H_3$ is perhaps more straightforward in physical space, where we discretize the interval $[0, 2 \pi)$ with 
\begin{equation}
x_n = \frac{2 \pi (n-1)}{N} \quad \text{for } n=1,\cdots,N
\end{equation}
where
\begin{equation}
N = 2 \Lambda \text{ or  } 2 \Lambda +1
\end{equation}
Both are valid choices for using standard real FFT packages.
%$x_n = 2 \pi (n-1)/N$ for $n=1,\cdots,N$, and $N$ is either $2 \Lambda$ or $2 \Lambda + 1$ (both are valid choices for using standard real FFT packages). 
In physical space, we have
\begin{equation}
\label{H3phys}
H_3 = \frac{1}{6} \intt \Proj (u^3) \dx \approx \frac{1}{6N} \sumn u_n^3
\end{equation}
The second relationship is only an approximation because de-aliasing would be required to make the right-hand side exactly equal to \eqref{H3dirsum}. Nonetheless, let us proceed with the estimate. For fixed energy, $E=E_0$, we require
\begin{equation}
E_0 = \intt \Proj(u^2) \dx = \frac{1}{N} \sumn u_n^2
\end{equation}
(I believe the above holds with no approximation being made due to the discrete Parseval identity)

Clearly, a maximizer of the right-hand-side of \eqref{H3phys} subject to $M=0$ and fixed $E=E_0$ is an approximate Dirac delta function --- i.e. a function with one extreme positive value, and the rest equal negative values,
\begin{align}
& w_1 = \sqrt{2 E_0 (N-1)}		\\
& w_n = -\sqrt{\frac{2 E_0}{N-1}}	\qquad \text{for } n \ne 1
\end{align}
This function is a maximizer because any slight modification (subject to keeping $M=0$ and $E=E_0$) would reduce the value of $H_3$.
Calculating the discrete Fourier transform of this function directly gives
\begin{align}
& \hat{w}_0 = 0	\\
& \hat{w}_k = \sqrt{\frac{2 E_0}{N-1}} \qquad \text{for } k \ne 1
\end{align}
Notice that if we choose $N = 2\Lambda + 1$, then our delta-function in physical space corresponds exactly to the Dirichlet kernel, $w(x) = \uDir(x)$. Now, if we calculate $H_3(w)$ purely in physical space using the approximation in \eqref{H3phys}, we get
\begin{equation}
H_3(w) = \frac{\sqrt{2} E_0^{3/2}}{3} \left( \frac{N-2}{\sqrt{N-1}} \right)
\end{equation}
Taking $N = 2\Lambda + 1$ gives
\begin{equation}
\label{H3physest}
H_3(w) = \frac{2 E_0^{3/2}}{3} \left( \Lambda^{1/2} - \Lambda^{-1/2} \right)
\end{equation}
So while \eqref{H3Dir} and \eqref{H3physest} take very similar forms, they differ by a multiplicative factor: 1/2 in Eq.~\eqref{H3Dir} versus 2/3 in Eq.~\eqref{H3physest}. Therefore, I suppose the error made by failing to de-alias in Eq.~\eqref{H3phys} is not negligible.

So the impasse here as follows: It is straightforward to maximize the approximate form of $H_3$ in physical space (the right-hand-side of \eqref{H3phys}), and justifying this maximization can be made completely rigorous in my mind. However, the approximation made by failing to di-alias is apparently not that small. On the other hand, we can calculate $H_3$ exactly in spectral space, but it is not clear how to maximize $H_3$ subject to the constraints --- I can only make a guess at the maximizer. To overcome this issue, one might naturally want to use the de-aliased form of $H_3$ in physical space (by taking $N = 4 \Lambda$ for example), but the problem here is that all discretized functions are not permissible. Only functions that correspond to a Fourier series truncated at wavenumber $\Lambda$ are permissible.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TKDV
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The truncated KdV equations}

We will be using the truncated KdV equations (TKDV) with variable depth, in particular an abrupt depth change (ADC). The upstream depth is normalized to $1$ and the downstream (dimensionless) depth is $D_0$. Accordingly, the upstream and downstream Hamiltonians are given by
\begin{align}
\label{Hamiltonian}
& H^- = E_0^{1/2} H_3 - H_2 \\
& H^+ = D_0^{-13/4} E_0^{1/2} H_3 - D_0^{3/2} H_2 \\
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Numerical Methods
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}

Here we describe the statistical numerical methods. The deterministic simulations of TKDV are described elsewhere. Below the microstate refers to the spectrum of displacement $\{ \uhat_k \}_{k=1}^{\Lambda}$.

\subsection{Sampling from microcanonical energy distribution}

We first consider sampling microstates from a microcanonical energy distribution with $E=E_0$ fixed and $M=0$ as always,
\begin{equation}
\Gibbs^0 = d\nu \propto \delta(E-E_0) \delta(M)
\end{equation}
This can be considered a canonical distribution in the Hamiltonian with zero inverse temperature (hence the notation $\Gibbs^0$). At times, it is easier to denote the distribution $d\nu$, so we will use both interchangeably. Note this is simply the uniform distribution on the hyper-sphere of $E=E_0$.

To sample from this distribution, we use the following algorithm
\begin{enumerate}
\item Sample each $\uhat_k$ for $k=1,\cdots,\Lambda$ from a normal distribution $N(0,1)$.
\item Compute $E = \sumk \abs{\uhat_k}^2$
\item Normalize each component by multiplying by $\sqrt{E_0/E}$.
\end{enumerate}
This algorithm was proven to converge to the measure $d\nu$ in \cite{abramov2003hamiltonian}. Throughout this paper, we normalize the energy for the sampling distribution to unity. This is the reason why $E_0$ appears in the Hamiltonian in \eqref{Hamiltonian}. Hence, we will always be sampling from
\begin{equation}
\Gibbs^0 \propto \delta(E-1) \delta(M)
\end{equation}

\subsection{Acceptance/rejection algorithm to sample from a mixed microcanonical/canonical distribution.}

Now consider the problem of sampling from a distribution that is microcanonical in energy and canonical in the Hamiltonian
\begin{equation}
\Gibbs^{\pm} \propto \delta(E-1) \delta(M) e^{-\theta^{\pm} H^{\pm}}
\end{equation}
Here, the Hamiltonian can be either the upstream or downstream, $H = H^-$ or $H^+$ and the inverse temperature $\theta$ can also take different values upstream and downstream, $\theta^-$ and $\theta^+$. To sample from this distribution, we use the following acceptance/rejection algorithm.
\begin{enumerate}
\item Sample a large number of microstates from $\Gibbs^0$.
\item For each, compute $H_3$ and $H_2$. Note: this is the most expensive step in the algorithm. We will therefore store all the values of $H_3$ and $H_2$ so they need not be recomputed.
\item For each sample, compute $H^+$ and $H^-$.
\item Assign a provisional acceptance rate as $e^{- \theta H}$ (for either upstream or downstream). The only issue is this acceptance rate is not yet normalized.
\item Find the maximum of the acceptance rate over all samples $m = \max{e^{-\theta H}}$.
\item Use this value to normalize the acceptance rates, $e^{- \theta H}/m$.
\item Decide to accept or reject each microstate by picking a uniform variable $v \in [0,1]$ and testing if $e^{- \theta H}/m < v$.
\item Store the microstates, as well as the values of $H_3$ and $H_2$, that have been accepted.
\end{enumerate}
The above is a conceptual description of the algorithm. In practice, it is implemented in a way to be computational efficient.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Results
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminary results}
We show some preliminary results below. 
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{Hist1}
\caption{Left: Sampling $H^-$ from the incoming Gibbs distribution $G^-$ with three different inverse temperatures $\theta^-$ (see legend). Right: Sampling the $H^+$ from an outgoing distribution $G^+$ with the same inverse temperatures.
In these tests, $\Lambda = 10$, $E_0 = 1$, $\lambda = 1$, $D_0 = 0.6$, and the number of samples is $1 \times 10^7$.}
\label{Hist1}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{Hist2}
\caption{Sensitivity to changes in $D_0$.
Left: Downstream histogram with $D_0 = 0.6$ (same as in Fig.~\ref{Hist1}). Right: Same with $D_0 = 0.4$. Notice the histogram changes significantly.
}
\label{Hist2}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
%Notice that even though the depth ratio changes only slightly, the histograms change significantly. The upstream histograms do not depend on $D_0$ and are therefore not shown.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{Hist3}
\caption{Sensitivity to changes in the energy $E_0$.
Left: Downstream histogram with $E_0 = 1$ (same as in Fig.~\ref{Hist1}). Right: Same with $E_0 = 5$. Notice a slight change in the histogram.}
\label{Hist3}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
 %Notice that the energy changes substantially, but the histograms change only slightly. While the upstream histogram ostensible depends on $E_0$, the changes are nearly imperceptible and therefore not shown.

  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{microup0}
\caption{Sampling the upstream microstates with $\theta^- = 0$ (and $E_0 = 4$). Left: histogram of the surface displacement, $u^-$. Right: ensemble average of the spectrum, $\mean{\abs{\uhat_k}}$. Notice symmetric distribution of the displacement and the equipartition of energy amongst modes.}
\label{microup0}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^

 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{microup3}
\caption{The upstream microstates with $\theta^- = -0.3$ (and $E_0 = 4$ as before). Notice the distinct tilt to the spectrum due to $\theta^- < 0$.}
\label{microup3}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^

  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}[p]%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{microdn3}
\caption{Examination of the downstream microstates with $\theta^+ = -0.3$ and $D_0 = 0.5$ ($E_0 = 4$ as before). Notice the skewness of $u$!!! The spectrum is tilted, though not as much as it is upstream.}
\label{microdn3}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^

\newpage
\newpage
\section{Experimental parameters}

Some relevant parameters from the experiment are given in the table.

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{Some relevant experimental parameters} 
\vspace{0.3 pc}
\begin{tabular}{c l l}
\hline
\hspace{0.5pc} Symbol
\hspace{0.5pc} & Meaning 
\hspace{0.5pc} & Value \\
\hline
N/A		& Length of the experimental tank (in wave direction)	& 6 m	\\
$L_0$	& Width of the experimental tank	 (transverse)		& $L_0 = $ 20 cm	\\
$f_0$	& Dominant forcing frequency in experiments	& $f_0 = $ 2 Hz	\\
$\hm$	& Depth on deep side						& $\hm = 12.5$ cm \\
$\hp$	& Depth on shallow side					& $\hp = 3$ cm \\
$\lamm$	& Wavelength induced by 2 Hz forcing on deep side	& $\lamm = 38$ cm	\\
$\lamp$	& Wavelength induced by 2 Hz forcing on shallow side	& $\lamp = 25$ cm	\\
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

\begin{comment}
% Majda's notes typed up
\section{Deterministic model for waves over abrupt depth change}

The surface-displacement is denoted $u(x,t)$.

Modeling assumptions
\begin{enumerate}
\item Waves in a narrow channel (width $L_0$) are nearly unidirectional and $2 \pi L_0$ periodic (purely a mathematical idealization, not realistic I believe). Where $L_0 < \lambda$, with $\lambda$ a characteristic wavelength. As seen in the table, this later assumption is reasonable since $L_0 / \lamm \approx 0.5$
\item The role of the paddle forcing is to generate a state with `zero momentum', $M = \intt u \, dx = 0$ (this assumption is strongly supported by experimental measurements pointwise!) and with some given energy $E_0 = \frac{1}{2} \intt u^2 \, dx$ (not an assumption, just a statement). Here, $E_0$ is a function of the strength of the paddle forcing.
\item Assume a Galerkin truncation of $u(x,t) \approx u_{\Lambda}(x,t)$, where $\Lambda$ (fixed) will probably be 20 or 30.
\item The `time' coordinate $t$  corresponds physically to a downstream spatial coordinate (i.e.~$X \approx \eps x$ in Johnson's notation). The abrupt depth change (ADC) corresponds to $t = t_0$ or $T_{ADC}$.
\end{enumerate}
\end{comment}

\bibliographystyle{plain}
\bibliography{Notesbib}

\end{document}

### 2018-08-30 SamplingCodes/main.jl

@@ -1,0 +1,34 @@
+include("routines.jl")
+
+#= Compute the two Hamiltonians (upstream and downstream)
+as sampled from the Gibbs0 distribution. =# 
+function main(nmodes::Int, nsamples::Int, invtemp::Float64, 
+		E0::Float64, D0::Float64, suffix::AbstractString)
+	savemicro = true
+	cput0 = time()
+	println("\n\n")
+	# Sample from a Gibbs distribution.
+	H3vec,H2vec,uhacc = gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro)
+	naccsamples = endof(H3vec)
+	# Convert each accepted uhat to physical space for analysis.
+	uacc = getuacc(uhacc)
+	uhavg = getuhavg(uhacc)
+	cputime = time()-cput0
+	# Write Hamiltonian data to a file.
+	println("\nWriting data to output file.")
+	hamlabel = "# Hamiltonian data: nsamples, nmodes, invtemp, E0, D0, H3, H2, uhavg"
+	hamdata = [hamlabel; naccsamples; nmodes; invtemp; E0; D0; H3vec; H2vec; uhavg]
+	hamfile = string("../SamplingData/hamdata",suffix,".txt")
+	writedata(hamdata, hamfile)
+	# Write uacc data to a file.
+	#udata = [uacc [real(uhacc); imag(uhacc)] ]
+	ufile = string("../SamplingData/udata",suffix,".txt")
+	writedata(uacc, ufile)
+	# Print the final CPU time.
+	println("The CPU time is ", signif(cputime/60,2), " minutes.")
+	return
+end
+
+main(10, 1*10^6, -0.0, 4., 1.0, "up")
+#main(10, 2*10^7, -0.3, 4., 1.0, "up")	#0.078%
+#main(10, 2*10^7, -0.3, 4., 0.5, "dn") 	#0.074%

### 2018-08-30 SamplingCodes/misc.jl

@@ -1,0 +1,121 @@
+
+
+#= Might still finish.
+Find indices corresponding to the largest values of H3 
+and smallest values of H2.=#
+function findind(H3::Vector{Float64},H2::Vector{Float64})
+	# FILL IN
+	return H3ind, H2ind
+end
+
+
+
+#----- GOOD ROUTINES THAT WORKED BUT THAT HAVE BEEN SUPERSEDED ----#
+# OBSELETE since I wrote gibbs for arbitrary invers temperature.
+#= Sample from a microcanonical distribution with fixed energy E=1.0,
+i.e. a Gibbs distribution with zero inverse temperature. =#
+function gibbs0(nmodes::Int, nsamples::Int)
+	# Get all the random samples at once.
+	rvar = randn(nmodes,2,nsamples)
+	H3 = zeros(Float64,nsamples)
+	H2 = zeros(Float64,nsamples)
+	# Compute H3 and H2 for each microstate. This is the bottleneck.
+	# TO DO: parallelize this for loop!!!
+	for nn=1:nsamples
+		uhat = getuhat(rvar,nn)
+		H3[nn] = ham3(uhat)
+		H2[nn] = ham2(uhat)
+		if mod(nn, 10^5) == 0
+			println("Computation is ", signif(100*nn/nsamples,3), "% completed.")
+		end
+	end
+	return H3, H2
+end
+# OBSELETE
+#= Compute H3 using direction summation in spectral space. =#
+function ham3sum(uhat::Vector{Complex128})
+	nmodes = endof(uhat)
+	H3 = 0.
+	# Compute H3 directly with a double sum over k1 and k2.
+	for k1 = 1:nmodes
+		for k2 = 1:nmodes
+			#= Case: k1 and k2 have same sign. 
+			Consider k1,k2>0, then add CC via 2*Real =#
+			k3 = k1 + k2
+			if k3 <= nmodes
+				H3 += 2*real(uhat[k1]*uhat[k2]*conj(uhat[k3]))
+			end
+			#= Case: k1 and k2 have opposite signs.
+			Consider k1>0 & k2<0, then add CC via 2*Real =#
+			k3 = k2-k1
+			if (abs(k3) <= nmodes)
+				# Case abs(k2)>abs(k1)
+				if k3 > 0
+					H3 += 2*real(uhat[k1]*conj(uhat[k2])*uhat[k3])
+				# Case abs(k1)>abs(k2)
[…]

### 2018-08-30 SamplingCodes/routines.jl

@@ -1,0 +1,165 @@
+using Winston
+using Distributions
+
+#---------- IO Routines ----------#
+function readvec(file::AbstractString)
+	iostream = open(file, "r")
+	vec1 = readdlm(iostream)[:,1]
+	close(iostream)
+	return vec1
+end
+function writedata(data::Array, filename::AbstractString)
+	iostream = open(filename, "w")
+	writedlm(iostream, data)
+	close(iostream)
+	return
+end
+#---------------------------------#
+
+#---------- Real FFT Routines ----------#
+#= Compute the FFT between physical and spectral space assuming
+1) the signal uu is real and 2) uu has zero mean (i.e. momentum M=0)
+Note: uu has length npoints = 2*nmodes. =#
+#= Basic realfft to go from uu to uhat. =#
+function realfft(uu::Vector{Float64})
+	uhat = rfft(uu)/endof(uu)
+	assert(abs(uhat[1])/maxabs(uhat) < 1e-6) 
+	return uhat[2:end]
+end
+#= Basic irealfft to go from uhat to uu (no upsampling). =#
+function irealfft(uhat::Vector{Complex128})
+	nmodes = endof(uhat)
+	uu = irfft([0; uhat], npoints(nmodes))
+	return uu*endof(uu)
+end
+#= Upsampled version of irealfft. =#
+function ifftup(uhat::Vector{Complex128})
+	uhat = [uhat; zeros(eltype(uhat), endof(uhat))]
+	return irealfft(uhat)
+end
+#= Valid choices for npts are 2*nmodes or 2*nmodes+1, 
+but I need a function to keep my usage consistent.
+Note: for H3test npts = 2*modes is slightly more accurate. =#
+function npoints(nmodes::Int)
+	return 2*nmodes
+end
+#---------------------------------------#
+
+#---------- Hamiltonian Routines ----------#
+#= In all routines, the physical signal is assumed to be real 
+with zero momentum, M = int u dx = 0. Therefore, the Fourier transform 
+only requires the modes k=1,...,nmodes. The negative modes are given 
+by the complex conjugates (CC), and the k=0 mode vanishes due to M=0. 
+I used the shorthand convention int = (1/2*pi) int_0^{2*pi}. 
+and nmodes is the same as Lambda. =#
[…]

### 2018-08-30 SamplingCodes/tests.jl

@@ -1,0 +1,207 @@
+include("routines.jl")
+
+# Simple test for gibbs
+function gibbs0test()
+	Lambda = 8
+	etah = gibbs0(Lambda)
+	Ecomp = energy(etah)
+	E0 = 1.0
+	reldiff = abs((E0-Ecomp)/E0)
+	println("\nGibbs0 test")
+	println("The desired energy is E0 = ", signif(E0,3))
+	println("The energy of etahat is ", signif(Ecomp,3))
+	println("The relative difference is ", signif(reldiff,3))
+end
+
+#= Test for ham3 and ham3sum.
+If eta = 2*a1*cos(x) - 2*b1*sin(x) + 2*a2*cos(2x) - 2*b1*sin(2x)
+then eta0 = 0, eta1 = a1 + im*b1, eta2 = a2 + im*b2 
+and H3 = a1^2*a2 + 2*a1*b1*b2 - a2*b1^2 =#
+function H3test()
+	# Parameters
+	nmodes = 8
+	a1 = 4.3
+	b1 = -1.2
+	a2 = 2.3
+	b2 = 4.5
+	# Calculations
+	H3exact = a1^2*a2 + 2*a1*b1*b2 - a2*b1^2
+	etah = zeros(Complex128,nmodes)
+	etah[1] = a1 + im*b1
+	etah[2] = a2 + im*b2
+	H3comp = ham3(etah)
+	H3sum = ham3sum(etah)
+	H3err = abs((H3exact-H3comp)/H3exact)
+	H3sumerr = abs((H3exact-H3sum)/H3exact)
+	# Print statements
+	println("\nH3 test")
+	println("The exact H3 is ", signif(H3exact,4))
+	println("H3 computed by fft to physical space is ", signif(H3comp,4))
+	println("H3 computed by direct summation in spectral space is ", signif(H3sum,4))
+	println("Relative error for fft: ", signif(H3err,3))
+	println("Relative error for direct: ", signif(H3sumerr,3))
+	return
+end
+#= Test for ham2 .=#
+function H2test()
+	# Parameters
+	nmodes = 4
+	a1 = 0.1
+	b1 = -1.2
+	a2 = -0.4
+	b2 = 0.8
+	# Calculations
+	H2exact = a1^2 + b1^2 + 4*a2^2 + 4*b2^2
+	etah = zeros(Complex128,nmodes)
+	etah[1] = a1 + im*b1
[…]

### 2018-08-30 SamplingData/GibbsGeneral.tank/Attachments/rowmean/rowmean/main.cpp

@@ -1,0 +1,66 @@
+// #include "DTSource.h"
+#include "DTSaveError.h"
+
+#include "DTArguments.h"
+#include "DTDataFile.h"
+#include "DTDictionary.h"
+#include "DTDoubleArray.h"
+#include "DTProgress.h"
+#include "DTSeriesNumberList.h"
+
+//////////////////////////////////////////////////////////////////////////////
+//    Main routine
+//////////////////////////////////////////////////////////////////////////////
+
+void Computation(const DTDoubleArray &uhmag,DTSeriesNumberList &computed);
+
+int main(int argc,const char *argv[])
+{
+    DTSetArguments(argc,argv);
+
+    DTDataFile inputFile("Input.dtbin",DTFile::ReadOnly);
+    DTDataFile outputFile("Output.dtbin",DTFile::NewReadWrite);
+    // Input variables.
+    DTDoubleArray uhmag = inputFile.ReadDoubleArray("uhmag");
+
+    // Output series.
+    DTSeriesNumberList computed(outputFile,"Var");
+    if (DTArgumentIncludesFlag("saveInput")) { // Add -saveInput to the argument list to save the input in the output file.
+        WriteOne(outputFile,"uhmag",uhmag);
+    }
+
+
+    // The computation.
+    clock_t t_before = clock();
+    Computation(uhmag,computed);
+    clock_t t_after = clock();
+    double exec_time = double(t_after-t_before)/double(CLOCKS_PER_SEC);
+
+    // The execution time.
+    outputFile.Save(exec_time,"ExecutionTime");
+    outputFile.Save("Real Number","Seq_ExecutionTime");
+
+    // The errors.
+    DTSaveError(outputFile,"ExecutionErrors");
+    outputFile.Save("StringList","Seq_ExecutionErrors");
+
+    outputFile.SaveIndex();
+
+    return 0;
+}
+
+//////////////////////////////////////////////////////////////////////////////
+//    Computational routine
+//////////////////////////////////////////////////////////////////////////////
+
+void Computation(const DTDoubleArray &uhmag,DTSeriesNumberList &computed)
+{
+    // Insert your code here.
+
+    DTProgress progress;
+
+    // Inside the loop, do
+    //     progress.UpdatePercentage(fraction);
+    //     computed.Add(returnStructure,time); // Call with time>=0 and strictly increasing.
+
+}

### 2018-08-30 KdV_SpectralCodes/SolverKdV_SpectralRK4.m
pseudo-spectral code for KdV

@@ -1,0 +1,139 @@
+% This script solves KdV equation in a periodic domain
+% u_t=-6uu_x-u_xxx, -L<=x<=L
+% using pseudo-spectral & explicit and implicit RK4 scheme (non-symplectic)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Set simulation parameters
+J = 256;      % Number of points in each direction, J devidable by 4
+dt = 1E-3;    % initial time step size
+Nt = 10e3;    % Number of time steps
+ulim = 1.5E5; % if any u > ulim, simulation stops
+Ls = 20;       % half domain size
+
+% Put useful stuff into a struct
+params = struct('Ls',Ls, 'J',J,'dt',dt, 'Nt',Nt);
+x=(-Ls:2*Ls/J:Ls-2*Ls/J)';
+
+% Set up implicit dispersion
+k = (pi/Ls)*[0:J/2 -J/2+1:-1]'; % wavenumbers
+Lk = 1i*k.^3;
+Dk = 1i*k;
+
+% Initialize
+t = 0;
+% rng(15); u0 = randn(J,1); 
+u0 = 2*sech(x).^2;
+uk0 = fft(u0); uk0(J/2+1)=0;
+uk = uk0;
+
+figure(10)
+plot(x,u0); ylim([-.5 2.5]);
+title('solution of KdV equation, non-symplectic')
+
+% Diagnostics
+countDiag = 10;  % Compute diagnostics every countDiag steps
+T = zeros(1,Nt/countDiag);
+ut = zeros(J,Nt/countDiag);
+mass = zeros(1,Nt/countDiag);
+energy = zeros(1,Nt/countDiag);
+hamiltonian = zeros(1,Nt/countDiag);
+
+% adaptive stepping stuff:
+tol= 1E-1;
+r0 = .8*tol;
+
+% Main loop 
+% tic;
+for ii=1:Nt
+    if mod(ii,countDiag)==0
+        if any(isnan(uk(:))), break, end
+        T(ii/countDiag)=t;
+        ut(:,ii/countDiag)=u;
+        
+        mass(ii/countDiag)=sum(u)*2*Ls/J;
+        energy(ii/countDiag) = sum(abs(uk).^2)*2*Ls/J^2; 
[…]

### 2018-08-30 KdV_SpectralCodes/SolverKdV_SymplecticGRK4.m

@@ -1,0 +1,116 @@
+% This script solves KdV equation in a periodic domain
+% u_t=-6uu_x-u_xxx, -L<=x<=L
+% using pseudo-spectral & symplectic Generalized Runge-Kutta
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Set simulation parameters
+J = 256;      % Number of points in each direction, J devidable by 4
+dt = 1E-3;    % initial time step size
+Nt = 10e3;    % Number of time steps
+ulim = 1.5E5; % if any u > ulim, simulation stops
+Ls = 20;       % half domain size
+A=[1/4,1/4-sqrt(3)/6;1/4+sqrt(3)/6,1/4];
+
+% Put useful stuff into a struct
+params = struct('Ls',Ls, 'J',J,'dt',dt, 'Nt',Nt,'A',A);
+x=(-Ls:2*Ls/J:Ls-2*Ls/J)';
+
+% Set up implicit dispersion
+k = (pi/Ls)*[0:J/2 -J/2+1:-1]'; % wavenumbers
+Lk = 1i*k.^3;
+Dk = 1i*k;
+
+% Initialize
+t = 0;
+% rng(15); u0 = randn(J,1); 
+u0 = 2*sech(x).^2;
+uk0 = fft(u0); uk0(J/2+1)=0;
+uk = uk0;
+ukm1 = uk0;  % at time n-1
+ukm2 = uk0;  % at time n-2
+% interpolatin function for the initial iteration input
+interp = @(u0,um1,um2,w) (.5*w*(w+3)+1).*u0 - w*(w+2).*um1 +.5*w*(w+1).*um2;
+
+figure(10)
+plot(x,u0); ylim([-.5 2.5]);
+title('solution of KdV equation, symplectic')
+
+% Diagnostics
+countDiag = 10;  % Compute diagnostics every countDiag steps
+T = zeros(1,Nt/countDiag);
+ut = zeros(J,Nt/countDiag);
+mass = zeros(1,Nt/countDiag);
+energy = zeros(1,Nt/countDiag);
+hamiltonian = zeros(1,Nt/countDiag);
+
+
+% Main loop 
+% tic;
+for ii=1:Nt
+    if mod(ii,countDiag)==0
+        if any(isnan(uk(:))), break, end
+        T(ii/countDiag) = t;
[…]

### 2018-08-30 KdV_SpectralCodes/SolverKdV_SymplecticM4a.m

@@ -1,0 +1,130 @@
+% This script solves KdV equation in a periodic domain
+% u_t=-6uu_x-u_xxx, -L<=x<=L
+% using pseudo-spectral & symplectic M4a integrator
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Set simulation parameters
+J = 256;      % Number of points in each direction, J devidable by 4
+dt = 1E-3;    % initial time step size
+Nt = 10e3;    % Number of time steps
+ulim = 1.5E5; % if any u > ulim, simulation stops
+Ls = 20;       % half domain size
+w1=(2+2^(1/3)+2^(-1/3))/3;
+w2=1-2*w1;     % integration fractional time steps
+
+% Put useful stuff into a struct
+params = struct('Ls',Ls, 'J',J,'dt',dt, 'Nt',Nt);
+x=(-Ls:2*Ls/J:Ls-2*Ls/J)';
+
+% Set up implicit dispersion
+k = (pi/Ls)*[0:J/2 -J/2+1:-1]'; % wavenumbers
+Lk = 1i*k.^3;
+Dk = 1i*k;
+
+% Initialize
+t = 0;
+% rng(15); u0 = randn(J,1); 
+u0 = 2*sech(x).^2;
+uk0 = fft(u0); uk0(J/2+1)=0;
+uk = uk0;
+ukm1 = uk0;  % at time n-1
+ukm2 = uk0;  % at time n-2
+% interpolatin function for the initial iteration input
+interp = @(u0,um1,um2,w) (.5*w*(w+3)+1).*u0 - w*(w+2).*um1 +.5*w*(w+1).*um2;
+
+figure(10)
+plot(x,u0); ylim([-.5 2.5]);
+title('solution of KdV equation, symplectic')
+
+% Diagnostics
+countDiag = 10;  % Compute diagnostics every countDiag steps
+T = zeros(1,Nt/countDiag);
+ut = zeros(J,Nt/countDiag);
+mass = zeros(1,Nt/countDiag);
+energy = zeros(1,Nt/countDiag);
+hamiltonian = zeros(1,Nt/countDiag);
+
+
+% Main loop 
+% tic;
+for ii=1:Nt
+    if mod(ii,countDiag)==0
+        if any(isnan(uk(:))), break, end
+        T(ii/countDiag)=t;
[…]

### 2018-08-30 KdV_SpectralCodes/iterativeKdV_GRK4.m

@@ -1,0 +1,45 @@
+function [uk1,uk2,ik]=iterativeKdV_GRK4(uk10,uk20, uk,p)
+% Function iterates to solve the solution for Z-Y=tau/2*F(Z)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+dt=p.dt; A=p.A;
+
+k = (pi/p.Ls)*[0:p.J/2 -p.J/2+1:-1]'; % wavenumbers
+Lk = 1i*k.^3;
+Dk = 1i*k;
+tol=5e-10; % iteration tolerance
+
+uk1=uk10; uk1p=uk1;
+uk2=uk20; uk2p=uk2;
+err1p=0; err2p=0;
+Ni=100;  % maximum interation steps
+for ll=1:Ni
+    u2k1 = u2k_dealiasing(uk1,p);
+    u2k2 = u2k_dealiasing(uk2,p);
+    M1 = 1./((1-dt*A(1,1).*Lk).*(1-dt*A(2,2).*Lk) - (dt*A(2,1).*Lk).*(dt*A(1,2).*Lk));
+    M2 = 1./((1-dt*A(2,2).*Lk).*(1-dt*A(1,1).*Lk) - (dt*A(1,2).*Lk).*(dt*A(2,1).*Lk));
+    
+    uk1 = M1.*(uk-3*dt*A(1,1).*Dk.*u2k1-3*dt*A(1,2).*Dk.*u2k2).*(1-dt*A(2,2).*Lk) ...
+         +M1.*(uk-3*dt*A(2,1).*Dk.*u2k1-3*dt*A(2,2).*Dk.*u2k2).*(dt*A(1,2).*Lk);
+    uk2 = M2.*(uk-3*dt*A(2,1).*Dk.*u2k1-3*dt*A(2,2).*Dk.*u2k2).*(1-dt*A(1,1).*Lk) ...
+         +M2.*(uk-3*dt*A(1,1).*Dk.*u2k1-3*dt*A(1,2).*Dk.*u2k2).*(dt*A(2,1).*Lk);
+    
+    
+    err1 = sum(abs(uk1-uk1p).^2);
+    err2 = sum(abs(uk2-uk2p).^2);
+    if sqrt(err1)<tol && sqrt(err2)<tol && err1+err2>=err1p+err2p
[…]

### 2018-08-30 KdV_SpectralCodes/iterativeKdV_M4a.m

@@ -1,0 +1,31 @@
+function [zk,ik]=iterativeKdV_M4a(zk0, tau,yk,p)
+% Function iterates to solve the solution for Z-Y=tau/2*F(Z)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+k = (pi/p.Ls)*[0:p.J/2 -p.J/2+1:-1]'; % wavenumbers
+Lk = 1i*k.^3;
+Dk = 1i*k;
+tol=5e-10; % iteration tolerance
+
+zk1=zk0;
+zk2=zk0;
+err0=0;
+Ni=100;  % maximum interation steps
+for ll=1:Ni
+    M = 1./(1-tau/2*Lk);
+    zk2=M.*(yk-tau/2*3*Dk.*u2k_dealiasing(zk1,p));
+    
+    err=sum(abs(zk2-zk1).^2);
+    if err<tol && err>=err0
+%         display(['iteration steps N = ',num2str(ll)]);
+        ik=0;
+        break;
+    end
+    err0=err;
+    zk1=zk2;
+    if ll==Ni
+        disp(['maximum interation N = ',num2str(ll),' reached!']);
+        ik=1;
+    end
+end
+zk=zk2;

### 2018-09-03 SamplingCodes/main.jl
NM

@@ -30,5 +30,5 @@ function main(nmodes::Int, nsamples::Int, invtemp::Float64,
 end
 
 main(10, 1*10^6, -0.0, 4., 1.0, "up")
-#main(10, 2*10^7, -0.3, 4., 1.0, "up")	#0.078%
-#main(10, 2*10^7, -0.3, 4., 0.5, "dn") 	#0.074%
+#main(10, 2*10^7, -0.3, 4., 1.0, "up")
+#main(10, 2*10^7, -0.3, 4., 0.5, "dn")

### 2018-09-03 SamplingCodes/routines.jl

@@ -1,4 +1,3 @@
-using Winston
 using Distributions
 
 #---------- IO Routines ----------#
@@ -161,5 +160,4 @@ function getuhavg(uhacc::Array{Complex128})
 	return uhavg
 end
 #---------------------------------------#
-
 

### 2018-09-03 SamplingCodes/main.jl
Touching up small things.

@@ -5,19 +5,25 @@ as sampled from the Gibbs0 distribution. =#
 function main(nmodes::Int, nsamples::Int, invtemp::Float64, 
 		E0::Float64, D0::Float64, suffix::AbstractString)
 	savemicro = true
-	cput0 = time()
+	### cput0 = time()
 	println("\n\n")
 	# Sample from a Gibbs distribution.
-	H3vec,H2vec,uhacc = gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro)
-	naccsamples = endof(H3vec)
+	cputime = @elapsed( (H3vec,H2vec,uhacc) = gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro) )
+	naccept = endof(H3vec)
+	acceptrate = signif(100*naccept/nsamples,2)
+	cputime = signif(cputime/60,2)
 	# Convert each accepted uhat to physical space for analysis.
 	uacc = getuacc(uhacc)
 	uhavg = getuhavg(uhacc)
-	cputime = time()-cput0
+	### cputime = time()-cput0
 	# Write Hamiltonian data to a file.
 	println("\nWriting data to output file.")
-	hamlabel = "# Hamiltonian data: nsamples, nmodes, invtemp, E0, D0, H3, H2, uhavg"
-	hamdata = [hamlabel; naccsamples; nmodes; invtemp; E0; D0; H3vec; H2vec; uhavg]
+	label1 = "# Input parameters: nmodes, nsamples, invtemp, E0, D0"
+	label2 = "# Calculated parameters: number of accepted samples, acceptance rate (%), CPU time (minutes)"
+	label3 = "# Computed data: H3, H2, uhavg"
+	hamdata = [label1; nmodes; nsamples; invtemp; E0; D0;
+			label2; naccept; acceptrate; cputime;
+			label3; H3vec; H2vec; uhavg]
 	hamfile = string("../SamplingData/hamdata",suffix,".txt")
 	writedata(hamdata, hamfile)
 	# Write uacc data to a file.
@@ -25,10 +31,10 @@ function main(nmodes::Int, nsamples::Int, invtemp::Float64,
 	ufile = string("../SamplingData/udata",suffix,".txt")
 	writedata(uacc, ufile)
 	# Print the final CPU time.
-	println("The CPU time is ", signif(cputime/60,2), " minutes.")
[…]

### 2018-09-03 SamplingCodes/routines.jl

@@ -22,7 +22,7 @@ Note: uu has length npoints = 2*nmodes. =#
 #= Basic realfft to go from uu to uhat. =#
 function realfft(uu::Vector{Float64})
 	uhat = rfft(uu)/endof(uu)
-	assert(abs(uhat[1])/maxabs(uhat) < 1e-6) 
+	assert(abs(uhat[1])/maximum(abs,uhat) < 1e-6) 
 	return uhat[2:end]
 end
 #= Basic irealfft to go from uhat to uu (no upsampling). =#
@@ -155,7 +155,7 @@ function getuhavg(uhacc::Array{Complex128})
 	nmodes = size(uhacc)[1]
 	uhavg = zeros(Float64,nmodes)
 	for kk=1:nmodes
-		uhavg[kk] = mean(abs(uhacc[kk,:]))
+		uhavg[kk] = mean(abs.(uhacc[kk,:]))
 	end
 	return uhavg
 end

### 2018-09-03 SamplingCodes/main.jl
Modifying codes to sample from both upstream and downstream distributions in same pass.

@@ -2,39 +2,48 @@ include("routines.jl")
 
 #= Compute the two Hamiltonians (upstream and downstream)
 as sampled from the Gibbs0 distribution. =# 
-function main(nmodes::Int, nsamples::Int, invtemp::Float64, 
-		E0::Float64, D0::Float64, suffix::AbstractString)
+function main(nmodes::Int, nsamples::Int, invtemp::Float64, E0::Float64, D0::Float64)
 	savemicro = true
-	### cput0 = time()
 	println("\n\n")
 	# Sample from a Gibbs distribution.
-	cputime = @elapsed( (H3vec,H2vec,uhacc) = gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro) )
-	naccept = endof(H3vec)
-	acceptrate = signif(100*naccept/nsamples,2)
+	cputime = @elapsed( (H3up,H2up,uhup,H3dn,H2dn,uhdn) = 
+				gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro) )
 	cputime = signif(cputime/60,2)
-	# Convert each accepted uhat to physical space for analysis.
-	uacc = getuacc(uhacc)
-	uhavg = getuhavg(uhacc)
-	### cputime = time()-cput0
-	# Write Hamiltonian data to a file.
-	println("\nWriting data to output file.")
-	label1 = "# Input parameters: nmodes, nsamples, invtemp, E0, D0"
-	label2 = "# Calculated parameters: number of accepted samples, acceptance rate (%), CPU time (minutes)"
-	label3 = "# Computed data: H3, H2, uhavg"
-	hamdata = [label1; nmodes; nsamples; invtemp; E0; D0;
-			label2; naccept; acceptrate; cputime;
-			label3; H3vec; H2vec; uhavg]
-	hamfile = string("../SamplingData/hamdata",suffix,".txt")
-	writedata(hamdata, hamfile)
-	# Write uacc data to a file.
-	#udata = [uacc [real(uhacc); imag(uhacc)] ]
-	ufile = string("../SamplingData/udata",suffix,".txt")
-	writedata(uacc, ufile)
+
+	# Function to save the data.
+	function save_ham_data(H3::Vector{Float64}, H2::Vector{Float64}, 
+			uhat::Vector{Float64}, suffix::AbstractString)
+		# Calculate the acceptance rate.
+		naccept = endof(H3)
[…]

### 2018-09-03 SamplingCodes/routines.jl

@@ -82,10 +82,9 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 	# Get all the random samples at once.
 	rvar = randn(nmodes,2,nsamples)
 	# Allocate space.
-	H3all = zeros(Float64,nsamples)
-	H2all = zeros(Float64,nsamples)
-	aratevec = zeros(Float64,nsamples)
-	maxaccept = 0.
+	H3all, H2all = [zeros(Float64,nsamples) for nn=1:2]
+	aupvec, adnvec = [zeros(Float64,nsamples) for nn=1:2]
+	maxup, maxdn = 0.,0.
 	# Compute H3 and H2 for each to estimate the acceptance rate.
 	# TO DO: parallelize this for loop!!! It is the bottleneck!
 	for nn=1:nsamples
@@ -93,30 +92,35 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 		H3 = ham3(uhat)
 		H2 = ham2(uhat)
 		# Compute the Hamiltonian (for upstream D0=1)
-		ham = D0^(-13/4)*sqrt(E0)*H3 - D0^(3/2)*H2
+		hamup = sqrt(E0)*H3 - H2
+		hamdn = D0^(-13/4)*sqrt(E0)*H3 - D0^(3/2)*H2
 		# Compute the unnormalized acceptance rate and maximize it.
-		acceptrate = exp(-invtemp * ham)
-		maxaccept = max(maxaccept, acceptrate)
+		accept_up = exp(-invtemp * hamup)
+		accept_dn = exp(-invtemp * hamdn)
+		maxup = max(maxup, accept_up)
+		maxdn = max(maxdn, accept_dn)
 		# Save H3 and H2 so as to not recompute.
 		H3all[nn] = H3
 		H2all[nn] = H2
-		aratevec[nn] = acceptrate
+		aupvec[nn] = accept_up
+		adnvec[nn] = accept_dn
 		# Print progress.
 		if mod(nn, 10^4) == 0
 			println("Initial sampling is ", signif(100*nn/nsamples,3), "% completed.")
 		end
 	end
 	# Now that the normalization is known, accept or reject each.
-	H3acc = zeros(Float64,0)
-	H2acc = zeros(Float64,0)
-	savemicro? uhacc = zeros(Complex128,nmodes,nsamples):0
-	count = 0
-	for nn=1:nsamples
[…]

### 2018-09-03 SamplingCodes/main.jl
Reverting back to old way. Found a problem with the new idea.

@@ -2,48 +2,39 @@ include("routines.jl")
 
 #= Compute the two Hamiltonians (upstream and downstream)
 as sampled from the Gibbs0 distribution. =# 
-function main(nmodes::Int, nsamples::Int, invtemp::Float64, E0::Float64, D0::Float64)
+function main(nmodes::Int, nsamples::Int, invtemp::Float64, 
+		E0::Float64, D0::Float64, suffix::AbstractString)
 	savemicro = true
+	### cput0 = time()
 	println("\n\n")
 	# Sample from a Gibbs distribution.
-	cputime = @elapsed( (H3up,H2up,uhup,H3dn,H2dn,uhdn) = 
-				gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro) )
+	cputime = @elapsed( (H3vec,H2vec,uhacc) = gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro) )
+	naccept = endof(H3vec)
+	acceptrate = signif(100*naccept/nsamples,2)
 	cputime = signif(cputime/60,2)
-
-	# Function to save the data.
-	function save_ham_data(H3::Vector{Float64}, H2::Vector{Float64}, 
-			uhat::Vector{Float64}, suffix::AbstractString)
-		# Calculate the acceptance rate.
-		naccept = endof(H3)
-		acceptrate = signif(100*naccept/nsamples,2)
-		# Convert each accepted uhat to physical space for analysis.
-		uu = getuacc(uhat)
-		uhavg = getuhavg(uhat)
-		# Write Hamiltonian data to a file.
-		println("\nWriting data to output file.")
-		label0 = string("# ", suffix, "stream data")
-		label1 = "# Input parameters: nmodes, nsamples, invtemp, E0, D0"
-		label2 = "# Calculated parameters: number of accepted samples, acceptance rate (%), CPU time (minutes)"
-		label3 = "# Computed data: H3, H2, uhavg"
-		hamdata = [label0; label1; nmodes; nsamples; invtemp; E0; D0;
-				label2; naccept; acceptrate; cputime;
-				label3; H3vec; H2vec; uhavg]
-		hamfile = string("../SamplingData/hamdata",suffix,".txt")
-		writedata(hamdata, hamfile)
-		# Write uacc data to a file.
-		#udata = [uacc [real(uhacc); imag(uhacc)] ]
[…]

### 2018-09-03 SamplingCodes/routines.jl

@@ -111,14 +111,15 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 	end
 	# Now that the normalization is known, accept or reject each.
 	H3accup, H2accup, H3accdn, H2accdn = [zeros(Float64,0) for nn=1:4]
-	if savemicro
-		uhaccup,uhaccdn = [zeros(Complex128,nmodes,nsamples) for nn=1:2]
-	end
+
+	savemicro? uhacc = zeros(Complex128,nmodes,nsamples):0 # CHECK
+
 	countup = [0]
 	countdn = [0]
+
 	# Little function to accept or reject each sample.
 	function accept!(accept_rate::Float64, nn::Int, count::Array{Int}, 
-			H3acc::Vector{Float64}, H2acc::Vector{Float64}, uhacc::Array{Complex128})
+			H3acc::Vector{Float64}, H2acc::Vector{Float64}, uhacc::Vector{Float64})
 		univar = rand()
 		if univar <= accept_rate
 			count += 1
@@ -130,9 +131,8 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 				uhat = getuhat(rvar,nn)
 				uhacc[:,count] = uhat[:]
 			end
-		end
-	end
-	# Loop through the samples and accept or reject each.
+		end	
+
 	for nn=1:nsamples
 		# The normalized acceptance rates.
 		accept_up = aupvec[nn]/maxup
@@ -140,25 +140,32 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 		# Accept or reject each sample.
 		accept!(accept_up, nn, countup, H3accup, H2accup, uhaccup)
 		accept!(accept_dn, nn, countdn, H3accdn, H2accdn, uhaccdn)
+
 		# Accept or reject based on a uniform random variable.
 		univarup = rand()
 		univardn = rand()
+
+		if univar <= accept_up
+			countup += 1
+			# Save H3 and H2
+			push!(H3accup, H3all[nn])
+			push!(H2accup, H2all[nn])
+			# Also save the microstate if requested.
+			if savemicro
+				uhat = getuhat(rvar,nn)
+				uhacc[:,count] = uhat[:]
+			end
+		end
 		# Print progress.
 		if mod(nn, 10^4) == 0
 			println("Acceptance/rejection loop is ", signif(100*nn/nsamples,3), "% completed.")
[…]

### 2018-09-03 SamplingCodes/main.jl
Reverting

@@ -5,33 +5,28 @@ as sampled from the Gibbs0 distribution. =#
 function main(nmodes::Int, nsamples::Int, invtemp::Float64, 
 		E0::Float64, D0::Float64, suffix::AbstractString)
 	savemicro = true
-	### cput0 = time()
 	println("\n\n")
 	# Sample from a Gibbs distribution.
-	cputime = @elapsed( (H3vec,H2vec,uhacc) = gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro) )
-	naccept = endof(H3vec)
-	acceptrate = signif(100*naccept/nsamples,2)
-	cputime = signif(cputime/60,2)
+	cputime = @elapsed (H3vec,H2vec,uhacc) = gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro)
+	cputime = signif(cputime/60,2)	
+	naccsamples = endof(H3vec)
 	# Convert each accepted uhat to physical space for analysis.
 	uacc = getuacc(uhacc)
 	uhavg = getuhavg(uhacc)
-	### cputime = time()-cput0
 	# Write Hamiltonian data to a file.
 	println("\nWriting data to output file.")
+	hamfile = string("../SamplingData/ham",suffix,".txt")
 	label1 = "# Input parameters: nmodes, nsamples, invtemp, E0, D0"
 	label2 = "# Calculated parameters: number of accepted samples, acceptance rate (%), CPU time (minutes)"
-	label3 = "# Computed data: H3, H2, uhavg"
+	label3 = "# Computed data: H3, H2, uhavg"	hamdata = [hamlabel; naccsamples; nmodes; invtemp; E0; D0; H3vec; H2vec; uhavg]
 	hamdata = [label1; nmodes; nsamples; invtemp; E0; D0;
-			label2; naccept; acceptrate; cputime;
-			label3; H3vec; H2vec; uhavg]
-	hamfile = string("../SamplingData/hamdata",suffix,".txt")
+				label2; naccept; acceptrate; cputime; label3; H3; H2; uhavg]	
 	writedata(hamdata, hamfile)
 	# Write uacc data to a file.
-	#udata = [uacc [real(uhacc); imag(uhacc)] ]
-	ufile = string("../SamplingData/udata",suffix,".txt")
+	ufile = string("../SamplingData/u",suffix,".txt")
 	writedata(uacc, ufile)
 	# Print the final CPU time.
[…]

### 2018-09-03 SamplingCodes/routines.jl

@@ -1,3 +1,4 @@
+using Winston
 using Distributions
 
 #---------- IO Routines ----------#
@@ -22,7 +23,7 @@ Note: uu has length npoints = 2*nmodes. =#
 #= Basic realfft to go from uu to uhat. =#
 function realfft(uu::Vector{Float64})
 	uhat = rfft(uu)/endof(uu)
-	assert(abs(uhat[1])/maximum(abs,uhat) < 1e-6) 
+	assert(abs(uhat[1])/max.(abs,uhat) < 1e-6) 
 	return uhat[2:end]
 end
 #= Basic irealfft to go from uhat to uu (no upsampling). =#
@@ -82,9 +83,10 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 	# Get all the random samples at once.
 	rvar = randn(nmodes,2,nsamples)
 	# Allocate space.
-	H3all, H2all = [zeros(Float64,nsamples) for nn=1:2]
-	aupvec, adnvec = [zeros(Float64,nsamples) for nn=1:2]
-	maxup, maxdn = 0.,0.
+	H3all = zeros(Float64,nsamples)
+	H2all = zeros(Float64,nsamples)
+	aratevec = zeros(Float64,nsamples)
+	maxaccept = 0.
 	# Compute H3 and H2 for each to estimate the acceptance rate.
 	# TO DO: parallelize this for loop!!! It is the bottleneck!
 	for nn=1:nsamples
@@ -92,64 +94,34 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 		H3 = ham3(uhat)
 		H2 = ham2(uhat)
 		# Compute the Hamiltonian (for upstream D0=1)
-		hamup = sqrt(E0)*H3 - H2
-		hamdn = D0^(-13/4)*sqrt(E0)*H3 - D0^(3/2)*H2
+		ham = D0^(-13/4)*sqrt(E0)*H3 - D0^(3/2)*H2
 		# Compute the unnormalized acceptance rate and maximize it.
-		accept_up = exp(-invtemp * hamup)
-		accept_dn = exp(-invtemp * hamdn)
-		maxup = max(maxup, accept_up)
-		maxdn = max(maxdn, accept_dn)
+		acceptrate = exp(-invtemp * ham)
+		maxaccept = max(maxaccept, acceptrate)
[…]

### 2018-09-03 SamplingCodes/main.jl
NM

@@ -9,7 +9,8 @@ function main(nmodes::Int, nsamples::Int, invtemp::Float64,
 	# Sample from a Gibbs distribution.
 	cputime = @elapsed (H3vec,H2vec,uhacc) = gibbs(nmodes,nsamples,invtemp,E0,D0,savemicro)
 	cputime = signif(cputime/60,2)	
-	naccsamples = endof(H3vec)
+	naccept = endof(H3vec)
+	acceptrate = signif(100*naccept/nsamples,2)
 	# Convert each accepted uhat to physical space for analysis.
 	uacc = getuacc(uhacc)
 	uhavg = getuhavg(uhacc)
@@ -18,9 +19,9 @@ function main(nmodes::Int, nsamples::Int, invtemp::Float64,
 	hamfile = string("../SamplingData/ham",suffix,".txt")
 	label1 = "# Input parameters: nmodes, nsamples, invtemp, E0, D0"
 	label2 = "# Calculated parameters: number of accepted samples, acceptance rate (%), CPU time (minutes)"
-	label3 = "# Computed data: H3, H2, uhavg"	hamdata = [hamlabel; naccsamples; nmodes; invtemp; E0; D0; H3vec; H2vec; uhavg]
+	label3 = "# Computed data: H3, H2, uhavg"	
 	hamdata = [label1; nmodes; nsamples; invtemp; E0; D0;
-				label2; naccept; acceptrate; cputime; label3; H3; H2; uhavg]	
+				label2; naccept; acceptrate; cputime; label3; H3vec; H2vec; uhavg]	
 	writedata(hamdata, hamfile)
 	# Write uacc data to a file.
 	ufile = string("../SamplingData/u",suffix,".txt")
@@ -30,6 +31,8 @@ function main(nmodes::Int, nsamples::Int, invtemp::Float64,
 	return
 end
 
-main(10, 1*10^6, -0.0, 4., 1.0, "up")
+main(10, 1*10^5, -0.0, 4., 1.0, "up")
+
+#main(10, 1*10^6, -0.0, 4., 1.0, "up")
 #main(10, 2*10^7, -0.3, 4., 1.0, "up")
 #main(10, 2*10^7, -0.3, 4., 0.5, "dn")

### 2018-09-03 SamplingCodes/routines.jl

@@ -1,4 +1,3 @@
-using Winston
 using Distributions
 
 #---------- IO Routines ----------#

### 2018-09-03 SamplingCodes/routines.jl
NM

@@ -82,9 +82,7 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 	# Get all the random samples at once.
 	rvar = randn(nmodes,2,nsamples)
 	# Allocate space.
-	H3all = zeros(Float64,nsamples)
-	H2all = zeros(Float64,nsamples)
-	aratevec = zeros(Float64,nsamples)
+	H3all,H2all,aratevec = [zeros(Float64,nsamples) for nn=1:3]
 	maxaccept = 0.
 	# Compute H3 and H2 for each to estimate the acceptance rate.
 	# TO DO: parallelize this for loop!!! It is the bottleneck!
@@ -107,8 +105,7 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 		end
 	end
 	# Now that the normalization is known, accept or reject each.
-	H3acc = zeros(Float64,0)
-	H2acc = zeros(Float64,0)
+	H3acc,H2acc = [zeros(Float64,0) for nn=1:2]
 	savemicro? uhacc = zeros(Complex128,nmodes,nsamples):0
 	count = 0
 	for nn=1:nsamples

### 2018-09-03 SamplingCodes/main.jl
NM

@@ -27,7 +27,7 @@ function main(nmodes::Int, nsamples::Int, invtemp::Float64,
 	ufile = string("../SamplingData/u",suffix,".txt")
 	writedata(uacc, ufile)
 	# Print the final CPU time.
-	println("The CPU time is ", signif(cputime/60,2), " minutes.")
+	println("The CPU time is ", signif(cputime,2), " minutes.")
 	return
 end
 

### 2018-09-03 SamplingCodes/main.jl
NM

@@ -31,8 +31,8 @@ function main(nmodes::Int, nsamples::Int, invtemp::Float64,
 	return
 end
 
-main(10, 1*10^5, -0.0, 4., 1.0, "up")
+#main(10, 1*10^5, -0.05, 4., 1.0, "up")
 
-#main(10, 1*10^6, -0.0, 4., 1.0, "up")
+main(10, 1*10^6, -0.0, 4., 1.0, "up")
 #main(10, 2*10^7, -0.3, 4., 1.0, "up")
 #main(10, 2*10^7, -0.3, 4., 0.5, "dn")

### 2018-09-03 SamplingCodes/main.jl
Working towards the matching condition.

@@ -31,8 +31,8 @@ function main(nmodes::Int, nsamples::Int, invtemp::Float64,
 	return
 end
 
-#main(10, 1*10^5, -0.05, 4., 1.0, "up")
+main(10, 1*10^5, -0.05, 4., 1.0, "up")
 
-main(10, 1*10^6, -0.0, 4., 1.0, "up")
+#main(10, 1*10^6, -0.0, 4., 1.0, "up")
 #main(10, 2*10^7, -0.3, 4., 1.0, "up")
 #main(10, 2*10^7, -0.3, 4., 0.5, "dn")

### 2018-09-03 SamplingCodes/routines.jl

@@ -76,6 +76,25 @@ function getuhat(rvar::Array{Float64}, nn::Int)
 	return uhat/sqrt(energy(uhat))
 end
 
+#= Sample from a microcanonical distribution (Gibbs with zero inverse temp) =#
+function microcan(nmodes::Int, nsamples::Int)
+	# Get all the random samples at once.
+	rvar = randn(nmodes,2,nsamples)
+	# Allocate space.
+	H3vec,H2vec = [zeros(Float64,nsamples) for nn=1:2]
+	# Compute H3 and H2 for each to estimate the acceptance rate.
+	# TO DO: parallelize this for loop!!! It is the bottleneck!
+	for nn=1:nsamples
+		uhat = getuhat(rvar,nn)
+		H3vec[nn] = ham3(uhat)
+		H2vec[nn] = ham2(uhat)
+		if mod(nn, 10^4) == 0
+			println("Initial sampling is ", signif(100*nn/nsamples,3), "% completed.")
+		end
+	end
+end
+
+
 #= Sample from a Gibbs distribution with arbitrary inverse temperature. =#
 function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0., 
 		E0::Float64=1., D0::Float64=1., savemicro::Bool=false)

### 2018-09-03 SamplingCodes/routines.jl
NM

@@ -75,13 +75,12 @@ function getuhat(rvar::Array{Float64}, nn::Int)
 	uhat = rvar[:,1,nn]+im*rvar[:,2,nn]
 	return uhat/sqrt(energy(uhat))
 end
-
-#= Sample from a microcanonical distribution (Gibbs with zero inverse temp) =#
+#= Sample from a microcanonical distribution (Gibbs with zero inverse temperature). =#
 function microcan(nmodes::Int, nsamples::Int)
 	# Get all the random samples at once.
 	rvar = randn(nmodes,2,nsamples)
 	# Allocate space.
-	H3vec,H2vec = [zeros(Float64,nsamples) for nn=1:2]
+	H3vec, H2vec = [zeros(Float64,nsamples) for nn=1:2]
 	# Compute H3 and H2 for each to estimate the acceptance rate.
 	# TO DO: parallelize this for loop!!! It is the bottleneck!
 	for nn=1:nsamples
@@ -92,45 +91,31 @@ function microcan(nmodes::Int, nsamples::Int)
 			println("Initial sampling is ", signif(100*nn/nsamples,3), "% completed.")
 		end
 	end
+	return H3vec, H2vec, rvar
 end
-
 
 #= Sample from a Gibbs distribution with arbitrary inverse temperature. =#
 function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0., 
 		E0::Float64=1., D0::Float64=1., savemicro::Bool=false)
-	# Get all the random samples at once.
-	rvar = randn(nmodes,2,nsamples)
-	# Allocate space.
-	H3all,H2all,aratevec = [zeros(Float64,nsamples) for nn=1:3]
+	# Sample H3 and H2 from the microcanonical distribution.
+	H3all,H2all,rvar = microcan(nmodes,nsamples)
+	# Determine the accpetance rate based on the Hamiltonian.
+	aratevec = zeros(Float64,nsamples)
 	maxaccept = 0.
-	# Compute H3 and H2 for each to estimate the acceptance rate.
-	# TO DO: parallelize this for loop!!! It is the bottleneck!
 	for nn=1:nsamples
-		uhat = getuhat(rvar,nn)
-		H3 = ham3(uhat)
-		H2 = ham2(uhat)
 		# Compute the Hamiltonian (for upstream D0=1)
-		ham = D0^(-13/4)*sqrt(E0)*H3 - D0^(3/2)*H2
[…]

### 2018-09-03 SamplingCodes/main.jl
NM

@@ -2,7 +2,7 @@ include("routines.jl")
 
 #= Compute the two Hamiltonians (upstream and downstream)
 as sampled from the Gibbs0 distribution. =# 
-function main(nmodes::Int, nsamples::Int, invtemp::Float64, 
+function sample_gibbs(nmodes::Int, nsamples::Int, invtemp::Float64, 
 		E0::Float64, D0::Float64, suffix::AbstractString)
 	savemicro = true
 	println("\n\n")
@@ -31,8 +31,15 @@ function main(nmodes::Int, nsamples::Int, invtemp::Float64,
 	return
 end
 
-main(10, 1*10^5, -0.05, 4., 1.0, "up")
+# Testing case:
+#sample_gibbs(10, 1*10^5, -0.05, 4., 1.0, "up")
+# Three main runs:
+#sample_gibbs(10, 1*10^6, -0.0, 4., 1.0, "up")
+#sample_gibbs(10, 2*10^7, -0.3, 4., 1.0, "up")
+#sample_gibbs(10, 2*10^7, -0.3, 4., 0.5, "dn")
 
-#main(10, 1*10^6, -0.0, 4., 1.0, "up")
-#main(10, 2*10^7, -0.3, 4., 1.0, "up")
-#main(10, 2*10^7, -0.3, 4., 0.5, "dn")
+function aaa()
+	microcan(nmodes,nsamples)
+
+
+

### 2018-09-03 SamplingCodes/routines.jl

@@ -104,7 +104,7 @@ function gibbs(nmodes::Int, nsamples::Int, invtemp::Float64=0.,
 	maxaccept = 0.
 	for nn=1:nsamples
 		# Compute the Hamiltonian (for upstream D0=1)
-		ham = D0^(-13/4)*sqrt(E0)*H3vec[nn] - D0^(3/2)*H2vec[nn]
+		ham = D0^(-13/4)*sqrt(E0)*H3all[nn] - D0^(3/2)*H2all[nn]
 		# Compute the unnormalized acceptance rate and find the maximum.
 		aratevec[nn] = exp(-invtemp * ham)
 		maxaccept = max(maxaccept, aratevec[nn])

### 2018-09-03 SamplingCodes/main.jl
NM

@@ -38,8 +38,25 @@ end
 #sample_gibbs(10, 2*10^7, -0.3, 4., 1.0, "up")
 #sample_gibbs(10, 2*10^7, -0.3, 4., 0.5, "dn")
 
-function aaa()
-	microcan(nmodes,nsamples)
+function HdnGup(H3vec::Vector{Float64}, H2vec::Vector{Float64}, 
+		thup::Float64, E0::Float64, D0::Float64)
+	Hdn_mean, normconst = 0.,0.
+	for nn=1:endof(H3vec)
+		# Compute the upstream and downstream Hamiltonians.
+		Hup = sqrt(E0)*H3vec[nn] - H2vec[nn]
+		Hdn = D0^(-13/4)*sqrt(E0)*H3vec[nn] - D0^(3/2)*H2vec[nn]
+		# Compute the mean of Hdn under Gup.
+		Hdn_mean += exp(-thup*Hup) * Hdn
+		normconst += exp(-thup*Hup)
+	end
+	return Hdn_mean/normconst
+end
+
+function main()
+	H3vec, H2vec, rvar = microcan(nmodes,nsamples)
+	HdnGup()
+
+
 
 
 

### 2018-09-03 SamplingCodes/main.jl
Writing the code to enforce the statistical matching condition.

@@ -38,23 +38,33 @@ end
 #sample_gibbs(10, 2*10^7, -0.3, 4., 1.0, "up")
 #sample_gibbs(10, 2*10^7, -0.3, 4., 0.5, "dn")
 
-function HdnGup(H3vec::Vector{Float64}, H2vec::Vector{Float64}, 
-		thup::Float64, E0::Float64, D0::Float64)
-	Hdn_mean, normconst = 0.,0.
+function meanham(H3vec::Vector{Float64}, H2vec::Vector{Float64}, 
+		theta::Float64, E0::Float64, D0::Float64, Gup::Bool)
+	hamdnmean, normconst = 0.,0.
 	for nn=1:endof(H3vec)
 		# Compute the upstream and downstream Hamiltonians.
-		Hup = sqrt(E0)*H3vec[nn] - H2vec[nn]
-		Hdn = D0^(-13/4)*sqrt(E0)*H3vec[nn] - D0^(3/2)*H2vec[nn]
+		hamup = sqrt(E0)*H3vec[nn] - H2vec[nn]
+		hamdn = D0^(-13/4)*sqrt(E0)*H3vec[nn] - D0^(3/2)*H2vec[nn]
+		# Decide whether to use Gup or Gdn
+		Gup? ham = hamup : ham = hamdn
 		# Compute the mean of Hdn under Gup.
-		Hdn_mean += exp(-thup*Hup) * Hdn
-		normconst += exp(-thup*Hup)
+		hamdnmean += exp(-theta*ham) * hamdn
+		normconst += exp(-theta*ham)
 	end
-	return Hdn_mean/normconst
+	return hamdnmean/normconst
 end
 
-function main()
+function main(nmodes::Int, nsamples::Int, thup::Float64, E0::Float64, D0::Float64)
+	# Sample H3 and H2 from a microcanonical distribution.
 	H3vec, H2vec, rvar = microcan(nmodes,nsamples)
-	HdnGup()
+	# Compute the upstream mean of the downstream Hamiltonian.
+	meanup = meanham(H3vec,H2vec,thup,E0,D0,true)
+	# Define a function for the dowstream mean of the downstream Hamiltonian.
+	function meandn(thdn::Float64)
+		return meanham(H3vec,H2vec,thdn,E0,D0,false)
+	end
+
+end
 
 
 

### 2018-09-03 SamplingCodes/main.jl
NM

@@ -54,19 +54,19 @@ function meanham(H3vec::Vector{Float64}, H2vec::Vector{Float64},
 	return hamdnmean/normconst
 end
 
-function main(nmodes::Int, nsamples::Int, thup::Float64, E0::Float64, D0::Float64)
+function matchmean(nmodes::Int, nsamples::Int, thup::Float64, E0::Float64, D0::Float64)
 	# Sample H3 and H2 from a microcanonical distribution.
 	H3vec, H2vec, rvar = microcan(nmodes,nsamples)
 	# Compute the upstream mean of the downstream Hamiltonian.
 	meanup = meanham(H3vec,H2vec,thup,E0,D0,true)
-	# Define a function for the dowstream mean of the downstream Hamiltonian.
-	function meandn(thdn::Float64)
-		return meanham(H3vec,H2vec,thdn,E0,D0,false)
-	end
-
+	# Define a function for the difference in the up/downstream means.
+	function diffmean(thdn::Float64) = meanham(H3vec,H2vec,thdn,E0,D0,false) - meanup
+	# Find thdn via a root of diffmean
+	thdn = find_zero(diffmean, thup, Roots.Newton())
+	return thdn
 end
 
-
+matchmean(10, 1*10^5, -0.1, 4., 1.)
 
 
 

### 2018-09-03 SamplingCodes/main.jl
The code to match the upstream and downstream means seems to work. Tomorrow I will test it and produce some results.

@@ -39,14 +39,14 @@ end
 #sample_gibbs(10, 2*10^7, -0.3, 4., 0.5, "dn")
 
 function meanham(H3vec::Vector{Float64}, H2vec::Vector{Float64}, 
-		theta::Float64, E0::Float64, D0::Float64, Gup::Bool)
+		theta::Float64, E0::Float64, D0::Float64, gibbsup::Bool)
 	hamdnmean, normconst = 0.,0.
 	for nn=1:endof(H3vec)
 		# Compute the upstream and downstream Hamiltonians.
 		hamup = sqrt(E0)*H3vec[nn] - H2vec[nn]
 		hamdn = D0^(-13/4)*sqrt(E0)*H3vec[nn] - D0^(3/2)*H2vec[nn]
 		# Decide whether to use Gup or Gdn
-		Gup? ham = hamup : ham = hamdn
+		gibbsup? ham = hamup : ham = hamdn
 		# Compute the mean of Hdn under Gup.
 		hamdnmean += exp(-theta*ham) * hamdn
 		normconst += exp(-theta*ham)
@@ -60,13 +60,13 @@ function matchmean(nmodes::Int, nsamples::Int, thup::Float64, E0::Float64, D0::F
 	# Compute the upstream mean of the downstream Hamiltonian.
 	meanup = meanham(H3vec,H2vec,thup,E0,D0,true)
 	# Define a function for the difference in the up/downstream means.
-	function diffmean(thdn::Float64) = meanham(H3vec,H2vec,thdn,E0,D0,false) - meanup
+	diffmean(thdn::Float64) = meanham(H3vec,H2vec,thdn,E0,D0,false) - meanup
 	# Find thdn via a root of diffmean
-	thdn = find_zero(diffmean, thup, Roots.Newton())
+	thdn = find_zero(diffmean, thup, Order1())
 	return thdn
 end
 
-matchmean(10, 1*10^5, -0.1, 4., 1.)
+matchmean(10, 1*10^5, -0.1, 4., 0.5)
 
 
 

### 2018-09-03 SamplingCodes/routines.jl

@@ -1,4 +1,5 @@
 using Distributions
+using Roots
 
 #---------- IO Routines ----------#
 function readvec(file::AbstractString)

### 2018-09-04 LatexNotes/Notes.tex
Small changes to plots in latex notes.
Working on latex notes.

@@
 \subsection{Sampling from microcanonical energy distribution}
+\label{sec_microcan}
 
@@
 \subsection{Acceptance/rejection algorithm to sample from a mixed microcanonical/canonical distribution.}
+\label{sec_accrej}
 
@@
 \end{enumerate}
-The above is a conceptual description of the algorithm. In practice, it is implemented in a way to be computational efficient.
+The above is a conceptual description of the algorithm.
+In practice, it is implemented in a way to be computational efficient.
+In particular, the most expensive step is computing $H_3$ and $H_2$ for a large number of microstates sampled from the microcanonical distribution.
+We therefore perform this step as precomputation, and save the values of $H_3$ and $H_2$.
+Afterwards, we can use this master list to perform the acceptance/rejection steps for a number of inverse temperatures, $\theta$, as well as values of $E_0$ and $D_0$.
+
+\subsection{Enforcing the statistical matching condition across ADC}
+\label{sec_match}
+
+Importantly, we must enforce the statistical matching condition at the abrupt depth change
+\begin{equation}
+\label{statmatch}
+\mean{H^+}_{\Gibbs^-} = \mean{H^+}_{\Gibbs^+}
+\end{equation}
+We regard the upstream inverse temperature $\theta^-$ as given, and the downstream value $\theta^+$ as the main unknown to be solved for.
+First, note that the ensemble average is given by the formula
+\begin{equation}
+\label{meanformula}
+\mean{H^+}_{\Gibbs^{\pm}} = \frac{ \sum H^+ \exp(- \theta^{\pm} H^{\pm}) } {\sum \exp(-\theta^{\pm} H^{\pm}) }
+\end{equation}
+where the sum is taken over the ensemble. We therefore enforcing matching condition \eqref{statmatch} as follows
+
+\begin{enumerate}
+\item Sample from a large number of microstates (e.g.~$10^8$) from the microcanonical distribution as described above. Compute and save the values of $H_3$ and $H_2$ for each microstate.
+\item For prescribed values of $\theta^-$, $E_0$, and $D_0$, compute the value $\mean{H^+}_{\Gibbs^-}$ using \eqref{meanformula}.
+\item Define a function to compute the difference $F(\theta^+) = \mean{H^+}_{\Gibbs^-} - \mean{H^+}_{\Gibbs^+}$ for a given $\theta^+$.
+\item Find a root of $F$ (e.g.~using the secant method) to determine the value of $\theta^+$ that satisfies \eqref{statmatch}.
+\end{enumerate}
+Once the value of $\theta^+$ is determined, we sample from the mixed canonical/microcanonical distributions using the algorithm described in the previous section.
+Note that in practice, we perform step 1 of generating a large number of microsamples only once, and then performing the rest of the algorithm for several different $\theta^-$ values, all using the same list of $H_3$ and $H_2$ values.
 
@@
 \section{Preliminary results}
+
 We show some preliminary results below.
+First, Figs.~\ref{Hist1}--\ref{Hist3} are histograms of the upstream and downstream Hamiltonians, $H^-$ and $H^+$, respectively, for various values of inverse temperature $\theta^{\pm}$, energy $E_0$, and depth ratio $D_0$.
+These histograms are made in a very simply way, namely by sampling from the microcanonical distribution (as described in Section \ref{sec_microcan}) and then multiplying by the appropriate factor $\exp(-\theta^{\pm} H^{\pm})$ (and normalizing).
+There is nothing complex at all in generating these histograms (not even an acceptance/rejection algorithm) and they are only made to give us some initial idea of what is going on.
+
+In these tests, I chose the default values $\Lambda = 10$, $E_0 = 1$, and $D_0 = 0.6$ ($\lambda$ is set to unity throughout).
+Fig.~\ref{Hist1} shows the upstream and downstream histograms generated with these default values for the inverse temperatures $\theta^{\pm} = 0, \pm 0.3$.
+First, notice that $\theta$ can shift the mean of $H$ significantly, with negative $\theta$ shifting the mean closer to zero.
+Second, notice that the downstream histogram has a mean much closer to zero, and this histogram is more concentrated.
+
+In figure \ref{Hist2} we vary the depth ratio.
+The left shows $D_0 = 0.6$ as before, and the right shows $D_0 = 0.4$.
+Notice a slight change in the depth ratio can alter the histograms significantly.
+In figure \ref{Hist3} we vary the energy $E_0$ from the default value of 1 (left) to the value 4.
+Notice that a significant change in $E_0$ only causes slight changes in the histograms.
+
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
@@
 \caption{Sensitivity to changes in the energy $E_0$.
-Left: Downstream histogram with $E_0 = 1$ (same as in Fig.~\ref{Hist1}). Right: Same with $E_0 = 5$. Notice a slight change in the histogram.}
+Left: Downstream histogram with $E_0 = 1$ (same as in Fig.~\ref{Hist1}). Right: Same with $E_0 = 4$. Notice a slight change in the histogram.}
 \label{Hist3}
@@
 While the upstream histogram ostensible depends on $E_0$, the changes are nearly imperceptible and therefore not shown.
+
+ Next, in Figs.~\ref{microup0}--\ref{microdn3}, we examine various aspects of the microstates under given Gibbs distributions with possibly non-zero inverse temperatures.
+For this task, we need to use the acceptance/rejection algorithm described in Section \ref{sec_accrej}.
+In these tests, we again fix $\Lambda = 10$.
+
+ First, Fig.~\ref{microup0} shows features of the microstate with zero inverse temperature (i.e.~the microcanonical distribution).
+The left panel shows a histogram of the displacement $u$ (in physical space).
+Notice the histogram is symmetric and resembles a Gaussian distribution.
+The right panel shows the $\mean{\abs{\uhat_k}}$ versus wavenumber $k$, which illustrates how the energy is distributed amongst the modes.
+Notice that the energy is evenly distributed amongst modes in this case, i.e.~a equidistribution of energy.
+
+ In Fig.~\ref{microup3} we show the same plots but with $\theta^- = -0.3$.
+The histogram of $u$ is shown on the left.
+While this histogram still appears nearly symmetric, close examination reveals a slight positive skewness.
+The right shows $\mean{\abs{\uhat_k}}$.
+Notice that, in this case of negative inverse temperature, the spectrum is decidedly tilted, with less energy present in the higher modes.
+
+ Figure \ref{microdn3} shows the same plots for the {\em downstream} distributions with $\theta^+ = -0.3$.
+Here, we have not enforced the statistical matching condition \eqref{statmatch}; we simply chose a value of $\theta^+$.
+First, notice that the histogram of $u$ shows a definite positive skewness.
+This observation is consistent with the main finding in the experiments!
+Second, the spectrum is tilted as before, but the tilt is not as strong as it is upstream.
+Both of these observations are consistent with the weights of $H_3$ and $H_2$.
+Downstream, the weights strongly favor $H_3$.
+Therefore, to have a better chance of being selected from the distribution $\Gibbs^{+}$, it is more advantageous to have large skewness of $u$ than it is to have small variance of $u_x$.
 
@@
 \caption{Examination of the downstream microstates with $\theta^+ = -0.3$ and $D_0 = 0.5$ ($E_0 = 4$ as before).
-Notice the skewness of $u$!!!
+Notice the skewness of $u$!
 The spectrum is tilted, though not as much as it is upstream.}
 \label{microdn3}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
+
+Now, figures \ref{thetaplot}--\ref{match_micro} show results obtained by satisfying the matching condition \eqref{statmatch} using the algorithm described in \ref{sec_match}.
+In these tests, I use $\Lambda = 8$ and the provisional number of samples is $10^8$ (the number of accepted samples will be much lower).
+First, Fig.~\ref{thetaplot} shows, for given values of $\theta^-$, the corresponding values of $\theta^+$ as determined by the matching condition.
+For one, $\theta^- = 0$ corresponds to $\theta^+ = 0$ which is a good sanity check.
+Second, notice that the signs of $\theta^-$ and $\theta^+$ always agree, which is a good thing.
+The relationship is not a simply linear one, as the curvature is significant.
+
+ Next, Fig.~\ref{matchH} shows upstream and downstream histograms of $H$ with the matching condition enforced.
+The black curve is the histogram of $\Gibbs^-(H^-)$, while the colored curves are histograms of $H+$.
+The red curve corresponds to $\Gibbs^-(H^+)$ and the blue to $\Gibbs^+(H^+)$.
+The matching condition says that the means of these two histograms should be the same, which can be verified visually.
+Notice, the main difference is that the downstream histogram $\Gibbs^+(H^+)$ is more spread out.
+
+Figure \ref{match_micro} shows features of the underlying microstates.
+On the left, we show histograms of the displacement $u$, as we did in Figs.~\ref{microup0}--\ref{microdn3}, but this time {\em we have enforced the matching condition} \eqref{statmatch}.
+Notice the upstream $u$-histogram is nearly symmetric, while the downstream histogram is noticeable skewed in the positive direction!
+The right panel shows the ensemble-averaged spectrum $\mean{\abs{\uhat_k}}$.
+Both show a tilt in the spectrum, with the upstream tilt being more significant than downstream.
+
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[htbp]
+\begin{center}
+\includegraphics[width = 0.55 \textwidth]{thetaplot}
+\caption{For given values of $\theta^-$, we compute $\theta^+$ to satisfy the matching condition.}
+\label{thetaplot}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
+
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[htbp]
+\begin{center}
+\includegraphics[width = 0.75 \textwidth]{matchH}
+\caption{Histograms of $H^\pm$ under the two Gibbs measures $\Gibbs^{\pm}$ with the matching condition enforced.}
+\label{matchH}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
+
+   %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[htbp]
+\begin{center}
+\includegraphics[width = 0.95 \textwidth]{match_micro}
+\caption{Microstate features with the matching condition enforced.
+Left: the histogram of displacement $u$ upstream (red) and downstream (blue).
+Notice the downstream histogram has a definite positive skewness.
+Right: the ensemble-averaged spectra.
+Both spectra are tilted, with the upstream tilt being more significant.}
+\label{match_micro}
 \end{center}

### 2018-09-04 LatexNotes/Notes.tex
NM

@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-%Notice that even though the depth ratio changes only slightly, the histograms change significantly. The upstream histograms do not depend on $D_0$ and are therefore not shown.
 
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
- %Notice that the energy changes substantially, but the histograms change only slightly.
-While the upstream histogram ostensible depends on $E_0$, the changes are nearly imperceptible and therefore not shown.
 
@@
 Now, figures \ref{thetaplot}--\ref{match_micro} show results obtained by satisfying the matching condition \eqref{statmatch} using the algorithm described in \ref{sec_match}.
-In these tests, I use $\Lambda = 8$ and the provisional number of samples is $10^8$ (the number of accepted samples will be much lower).
+In these tests, we use $\Lambda = 8$ and the provisional number of samples is $10^8$ (the number of accepted samples will be much lower).
+We also fix $E_0 = 4$ and $D_0 = 0.5$.
+
 First, Fig.~\ref{thetaplot} shows, for given values of $\theta^-$, the corresponding values of $\theta^+$ as determined by the matching condition.
@@
 Both show a tilt in the spectrum, with the upstream tilt being more significant than downstream.
+
+We remark that simulations have been performed with larger values of $\Lambda$, for example $\Lambda = 16$ and 20.
+For the provisional number of samples fixed, increasing $\Lambda$ does not drastically increase the computational time.
+However, the main downside to increasing $\Lambda$ is that the acceptance rate for the Gibbs distributions is substantially smaller, and therefore the statistics are less resolved.
 

### 2018-09-05 LatexNotes/Notes.tex
NM

@@
 
-Now, figures \ref{thetaplot}--\ref{match_micro} show results obtained by satisfying the matching condition \eqref{statmatch} using the algorithm described in \ref{sec_match}.
+Now, figures \ref{match_theta}--\ref{match_micro} show results obtained by satisfying the matching condition \eqref{statmatch} using the algorithm described in \ref{sec_match}.
 In these tests, we use $\Lambda = 8$ and the provisional number of samples is $10^8$ (the number of accepted samples will be much lower).
@@
 
-First, Fig.~\ref{thetaplot} shows, for given values of $\theta^-$, the corresponding values of $\theta^+$ as determined by the matching condition.
+First, Fig.~\ref{match_theta} shows, for given values of $\theta^-$, the corresponding values of $\theta^+$ as determined by the matching condition.
 For one, $\theta^- = 0$ corresponds to $\theta^+ = 0$ which is a good sanity check.
@@
 
+Little remark for the very careful reader: the plot in Fig.
+\ref{match_theta} is out of date (used $\Lambda=16$) and is not consistent with Figs.~\ref{matchH} and \ref{match_micro} (which used $\Lambda = 8$).
+
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
@@
 \begin{center}
-\includegraphics[width = 0.55 \textwidth]{thetaplot}
+\includegraphics[width = 0.55 \textwidth]{match_theta}
 \caption{For given values of $\theta^-$, we compute $\theta^+$ to satisfy the matching condition.}
-\label{thetaplot}
+\label{match_theta}
 \end{center}

### 2018-09-19 LatexNotes/Notes.tex
NM

@@
 
-Little remark for the very careful reader: the plot in Fig.
-\ref{match_theta} is out of date (used $\Lambda=16$) and is not consistent with Figs.~\ref{matchH} and \ref{match_micro} (which used $\Lambda = 8$).
-
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
@@
 \label{match_micro}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
+
+\np
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Results
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Big-time results}
+
+Now lets do some more serious runs with larger $\Lambda$.
+The major challenge is that the acceptance rate plummets as $\Lambda$ grows, meaning that we either have to take many more samples or develop a smarter strategy.
+For now, we will do the first.
+To run more samples, I had to modify the organization of the code.
+I am still exploiting the ability to precompute a list of $H_3$ and $H_2$ values, but I cannot precompute and store the entire list at once because it will not fit in memory.
+So instead, I take several sweeps each with a manageable number of samples (usually about $10^7$).
+
+Figure \ref{Lambda16Run} shows a run with $\Lambda = 16$, with $10^7$ samples per sweep and 200 sweeps for a total of $2 \times 10^9$ nominal samples.
+The main parameter to vary is the upstream inverse temperature $\theta^-$.
+More negative $\theta^-$ gives a more significantly skewed distribution of $u$, but also results in noisier statistics due to the lower acceptance rate.
+I therefore chose $\theta^- = -0.2$ since it gives a good balance of the two.
+The top panel of the figure shows $\theta^+$ as a function of $\theta^-$ and the histograms for $\theta^- = -0.2$.
+The bottom panel shows microstate information, including the pdf of displacement, $u$, and the spectrum.
+Notice the definite skew of the downstream displacement and the decay of both upstream and downstream spectra.
+For this run, the upstream acceptance rate was $1\%$ and the downstream rate was $0.08\%$.
+
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[htbp]
+\begin{center}
+\includegraphics[width = 0.95 \textwidth]{Lambda16Run}
+\caption{Big run with $\Lambda = 16$ and $2 \times 10^9$ nominal samples.}
+\label{Lambda16Run}
 \end{center}

### 2018-09-22 LatexNotes/Notes.tex
NM
Updating latex notes. Writing up derivation of KdV so that I can calculate powers of E_0 and D_0

@@
 
+\subsection{Nondimensionalization of the Euler system leading to KdV}
+
+I find Johnson's derivation of KdV difficult to follow (he keeps changing his mind about what scales to use) so I am doing it myself.
+We start with the incompressible Euler equations in dimensional form
+\begin{align}
+\frac{Du}{Dt} = -\frac{1}{\rho} \pd{P}{x} \, , \qquad
+\frac{Dw}{Dt} = -\frac{1}{\rho} \pd{P}{z} - g \, , \qquad
+u_x + w_z = 0
+\end{align}
+which are supplemented with free-surface conditions
+\begin{align}
+&w = h_t + u h_x	\quad \text{on } z = h \\
+&\phi_t + \frac{1}{2} (u^2 + w^2) + g(h-h_0) = 0  	\quad \text{on } z = h
+\end{align}
+and bottom conditions
+\begin{equation}
+w = b_t + u b_x		\quad \text{on } z = b
+\end{equation}
+Henceforth, we will assume the bottom topography is fixed so that the above reduces to $w = u b_x$.
+
+We non-dimensionalize the above system by choosing a length scale of $h_0$ (the upstream depth) and a timescale of $\sqrt{h_0/g}$, which gives a velocity scale of $\sqrt{gh_0}$. 
+Note: Johnson uses separate scales for horizontal versus vertical lengths and velocities, but then we rescales everything to undo this anyway.
+Instead, I am directly scaling all lengths and velocities on the same scales.
+We write the free surface as 
+\begin{equation}
+h = h_0 + a \eta
+\end{equation}
+where $a$ is a characteristic amplitude. Thus, $\eta$ is dimensionless. We write the pressure as
+\begin{equation}
+P = P_a + \rho g (h_0 - z) + \rho g h_0 p
+\end{equation}
+so that $p$ is dimensionless. Using these scales, the dimensionless Euler equations are
+\begin{align}
+\frac{Du}{Dt} = -p_x \, , \qquad
+\frac{Dw}{Dt} = -p_z \, , \qquad
+u_x + w_z = 0
+\end{align}
+The free-surface conditions are
+\begin{align}
+&w = \eps (\eta_t + u \eta_x) \quad \text{on } z = 1 + \eps \eta \\
+&\phi_t + \frac{1}{2} (u^2 + w^2) + \eps \eta = 0 \quad \text{on } z = 1 + \eps \eta
+\end{align}
+The bottom condition is
+\begin{equation}
+w = u b_x \quad \text{on } z = b
+\end{equation}
+Here, $\eps = a / h_0$ is a small parameter that measures the ratio of wave amplitude to depth.
+
+\subsection{Hamiltonian structure}
 We will be using the truncated KdV equations (TKDV) with variable depth, in particular an abrupt depth change (ADC).
@@
 Notice the definite skew of the downstream displacement and the decay of both upstream and downstream spectra.
-For this run, the upstream acceptance rate was $1\%$ and the downstream rate was $0.08\%$.
+For this run, the upstream acceptance rate was $0.01\%$ and the downstream rate was $0.008\%$.
+The total CPU time was about 33 hours.
 

### 2018-09-23 LatexNotes/Notes.tex
Sorting out KdV scales.

--- 2018-09-19 LatexNotes/Notes.tex
@@
 
+\subsection{The KdV equation}
+
+Following Johnson's derivation the dimensionless KdV equation is (on page 208)
+\begin{equation}
+2 \eta_{\tau} + 3 \eta \eta_{\xi} + \frac{1}{3} \eta_{\xi \xi \xi} = 0
+\end{equation}
+I had to work backwards quite a bit to figure out how this was non-dimensionalized (Johnson keeps rescaling things over and over again). Ultimately, the characteristic scales are
+\begin{align}
+& Z^* = h_0 , \quad X^* = \eps^{-1/2} h_0 , \quad T^* = \eps^{-1/2} \sqrt{h_0/g} , \\
+& W^* = \eps^{3/2} \sqrt{g h_0} , \quad U^* = \eps \sqrt{g h_0} , \quad P^* = \eps \rho g h_0
+\end{align}
+where $\eps$ is the ratio of characteristic amplitude to depth
+\begin{equation}
+\eps = a / h_0
+\end{equation}
+The variables $\tau$ and $\xi$ are
+\begin{align}
+\tau = \eps t , \quad \xi = x-t
+\end{align}
+Remember, the privileged KdV scaling is $\delta^2 \sim \eps$ where $\delta = h_0 / \lambda$.
+If this scaling is inserted into above, then $X^* \sim \lambda$, $T^* \sim \lambda/\sqrt{g h_0}$ which all makes sense.
+
+\subsection{Hamiltonian structure}
 We will be using the truncated KdV equations (TKDV) with variable depth, in particular an abrupt depth change (ADC).

### 2018-09-25 LatexNotes/Notes.tex
NM
Typing up notes on non-dimensionalization.

@@
 \newcommand{\hp}{h^{+}}
-
+\newcommand{\etastd}{\eta_{\text std}}
 % Theory for truncated KdV and stat mech
@@
 
+\subsubsection{Johnson's formulation}
 Following Johnson's derivation the dimensionless KdV equation is (on page 208)
@@
 If this scaling is inserted into above, then $X^* \sim \lambda$, $T^* \sim \lambda/\sqrt{g h_0}$ which all makes sense.
+
+If we convert KdV back into dimensional form, we get
+\begin{equation}
+2 \eta_t + \frac{3 c}{h_0} \eta \eta_{\xi} + \frac{c h_0^2}{3} \eta_{\xi \xi \xi} = 0
+\end{equation}
+Here, $\eta$ has units of length, $t$ time, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$.
+
+\subsubsection{My own reformulation}
+
+We now seek to non-dimensionalize the KdV equation, but not exactly the same way that Johnson did. Some things we would like to accomplish are:
+\begin{enumerate}
+\item When recast in dimensionless variables, $\eta$ should be periodic on the interval $[0, 2\pi)$ in order to make the Fourier notation simple.
+\item We want to rescale $\eta$ to have unit energy in order to make the Hamiltonian sampling simple.
+\end{enumerate}
+First, we observe there are a few lengthscales in the experiment that may be relevant: the upstream depth $h^- = 12.5$ cm, the tank width $L_0 = 20$ cm, the peak wavelength $\lambda_p^- = 38$ cm (induced by the peak forcing frequency as calculated by the dispersion relation), and possibly the maximum wavelength present $\lambda_{max}^{-} = ?$.
+First, since we want $\eta$ to be periodic on $[0,2 \pi)$ in dimensionless variables, the most logical choice for scaling $x$ is a wavelength, either the peak or max.
+Lets choose the peak.
+Next, since we want dimensionless $\eta$ to have unit energy, we {\em must} scale $\eta$ by the characteristic amplitude $a = \etastd = \mean{\eta^2}^{1/2}$, where the mean is taken over a sufficiently long time or distance.
+We therefore introduce dimensionless variables, denoted by a tilde
+\begin{align}
+& x = \frac{\lambda_p}{2\pi} \tilde{x} , \quad 
+t = \frac{\lambda_p}{2\pi c} \tilde{t} , \quad
+\eta = a \tilde{\eta} \\
+& \xi = x-ct = \frac{\lambda_p}{2\pi} (\tilde{x} - \tilde{t}) = \frac{\lambda_p}{2\pi} \tilde{\xi}
+\end{align}
+
+Then in dimensionless variables (after dropping the tildes), the system is given by
+\begin{equation}
+\eta_t + \frac{3}{2}\eps \eta \eta_{\xi} + \frac{2 \pi^2}{3} \delta^2 \eta_{\xi \xi \xi} = 0
+\end{equation}
+where we have the dimensionless variables
+\begin{align}
+\eps = a/h_0 , \quad \delta = h_0/\lambda_p
+\end{align}
+In these dimensionless variables, we have that $\eta$ is periodic for $x \in [0, 2\pi)$ and we have $\mean{\eta^2} = 1$, just as intended.
+Note, in the experiments, $\delta = 0.33$ and $\eps$ lies in the range $2.4 \times 10^{-3}$ to $2.4 \times 10^{-2}$.
+
+We will eventually want to rescale $t$, to look at long-time scales as is typically done in KdV analysis. Accordingly, we will define
+\begin{equation}
+\tau = \beta t
+\end{equation}
+where $\beta$ is usually either $\eps$ or $\delta^2$ (or we can choose it to be something more appropriate for the experiments).
+Of course, the effect of this rescaling would be to introduce a factor of $\beta$ in front of the first term $\eta_{\tau}$.
+
+Finally, since we want to study variable-depth KdV, we will ultimately want to interchange the roles of space and time as Johnson does on p.
+269.
+So we will have to modify some of this non-dimensionalization.
+I had to wrap my mind around the classical case first though.
+
+\subsubsection{Comparison to variables in Andy's original notes}
+In Andy's original notes (as well as in Di Qi's latex notes), the width of the tank $L_0$ was used for non-dimensionalizing $x$ (as well as $\xi$).
+In comparing our two formulations, the main subtlety is the energy, defined as
+\begin{equation}
+E_0 = \frac{1}{2} \int_{0}^{2 \pi L_0} u^2 dx
+\end{equation}
+where $u$ is displacement and has units of length.
+Thus, $E_0$ has units of length cubed (it took me a while to realize this is the main subtlety in reconciling the two formulations).
+The standard deviation of displacement is
+\begin{equation}
+u_{std} = \frac{1}{\sqrt{2 \pi L_0}} \left( \int_{0}^{2\pi L_0} u^2 dx \right)^{1/2}
+\end{equation}
+or in Andy's / Di Qi's notation
+\begin{equation}
+u_{std} = const \times L_0^{-1/2} E_0^{1/2}
+\end{equation}
+where I really don't care what the constant is. The main thing I wanted to see was why $L_0^{-1/2}$ appears, and the above shows this.
+
+The main difficulty I have with this formulation is that there is no reason to expect the spatial period of $u$ to have anything to do with the tank width $L_0$.
+More logically, the spatial period of $u$ should correspond to a characteristic wavelength in the problem, either the peak or maximum wavelength.
+The latter is more likely to be true in a literal sense, but I believe the peak wavelength provides a more reasonable scale for non-dimensionalization.
 

### 2018-09-25 LatexNotes/Notes.tex
NM

@@
 
+\np
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsubsection{My own reformulation}
+\subsubsection{My reformulation}
 
@@
 In these dimensionless variables, we have that $\eta$ is periodic for $x \in [0, 2\pi)$ and we have $\mean{\eta^2} = 1$, just as intended.
-Note, in the experiments, $\delta = 0.33$ and $\eps$ lies in the range $2.4 \times 10^{-3}$ to $2.4 \times 10^{-2}$.
+Note, in the experiments, $\delta = 0.33$ and $\eps$ lies in the range $2.4 \times 10^{-3}$--$2.4 \times 10^{-2}$.
 
@@
 
-The main difficulty I have with this formulation is that there is no reason to expect the spatial period of $u$ to have anything to do with the tank width $L_0$.
+The main difficulty I have with this formulation is that there is no reason to expect the spatial period of $u$ to be related to the tank width $L_0$.
 More logically, the spatial period of $u$ should correspond to a characteristic wavelength in the problem, either the peak or maximum wavelength.
@@
 
+Otherwise, the two formulations are completely equivalent.
+Notice that $\mu \lambda^{-3}$ in Andy's notation plays the role of $\delta^2$ in my notation, and $E_0^{1/2} \lambda^{-3/2}$ in Andy's notation plays the role of $\eps$ in my notation.
+
+\np
 \subsection{Hamiltonian structure}

### 2018-09-26 LatexNotes/Notes.tex
NM

@@
 \subsubsection{Johnson's formulation}
-Following Johnson's derivation the dimensionless KdV equation is (on page 208)
+Following Johnson's derivation, the dimensionless KdV equation is (on page 208)
 \begin{equation}
@@
 \begin{align}
-& Z^* = h_0 , \quad X^* = \eps^{-1/2} h_0 , \quad T^* = \eps^{-1/2} \sqrt{h_0/g} , \\
-& W^* = \eps^{3/2} \sqrt{g h_0} , \quad U^* = \eps \sqrt{g h_0} , \quad P^* = \eps \rho g h_0
+& X^* = \eps^{-1/2} h_0 ,\quad T^* = \eps^{-1/2} \sqrt{h_0/g} , \quad \eta^* = \eps h_0, \\
+& Z^* = h_0 , \quad U^* = \eps \sqrt{g h_0} , \quad W^* = \eps^{3/2} \sqrt{g h_0} , \quad P^* = \eps \rho g h_0
 \end{align}
@@
 Notice that $\mu \lambda^{-3}$ in Andy's notation plays the role of $\delta^2$ in my notation, and $E_0^{1/2} \lambda^{-3/2}$ in Andy's notation plays the role of $\eps$ in my notation.
+
+\subsubsection{Reformulating variable-depth KdV}
+
+We now seek to formulate the variable-depth KdV in a similar fashion. We start with Johnson's Eq. (3.148), which is written in dimensionless variables
+\begin{equation}
+2 \sqrt{D} \eta_X + \frac{1}{2} \frac{D'(X)}{\sqrt{D}} \eta + \frac{3}{D} \eta \eta_\xi + \frac{1}{3} D \eta_{\xi \xi \xi} = 0
+\end{equation}
+The first task is to convert this to dimensional form. For this, we now denote dimensionless variables with a tilde and use the relationships
+\begin{align}
+\eta = \eps h_0 \tilde{\eta} , \quad x = \eps^{-1/2} h_0 \tilde{x} = \eps^{-3/2} h_0 \tilde{X}
+\end{align}
+The trickiest part relates to $\xi$.
 

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \maketitle
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Experimental Preliminaries
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experimental parameters}
+
+Some relevant parameters from the experiment are given in the table.
+
+% Table
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{table}%[htbp]
+\begin{center}
+\caption{Basic experimental parameters and scales.}
+\label{paramtable}
+\begin{tabular}{l l l}
+\hline Parameter & Upstream  value \,\, & Downstream value \\
+\hline
+peak forcing frequency			& 2 Hz			& N/A \\
+depth, $h$						& 12.5 cm			& 3 cm \\
+peak wavelength, $\lambda_p$		& 38 cm			& 25 cm \\
+peak wavenumber, $k_p$			& 0.17 rad/cm		& 0.25 rad/cm \\
+dimensionless depth, $k_p h$\,\,\,		& 2.1			& 0.76 \\
+\end{tabular}
+\end{center}
+\end{table}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 
@@
 \subsection{The KdV equation}
+
+\subsubsection{Non-dimensionalization}
+
+The KdV equation in dimensional form is
+\begin{equation}
+2 \eta_t + \frac{3 c}{h_0} \eta \eta_{\xi} + \frac{c h_0^2}{3} \eta_{\xi \xi \xi} = 0
+\end{equation}
+Here, $\eta$ has units of length, $t$ time, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$. I obtained this equation by starting from Johnson's dimensionless equation ()
 
@@
 
-\newpage
-\newpage
-\section{Experimental parameters}
-
-Some relevant parameters from the experiment are given in the table.
-
-% Table
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{table}%[htbp]
-\begin{center}
-\caption{Some relevant experimental parameters} 
-\vspace{0.3 pc}
-\begin{tabular}{c l l}
-\hline
-\hspace{0.5pc} Symbol
-\hspace{0.5pc} & Meaning 
-\hspace{0.5pc} & Value \\
-\hline
-N/A		& Length of the experimental tank (in wave direction)	& 6 m	\\
-$L_0$	& Width of the experimental tank	 (transverse)		& $L_0 = $ 20 cm	\\
-$f_0$	& Dominant forcing frequency in experiments	& $f_0 = $ 2 Hz	\\
-$\hm$	& Depth on deep side						& $\hm = 12.5$ cm \\
-$\hp$	& Depth on shallow side					& $\hp = 3$ cm \\
-$\lamm$	& Wavelength induced by 2 Hz forcing on deep side	& $\lamm = 38$ cm	\\
-$\lamp$	& Wavelength induced by 2 Hz forcing on shallow side	& $\lamp = 25$ cm	\\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-\begin{comment}
-% Majda's notes typed up
-\section{Deterministic model for waves over abrupt depth change}
-
-The surface-displacement is denoted $u(x,t)$.
-
-Modeling assumptions
-\begin{enumerate}
-\item Waves in a narrow channel (width $L_0$) are nearly unidirectional and $2 \pi L_0$ periodic (purely a mathematical idealization, not realistic I believe).
-Where $L_0 < \lambda$, with $\lambda$ a characteristic wavelength.
-As seen in the table, this later assumption is reasonable since $L_0 / \lamm \approx 0.5$
-\item The role of the paddle forcing is to generate a state with `zero momentum', $M = \intt u \, dx = 0$ (this assumption is strongly supported by experimental measurements pointwise!) and with some given energy $E_0 = \frac{1}{2} \intt u^2 \, dx$ (not an assumption, just a statement).
-Here, $E_0$ is a function of the strength of the paddle forcing.
-\item Assume a Galerkin truncation of $u(x,t) \approx u_{\Lambda}(x,t)$, where $\Lambda$ (fixed) will probably be 20 or 30.
-\item The `time' coordinate $t$  corresponds physically to a downstream spatial coordinate (i.e.~$X \approx \eps x$ in Johnson's notation).
-The abrupt depth change (ADC) corresponds to $t = t_0$ or $T_{ADC}$.
-\end{enumerate}
-\end{comment}
-
 \bibliographystyle{plain}

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \end{equation}
-Here, $\eta$ has units of length, $t$ time, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$. I obtained this equation by starting from Johnson's dimensionless equation ()
+Here, $\eta$ has units of length, $t$ time, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$.
+I obtained this equation by starting from Johnson's dimensionless equation (3.28) on page 208 and working backwards to untangle the scales that he had used to non-dimensionalize.
 
+Let's establish some basic scales and notation:
+% Table
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{table}[h]%[htbp]
+\begin{center}
+\caption{Table.}
+\label{scale_table}
+\begin{tabular}{l l l l}
+\hline Description & Notation & Upstream  value \,\, & Downstream value \\
+\hline
+Characteristic wave amplitude	& $a = \etastd = \mean{\eta^2}^{1/2}$	& varied	& varied \\
+Peak forcing frequency		& $f_p$				& 2 Hz		& N/A \\
+Depth					& $h_0^{\pm}$			& 12.5 cm		& 3 cm \\
+Peak wavelength			& $\lambda_p^{\pm}$	& 38 cm		& 25 cm \\
+%Peak wavenumber			& $k_p$				& 0.17 rad/cm	& 0.25 rad/cm \\
+%Dimensionless depth			& $k_p h$				& 2.1		& 0.76
+\end{tabular}
+\end{center}
+\end{table}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+\vsp{10}
 \subsubsection{Johnson's formulation}

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \end{equation}
-Here, $\eta$ has units of length, $t$ time, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$.
+Here, $\eta$ has units of length, $t$ time, $h_0$ is the depth, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$.
 I obtained this equation by starting from Johnson's dimensionless equation (3.28) on page 208 and working backwards to untangle the scales that he had used to non-dimensionalize.
+Alternatively, when the first term is expressed as a spatial derivative, the dimensional KdV equation is
+\begin{equation}
+2 \eta_x + \frac{3}{h_0} \eta \eta_{\xi} + \frac{h_0^2}{3} \eta_{\xi \xi \xi} = 0
+\end{equation}
+Notice the only difference is the absence of the wavespeed $c$ and the first term is now a spatial derivative.
+
+The important thing to remember is that Johnson uses these characteristic scales for non-dimensionalizing (characteristic scales denoted by a *)
+\begin{align}
+\eta^* = \eps h_0 , \quad
+x^* = \eps^{-1/2} h_0 , \quad
+t^* = \eps^{-1/2} h_0/c
+\end{align}
+where $h_0$ is a characteristic depth (say the upstream depth), $c = \sqrt{g h_0}$ the corresponding wave speed, and $\eps = a / h_0$ is the ratio of wave amplitude to depth.
+Roughly speaking, we want to alter Johnson's non-dimensionalization so that $x$ is scaled on some characteristic wavelength (and $t$ scaled accordingly).
+Importantly, we need to account for variable depth.
 

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 
-\subsubsection{Non-dimensionalization}
+\subsubsection{The constant-depth KdV equation in dimensional form}
 
-The KdV equation in dimensional form is
+The constant-depth KdV equation in dimensional form is
 \begin{equation}
@@
 Peak forcing frequency		& $f_p$				& 2 Hz		& N/A \\
-Depth					& $h_0^{\pm}$			& 12.5 cm		& 3 cm \\
+Depth					& $h^{\pm}$			& 12.5 cm		& 3 cm \\
 Peak wavelength			& $\lambda_p^{\pm}$	& 38 cm		& 25 cm \\
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+\subsubsection{The variable-depth KdV equation}
+
+We start with the variable-depth KdV equation given by Johnson in Eq.(3.149) on page 272 in dimensionless form 
+\begin{equation}
+2 H_X + \frac{3}{D^{7/4}} H H_\xi + \frac{1}{3} D^{1/2} H_{\xi \xi \xi} = 0
+\end{equation}
+All variables above are dimensionless, and by working backwards I figured out how they relate to the original, dimensional variables. Below, the dimensional quantities are indicated with an overbar
+\begin{align}
+& H = D^{1/4} \eta = \left( \frac{d}{h_0} \right)^{1/4} \frac{\bar{\eta}}{\eps h_0} \\
+& X = \eps x = \frac{\eps^{3/2}}{h_0} \bar{x} \\
+& \xi = \frac{\eps^{1/2}}{\sqrt{d h_0}} (\bar{x} - c \bar{t})
+\end{align}
+Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the variable depth ratio.
+
+There are a few important things to remember when we reformulate this equation
+\begin{enumerate}
+\item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
+\item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change.
+We have to do this so that the assumption of spatial periodicity will be reasonably satisfies.
+\end{enumerate}
+To achieve 2, we can be guided by the characteristic wavelength from shallow water theory
+\begin{equation}
+\lambda_c = {c}/{f_c} = \sqrt{g d}/f_c
+\end{equation}
+where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed (in shallow-water theory).
+It is tempting to simply use the peak forcing frequency $f_p$ for $f_c$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$.
+Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments).
+Therefore, let us choose
+\begin{equation}
+\lambda_c = M \sqrt{gd} / f_p
+\end{equation}
+where $M$ is an integer (on the order of 10 or so).
 

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \end{equation}
-All variables above are dimensionless, and by working backwards I figured out how they relate to the original, dimensional variables. Below, the dimensional quantities are indicated with an overbar
+All variables above are dimensionless, and by working backwards we can figure out how they relate to the original, dimensional variables.
+Below, the dimensional quantities are indicated with an overbar wherever there is the possibility of ambiguity
 \begin{align}
@@
 \end{equation}
-where $M$ is an integer (on the order of 10 or so).
+where $M$ is an integer (on the order of 10 or so). Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient.
 

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
+\item We want to normalize the standard deviation (or energy) of $\eta$ and therefore want to keep Johnson's choice to scale surface displacement on $a = \eps h_0$.
 \item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change.
-We have to do this so that the assumption of spatial periodicity will be reasonably satisfies.
+We have to do this so that the assumption of spatial periodicity will be reasonably satisfied.
 \end{enumerate}
-To achieve 2, we can be guided by the characteristic wavelength from shallow water theory
+To achieve 3, we can be guided by the characteristic wavelength from shallow water theory
 \begin{equation}
@@
 where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed (in shallow-water theory).
-It is tempting to simply use the peak forcing frequency $f_p$ for $f_c$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$.
+It is tempting to simply set $f_c$ to the peak forcing frequency $f_p$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$.
 Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments).
@@
 \vsp{10}
-\subsubsection{Johnson's formulation}
-Following Johnson's derivation, the dimensionless KdV equation is (on page 208)
-\begin{equation}
-2 \eta_{\tau} + 3 \eta \eta_{\xi} + \frac{1}{3} \eta_{\xi \xi \xi} = 0
-\end{equation}
-I had to work backwards quite a bit to figure out how this was non-dimensionalized (Johnson keeps rescaling things over and over again). Ultimately, the characteristic scales are
-\begin{align}
-& X^* = \eps^{-1/2} h_0 ,\quad T^* = \eps^{-1/2} \sqrt{h_0/g} , \quad \eta^* = \eps h_0, \\
-& Z^* = h_0 , \quad U^* = \eps \sqrt{g h_0} , \quad W^* = \eps^{3/2} \sqrt{g h_0} , \quad P^* = \eps \rho g h_0
-\end{align}
-where $\eps$ is the ratio of characteristic amplitude to depth
-\begin{equation}
-\eps = a / h_0
-\end{equation}
-The variables $\tau$ and $\xi$ are
-\begin{align}
-\tau = \eps t , \quad \xi = x-t
-\end{align}
-Remember, the privileged KdV scaling is $\delta^2 \sim \eps$ where $\delta = h_0 / \lambda$.
-If this scaling is inserted into above, then $X^* \sim \lambda$, $T^* \sim \lambda/\sqrt{g h_0}$ which all makes sense.
 
-If we convert KdV back into dimensional form, we get
-\begin{equation}
-2 \eta_t + \frac{3 c}{h_0} \eta \eta_{\xi} + \frac{c h_0^2}{3} \eta_{\xi \xi \xi} = 0
-\end{equation}
-Here, $\eta$ has units of length, $t$ time, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$.
+\subsubsection{Miscellaneous }
 
@@
 
-We now seek to non-dimensionalize the KdV equation, but not exactly the same way that Johnson did. Some things we would like to accomplish are:
-\begin{enumerate}
-\item When recast in dimensionless variables, $\eta$ should be periodic on the interval $[0, 2\pi)$ in order to make the Fourier notation simple.
-\item We want to rescale $\eta$ to have unit energy in order to make the Hamiltonian sampling simple.
-\end{enumerate}
-First, we observe there are a few lengthscales in the experiment that may be relevant: the upstream depth $h^- = 12.5$ cm, the tank width $L_0 = 20$ cm, the peak wavelength $\lambda_p^- = 38$ cm (induced by the peak forcing frequency as calculated by the dispersion relation), and possibly the maximum wavelength present $\lambda_{max}^{-} = ?$.
-First, since we want $\eta$ to be periodic on $[0,2 \pi)$ in dimensionless variables, the most logical choice for scaling $x$ is a wavelength, either the peak or max.
-Lets choose the peak.
-Next, since we want dimensionless $\eta$ to have unit energy, we {\em must} scale $\eta$ by the characteristic amplitude $a = \etastd = \mean{\eta^2}^{1/2}$, where the mean is taken over a sufficiently long time or distance.
 We therefore introduce dimensionless variables, denoted by a tilde
@@
 \end{align}
-
 Then in dimensionless variables (after dropping the tildes), the system is given by
@@
 Of course, the effect of this rescaling would be to introduce a factor of $\beta$ in front of the first term $\eta_{\tau}$.
-
-Finally, since we want to study variable-depth KdV, we will ultimately want to interchange the roles of space and time as Johnson does on p.
-269.
-So we will have to modify some of this non-dimensionalization.
-I had to wrap my mind around the classical case first though.
-
-\subsubsection{Comparison to variables in Andy's original notes}
-In Andy's original notes (as well as in Di Qi's latex notes), the width of the tank $L_0$ was used for non-dimensionalizing $x$ (as well as $\xi$).
-In comparing our two formulations, the main subtlety is the energy, defined as
-\begin{equation}
-E_0 = \frac{1}{2} \int_{0}^{2 \pi L_0} u^2 dx
-\end{equation}
-where $u$ is displacement and has units of length.
-Thus, $E_0$ has units of length cubed (it took me a while to realize this is the main subtlety in reconciling the two formulations).
-The standard deviation of displacement is
-\begin{equation}
-u_{std} = \frac{1}{\sqrt{2 \pi L_0}} \left( \int_{0}^{2\pi L_0} u^2 dx \right)^{1/2}
-\end{equation}
-or in Andy's / Di Qi's notation
-\begin{equation}
-u_{std} = const \times L_0^{-1/2} E_0^{1/2}
-\end{equation}
-where I really don't care what the constant is. The main thing I wanted to see was why $L_0^{-1/2}$ appears, and the above shows this.
-
-The main difficulty I have with this formulation is that there is no reason to expect the spatial period of $u$ to be related to the tank width $L_0$.
-More logically, the spatial period of $u$ should correspond to a characteristic wavelength in the problem, either the peak or maximum wavelength.
-The latter is more likely to be true in a literal sense, but I believe the peak wavelength provides a more reasonable scale for non-dimensionalization.
-
-Otherwise, the two formulations are completely equivalent.
-Notice that $\mu \lambda^{-3}$ in Andy's notation plays the role of $\delta^2$ in my notation, and $E_0^{1/2} \lambda^{-3/2}$ in Andy's notation plays the role of $\eps$ in my notation.
-
-\subsubsection{Reformulating variable-depth KdV}
-
-We now seek to formulate the variable-depth KdV in a similar fashion. We start with Johnson's Eq. (3.148), which is written in dimensionless variables
-\begin{equation}
-2 \sqrt{D} \eta_X + \frac{1}{2} \frac{D'(X)}{\sqrt{D}} \eta + \frac{3}{D} \eta \eta_\xi + \frac{1}{3} D \eta_{\xi \xi \xi} = 0
-\end{equation}
-The first task is to convert this to dimensional form. For this, we now denote dimensionless variables with a tilde and use the relationships
-\begin{align}
-\eta = \eps h_0 \tilde{\eta} , \quad x = \eps^{-1/2} h_0 \tilde{x} = \eps^{-3/2} h_0 \tilde{X}
-\end{align}
-The trickiest part relates to $\xi$.
 

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \end{align}
-Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the variable depth ratio.
+Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the depth ratio.
 
-There are a few important things to remember when we reformulate this equation
+There are a few important things to remember when we reformulate this equation:
 \begin{enumerate}
@@
 \item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change.
-We have to do this so that the assumption of spatial periodicity will be reasonably satisfied.
+We have to do this so that the assumption of spatial periodicity will be reasonably satisfied on both sides.
 \end{enumerate}
-To achieve 3, we can be guided by the characteristic wavelength from shallow water theory
+To achieve 3, we can use the characteristic wavelength from shallow water theory as a guide,
 \begin{equation}
@@
 Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments).
-Therefore, let us choose
+We therefore choose
 \begin{equation}

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \begin{equation}
+\label{varKdVJohnson}
 2 H_X + \frac{3}{D^{7/4}} H H_\xi + \frac{1}{3} D^{1/2} H_{\xi \xi \xi} = 0
@@
 \begin{equation}
-\lambda_c = M \sqrt{gd} / f_p
+\lambda_c = M \lambda_p = M \sqrt{gd} / f_p
 \end{equation}
-where $M$ is an integer (on the order of 10 or so). Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient.
+where $M$ is an integer (on the order of 10 or so).
+Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient.
+Thus, the main change we need to make is to introduce the dimensionless variable
+\begin{equation}
+\zeta = \frac{2\pi}{\lambda_c} (\bar{x} - c \bar{t}) 
+= \frac{2 \pi \delta}{M \eps^{1/2}} \, \xi
+\end{equation}
+where
+\begin{equation}
+\delta = h_0 / \lambda_0 = f_p \sqrt{h_0/g}
+\end{equation}
+To clean things up a little, lets also introduce a slightly different long-space coordinate that uses the upstream wavelength $\lambda_0$ as a characteristic scale instead of $h_0$,
+\begin{equation}
+s = \frac{2 \pi \beta}{\lambda_0} \bar{x} = 2 \pi \beta \delta \eps^{-3/2} X
+\end{equation}
+where $\beta$ is a small parameter that can be taken to be, for example, $\eps$ or $\delta^2$ or something else.
+
+Then Eq.~\eqref{varKdVJohnson} becomes
+\begin{equation}
+\label{varKdVugly}
+2 \beta H_s + \frac{3 \eps}{M} D^{-7/4} H H_\zeta
++ \frac{4 \pi^2 \delta^2}{3 M^3} D^{1/2} H_{\zeta \zeta \zeta} = 0
+\end{equation}
+
+\subsubsection{Final formulation}
+Equation \eqref{varKdVugly} is essentially in the form we want to use, except that the notation is bad.
+In particular, we will want to reserve the symbol $H$ for the Hamiltonian.
+So making superficial changes to the notation gives
+\begin{align}
+2 \beta u_s + \frac{3 \eps}{M} D^{-7/4} \, u u_x
++ \frac{4 \pi^2 \delta^2}{3 M^3} D^{1/2} \, u_{xxx} = 0
+\end{align}
+The definition of each of these dimensionless variables in terms of the original, physical ones are
+\begin{align}
+& u = D^{1/4} \, \bar{\eta}/a \\
+& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
+& s = 2 \pi \beta \lambda_0^{-1} \bar{x}
+\end{align}
 

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \begin{equation}
-s = \frac{2 \pi \beta}{\lambda_0} \bar{x} = 2 \pi \beta \delta \eps^{-3/2} X
+s = \frac{2 \pi \beta}{M \lambda_0} \bar{x} = 2 \pi \beta M^{-1} \delta \eps^{-3/2} X
 \end{equation}
@@
 \label{varKdVugly}
-2 \beta H_s + \frac{3 \eps}{M} D^{-7/4} H H_\zeta
-+ \frac{4 \pi^2 \delta^2}{3 M^3} D^{1/2} H_{\zeta \zeta \zeta} = 0
+2 \beta H_s + 3 \eps D^{-7/4} H H_\zeta
++ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} H_{\zeta \zeta \zeta} = 0
 \end{equation}
@@
 \begin{align}
-2 \beta u_s + \frac{3 \eps}{M} D^{-7/4} \, u u_x
-+ \frac{4 \pi^2 \delta^2}{3 M^3} D^{1/2} \, u_{xxx} = 0
+2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
++ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} \, u_{xxx} = 0
 \end{align}
-The definition of each of these dimensionless variables in terms of the original, physical ones are
+Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.
+The definitions of each of the dimensionless variables in terms of the original, physical ones (denoted with a bar) are
 \begin{align}
@@
 & x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
-& s = 2 \pi \beta \lambda_0^{-1} \bar{x}
+& s = 2 \pi \beta M^{-1} \lambda_0^{-1} \bar{x}
+\end{align}
+A list of the fixed parameters is
+\begin{align}
+& \lambda_0 = \sqrt{g h_0}/f_p \quad \text{(upstream wavelength)} \\
+& \eps = a / h_0 \quad \text{(dimensionless ratio of wave amplitude to depth)} \\
+& \delta = h_0 / \lambda_0 \quad \text{(dimensionless ratio of depth to upstream wavelength)} \\
+& \beta	\quad \text{(an unspecified small parameter, usually $\eps$ or $\delta^2$)}
+\end{align}
+Here is a list of the parameters that vary in space due to the fact that depth $d$ varies in space
+\begin{align}
+&\lambda_c = M \lambda_p = M \sqrt{gd}/f_p \quad \text{(characteristic wavelength as it varies in space)} \\
+&D = d/h_0 \quad \text{(dimensionless depth ratio as it varies in space)} \\	
 \end{align}

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \end{align}
-
-\vsp{10}
-
-\subsubsection{Miscellaneous }
-
-\subsubsection{My reformulation}
-
-We therefore introduce dimensionless variables, denoted by a tilde
-\begin{align}
-& x = \frac{\lambda_p}{2\pi} \tilde{x} , \quad 
-t = \frac{\lambda_p}{2\pi c} \tilde{t} , \quad
-\eta = a \tilde{\eta} \\
-& \xi = x-ct = \frac{\lambda_p}{2\pi} (\tilde{x} - \tilde{t}) = \frac{\lambda_p}{2\pi} \tilde{\xi}
-\end{align}
-Then in dimensionless variables (after dropping the tildes), the system is given by
-\begin{equation}
-\eta_t + \frac{3}{2}\eps \eta \eta_{\xi} + \frac{2 \pi^2}{3} \delta^2 \eta_{\xi \xi \xi} = 0
-\end{equation}
-where we have the dimensionless variables
-\begin{align}
-\eps = a/h_0 , \quad \delta = h_0/\lambda_p
-\end{align}
-In these dimensionless variables, we have that $\eta$ is periodic for $x \in [0, 2\pi)$ and we have $\mean{\eta^2} = 1$, just as intended.
 Note, in the experiments, $\delta = 0.33$ and $\eps$ lies in the range $2.4 \times 10^{-3}$--$2.4 \times 10^{-2}$.
-
-We will eventually want to rescale $t$, to look at long-time scales as is typically done in KdV analysis. Accordingly, we will define
-\begin{equation}
-\tau = \beta t
-\end{equation}
-where $\beta$ is usually either $\eps$ or $\delta^2$ (or we can choose it to be something more appropriate for the experiments).
-Of course, the effect of this rescaling would be to introduce a factor of $\beta$ in front of the first term $\eta_{\tau}$.
 

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the depth ratio.
+The last relationship above for $\xi$ is valid for the case of two depths (I am not sure if it is valid in the more general case of variable depth).
 
@@
 \begin{align}
+& c = \sqrt{gd} \quad \text{(the wavespeed as it varies in space)} \\
 &\lambda_c = M \lambda_p = M \sqrt{gd}/f_p \quad \text{(characteristic wavelength as it varies in space)} \\

### 2018-09-27 LatexNotes/Notes.tex
NM

@@
 \begin{align}
+\label{varKdV1}
 2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
@@
 
-\np
-\subsection{Hamiltonian structure}
-We will be using the truncated KdV equations (TKDV) with variable depth, in particular an abrupt depth change (ADC).
-The upstream depth is normalized to $1$ and the downstream (dimensionless) depth is $D_0$.
-Accordingly, the upstream and downstream Hamiltonians are given by
+Let's make one more cosmetic change. Divide Eq.~\eqref{varKdV1} by $3 \delta^2$ and set $\beta = 3\delta^2/2$ to get
+\begin{align}
+\label{varKdV}
+u_s + \eps \delta^{-2} D^{-7/4} \, u u_x
++ \mu D^{1/2} \, u_{xxx} = 0
+\end{align}
+where
+\begin{equation}
+\mu = \frac{4 \pi^2}{9 M^2}
+\end{equation}
+
+Then the Hamiltonian structure is given by
 \begin{align}
 \label{Hamiltonian}
-& H^- = E_0^{1/2} H_3 - H_2 \\
-& H^+ = D_0^{-13/4} E_0^{1/2} H_3 - D_0^{3/2} H_2 \\
+& H^- = \eps \delta^{-2} H_3 - \mu H_2 \\
+& H^+ = \eps \delta^{-2} D_0^{-7/4} H_3 - \mu D_0^{1/2} H_2
 \end{align}
 
+\np
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-09-28 LatexNotes/Notes.tex
NM

@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% KdV
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The KdV equations}
+
+\subsection{The constant-depth KdV equation in dimensional form}
+
+The constant-depth KdV equation in dimensional form is
+\begin{equation}
+2 \eta_t + \frac{3 c}{h_0} \eta \eta_{\xi} + \frac{c h_0^2}{3} \eta_{\xi \xi \xi} = 0
+\end{equation}
+Here, $\eta$ has units of length, $t$ time, $h_0$ is the depth, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$.
+I obtained this equation by starting from Johnson's dimensionless equation (3.28) on page 208 and working backwards to untangle the scales that he had used to non-dimensionalize.
+Alternatively, when the first term is expressed as a spatial derivative, the dimensional KdV equation is
+\begin{equation}
+2 \eta_x + \frac{3}{h_0} \eta \eta_{\xi} + \frac{h_0^2}{3} \eta_{\xi \xi \xi} = 0
+\end{equation}
+Notice the only difference is the absence of the wavespeed $c$ and the first term is now a spatial derivative.
+
+The important thing to remember is that Johnson uses these characteristic scales for non-dimensionalizing (characteristic scales denoted by a *)
+\begin{align}
+\eta^* = \eps h_0 , \quad
+x^* = \eps^{-1/2} h_0 , \quad
+t^* = \eps^{-1/2} h_0/c
+\end{align}
+where $h_0$ is a characteristic depth (say the upstream depth), $c = \sqrt{g h_0}$ the corresponding wave speed, and $\eps = a / h_0$ is the ratio of wave amplitude to depth.
+We want to alter Johnson's non-dimensionalization so that $x$ is scaled on some characteristic wavelength (and $t$ scaled accordingly).
+Importantly, we need to account for variable depth.
+
+Let's establish some basic scales and notation:
+% Table
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{table}[h]%[htbp]
+\begin{center}
+\caption{Table.}
+\label{scale_table}
+\begin{tabular}{l l l l}
+\hline Description & Notation & Upstream  value \,\, & Downstream value \\
+\hline
+Characteristic wave amplitude	& $a = \etastd = \mean{\eta^2}^{1/2}$	& varied	& varied \\
+Peak forcing frequency		& $f_p$				& 2 Hz		& N/A \\
+Depth					& $h^{\pm}$			& 12.5 cm		& 3 cm \\
+Peak wavelength			& $\lambda_p^{\pm}$	& 38 cm		& 25 cm \\
+%Peak wavenumber			& $k_p$				& 0.17 rad/cm	& 0.25 rad/cm \\
+%Dimensionless depth			& $k_p h$				& 2.1		& 0.76
+\end{tabular}
+\end{center}
+\end{table}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+\subsection{The variable-depth KdV equation}
+
+We start with the variable-depth KdV equation given by Johnson in Eq.(3.149) on page 272 in dimensionless form 
+\begin{equation}
+\label{varKdVJohnson}
+2 H_X + \frac{3}{D^{7/4}} H H_\xi + \frac{1}{3} D^{1/2} H_{\xi \xi \xi} = 0
+\end{equation}
+All variables above are dimensionless, and by working backwards we can figure out how they relate to the original, dimensional variables.
+Below, the dimensional quantities are indicated with an overbar wherever there is the possibility of ambiguity
+\begin{align}
+& H = D^{1/4} \eta = \left( \frac{d}{h_0} \right)^{1/4} \frac{\bar{\eta}}{\eps h_0} \\
+& X = \eps x = \frac{\eps^{3/2}}{h_0} \bar{x} \\
+& \xi = \frac{\eps^{1/2}}{\sqrt{d h_0}} (\bar{x} - c \bar{t})
+\end{align}
+Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the depth ratio.
+The last relationship above for $\xi$ is valid for the case of two depths (though perhaps not in the general case).
+
+There are a few important things to remember when we reformulate this equation:
+\begin{enumerate}
+\item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
+\item We want to normalize the standard deviation (or energy) of $\eta$ and therefore want to keep Johnson's choice to scale surface displacement on $a = \eps h_0$.
+\item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change.
+We have to do this so that the assumption of spatial periodicity will be reasonably satisfied on both sides.
+\end{enumerate}
+To achieve 3, we can use the characteristic wavelength from shallow water theory,
+\begin{equation}
+\lambda_c = {c}/{f_c} = \sqrt{g d}/f_c
+\end{equation}
+where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed.
+It is tempting to simply set $f_c$ to the peak forcing frequency $f_p$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$.
+Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments).
+We therefore choose
+\begin{equation}
+\lambda_c = M \lambda_p = M \sqrt{gd} / f_p
+\end{equation}
+where $M$ is an integer (on the order of 10 or so).
+Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient.
+Thus, the main change we need to make is to introduce the dimensionless variable
+\begin{equation}
+\zeta = \frac{2\pi}{\lambda_c} (\bar{x} - c \bar{t}) 
+= \frac{2 \pi \delta}{M \eps^{1/2}} \, \xi
+\end{equation}
+where
+\begin{equation}
+\delta = h_0 / \lambda_0 = f_p \sqrt{h_0/g}
+\end{equation}
+To clean things up a little, lets also introduce a slightly different long-space coordinate that uses the upstream wavelength $\lambda_0$ as a characteristic scale instead of $h_0$,
+\begin{equation}
+s = \frac{2 \pi \beta}{M \lambda_0} \bar{x} = 2 \pi \beta M^{-1} \delta \eps^{-3/2} X
+\end{equation}
+where $\beta$ is a small parameter that can be taken to be, for example, $\eps$ or $\delta^2$ or something else.
+
+Then Eq.~\eqref{varKdVJohnson} becomes
+\begin{equation}
+\label{varKdVugly}
+2 \beta H_s + 3 \eps D^{-7/4} H H_\zeta
++ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} H_{\zeta \zeta \zeta} = 0
+\end{equation}
+
+\subsection{Final formulation}
+Equation \eqref{varKdVugly} is essentially in the form we want to use, except that the notation is bad.
+In particular, we will want to reserve the symbol $H$ for the Hamiltonian.
+So making superficial changes to the notation gives
+\begin{align}
+\label{varKdV1}
+2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
++ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} \, u_{xxx} = 0
+\end{align}
+Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.
+The definitions of each of the dimensionless variables in terms of the original, physical ones (denoted with a bar) are
+\begin{align}
+& u = D^{1/4} \, \bar{\eta}/a \\
+& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
+& s = 2 \pi \beta M^{-1} \lambda_0^{-1} \bar{x}
+\end{align}
+A list of the fixed parameters is
+\begin{align}
+& \lambda_0 = \sqrt{g h_0}/f_p \quad \text{(upstream wavelength)} \\
+& \eps = a / h_0 \quad \text{(dimensionless ratio of wave amplitude to depth)} \\
+& \delta = h_0 / \lambda_0 \quad \text{(dimensionless ratio of depth to upstream wavelength)} \\
+& \beta	\quad \text{(an unspecified small parameter, usually $\eps$ or $\delta^2$)}
+\end{align}
+Here is a list of the parameters that vary in space due to the fact that depth $d$ varies in space
+\begin{align}
+& c = \sqrt{gd} \quad \text{(the wavespeed as it varies in space)} \\
+&\lambda_c = M \lambda_p = M \sqrt{gd}/f_p \quad \text{(characteristic wavelength as it varies in space)} \\
+&D = d/h_0 \quad \text{(dimensionless depth ratio as it varies in space)} \\	
+\end{align}
+Note, in the experiments, $\delta = 0.33$ and $\eps$ lies in the range $2.4 \times 10^{-3}$--$2.4 \times 10^{-2}$.
+
+Let's make one more cosmetic change. Divide Eq.~\eqref{varKdV1} by $3 \delta^2$ and set $\beta = 3\delta^2/2$ to get
+\begin{align}
+\label{varKdV}
+u_s + \eps \delta^{-2} D^{-7/4} \, u u_x
++ \mu D^{1/2} \, u_{xxx} = 0
+\end{align}
+where
+\begin{equation}
+\mu = \frac{4 \pi^2}{9 M^2}
+\end{equation}
+
+Then the Hamiltonian structure is given by
+\begin{align}
+\label{Hamiltonian}
+& H^- = \eps \delta^{-2} H_3 - \mu H_2 \\
+& H^+ = \eps \delta^{-2} D_0^{-7/4} H_3 - \mu D_0^{1/2} H_2
+\end{align}
+
+Note: $\eps \delta^{-2} = a f_p^2 / g$ which is an interesting parameter. It suggests an equivalence between driving amplitude and driving frequency.
 
@@
 Only functions that correspond to a Fourier series truncated at wavenumber $\Lambda$ are permissible.
-
-\np
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% TKDV
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The truncated KdV equations}
-
-\subsection{The KdV equation}
-
-\subsubsection{The constant-depth KdV equation in dimensional form}
-
-The constant-depth KdV equation in dimensional form is
-\begin{equation}
-2 \eta_t + \frac{3 c}{h_0} \eta \eta_{\xi} + \frac{c h_0^2}{3} \eta_{\xi \xi \xi} = 0
-\end{equation}
-Here, $\eta$ has units of length, $t$ time, $h_0$ is the depth, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$.
-I obtained this equation by starting from Johnson's dimensionless equation (3.28) on page 208 and working backwards to untangle the scales that he had used to non-dimensionalize.
-Alternatively, when the first term is expressed as a spatial derivative, the dimensional KdV equation is
-\begin{equation}
-2 \eta_x + \frac{3}{h_0} \eta \eta_{\xi} + \frac{h_0^2}{3} \eta_{\xi \xi \xi} = 0
-\end{equation}
-Notice the only difference is the absence of the wavespeed $c$ and the first term is now a spatial derivative.
-
-The important thing to remember is that Johnson uses these characteristic scales for non-dimensionalizing (characteristic scales denoted by a *)
-\begin{align}
-\eta^* = \eps h_0 , \quad
-x^* = \eps^{-1/2} h_0 , \quad
-t^* = \eps^{-1/2} h_0/c
-\end{align}
-where $h_0$ is a characteristic depth (say the upstream depth), $c = \sqrt{g h_0}$ the corresponding wave speed, and $\eps = a / h_0$ is the ratio of wave amplitude to depth.
-Roughly speaking, we want to alter Johnson's non-dimensionalization so that $x$ is scaled on some characteristic wavelength (and $t$ scaled accordingly).
-Importantly, we need to account for variable depth.
-
-Let's establish some basic scales and notation:
-% Table
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{table}[h]%[htbp]
-\begin{center}
-\caption{Table.}
-\label{scale_table}
-\begin{tabular}{l l l l}
-\hline Description & Notation & Upstream  value \,\, & Downstream value \\
-\hline
-Characteristic wave amplitude	& $a = \etastd = \mean{\eta^2}^{1/2}$	& varied	& varied \\
-Peak forcing frequency		& $f_p$				& 2 Hz		& N/A \\
-Depth					& $h^{\pm}$			& 12.5 cm		& 3 cm \\
-Peak wavelength			& $\lambda_p^{\pm}$	& 38 cm		& 25 cm \\
-%Peak wavenumber			& $k_p$				& 0.17 rad/cm	& 0.25 rad/cm \\
-%Dimensionless depth			& $k_p h$				& 2.1		& 0.76
-\end{tabular}
-\end{center}
-\end{table}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-\subsubsection{The variable-depth KdV equation}
-
-We start with the variable-depth KdV equation given by Johnson in Eq.(3.149) on page 272 in dimensionless form 
-\begin{equation}
-\label{varKdVJohnson}
-2 H_X + \frac{3}{D^{7/4}} H H_\xi + \frac{1}{3} D^{1/2} H_{\xi \xi \xi} = 0
-\end{equation}
-All variables above are dimensionless, and by working backwards we can figure out how they relate to the original, dimensional variables.
-Below, the dimensional quantities are indicated with an overbar wherever there is the possibility of ambiguity
-\begin{align}
-& H = D^{1/4} \eta = \left( \frac{d}{h_0} \right)^{1/4} \frac{\bar{\eta}}{\eps h_0} \\
-& X = \eps x = \frac{\eps^{3/2}}{h_0} \bar{x} \\
-& \xi = \frac{\eps^{1/2}}{\sqrt{d h_0}} (\bar{x} - c \bar{t})
-\end{align}
-Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the depth ratio.
-The last relationship above for $\xi$ is valid for the case of two depths (I am not sure if it is valid in the more general case of variable depth).
-
-There are a few important things to remember when we reformulate this equation:
-\begin{enumerate}
-\item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
-\item We want to normalize the standard deviation (or energy) of $\eta$ and therefore want to keep Johnson's choice to scale surface displacement on $a = \eps h_0$.
-\item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change.
-We have to do this so that the assumption of spatial periodicity will be reasonably satisfied on both sides.
-\end{enumerate}
-To achieve 3, we can use the characteristic wavelength from shallow water theory as a guide,
-\begin{equation}
-\lambda_c = {c}/{f_c} = \sqrt{g d}/f_c
-\end{equation}
-where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed (in shallow-water theory).
-It is tempting to simply set $f_c$ to the peak forcing frequency $f_p$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$.
-Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments).
-We therefore choose
-\begin{equation}
-\lambda_c = M \lambda_p = M \sqrt{gd} / f_p
-\end{equation}
-where $M$ is an integer (on the order of 10 or so).
-Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient.
-Thus, the main change we need to make is to introduce the dimensionless variable
-\begin{equation}
-\zeta = \frac{2\pi}{\lambda_c} (\bar{x} - c \bar{t}) 
-= \frac{2 \pi \delta}{M \eps^{1/2}} \, \xi
-\end{equation}
-where
-\begin{equation}
-\delta = h_0 / \lambda_0 = f_p \sqrt{h_0/g}
-\end{equation}
-To clean things up a little, lets also introduce a slightly different long-space coordinate that uses the upstream wavelength $\lambda_0$ as a characteristic scale instead of $h_0$,
-\begin{equation}
-s = \frac{2 \pi \beta}{M \lambda_0} \bar{x} = 2 \pi \beta M^{-1} \delta \eps^{-3/2} X
-\end{equation}
-where $\beta$ is a small parameter that can be taken to be, for example, $\eps$ or $\delta^2$ or something else.
-
-Then Eq.~\eqref{varKdVJohnson} becomes
-\begin{equation}
-\label{varKdVugly}
-2 \beta H_s + 3 \eps D^{-7/4} H H_\zeta
-+ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} H_{\zeta \zeta \zeta} = 0
-\end{equation}
-
-\subsubsection{Final formulation}
-Equation \eqref{varKdVugly} is essentially in the form we want to use, except that the notation is bad.
-In particular, we will want to reserve the symbol $H$ for the Hamiltonian.
-So making superficial changes to the notation gives
-\begin{align}
-\label{varKdV1}
-2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
-+ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} \, u_{xxx} = 0
-\end{align}
-Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.
-The definitions of each of the dimensionless variables in terms of the original, physical ones (denoted with a bar) are
-\begin{align}
-& u = D^{1/4} \, \bar{\eta}/a \\
-& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
-& s = 2 \pi \beta M^{-1} \lambda_0^{-1} \bar{x}
-\end{align}
-A list of the fixed parameters is
-\begin{align}
-& \lambda_0 = \sqrt{g h_0}/f_p \quad \text{(upstream wavelength)} \\
-& \eps = a / h_0 \quad \text{(dimensionless ratio of wave amplitude to depth)} \\
-& \delta = h_0 / \lambda_0 \quad \text{(dimensionless ratio of depth to upstream wavelength)} \\
-& \beta	\quad \text{(an unspecified small parameter, usually $\eps$ or $\delta^2$)}
-\end{align}
-Here is a list of the parameters that vary in space due to the fact that depth $d$ varies in space
-\begin{align}
-& c = \sqrt{gd} \quad \text{(the wavespeed as it varies in space)} \\
-&\lambda_c = M \lambda_p = M \sqrt{gd}/f_p \quad \text{(characteristic wavelength as it varies in space)} \\
-&D = d/h_0 \quad \text{(dimensionless depth ratio as it varies in space)} \\	
-\end{align}
-Note, in the experiments, $\delta = 0.33$ and $\eps$ lies in the range $2.4 \times 10^{-3}$--$2.4 \times 10^{-2}$.
-
-Let's make one more cosmetic change. Divide Eq.~\eqref{varKdV1} by $3 \delta^2$ and set $\beta = 3\delta^2/2$ to get
-\begin{align}
-\label{varKdV}
-u_s + \eps \delta^{-2} D^{-7/4} \, u u_x
-+ \mu D^{1/2} \, u_{xxx} = 0
-\end{align}
-where
-\begin{equation}
-\mu = \frac{4 \pi^2}{9 M^2}
-\end{equation}
-
-Then the Hamiltonian structure is given by
-\begin{align}
-\label{Hamiltonian}
-& H^- = \eps \delta^{-2} H_3 - \mu H_2 \\
-& H^+ = \eps \delta^{-2} D_0^{-7/4} H_3 - \mu D_0^{1/2} H_2
-\end{align}
 

### 2018-09-28 LatexNotes/Notes.tex
NM

@@
 \newcommand{\Gibbs}{\mathcal{G}}
+\newcommand{\ampp}{\mathcal{A}}
 
@@
 \caption{Basic experimental parameters and scales.}
-\label{paramtable}
+\label{exptable}
 \begin{tabular}{l l l}
@@
 \hline
-peak forcing frequency			& 2 Hz			& N/A \\
+peak forcing frequency, $f_p$\,\,		& 2 Hz			& N/A \\
 depth, $h$						& 12.5 cm			& 3 cm \\
@@
 where $h_0$ is a characteristic depth (say the upstream depth), $c = \sqrt{g h_0}$ the corresponding wave speed, and $\eps = a / h_0$ is the ratio of wave amplitude to depth.
-We want to alter Johnson's non-dimensionalization so that $x$ is scaled on some characteristic wavelength (and $t$ scaled accordingly).
+We want to alter Johnson's non-dimensionalization so that $x$ is instead scaled on a characteristic wavelength (and $t$ scaled accordingly).
 Importantly, we need to account for variable depth.
-
-Let's establish some basic scales and notation:
-% Table
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{table}[h]%[htbp]
-\begin{center}
-\caption{Table.}
-\label{scale_table}
-\begin{tabular}{l l l l}
-\hline Description & Notation & Upstream  value \,\, & Downstream value \\
-\hline
-Characteristic wave amplitude	& $a = \etastd = \mean{\eta^2}^{1/2}$	& varied	& varied \\
-Peak forcing frequency		& $f_p$				& 2 Hz		& N/A \\
-Depth					& $h^{\pm}$			& 12.5 cm		& 3 cm \\
-Peak wavelength			& $\lambda_p^{\pm}$	& 38 cm		& 25 cm \\
-%Peak wavenumber			& $k_p$				& 0.17 rad/cm	& 0.25 rad/cm \\
-%Dimensionless depth			& $k_p h$				& 2.1		& 0.76
-\end{tabular}
-\end{center}
-\end{table}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 
@@
 \end{equation}
-where $\beta$ is a small parameter that can be taken to be, for example, $\eps$ or $\delta^2$ or something else.
+where $\beta$ is a small parameter that scale like $\eps$ or $\delta^2$ for example.
 
@@
 
-\subsection{Final formulation}
+\subsection{Final formulation of variable-depth KdV}
+
 Equation \eqref{varKdVugly} is essentially in the form we want to use, except that the notation is bad.
 In particular, we will want to reserve the symbol $H$ for the Hamiltonian.
-So making superficial changes to the notation gives
+So lets introduce some new variables
+\begin{align}
+& u = D^{1/4} \, \bar{\eta}/a \\
+& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
+& s = 2 \pi \beta M^{-1} \lambda_0^{-1} \bar{x}
+\end{align}
+Above, the original physical variables are denoted with a bar. Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.
+
+% Table
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{table}[h]%[htbp]
+\begin{center}
+\caption{Table of parameters}
+\label{paramtable}
+\begin{tabular}{l l l l}
+\hline \multicolumn{4} { c }{Parameters that are constant throughout the domain} \\
+\hline Description & Notation and definition & Units & Value in experiments \\
+\hline
+Peak forcing frequency		& $f_p$						& frequency	& 2 Hz \\
+Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2}$		& length		& 0.03--0.3 cm \\
+Upstream wavelength			& $\lambda_0 = \sqrt{g h_0}/f_p$	& length		& 56 cm \\
+Amplitude-to-depth ratio		& $\eps = a / h_0$				& dimensionless	& 0.0024--0.024 \\
+Depth-to-wavelength ratio		& $\delta = h_0 / \lambda_0$		& dimensionless		& 0.22 \\
+Amplitude parameter			& $\ampp = \eps \delta^{-2} = ag h_0^{-2} f_p^{-2}$	
+& dimensionless		& 0.05--0.5\\
+Catch-all parameter		& $\mu = 4 \pi^2 / (9 M^2)$			& dimensionless		& 0.27 if $M=4$ \\
+\hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
+\hline Description & Notation and definition & Value in experiments \\
+\hline
+Local depth			& $d$					& length \\
+Local wavespeed		& $c = \sqrt{gd}$			& speed \\
+Local peak wavelength	& $\lambda_p = \sqrt{gd}/f_p$	& length \\
+Characteristic wavelength	& $\lambda_c = M \lambda_p$	& length \\
+Depth ratio			& $D = d/h_0$				& dimensionless
+\end{tabular}
+\end{center}
+\end{table}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+In Table \ref{paramtable} we provide a list of all the relevant parameters needed to convert these dimensional variables to physical ones.
+Note that the wavelength $\lambda_0$ is predicted from shallow-water theory and so it differs somewhat from the value in Table \ref{exptable} obtained from the dispersion relation.
+In terms of the new variables,  Eq.~\eqref{varKdVugly} becomes
 \begin{align}
@@
 \end{align}
-Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.
-The definitions of each of the dimensionless variables in terms of the original, physical ones (denoted with a bar) are
+
+Let's clean it up now a little bit. Let us set $\beta = 3 \delta^2/2$ and introduce the parameters
 \begin{align}
-& u = D^{1/4} \, \bar{\eta}/a \\
-& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
-& s = 2 \pi \beta M^{-1} \lambda_0^{-1} \bar{x}
+\ampp = \eps \delta^{-2} = \frac{ag}{h_0^2 f_p^2} , \qquad
+\mu = \frac{4 \pi^2}{9 M^2}
 \end{align}
-A list of the fixed parameters is
-\begin{align}
-& \lambda_0 = \sqrt{g h_0}/f_p \quad \text{(upstream wavelength)} \\
-& \eps = a / h_0 \quad \text{(dimensionless ratio of wave amplitude to depth)} \\
-& \delta = h_0 / \lambda_0 \quad \text{(dimensionless ratio of depth to upstream wavelength)} \\
-& \beta	\quad \text{(an unspecified small parameter, usually $\eps$ or $\delta^2$)}
-\end{align}
-Here is a list of the parameters that vary in space due to the fact that depth $d$ varies in space
-\begin{align}
-& c = \sqrt{gd} \quad \text{(the wavespeed as it varies in space)} \\
-&\lambda_c = M \lambda_p = M \sqrt{gd}/f_p \quad \text{(characteristic wavelength as it varies in space)} \\
-&D = d/h_0 \quad \text{(dimensionless depth ratio as it varies in space)} \\	
-\end{align}
-Note, in the experiments, $\delta = 0.33$ and $\eps$ lies in the range $2.4 \times 10^{-3}$--$2.4 \times 10^{-2}$.
-
-Let's make one more cosmetic change. Divide Eq.~\eqref{varKdV1} by $3 \delta^2$ and set $\beta = 3\delta^2/2$ to get
+The Eq.~\eqref{varKdV1} becomes
 \begin{align}
 \label{varKdV}
-u_s + \eps \delta^{-2} D^{-7/4} \, u u_x
+u_s + \ampp D^{-7/4} \, u u_x
 + \mu D^{1/2} \, u_{xxx} = 0
 \end{align}
-where
-\begin{equation}
-\mu = \frac{4 \pi^2}{9 M^2}
-\end{equation}
 
@@
 \label{Hamiltonian}
-& H^- = \eps \delta^{-2} H_3 - \mu H_2 \\
-& H^+ = \eps \delta^{-2} D_0^{-7/4} H_3 - \mu D_0^{1/2} H_2
+& H^- = \ampp H_3 - \mu H_2 \\
+& H^+ = \ampp D_0^{-7/4} H_3 - \mu D_0^{1/2} H_2
 \end{align}
-
-Note: $\eps \delta^{-2} = a f_p^2 / g$ which is an interesting parameter. It suggests an equivalence between driving amplitude and driving frequency.
 

### 2018-09-29 LatexNotes/Notes.tex
NM

@@
 \newcommand{\ampp}{\mathcal{A}}
+\newcommand{\lamfac}{N}
 
@@
 \begin{equation}
-\lambda_c = M \lambda_p = M \sqrt{gd} / f_p
+\lambda_c = \lamfac \lambda_p = \lamfac \sqrt{gd} / f_p
 \end{equation}
-where $M$ is an integer (on the order of 10 or so).
+where $\lamfac$ is an integer (on the order of 10 or so).
 Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient.
@@
 \zeta = \frac{2\pi}{\lambda_c} (\bar{x} - c \bar{t}) 
-= \frac{2 \pi \delta}{M \eps^{1/2}} \, \xi
+= \frac{2 \pi \delta}{\lamfac \eps^{1/2}} \, \xi
 \end{equation}
@@
 \begin{equation}
-s = \frac{2 \pi \beta}{M \lambda_0} \bar{x} = 2 \pi \beta M^{-1} \delta \eps^{-3/2} X
+s = \frac{2 \pi \beta}{\lamfac \lambda_0} \bar{x} = 2 \pi \beta \lamfac^{-1} \delta \eps^{-3/2} X
 \end{equation}
@@
 & x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
-& s = 2 \pi \beta M^{-1} \lambda_0^{-1} \bar{x}
+& s = 2 \pi \beta \lamfac^{-1} \lambda_0^{-1} \bar{x}
 \end{align}
@@
 \hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
-\hline Description & Notation and definition & Value in experiments \\
+\hline Description & Notation and definition & Units \\
 \hline
@@
 2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
-+ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} \, u_{xxx} = 0
++ \frac{4 \pi^2 \delta^2}{3 \lamfac^2} D^{1/2} \, u_{xxx} = 0
 \end{align}
@@
 
+\subsection{Hamiltonian structure}
 Then the Hamiltonian structure is given by

### 2018-09-29 LatexNotes/Notes.tex
NM

@@
 & dimensionless		& 0.05--0.5\\
-Catch-all parameter		& $\mu = 4 \pi^2 / (9 M^2)$			& dimensionless		& 0.27 if $M=4$ \\
+Catch-all parameter		& $\mu = 4 \pi^2 / (9 \lamfac^2)$			& dimensionless		& 0.27 if $\lamfac=4$ \\
 \hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
@@
 Local peak wavelength	& $\lambda_p = \sqrt{gd}/f_p$	& length \\
-Characteristic wavelength	& $\lambda_c = M \lambda_p$	& length \\
+Characteristic wavelength	& $\lambda_c = \lamfac \lambda_p$	& length \\
 Depth ratio			& $D = d/h_0$				& dimensionless
@@
 \ampp = \eps \delta^{-2} = \frac{ag}{h_0^2 f_p^2} , \qquad
-\mu = \frac{4 \pi^2}{9 M^2}
+\mu = \frac{4 \pi^2}{9 \lamfac^2}
 \end{align}

### 2018-11-23 paper1_tKdV/SI_appendix/SI_tKdV.tex
source files for paper1

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnassupportinginfo} % Choose template 
%\readytosubmit %% Uncomment this line before submitting, so that the instruction page is removed.
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\usepackage{units}
\usepackage{array}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{multirow}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\theoremstyle{plain}
\newtheorem{lalgorithm}[thm]{\protect\algorithmname}

\makeatother

\usepackage{babel}
\providecommand{\algorithmname}{Algorithm}
\providecommand{\theoremname}{Theorem}

\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}

\title{A Statistical Dynamical Model to Predict
Extreme Events and Anomalous Features in Shallow Water Waves with
Abrupt Depth Change}
\author{Andrew J. Majda, M. N. J. Moore, and Di Qi}

%\correspondingauthor{\textsuperscript{a} Department of Mathematics and Center for %Atmosphere and Ocean Science, Courant Institute of Mathematical Sciences, New
%York University, New York, NY 10012 \\
%\textsuperscript{b} Department of Mathematics and Geophysical Fluid
%Dynamics Institute, Florida State University, Tallahassee, FL}

% Please give the surname of the lead author for the running footer
%\leadauthor{Majda}

%\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: qidi@cims.nyu.edu, jonjon@cims.nyu.edu}

\begin{document}
\appendix
% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
%\verticaladjustment{-2pt}

\maketitle
%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

\section{Details about the TKdV model non-dimensionalization}

We provide more details about the derivation and properties for the
truncated KdV equation with abrupt depth change used as the prediction
model in the \textcolor{blue}{\emph{main text}}. A unified formulation with dependence on
different parameter values is provided.

\subsection{Mathematical formulation of the truncated KdV equation as a Hamiltonian
system}

The classic \emph{Korteweg-de Vries} (KdV) equation \cite{johnson1997modern}
can be written in the standard form as
\begin{equation}
u_{t}+uu_{x}+u_{xxx}=0,\quad x\in\left[-\pi L_{0},\pi L_{0}\right].\label{eq:KdV}
\end{equation}
The state variable $u\left(x,t\right)$ for the leading-order surface
wave disturbance is defined on a periodic geometry of length $2\pi L_{0}$.
The KdV equation [\ref{eq:KdV}] can be also recognized as a \emph{Hamiltonian
system} as
\begin{equation}
\dot{u}=\left\{ u,\mathcal{H}\right\} =\mathcal{J}\frac{\delta\mathcal{H}}{\delta u},\quad\mathcal{J}=-\partial_{x},\;\mathcal{H}=\int_{-\pi L_{0}}^{\pi L_{0}}\left(\frac{1}{6}u^{3}-\frac{1}{2}u_{x}^{2}\right)dx.\label{eq:hamiltonian}
\end{equation}
The Poisson bracket is defined by the symplectic operator $\mathcal{J}$
\[
\left\{ \mathcal{F},\mathcal{G}\right\} =\int_{-\pi L_{0}}^{\pi L_{0}}\frac{\delta\mathcal{F}}{\delta u}\mathcal{J}\frac{\delta\mathcal{G}}{\delta u}dx,
\]
which forms a skew-symmetric and bilinear form satisfying the Jacobi
identity, $\left\{ \left\{ \mathcal{F},\mathcal{G}\right\} ,\mathcal{H}\right\} +\left\{ \left\{ \mathcal{H},\mathcal{F}\right\} ,\mathcal{G}\right\} +\left\{ \left\{ \mathcal{G},\mathcal{H}\right\} ,\mathcal{F}\right\} =0$,
acting on functionals $\mathcal{F}\left(u\right)$ and $\mathcal{G}\left(u\right)$.
The evolution of any functional $\mathcal{F}\left(u\right)$ obeys
the dynamical equation
\[
\mathcal{F}_{t}=\left\{ \mathcal{F},\mathcal{H}\right\} .
\]
Immediately, we have the conservation of the Hamiltonian in [\ref{eq:hamiltonian}],
$\mathcal{H}_{t}=\left\{ \mathcal{H},\mathcal{H}\right\} =0$. Besides
the Hamiltonian $\mathcal{H}$, two other important conserved quantities
in the KdV equation are the momentum $\mathcal{M}$ and energy $\mathcal{E}$
defined as 
\[
\mathcal{M}\left(u\right)=\int_{-\pi L_{0}}^{\pi L_{0}}udx,\quad\mathcal{E}\left(u\right)=\frac{1}{2}\int_{-\pi L_{0}}^{\pi L_{0}}u^{2}dx.
\]

In applications using the KdV equation for modeling water waves, it
is useful to consider a normalized version of the equation. For convenience,
the total momentum is normalized to zero $\mathcal{M}=0$ without
loss of generality due to the Galilean invariance and the total energy
is rescaled to unity $\mathcal{E}=1$. To achieve this, consider the
following change of variables
\[
t=\tilde{t},\;x=L_{0}\tilde{x}+M_{0}\tilde{t},\quad u=E_{0}^{1/2}L_{0}^{-1/2}\tilde{u}+M_{0},
\]
where $M_{0}=\int_{-\pi L_{0}}^{\pi L_{0}}udx$ is the conserved total
momentum and $E_{0}=\frac{1}{2}\int_{-\pi L_{0}}^{\pi L_{0}}u^{2}dx-\pi L_{0}M_{0}^{2}$
is the conserved total energy from the original system [\ref{eq:KdV}],
and $L_{0}$ defines the characteristic length scale of the system.
The additional shift in time $M_{0}t$ in the new coordinate creates
the Doppler shift from the non-zero mean momentum $M_{0}$. The \emph{normalized
KdV equation with zero momentum and unit total energy} for the scaled
state variable $\tilde{u}\left(\tilde{x},\tilde{t}\right)$ becomes
\begin{equation}
\tilde{u}_{\tilde{t}}+E_{0}^{1/2}L_{0}^{-3/2}\tilde{u}\tilde{u}_{\tilde{x}}+L_{0}^{-3}\tilde{u}_{\tilde{x}\tilde{x}\tilde{x}}=0,\quad\tilde{x}\in\left[-\pi,\pi\right],\label{eq:KdV_unit}
\end{equation}
where the system is defined on the normalized periodic domain, and
we have the normalized conservatives 
\[
\mathcal{M}=\int_{-\pi}^{\pi}\tilde{u}d\tilde{x}=0,\quad\mathcal{E}=\frac{1}{2}\int_{-\pi}^{\pi}\tilde{u}^{2}d\tilde{x}=1.
\]
The corresponding Hamiltonian functional becomes the difference between
the cubic term $H_{3}$ and the quadratic term $H_{2}$
\begin{equation}
\tilde{\mathcal{H}}\left(\tilde{u}\right)=E_{0}^{1/2}L_{0}^{-3/2}H_{3}\left(\tilde{u}\right)-L_{0}^{-3}H_{2}\left(\tilde{u}\right),\quad H_{3}\left(\tilde{u}\right)=\frac{1}{6}\int_{0}^{2\pi}\tilde{u}^{3}d\tilde{x},\;H_{2}\left(\tilde{u}\right)=\frac{1}{2}\int_{0}^{2\pi}\tilde{u}_{\tilde{x}}^{2}d\tilde{x}.\label{eq:Hamil_unit}
\end{equation}

Notice that the total momentum $M_{0}$ has no explicit contribution
in the rescaled equation. $E_{0}$ is the total energy of the system
depending on the domain size $L_{0}$, while $E_{0}/\left(2\pi L_{0}\right)$
defines the energy density at unit scale length. The entire dynamics
of the equation is determined by setting the two parameters $\left(E_{0},L_{0}\right)$.
The choice of these parameter values will be discussed next based
on a proper non-dimensinalization of the physical model. The advantage
of adopting the normalized formulation [\ref{eq:KdV_unit}] with Hamiltonian
[\ref{eq:Hamil_unit}] is that it enables us to easily control the
different cases with changing energy from a unified model setup.

\subsubsection*{Truncated KdV equation on the spectral domain}

To investigate the turbulent dynamics generated from the KdV equation,
usually a Galerkin projection $\mathcal{P}_{\Lambda}$ is applied
to the state variable $u$ with a high wavenumber truncation up to
$\Lambda$
\begin{equation}
u_{\Lambda}\left(x,t\right)\equiv\mathcal{P}_{\Lambda}u=\sum_{\left|k\right|\leq\Lambda}\hat{u}_{k}\left(t\right)e^{ikx},\label{eq:trunc_u}
\end{equation}
with in total $J=2\Lambda+1$ grid points. Accordingly,
the truncated version of the KdV equation (TKdV) can be formulated
by projecting the continuous equation [\ref{eq:KdV_unit}] to the
truncated subspace as
\begin{equation}
\frac{\partial}{\partial t}u_{\Lambda}+\frac{1}{2}E_{0}^{1/2}L_{0}^{-3/2}\frac{\partial}{\partial x}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{2}\right)+L_{0}^{-3}\frac{\partial^{3}}{\partial x^{3}}u_{\Lambda}=0.\label{eq:tKdV}
\end{equation}
The additional projection in front of the quadratic term $u_{\Lambda}^{2}$
is used to remove the aliasing modes that go beyond the range $\left|k\right|>\Lambda$.

The TKdV equation [\ref{eq:tKdV}] can be also written in the spectral
domain for each Fourier mode $\hat{u}_{k}$ as
\[
\frac{d\hat{u}_{k}}{dt}=-\frac{ik}{2}E_{0}^{1/2}L_{0}^{-3/2}\sum_{\left|m\right|\leq\Lambda}\hat{u}_{m}\hat{u}_{m-k}^{*}+ik^{3}L_{0}^{-3}\hat{u}_{k}.
\]
It can be shown that the three conserved quantities above are still
conserved in this truncated system. With the spectral representation,
the conserved momentum and energy set the constraints on the spectral
coefficients 
\[
\mathcal{M}_{\Lambda}=\int_{0}^{2\pi}u_{\Lambda}dx=\left(2\pi\right)\hat{u}_{0}=0,\quad\mathcal{E}_{\Lambda}=\frac{1}{2}\int_{0}^{2\pi}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{2}\right)dx=2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1.
\]
The discretized Hamiltonian can be written accordingly as
\[
\mathcal{H}_{\Lambda}=E_{0}^{1/2}L_{0}^{-3/2}H_{3,\Lambda}-L_{0}^{-3}H_{2,\Lambda},\quad H_{3,\Lambda}=\frac{\pi}{3}\sum_{\left|m\right|,\left|n\right|\leq\Lambda}\hat{u}_{m}\hat{u}_{n}\hat{u}_{m+n}^{*},\;H_{2,\Lambda}=\pi\sum_{\left|k\right|\leq\Lambda}k^{2}\left|\hat{u}_{k}\right|^{2}.
\]
Especially, the Hamiltonian structure of the previous continuous equation
is maintained in this semi-discrete TKdV equation. The truncated equation
[\ref{eq:tKdV}] is still a Hamiltonian system with the corresponding
discrete Hamiltonian $\mathcal{H}_{\Lambda}$. Furthermore, the truncated
system [\ref{eq:tKdV}] satisfies the Liouville property \cite{abramov2003hamiltonian,majda2006nonlinear},
thus equilibrium statistical mechanism can be constructed based on
the conserved quantities.

\subsection{The rescaled TKdV model with non-dimensionalized parameters}

Since the KdV equation is derived from the leading-order terms in
the asymptotic expansion of the Euler equations, we find the sizes
of the model parameters directly from the leading-order assumption
for characterizing the physical problem from the experiments.

\subsubsection*{The KdV equation with depth dependence}

The formulation for the KdV equation with a depth dependence $D_{0}$
can be introduced using the new coordinate system $\left(X,\xi\right)$
for the convenience of derivation as
\begin{equation}
X=\epsilon x,\quad\xi=D_{0}^{-1/2}x-t.\label{eq:variables}
\end{equation}
The equation is used to describe the far-field and long-time variability
at $X=O\left(1\right),\xi=O\left(1\right)$ as $x\rightarrow\infty$
and $t\rightarrow\infty$. With the water depth dependence, the leading-order
wave disturbance $\eta\left(\xi,X\right)$ obeys the generalized KdV
equation (derivation can be found in \cite{johnson1997modern})
\begin{equation}
D_{0}^{1/4}\eta_{X}+\frac{3}{2}D_{0}{}^{-5/4}\eta\eta_{\xi}+\frac{1}{6}D_{0}{}^{3/4}\eta_{\xi\xi\xi}=0.\label{eq:KdV_depth}
\end{equation}
It needs to be emphasized that the spatial variable $X$ here plays
the original role of the time variable $t$, and the new variable
$\xi$ tracks the right-moving waves along the characteristics in
the new equation.

\subsubsection*{Model parameters from non-dimensionalization}

To derive the model parameter values, we first propose scales in the
state variables according to the dimensional formulation. Then the
parameter values in the rescaled model can be found by comparing the
equation with its non-dimensionalized version. For the basic scales
in the problem, we consider: i) the typical depth of the water tank
$H_{0}$; ii) the surface wave disturbance amplitude $a$; and iii)
the characteristic surface wavelength $\lambda_{c}=c_{0}/f_{c}$ with
$f_{c}$ the characteristic forcing frequency and the gravity wave
speed $c_{0}=\sqrt{gH_{0}}$. The important characterizing parameters
that can be measured from the experiments include 
\begin{equation}
\epsilon=\frac{a}{H_{0}},\quad\delta=\frac{H_{0}}{\lambda_{c}},\quad D_{0}=\frac{d}{H_{0}}.\label{eq:params}
\end{equation}
Above, $\epsilon$ is the wave amplitude to water depth ratio; $\delta$
is the water depth to wavelength scale ratio; and $D_{0}$ defines
the normalized wave depth ratio in the depth change from $d=H_{0}$
to $d<H_{0}$. The interpretations and reference values of these model
parameters are based on the experimental setup \cite{bolles2018anomalous}.

We are interested in the surface waves at the limit $\epsilon\rightarrow0,\delta\rightarrow0$,
so that the amplitude of the surface waves stays small and the shallow
water approximation is valid. Then we introduce the characteristic
physical scales for the surface disturbance $\eta$, spatial variable
$x$, and temporal variable $t$, based on the scales in \cite{johnson1997modern}
\[
\left[\eta\right]=\epsilon H_{0},\;\left[x\right]=\epsilon^{-1/2}H_{0},\;\left[t\right]=\epsilon^{-1/2}H_{0}/c_{0}.
\]
We use $\left(\eta;X,\xi\right)$ to represent the variables in the
dimensional formulation [\ref{eq:KdV_depth}]; and $\left(\tilde{u};\tilde{t},\tilde{x}\right)$
for the final non-dimensionalized variables in [\ref{eq:KdV_unit}].
The above scales give the relations between the two sets of variables
\[
\tilde{u}=U^{-1}\left[\epsilon H_{0}\right]\eta,\;\tilde{x}=L_{d}^{-1}\left[\epsilon^{-\frac{1}{2}}D_{0}^{\frac{1}{2}}H_{0}\right]\xi,\;\tilde{t}=L_{s}^{-1}\left[\epsilon^{-\frac{3}{2}}H_{0}\right]X,
\]
where $\left(U,L_{d},L_{s}\right)$ are the scales in $\left(\eta,\xi,X\right)$
to normalize the original system to the standard form with unit energy
and domain size $2\pi$ as in [\ref{eq:KdV_unit}]. We introduce the
length scale $L_{d}$ according to the characteristic wavenumber $\lambda_{c}$
and the spatial scale $L_{s}$ according to slow-varying far-field
variables
\begin{equation}
L_{d}=M\lambda_{c}=\frac{M\sqrt{gd}}{f_{c}},\quad L_{s}=\epsilon^{-3/2}H_{0}=\frac{\sqrt{H_{0}^{5}}}{\sqrt{a^{3}}},\quad\gamma=\frac{U}{a}=\frac{U}{\epsilon H_{0}}.\label{eq:meas_scales}
\end{equation}
where $M$ is an integer representing a wave package of multiple wavelengths.
Furthermore, $\gamma$ defines the additional ratio that normalizes
the total energy in $u$ to unit. Substituting the above relations
into the equation [\ref{eq:KdV_depth}], we find the final non-dimensionalized
equation about $\tilde{u}\left(\tilde{t},\tilde{x}\right)$
\begin{equation}
\frac{\partial\tilde{u}}{\partial\tilde{t}}+\epsilon^{-\frac{1}{2}}\frac{3\delta}{4\gamma M}D_{0}^{-\frac{3}{2}}\frac{\partial}{\partial\tilde{x}}\left(\tilde{u}^{2}\right)+\epsilon^{-\frac{3}{2}}\frac{\delta^{3}}{6M^{3}}D_{0}^{\frac{1}{2}}\frac{\partial^{3}\tilde{u}}{\partial\tilde{x}^{3}}=0.\label{eq:KdV_nondim}
\end{equation}
Therefore, we formulation the non-dimensional equations [\ref{eq:KdV_nondim}]
based on the non-dimensional parameters [\ref{eq:params}].

In the final step, we determine the parameter values in the computational
models [\ref{eq:KdV_unit}] by comparing with the corresponding non-dimensionalized
form [\ref{eq:KdV_nondim}] with the measured quantities. After proper
rearrangement and leaving the tildes in the normalized variables,
we find the \emph{normalized KdV equation} (as eqn. {[}1{]} in the
\textcolor{blue}{\emph{main text}}) 
\begin{equation}
u_{t}+E_{0}^{1/2}L_{0}^{-3/2}D_{0}^{-3/2}uu_{x}+L_{0}^{-3}D_{0}^{1/2}u_{xxx}=0,\quad x\in\left[-\pi,\pi\right],\label{eq:KdV_simu}
\end{equation}
The model parameters are determined by the non-dimensional measurements
\begin{equation}
L_{0}=6^{\frac{1}{3}}\left(M\epsilon^{\frac{1}{2}}\delta^{-1}\right),\;E_{0}=\frac{27}{2}\gamma^{-2}\left(M\epsilon^{\frac{1}{2}}\delta^{-1}\right).\label{eq:params-1}
\end{equation}
Notice that in the above model `$t$' is used to represent either
the far-field spatial variable `$X$' and `$x$' is used to represent
the wave location variable `$\xi$'. We also drop the `tildes' in
the variables for simplicity.

In addition, to estimate the parameter $M$ for the computational
domain size $2\pi L_{d}=2\pi M\lambda_{c}$ determined as $M$-multiple
of the characteristic wavelength $\lambda_{c}$, we consider the spatial
discretization $J=2\Lambda+1=32$ so that the smallest resolved scale
is comparable with the characteristic wavelength, that is, $2\pi M\lambda_{c}/J=\lambda_{c}$.
This estimation gives $M=\frac{J}{2\pi}\sim5$, inferring a computational
domain with 5 multiples of the characteristic wavelength. Using the
reference experimental measurements, we can calculate the reference
values for the model scales $L_{0}$ and $E_{0}$ used in the direct
numerical simulations. The wave amplitude to depth ratio $\epsilon$
changes in the range $\left[0.0024,0.024\right]$, thus we can estimate
the minimum and maximum values of the scales in Table \ref{tab:Reference-parameter-values}.
We adopt this basic guideline in choosing parameter values in the
numerical confirmation of the theory in the \textcolor{blue}{\emph{main text}}.

\begin{table}
\begin{centering}
\begin{tabular}{cccc}
\toprule 
$J=32$ &  & \multicolumn{2}{c}{$\epsilon\in\left[0.0024,0.024\right]$, $\delta=0.22$, $M=5$}\tabularnewline
\midrule
\midrule 
\multirow{1}{*}{$L_{0}$} &  & \multicolumn{2}{c}{2.0 -- 6.3}\tabularnewline
\midrule 
\multirow{2}{*}{$E_{0}$} & $\gamma=1$ & \multicolumn{2}{c}{47.53}\tabularnewline
\cmidrule{2-4} 
 & $\gamma=0.5$ & \multicolumn{2}{c}{190.1}\tabularnewline
\midrule 
\multirow{2}{*}{$D_{0}$} & upstream & \multicolumn{2}{c}{1}\tabularnewline
\cmidrule{2-4} 
 & downstream & \multicolumn{2}{c}{0.24}\tabularnewline
\bottomrule
\end{tabular}
\par\end{centering}
\caption{Reference parameter values for the numerical model. The computation
domain is set as $M=5$ multiples of the characteristic wavelength.\label{tab:Reference-parameter-values}}
\end{table}

\section{Sampling strategy and statistical link for the invariant measures}

Here we show the detailed strategy and algorithm used for computing
the sampled Gibbs invariant measure and the statistical matching at
the abrupt depth change shown in the \textcolor{blue}{\emph{main text}}. The invariant
Gibbs measure is defined based on microcanonical ensemble in the quadratic
energy $\mathcal{E}_{\Lambda}$ on the isosurface and canonical ensemble
in the Hamiltonian $\mathcal{H}_{\Lambda}$ \cite{majda2006nonlinear,bajars2013weakly}.
The \emph{invariant Gibbs measure} for the TKdV model [\ref{eq:KdV_simu}]
about the normalized state variable $u_{\Lambda}$ with unit energy
can be written explicitly as
\begin{equation}
\mathcal{G}_{\theta^{\pm}}\left(u_{\Lambda}\right)=C_{\theta^{\pm}}\exp\left(-\theta^{\pm}\left\{ h_{3}^{\pm}\sum_{1\leq\left|m\right|,\left|n\right|\leq\Lambda}\hat{u}_{m}\hat{u}_{n}\hat{u}_{m+n}^{*}-h_{2}^{\pm}\sum_{1\leq\left|k\right|\leq\Lambda}k^{2}\left|\hat{u}_{k}\right|^{2}\right\} \right)\delta\left(\pi\sum_{1\leq\left|k\right|\leq\Lambda}\left|\hat{u}_{k}\right|^{2}-1\right),\label{eq:gibbs}
\end{equation}
with the coefficients $h_{3}^{\pm}=\frac{\pi}{3}E_{0}^{1/2}L_{0}^{-3/2}D_{\pm}^{-3/2}$
and $h_{2}^{\pm}=\pi L_{0}^{-3}D_{\pm}^{1/2}$ depending on the model
parameters. The proper statistical matching condition is proposed to connect the
invariant measures in incoming and outgoing flow statistics with different
inverse temperatures $\theta^{\pm}$.

In predicting the statistical distribution of state variables after the abrupt depth change, we
start with the assumption that we have full knowledge about the
invariant distribution about the incoming flow field, $\mathcal{G}_{\theta^{-}}$
at equilibrium. Thus the parameters for $\theta^{-}$ are determined
with accuracy from the observations or statistical simulations in
the incoming flow, and the total energy $E_{0}$ is conserved. The
condition to get the outgoing flow statistics with parameter $\theta^{+}$
is through the following matching condition
\begin{equation}
\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta^{-}}}=\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta^{+}}},\label{eq:matching}
\end{equation}
with the Gibbs measures $\mathcal{G}_{\theta^{\pm}}$ before and after
the abrupt depth change defined in [\ref{eq:gibbs}]. Notice in the
statistical matching condition [\ref{eq:matching}], we don't require
the statistical information for the incoming Hamiltonian $H_{\Lambda}^{-}$.
This is considering the near-Gaussian statistics in the incoming waves
and its weak dependence on the Hamiltonian in the Gibbs measure. Therefore
the Hamiltonian before the depth change is only used to construct
the invariant measure, $\mathcal{G}_{\theta^{-}}$.

\subsection{Sampling algorithm for the incoming Gibbs measure}

First, the task is to compute the expectations $\left\langle H^{+}\right\rangle _{\mathcal{G}_{\theta^{\pm}}}$
under the proper invariant measures from [\ref{eq:gibbs}]. The central
question is how to find a method to properly sample the equilibrium
distribution according to the chosen inverse temperature
$\theta$. One method to directly sample the Gibbs measure is
described in \cite{abramov2003hamiltonian}, where a Metropolis-Hasting
Monte-Carlo sampling is applied. The following algorithm is adopted
to sample the mixed canonical-microcanonical ensemble distribution
[\ref{eq:gibbs}] for the numerical results in the \textcolor{blue}{\emph{main text}}:
\begin{algorithm}
\caption{Metropolis-Hastings sampling for the
mixed ensemble for $\left\{ \hat{u}_{k}\right\} _{k=1}^{\Lambda}$}\label{alg:Metropolis-Hastings}

%\begin{itemize}
\begin{algorithmic}[1]
\State Assign i.i.d standard Gaussian random variables to the $\Lambda$-dimensional
complex valued samples, $\vec{w}=\left\{ \hat{w}_{1},\cdots,\hat{w}_{\Lambda}\right\} \in\mathbb{C}^{\Lambda}$;
\State Normalize for each component $\hat{u}_{k}=\sqrt{\frac{1}{E}}\hat{w}_{k}$
with the total energy $E=2\pi\sum_{k=1}^{\Lambda}\left|\hat{w}_{k}\right|^{2}$;
\State Starting from $\vec{u}_{t}=\vec{u}$, generate a new candidate $\vec{w}$
according to a transition probability, $g\left(\vec{w}\mid\vec{u}\right)$
(a simple choice is $g\sim\exp\left(-\frac{\Lambda}{2}\left|\vec{w}-\vec{u}\right|^{2}\right)$),
and normalize $\vec{u}^{\prime}\leftarrow\sqrt{\frac{1}{E}}\vec{w}$;
\State Calculate the acceptance probability $a=\min\left\{ 1,\exp\left(-\theta\left[H\left(\vec{u}^{\prime}\right)-H\left(\vec{u}\right)\right]\right)\frac{g\left(\vec{u}\mid\vec{u}^{\prime}\right)}{g\left(\vec{u}^{\prime}\mid\vec{u}\right)}\right\} $;
\State Generate a uniform random number $r=\mathcal{U}\left[0,1\right]$.
Accept $\vec{u}_{t+1}=\vec{u}^{\prime}$ if $r\leq a$ and reject
$\vec{u}_{t+1}=\vec{u}$ if $r>a$;
\State Repeat the previous steps and sample at every $n=10$ steps.
%\end{itemize}
\end{algorithmic}
\end{algorithm}

We normalize the vector at each step to keep it on the fixed energy
surface. It is also safer in practice to skip the first $M=10^{3}$
samples from the start for the system to reach stationary. Then the
expectations can be calculated by averaging among all the samples
from the above algorithms
\[
\left\langle H_{\Lambda}^{+}\right\rangle _{\mu}=\frac{1}{N}\sum_{t=1}^{N}H_{\Lambda}^{+}\left(\vec{u}_{t}\right)=E_{0}^{1/2}L_{0}^{-3/2}D_{+}^{-3/2}\frac{1}{N}\sum_{t=1}^{N}H_{3}\left(\vec{u}_{t}\right)-L_{0}^{-3}D_{+}^{1/2}\frac{1}{N}\sum_{t=1}^{N}H_{2}\left(\vec{u}_{t}\right),
\]
where $H_{3}\left(\vec{u}\right)$ computes the cubic term $u^{3}$
and $H_{2}\left(\vec{u}\right)$ computes the wave slope energy $u_{x}^{2}$
according to the definition in [\ref{eq:hamiltonian}]. The PDFs are
achieved by ``bin counting'' from the histogram of the samples $\left\{ \vec{u}_{t}\right\} _{t=1}^{N}$.
As one illustration, Figure \ref{fig:Test-distributions} displays the
sampled statistics, the PDFs for the Hamiltonian $H_{\Lambda}$, the
PDFs for the state variable $u_{\Lambda}$, and the corresponding
energy spectra for each Fourier mode. Typically we test the range
of values of the inverse temperature from positive to negative $\theta=0.5,0.25,0,-0.25,-0.5$.
Supporting the claim in the \textcolor{blue}{\emph{main text}}, the negative inverse
temperature $\theta<0$ produces the proper physical regime for the
incoming flow statistics, where the energy spectrum is decaying from
larger scale to small scale. In contrast with $\theta>0$, the small
scale modes get more energetic and a flat top is generated in the
PDF of $u_{\Lambda}$ with two peaks. With $\theta=0$, the Gibbs
measure is the uniform distribution on the constant energy shell $E_{0}$,
thus equipartition of energy in each mode is expected.

\begin{figure}
\begin{centering}
\includegraphics[scale=0.27]{./pdf0_hamil}\includegraphics[scale=0.27]{./pdf0_u}\includegraphics[scale=0.27]{./spec0}
\par\end{centering}
\caption{Statistics generated from the Gibbs invariant measures with different
inverse temperatures $\theta$ and $D_{0}=1$. Positive values of
$\theta$ are shown in dashed lines and negative values of $\theta$
in solid lines.\label{fig:Test-distributions}}
\end{figure}

\subsection{Computing downstream parameter from the statistical matching}

In the second step, we need to enforce the the statistical matching
condition [\ref{eq:matching}] to discover the corresponding downstream
inverse temperature $\theta^{+}$. For the efficiency of the method
in finding the optimal value of $\theta_{*}^{+}$, we use a secant
method by solving a nonlinear equation following the algorithm:
%\addtocounter{algorithm}{1}
\begin{algorithm}[H]
\caption{Recovering the downstream inverse temperature
$\theta^{+}$ from the statistical matching condition}\label{alg:Recovering}
%\begin{itemize}
\begin{algorithmic}[1]
\State Pick up the upstream value of $\theta^{-}$ and compute the expectation
under incoming flow Gibbs measure $\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta^{-}}}$;
\State Define a function of $\theta^{+}$ as the difference of the upstream
and downstream expectations, $F\left(\theta^{+}\right)=\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta^{+}}}-\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta^{-}}}$;
\State Find the solution of $F\left(\theta^{+}\right)=0$ using the secant
method, so that the optimal downstream value $\theta_{*}^{+}$ is
found.
%\end{itemize}
\end{algorithmic}
\end{algorithm}

In addition, to remove stiffness with large values of $L_0$ and $E_0$, we introduce the normalized parameter $\tilde{\theta}=L_{0}^{-3/2}\theta$ so that $\mathcal{G}_{\tilde{\theta}}\sim\exp\left(-\tilde{\theta}\left[E_{0}^{1/2}D_0^{-3/2}H_{3}-L_{0}^{-3/2}D_0^{1/2}H_{2}\right]\right)$. In the numerical computations, we tune the parameter $\tilde{\theta}$ instead for the matching condition. Using this modified secant method, it is found that the optimal value can be
usually reached after 2 to 5 iteration steps as long as the scheme
is guaranteed non-stiff.

\subsubsection*{Numerical comparison of the matching results}

The solid black lines in Figure \ref{fig:Confirmation} plot the downstream
skewness $\kappa^{+}$ computed according to the matched downstream
$\theta^{+}$ as shown in Fig. 1 of the \textcolor{blue}{\emph{main text}}. Here as
a further comparison, the results with different total energy levels
$E_{0}$ starting from various various values of incoming $\theta^{-}$
are compared. Further, Table \ref{tab:Statistics-matching} compares
the incoming (with $\theta^{-}$) and outgoing (with $\theta^{+}$)
wave statistics using the parameters achieved from the statistical
matching condition. In the last two rows, the contributions of the
cubic term $H_{3}$ and quadratic term $H_{2}$ in the Hamiltonian
are compared in the incoming and outgoing flow statistics. It is found
that in the upstream statistics $\theta^{-}$, the quadratic part
$H_{2}$ is dominant, while in the downstream flow statistics, the
cubic part becomes important in the total Hamiltonian, implying stronger
non-Gaussian statistics.

As a further comparison, in the \textcolor{blue}{\emph{main text}} we also use the Gamma
distribution 
\[
\rho\left(u;k,\alpha\right)=\frac{e^{-k}\alpha^{-1}}{\Gamma\left(k\right)}\left(k+\alpha^{-1}u\right)^{k-1}e^{-\alpha^{-1}u},\quad\sigma^{2}=k\alpha^{2},\:\kappa_{3}=\frac{2}{\sqrt{k}},
\]
to fit the downstream state $u^{+}$ according to the measured skewness $\kappa_3$ and variance $\sigma^2$ for the parameters $\left(k,\alpha\right)$.
Table \ref{tab:Statistics-matching} also shows the fitting parameter
values according to the downstream statistics with $\theta^{+}$.
The Gamma fitting can also give an estimation for the kurtosis as
$\kappa_{4}=\frac{6}{k}+3$, which is shown to have good agreement
with the truth value $\kappa_4^{+}$ especially in the larger skewness
cases.

\begin{table}
\begin{centering}
\begin{tabular}{ccccc}
\toprule 
$\theta^{-}$ & $E_{0}=100$ & -0.5  & -0.3  & -0.1 \tabularnewline
\midrule
\midrule 
$\theta^{+}$ &  & -0.32  & -0.26  & -0.12 \tabularnewline
\midrule 
\multicolumn{2}{c}{$\left\langle H_{\Lambda}^{+}\right\rangle _{\theta^{-}}=\left\langle H_{\Lambda}^{+}\right\rangle _{\theta^{+}}$} & -9.57 & -15.72 & -28.17\tabularnewline
\midrule
\midrule 
\multirow{2}{*}{$\kappa_{3}$} & $\theta^{-}$ & 0.16 & 0.093 & 0.024\tabularnewline
\cmidrule{2-5} 
 & $\theta^{+}$ & 0.85 & 0.62 & 0.24\tabularnewline
\midrule
\multirow{1}{*}{$\kappa_{4}^{+}$} &  & 3.96 & 3.42 & 2.88\tabularnewline
\midrule
Gamma  & $\left(k,\alpha\right)$ & (5.5, 0.24) & (10.4, 0.18) & (68.0, 0.068)\tabularnewline
\cmidrule{2-5} 
fitting & $6/k+3$ & 4.08 & 3.58 & 3.08\tabularnewline
\midrule
\midrule 
\multirow{2}{*}{$\left\langle \tilde{H}_{3}\right\rangle _{\theta}$} & $\theta^{-}$ & 0.31 & 0.17 & 0.043\tabularnewline
\cmidrule{2-5} 
 & $\theta^{+}$ & 13.64 & 9.92 & 3.85\tabularnewline
\midrule 
\multirow{2}{*}{$\left\langle \tilde{H}_{2}\right\rangle _{\theta}$} & $\theta^{-}$ & 24.85 & 35.09 & 58.25\tabularnewline
\cmidrule{2-5} 
 & $\theta^{+}$ & 23.38 & 25.73 & 32.58\tabularnewline
\bottomrule
\end{tabular}
\par\end{centering}
\caption{Statistics computed from incoming and outgoing flow Gibbs measures
with $E_{0}=100$. The skewness $\kappa_{3}$ and kurtosis $\kappa_{4}$
predicted from the Gibbs measures with different $\theta^{-}$ are
compared. The outgoing statistics are compared with the Gamma function
fitting with parameter $\left(k,\alpha\right)$ and the corresponding
kurtosis $\kappa_{4}=6/k+3$. The expected values of the cubic and
quadratic terms in the Hamiltonian, $\tilde{H}_{3}=E_{0}^{1/2}L_{0}^{-3/2}D_{\pm}^{-3/2}\frac{1}{6}\int u^{3}$,
and $\tilde{H}_{2}=L_{0}^{-3}D_{\pm}^{1/2}\frac{1}{2}\int u_{x}^{2}$
are also compared.\label{tab:Statistics-matching}}
\end{table}

\subsubsection*{An explicit formula from the statistical matching condition}

In complementary to the \textcolor{blue}{\emph{main text}}, we offer a more detailed
derivation about the statistical link between the upstream and downstream
wave slope energy and the skewness. The statistical matching condition
with the downstream Hamiltonian can be written as
\[
\left\langle H_{\Lambda}^{+}\right\rangle _{\mu-}=\left\langle H_{\Lambda}^{+}\right\rangle _{\mu+},\;H_{\Lambda}^{+}=E_{0}^{1/2}L_{0}^{-3/2}D_{+}^{-3/2}H_{3}-L_{0}^{-3}D_{+}^{1/2}H_{2}.
\]
In general, we can have the expected value of $H_{3}$ and $H_{2}$
about any probability measure $\mu$ as
\[
\left\langle H_{3}\right\rangle _{\mu}=\frac{1}{6}\int_{0}^{2\pi}\left\langle u^{3}\right\rangle _{\mu-}dx=\frac{2\pi}{6J}\sum_{j=1}^{J}\left\langle u_{j}^{3}\right\rangle _{\mu},\;\left\langle H_{2}\right\rangle _{\mu}=\frac{1}{2}\int_{0}^{2\pi}\left\langle \left(u_{x}\right)^{2}\right\rangle _{\mu}dx=\frac{2\pi}{2}\sum_{\left|k\right|=1}^{\Lambda}k^{2}\left\langle \left|\hat{u}_{k}\right|^{2}\right\rangle _{\mu}.
\]
Above the quadratic term $H_{2}$ characterizes the \emph{slopes of
the surface waves}, $u_{x}$, in the incoming and outgoing flows.
Beside, the skewness the state variable $u$ at one grid point is
defined as the ratio between the third and second moments (note that
in the homogeneous model setup, we always have zero mean in the state
variable $\left\langle u\right\rangle \equiv0$)
\[
\kappa_{3}=\frac{\left\langle u_{j}^{3}\right\rangle _{\mu}}{\left\langle u_{j}^{2}\right\rangle _{\mu}^{\frac{3}{2}}},\quad\frac{1}{2}\int_{0}^{2\pi}\left\langle u^{2}\right\rangle _{\mu}dx=\frac{2\pi}{2J}\sum_{j=1}^{J}\left\langle u_{j}^{2}\right\rangle _{\mu}=1.
\]
Above the total energy of the system is normalized to one in the computations.
Now we use the two assumptions introduced in the main text:
\begin{itemize}
\item The upstream equilibrium measure $\mu_{-}$ has a near-Gaussian distribution,
so that the third moments in the cubic term stay in a small value
\[
\left\langle H_{3}\right\rangle _{\mu-}=\frac{1}{6}\int_{0}^{2\pi}\left\langle u^{3}\right\rangle _{\mu-}dx=\epsilon;
\]
\item The downstream equilibrium measure $\mu_{+}$ is homogeneous at each
physical grid point, so that the second and third moments are invariant
at each grid point
\[
\left\langle u_{j}^{2}\right\rangle _{\mu+}=\sigma^{2}=\pi^{-1},\quad\left\langle u_{j}^{3}\right\rangle _{\mu+}=\sigma^{3}\kappa_{3}=\pi^{-\frac{3}{2}}\kappa_{3}.
\]
\end{itemize}
With the homogeneous assumption above, the downstream cubic term can
be simplified as
\[
\left\langle H_{3}\right\rangle _{\mu+}=\frac{\pi}{3J}\sum_{j=1}^{J}\left\langle u_{j}^{3}\right\rangle _{\mu+}=\frac{1}{3}\pi^{-\frac{1}{2}}\kappa_{3}.
\]
With the above assumptions, the above statistical matching condition
can be simplified as
\begin{align*}
E_{0}^{1/2}L_{0}^{-3/2}D_{+}^{-3/2}\epsilon & -L_{0}^{-3}D_{+}^{1/2}\left\langle H_{2}\right\rangle _{\mu-}=E_{0}^{1/2}L_{0}^{-3/2}D_{+}^{-3/2}\left\langle H_{3}\right\rangle _{\mu+}-L_{0}^{-3}D_{+}^{1/2}\left\langle H_{2}\right\rangle _{\mu+}\\
\Rightarrow\; & E_{0}^{1/2}L_{0}^{-3/2}D_{+}^{-3/2}\epsilon+\pi L_{0}^{-3}D_{+}^{1/2}\sum_{1\leq\left|k\right|\leq\Lambda}k^{2}\left(r_{k}^{+}-r_{k}^{-}\right)=\frac{1}{3}\pi^{-1/2}E_{0}^{1/2}L_{0}^{-3/2}D_{+}^{-3/2}\kappa_{3},
\end{align*}
with $r_{k}^{-}=\left\langle \left|\hat{u}_{k}\right|^{2}\right\rangle _{\mu-}$
the variance in upstream statistics and $r_{k}^{+}=\left\langle \left|\hat{u}_{k}\right|^{2}\right\rangle _{\mu+}$
the statistics from downstream statistics. Therefore we can estimate
the skewness of the state variable $u$ from the energy spectra in
inflow and outflow statistics
\begin{equation}
\kappa_{3}^{+}=3\pi^{\frac{3}{2}}L_{0}^{-\frac{3}{2}}E_{0}^{-\frac{1}{2}}D_{+}^{2}\sum_{1\leq\left|k\right|\leq\Lambda}k^{2}\left(r_{k}^{+}-r_{k}^{-}\right)+3\pi^{\frac{1}{2}}\epsilon.\label{eq:skewness}
\end{equation}
With further assumption that the incoming flow gets zero third moment,
then the skewness above can be calculated by
\[
\kappa_{3}^{+}=C\sum_{1\leq\left|k\right|\leq\Lambda}k^{2}\left(r_{k}^{+}-r_{k}^{-}\right),\;C=3\pi^{\frac{3}{2}}L_{0}^{-\frac{3}{2}}E_{0}^{-\frac{1}{2}}D_{+}^{2},
\]
with the coefficient $C$ depending on the model setup. Especially,
we can see that a positive skewness can be generated from $r_{k}^{+}>r_{k}^{-}$
in small scale modes and $r_{k}^{+}<r_{k}^{-}$ in large scales. The
total energy of the system is always normalized to unit, $\pi\sum r_{k}^{+}=\pi\sum r_{k}^{-}=1$.

Notice that in [\ref{eq:skewness}], the spectral difference between
the incoming and outgoing flows
\[
\sum_{1\leq\left|k\right|\leq\Lambda}k^{2}\left(r_{k}^{+}-r_{k}^{-}\right)=\frac{1}{2\pi}\int_{0}^{2\pi}\left[\left\langle \left(u_{x}\right)^{2}\right\rangle _{\mu+}-\left\langle \left(u_{x}\right)^{2}\right\rangle _{\mu-}\right]dx,
\]
calibrates the statistics in the `waves slopes', $u_{x}$, before
and after the depth change. Therefore, the skewness in the surface
wave displacement statistics is connected with the waves slopes between
the incoming and outgoing waves. This is consistent with the physical
understanding in the flow field, since a stronger slope in the waves
always implies the larger skewness in the disturbance.

In Figure \ref{fig:Confirmation}, we compare the accuracy of the
theoretical estimation derived above using numerical tests. In the
regime with small incoming inverse temperature $\theta^{-}$, the
theoretical formula in [\ref{eq:skewness}] offers quite accurate
approximation of the third-order skewness using only information from
the second-order moments of the wave-slope spectrum. When $\theta^{-}$
grows larger in amplitude, the error without the incoming flow correction
becomes large. The improvement is obvious by adding the correction
from the incoming wave skewness $\kappa^{-}$. However, for the statistical
matching condition, there exists a maximum amplitude threshold of
$\theta^{-}$ for the matching to be valid (so that $\left\langle H^{-}\right\rangle $
stays negative). The maximum threshold value of $\theta^{-}$ is marked
by the dashed vertical line.

\begin{figure}
\begin{centering}
\includegraphics[scale=0.28]{./matchTheory_E50}\includegraphics[scale=0.28]{./matchTheory_E100}\includegraphics[scale=0.28]{./matchTheory_E200}
\par\end{centering}
\caption{Confirmation of the theoretical approximation of the skewness in the
state $u$ using energy in the wave slope $u_{x}$. The results with
and without the incoming flow skewness correction $\kappa_{3}^{-}$
are compared. The vertical line marks the value where the statistical
matching condition is valid.\label{fig:Confirmation} }
\end{figure}

\section{Symplectic scheme for direct numerical simulations of the TKdV equation}

Here we summarize the direct numerical methods used for the validation
of the statistical theory in the \textcolor{blue}{\emph{main text}}. In general, the
strategy by using direct numerical simulations to confirm the anomalous
statistics follows the steps:
\begin{enumerate}
\item Generate samples from the incoming equilibrium distribution $\mu_{\infty}^{-}$
according to Algorithm \ref{alg:Metropolis-Hastings} for the incoming
waves. The model parameters $\left(E_{0},L_{0},\Lambda\right)$ are
determined from the experimental measurements and statistics in the
incoming flow;
\item Compute the expectation of the outflow Hamiltonian, $\left\langle H_{\Lambda}^{+}\right\rangle _{\mu_{\infty}^{-}}$,
under the chosen incoming Gibbs measure $\mu_{\infty}^{-}=\mathcal{G}_{\theta^{-}}$;
\item The outgoing flow statistics can be predicted by both the equilibrium
statistical mechanism and the direction model simulations. 
\begin{enumerate}
\item Equilibrium statistical mechanism: find the optimal outgoing flow
parameter $\theta_{*}^{+}$ from matching the expectation in [\ref{eq:matching}]
with the inflow value according to Algorithm \ref{alg:Recovering}.
Then the new equilibrium distribution $\mu_{\infty}^{+}$ after the
abrupt depth change can be sampled using the same method;
\item Direct model simulations: run the dynamical model [\ref{eq:KdV_simu}]
with outflow depth $D_{+}<1$ starting from the state sampled from
the inflow distribution $\mathcal{G}_{\theta^{-}}$. The outgoing
flow statistics can be captured by the solution from the direction
simulations. 
\end{enumerate}
\end{enumerate}
Above we proposed two approaches to predict the statistics in the
downstream wave disturbance. The equilibrium statistical mechanism
is an easy way to confirm the matching condition and get the equilibrium
distribution in far-field long-time limit. On the other hand, the
direct simulations of the dynamical model can offer estimates about
the transient statistics before the far-field equilibrium state is
reached. These two approaches can be used to confirm each other in
the numerical results.

\subsection{Conservation of energy and Hamiltonian form the symplectic scheme}

For running the direct numerical simulations of the TKdV equations,
a proper symplectic integrator is necessary to guarantee the conservation
of the Hamiltonian and energy in time. As emphasized in the \textcolor{blue}{\emph{main
text}}, the conservation property plays a central role in the statistical
matching and the final equilibrium measure, otherwise the non-symplectic
schemes introduce high numerical dissipations. The symplectic scheme
used here for the time integration of equation [\ref{eq:KdV_simu}]
is the 4th-order midpoint method \cite{mclachlan1993symplectic}.
Let $\mathbf{u}^{n}$ be the solution at time $t_{n}$ and $\mathbf{u}^{n+1}$
be the solution at next time step $t_{n+1}=t_{n}+\Delta t$, and $\mathbf{F}$the
discretized operator for the spectral modes. The midpoint method has
two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$
\[
\begin{aligned}\mathbf{y}_{1}-\mathbf{u}^{n} & =w_{1}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{1}+\mathbf{u}^{n}\right]\right),\\
\mathbf{y}_{2}-\mathbf{y}_{1} & =w_{2}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{2}+\mathbf{y}_{1}\right]\right),\\
\mathbf{u}^{n+1}-\mathbf{y}_{2} & =w_{3}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{u}^{n+1}+\mathbf{y}_{2}\right]\right),
\end{aligned}
\]
with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$,
$w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$. Then the PDFs can be counted
from the solution $u_{\Lambda}\left(t\right)$, and the statistical
expectations can be calculated by averaging along the trajectory
\[
\left\langle F\left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T}F\left(u_{\Lambda}\left(t\right)\right)dt,
\]
given ergodicity in the solution with $u_{\Lambda}\left(t_{0}\right)$
already reaching the stationary state.

We compare the statistical spectra from the direct Monte Carlo model
simulation with the prediction from the equilibrium statistical mechanism
in Figure \ref{fig:Comparison-spec }. The flow energy spectra computed
both from the direct TKdV model simulations and from the Gibbs invariant
measure in the equilibrium statistical mechanism are compared. Starting
from two different $\theta^{-}$ initial states, the direct model
simulations offer the final equilibrium energy spectrum in a cimparable
decaying structure as predicted from the Gibbs measure and statistical
matching condition. This first comparison offers the consistency in
these two distinct approaches.

\begin{figure}
\begin{centering}
\includegraphics[scale=0.32]{./specEqui_th05}\includegraphics[scale=0.32]{./specEqui_th03}
\par\end{centering}
\caption{Comparison of the energy spectra computed from directly running the numerical 
model and the energy spectra predicted from the
Gibbs invariant measure in equilibrium statistical mechanism. The
model simulations start with the Gibbs measure with different
incoming flow parameter $\theta^{-}$.\label{fig:Comparison-spec }}
\end{figure}

\subsection{Ergodicity and mixing properties in the TKdV solutions}

Ergodicity is another crucial feature that needs to be confirmed in
the TKdV equation solution. To achieve reasonable statistics
from averaging the model time solutions, it is important to make sure
that the solution is turbulent mixing. Following we display the
direct numerical simulation solutions from the TKdV equation using
initial states sampled from different inverse temperatures.

\subsubsection*{Mixing solutions starting from negative inverse temperature $\theta<0$}

In the \textcolor{blue}{\emph{main text}}, the results shown always use negative inverse
temperature. As shown from \cite{abramov2003hamiltonian,bajars2013weakly}
and Figure \ref{fig:Test-distributions} before, this is a reasonable
choice to generate a decaying energy spectrum. Here as a further illustration,
we show that the negative inverse temperature case also generates
fast mixing solutions to calculate reasonable statistical averages.

In Figure \ref{fig:Realization1} and \ref{fig:Realization2}, we
display the single trajectory solutions for both downstream $D_{+}=0.24$
and upstream $D_{-}=1$ flows. The initial state is taken from the
samples in Algorithm \ref{alg:Metropolis-Hastings} with a negative
initial temperature $\theta<0$. The other parameters are kept the
same as in the main text. On the left panel, the contour plot for
the time-series of $u_{\Lambda}$ is shown for the downstream and
upstream flows. Both cases display turbulent dynamics, while the downstream
solution with $D_{+}<0$ generates more rare events and stronger turbulence.

For more detailed comparison, the right panels show one slice of the
time-series of $u_{\Lambda}\left(x_{1}\right)$ at one physical location
as well as the first two Fourier modes, $\hat{u}_{1},\hat{u}_{2}$,
in the spectral domain. The rare events in the downstream solution
in Figure \ref{fig:Realization1} is obvious with high peaks representing
the generation of skewed PDFs in the positive direction. In comparison,
the upstream solution in Figure \ref{fig:Realization2} displays almost
symmetric values in positive and negative disturbances, implying near-Gaussian
statistics in the incoming flow. The mixing properties are characterized
by the autocorrelation and decorrelation time of the state variable.
The decorrelation time $T_{k}$ and absolute decorrelation time $T_{k}^{\mathrm{abs}}$
together with the autocorrelation function $\mathcal{R}_{k}$ in each
spectral mode are defined as
\[
T_{k}=\left|\int\mathcal{R}_{k}\left(t\right)dt\right|,\;T_{k}^{\mathrm{abs}}=\int\left|\mathcal{R}_{k}\left(t\right)\right|dt,\quad\mathcal{R}_{k}\left(t\right)=\left\langle \hat{u}_{k}\left(0\right)\hat{u}_{k}\left(t\right)^{*}\right\rangle /\left\langle \left|\hat{u}_{k}\right|^{2}\right\rangle .
\]
Decaying autocorrelations in both the physical state and the spectral
modes confirm the fast mixing and turbulent dynamics in both the tested
regimes used in the main text (though weaker in upstream). We can
observe the much faster time scales in the downstream flow in the
autocorrelation functions in the corresponding time-series. The longer
decorrelation times in the upstream flow indicate oscillating autocorrelations
for a long time.

\begin{figure}[h]
\begin{centering}
\includegraphics[scale=0.35]{./snap_J32L3D024E100}\includegraphics[scale=0.35]{./tseries1_J32L3D024E100}
\par\end{centering}
\caption{Realization of the downstream flow solution with $D_{0}=0.24$ and
initial state from samples with negative inverse temperature.\label{fig:Realization1}}

\end{figure}
\begin{figure}[h]
\begin{centering}
\includegraphics[scale=0.35]{./snap1_J32L3D1E100}\includegraphics[scale=0.35]{./tseries1_J32L3D1E100}
\par\end{centering}
\caption{Realization of the upstream flow solution with $D_{0}=1$ and initial
state from samples with negative inverse temperature.\label{fig:Realization2}}
\end{figure}
As an additional comparison, we show the power spectra of the upstream
and downstream flows from the above test cases in Figure \ref{fig:Power-spectra}.
The power spectrum can be calculated by the Fourier transform of the
autocorrelation function
\[
S\left(\lambda\right)=\int_{-\infty}^{\infty}\mathcal{R}\left(\tau\right)e^{-i\lambda\tau}d\tau,
\]
which offers the characterization of the decay in power at each time
frequency. Again, consistent with the observations from experiments
\cite{bolles2018anomalous} and the numerics in the \textcolor{blue}{\emph{main text}},
the downstream power spectrum gets more energetic small time scales
at low frequency modes and a slower decay rate in the high frequency
modes compared with the upstream case. Also notice that the peak in
the incoming flow power spectrum is taken place at a much lower frequency.
This illustrates the slower mixing rate in the incoming flow state.

\begin{figure}[h]
\begin{centering}
\includegraphics[scale=0.32]{./power_up}\includegraphics[scale=0.32]{./power_down}
\par\end{centering}
\caption{Power spectra of the upstream and downstream states $u_{\Lambda}$.\label{fig:Power-spectra}}

\end{figure}

\subsubsection*{Non-mixing solutions starting from positive inverse temperature $\theta>0$}

In comparison, we also compute the solution starting from an initial
state with a positive inverse temperature $\theta>0$. In contrast to
the previous case in Figure \ref{fig:Realization2}, the only difference
in the present case is the initial state taken from the samples in
an invariant measure with positive inverse temperature. Figure \ref{fig:Realization3}
displays the corresponding solution in this 
$\theta>0$ case. In contrast to the previous mixing solution with
negative inverse temperature, drastic difference performance is generated
due to only the change in the initial state. The solution is recurrent in time with interacting small scales. As a result, the solution
in stationary waves lacks the proper mixing property and the autocorrelation just
oscillates for an extremely long time.

Accordingly in the autocorrelation functions and decorrelation times,
the autocorrelation functions just keep oscillating and all the modes
get long absolute decorrelation time (the decorrelation time is small
due to the cancellation from the oscillations). This non-mixing feature
with positive temperature supports the conclusion in the main text
that this regime is not appropriate for correctly representing the
physical statistics in the experimental setups. This simple test further
confirms the unrealistic energy spectra in the positive temperature
case with increasing energy in the smaller scales shown in \cite{bajars2013weakly}
and Figure \ref{fig:Test-distributions}.

\begin{figure}[h]
\begin{centering}
\includegraphics[scale=0.35]{./snap2_J32L3D1E100}\includegraphics[scale=0.35]{./tseries3_J32L3D1E100}
\par\end{centering}
\caption{Realization of the upstream flow solution with $D_{0}=1$ and initial
state from samples with positive inverse temperature.\label{fig:Realization3}}
\end{figure}
\

\bibliographystyle{plain}
\bibliography{refs_suppl}

\end{document}

### 2018-11-23 paper1_tKdV/tKdV_theory.tex
source files for paper1

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\usepackage{units}
\usepackage{subfig}
\newtheorem{remark}{Remark}

\title{A Statistical Dynamical Model to Predict Extreme Events and Anomalous Features in Shallow Water Waves with Abrupt Depth Change}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Andrew J. Majda}
\author[b]{M. N. J. Moore}
\author[a,1]{Di Qi}

\affil[a]{Department of Mathematics and Center for Atmosphere
and Ocean Science, Courant Institute of Mathematical Sciences, New
York University, New York, NY 10012}
\affil[b]{Department of Mathematics and Geophysical Fluid
Dynamics Institute, Florida State University, Tallahassee, FL}

% Please give the surname of the lead author for the running footer
\leadauthor{Majda}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Understanding and predicting extreme events and their anomalous statistics in complex nonlinear systems is a grand challenge in applied sciences as well as for engineering design. Recent controlled
laboratory experiments in weakly turbulent shallow water with abrupt
depth change exhibit a remarkable transition from nearly Gaussian
statistics to  extreme anomalous statistics with large
positive skewness of the surface height. We develop a statistical
dynamical model to explain and quantitatively predict the anomalous
statistical behavior. The incoming and outgoing  wares are modeled by the truncated Korteweg-de Vries equations statistically matched at the depth change.
The statistical matching of the known nearly Gaussian incoming Gibbs
state completely determines the predicted anomalous outgoing
Gibbs state, and successfully captures key features
of the experiment.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{A.J.M. designed the research. A.J.M., M.N.J.M., and D.Q. performed the research. A.J.M., M.N.J.M., and D.Q. wrote the paper.}
\authordeclaration{The authors declare no conflict of interest.}
%\equalauthors{\textsuperscript{1}A.J.M.(Author One) and D.Q. (Author Two) contributed equally to this work.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: qidi@cims.nyu.edu, jonjon@cims.nyu.edu}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{extreme anomalous event $|$ statistical TKdV model $|$ matching Gibbs measures $|$ surface wave displacement and slope}

\begin{abstract}
Understanding and predicting extreme events and their anomalous statistics
in complex nonlinear systems is a grand challenge in climate, material,
and neuroscience, as well as for engineering design. Recent controlled
laboratory experiments in weakly turbulent shallow water with abrupt
depth change (ADC) exhibit a remarkable transition from nearly Gaussian
statistics in incoming wave trains before the ADC to outgoing waves
trains after the ADC with extreme anomalous statistics with large
positive skewness of the surface height. Here we develop a statistical
dynamical model to explain and quantitatively predict the above anomalous
statistical behavior as experimental control parameters are varied.
The first step is to use incoming and outgoing truncated Korteweg-de
Vries (TKdV) equations matched in time at the ADC. The TKdV equation
is a Hamiltonian system which induces incoming and outgoing statistical
Gibbs invariant measures which are statistically matched at the ADC.
The statistical matching of the known nearly Gaussian incoming Gibbs
state at the ADC completely determines the predicted anomalous outgoing
Gibbs state, which can be calculated by a simple sampling algorithm, verified
by direct numerical simulations, and successfully captures key features
of the experiment. There is even an analytic formula for the anomalous
outgoing skewness. The strategy here should be useful for predicting
extreme anomalous statistical behavior in other dispersive media in
different settings.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{U}nderstanding and predicting extreme events and their anomalous statistics
in complex nonlinear systems is a grand challenge in climate, material
and neuroscience as well as for engineering design. This is a very
active contemporary topic in applied mathematics with qualitative
and quantitative models \cite{majda2012lessons,mohamad2018sequential,qi2016predicting,majda2015intermittency,majda2018model,majda2018simple,thual2016simple}
and novel numerical algorithms which overcome the curse of dimension
for extreme event prediction in large complex systems \cite{chen2018efficient,chen2017beating,chen2018conditional,mohamad2018sequential,qi2018predicting}.
The occurrence of Rogue waves as extreme events with different physical
settings of deep water \cite{adcock2014physics,cousins2015unsteady,farazmand2017reduced,onorato2001freak,dematteis2018rogue}
and shallow water \cite{sergeeva2011nonlinear,trulsen2012laboratory,viotti2014extreme}
is an important practical topic.

Recent controlled laboratory experiments in weakly turbulent shallow
water with abrupt depth change (ADC) exhibit a remarkable transition
from nearly Gaussian statistics in incoming wave trains before the
ADC to outgoing waves trains after the ADC with extreme anomalous
statistics with large positive skewness of the surface height \cite{bolles2018anomalous}.
Here we develop a statistical dynamical model to explain and quantitatively
predict the above anomalous statistical behavior as experimental control
parameters are varied. The first step is to use incoming and outgoing
truncated Korteweg-de Vries (TKdV) equations matched in time at the
ADC. The TKdV equation is a Hamiltonian system which induces incoming
and outgoing statistical Gibbs invariant measures which are statistically
matched at the ADC. The statistical matching of the known nearly Gaussian
incoming Gibbs state at the ADC completely determines the predicted
anomalous outgoing Gibbs state, which can be calculated by a simple
MCMC algorithm, verified by direct numerical simulations, and successfully
captures key features of the experiment. There is even an analytic
formula for the anomalous outgoing skewness. The strategy here should
be useful for predicting extreme anomalous statistical behavior in
other dispersive media in different settings.

\section{Experiments showing anomalous wave statistics by the abrupt depth change}

A series of experiments are carried out \cite{bolles2018anomalous}
studying the anomalous statistical behaviors in surface water waves
going through an abrupt depth transition. The unidirectional waves
propagate along a water tank over a step in the bottom topography,
and the surface displacements of the wave levels are measured at several
upstream and downstream locations. The wave field is excited by a
paddle wheel forcing with angle
\[
\theta\left(t\right)=\theta_{0}+\Delta\theta\sum_{n=1}^{N}a_{n}\cos\left(\omega_{n}t+\delta_{n}\right),\;E\sim\left(\Delta\theta\right)^{2}\sum_{n}a_{n}^{2}\omega_{n}^{2}.
\]
The total energy $E$ in the system is determined by the angle amplitude
$\Delta\theta$. Several observations are found in the anomalous wave
statistics:
\begin{itemize}
\item Distinct statistics are found between the incoming and outgoing wave
disturbances: the incoming waves display near-Gaussian statistics,
while the outgoing waves show skewness toward the positive displacement.
\item The non-Gaussian statistics is related with the total energy contained
in the system: larger driving amplitude $\Delta\theta$ will generate
stronger skewness in the PDFs.
\item The waves also show different characteristic peak wavelengths in incoming
and outgoing flows.
\end{itemize}

\section{Surface wave turbulence modeled by truncated KdV equation with depth dependence\label{sec:Surface-wave-turbulence}}

The surface wave turbulence is modeled by a one-dimensional deterministic
dynamical model. The Korteweg-de Vries (KdV) equation \cite{johnson1997modern}
is a leading-order approximation of the surface waves that are determined
by the balance of nonlinear and dispersive effects in an appropriate
far-field limit. Here, the KdV equation is truncated in the first
$\Lambda$ modes (with $J=2\Lambda+1$ grid points) to generate weakly
turbulent dynamics. Therefore, the surface disturbance is modeled by
the state variable $u_{\Lambda}^{\pm}\left(x,t\right)$ with superscript `$-$'
for the incoming waves and `$+$' for the outgoing waves. The Galerkin
truncated variable $u_{\Lambda}=\sum_{1\leq\left|k\right|\leq\Lambda}\hat{u}_{k}\left(t\right)e^{ikx}$
is normalized with zero mean $\hat{u}_{0}=0$ and unit energy $2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1$,
which are conserved quantities, and $u_{\Lambda}\equiv\mathcal{P}_{\Lambda}u$
denotes the subspace projection. The motion is governed by the truncated
KdV equation with depth change $D_{\pm}$ about the state variable
$u_{\Lambda}^{\pm}$
\begin{equation}
\frac{\partial u_{\Lambda}^{\pm}}{\partial t}+\frac{D_{\pm}^{-3/2}}{2}E_{0}^{1/2}L_{0}^{-3/2}\frac{\partial}{\partial x}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{\pm}\right)^{2}+D_{\pm}^{1/2}L_{0}^{-3}\frac{\partial^{3}u_{\Lambda}^{\pm}}{\partial x^{3}}=0,\label{eq:tKdV_nondim}
\end{equation}
on the normalized periodic domain $x\in\left[-\pi,\pi\right]$ with the conserved Hamiltonian decomposed into the difference of two
components containing the cubic and quadratic terms
\begin{align*}
\mathcal{H}_{\Lambda}^{\pm} & =  D_{\pm}^{-3/2}E_{0}^{1/2}L_{0}^{-3/2}H_{3}\left(u_{\Lambda}^{\pm}\right)-D_{\pm}^{1/2}L_{0}^{-3}H_{2}\left(u_{\Lambda}^{\pm}\right),\\
& H_{3}\left(u\right) =  \frac{1}{6}\int_{-\pi}^{\pi}u^{3}dx,\;H_{2}\left(u\right)=\frac{1}{2}\int_{-\pi}^{\pi}\left(\frac{\partial u}{\partial x}\right)^{2}dx.
\end{align*}
The model [\ref{eq:tKdV_nondim}] is non-dimensionalized in the periodic
domain. The depth is assumed to be unit $D_{-}=1$
before the abrupt depth change and becomes $D_{+}<1$ for the flows
after the change. We introduce the model parameters $\left(E_{0},L_{0},\Lambda\right)$
based on the following model assumptions:
\begin{itemize}
\item The wavenumber truncation $\Lambda$ is fixed in a moderate value
for generating weakly turbulent dynamics;
\item The state variable $u_{\Lambda}^{\pm}$ is normalized with zero mean
and unit energy, $\mathcal{M}\left(u_{\Lambda}\right)=0,\mathcal{E}\left(u_{\Lambda}\right)=1$,
conserved during the evolution, while $E_{0}$ characterizes the total
energy injected in the system based on the paddle amplitude $\left(\Delta\theta\right)^{2}$;
\item The length scale of the system is defined by $L_{0}$. The value is
chosen so that the resolved scale $\Delta x=2\pi L_{0}/J$ is comparable
with the the characteristic wave length $\lambda_{c}$ found from
the experiments.
\end{itemize}
The intuition for the distinct model dynamics comes from the balance
between the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
After the depth change, $D_{+}<1$, more weight is added in the cubic
term, $H_{3}$, for stronger nonlinearity and weaker dispersion for
the third-order derivative term reflected by the smaller coefficient
for $H_{2}$ in the Hamiltonian. Since $\frac{\partial u}{\partial x}$
is the slope of the wave height, $H_{2}\left(u\right)$ measures the
wave slope energy.

A \emph{deterministic matching condition} is given for the surface
displacement $u_{\Lambda}^{\pm}$ agreeing at the locations before
and after the abrupt depth change $T_{\mathrm{ADC}}$
\[
u_{\Lambda}^{-}\left(x,t\right)\mid_{t=T_{\mathrm{ADC}}-}=u_{\Lambda}^{+}\left(x,t\right)\mid_{t=T_{\mathrm{ADC}}+},
\]
assuming the abrupt depth change is met at $t=T_{\mathrm{ADC}}$.
Equation [\ref{eq:tKdV_nondim}] is not designed to capture the short
scale changes in rapid time. On the other hand, we are interested
in the model statistical transition before and after the depth change,
so it is reasonable to observe the suitable slow-time performance
in the large scale structures.

\subsection{Interpreting experimental parameters in the dynamical model}

The model parameters $\left(E_{0},L_{0},\Lambda\right)$ in [\ref{eq:tKdV_nondim}]
can be directly linked with the basic scales from the physical problem.
The important characterizing parameters measured from the experiments
include: $\epsilon=\frac{a}{H_{0}}$ the wave amplitude $a$ to water
depth $H_{0}$ ratio; $\delta=\frac{H_{0}}{\lambda_{c}}$ the water
depth to wavelength scale $\lambda_{c}$ ratio; and $D_{0}=\frac{d}{H_{0}}$
the normalized wave depth ratio with incoming flow depth $d=H_{0}$
to the outgoing flow depth $d<H_{0}$. The interpretations and reference
values of these model parameters are based on the experimental setup
\cite{bolles2018anomalous}. By comparing the characteristic physical
scales, the normalized TKdV equation [\ref{eq:tKdV_nondim}] can be
linked directly with the measured non-dimensional quantities by
\begin{equation}
L_{0}=6^{\frac{1}{3}}\left(M\epsilon^{\frac{1}{2}}\delta^{-1}\right),\;E_{0}=\frac{27}{2}\gamma^{-2}\left(M\epsilon^{\frac{1}{2}}\delta^{-1}\right),\label{eq:params}
\end{equation}
where $M$ defines the computational domain size $L_{d}=M\lambda_{c}$
as $M$-multiple of the characteristic wavelength $\lambda_{c}$,
and $\gamma=\frac{U}{a}$ represents the factor to normalize the total
energy in the state variable $u_{\Lambda}$ to one.

Consider the spatial discretization $J=2\Lambda+1$ so that the smallest
resolved scale is comparable with the characteristic wavelength
\[
\Delta x=\frac{2\pi M\lambda_{c}}{J}\lesssim\lambda_{c}\:\Rightarrow\:M=\frac{J}{2\pi}\sim5,\;J=32.
\]
Therefore in the practical numerical simulations, we pick $M=5$ and
$\gamma$ varies in the range $\left[0.5,1\right]$. Using the reference
experimental measurements \cite{bolles2018anomalous}, $\epsilon\in\left[0.0024,0.024\right],\delta\sim0.22$,
and $D_{0}$ changes from 1 to 0.24 before and after the depth change.
The reference values for the model scales can be estimated in the
range $L_{0}\in\left[2,6\right]$ and $E_{0}\in\left[50,200\right]$.
These are the values we will test in the direct numerical simulations.
See details about the derivation from scale analysis in \textcolor{blue}{\emph{SI Appendix, A}}.

\section{Equilibrium statistical mechanism for generating the stationary invariant measure}

Since the TKdV equation satisfies the Liouville property, the equilibrium
invariant measure can be described by an equilibrium statistical formulism
\cite{abramov2003hamiltonian,majda2006nonlinear,bajars2013weakly}
using a Gibbs measure with the conserved energy $\mathcal{E}_{\Lambda}$ and
Hamiltonian $\mathcal{H}_{\Lambda}$. The equilibrium invariant measure is dictated
by the conservation laws in the TKdV equation. In the case with fixed
total energy $E_{0}$, this is the \emph{mixed Gibbs measure} in the
truncated model with microcanonical energy and canonical Hamiltonian
ensembles \cite{abramov2003hamiltonian}
\begin{equation}
\mathcal{G}_{\theta}^{\pm}\left(u_{\Lambda}^{\pm};E_{0}\right)=C_{\theta}^{\pm}\exp\left(-\theta^{\pm}\mathcal{H}\left(u_{\Lambda}^{\pm}\right)\right)\delta\left(\mathcal{E}\left(u_{\Lambda}^{\pm}\right)-E_{0}\right),\label{eq:gibbs_mixed}
\end{equation}
with $\theta$ representing the ``inverse temperature''. The distinct
statistics in the upstream and downstream waves can be controlled
by the parameter value of the inverse temperature. The negative temperature
regime, $\theta^{\pm}<0$, is the appropriate regime to predict the
experiments as shown below. In the incoming flow field, the inverse
temperature $\theta^{-}$ is chosen so that $\mathcal{G}_{\theta}^{-}$
has Gaussian statistics. Using the above invariant measures [\ref{eq:gibbs_mixed}],
the expectation of any functional $F\left(u\right)$ can be computed
based on the Gibbs measure 
\[
\left\langle F\right\rangle _{\mathcal{G}_{\theta}}\equiv\int F\left(u\right)\mathcal{G}_{\theta}\left(u\right)du.
\]
The value of $\theta$ in the invariant measure is specified from
$\left\langle H_{\Lambda}\right\rangle _{\mathcal{G}_{\theta}}$ \cite{abramov2003hamiltonian,bajars2013weakly}.
The invariant measure also predicts an equilibrium energy spectrum
without running the TKdV equation directly. On the other hand, the
time autocorrelation and transient statistics about the state variable
$u_{\Lambda}$ cannot be recovered from the statistical theory.

\subsection{Statistical matching condition in invariant measures before and after the abrupt depth change}

The Gibbs measures $\mathcal{G}_{\theta}^{\pm}$ are defined based
on the different inverse temperatures $\theta^{\pm}$ on the two sides
of the solutions
\[
\begin{aligned}\mu_{t}^{-}\left(u_{\Lambda}^{-};D_{-}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}-}}=u_{0},\;t<T_{\mathrm{ADC}};\\
\mu_{t}^{+}\left(u_{\Lambda}^{+};D_{+}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}+}}=u_{0},\;t>T_{\mathrm{ADC}},
\end{aligned}
\]
where $u_{0}$ represents the deterministic matching condition between
the incoming and outgoing waves. The two distributions, $\mu_{t}^{-},\mu_{t}^{+}$
should also be matched at the depth change location $T_{\mathrm{ADC}}$,
so that
\[
\mu_{\infty}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{+}\left(u_{\Lambda}\right).
\]
In matching the flow statistics before and after the abrupt depth
change, first we use the conservation of the deterministic Hamiltonian
$H_{\Lambda}^{+}$ after the depth change. Then assuming ergodicity
\cite{abramov2003hamiltonian,majda2006nonlinear}, the statistical
expectation for the Hamiltonian $\left\langle H_{\Lambda}^{+}\right\rangle $
is conserved in time after the depth change at $t=T_{\mathrm{ADC}}$
and should stay in the same value as the system approaches equilibrium
as $t\rightarrow\infty$. The final statistical matching condition
to get the outgoing flow statistics with parameter $\theta^{+}$ can
be found by
\begin{equation}
\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{+}}=\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{-}},\label{eq:matching}
\end{equation}
with the outgoing flow Hamiltonian $H_{\Lambda}^{+}$ and the Gibbs
measures $\mathcal{G}_{\theta}^{\pm}$ before and after the abrupt
depth change.

\section{The nearly Gaussian incoming statistical state}

The incoming flow is always characterized by a near-Gaussian distribution
in the wave displacement. It is found that a physically consistent
Gibbs measure should take negative values in the inverse temperature
parameter $\theta<0$, where a proper distribution function and a
decaying energy spectrum are generated (see \cite{bajars2013weakly}
and \textcolor{blue}{\emph{SI Appendix, B.1}} for the explicit simulation results).
The upstream Gibbs measure $\mathcal{G}_{\theta}^{-}$ with $D_{-}=1$
displays a wide parameter regime in $\left(\theta^{-},E_{0}\right)$
with near-Gaussian statistics. In the left panel of Figure \ref{fig:matching}
(a), the inflow skewness $\kappa_3^{-}$ varies only within small amplitudes
among changing values of $E_{0}$ and $\theta^{-}$. The incoming
flow PDF then can be determined by picking the proper parameter value
$\theta^{-}$ in the near Gaussian regime with small skewness. In
contrast, the downstream Gibbs measure $\mathcal{G}_{\theta}^{+}$
with $D_{+}=0.24$ shown in the right panel of Figure \ref{fig:matching}
(a) generates much larger skewness $\kappa_3^{+}$ with highly skewed
PDFs as the absolute value of $\theta^{+}$ and the total energy level
$E_{0}$ increase in amplitude. The solid lines in Figure \ref{fig:matching}
(c) offer a further confirmation of the transition from near-Gaussian
statistics with tiny $\kappa_3^{-}$ to strongly skewed distribution
$\kappa_3^{+}$ after the depth change.

In the next step, the value of the downstream $\theta^{+}$ is determined
based on the matching condition [\ref{eq:matching}]. The expectation
$\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{-}}$
about the incoming flow Gibbs measure can be calculated according
to the predetermined parameter values of $\theta^{-}$ as well as
$E_{0}$ from the previous step. For the direct numerical experiments
shown later in Figure \ref{fig:Changes-stats}, we pick proper choices
of test parameter values as $L_{0}=6, E_{0}=100$ and $\theta^{-}=-0.1,-0.3,-0.5$.
More test cases with different system energy $E_{0}$ can be found
in \textcolor{blue}{\emph{SI Appendix, B.2}} where similar transition from near
Gaussian symmetric PDFs to skewed PDFs in the flow state $u_{\Lambda}^{\pm}$
can always be observed.

\subsection{Direct numerical model simulations}

Besides the prediction of equilibrium statistical measures from the
equilibrium statistical approach, another way to predict the downstream
model statistics is through running the dynamical model [\ref{eq:tKdV_nondim}]
directly. The TKdV equation is found to be ergodic with proper mixing
property as measured by the decay of autocorrelations as long as the
system starts from a negative inverse temperature state as described
before. For direct numerical simulations of the TKdV equations, a
proper symplectic integrator is required to guarantee the Hamiltonian
and energy are conserved in time. It is crucial to use the symplectic
scheme to guarantee the exact conservation of the energy and Hamiltonian
since they are playing the central role in generating the invariant
measure and the statistical matching. The symplectic schemes used
here for the time integration of the equation is the 4th-order midpoint
method \cite{mclachlan1993symplectic}. Details about the mixing properties
from different initial states and the numerical algorithm are described
in \textcolor{blue}{\emph{SI Appendix, C}}.

\begin{figure}[h]
\centering
\includegraphics[scale=1.]{./fig1}

\caption{Fist row: skewness from the Gibbs measures in incoming and outgoing
flow states with different values of total energy $E_{0}$ and inverse
temperature $\theta$ (notice the different scales in the incoming
and outgoing flows); Second row: outgoing flow parameter $\theta^{+}$
as a function of the incoming flow $\theta^{-}$ computed from the
statistical matching condition with three energy level $E_{0}$; Last
row: skewness in the outgoing flow with the matched value of $\theta^{+}$
as a function of the inflow parameter $\theta^{-}$ (the theoretical
predictions using [\ref{eq:skewness}] are compared).\label{fig:matching}}
\end{figure}

\section{Predicting extreme anomalous behavior after the ADC by statistical matching}

With the inflow statistics well described and the numerical scheme
set up, we are able to predict the downstream anomalous statistics
starting from the near-Gaussian incoming flow going through the abrupt
depth change from $D_{-}=1$ to $D_{+}=0.24$. First, we consider
the statistical prediction in the downstream equilibrium measure directly
from the matching condition. The downstream parameter value $\theta^{+}$
is determined by solving the nonlinear equation [\ref{eq:matching}]
as a function of $\theta^{+}$, $F\left(\theta^{+}\right)=\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{+}}\left(\theta^{+}\right)-\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{-}}=0$.
In the numerical approach, we adopt a modified secant method avoiding
the stiffness in the parameter regime (see the \textcolor{blue}{\emph{SI Appendix, B.2}}
for the algorithm). The fitted solution is plotted in Figure
\ref{fig:matching} (b) as a function of the proposed inflow $\theta^{-}$.
A nonlinear $\theta^{-}$--$\theta^{+}$ relation is discovered from
the matching condition. The downstream inverse temperature $\theta^{+}$
will finally saturate at some level. The corresponding downstream
skewness of the wave displacement $u_{\Lambda}$ predicted from the
statistical matching of Gibbs measures is plotted in Figure \ref{fig:matching} (c). In general, a large positive skewness
for outgoing flow $\kappa_3^{+}$ is predicted from the theory, while
the incoming flow skewness $\kappa_3^{-}$ is kept in a small value
in a wide range of $\theta^{-}$. Note that with $\theta^{-}\sim0$
(that is, using the microcanonical ensemble only with energy conservation),
the outflow statistics are also near Gaussian with weak skewness.
The skewness in the outflow statistics grows as the inflow parameter
value $\theta^{-}$ increases in amplitude.

For a second approach, we can use direct numerical simulations starting
from the initial state sampled from the incoming flow Gibbs measure
$\mathcal{G}_{\theta}^{-}$ and check the transient changes in the
model statistics. Figure \ref{fig:Changes-stats} illustrates the
change of statistics as the flow goes through the abrupt depth change.
The first row plots the changes in the skewness and kurtosis for the
state variable $u_{\Lambda}$ after the depth change at $t=0$. The
PDFs in the incoming and outgoing flow states are compared with three different
initial inverse temperatures $\theta^{-}$. After the depth changes
to $D_{0}=0.24$ abruptly at $t=0$, both the skewness and kurtosis
jump to a much larger value in a short time, implying the rapid transition
to a highly skewed non-Gaussian statistical regime after the depth
change. Further from Figure \ref{fig:Changes-stats}, different initial
skewness (but all relatively small) is set due to the various values
of $\theta^{-}$. With small $\theta^{-}=-0.1$, the change in the
skewness is not very obvious (see the second row of Figure \ref{fig:Changes-stats}
for the incoming and outgoing PDFs of $u_{\Lambda}$). In comparison,
if the incoming flow starts from the initial parameter $\theta^{-}=-0.3$
and $\theta^{-}=-0.5$, much larger increase in the skewness is induced
from the abrupt depth change. Furthermore, in the detailed plots in
the third row of Figure \ref{fig:Changes-stats} for the downstream PDFs
under logarithmic scale, fat tails towards the positive direction
can be observed, which represent the extreme events in the downstream
flow (see also Figure \ref{fig:Realization} for the time-series of
$u_{\Lambda}$).

As a result, the downstream statistics in final equilibrium predicted
from the direct numerical simulations here agree with the equilibrium
statistical mechanism prediction illustrated in Figure \ref{fig:matching}. The prediction from these two different approaches
confirm each other.

\begin{figure*}
\centering
\includegraphics[scale=.9]{./fig2}

\caption{Changes in the statistics of the flow state going through the abrupt
depth change. The initial ensemble is set with the incoming flow Gibbs
measure with different inverse temperature $\theta^{-}$. First row:
time evolution of the skewness and kurtosis. The abrupt depth change
is taking place at $t=0$; Second row: inflow and outflow PDFs of
$u_{\Lambda}$; Third row: the downstream PDFs fitted with Gamma distributions
with consistent variance and skewness (in log coordinate in $y$);
Last row: energy spectra in the incoming and outgoing flows.\label{fig:Changes-stats}}
\end{figure*}

\section{Analytic formula for the upstream skewness after the ADC}

A statistical link between the upstream and downstream energy spectra
can be found for an analytical prediction of the skewness in the flow
state $u$ after the ADC. The skewness of the state variable $u_{j}$
at one spatial grid point is defined as the ratio between the third
and second moments
\[
\kappa_{3}=\left\langle u_{j}^{3}\right\rangle _{\mu}/\left\langle u_{j}^{2}\right\rangle _{\mu}^{\frac{3}{2}}.
\]
Now we introduce mild assumptions on the distribution functions:
\begin{itemize}
\item The upstream equilibrium measure $\mu_{-}$ has a relatively small
skewness so that
\[
\left\langle H_{3}\right\rangle _{\mu_{-}}=\frac{1}{6}\int_{-\pi}^{\pi}\left\langle u^{3}\right\rangle _{\mu_{-}}dx\equiv\epsilon;
\]
\item The downstream equilibrium measure $\mu_{+}$ is homogeneous at each
physical grid point, so that the second and third moments are invariant
at each grid point
\[
\left\langle u_{j}^{2}\right\rangle _{\mu_{+}}=\sigma^{2}=\pi^{-1},\quad\left\langle u_{j}^{3}\right\rangle _{\mu_{+}}=\sigma^{3}\kappa_{3}=\pi^{-\frac{3}{2}}\kappa_{3}^{+}.
\]
\end{itemize}
Then the skewness of the downstream state variable $u_{\Lambda}^{+}$
after the ADC is given by the difference between the inflow and outflow
wave slope energy of $u_x$
\begin{equation}
\begin{aligned}
\kappa_{3}^{+} = &\frac{3}{2}\pi^{\frac{1}{2}}L_{0}^{-\frac{3}{2}}E_{0}^{-\frac{1}{2}}D_{+}^{2}\int_{-\pi}^{\pi}\left[\left\langle u_{x}^{2}\right\rangle _{\mu+}-\left\langle u_{x}^{2}\right\rangle _{\mu-}\right]dx\\
 & +3\pi^{\frac{1}{2}}\epsilon.\label{eq:skewness}
\end{aligned}
\end{equation}
The detailed derivation is shown in \textcolor{blue}{\emph{SI Appendix, B.2}}.
In particular, the downstream skewness with near-Gaussian inflow statistics
$\epsilon\ll1$ is positive if and only if the difference of the incoming and outgoing 
wave slope energy is positive. This means that there is more
small scale wave slope energy in the outgoing state. As an evidence, 
in the last row of Figure \ref{fig:Changes-stats} in all the weak
and strong skewness cases, the outflow energy spectrum always has
a slower decay rate than the inflow energy spectrum which possesses
stronger energy in larger scales and weaker energy in the smaller
scales.

In Figure \ref{fig:matching} (c), we compare the accuracy of the
theoretical estimation [\ref{eq:skewness}] with numerical tests.
In the regime with small incoming inverse temperature $\theta^{-}$,
the theoretical formula offers a quite accurate approximation of the
third-order skewness using only information from the second-order
moments of the wave-slope spectrum.

\section{Key features from experiments captured by the statistical dynamical model}

In this final section, we emphasize the crucial features generated
by the statistical dynamical model [\ref{eq:tKdV_nondim}] by making
comparison with the experimental observations in \cite{bolles2018anomalous}.
As from the scale analysis displayed in Section \ref{sec:Surface-wave-turbulence},
the theory is set in the same parameter regime as the experimental
setup.

\begin{figure*}[h]
\centering
\subfloat[downstream state $D_{0}=0.24$]{\includegraphics[scale=.95]{./fig3_1}}
\vspace{-1em}
\subfloat[upstream state $D_{0}=1$]{\includegraphics[scale=.95]{./fig3_2}}
\vspace{-1em}
\caption{Realization of the downstream and upstream flow solutions $u_{\Lambda}^{\pm}$. Note the larger vertical scale in the downstream time-series plot.\label{fig:Realization}}
\end{figure*}
\begin{itemize}
\item The transition from near-Gaussian to skewed non-Gaussian distribution
as well as the jump in both skewness and kurtosis observed in the
experiment observations (Fig. 1 of  \cite{bolles2018anomalous}) can
be characterized by the statistical model simulation results (see
the first and second row of Figure \ref{fig:Changes-stats}). Notice
that the difference in the decay of third and fourth moments in the far
end of the downstream regime from the experimental data is due to the
dissipation effect in the flow from the wave absorbers that is not modeled in the statistical
model here. The model simulation time-series plotted in Figure \ref{fig:Realization}
can be compared with the observed time sequences from experiments
(Fig. 1 of \cite{bolles2018anomalous}). The downstream simulation generates
waves with strong and frequent intermittency towards the positive
displacement, while the upstream waves show symmetric displacements
in two directions with at most small peaks in slow time. Even in the
time-series at a single location $x=0$, the long-time variation displays similar
structures. 
\item The downstream PDFs in experimental data are estimated with a Gamma
distribution in Fig. 2 of \cite{bolles2018anomalous}. Here in the same
way, we can fit the highly skewed outgoing flow PDFs from the numerical
results with the Gamma distribution
\[
\rho\left(u;k,\alpha\right)=\frac{e^{-k}\alpha^{-1}}{\Gamma\left(k\right)}\left(k+\alpha^{-1}u\right)^{k-1}e^{-\alpha^{-1}u}.
\]
The parameters $\left(k,\alpha\right)$ in the Gamma distribution
are fitted according to the measured statistics in skewness and variance,
that is, $\sigma^{2}=k\alpha^{2},\kappa_{3}=2/\sqrt{k}$. And the excess
kurtosis of the Gamma distribution can be recovered as $\kappa_{4}=6/k$.
As shown in the third row of Figure \ref{fig:Changes-stats}, excellent
agreement in the PDFs with the Gamma distributions
is reached in consistency with the experimental data observations.
The accuracy with this approximation increases as the initial inverse
temperature $\theta^{-}$ increases in value to generate more skewed
distribution functions.
\item The experiments also have the up and down stream power spectra in
time (Fig. 4 of \cite{bolles2018anomalous}), which shows more energy
at small time scales, i.e., a relatively slower decay rate in the
downstream compared with the upstream case. This is also observed
in the direct numerical simulations here (detailed results shown
in \textcolor{blue}{\emph{SI Appendix, C.2}}). The downstream state contains
more energetic high frequencies.
The peak frequency illustrates the occurrence of the transporting
waves along the water tank.
\end{itemize}

\section{Concluding discussion}

We have developed a statistical dynamical model to explain and predict
extreme events and anomalous features in shallow water waves with
abrupt depth change. The theory is based on the dynamical modeling
strategy consisting of the TKdV equation matched at the abrupt depth
change with conservation of energy and Hamiltonian. Predictions can
be made of the extreme events and anomalous features by matching incoming
and outgoing statistical Gibbs measures before and after the abrupt
depth transition. The statistical matching of the known nearly Gaussian
incoming Gibbs state completely determines the predicted anomalous
outgoing Gibbs state, which can be calculated by a simple sampling
algorithm, verified by direct numerical simulations, and successfully
captures key features of the experiment. An analytic formula for the
anomalous outgoing skewness is also derived. The strategy here should
be useful for predicting extreme statistical events in other dispersive
media in different settings.

\acknow{This research of A. J. M. is partially supported by the Office of
Naval Research through MURI N00014-16-1-2161. D. Q. is supported as
a postdoctoral fellow on the second grant. M.N.J.M }

\showacknow % Display the acknowledgments section

% \pnasbreak splits and balances the columns before the references.
% If you see unexpected formatting errors, try commenting out this line
% as it can run into problems with floats and footnotes on the final page.
\pnasbreak

% Bibliography
\bibliography{refs}

\end{document}

### 2018-11-23 Sampling_MH/sampling_MH.m

@@ -36,12 +36,12 @@ for ii = 1:N_iter
         uk_tilde = uk;
               
         if jj==1 || jj==Lambda/2+1
-        pert = sqrt(2*Lambda*E0)*randn(1);
+        pert = sqrt(.1*2*Lambda*E0)*randn(1);
         uk_tilde(jj) = uk(jj)+pert;
         E = .5*sum(abs(uk_tilde).^2)/Lambda^2;
         uk_tilde = sqrt(E0/E)*uk_tilde;
         else
-        pert = sqrt(2*Lambda*E0)*(randn(1)+1i*randn(1));
+        pert = sqrt(.1*2*Lambda*E0)*(randn(1)+1i*randn(1));
         uk_tilde(jj) = uk(jj)+pert;
         uk_tilde(end-jj+2) = conj(uk_tilde(jj));
         E = .5*sum(abs(uk_tilde).^2)/Lambda^2;
@@ -74,6 +74,6 @@ figure(1)
 plot(centers,counts/MC/(centers(2)-centers(1))); hold on;
 ylabel('PDF of Hamiltonian')
 title(['Lambda = ',num2str(Lambda),', E0 = ',num2str(E0),', D0 = ', num2str(D0), ', N = ',num2str(MC)]);
-figure
+figure(2)
 plot(kk(1:Lambda/2+1),enek(1:Lambda/2+1),'-o'); hold on;
 title('energy spectra <|u_k|^2>')

### 2018-11-25 paper1_tKdV/tKdV_theory.tex
NM

@@
 \newtheorem{remark}{Remark}
+\usepackage{todonotes}
 
@@
 for extreme event prediction in large complex systems \cite{chen2018efficient,chen2017beating,chen2018conditional,mohamad2018sequential,qi2018predicting}.
-The occurrence of Rogue waves as extreme events with different physical
+The occurrence of Rogue waves as extreme events within different physical
 settings of deep water \cite{adcock2014physics,cousins2015unsteady,farazmand2017reduced,onorato2001freak,dematteis2018rogue}
@@
 
-Recent controlled laboratory experiments in weakly turbulent shallow
-water with abrupt depth change (ADC) exhibit a remarkable transition
-from nearly Gaussian statistics in incoming wave trains before the
-ADC to outgoing waves trains after the ADC with extreme anomalous
-statistics with large positive skewness of the surface height \cite{bolles2018anomalous}.
+% ORIGINAL
+%Recent controlled laboratory experiments in weakly turbulent shallow
+%water with abrupt depth change (ADC) exhibit a remarkable transition
+%from nearly Gaussian statistics in incoming wave trains before the
+%ADC to outgoing waves trains after the ADC with extreme anomalous
+%statistics with large positive skewness of the surface height \cite{bolles2018anomalous}.
+
+Recent laboratory experiments in weakly turbulent shallow water reveal a remarkable transition from Gaussian to anomalous behavior as surface waves cross an abrupt depth change (ADC). 
+A normally-distributed incoming wave train, upstream of the ADC, transitions to a highly non-Gaussian outgoing wave train, downstream of the ADC, that exhibits large positive skewness of the surface height and more frequent extreme events \cite{bolles2018anomalous}.
 Here we develop a statistical dynamical model to explain and quantitatively
@@
 
+\todo[inline]{If you are interested, we could cite papers on `Optical rogue waves' and/or rogue waves in microwave systems here.}
+
 \section{Experiments showing anomalous wave statistics by the abrupt depth change}
@@
 studying the anomalous statistical behaviors in surface water waves
-going through an abrupt depth transition. The unidirectional waves
-propagate along a water tank over a step in the bottom topography,
+going through an abrupt depth transition. 
+The unidirectional waves propagate along a water tank over a step in the bottom topography,
 and the surface displacements of the wave levels are measured at several
@@
 waves along the water tank.
+
+\todo[inline]{I do not understand the last sentence above. Why does the peak frequency illustrate the occurrence of transporting water waves?}
+
 \end{itemize}
@@
 Naval Research through MURI N00014-16-1-2161. D. Q. is supported as
-a postdoctoral fellow on the second grant. M.N.J.M }
+a postdoctoral fellow on the second grant. M.N.J.M would like to acknowledge support from Simons grant 524259.}
 

### 2018-11-25 paper1_tKdV/tKdV_theory.tex
NM

@@
 Here we develop a statistical dynamical model to explain and quantitatively
-predict the above anomalous statistical behavior as experimental control
+predict this anomalous behavior as experimental control
 parameters are varied. The first step is to use incoming and outgoing
@@
 
-\section{Experiments showing anomalous wave statistics by the abrupt depth change}
+\section{Experiments showing anomalous wave statistics induced by an abrupt depth change}
 
-A series of experiments are carried out \cite{bolles2018anomalous}
-studying the anomalous statistical behaviors in surface water waves
-going through an abrupt depth transition. 
-The unidirectional waves propagate along a water tank over a step in the bottom topography,
-and the surface displacements of the wave levels are measured at several
-upstream and downstream locations. The wave field is excited by a
-paddle wheel forcing with angle
+% ORIGINAL
+%A series of experiments are carried out \cite{bolles2018anomalous}
+%studying the anomalous statistical behavior of surface water waves
+%crossing an abrupt depth transition. 
+%The unidirectional waves propagate along a water tank over a step in the bottom topography,
+%and the surface displacements of the wave levels are measured at several
+%upstream and downstream locations. The wave field is excited by a
+%paddle wheel forcing with angle
+
+Controlled laboratory experiments were carried out in  \cite{bolles2018anomalous} to examine the statistical behavior of surface waves crossing an ADC.
+In these experiments, nearly unidirectional waves are generated by a paddle wheel and propagate through a long, narrow wave tank.
+Midway through, the waves encounter a step in the bottom topography, and thus abruptly transition from one depth to another.
+The paddle wheel is forced with angle
 \[
@@
 \]
-The total energy $E$ in the system is determined by the angle amplitude
-$\Delta\theta$. Several observations are found in the anomalous wave
-statistics:
+where the weights $a_n$ are Gaussian in spectral space, and
+the total energy $E$ in the system is determined by the angle amplitude
+$\Delta\theta$. Optical measurements of the free surface reveal a number of surprising statistical features:
 \begin{itemize}
@@
 \item The non-Gaussian statistics is related with the total energy contained
-in the system: larger driving amplitude $\Delta\theta$ will generate
-stronger skewness in the PDFs.
+in the system: larger driving amplitudes $\Delta\theta$ generate
+stronger skewness in the outgoing PDFs.
 \item The waves also show different characteristic peak wavelengths in incoming
 and outgoing flows.
+
+\todo[inline]{I am not sure we need to include this last item.
+It is not really an experimental observation, simply a calculation that can be done using the dispersion relation and the two depths.
+Could we simply take this last item out?}
+
 \end{itemize}

### 2018-11-25 paper1_tKdV/tKdV_theory.tex
NM

@@
 and quantitative models \cite{majda2012lessons,mohamad2018sequential,qi2016predicting,majda2015intermittency,majda2018model,majda2018simple,thual2016simple}
-and novel numerical algorithms which overcome the curse of dimension
+and novel numerical algorithms which overcome the curse of dimensionality
 for extreme event prediction in large complex systems \cite{chen2018efficient,chen2017beating,chen2018conditional,mohamad2018sequential,qi2018predicting}.
@@
 
-\todo[inline]{If you are interested, we could cite papers on `Optical rogue waves' and/or rogue waves in microwave systems here.}
+\todo[inline]{If you like, we could cite papers on `Optical rogue waves' and/or rogue waves in microwave systems here.}
 
@@
 \]
-where the weights $a_n$ are Gaussian in spectral space, and
-the total energy $E$ in the system is determined by the angle amplitude
+where the weights $a_n$ are Gaussian in spectral space.
+The peak frequency $\omega_p$ gives rise to a characteristic wavelength $\lambda_c$ of water waves, which can be estimated using the dispersion relation.
+Above, the total energy $E$ in the system is determined by the angle amplitude
 $\Delta\theta$. Optical measurements of the free surface reveal a number of surprising statistical features:
@@
 \item Distinct statistics are found between the incoming and outgoing wave
-disturbances: the incoming waves display near-Gaussian statistics,
-while the outgoing waves show skewness toward the positive displacement.
-\item The non-Gaussian statistics is related with the total energy contained
-in the system: larger driving amplitudes $\Delta\theta$ generate
-stronger skewness in the outgoing PDFs.
+disturbances: incoming waves display near-Gaussian statistics,
+while outgoing waves skew strongly towards positive displacement.
+\item The degree of non-Gaussianity present in the outgoing wave field depends on the total energy of the system: 
+larger driving amplitudes $\Delta\theta$ generate stronger skewness in the surface displacement PDFs.
 \item The waves also show different characteristic peak wavelengths in incoming
@@
 \todo[inline]{I am not sure we need to include this last item.
-It is not really an experimental observation, simply a calculation that can be done using the dispersion relation and the two depths.
-Could we simply take this last item out?}
+It is not really an experimental observation, simply a calculation that can be done using the dispersion relation and the values of the two depths.
+Could we simply take this last item out?
+Would you prefer to replace it with another observation?
+I could easily come up with one.
+Perhaps a comment on the decay rate of the power spectra.}
 
@@
 The surface wave turbulence is modeled by a one-dimensional deterministic
-dynamical model. The Korteweg-de Vries (KdV) equation \cite{johnson1997modern}
+dynamical model. 
+\todo[inline]{This first sentence seems awkward to me.}
+The Korteweg-de Vries (KdV) equation \cite{johnson1997modern}
 is a leading-order approximation of the surface waves that are determined
@@
 $\Lambda$ modes (with $J=2\Lambda+1$ grid points) to generate weakly
-turbulent dynamics. Therefore, the surface disturbance is modeled by
+turbulent dynamics. The surface displacement is described by
 the state variable $u_{\Lambda}^{\pm}\left(x,t\right)$ with superscript `$-$'
@@
 which are conserved quantities, and $u_{\Lambda}\equiv\mathcal{P}_{\Lambda}u$
-denotes the subspace projection. The motion is governed by the truncated
-KdV equation with depth change $D_{\pm}$ about the state variable
-$u_{\Lambda}^{\pm}$
+denotes the subspace projection. The evolution of  $u_{\Lambda}^{\pm}$ is governed by the truncated
+KdV equation with depth change $D_{\pm}$
 \begin{equation}
@@
 \end{equation}
-on the normalized periodic domain $x\in\left[-\pi,\pi\right]$ with the conserved Hamiltonian decomposed into the difference of two
-components containing the cubic and quadratic terms
+on the normalized periodic domain $x\in\left[-\pi,\pi\right]$. The conserved Hamiltonian can be decomposed into a cubic and a quadratic term
 \begin{align*}
@@
 \end{align*}
-The model [\ref{eq:tKdV_nondim}] is non-dimensionalized in the periodic
+Equation [\ref{eq:tKdV_nondim}] is non-dimensionalized in the periodic
 domain. The depth is assumed to be unit $D_{-}=1$
-before the abrupt depth change and becomes $D_{+}<1$ for the flows
-after the change. We introduce the model parameters $\left(E_{0},L_{0},\Lambda\right)$
-based on the following model assumptions:
+before the ADC and becomes $D_{+}<1$
+after the ADC. We introduce parameters $\left(E_{0},L_{0},\Lambda\right)$
+based on the following assumptions:
 \begin{itemize}
@@
 and unit energy, $\mathcal{M}\left(u_{\Lambda}\right)=0,\mathcal{E}\left(u_{\Lambda}\right)=1$,
-conserved during the evolution, while $E_{0}$ characterizes the total
-energy injected in the system based on the paddle amplitude $\left(\Delta\theta\right)^{2}$;
-\item The length scale of the system is defined by $L_{0}$. The value is
+which are conserved during evolution. Meanwhile, $E_{0}$ characterizes the total
+energy injected into the system based on the paddle amplitude $\left(\Delta\theta\right)^{2}$;
+\item The length scale of the system is denoted $L_{0}$. The value is
 chosen so that the resolved scale $\Delta x=2\pi L_{0}/J$ is comparable
-with the the characteristic wave length $\lambda_{c}$ found from
+to the the characteristic wave length $\lambda_{c}$ in
 the experiments.
 \end{itemize}
-The intuition for the distinct model dynamics comes from the balance
-between the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
-After the depth change, $D_{+}<1$, more weight is added in the cubic
-term, $H_{3}$, for stronger nonlinearity and weaker dispersion for
-the third-order derivative term reflected by the smaller coefficient
-for $H_{2}$ in the Hamiltonian. Since $\frac{\partial u}{\partial x}$
-is the slope of the wave height, $H_{2}\left(u\right)$ measures the
+% ORIGINAL
+%The intuition for the distinct model dynamics comes from the balance
+%between the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
+%After the depth change, $D_{+}<1$, more weight is added in the cubic
+%term, $H_{3}$, for stronger nonlinearity and weaker dispersion for
+%the third-order derivative term reflected by the smaller coefficient
+%for $H_{2}$ in the Hamiltonian. 
+Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the balance of the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
+The depth change, $D_{+}<1$, increases the weight of $H_3$ and decreases that of $H_2$, thus promoting the effects of nonlinearity and reducing dispersion.
+Since $\frac{\partial u}{\partial x}$ is the slope of the wave height, $H_{2}\left(u\right)$ measures the
 wave slope energy.
 
-A \emph{deterministic matching condition} is given for the surface
-displacement $u_{\Lambda}^{\pm}$ agreeing at the locations before
-and after the abrupt depth change $T_{\mathrm{ADC}}$
+A \emph{deterministic matching condition} is applied to the surface displacement $u_{\Lambda}^{\pm}$ to link the incoming and outgoing wave trains.
+Assuming the abrupt depth change is met at $t=T_{\mathrm{ADC}}$, the matching condition is given by
 \[
@@
 \]
-assuming the abrupt depth change is met at $t=T_{\mathrm{ADC}}$.
 Equation [\ref{eq:tKdV_nondim}] is not designed to capture the short
-scale changes in rapid time. On the other hand, we are interested
-in the model statistical transition before and after the depth change,
-so it is reasonable to observe the suitable slow-time performance
-in the large scale structures.
+scale changes in rapid time. 
+Rather, since we are interested in modeling statistics before and after the depth change, we will examine the long-time dynamics of large-scale structures.
+
+\todo[inline]{I modified the last sentence above. Feel free to revert to your original sentence if you do not like mine.}
 
@@
 The model parameters $\left(E_{0},L_{0},\Lambda\right)$ in [\ref{eq:tKdV_nondim}]
-can be directly linked with the basic scales from the physical problem.
+can be directly linked to the basic scales from the physical problem.
 The important characterizing parameters measured from the experiments
@@
 
+\todo[inline]{Is there a difference between $L_0$ and $L_d$? typo?}
+
 Consider the spatial discretization $J=2\Lambda+1$ so that the smallest
@@
 
+\todo[inline]{In this section title, did you mean `statistical mechanism' or `statistical mechanics'?}
+
 Since the TKdV equation satisfies the Liouville property, the equilibrium
@@
 statistics in the upstream and downstream waves can be controlled
-by the parameter value of the inverse temperature. The negative temperature
-regime, $\theta^{\pm}<0$, is the appropriate regime to predict the
+by the value of $\theta$. Negative temperature, $\theta^{\pm}<0$, is the appropriate regime to predict the
 experiments as shown below. In the incoming flow field, the inverse
@@
 \]
+\todo[inline]{I do not believe we have defined $\mu_{\infty}^{-}$ anywhere.}
 In matching the flow statistics before and after the abrupt depth
-change, first we use the conservation of the deterministic Hamiltonian
+change, we first use the conservation of the deterministic Hamiltonian
 $H_{\Lambda}^{+}$ after the depth change. Then assuming ergodicity
@@
 with near-Gaussian statistics. In the left panel of Figure \ref{fig:matching}
-(a), the inflow skewness $\kappa_3^{-}$ varies only within small amplitudes
-among changing values of $E_{0}$ and $\theta^{-}$. The incoming
+(a), the inflow skewness $\kappa_3^{-}$ varies only slightly 
+with changing values of $E_{0}$ and $\theta^{-}$. The incoming
 flow PDF then can be determined by picking the proper parameter value
@@
 with $D_{+}=0.24$ shown in the right panel of Figure \ref{fig:matching}
-(a) generates much larger skewness $\kappa_3^{+}$ with highly skewed
-PDFs as the absolute value of $\theta^{+}$ and the total energy level
-$E_{0}$ increase in amplitude. The solid lines in Figure \ref{fig:matching}
+(a) generates much larger skewness $\kappa_3^{+}$ 
+as the absolute value of $\theta^{+}$ and the total energy level
+$E_{0}$ increases. The solid lines in Figure \ref{fig:matching}
 (c) offer a further confirmation of the transition from near-Gaussian
-statistics with tiny $\kappa_3^{-}$ to strongly skewed distribution
+statistics with small $\kappa_3^{-}$ to a strongly skewed distribution
 $\kappa_3^{+}$ after the depth change.
@@
 confirm each other.
+\todo[inline]{Again, do we mean statistical `mechanism' or `mechanics'?}
 
@@
 We have developed a statistical dynamical model to explain and predict
-extreme events and anomalous features in shallow water waves with
+extreme events and anomalous features of shallow water waves crossing an
 abrupt depth change. The theory is based on the dynamical modeling

### 2018-11-25 paper1_tKdV/SI_appendix/SI_tKdV.tex
DQ

@@
 system [\ref{eq:tKdV}] satisfies the Liouville property \cite{abramov2003hamiltonian,majda2006nonlinear},
-thus equilibrium statistical mechanism can be constructed based on
+thus equilibrium statistical mechanics can be constructed based on
 the conserved quantities.
@@
 \item The outgoing flow statistics can be predicted by both the equilibrium
-statistical mechanism and the direction model simulations. 
+statistical mechanics and the direction model simulations. 
 \begin{enumerate}
-\item Equilibrium statistical mechanism: find the optimal outgoing flow
+\item Equilibrium statistical mechanics: find the optimal outgoing flow
 parameter $\theta_{*}^{+}$ from matching the expectation in [\ref{eq:matching}]
@@
 Above we proposed two approaches to predict the statistics in the
-downstream wave disturbance. The equilibrium statistical mechanism
+downstream wave disturbance. The equilibrium statistical mechanics
 is an easy way to confirm the matching condition and get the equilibrium
@@
 We compare the statistical spectra from the direct Monte Carlo model
-simulation with the prediction from the equilibrium statistical mechanism
+simulation with the prediction from the equilibrium statistical mechanics
 in Figure \ref{fig:Comparison-spec }. The flow energy spectra computed
 both from the direct TKdV model simulations and from the Gibbs invariant
-measure in the equilibrium statistical mechanism are compared. Starting
+measure in the equilibrium statistical mechanics are compared. Starting
 from two different $\theta^{-}$ initial states, the direct model
@@
 model and the energy spectra predicted from the
-Gibbs invariant measure in equilibrium statistical mechanism. The
+Gibbs invariant measure in equilibrium statistical mechanics. The
 model simulations start with the Gibbs measure with different

### 2018-11-25 paper1_tKdV/tKdV_theory.tex
DQ

@@
 be useful for predicting extreme anomalous statistical behavior in
-other dispersive media in different settings.
+other dispersive media in different settings  \cite{solli2007optical,hohmann2010freak}.
 
-\todo[inline]{If you like, we could cite papers on `Optical rogue waves' and/or rogue waves in microwave systems here.}
+\todo[inline]{If you like, we could cite papers on `Optical rogue waves' and/or rogue waves in microwave systems here. (references added)}
 
@@
 larger driving amplitudes $\Delta\theta$ generate stronger skewness in the surface displacement PDFs.
-\item The waves also show different characteristic peak wavelengths in incoming
-and outgoing flows.
+\item \textcolor{blue}{The waves also show different slopes in the power spectra at upstream and downstream locations}.
 
@@
 I could easily come up with one.
-Perhaps a comment on the decay rate of the power spectra.}
+Perhaps a comment on the decay rate of the power spectra.
+(I will confirm with Andy whether he would just delete this item or keep something else like the one above)}
 
@@
 
-\todo[inline]{I modified the last sentence above. Feel free to revert to your original sentence if you do not like mine.}
+\todo[inline]{I modified the last sentence above. Feel free to revert to your original sentence if you do not like mine. (the change looks good)}
 
@@
 \end{equation}
-where $M$ defines the computational domain size $L_{d}=M\lambda_{c}$
+where $M$ defines the computational domain size $M\lambda_{c}$
 as $M$-multiple of the characteristic wavelength $\lambda_{c}$,
@@
 
-\todo[inline]{Is there a difference between $L_0$ and $L_d$? typo?}
+\todo[inline]{Is there a difference between $L_0$ and $L_d$?
+typo?
+(These two are actually different.
+$L_d$ is the `true' length scale defined in SI eqn.
+[10].
+I guess here it's better to delete this notation to avoid confusion.)}
 
@@
 
-\section{Equilibrium statistical mechanism for generating the stationary invariant measure}
+\section{Equilibrium statistical \textcolor{blue}{mechanics} for generating the stationary invariant measure}
 
-\todo[inline]{In this section title, did you mean `statistical mechanism' or `statistical mechanics'?}
+\todo[inline]{In this section title, did you mean `statistical mechanism' or `statistical mechanics'? (typos corrected)}
 
@@
 should also be matched at the depth change location $T_{\mathrm{ADC}}$,
-so that
+so that,
 \[
-\mu_{\infty}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{+}\left(u_{\Lambda}\right).
+\mu_{t=T_{\mathrm{ADC}}}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{+}\left(u_{\Lambda}\right).
 \]
-\todo[inline]{I do not believe we have defined $\mu_{\infty}^{-}$ anywhere.}
+\todo[inline]{I do not believe we have defined $\mu_{\infty}^{-}$ anywhere. (removed to avoid confusion)}
 In matching the flow statistics before and after the abrupt depth
@@
 from the direct numerical simulations here agree with the equilibrium
-statistical mechanism prediction illustrated in Figure \ref{fig:matching}. The prediction from these two different approaches
+statistical mechanics prediction illustrated in Figure \ref{fig:matching}. The prediction from these two different approaches
 confirm each other.
-\todo[inline]{Again, do we mean statistical `mechanism' or `mechanics'?}
+\todo[inline]{Again, do we mean statistical `mechanism' or `mechanics'? (should be `mechanics')}
 
@@
 more energetic high frequencies.
-The peak frequency illustrates the occurrence of the transporting
-waves along the water tank.
+\textcolor{blue}{The peak frequency illustrates the characteristic time scale of the transporting
+wave trains along the water tank}.
 

### 2018-11-26 paper1_tKdV/tKdV_theory.tex
NM

@@
 
-\todo[inline]{Is there a difference between $L_0$ and $L_d$?
-typo?
-(These two are actually different.
-$L_d$ is the `true' length scale defined in SI eqn.
-[10].
-I guess here it's better to delete this notation to avoid confusion.)}
-
 Consider the spatial discretization $J=2\Lambda+1$ so that the smallest
@@
 
-\section{Equilibrium statistical \textcolor{blue}{mechanics} for generating the stationary invariant measure}
-
-\todo[inline]{In this section title, did you mean `statistical mechanism' or `statistical mechanics'? (typos corrected)}
+\section{Equilibrium statistical mechanics for generating the stationary invariant measure}
 
@@
 \]
-\todo[inline]{I do not believe we have defined $\mu_{\infty}^{-}$ anywhere. (removed to avoid confusion)}
 In matching the flow statistics before and after the abrupt depth
@@
 confirm each other.
-\todo[inline]{Again, do we mean statistical `mechanism' or `mechanics'? (should be `mechanics')}
 

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
DQ_fixFigures
NM

@@
 is a Hamiltonian system which induces incoming and outgoing statistical
-Gibbs invariant measures which are statistically matched at the ADC.
+Gibbs invariant measures.
 The statistical matching of the known nearly Gaussian incoming Gibbs
@@
 outgoing skewness. The strategy here should be useful for predicting
-extreme anomalous statistical behavior in other dispersive media in
-different settings.
+extreme anomalous statistical behavior in other dispersive media.
 \end{abstract}
@@
 ADC. The TKdV equation is a Hamiltonian system which induces incoming
-and outgoing statistical Gibbs invariant measures which are statistically
-matched at the ADC. The statistical matching of the known nearly Gaussian
+and outgoing Gibbs invariant measures. The statistical matching of the known nearly Gaussian
 incoming Gibbs state at the ADC completely determines the predicted
@@
 other dispersive media in different settings  \cite{solli2007optical,hohmann2010freak}.
-
-\todo[inline]{If you like, we could cite papers on `Optical rogue waves' and/or rogue waves in microwave systems here. (references added)}
 
@@
 In these experiments, nearly unidirectional waves are generated by a paddle wheel and propagate through a long, narrow wave tank.
-Midway through, the waves encounter a step in the bottom topography, and thus abruptly transition from one depth to another.
-The paddle wheel is forced with angle
+Midway through, the waves encounter a step in the bottom topography, and thus abruptly transition to a shallower depth.
+The paddle wheel is forced with a pseudo-random signal intended to mimic a Gaussian random sea.
+In particular, the paddle angle is specified as
+%The angle of the paddle wheel $\theta(t)$ is prescribed as a pseudo-random signal that mimics a Gaussian random sea 
 \[
@@
 \centering
-\includegraphics[scale=1.]{./fig1}
+\subfloat[skewness from the Gibbs measure]{\includegraphics[scale=1.]{./fig1_1}}
+\vspace{-1em}
+\subfloat[matching for downstream $\theta^{+}$]{\includegraphics[scale=1.]{./fig1_2}}
+\vspace{-1em}
+\subfloat[prediction of downstream skewness]{\includegraphics[scale=1.]{./fig1_3}}
 

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

@@
 Midway through, the waves encounter a step in the bottom topography, and thus abruptly transition to a shallower depth.
-The paddle wheel is forced with a pseudo-random signal intended to mimic a Gaussian random sea.
+The paddle wheel is forced with a pseudo-random signal intended to mimic a Gaussian random sea upstream of the ADC.
 In particular, the paddle angle is specified as
-%The angle of the paddle wheel $\theta(t)$ is prescribed as a pseudo-random signal that mimics a Gaussian random sea 
 \[
@@
 \]
-where the weights $a_n$ are Gaussian in spectral space.
-The peak frequency $\omega_p$ gives rise to a characteristic wavelength $\lambda_c$ of water waves, which can be estimated using the dispersion relation.
-Above, the total energy $E$ in the system is determined by the angle amplitude
-$\Delta\theta$. Optical measurements of the free surface reveal a number of surprising statistical features:
+where the weights $a_n$ are Gaussian in spectral space with peak frequency $\omega_p$ and the phases $\delta_n$ are uniformly distributed random variables.
+The peak frequency gives rise to a characteristic wavelength $\lambda_c$ which can be estimated from the dispersion relation.
+The energy $E$ injected into the system is determined by the angle amplitude $\Delta\theta$, which is the main control parameter varied in \cite{bolles2018anomalous}.
+Optical measurements of the free surface reveal a number of surprising statistical features:
 \begin{itemize}
@@
 while outgoing waves skew strongly towards positive displacement.
-\item The degree of non-Gaussianity present in the outgoing wave field depends on the total energy of the system: 
-larger driving amplitudes $\Delta\theta$ generate stronger skewness in the surface displacement PDFs.
-\item \textcolor{blue}{The waves also show different slopes in the power spectra at upstream and downstream locations}.
-
-\todo[inline]{I am not sure we need to include this last item.
-It is not really an experimental observation, simply a calculation that can be done using the dispersion relation and the values of the two depths.
-Could we simply take this last item out?
-Would you prefer to replace it with another observation?
-I could easily come up with one.
-Perhaps a comment on the decay rate of the power spectra.
-(I will confirm with Andy whether he would just delete this item or keep something else like the one above)}
-
+\item The degree of non-Gaussianity present in the outgoing waves depends on the injected energy $E$: 
+larger energies generate stronger skewness in the surface displacement PDFs and more extreme events.
+\item \textcolor{blue} {Compared to the incoming wave train, the power spectrum of the outgoing wave field decays more slowly, which indicates that the anomalous behavior is associated with an elevated level of high frequencies.}
+%{The waves also show different slopes in the power spectra at upstream and downstream locations}.
 \end{itemize}

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

@@
 
-The surface wave turbulence is modeled by a one-dimensional deterministic
-dynamical model. 
-\todo[inline]{This first sentence seems awkward to me.}
 The Korteweg-de Vries (KdV) equation \cite{johnson1997modern}
-is a leading-order approximation of the surface waves that are determined
+is a leading-order approximation of surface waves that are determined
 by the balance of nonlinear and dispersive effects in an appropriate
@@
 \begin{equation}
-\frac{\partial u_{\Lambda}^{\pm}}{\partial t}+\frac{D_{\pm}^{-3/2}}{2}E_{0}^{1/2}L_{0}^{-3/2}\frac{\partial}{\partial x}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{\pm}\right)^{2}+D_{\pm}^{1/2}L_{0}^{-3}\frac{\partial^{3}u_{\Lambda}^{\pm}}{\partial x^{3}}=0,\label{eq:tKdV_nondim}
+\frac{\partial u_{\Lambda}^{\pm}}{\partial t}+\frac{D_{\pm}^{-3/2}}{2}E_{0}^{1/2}L_{0}^{-3/2}\frac{\partial}{\partial x}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{\pm}\right)^{2}+D_{\pm}^{1/2}L_{0}^{-3}\frac{\partial^{3}u_{\Lambda}^{\pm}}{\partial x^{3}}=0.
+\label{eq:tKdV_nondim}
 \end{equation}
-on the normalized periodic domain $x\in\left[-\pi,\pi\right]$. The conserved Hamiltonian can be decomposed into a cubic and a quadratic term
+Equation [\ref{eq:tKdV_nondim}] is non-dimensionalized on the periodic
+domain $x\in\left[-\pi,\pi\right]$. The depth is assumed to be unit $D_{-}=1$ before the ADC and becomes $D_{+}<1$ after the ADC. 
+The conserved Hamiltonian can be decomposed into a cubic and a quadratic term
 \begin{align*}
@@
 \end{align*}
-Equation [\ref{eq:tKdV_nondim}] is non-dimensionalized in the periodic
-domain. The depth is assumed to be unit $D_{-}=1$
-before the ADC and becomes $D_{+}<1$
-after the ADC. We introduce parameters $\left(E_{0},L_{0},\Lambda\right)$
+We introduce parameters $\left(E_{0},L_{0},\Lambda\right)$
 based on the following assumptions:
 \begin{itemize}
-\item The wavenumber truncation $\Lambda$ is fixed in a moderate value
+\item The wavenumber truncation $\Lambda$ is fixed at a moderate value
 for generating weakly turbulent dynamics;
@@
 which are conserved during evolution. Meanwhile, $E_{0}$ characterizes the total
-energy injected into the system based on the paddle amplitude $\left(\Delta\theta\right)^{2}$;
-\item The length scale of the system is denoted $L_{0}$. The value is
-chosen so that the resolved scale $\Delta x=2\pi L_{0}/J$ is comparable
-to the the characteristic wave length $\lambda_{c}$ in
-the experiments.
+energy injected into the system based on the driving amplitude $\Delta \theta$;
+\item The length scale of the system $L_{0}$ is chosen so that the resolved scale $\Delta x=2\pi L_{0}/J$ is comparable to the the characteristic wave length $\lambda_{c}$ from the experiments.
 \end{itemize}
@@
 %for $H_{2}$ in the Hamiltonian. 
-Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the balance of the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
+Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the relative contributions of the cubic and quadratic terms to the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
 The depth change, $D_{+}<1$, increases the weight of $H_3$ and decreases that of $H_2$, thus promoting the effects of nonlinearity and reducing dispersion.
@@
 scale changes in rapid time. 
-Rather, since we are interested in modeling statistics before and after the depth change, we will examine the long-time dynamics of large-scale structures.
-
-\todo[inline]{I modified the last sentence above. Feel free to revert to your original sentence if you do not like mine. (the change looks good)}
+Rather, since we are interested in modeling statistics before and after the ADC, we will examine the long-time dynamics of large-scale structures.
 

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

@@
 
-The Korteweg-de Vries (KdV) equation \cite{johnson1997modern}
-is a leading-order approximation of surface waves that are determined
-by the balance of nonlinear and dispersive effects in an appropriate
-far-field limit. Here, the KdV equation is truncated in the first
-$\Lambda$ modes (with $J=2\Lambda+1$ grid points) to generate weakly
-turbulent dynamics. The surface displacement is described by
-the state variable $u_{\Lambda}^{\pm}\left(x,t\right)$ with superscript `$-$'
-for the incoming waves and `$+$' for the outgoing waves. The Galerkin
-truncated variable $u_{\Lambda}=\sum_{1\leq\left|k\right|\leq\Lambda}\hat{u}_{k}\left(t\right)e^{ikx}$
-is normalized with zero mean $\hat{u}_{0}=0$ and unit energy $2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1$,
-which are conserved quantities, and $u_{\Lambda}\equiv\mathcal{P}_{\Lambda}u$
-denotes the subspace projection. The evolution of  $u_{\Lambda}^{\pm}$ is governed by the truncated
-KdV equation with depth change $D_{\pm}$
+The Korteweg-de Vries (KdV) equation is a one-dimensional, deterministic model capable of describing {\color{blue} (weak?)} surface wave turbulence.
+More specifically, KdV is leading-order approximation for surface waves governed by a balance of nonlinear and dispersive effects, valid in an appropriate far-field limit \cite{johnson1997modern}.
+Moreover, KdV has been adapted to describe waves propagating over variable depth  \cite{johnson1997modern}.
+Here, we consider the variable-depth KdV equation truncated at wavenumber $\Lambda$ (with $J=2\Lambda+1$ grid points) in order to generate weakly turbulent dynamics.
+The surface displacement is described by the state variable $u_{\Lambda}^{\pm}\left(x,t\right)$ with superscript `$-$' for the incoming waves and `$+$' for the outgoing waves.
+The Galerkin truncated variable $u_{\Lambda}=\sum_{1\leq\left|k\right|\leq\Lambda}\hat{u}_{k}\left(t\right)e^{ikx}$ is normalized with zero mean $\hat{u}_{0}=0$ and unit energy $2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1$, which are conserved quantities.
+Here, $u_{\Lambda}\equiv\mathcal{P}_{\Lambda}u$ denotes the subspace projection.
+The evolution of  $u_{\Lambda}^{\pm}$ is governed by the truncated KdV equation with depth change $D_{\pm}$
 \begin{equation}
@@
 Equation [\ref{eq:tKdV_nondim}] is non-dimensionalized on the periodic
-domain $x\in\left[-\pi,\pi\right]$. The depth is assumed to be unit $D_{-}=1$ before the ADC and becomes $D_{+}<1$ after the ADC. 
-The conserved Hamiltonian can be decomposed into a cubic and a quadratic term
+domain $x\in\left[-\pi,\pi\right]$. The depth is assumed to be unit $D_{-}=1$ before the ADC and $D_{+}<1$ after the ADC. 
+The conserved Hamiltonian can be decomposed as
 \begin{align*}
@@
 \end{align*}
+where we refer to $H_3$ as the cubic term and $H_2$ the quadratic term.
 We introduce parameters $\left(E_{0},L_{0},\Lambda\right)$

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

@@
 \item The wavenumber truncation $\Lambda$ is fixed at a moderate value
-for generating weakly turbulent dynamics;
+for generating weakly turbulent dynamics.
 \item The state variable $u_{\Lambda}^{\pm}$ is normalized with zero mean
@@
 which are conserved during evolution. Meanwhile, $E_{0}$ characterizes the total
-energy injected into the system based on the driving amplitude $\Delta \theta$;
+energy injected into the system based on the driving amplitude $\Delta \theta$.
 \item The length scale of the system $L_{0}$ is chosen so that the resolved scale $\Delta x=2\pi L_{0}/J$ is comparable to the the characteristic wave length $\lambda_{c}$ from the experiments.
@@
 %for $H_{2}$ in the Hamiltonian. 
-Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the relative contributions of the cubic and quadratic terms to the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
-The depth change, $D_{+}<1$, increases the weight of $H_3$ and decreases that of $H_2$, thus promoting the effects of nonlinearity and reducing dispersion.
+Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the relative contributions of $H_3$ and $H_2$ in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
+The depth change, $D_{+}<1$, increases the weight of $H_3$ and decreases that of $H_2$, thereby promoting the effects of nonlinearity over dispersion and creating conditions favorable for extreme events.
 Since $\frac{\partial u}{\partial x}$ is the slope of the wave height, $H_{2}\left(u\right)$ measures the

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

@@
 distribution functions.
-\item The experiments also have the up and down stream power spectra in
-time (Fig. 4 of \cite{bolles2018anomalous}), which shows more energy
-at small time scales, i.e., a relatively slower decay rate in the
-downstream compared with the upstream case. This is also observed
+\item Experimental measurements of the power spectra (Fig.
+4 of \cite{bolles2018anomalous}) reveal the downstream measurements to contain more energy at small scales, i.e.
+a relatively slower decay rate of the spectrum.
+This result is also observed
 in the direct numerical simulations here (detailed results shown
-in \textcolor{blue}{\emph{SI Appendix, C.2}}). The downstream state contains
+in \textcolor{blue}{\emph{SI Appendix, C.2}}), as the outgoing state contains
 more energetic high frequencies.
-\textcolor{blue}{The peak frequency illustrates the characteristic time scale of the transporting
-wave trains along the water tank}.
-
-\todo[inline]{I do not understand the last sentence above. Why does the peak frequency illustrate the occurrence of transporting water waves?}
 

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

@@
 in complex nonlinear systems is a grand challenge in climate, material,
-and neuroscience, as well as for engineering design. Recent controlled
-laboratory experiments in weakly turbulent shallow water with abrupt
-depth change (ADC) exhibit a remarkable transition from nearly Gaussian
-statistics in incoming wave trains before the ADC to outgoing waves
-trains after the ADC with extreme anomalous statistics with large
-positive skewness of the surface height. Here we develop a statistical
+and neuroscience, as well as for engineering design. 
+Recent laboratory experiments in weakly turbulent shallow water reveal a remarkable transition from Gaussian to anomalous behavior as surface waves cross an abrupt depth change (ADC).
+Downstream of the ADC, PDFs of  surface displacement exhibit strong positive skewness, which is accompanied by an elevated level of extreme events.
+Here we develop a statistical
 dynamical model to explain and quantitatively predict the above anomalous

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

@@
 Recent laboratory experiments in weakly turbulent shallow water reveal a remarkable transition from Gaussian to anomalous behavior as surface waves cross an abrupt depth change (ADC).
-Downstream of the ADC, PDFs of  surface displacement exhibit strong positive skewness, which is accompanied by an elevated level of extreme events.
+Downstream of the ADC, PDFs of  surface displacement exhibit strong positive skewness, accompanied by an elevated level of extreme events.
 Here we develop a statistical
@@
 \end{abstract}
+
+% ORIGINAL SENTENCE IN ABSTRACT
+%Recent controlled laboratory experiments in weakly turbulent shallow water with abrupt depth change (ADC) exhibit a remarkable transition from nearly Gaussian statistics in incoming wave trains before the ADC to outgoing waves trains after the ADC with extreme anomalous statistics with large positive skewness of the surface height.
 
@@
 
+\cite{bolles2018anomalous}
+
 The model parameters $\left(E_{0},L_{0},\Lambda\right)$ in [\ref{eq:tKdV_nondim}]
 can be directly linked to the basic scales from the physical problem.
-The important characterizing parameters measured from the experiments
-include: $\epsilon=\frac{a}{H_{0}}$ the wave amplitude $a$ to water
+The important parameters that characterize the experiments
+include: $\epsilon=\frac{a}{H_{0}}$ the wave amplitude $a$ to (upstream) water
 depth $H_{0}$ ratio; $\delta=\frac{H_{0}}{\lambda_{c}}$ the water
-depth to wavelength scale $\lambda_{c}$ ratio; and $D_{0}=\frac{d}{H_{0}}$
-the normalized wave depth ratio with incoming flow depth $d=H_{0}$
-to the outgoing flow depth $d<H_{0}$. The interpretations and reference
-values of these model parameters are based on the experimental setup
-\cite{bolles2018anomalous}. By comparing the characteristic physical
-scales, the normalized TKdV equation [\ref{eq:tKdV_nondim}] can be
-linked directly with the measured non-dimensional quantities by
+depth to wavelength $\lambda_{c}$ ratio; and $D_{0}=\frac{d}{H_{0}}$ 
+the depth ratio with upstream value $d=H_{0}$
+and downstream value $d<H_{0}$. 
+By comparing the characteristic physical scales, the normalized TKdV equation [\ref{eq:tKdV_nondim}] can be
+linked directly to the above experimental parameters via
 \begin{equation}
@@
 as $M$-multiple of the characteristic wavelength $\lambda_{c}$,
-and $\gamma=\frac{U}{a}$ represents the factor to normalize the total
-energy in the state variable $u_{\Lambda}$ to one.
+and $\gamma=\frac{U}{a}$ normalizes the total
+energy of the state variable $u_{\Lambda}$ to one.
+\todo[inline]{What is the variable $U$? Have we defined it?}
 
-Consider the spatial discretization $J=2\Lambda+1$ so that the smallest
-resolved scale is comparable with the characteristic wavelength
+Consider a spatial discretization with $J=2\Lambda+1$ grid points, so that the smallest
+resolved scale is comparable to the characteristic wavelength
 \[
@@
 \]
-Therefore in the practical numerical simulations, we pick $M=5$ and
+In the practical numerical simulations, we select $M=5$ and
 $\gamma$ varies in the range $\left[0.5,1\right]$. Using the reference

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

@@
 
-\cite{bolles2018anomalous}
-
 The model parameters $\left(E_{0},L_{0},\Lambda\right)$ in [\ref{eq:tKdV_nondim}]
 can be directly linked to the basic scales from the physical problem.
-The important parameters that characterize the experiments
+The important parameters that characterize the experiments \cite{bolles2018anomalous}
 include: $\epsilon=\frac{a}{H_{0}}$ the wave amplitude $a$ to (upstream) water
@@
 By comparing the characteristic physical scales, the normalized TKdV equation [\ref{eq:tKdV_nondim}] can be
-linked directly to the above experimental parameters via
+linked to these experimental parameters via
 \begin{equation}
@@
 In the practical numerical simulations, we select $M=5$ and
-$\gamma$ varies in the range $\left[0.5,1\right]$. Using the reference
-experimental measurements \cite{bolles2018anomalous}, $\epsilon\in\left[0.0024,0.024\right],\delta\sim0.22$,
-and $D_{0}$ changes from 1 to 0.24 before and after the depth change.
-The reference values for the model scales can be estimated in the
-range $L_{0}\in\left[2,6\right]$ and $E_{0}\in\left[50,200\right]$.
+let $\gamma$ vary in the range $\left[0.5,1\right]$. Using reference values from the experiments, 
+\cite{bolles2018anomalous}, we find $\epsilon\in\left[0.0024,0.024\right], \delta\sim0.22$,
+and $D_{0}$ changes from 1 to 0.24 at the ADC.
+These values yield the following the following ranges for the model parameters: $L_{0}\in\left[2,6\right]$ and $E_{0}\in\left[50,200\right]$.
 These are the values we will test in the direct numerical simulations.
@@
 Since the TKdV equation satisfies the Liouville property, the equilibrium
-invariant measure can be described by an equilibrium statistical formulism
+invariant measure can be described by a statistical formalism
 \cite{abramov2003hamiltonian,majda2006nonlinear,bajars2013weakly}
-using a Gibbs measure with the conserved energy $\mathcal{E}_{\Lambda}$ and
+based on a Gibbs measure with the conserved energy $\mathcal{E}_{\Lambda}$ and
 Hamiltonian $\mathcal{H}_{\Lambda}$. The equilibrium invariant measure is dictated
-by the conservation laws in the TKdV equation. In the case with fixed
-total energy $E_{0}$, this is the \emph{mixed Gibbs measure} in the
-truncated model with microcanonical energy and canonical Hamiltonian
+by the conservation laws in the TKdV equation. In the case of fixed energy $E_{0}$, 
+this is the \emph{mixed Gibbs measure} with microcanonical energy and canonical Hamiltonian
 ensembles \cite{abramov2003hamiltonian}
@@
 \end{equation}
-with $\theta$ representing the ``inverse temperature''. The distinct
-statistics in the upstream and downstream waves can be controlled
-by the value of $\theta$. Negative temperature, $\theta^{\pm}<0$, is the appropriate regime to predict the
-experiments as shown below. In the incoming flow field, the inverse
-temperature $\theta^{-}$ is chosen so that $\mathcal{G}_{\theta}^{-}$
-has Gaussian statistics. Using the above invariant measures [\ref{eq:gibbs_mixed}],
-the expectation of any functional $F\left(u\right)$ can be computed
-based on the Gibbs measure 
+where $\theta$ represents the ``inverse temperature''. The distinct
+statistics in the upstream and downstream waves can be controlled $\theta$. 
+As shown below, we find that negative temperature, $\theta^{\pm}<0$, is the appropriate regime to predict the
+experiments. In the incoming wave field, $\theta^{-}$ is chosen so that $\mathcal{G}_{\theta}^{-}$
+has nearly Gaussian statistics. Using the above invariant measures [\ref{eq:gibbs_mixed}],
+the expectation of any functional $F\left(u\right)$ can be computed as
 \[
@@
 $\left\langle H_{\Lambda}\right\rangle _{\mathcal{G}_{\theta}}$ \cite{abramov2003hamiltonian,bajars2013weakly}.
-The invariant measure also predicts an equilibrium energy spectrum
-without running the TKdV equation directly. On the other hand, the
-time autocorrelation and transient statistics about the state variable
-$u_{\Lambda}$ cannot be recovered from the statistical theory.
+\todo[inline]{I do not quite understand the above sentence.}
+In addition to producing equilibrium PDFs of $u_{\Lambda}$, the invariant measure can be used to predict the equilibrium energy spectrum without the direct simulation of TKdV.
+Direct simulation, however, is required to recover transient statistics of $u_{\Lambda}$ and time autocorrelations.
+%The invariant measure also predicts an equilibrium energy spectrum without running the TKdV equation directly.
+On the other hand, the time autocorrelation and transient statistics about the state variable $u_{\Lambda}$ cannot be recovered from the statistical theory.
 
-\subsection{Statistical matching condition in invariant measures before and after the abrupt depth change}
+\subsection{Statistical matching condition of the invariant measures before and after the abrupt depth change}
 
@@
 the incoming and outgoing waves. The two distributions, $\mu_{t}^{-},\mu_{t}^{+}$
-should also be matched at the depth change location $T_{\mathrm{ADC}}$,
-so that,
+should be matched at $T_{\mathrm{ADC}}$ giving
 \[
@@
 
-The incoming flow is always characterized by a near-Gaussian distribution
-in the wave displacement. It is found that a physically consistent
+For the parameters explored in \cite{bolles2018anomalous}, the incoming wave field 
+is always characterized by a near-Gaussian distribution
+of the surface displacement. It is found that a physically consistent
 Gibbs measure should take negative values in the inverse temperature

### 2018-11-29 paper1_tKdV/tKdV_theory.tex
DQ_final

@@
 larger energies generate stronger skewness in the surface displacement PDFs and more extreme events.
-\item \textcolor{blue} {Compared to the incoming wave train, the power spectrum of the outgoing wave field decays more slowly, which indicates that the anomalous behavior is associated with an elevated level of high frequencies.}
+\item Compared to the incoming wave train, the power spectrum of the outgoing wave field decays more slowly, which indicates that the anomalous behavior is associated with an elevated level of high frequencies.
 %{The waves also show different slopes in the power spectra at upstream and downstream locations}.
@@
 
-The Korteweg-de Vries (KdV) equation is a one-dimensional, deterministic model capable of describing {\color{blue} (weak?)} surface wave turbulence.
+The Korteweg-de Vries (KdV) equation is a one-dimensional, deterministic model capable of describing surface wave turbulence.
 More specifically, KdV is leading-order approximation for surface waves governed by a balance of nonlinear and dispersive effects, valid in an appropriate far-field limit \cite{johnson1997modern}.
@@
 %for $H_{2}$ in the Hamiltonian. 
-Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the relative contributions of $H_3$ and $H_2$ in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
+Some intuition for how equation [\ref{eq:tKdV_nondim}] produces different dynamics on either side of the ADC can be gained by considering the relative contributions of $H_3$ and $H_2$ in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
 The depth change, $D_{+}<1$, increases the weight of $H_3$ and decreases that of $H_2$, thereby promoting the effects of nonlinearity over dispersion and creating conditions favorable for extreme events.
@@
 as $M$-multiple of the characteristic wavelength $\lambda_{c}$,
-and $\gamma=\frac{U}{a}$ normalizes the total
-energy of the state variable $u_{\Lambda}$ to one.
-\todo[inline]{What is the variable $U$? Have we defined it?}
+and $\gamma=\frac{U}{a}$ with $U$ a scaling factor for the state variable $u_{\Lambda}$ normalizes the total energy of the system to one.
 
@@
 $\left\langle H_{\Lambda}\right\rangle _{\mathcal{G}_{\theta}}$ \cite{abramov2003hamiltonian,bajars2013weakly}.
-\todo[inline]{I do not quite understand the above sentence.}
+
 In addition to producing equilibrium PDFs of $u_{\Lambda}$, the invariant measure can be used to predict the equilibrium energy spectrum without the direct simulation of TKdV.
@@
 Naval Research through MURI N00014-16-1-2161. D. Q. is supported as
-a postdoctoral fellow on the second grant. M.N.J.M would like to acknowledge support from Simons grant 524259.}
+a postdoctoral fellow on the grant. M.N.J.M would like to acknowledge support from Simons grant 524259.}
 

### 2018-11-29 paper1_tKdV/tKdV_theory.tex
NM

@@
 \affil[b]{Department of Mathematics and Geophysical Fluid
-Dynamics Institute, Florida State University, Tallahassee, FL}
+Dynamics Institute, Florida State University, Tallahassee, FL 32306}
 
@@
 and $D_{0}$ changes from 1 to 0.24 at the ADC.
-These values yield the following the following ranges for the model parameters: $L_{0}\in\left[2,6\right]$ and $E_{0}\in\left[50,200\right]$.
+These values yield the following ranges for the model parameters: $L_{0}\in\left[2,6\right]$ and $E_{0}\in\left[50,200\right]$.
 These are the values we will test in the direct numerical simulations.
@@
 is conserved in time after the depth change at $t=T_{\mathrm{ADC}}$
-and should stay in the same value as the system approaches equilibrium
+and should remain at this value as the system approaches equilibrium
 as $t\rightarrow\infty$. The final statistical matching condition

### 2019-05-05 LatexNotes/Notes.tex
NM

@@
 \end{equation}
+
+\subsection{The optimizer of the Hamiltonian is a traveling wave solution}
+For the acceptance-rejection algorithm presented below, it will be important to find the maximal value of the Hamiltonian subject to the constraints of fixed energy $E = E_0$ and zero momentum $M = 0$.
+For simplicity, consider the KdV equation
+\begin{equation}
+u_t + u u_x + \mu u_{xxx} = 0
+\end{equation}
+with corresponding Hamiltonian
+\begin{equation}
+H = \frac{1}{6} \intt u^3 dx - \frac{\mu}{2} \intt u_x^2 dx
+\end{equation}
+
+Consider the functional
+\begin{equation}
+F[u] = H[u] + \lambda_1 (E-E_0) + \lambda_2 (M-0)
+\end{equation}
+where $\lambda_1$ and $\lambda_2$ are Lagrangian multipliers. Taking the variational derivative gives
+\begin{equation}
+\frac{\delta F}{\delta u} = \frac{1}{2} u^2 + \mu u_{xx} + \lambda_1 u + \lambda_2 = 0
+\end{equation}
+which is a second-order nonlinear ODE for the optimizer $u = U(x)$. Note that this is identical to the condition for a traveling wave solution!!!
 

### 2019-06-12 WavesData-2019/LatexNotes/notes.tex
Added new notes

--- 2018-09-29 LatexNotes/Notes.tex
@@
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
+%\documentclass[preprint, 10pt]{elsarticle}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%----------------------------------------------%
 %% LATEX DEFINITIONS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%----------------------------------------------%
 
@@
 
+% New
+\newcommand{\Hp}{H^{+}}
+\newcommand{\Hm}{H^{-}}
+\newcommand{\Gp}{\Gibbs^{+}}
+\newcommand{\Gm}{\Gibbs^{-}}
+
 % Dashed integral
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%----------------------------------------------%
 %% TITLE
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%----------------------------------------------%
 \begin{document}
-\title{Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change}
-\author{}
+%Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change
+\title{The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes}
+
+%\author{
+% Tyler Bolles
+%Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
+%M.~N.~J.~Moore\thanks{Florida State University}, 
+%Di Qi\thanks{Courant Institute of Mathematical Sciences} }
 \maketitle
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Experimental Preliminaries
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experimental parameters}
+%----------------------------------------------------------%
+% Mathematical framework
+%----------------------------------------------------------%
+\section{The truncated KdV framework}
 
-Some relevant parameters from the experiment are given in the table.
-
-% Table
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{table}%[htbp]
-\begin{center}
-\caption{Basic experimental parameters and scales.}
-\label{exptable}
-\begin{tabular}{l l l}
-\hline Parameter & Upstream  value \,\, & Downstream value \\
-\hline
-peak forcing frequency, $f_p$\,\,		& 2 Hz			& N/A \\
-depth, $h$						& 12.5 cm			& 3 cm \\
-peak wavelength, $\lambda_p$		& 38 cm			& 25 cm \\
-peak wavenumber, $k_p$			& 0.17 rad/cm		& 0.25 rad/cm \\
-dimensionless depth, $k_p h$\,\,\,		& 2.1			& 0.76 \\
-\end{tabular}
-\end{center}
-\end{table}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% KdV
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The KdV equations}
-
-\subsection{The constant-depth KdV equation in dimensional form}
-
-The constant-depth KdV equation in dimensional form is
+\subsection{KdV with depth change}
+In dimensional form the Korteweg–de Vries equation is given by
 \begin{equation}
@@
 \end{equation}
-Here, $\eta$ has units of length, $t$ time, $h_0$ is the depth, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$.
-I obtained this equation by starting from Johnson's dimensionless equation (3.28) on page 208 and working backwards to untangle the scales that he had used to non-dimensionalize.
-Alternatively, when the first term is expressed as a spatial derivative, the dimensional KdV equation is
-\begin{equation}
-2 \eta_x + \frac{3}{h_0} \eta \eta_{\xi} + \frac{h_0^2}{3} \eta_{\xi \xi \xi} = 0
-\end{equation}
-Notice the only difference is the absence of the wavespeed $c$ and the first term is now a spatial derivative.
 
-The important thing to remember is that Johnson uses these characteristic scales for non-dimensionalizing (characteristic scales denoted by a *)
+\subsection{Nondimensionalization}
+
+The dimensionless KdV equation is
 \begin{align}
-\eta^* = \eps h_0 , \quad
-x^* = \eps^{-1/2} h_0 , \quad
-t^* = \eps^{-1/2} h_0/c
+\label{varKdV}
+u_t + \ampp D^{-3/2} \, u u_x + \mu D^{1/2} \, u_{xxx} = 0
 \end{align}
-where $h_0$ is a characteristic depth (say the upstream depth), $c = \sqrt{g h_0}$ the corresponding wave speed, and $\eps = a / h_0$ is the ratio of wave amplitude to depth.
-We want to alter Johnson's non-dimensionalization so that $x$ is instead scaled on a characteristic wavelength (and $t$ scaled accordingly).
-Importantly, we need to account for variable depth.
+with the dimensionless parameters
+\begin{align}
+\ampp = \eps \delta^{-2} = \frac{ag}{h_0^2 f_p^2} , \qquad
+\mu = \frac{4 \pi^2}{9 \lamfac^2}
+\end{align}
+Here, $\lamfac$ is the number of peak wavelengths in the periodic domain (perhaps we should set $\lamfac=1$, see note below).
+Above, the displacement $u$ has been scaled on the characteristic wave amplitude $a = \mean{\eta^2}^{1/2}$, and position $x$ has been scaled on the upstream characteristic wavelength $\lambda_0 = \sqrt{g h_0}/f_p$.
+Then the Hamiltonian is given by
+\begin{align}
+\label{Hamiltonian}
+& \Hm = \ampp H_3 - \mu H_2 \\
+& \Hp = \ampp D_0^{-3/2} H_3 - \mu D_0^{1/2} H_2
+\end{align}
+where
+\begin{align}
+&H_3 = \intt u^3 \dx	\, , \\
+&H_2 = \intt u_x^2 \dx	\, .
+\end{align}
 
-\subsection{The variable-depth KdV equation}
+Note: after giving it some deliberation, I believe we should use $\lamfac=1$.
+My reason is that the peak in the wave spectrum occurs at a frequency of $f_p$ or wavelength of $\lambda_p$.
+In the tKdV Gibbs measure, the spectrum decays monotonically, so that the peak is at the lowest resolved frequency (or largest resolved wavelength).
+Thus, we want $\lambda_p$ to correspond to the largest resolved wavelength, i.e.~the length of the periodic domain in the tKdV framework.
 
-We start with the variable-depth KdV equation given by Johnson in Eq.(3.149) on page 272 in dimensionless form 
-\begin{equation}
-\label{varKdVJohnson}
-2 H_X + \frac{3}{D^{7/4}} H H_\xi + \frac{1}{3} D^{1/2} H_{\xi \xi \xi} = 0
-\end{equation}
-All variables above are dimensionless, and by working backwards we can figure out how they relate to the original, dimensional variables.
-Below, the dimensional quantities are indicated with an overbar wherever there is the possibility of ambiguity
-\begin{align}
-& H = D^{1/4} \eta = \left( \frac{d}{h_0} \right)^{1/4} \frac{\bar{\eta}}{\eps h_0} \\
-& X = \eps x = \frac{\eps^{3/2}}{h_0} \bar{x} \\
-& \xi = \frac{\eps^{1/2}}{\sqrt{d h_0}} (\bar{x} - c \bar{t})
-\end{align}
-Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the depth ratio.
-The last relationship above for $\xi$ is valid for the case of two depths (though perhaps not in the general case).
-
-There are a few important things to remember when we reformulate this equation:
-\begin{enumerate}
-\item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
-\item We want to normalize the standard deviation (or energy) of $\eta$ and therefore want to keep Johnson's choice to scale surface displacement on $a = \eps h_0$.
-\item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change.
-We have to do this so that the assumption of spatial periodicity will be reasonably satisfied on both sides.
-\end{enumerate}
-To achieve 3, we can use the characteristic wavelength from shallow water theory,
-\begin{equation}
-\lambda_c = {c}/{f_c} = \sqrt{g d}/f_c
-\end{equation}
-where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed.
-It is tempting to simply set $f_c$ to the peak forcing frequency $f_p$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$.
-Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments).
-We therefore choose
-\begin{equation}
-\lambda_c = \lamfac \lambda_p = \lamfac \sqrt{gd} / f_p
-\end{equation}
-where $\lamfac$ is an integer (on the order of 10 or so).
-Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient.
-Thus, the main change we need to make is to introduce the dimensionless variable
-\begin{equation}
-\zeta = \frac{2\pi}{\lambda_c} (\bar{x} - c \bar{t}) 
-= \frac{2 \pi \delta}{\lamfac \eps^{1/2}} \, \xi
-\end{equation}
-where
-\begin{equation}
-\delta = h_0 / \lambda_0 = f_p \sqrt{h_0/g}
-\end{equation}
-To clean things up a little, lets also introduce a slightly different long-space coordinate that uses the upstream wavelength $\lambda_0$ as a characteristic scale instead of $h_0$,
-\begin{equation}
-s = \frac{2 \pi \beta}{\lamfac \lambda_0} \bar{x} = 2 \pi \beta \lamfac^{-1} \delta \eps^{-3/2} X
-\end{equation}
-where $\beta$ is a small parameter that scale like $\eps$ or $\delta^2$ for example.
-
-Then Eq.~\eqref{varKdVJohnson} becomes
-\begin{equation}
-\label{varKdVugly}
-2 \beta H_s + 3 \eps D^{-7/4} H H_\zeta
-+ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} H_{\zeta \zeta \zeta} = 0
-\end{equation}
-
-\subsection{Final formulation of variable-depth KdV}
-
-Equation \eqref{varKdVugly} is essentially in the form we want to use, except that the notation is bad.
-In particular, we will want to reserve the symbol $H$ for the Hamiltonian.
-So lets introduce some new variables
-\begin{align}
-& u = D^{1/4} \, \bar{\eta}/a \\
-& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
-& s = 2 \pi \beta \lamfac^{-1} \lambda_0^{-1} \bar{x}
-\end{align}
-Above, the original physical variables are denoted with a bar. Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.
+%Remark: We note that another possibility is to match the quantity $D^{1/4} \eta$ at the abrupt depth change as would be consistent with the analysis of Johnson.
+If we were to do this, we would obtain only slightly different powers of $D$ in the coefficients: $D^{-7/4}$ and $D^{1/2}$ respectively.
 
@@
 
-In Table \ref{paramtable} we provide a list of all the relevant parameters needed to convert these dimensional variables to physical ones.
-Note that the wavelength $\lambda_0$ is predicted from shallow-water theory and so it differs somewhat from the value in Table \ref{exptable} obtained from the dispersion relation.
-In terms of the new variables,  Eq.~\eqref{varKdVugly} becomes
-\begin{align}
-\label{varKdV1}
-2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
-+ \frac{4 \pi^2 \delta^2}{3 \lamfac^2} D^{1/2} \, u_{xxx} = 0
-\end{align}
+\subsection{Matching condition}
 
-Let's clean it up now a little bit. Let us set $\beta = 3 \delta^2/2$ and introduce the parameters
-\begin{align}
-\ampp = \eps \delta^{-2} = \frac{ag}{h_0^2 f_p^2} , \qquad
-\mu = \frac{4 \pi^2}{9 \lamfac^2}
-\end{align}
-The Eq.~\eqref{varKdV1} becomes
-\begin{align}
-\label{varKdV}
-u_s + \ampp D^{-7/4} \, u u_x
-+ \mu D^{1/2} \, u_{xxx} = 0
-\end{align}
-
-\subsection{Hamiltonian structure}
-Then the Hamiltonian structure is given by
-\begin{align}
-\label{Hamiltonian}
-& H^- = \ampp H_3 - \mu H_2 \\
-& H^+ = \ampp D_0^{-7/4} H_3 - \mu D_0^{1/2} H_2
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Mathematical Preliminaries
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Mathematical preliminaries}
-
-\subsection{Truncated Fourier series}
-
-We will be dealing with the truncated KdV equation (TKDV). 
-Consider a real-valued function $u(x): [0, 2\pi] \to \RR$ given by a truncated Fourier series
+The statistical matching condition is
 \begin{equation}
-u(x) = \sum_{k= -\Lambda}^{\Lambda} \uhat_k e^{i k x}	\qquad \text{for } 0 \le x \le 2\pi
-\end{equation}
-where
-\begin{equation}
-\uhat_k = \intt u(x) e^{-i k x} \dx
-\end{equation}
-Above, and throughout the paper, we use the notation
-\begin{equation}
-\intt \cdot \, \dx  = \frac{1}{2\pi} \int_0^{2 \pi} \cdot \, \dx
-\end{equation}
-We will be dealing with zero-momentum functions
-\begin{equation}
-M = \intt u(x) \dx = 0
-\end{equation}
-Let $\Fspace$ denote the space of all zero-momentum functions that can be represented by the above truncated Fourier Series (with maximum wavenumber $\Lambda$).
-
-Since any element $u \in \Fspace$ is real valued and has zero momentum, it can be written as
-\begin{equation}
-u(x) = \sumk \uhat_k e^{i k x} + \uhat_k^{*} e^{-i k x} 
-= 2 \Real \sumk \uhat_k e^{i k x}
-\end{equation}
-This implies that $\dim(\Fspace) = \Lambda$ as a complex vector space (or the dimension of the equivalent real vector space is $2\Lambda$).
-For any $u \in \Fspace$, we introduce the energy
-\begin{equation}
-E = \frac{1}{2} \intt \Proj(u^2) \dx = \sumk \abs{\uhat_k}^2
-\end{equation}
-where $\Proj$ is the projection onto $\Fspace$.
-In order to define the Hamiltonian for TKDV, we introduce the following components
-\begin{equation}
-H_3 = \frac{1}{6} \intt \Proj(u^3) \dx \qquad H_2 = \frac{1}{2} \intt \Proj(u_x^2) \dx
-\end{equation}
-We note that $H_3$ can be written explicitly in Fourier space as \cite{abramov2003hamiltonian} 
-\begin{equation}
-\label{H3dirsum}
-H_3 = \frac{1}{6} \sum_{\substack{ k_1 + k_2 + k_3 = 0 \\ \abs{k_1}, \abs{k_2}, \abs{k_3} \le \Lambda}}\uhat_{k_1} \uhat_{k_2} \uhat_{k_3}
+\mean{\Hp}_{\Gm} = \mean{\Hp}_{\Gp}
 \end{equation}
 
-\subsection{Some useful estimates/bounds for $H_2$ and $H_3$}
+%----------------------------------------------------------%
+% Comparison with experiments
+%----------------------------------------------------------%
+\section{Comparison with experiments}
 
-Since we examine the statistical mechanics of TKDV, it is valuable to have some a priori bounds and estimates for the components of the Hamiltonian, $H_2$ and $H_3$.
-In this section, we frequently appeal to the idea of an equipartition of energy among all the modes.
-For energy $E_0$, if the equipartition holds, then
-\begin{equation}
-\abs{u_k} = \sqrt{\frac{E_0}{\Lambda}} \qquad \text{for } k=1,\cdots,\Lambda
-\end{equation}
-We might expect equipartition of energy to hold, for example, in the case of long-time averages or ensemble averages (appealing to ergodicity).
+\subsection{Comparison of basic features}
 
-\subsubsection{Expected value of $H_2$ given $E=E_0$}
+\subsection{New experimental measurements guided by theory}
 
-First, we consider $H_2$. In Fourier space, $u_x$ is represented as
-\begin{equation}
-u_x(x) = 2 \Real \sumk i k \uhat_k e^{i k x}
-\end{equation}
-Thus, for equipartitioned energy, we have
-\begin{equation}
-H_2 = \sumk \abs{k \uhat_k}^2 = \frac{E_0}{\Lambda} \sumk k^2 = \frac{E_0}{6} (\Lambda+1)(2 \Lambda+1) \sim \frac{E_0 \Lambda^2}{3} \text{for } \Lambda \gg 1
-\end{equation}
-The above provides a useful estimate for the expected value of $H_2$ (to test the numerics for example).
-
-\subsubsection{Upper bound on $H_3$ given $E=E_0$}
-
-Next, consider $H_3$.
-By the symmetry of the cubic form, the expected value of $H_3$ is zero.
-However, to sample from Gibbs measures with non-vanishing inverse temperatures using an acceptance/rejection based algorithm, we will need to estimate the maximum allowed value of $H_3$ on the hypersphere of fixed energy $E = E_0$ so that the acceptance rate can be normalized.
-Note that the set $E=E_0$ is compact and therefore a maximum is guaranteed to exist.
-A reasonable guess for function that maximizes $H_3$ is something approximating a Dirac delta function, which naturally motivates the Dirichlet kernel
-\begin{equation}
-\label{Dkern}
-\uDir(x) = \sqrt{\frac{E_0}{\Lambda}}\sumk e^{i k x}
-\end{equation}
-We can compute $H_3(\uDir)$ by using the direct summation formula \eqref{H3dirsum}.
-Since $\uhat_k = \sqrt{E_0/\Lambda}$, each term in this sum takes the value $(E_0/\Lambda)^{3/2}$, and so we simply need to count the number of terms present in this sum.
-A careful counting (being sure to dismiss any terms in which one of $k_1, k_2, k_3$ is zero or one of these terms exceeds $\Lambda$) gives
-\begin{equation}
-\label{H3Dir}
-H_3(\uDir) = \frac{1}{2} E_0^{3/2} (\Lambda^{1/2} - \Lambda^{-1/2})
-\end{equation}
-Certainly, it must be that
-\begin{equation}
-\max H_3 \ge H_3(\uDir) 
-\end{equation}
-My conjecture is that
-\begin{equation}
-\max H_3 = H_3(\uDir) 
-\end{equation}
-or at least that this relationship holds asymptotically for $\Lambda \gg 1$.
-Preliminary numerical tests seem to support this conjecture for tested values $\Lambda = 6, 8, 10, 20$.
-In fact, the maximum of $H_3$ encountered over about $10^6$ samples is usually significantly smaller than the value predicted in \eqref{H3Dir}.
-However, for $\Lambda = 4$, my tests encounter values of $H_3$ slightly larger than the predicted value.
-
-Maximizing $H_3$ is perhaps more straightforward in physical space, where we discretize the interval $[0, 2 \pi)$ with 
-\begin{equation}
-x_n = \frac{2 \pi (n-1)}{N} \quad \text{for } n=1,\cdots,N
-\end{equation}
-where
-\begin{equation}
-N = 2 \Lambda \text{ or  } 2 \Lambda +1
-\end{equation}
-Both are valid choices for using standard real FFT packages.
-%$x_n = 2 \pi (n-1)/N$ for $n=1,\cdots,N$, and $N$ is either $2 \Lambda$ or $2 \Lambda + 1$ (both are valid choices for using standard real FFT packages). 
-In physical space, we have
-\begin{equation}
-\label{H3phys}
-H_3 = \frac{1}{6} \intt \Proj (u^3) \dx \approx \frac{1}{6N} \sumn u_n^3
-\end{equation}
-The second relationship is only an approximation because de-aliasing would be required to make the right-hand side exactly equal to \eqref{H3dirsum}.
-Nonetheless, let us proceed with the estimate.
-For fixed energy, $E=E_0$, we require
-\begin{equation}
-E_0 = \intt \Proj(u^2) \dx = \frac{1}{N} \sumn u_n^2
-\end{equation}
-(I believe the above holds with no approximation being made due to the discrete Parseval identity)
-
-Clearly, a maximizer of the right-hand-side of \eqref{H3phys} subject to $M=0$ and fixed $E=E_0$ is an approximate Dirac delta function --- i.e.
-a function with one extreme positive value, and the rest equal negative values,
-\begin{align}
-& w_1 = \sqrt{2 E_0 (N-1)}		\\
-& w_n = -\sqrt{\frac{2 E_0}{N-1}}	\qquad \text{for } n \ne 1
-\end{align}
-This function is a maximizer because any slight modification (subject to keeping $M=0$ and $E=E_0$) would reduce the value of $H_3$.
-Calculating the discrete Fourier transform of this function directly gives
-\begin{align}
-& \hat{w}_0 = 0	\\
-& \hat{w}_k = \sqrt{\frac{2 E_0}{N-1}} \qquad \text{for } k \ne 1
-\end{align}
-Notice that if we choose $N = 2\Lambda + 1$, then our delta-function in physical space corresponds exactly to the Dirichlet kernel, $w(x) = \uDir(x)$.
-Now, if we calculate $H_3(w)$ purely in physical space using the approximation in \eqref{H3phys}, we get
-\begin{equation}
-H_3(w) = \frac{\sqrt{2} E_0^{3/2}}{3} \left( \frac{N-2}{\sqrt{N-1}} \right)
-\end{equation}
-Taking $N = 2\Lambda + 1$ gives
-\begin{equation}
-\label{H3physest}
-H_3(w) = \frac{2 E_0^{3/2}}{3} \left( \Lambda^{1/2} - \Lambda^{-1/2} \right)
-\end{equation}
-So while \eqref{H3Dir} and \eqref{H3physest} take very similar forms, they differ by a multiplicative factor: 1/2 in Eq.~\eqref{H3Dir} versus 2/3 in Eq.~\eqref{H3physest}.
-Therefore, I suppose the error made by failing to de-alias in Eq.~\eqref{H3phys} is not negligible.
-
-So the impasse here as follows: It is straightforward to maximize the approximate form of $H_3$ in physical space (the right-hand-side of \eqref{H3phys}), and justifying this maximization can be made completely rigorous in my mind.
-However, the approximation made by failing to di-alias is apparently not that small.
-On the other hand, we can calculate $H_3$ exactly in spectral space, but it is not clear how to maximize $H_3$ subject to the constraints --- I can only make a guess at the maximizer.
-To overcome this issue, one might naturally want to use the de-aliased form of $H_3$ in physical space (by taking $N = 4 \Lambda$ for example), but the problem here is that all discretized functions are not permissible.
-Only functions that correspond to a Fourier series truncated at wavenumber $\Lambda$ are permissible.
-
-\np
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Numerical Methods
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Numerical methods}
-
-Here we describe the statistical numerical methods.
-The deterministic simulations of TKDV are described elsewhere.
-Below the microstate refers to the spectrum of displacement $\{ \uhat_k \}_{k=1}^{\Lambda}$.
-
-\subsection{Sampling from microcanonical energy distribution}
-\label{sec_microcan}
-
-We first consider sampling microstates from a microcanonical energy distribution with $E=E_0$ fixed and $M=0$ as always,
-\begin{equation}
-\Gibbs^0 = d\nu \propto \delta(E-E_0) \delta(M)
-\end{equation}
-This can be considered a canonical distribution in the Hamiltonian with zero inverse temperature (hence the notation $\Gibbs^0$).
-At times, it is easier to denote the distribution $d\nu$, so we will use both interchangeably.
-Note this is simply the uniform distribution on the hyper-sphere of $E=E_0$.
-
-To sample from this distribution, we use the following algorithm
-\begin{enumerate}
-\item Sample each $\uhat_k$ for $k=1,\cdots,\Lambda$ from a normal distribution $N(0,1)$.
-\item Compute $E = \sumk \abs{\uhat_k}^2$
-\item Normalize each component by multiplying by $\sqrt{E_0/E}$.
-\end{enumerate}
-This algorithm was proven to converge to the measure $d\nu$ in \cite{abramov2003hamiltonian}.
-Throughout this paper, we normalize the energy for the sampling distribution to unity.
-This is the reason why $E_0$ appears in the Hamiltonian in \eqref{Hamiltonian}.
-Hence, we will always be sampling from
-\begin{equation}
-\Gibbs^0 \propto \delta(E-1) \delta(M)
-\end{equation}
-
-\subsection{Acceptance/rejection algorithm to sample from a mixed microcanonical/canonical distribution.}
-\label{sec_accrej}
-
-Now consider the problem of sampling from a distribution that is microcanonical in energy and canonical in the Hamiltonian
-\begin{equation}
-\Gibbs^{\pm} \propto \delta(E-1) \delta(M) e^{-\theta^{\pm} H^{\pm}}
-\end{equation}
-Here, the Hamiltonian can be either the upstream or downstream, $H = H^-$ or $H^+$ and the inverse temperature $\theta$ can also take different values upstream and downstream, $\theta^-$ and $\theta^+$.
-To sample from this distribution, we use the following acceptance/rejection algorithm.
-\begin{enumerate}
-\item Sample a large number of microstates from $\Gibbs^0$.
-\item For each, compute $H_3$ and $H_2$. Note: this is the most expensive step in the algorithm. We will therefore store all the values of $H_3$ and $H_2$ so they need not be recomputed.
-\item For each sample, compute $H^+$ and $H^-$.
-\item Assign a provisional acceptance rate as $e^{- \theta H}$ (for either upstream or downstream). The only issue is this acceptance rate is not yet normalized.
-\item Find the maximum of the acceptance rate over all samples $m = \max{e^{-\theta H}}$.
-\item Use this value to normalize the acceptance rates, $e^{- \theta H}/m$.
-\item Decide to accept or reject each microstate by picking a uniform variable $v \in [0,1]$ and testing if $e^{- \theta H}/m < v$.
-\item Store the microstates, as well as the values of $H_3$ and $H_2$, that have been accepted.
-\end{enumerate}
-The above is a conceptual description of the algorithm.
-In practice, it is implemented in a way to be computational efficient.
-In particular, the most expensive step is computing $H_3$ and $H_2$ for a large number of microstates sampled from the microcanonical distribution.
-We therefore perform this step as precomputation, and save the values of $H_3$ and $H_2$.
-Afterwards, we can use this master list to perform the acceptance/rejection steps for a number of inverse temperatures, $\theta$, as well as values of $E_0$ and $D_0$.
-
-\subsection{Enforcing the statistical matching condition across ADC}
-\label{sec_match}
-
-Importantly, we must enforce the statistical matching condition at the abrupt depth change
-\begin{equation}
-\label{statmatch}
-\mean{H^+}_{\Gibbs^-} = \mean{H^+}_{\Gibbs^+}
-\end{equation}
-We regard the upstream inverse temperature $\theta^-$ as given, and the downstream value $\theta^+$ as the main unknown to be solved for.
-First, note that the ensemble average is given by the formula
-\begin{equation}
-\label{meanformula}
-\mean{H^+}_{\Gibbs^{\pm}} = \frac{ \sum H^+ \exp(- \theta^{\pm} H^{\pm}) } {\sum \exp(-\theta^{\pm} H^{\pm}) }
-\end{equation}
-where the sum is taken over the ensemble. We therefore enforcing matching condition \eqref{statmatch} as follows
-
-\begin{enumerate}
-\item Sample from a large number of microstates (e.g.~$10^8$) from the microcanonical distribution as described above. Compute and save the values of $H_3$ and $H_2$ for each microstate.
-\item For prescribed values of $\theta^-$, $E_0$, and $D_0$, compute the value $\mean{H^+}_{\Gibbs^-}$ using \eqref{meanformula}.
-\item Define a function to compute the difference $F(\theta^+) = \mean{H^+}_{\Gibbs^-} - \mean{H^+}_{\Gibbs^+}$ for a given $\theta^+$.
-\item Find a root of $F$ (e.g.~using the secant method) to determine the value of $\theta^+$ that satisfies \eqref{statmatch}.
-\end{enumerate}
-Once the value of $\theta^+$ is determined, we sample from the mixed canonical/microcanonical distributions using the algorithm described in the previous section.
-Note that in practice, we perform step 1 of generating a large number of microsamples only once, and then performing the rest of the algorithm for several different $\theta^-$ values, all using the same list of $H_3$ and $H_2$ values.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Results
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Preliminary results}
-
-We show some preliminary results below.
-First, Figs.~\ref{Hist1}--\ref{Hist3} are histograms of the upstream and downstream Hamiltonians, $H^-$ and $H^+$, respectively, for various values of inverse temperature $\theta^{\pm}$, energy $E_0$, and depth ratio $D_0$.
-These histograms are made in a very simply way, namely by sampling from the microcanonical distribution (as described in Section \ref{sec_microcan}) and then multiplying by the appropriate factor $\exp(-\theta^{\pm} H^{\pm})$ (and normalizing).
-There is nothing complex at all in generating these histograms (not even an acceptance/rejection algorithm) and they are only made to give us some initial idea of what is going on.
-
-In these tests, I chose the default values $\Lambda = 10$, $E_0 = 1$, and $D_0 = 0.6$ ($\lambda$ is set to unity throughout).
-Fig.~\ref{Hist1} shows the upstream and downstream histograms generated with these default values for the inverse temperatures $\theta^{\pm} = 0, \pm 0.3$.
-First, notice that $\theta$ can shift the mean of $H$ significantly, with negative $\theta$ shifting the mean closer to zero.
-Second, notice that the downstream histogram has a mean much closer to zero, and this histogram is more concentrated.
-
-In figure \ref{Hist2} we vary the depth ratio.
-The left shows $D_0 = 0.6$ as before, and the right shows $D_0 = 0.4$.
-Notice a slight change in the depth ratio can alter the histograms significantly.
-In figure \ref{Hist3} we vary the energy $E_0$ from the default value of 1 (left) to the value 4.
-Notice that a significant change in $E_0$ only causes slight changes in the histograms.
-
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}[p]%[htbp]
-\begin{center}
-\includegraphics[width = 0.8 \textwidth]{Hist1}
-\caption{Left: Sampling $H^-$ from the incoming Gibbs distribution $G^-$ with three different inverse temperatures $\theta^-$ (see legend).
-Right: Sampling the $H^+$ from an outgoing distribution $G^+$ with the same inverse temperatures.
-In these tests, $\Lambda = 10$, $E_0 = 1$, $\lambda = 1$, $D_0 = 0.6$, and the number of samples is $1 \times 10^7$.}
-\label{Hist1}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}[p]%[htbp]
-\begin{center}
-\includegraphics[width = 0.8 \textwidth]{Hist2}
-\caption{Sensitivity to changes in $D_0$.
-Left: Downstream histogram with $D_0 = 0.6$ (same as in Fig.~\ref{Hist1}). Right: Same with $D_0 = 0.4$. Notice the histogram changes significantly.
-}
-\label{Hist2}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}[p]%[htbp]
-\begin{center}
-\includegraphics[width = 0.8 \textwidth]{Hist3}
-\caption{Sensitivity to changes in the energy $E_0$.
-Left: Downstream histogram with $E_0 = 1$ (same as in Fig.~\ref{Hist1}). Right: Same with $E_0 = 4$. Notice a slight change in the histogram.}
-\label{Hist3}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-
- Next, in Figs.~\ref{microup0}--\ref{microdn3}, we examine various aspects of the microstates under given Gibbs distributions with possibly non-zero inverse temperatures.
-For this task, we need to use the acceptance/rejection algorithm described in Section \ref{sec_accrej}.
-In these tests, we again fix $\Lambda = 10$.
-
- First, Fig.~\ref{microup0} shows features of the microstate with zero inverse temperature (i.e.~the microcanonical distribution).
-The left panel shows a histogram of the displacement $u$ (in physical space).
-Notice the histogram is symmetric and resembles a Gaussian distribution.
-The right panel shows the $\mean{\abs{\uhat_k}}$ versus wavenumber $k$, which illustrates how the energy is distributed amongst the modes.
-Notice that the energy is evenly distributed amongst modes in this case, i.e.~a equidistribution of energy.
-
- In Fig.~\ref{microup3} we show the same plots but with $\theta^- = -0.3$.
-The histogram of $u$ is shown on the left.
-While this histogram still appears nearly symmetric, close examination reveals a slight positive skewness.
-The right shows $\mean{\abs{\uhat_k}}$.
-Notice that, in this case of negative inverse temperature, the spectrum is decidedly tilted, with less energy present in the higher modes.
-
- Figure \ref{microdn3} shows the same plots for the {\em downstream} distributions with $\theta^+ = -0.3$.
-Here, we have not enforced the statistical matching condition \eqref{statmatch}; we simply chose a value of $\theta^+$.
-First, notice that the histogram of $u$ shows a definite positive skewness.
-This observation is consistent with the main finding in the experiments!
-Second, the spectrum is tilted as before, but the tilt is not as strong as it is upstream.
-Both of these observations are consistent with the weights of $H_3$ and $H_2$.
-Downstream, the weights strongly favor $H_3$.
-Therefore, to have a better chance of being selected from the distribution $\Gibbs^{+}$, it is more advantageous to have large skewness of $u$ than it is to have small variance of $u_x$.
-
-  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}[p]%[htbp]
-\begin{center}
-\includegraphics[width = 0.8 \textwidth]{microup0}
-\caption{Sampling the upstream microstates with $\theta^- = 0$ (and $E_0 = 4$).
-Left: histogram of the surface displacement, $u^-$.
-Right: ensemble average of the spectrum, $\mean{\abs{\uhat_k}}$.
-Notice symmetric distribution of the displacement and the equipartition of energy amongst modes.}
-\label{microup0}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}[p]%[htbp]
-\begin{center}
-\includegraphics[width = 0.8 \textwidth]{microup3}
-\caption{The upstream microstates with $\theta^- = -0.3$ (and $E_0 = 4$ as before). Notice the distinct tilt to the spectrum due to $\theta^- < 0$.}
-\label{microup3}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-
-  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}[p]%[htbp]
-\begin{center}
-\includegraphics[width = 0.8 \textwidth]{microdn3}
-\caption{Examination of the downstream microstates with $\theta^+ = -0.3$ and $D_0 = 0.5$ ($E_0 = 4$ as before).
-Notice the skewness of $u$!
-The spectrum is tilted, though not as much as it is upstream.}
-\label{microdn3}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-
-Now, figures \ref{match_theta}--\ref{match_micro} show results obtained by satisfying the matching condition \eqref{statmatch} using the algorithm described in \ref{sec_match}.
-In these tests, we use $\Lambda = 8$ and the provisional number of samples is $10^8$ (the number of accepted samples will be much lower).
-We also fix $E_0 = 4$ and $D_0 = 0.5$.
-
-First, Fig.~\ref{match_theta} shows, for given values of $\theta^-$, the corresponding values of $\theta^+$ as determined by the matching condition.
-For one, $\theta^- = 0$ corresponds to $\theta^+ = 0$ which is a good sanity check.
-Second, notice that the signs of $\theta^-$ and $\theta^+$ always agree, which is a good thing.
-The relationship is not a simply linear one, as the curvature is significant.
-
- Next, Fig.~\ref{matchH} shows upstream and downstream histograms of $H$ with the matching condition enforced.
-The black curve is the histogram of $\Gibbs^-(H^-)$, while the colored curves are histograms of $H+$.
-The red curve corresponds to $\Gibbs^-(H^+)$ and the blue to $\Gibbs^+(H^+)$.
-The matching condition says that the means of these two histograms should be the same, which can be verified visually.
-Notice, the main difference is that the downstream histogram $\Gibbs^+(H^+)$ is more spread out.
-
-Figure \ref{match_micro} shows features of the underlying microstates.
-On the left, we show histograms of the displacement $u$, as we did in Figs.~\ref{microup0}--\ref{microdn3}, but this time {\em we have enforced the matching condition} \eqref{statmatch}.
-Notice the upstream $u$-histogram is nearly symmetric, while the downstream histogram is noticeable skewed in the positive direction!
-The right panel shows the ensemble-averaged spectrum $\mean{\abs{\uhat_k}}$.
-Both show a tilt in the spectrum, with the upstream tilt being more significant than downstream.
-
-We remark that simulations have been performed with larger values of $\Lambda$, for example $\Lambda = 16$ and 20.
-For the provisional number of samples fixed, increasing $\Lambda$ does not drastically increase the computational time.
-However, the main downside to increasing $\Lambda$ is that the acceptance rate for the Gibbs distributions is substantially smaller, and therefore the statistics are less resolved.
-
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[htbp]
-\begin{center}
-\includegraphics[width = 0.55 \textwidth]{match_theta}
-\caption{For given values of $\theta^-$, we compute $\theta^+$ to satisfy the matching condition.}
-\label{match_theta}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[htbp]
-\begin{center}
-\includegraphics[width = 0.75 \textwidth]{matchH}
-\caption{Histograms of $H^\pm$ under the two Gibbs measures $\Gibbs^{\pm}$ with the matching condition enforced.}
-\label{matchH}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[htbp]
-\begin{center}
-\includegraphics[width = 0.95 \textwidth]{match_micro}
-\caption{Microstate features with the matching condition enforced.
-Left: the histogram of displacement $u$ upstream (red) and downstream (blue).
-Notice the downstream histogram has a definite positive skewness.
-Right: the ensemble-averaged spectra.
-Both spectra are tilted, with the upstream tilt being more significant.}
-\label{match_micro}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-
-\np
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Results
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Big-time results}
-
-Now lets do some more serious runs with larger $\Lambda$.
-The major challenge is that the acceptance rate plummets as $\Lambda$ grows, meaning that we either have to take many more samples or develop a smarter strategy.
-For now, we will do the first.
-To run more samples, I had to modify the organization of the code.
-I am still exploiting the ability to precompute a list of $H_3$ and $H_2$ values, but I cannot precompute and store the entire list at once because it will not fit in memory.
-So instead, I take several sweeps each with a manageable number of samples (usually about $10^7$).
-
-Figure \ref{Lambda16Run} shows a run with $\Lambda = 16$, with $10^7$ samples per sweep and 200 sweeps for a total of $2 \times 10^9$ nominal samples.
-The main parameter to vary is the upstream inverse temperature $\theta^-$.
-More negative $\theta^-$ gives a more significantly skewed distribution of $u$, but also results in noisier statistics due to the lower acceptance rate.
-I therefore chose $\theta^- = -0.2$ since it gives a good balance of the two.
-The top panel of the figure shows $\theta^+$ as a function of $\theta^-$ and the histograms for $\theta^- = -0.2$.
-The bottom panel shows microstate information, including the pdf of displacement, $u$, and the spectrum.
-Notice the definite skew of the downstream displacement and the decay of both upstream and downstream spectra.
-For this run, the upstream acceptance rate was $0.01\%$ and the downstream rate was $0.008\%$.
-The total CPU time was about 33 hours.
-
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[htbp]
-\begin{center}
-\includegraphics[width = 0.95 \textwidth]{Lambda16Run}
-\caption{Big run with $\Lambda = 16$ and $2 \times 10^9$ nominal samples.}
-\label{Lambda16Run}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-
-\bibliographystyle{plain}
-\bibliography{Notesbib}
+%\bibliographystyle{plain}
+%\bibliography{Notesbib}
 

### 2019-06-17 WavesData-2019/LatexNotes/notes.tex
NMM
NM

@@
 \newcommand{\Gm}{\Gibbs^{-}}
+\newcommand{\meanp}[1]{\mean{#1}_{+}}
+\newcommand{\meanm}[1]{\mean{#1}_{-}}
 
@@
 \begin{align}
+\label{ndKdV}
 \ampp = \eps \delta^{-2} = \frac{ag}{h_0^2 f_p^2} , \qquad
@@
 \end{align}
-Here, $\lamfac$ is the number of peak wavelengths in the periodic domain (perhaps we should set $\lamfac=1$, see note below).
-Above, the displacement $u$ has been scaled on the characteristic wave amplitude $a = \mean{\eta^2}^{1/2}$, and position $x$ has been scaled on the upstream characteristic wavelength $\lambda_0 = \sqrt{g h_0}/f_p$.
+Here, the displacement has been scaled on the characteristic wave amplitude, $u = \eta/a$, where $a = \etastd= \mean{\eta^2}^{1/2}$.
+Position $x$ has been scaled on the upstream characteristic wavelength $\lambda_0 = \sqrt{g h_0}/f_p$. As a consequence of the scaling for displacement, we have fixed energy
+\begin{equation}
+\frac{1}{2} \intt u^2 \dx = \frac{1}{2}
+\end{equation}
+
+In \eqref{ndKdV} , $\lamfac$ is the number of peak wavelengths in the periodic domain (perhaps we should set $\lamfac=1$, see note below).
 Then the Hamiltonian is given by
@@
 \begin{align}
-&H_3 = \intt u^3 \dx	\, , \\
-&H_2 = \intt u_x^2 \dx	\, .
+&H_3 = \frac{1}{6} \intt u^3 \dx	\, , \\
+&H_2 = \frac{1}{2} \intt u_x^2 \dx	\, .
 \end{align}
@@
 \begin{equation}
-\mean{\Hp}_{\Gm} = \mean{\Hp}_{\Gp}
+\meanm{\Hp} = \meanp{\Hp}
+\end{equation}
+An experimental observation is that the incoming skewness is small, thus
+\begin{equation}
+\meanm{H_3} \approx 0
+\end{equation}
+which immediately gives
+\begin{equation}
+\ampp D_0^{-3/2} \meanp{H_3} = \mu D_0^{1/2} \left( \meanp{H_2} - \meanm{H_2} \right)
 \end{equation}

### 2019-06-22 WavesData-2019/LatexNotes/notes.tex
NM

@@
 Here, the displacement has been scaled on the characteristic wave amplitude, $u = \eta/a$, where $a = \etastd= \mean{\eta^2}^{1/2}$.
-Position $x$ has been scaled on the upstream characteristic wavelength $\lambda_0 = \sqrt{g h_0}/f_p$. As a consequence of the scaling for displacement, we have fixed energy
+As a consequence, $u$ has fixed energy
 \begin{equation}
-\frac{1}{2} \intt u^2 \dx = \frac{1}{2}
+E[u] = \frac{1}{2} \intt u^2 \dx = \frac{1}{2}
 \end{equation}
 
-In \eqref{ndKdV} , $\lamfac$ is the number of peak wavelengths in the periodic domain (perhaps we should set $\lamfac=1$, see note below).
+In \eqref{ndKdV}, position $x$ has been scaled on the upstream characteristic wavelength $\lambda_0 = \sqrt{g h_0}/f_p$.
+Further, $\lamfac$ is the number of peak wavelengths in the periodic domain (perhaps we should set $\lamfac=1$, see note below).
 Then the Hamiltonian is given by
@@
 \end{equation}
+or
+\begin{equation}
+\meanp{H_3} = \frac{\mu D_0^{2}}{\ampp} \left( \meanp{H_2} - \meanm{H_2} \right)
+\end{equation}
+
+Converting to dimensional (i.e.~experimental) values gives
+\begin{equation}
+\frac{\text{skew}(\eta)}{\Delta \mean{\eta_x^2}} =
+\frac{3 \mu h_0^2 f_p^2 D_0^2 \lambda^2}{g a^3}
+\end{equation}
 

### 2019-06-22 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \newcommand{\meanm}[1]{\mean{#1}_{-}}
+\newcommand{\skw}{\text{skew}}
+\newcommand{\var}{\text{var}}
 
@@
 & \Hm = \ampp H_3 - \mu H_2 \\
-& \Hp = \ampp D_0^{-3/2} H_3 - \mu D_0^{1/2} H_2
+& \Hp = \ampp D^{-3/2} H_3 - \mu D^{1/2} H_2
 \end{align}
@@
 Peak forcing frequency		& $f_p$						& frequency	& 2 Hz \\
-Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2}$		& length		& 0.03--0.3 cm \\
+Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2} = \etastd$		& length		& 0.03--0.3 cm \\
 Upstream wavelength			& $\lambda_0 = \sqrt{g h_0}/f_p$	& length		& 56 cm \\
@@
 Depth-to-wavelength ratio		& $\delta = h_0 / \lambda_0$		& dimensionless		& 0.22 \\
-Amplitude parameter			& $\ampp = \eps \delta^{-2} = ag h_0^{-2} f_p^{-2}$	
+Amplitude parameter			& $\ampp = \eps \delta^{-2} = a \lambda_0^2 h_0^{-3} = ag h_0^{-2} f_p^{-2}$	
 & dimensionless		& 0.05--0.5\\
@@
 \begin{equation}
-\ampp D_0^{-3/2} \meanp{H_3} = \mu D_0^{1/2} \left( \meanp{H_2} - \meanm{H_2} \right)
+\ampp D^{-3/2} \meanp{H_3} = \mu D^{1/2} \left( \meanp{H_2} - \meanm{H_2} \right)
 \end{equation}
-or
+or equivalently
 \begin{equation}
-\meanp{H_3} = \frac{\mu D_0^{2}}{\ampp} \left( \meanp{H_2} - \meanm{H_2} \right)
+\meanp{H_3} = \frac{\mu D^{2}}{\ampp} \, \Delta \mean{H_2}
 \end{equation}
+where $\Delta \mean{H_2} =  \meanp{H_2} - \meanm{H_2}$  indicates the difference between the upstream and downstream values.
 
-Converting to dimensional (i.e.~experimental) values gives
+To convert to dimensional, i.e.~experimental, values we use
+\begin{align}
+& \intt u^3 \dx = \frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
+& \intt u_x^2 \dx = \left(\frac{\lambda_0}{a} \right)^2 \mean{\eta_x^2} 
+= \left(\frac{\lambda_0}{a} \right)^2 \var(\eta_x)
+\end{align}
+Then we have the relationship
 \begin{equation}
-\frac{\text{skew}(\eta)}{\Delta \mean{\eta_x^2}} =
-\frac{3 \mu h_0^2 f_p^2 D_0^2 \lambda^2}{g a^3}
+\frac{\skw(\eta)}{\Delta \var(\eta_x)} =
+3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
 \end{equation}
+In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude (as well as the square of the depth ratio).
 

### 2019-06-25 WavesData-2019/LatexNotes/notes.tex
NM
N

@@
 
+% New
+\newcommand{\depth}{d}
+\newcommand{\dpm}{\depth^{-}}
+\newcommand{\dpp}{\depth^{+}}
+
 % Dashed integral
@@
 \subsection{KdV with depth change}
-In dimensional form the Korteweg–de Vries equation is given by
+Consider waves propagating unidirectionally in shallow water.
+Consider the surface displacement $\eta(x,t)$ and the reference frame moving with the local wave speed $\xi = x - ct$, where $c = \sqrt{g \depth}$ is the wave speed, $g$ gravity, and $\depth$ the local depth.
+To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation, which in dimensional form is given by
 \begin{equation}
-2 \eta_t + \frac{3 c}{h_0} \eta \eta_{\xi} + \frac{c h_0^2}{3} \eta_{\xi \xi \xi} = 0
+2 \eta_t + \frac{3 c}{\depth} \eta \eta_{\xi} + \frac{c \depth^2}{3} \eta_{\xi \xi \xi} = 0
 \end{equation}
+
+We will primarily consider the case in which waves originate from a region of constant depth, encounter an abrupt depth change, and then continue in another region of constant depth.
+Thus, depth will be piecewise constant
+\begin{align}
+\depth = 
+\begin{cases}
+\it \dpm \quad \mbox{if } x<0 \\
+\it \dpp \quad \mbox{if } x>0
+\end{cases}
+\end{align}
+Most often, we will consider waves moving into shallower depth, so that $\dpp < \dpm$.
 
@@
 
+We introduce the following characteristic scales
+\begin{align}
+\mbox{\em characteristic wave amplitude} \qquad
+a = \etastd = \mean{\eta^2}^{1/2} \\
+\mbox{\em characteristic wave length} \qquad
+ = \etastd = \mean{\eta^2}^{1/2}
+\end{align}
+
+\vsp{10}
 The dimensionless KdV equation is
@@
 If we were to do this, we would obtain only slightly different powers of $D$ in the coefficients: $D^{-7/4}$ and $D^{1/2}$ respectively.
+
+NOTE: We need to include the bandwidth, 2 Hz, in the theory somehow. For example, the peak forcing frequency and bandwidth together can be used to estimate the decay in the spectrum.
 
@@
 \begin{equation}
+\label{H3H2}
 \meanp{H_3} = \frac{\mu D^{2}}{\ampp} \, \Delta \mean{H_2}
@@
 \begin{align}
-& \intt u^3 \dx = \frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
+& \intt u^3 \dx = \frac{\mean{\eta^3}}{a^3} = 
+\frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
 & \intt u_x^2 \dx = \left(\frac{\lambda_0}{a} \right)^2 \mean{\eta_x^2} 
@@
 \end{align}
-Then we have the relationship
+Then \eqref{H3H2} gives the relationship
 \begin{equation}

### 2019-06-25 WavesData-2019/LatexNotes/notes.tex
NM

@@
 
-% Experimental parameters
-\newcommand{\lamm}{\lambda^{-}}
-\newcommand{\lamp}{\lambda^{+}}
-\newcommand{\hm}{h^{-}}
-\newcommand{\hp}{h^{+}}
+% New
 \newcommand{\etastd}{\eta_{\text std}}
-% Theory for truncated KdV and stat mech
-\newcommand{\uhat}{\hat{u}}
-\newcommand{\RR}{\mathbb{R}}
-\newcommand{\Real}{\text{Re}}
-\newcommand{\Fspace}{\mathscr{F}_{\Lambda}}
-\newcommand{\Proj}{P_{\Lambda}}
-\newcommand{\sumk}{\sum_{k=1}^{\Lambda}}
-\newcommand{\sumn}{\sum_{n=1}^{N}}
-\newcommand{\uDir}{u_{\text{Dir}}}
+\newcommand{\ampp}{\mathcal{A}}
 \newcommand{\Gibbs}{\mathcal{G}}
-\newcommand{\ampp}{\mathcal{A}}
 \newcommand{\lamfac}{N}
-
 % New
@@
 \newcommand{\depth}{d}
-\newcommand{\dpm}{\depth^{-}}
-\newcommand{\dpp}{\depth^{+}}
+\newcommand{\dup}{\depth^{-}}
+\newcommand{\ddn}{\depth^{+}}
+\newcommand{\lamup}{\lambda^{-}}
+\newcommand{\lamdn}{\lambda^{+}}
+\newcommand{\freqp}{f_p}
 
@@
 
-\subsection{KdV with depth change}
+\subsection{The Korteweg–de Vries equation with variable depth}
 Consider waves propagating unidirectionally in shallow water.
 Consider the surface displacement $\eta(x,t)$ and the reference frame moving with the local wave speed $\xi = x - ct$, where $c = \sqrt{g \depth}$ is the wave speed, $g$ gravity, and $\depth$ the local depth.
-To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation, which in dimensional form is given by
+To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation (KdV), which in dimensional form is given by
 \begin{equation}
@@
 \begin{cases}
-\it \dpm \quad \mbox{if } x<0 \\
-\it \dpp \quad \mbox{if } x>0
+\it \dup \quad \mbox{if } x<0 \\
+\it \ddn \quad \mbox{if } x>0
 \end{cases}
 \end{align}
-Most often, we will consider waves moving into shallower depth, so that $\dpp < \dpm$.
+Most often, we will consider waves moving into shallower depth, so that $\ddn < \dup$. The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$.
 
-\subsection{Nondimensionalization}
+\subsection{Nondimensionalization of KdV}
 
@@
 \begin{align}
-\mbox{\em characteristic wave amplitude} \qquad
-a = \etastd = \mean{\eta^2}^{1/2} \\
-\mbox{\em characteristic wave length} \qquad
- = \etastd = \mean{\eta^2}^{1/2}
+&a = \etastd = \mean{\eta^2}^{1/2} 
+&&\mbox{\em upstream characteristic wave amplitude} \\
+&\lamup = \sqrt{g \dup} / \freqp
+&&\mbox{\em upstream characteristic wave length}
 \end{align}

### 2019-06-25 WavesData-2019/LatexNotes/notes.tex
NM

@@
 % New
+\newcommand{\depth}{d}
+\newcommand{\dup}{\depth^{-}}
+\newcommand{\ddn}{\depth^{+}}
+\newcommand{\lamup}{\lambda^{-}}
+\newcommand{\lamdn}{\lambda^{+}}
+\newcommand{\freqp}{f_p}
+% New
+\newcommand{\meanup}[1]{\mean{#1}_{-}}
+\newcommand{\meandn}[1]{\mean{#1}_{+}}
+\newcommand{\skw}{\text{skew}}
+\newcommand{\var}{\text{var}}
+
+% New
 \newcommand{\etastd}{\eta_{\text std}}
@@
 \newcommand{\Gm}{\Gibbs^{-}}
-\newcommand{\meanp}[1]{\mean{#1}_{+}}
-\newcommand{\meanm}[1]{\mean{#1}_{-}}
-\newcommand{\skw}{\text{skew}}
-\newcommand{\var}{\text{var}}
-
-% New
-\newcommand{\depth}{d}
-\newcommand{\dup}{\depth^{-}}
-\newcommand{\ddn}{\depth^{+}}
-\newcommand{\lamup}{\lambda^{-}}
-\newcommand{\lamdn}{\lambda^{+}}
-\newcommand{\freqp}{f_p}
 
@@
 
-\subsection{Nondimensionalization of KdV}
+\subsection{Dimensionless variables}
 
@@
 \begin{align}
-&a = \etastd = \mean{\eta^2}^{1/2} 
+&a = \mean{\eta^2}^{1/2} 
 &&\mbox{\em upstream characteristic wave amplitude} \\
@@
 &&\mbox{\em upstream characteristic wave length}
+\end{align}
+and the dimensionless variables
+\begin{align}
+&u = \eta/a
+&&\mbox{\em dimensionless displacement} \\
+&\tilde{x} = \xi/\lamup = (x-ct)/\lamup
+&&\mbox{\em dimensionless position} \\
+& \tilde{t} = 
 \end{align}
@@
 \begin{equation}
-\meanm{\Hp} = \meanp{\Hp}
+\meanup{\Hp} = \meandn{\Hp}
 \end{equation}
@@
 \begin{equation}
-\meanm{H_3} \approx 0
+\meanup{H_3} \approx 0
 \end{equation}
@@
 \begin{equation}
-\ampp D^{-3/2} \meanp{H_3} = \mu D^{1/2} \left( \meanp{H_2} - \meanm{H_2} \right)
+\ampp D^{-3/2} \meandn{H_3} = \mu D^{1/2} \left( \meandn{H_2} - \meanup{H_2} \right)
 \end{equation}
@@
 \label{H3H2}
-\meanp{H_3} = \frac{\mu D^{2}}{\ampp} \, \Delta \mean{H_2}
+\meandn{H_3} = \frac{\mu D^{2}}{\ampp} \, \Delta \mean{H_2}
 \end{equation}
-where $\Delta \mean{H_2} =  \meanp{H_2} - \meanm{H_2}$  indicates the difference between the upstream and downstream values.
+where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
 

### 2019-06-27 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \end{equation}
+% The coefficients and scales can be verified on Wolfram KdV page (but what about the sign?). I would like an official source as confirmation though.
 
@@
 &&\mbox{\em dimensionless displacement} \\
-&\tilde{x} = \xi/\lamup = (x-ct)/\lamup
+&\tilde{x} = \frac{2 \pi \xi}{\lamfac \lamup} = \frac{2 \pi (x-ct)}{\lamfac \lamup}
 &&\mbox{\em dimensionless position} \\

### 2019-06-28 WavesData-2019/LatexNotes/notes.tex
NM

@@
 
-% New
-\newcommand{\depth}{d}
+% Basic physical parameters and scales
+\newcommand{\depth}{h}
 \newcommand{\dup}{\depth^{-}}
 \newcommand{\ddn}{\depth^{+}}
-\newcommand{\lamup}{\lambda^{-}}
-\newcommand{\lamdn}{\lambda^{+}}
 \newcommand{\freqp}{f_p}
-% New
+\newcommand{\lam}{\lambda}
+\newcommand{\lamfac}{N}
+% More parameters
+\newcommand{\etastd}{\eta_{\text std}}
+\newcommand{\amp}{a}
+% Statistical quantities
+\newcommand{\skw}{\text{skew}}
+\newcommand{\var}{\text{var}}
+% Dimensionless parameters
+\newcommand{\epsup}{\eps_0}
+\newcommand{\delup}{\delta_0}
+\newcommand{\drat}{D}
+\newcommand{\ampp}{\mathcal{A}}
+% Hamiltonian and Gibbs stuff
+\newcommand{\Hp}{H^{+}}
+\newcommand{\Hm}{H^{-}}
+\newcommand{\Gibbs}{\mathcal{G}}
+\newcommand{\Gp}{\Gibbs^{+}}
+\newcommand{\Gm}{\Gibbs^{-}}
 \newcommand{\meanup}[1]{\mean{#1}_{-}}
 \newcommand{\meandn}[1]{\mean{#1}_{+}}
-\newcommand{\skw}{\text{skew}}
-\newcommand{\var}{\text{var}}
-
-% New
-\newcommand{\etastd}{\eta_{\text std}}
-\newcommand{\ampp}{\mathcal{A}}
-\newcommand{\Gibbs}{\mathcal{G}}
-\newcommand{\lamfac}{N}
-% New
-\newcommand{\Hp}{H^{+}}
-\newcommand{\Hm}{H^{-}}
-\newcommand{\Gp}{\Gibbs^{+}}
-\newcommand{\Gm}{\Gibbs^{-}}
 
@@
 \begin{align}
-&a = \mean{\eta^2}^{1/2} 
-&&\mbox{\em upstream characteristic wave amplitude} \\
-&\lamup = \sqrt{g \dup} / \freqp
-&&\mbox{\em upstream characteristic wave length}
+&\amp = \mean{\eta^2}^{1/2} 
+&&\mbox{\em characteristic wave amplitude} \\
+&\lam = \sqrt{g \depth} / \freqp
+&&\mbox{\em characteristic wave length}
 \end{align}
-and the dimensionless variables
+where $\mean{\cdot}$ indicates the mean of a quantity. Note that the characteristic wavelength $\lam$ takes different values upstream and downstream of the ADC.
+We introduce the following dimensionless variables
 \begin{align}
-&u = \eta/a
-&&\mbox{\em dimensionless displacement} \\
-&\tilde{x} = \frac{2 \pi \xi}{\lamfac \lamup} = \frac{2 \pi (x-ct)}{\lamfac \lamup}
+&u = \eta/\amp
+&&\mbox{\em dimensionless surface displacement} \\
+&\tilde{x} = \frac{2 \pi \xi}{\lamfac \lam} = \frac{2 \pi (x-ct)}{\lamfac \lam}
 &&\mbox{\em dimensionless position} \\
-& \tilde{t} = 
+& \tilde{t} = \freqp^{-1}
 \end{align}
+Then the dimensionless KdV becomes
+\begin{equation}
+u_t + \frac{3 \pi}{\lamfac} \epsup \drat^{-1} u u_x + \frac{4 \pi^3}{3 \lamfac^3} \delup^2 \drat u_{xxx} = 0
+\end{equation}
+
+Sidenote: If one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
+\begin{equation}
+2 u_t + 3 \eps u u_x + \frac{\delta^2}{3} u_{xxx} = 0
+\end{equation}
+where $\eps = a/h$ and $\delta = h/\lam$.
 

### 2019-07-10 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \newcommand{\lam}{\lambda}
+\newcommand{\lamup}{\lam^{-}}
 \newcommand{\lamfac}{N}
@@
 &\tilde{x} = \frac{2 \pi \xi}{\lamfac \lam} = \frac{2 \pi (x-ct)}{\lamfac \lam}
-&&\mbox{\em dimensionless position} \\
+&&\mbox{\em dimensionless position in moving frame} \\
 & \tilde{t} = \freqp^{-1}
+&&\mbox{\em dimensionless time} \\
 \end{align}
@@
 \end{equation}
+where $\epsup = \amp/\dup$, $\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g}$, and $\drat = {\depth}/{\dup}$ is the local dimensionless depth.
 
+\vsp{2}
 Sidenote: If one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
@@
 
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[htbp]
+\begin{center}
+\includegraphics[width = 0.80 \textwidth]{../Figs/SkewAmp.pdf}
+\caption{\label{fig1} 
+The ratio ${\skw(\eta)}/{\Delta \var(\eta_x)}$ plots against dimensionless wave amplitude.
+}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
 %----------------------------------------------------------%

### 2019-07-10 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \end{align}
-where $\mean{\cdot}$ indicates the mean of a quantity. Note that the characteristic wavelength $\lam$ takes different values upstream and downstream of the ADC.
+where $\mean{\cdot}$ indicates the mean of a quantity. Note that the characteristic wavelength $\lam = \lam(x)$ takes different values upstream and downstream of the ADC.
 We introduce the following dimensionless variables
@@
 &&\mbox{\em dimensionless position in moving frame} \\
-& \tilde{t} = \freqp^{-1}
+& \tilde{t} = t \freqp
 &&\mbox{\em dimensionless time} \\
@@
 \vsp{10}
+Old notes:
 The dimensionless KdV equation is

### 2019-07-12 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \end{align}
-Most often, we will consider waves moving into shallower depth, so that $\ddn < \dup$. The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$.
+Most often, we will consider waves moving into shallower depth, so that $\dup > \ddn$. The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$.
 
@@
 \vsp{10}
-Old notes:
+Old notes here (I need to make sure everything is consistent and then remove this section below in favor of the newer, simpler writeup above).
 The dimensionless KdV equation is

### 2019-07-12 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \begin{align}
+\label{ampdefn}
 &\amp = \mean{\eta^2}^{1/2} 
 &&\mbox{\em characteristic wave amplitude} \\
+\label{lamdefn}
 &\lam = \sqrt{g \depth} / \freqp
@@
 \end{align}
-Then the dimensionless KdV becomes
+We will consider the physical domain $\xi \in [0, \lamfac \lam]$, i.e.~an integer multiple of the characteristic wavelength.
+Then (after a shift) the dimensionless domain can be taken to be $\tilde{x} \in [-\pi,\pi]$.
+
+The dimensionless KdV equation takes the form
 \begin{equation}
-u_t + \frac{3 \pi}{\lamfac} \epsup \drat^{-1} u u_x + \frac{4 \pi^3}{3 \lamfac^3} \delup^2 \drat u_{xxx} = 0
+u_t + \frac{3 \pi}{\lamfac} \epsup \drat^{-1} \, u u_x + \frac{4 \pi^3}{3 \lamfac^3} \delup^2 \drat \, u_{xxx} = 0
+\qquad \text{for } x \in [-\pi,\pi]
 \end{equation}
-where $\epsup = \amp/\dup$, $\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g}$, and $\drat = {\depth}/{\dup}$ is the local dimensionless depth.
+where dimensionless parameters are as follows
+\begin{align}
+&\epsup = \amp/\dup \, , \\
+&\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g} \, \\
+&\drat(x) = {\depth(x)}/{\dup} \, .
+\end{align}
+$\drat = \drat(x)$ is the local dimensionless depth.
 
-\vsp{2}
-Sidenote: If one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
+\begin{comment}
+{\bf Sidenote}: In the case of constant depth, if one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
 \begin{equation}
@@
 where $\eps = a/h$ and $\delta = h/\lam$.
+\end{comment}
+
+\subsection{Hamiltonian analysis}
+
+Note, that as a consequence of the definition of $\amp$ in \eqref{ampdefn}, we have that
+\begin{equation}
+E[u] = \frac{1}{2} \frac{1}{2 \pi} \int_{-\pi}^{\pi} u^2 \dx = \frac{1}{2}
+\end{equation}
 
 \vsp{10}
-Old notes here (I need to make sure everything is consistent and then remove this section below in favor of the newer, simpler writeup above).
-The dimensionless KdV equation is
-\begin{align}
-\label{varKdV}
-u_t + \ampp D^{-3/2} \, u u_x + \mu D^{1/2} \, u_{xxx} = 0
-\end{align}
-with the dimensionless parameters
-\begin{align}
-\label{ndKdV}
-\ampp = \eps \delta^{-2} = \frac{ag}{h_0^2 f_p^2} , \qquad
-\mu = \frac{4 \pi^2}{9 \lamfac^2}
-\end{align}
-Here, the displacement has been scaled on the characteristic wave amplitude, $u = \eta/a$, where $a = \etastd= \mean{\eta^2}^{1/2}$.
-As a consequence, $u$ has fixed energy
-\begin{equation}
-E[u] = \frac{1}{2} \intt u^2 \dx = \frac{1}{2}
-\end{equation}
 
-In \eqref{ndKdV}, position $x$ has been scaled on the upstream characteristic wavelength $\lambda_0 = \sqrt{g h_0}/f_p$.
-Further, $\lamfac$ is the number of peak wavelengths in the periodic domain (perhaps we should set $\lamfac=1$, see note below).
 Then the Hamiltonian is given by

### 2019-07-12 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \newcommand{\eps}{\varepsilon}
+\newcommand{\defeq}{\mathrel{\mathop:}=}
 
@@
 \newcommand{\depth}{h}
-\newcommand{\dup}{\depth^{-}}
-\newcommand{\ddn}{\depth^{+}}
+\newcommand{\dup}{\depth_{-}}
+\newcommand{\ddn}{\depth_{+}}
 \newcommand{\freqp}{f_p}
 \newcommand{\lam}{\lambda}
-\newcommand{\lamup}{\lam^{-}}
+\newcommand{\lamup}{\lam_{-}}
 \newcommand{\lamfac}{N}
@@
 
-The dimensionless KdV equation takes the form
+Dropping the tildes, the dimensionless KdV equation takes the form
 \begin{equation}
@@
 \end{equation}
-where dimensionless parameters are as follows
+where dimensionless variables are as follows
 \begin{align}
@@
 \end{align}
-$\drat = \drat(x)$ is the local dimensionless depth.
+Here, $\drat = \drat(x)$ is the local dimensionless depth.
 
@@
 
+% Table
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{table}[h]%[htbp]
+\begin{center}
+\caption{Table of parameters}
+\label{paramtable}
+\begin{tabular}{l l l l}
+\hline \multicolumn{4} { c }{Parameters that are constant throughout the domain} \\
+\hline Description & Notation and definition & Units & Value in experiments \\
+\hline
+Peak forcing frequency		& $f_p$						& frequency	& 2 Hz \\
+Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2} = \etastd$		& length		& 0.03--0.3 cm \\
+Upstream wavelength			& $\lamup = \sqrt{g \dup}/f_p$	& length		& 56 cm \\
+Amplitude-to-depth ratio		& $\epsup = a / \dup$			& dimensionless	& 0.0024--0.024 \\
+Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& dimensionless		& 0.22 \\
+%Amplitude parameter			& $\ampp = \eps \delta^{-2} = a \lambda_0^2 h_0^{-3} = ag h_0^{-2} f_p^{-2}$	& dimensionless		& 0.05--0.5\\
+%Catch-all parameter		& $\mu = 4 \pi^2 / (9 \lamfac^2)$			& dimensionless		& 0.27 if $\lamfac=4$ \\
+\hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
+\hline Description & Notation and definition & Units \\
+\hline
+Local depth			& $\depth$				& length \\
+Local wavespeed		& $c = \sqrt{g \depth}$		& speed \\
+Local peak wavelength	& $\lam = \sqrt{gd}/f_p$	& length \\
+%Characteristic wavelength	& $\lambda_c = \lamfac \lambda_p$	& length \\
+Depth ratio			& $\drat = \depth/\dup$		& dimensionless
+\end{tabular}
+\end{center}
+\end{table}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+\subsubsection{Alternative KdV formulation following Johnson}
+The above makes the implicit assumption of continuous surface displacement crossing the ADC.
+This is a very physically motivated assumption, but there is a small amount of debate surrounding the assumption.
+An alternative is to appeal to Green's law and assume the quantity $\depth^{1/4} \eta$ to be continuous across the ADC, which necessarily implies a small discontinuity in the surface (which of course cannot be true literally).
+The dimensional KdV equation would then be expressed as
+\begin{equation}
+2 (\depth^{1/4} \eta)_t + \frac{3 c}{\depth^{5/4}} (\depth^{1/4} \eta) (\depth^{1/4} \eta)_{\xi} + \frac{c \depth^2}{3} (\depth^{1/4} \eta)_{\xi \xi \xi} = 0
+\end{equation}
+This is the approach taken in Johnson's book. This assumption results in a small modification of the coefficients in the dimensionless KdV. In particular, it gives
+\begin{equation}
+u_t + \frac{3 \pi}{\lamfac} \epsup \drat^{-5/4} \, u u_x + \frac{4 \pi^3}{3 \lamfac^3} \delup^2 \drat \, u_{xxx} = 0
+\qquad \text{for } x \in [-\pi,\pi]
+\end{equation}
+where the state variable is
+\begin{equation}
+u = \frac{\depth^{1/4} \, \eta}{ \dup^{1/4} \, \amp}
+\end{equation}
+Notice the only changes are the definition of the state variable and the small change in the power of $\drat$ in the second term (changes from $\drat^{-1}$ to $\drat^{-5/4}$).
+
 \subsection{Hamiltonian analysis}
 
-Note, that as a consequence of the definition of $\amp$ in \eqref{ampdefn}, we have that
+The domain of consideration is $x \in [-\pi,\pi]$. We first define the spatial average of any quantity as
 \begin{equation}
-E[u] = \frac{1}{2} \frac{1}{2 \pi} \int_{-\pi}^{\pi} u^2 \dx = \frac{1}{2}
+\mean{q} = \frac{1}{2\pi} \int_{-\pi}^{\pi} q \dx \, .
+\end{equation}
+Note that from the definition of the characteristic amplitude in \eqref{ampdefn}, we have $\mean{u^2} = 1$. Consequently, the energy of $u$ is fixed as
+\begin{equation}
+E[u] \defeq \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = \pi
 \end{equation}
@@
 NOTE: We need to include the bandwidth, 2 Hz, in the theory somehow. For example, the peak forcing frequency and bandwidth together can be used to estimate the decay in the spectrum.
-
-% Table
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{table}[h]%[htbp]
-\begin{center}
-\caption{Table of parameters}
-\label{paramtable}
-\begin{tabular}{l l l l}
-\hline \multicolumn{4} { c }{Parameters that are constant throughout the domain} \\
-\hline Description & Notation and definition & Units & Value in experiments \\
-\hline
-Peak forcing frequency		& $f_p$						& frequency	& 2 Hz \\
-Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2} = \etastd$		& length		& 0.03--0.3 cm \\
-Upstream wavelength			& $\lambda_0 = \sqrt{g h_0}/f_p$	& length		& 56 cm \\
-Amplitude-to-depth ratio		& $\eps = a / h_0$				& dimensionless	& 0.0024--0.024 \\
-Depth-to-wavelength ratio		& $\delta = h_0 / \lambda_0$		& dimensionless		& 0.22 \\
-Amplitude parameter			& $\ampp = \eps \delta^{-2} = a \lambda_0^2 h_0^{-3} = ag h_0^{-2} f_p^{-2}$	
-& dimensionless		& 0.05--0.5\\
-Catch-all parameter		& $\mu = 4 \pi^2 / (9 \lamfac^2)$			& dimensionless		& 0.27 if $\lamfac=4$ \\
-\hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
-\hline Description & Notation and definition & Units \\
-\hline
-Local depth			& $d$					& length \\
-Local wavespeed		& $c = \sqrt{gd}$			& speed \\
-Local peak wavelength	& $\lambda_p = \sqrt{gd}/f_p$	& length \\
-Characteristic wavelength	& $\lambda_c = \lamfac \lambda_p$	& length \\
-Depth ratio			& $D = d/h_0$				& dimensionless
-\end{tabular}
-\end{center}
-\end{table}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 

### 2019-07-15 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \end{align}
-Most often, we will consider waves moving into shallower depth, so that $\dup > \ddn$. The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$.
+Most often, we will consider waves moving into shallower depth, so that $\dup > \ddn$.
+The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$, which gives rise to peak wavelength $\lam = c/\freqp = \sqrt{g \depth} / \freqp$.
 
+% Nondimensionalization
 \subsection{Dimensionless variables}
-
+\label{nondim}
 We introduce the following characteristic scales
@@
 &&\mbox{\em dimensionless position in moving frame} \\
-& \tilde{t} = t \freqp
+& \tilde{t} = t \freqp / \lamfac
 &&\mbox{\em dimensionless time} \\
 \end{align}
-We will consider the physical domain $\xi \in [0, \lamfac \lam]$, i.e.~an integer multiple of the characteristic wavelength.
-Then (after a shift) the dimensionless domain can be taken to be $\tilde{x} \in [-\pi,\pi]$.
+We will consider the physical domain $\xi \in [0, \lamfac \lam]$, which is an integer multiple of the characteristic wavelength where $\lamfac$ will be selected later (probably about 5).
+Above, we have chosen the scale to normalize position as $\lamfac \lam/(2\pi)$ so that the dimensionless domain can be taken (after a shift) to be $\tilde{x} \in [-\pi,\pi]$.
+Accordingly, we have also chosen the time scale as $\lamfac/\freqp$, so that a wave traveling at the local wavespeed $c = \lam \freqp = \sqrt{g \depth}$ crosses the physical domain $x \in [0, \lamfac \lam]$ in one unit of dimensionless time.
+Said differently, the dimensionless wave speed is $(2 \pi)^{-1}$ so that a wave crosses the entire domain $[-\pi,\pi]$ in unit time.
 
@@
 \begin{equation}
-u_t + \frac{3 \pi}{\lamfac} \epsup \drat^{-1} \, u u_x + \frac{4 \pi^3}{3 \lamfac^3} \delup^2 \drat \, u_{xxx} = 0
+u_t + {3 \pi} \epsup \drat^{-1} \, u u_x + \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \drat \, u_{xxx} = 0
 \qquad \text{for } x \in [-\pi,\pi]
@@
 \end{align}
-Here, $\drat = \drat(x)$ is the local dimensionless depth.
+Here, $\drat = \drat(x)$ is the local dimensionless depth. See Table \ref{paramtable} for a summary of important parameters.
 
@@
 \label{paramtable}
-\begin{tabular}{l l l l}
-\hline \multicolumn{4} { c }{Parameters that are constant throughout the domain} \\
-\hline Description & Notation and definition & Units & Value in experiments \\
+\begin{tabular}{l l l}
+\hline \multicolumn{3} { c }{Parameters that are constant in a single experiment} \\
+\hline Description & Notation and definition & Value in experiments \\
 \hline
-Peak forcing frequency		& $f_p$						& frequency	& 2 Hz \\
-Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2} = \etastd$		& length		& 0.03--0.3 cm \\
-Upstream wavelength			& $\lamup = \sqrt{g \dup}/f_p$	& length		& 56 cm \\
-Amplitude-to-depth ratio		& $\epsup = a / \dup$			& dimensionless	& 0.0024--0.024 \\
-Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& dimensionless		& 0.22 \\
-%Amplitude parameter			& $\ampp = \eps \delta^{-2} = a \lambda_0^2 h_0^{-3} = ag h_0^{-2} f_p^{-2}$	& dimensionless		& 0.05--0.5\\
-%Catch-all parameter		& $\mu = 4 \pi^2 / (9 \lamfac^2)$			& dimensionless		& 0.27 if $\lamfac=4$ \\
-\hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
-\hline Description & Notation and definition & Units \\
+Peak forcing frequency		& $f_p$						& 2 Hz \\
+Upstream depth			& $\dup$						& 12.5 cm \\
+Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
+Upstream wavelength		& $\lamup = \sqrt{g \dup}/f_p$		& 56 cm \\
+Amplitude-to-depth ratio		& $\epsup = a / \dup$			& 0.0024--0.024 \\
+Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
+\hline \multicolumn{3} { c }{Parameters that change value across ADC} \\
+\hline Description & Notation and definition \\
 \hline
-Local depth			& $\depth$				& length \\
-Local wavespeed		& $c = \sqrt{g \depth}$		& speed \\
-Local peak wavelength	& $\lam = \sqrt{gd}/f_p$	& length \\
-%Characteristic wavelength	& $\lambda_c = \lamfac \lambda_p$	& length \\
-Depth ratio			& $\drat = \depth/\dup$		& dimensionless
+Local depth			& $\depth$			\\
+Local wavespeed		& $c = \sqrt{g \depth}$	\\
+Local peak wavelength	& $\lam = \sqrt{gd}/f_p$	\\
+Depth ratio			& $\drat = \depth/\dup$	
 \end{tabular}
@@
 \subsubsection{Alternative KdV formulation following Johnson}
-The above makes the implicit assumption of continuous surface displacement crossing the ADC.
+Section \ref{nondim} makes the implicit assumption of continuous surface displacement crossing the ADC.
 This is a very physically motivated assumption, but there is a small amount of debate surrounding the assumption.
-An alternative is to appeal to Green's law and assume the quantity $\depth^{1/4} \eta$ to be continuous across the ADC, which necessarily implies a small discontinuity in the surface (which of course cannot be true literally).
+An alternative is to appeal to Green's law and assume the quantity $\depth^{1/4} \eta$ to be continuous across the ADC, which necessarily implies a small discontinuity in the surface (which of course cannot be true literally), as is the approach taken in Johnson't book.
 The dimensional KdV equation would then be expressed as
@@
 \end{equation}
-This is the approach taken in Johnson's book. This assumption results in a small modification of the coefficients in the dimensionless KdV. In particular, it gives
+Then, the dimensionless state variable is instead 
+\begin{equation}
+u = \frac{\depth^{1/4} \, \eta}{ \dup^{1/4} \, \amp}
+\end{equation}
+and the dimensionless KdV equation takes the form
 \begin{equation}
@@
 \end{equation}
-where the state variable is
-\begin{equation}
-u = \frac{\depth^{1/4} \, \eta}{ \dup^{1/4} \, \amp}
-\end{equation}
-Notice the only changes are the definition of the state variable and the small change in the power of $\drat$ in the second term (changes from $\drat^{-1}$ to $\drat^{-5/4}$).
+Notice the only difference is a small change in the power of $\drat$ in the second term (changes from $\drat^{-1}$ to $\drat^{-5/4}$).
+We will not use this alternative formulation.
 
@@
 
-The domain of consideration is $x \in [-\pi,\pi]$. We first define the spatial average of any quantity as
+The domain of consideration is $x \in [-\pi,\pi]$. First, the spatial average of any quantity is given by
 \begin{equation}
@@
 \end{equation}
-Note that from the definition of the characteristic amplitude in \eqref{ampdefn}, we have $\mean{u^2} = 1$. Consequently, the energy of $u$ is fixed as
+Appealing to near Ergodicity, we will typically assume equivalence between spatial, temporal, and ensemble averages.
+Note that from the definition of the characteristic amplitude in \eqref{ampdefn}, we have $\mean{u^2} = 1$.
+Consequently, the energy of $u$ is fixed as
 \begin{equation}

### 2019-07-15 WavesData-2019/LatexNotes/notes.tex
NM

@@
 % Hamiltonian and Gibbs stuff
-\newcommand{\Hp}{H^{+}}
-\newcommand{\Hm}{H^{-}}
+\newcommand{\Ham}{H}
+\newcommand{\Hp}{\Ham^{+}}
+\newcommand{\Hm}{\Ham^{-}}
 \newcommand{\Gibbs}{\mathcal{G}}
@@
 \label{nondim}
-We introduce the following characteristic scales
-\begin{align}
-\label{ampdefn}
-&\amp = \mean{\eta^2}^{1/2} 
-&&\mbox{\em characteristic wave amplitude} \\
-\label{lamdefn}
-&\lam = \sqrt{g \depth} / \freqp
-&&\mbox{\em characteristic wave length}
-\end{align}
-where $\mean{\cdot}$ indicates the mean of a quantity. Note that the characteristic wavelength $\lam = \lam(x)$ takes different values upstream and downstream of the ADC.
-We introduce the following dimensionless variables
-\begin{align}
-&u = \eta/\amp
-&&\mbox{\em dimensionless surface displacement} \\
-&\tilde{x} = \frac{2 \pi \xi}{\lamfac \lam} = \frac{2 \pi (x-ct)}{\lamfac \lam}
-&&\mbox{\em dimensionless position in moving frame} \\
-& \tilde{t} = t \freqp / \lamfac
-&&\mbox{\em dimensionless time} \\
-\end{align}
-We will consider the physical domain $\xi \in [0, \lamfac \lam]$, which is an integer multiple of the characteristic wavelength where $\lamfac$ will be selected later (probably about 5).
-Above, we have chosen the scale to normalize position as $\lamfac \lam/(2\pi)$ so that the dimensionless domain can be taken (after a shift) to be $\tilde{x} \in [-\pi,\pi]$.
-Accordingly, we have also chosen the time scale as $\lamfac/\freqp$, so that a wave traveling at the local wavespeed $c = \lam \freqp = \sqrt{g \depth}$ crosses the physical domain $x \in [0, \lamfac \lam]$ in one unit of dimensionless time.
-Said differently, the dimensionless wave speed is $(2 \pi)^{-1}$ so that a wave crosses the entire domain $[-\pi,\pi]$ in unit time.
-
-Dropping the tildes, the dimensionless KdV equation takes the form
-\begin{equation}
-u_t + {3 \pi} \epsup \drat^{-1} \, u u_x + \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \drat \, u_{xxx} = 0
-\qquad \text{for } x \in [-\pi,\pi]
-\end{equation}
-where dimensionless variables are as follows
-\begin{align}
-&\epsup = \amp/\dup \, , \\
-&\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g} \, \\
-&\drat(x) = {\depth(x)}/{\dup} \, .
-\end{align}
-Here, $\drat = \drat(x)$ is the local dimensionless depth. See Table \ref{paramtable} for a summary of important parameters.
-
-\begin{comment}
-{\bf Sidenote}: In the case of constant depth, if one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
-\begin{equation}
-2 u_t + 3 \eps u u_x + \frac{\delta^2}{3} u_{xxx} = 0
-\end{equation}
-where $\eps = a/h$ and $\delta = h/\lam$.
-\end{comment}
 
@@
 
+We introduce the following characteristic scales
+\begin{align}
+\label{ampdefn}
+&\amp = \mean{\eta^2}^{1/2} 
+&&\mbox{\em characteristic wave amplitude} \\
+\label{lamdefn}
+&\lam = \sqrt{g \depth} / \freqp
+&&\mbox{\em characteristic wave length}
+\end{align}
+where $\mean{\cdot}$ indicates the mean of a quantity.
+Note that the characteristic wavelength $\lam = \lam(x)$ takes different values upstream and downstream of the ADC.
+We remark that experimental measurements indicate that the characteristic amplitude $\amp = \mean{\eta^2}^{1/2}$ is nearly the same on both values of the ADC.
+Hence, we will not distinguish between upstream and downstream values of $\amp$.
+We introduce the following dimensionless variables
+\begin{align}
+&u = \eta/\amp
+&&\mbox{\em dimensionless surface displacement} \\
+&\tilde{x} = \frac{2 \pi \xi}{\lamfac \lam} = \frac{2 \pi (x-ct)}{\lamfac \lam}
+&&\mbox{\em dimensionless position in moving frame} \\
+& \tilde{t} = t \freqp / \lamfac
+&&\mbox{\em dimensionless time} \\
+\end{align}
+We will consider the physical domain $\xi \in [0, \lamfac \lam]$, which is an integer multiple of the characteristic wavelength where $\lamfac$ will be selected later (probably about 5).
+Above, we have chosen the scale to normalize position as $\lamfac \lam/(2\pi)$ so that the dimensionless domain can be taken (after a shift) to be $\tilde{x} \in [-\pi,\pi]$.
+Accordingly, we have also chosen the time scale as $\lamfac/\freqp$, so that a wave traveling at the local wavespeed $c = \lam \freqp = \sqrt{g \depth}$ crosses the physical domain $x \in [0, \lamfac \lam]$ in one unit of dimensionless time.
+Said differently, the dimensionless wave speed is $(2 \pi)^{-1}$ so that a wave crosses the domain $[-\pi,\pi]$ in unit time.
+
+Dropping the tildes, the dimensionless KdV equation takes the form
+\begin{equation}
+\label{dimlessKdV}
+u_t + {3 \pi} \epsup \drat^{-1} \, u u_x + \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \drat \, u_{xxx} = 0
+\qquad \text{for } x \in [-\pi,\pi]
+\end{equation}
+where dimensionless variables are as follows
+\begin{align}
+&\epsup = \amp/\dup \, , \\
+&\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g} \, \\
+&\drat(x) = {\depth(x)}/{\dup} \, .
+\end{align}
+Here, $\drat = \drat(x)$ is the local dimensionless depth. See Table \ref{paramtable} for a summary of important parameters.
+
+\begin{comment}
+{\bf Sidenote}: In the case of constant depth, if one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
+\begin{equation}
+2 u_t + 3 \eps u u_x + \frac{\delta^2}{3} u_{xxx} = 0
+\end{equation}
+where $\eps = a/h$ and $\delta = h/\lam$.
+\end{comment}
+
 \subsubsection{Alternative KdV formulation following Johnson}
@@
 
-\subsection{Hamiltonian analysis}
+\subsection{Hamiltonian structure}
 
-The domain of consideration is $x \in [-\pi,\pi]$. First, the spatial average of any quantity is given by
+We now consider the Hamiltonian structure of the dimensionless variable-depth KdV equation \eqref{dimlessKdV}, over dimensionless domain $x \in [-\pi,\pi]$.
+First, the spatial average of any quantity is given by
 \begin{equation}
@@
 Appealing to near Ergodicity, we will typically assume equivalence between spatial, temporal, and ensemble averages.
-Note that from the definition of the characteristic amplitude in \eqref{ampdefn}, we have $\mean{u^2} = 1$.
-Consequently, the energy of $u$ is fixed as
+
+We measure the state variable $u$ as the displacement of the surface from equilibrium, so that $\mean{u} = 0$. Consequently, the momentum of $u$ vanishes
+\begin{equation}
+M[u] \defeq \int_{-\pi}^{\pi} u \dx = 0
+\end{equation}
+Next, the definition of the characteristic amplitude in \eqref{ampdefn} implies that $\mean{u^2} = 1$. Consequently, the energy of $u$ is fixed as
 \begin{equation}
@@
 
-\vsp{10}
-
-Then the Hamiltonian is given by
+We introduce the cubic and quadratic Hamiltonian components as
 \begin{align}
-\label{Hamiltonian}
-& \Hm = \ampp H_3 - \mu H_2 \\
-& \Hp = \ampp D^{-3/2} H_3 - \mu D^{1/2} H_2
-\end{align}
-where
-\begin{align}
-&H_3 = \frac{1}{6} \intt u^3 \dx	\, , \\
-&H_2 = \frac{1}{2} \intt u_x^2 \dx	\, .
+H_3 = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
+H_2 = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
 \end{align}
 
-Note: after giving it some deliberation, I believe we should use $\lamfac=1$.
-My reason is that the peak in the wave spectrum occurs at a frequency of $f_p$ or wavelength of $\lambda_p$.
+Then the Hamiltonian of \eqref{dimlessKdV} is given by
+\begin{equation}
+\Ham = 3 \pi \epsup \drat^{-1} \, H_3 - \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \drat \, H_2
+\end{equation}
+
+\begin{comment}
+Old Note (around May 2019): After giving it some deliberation, I believe we should use $\lamfac=1$.
+My reason is that the peak in the wave spectrum occurs at a frequency of $\freqp$ or wavelength of $\lam$.
 In the tKdV Gibbs measure, the spectrum decays monotonically, so that the peak is at the lowest resolved frequency (or largest resolved wavelength).
-Thus, we want $\lambda_p$ to correspond to the largest resolved wavelength, i.e.~the length of the periodic domain in the tKdV framework.
+Thus, we want $\lam$ to correspond to the largest resolved wavelength, i.e.~the length of the periodic domain in the tKdV framework.
 
-%Remark: We note that another possibility is to match the quantity $D^{1/4} \eta$ at the abrupt depth change as would be consistent with the analysis of Johnson.
-If we were to do this, we would obtain only slightly different powers of $D$ in the coefficients: $D^{-7/4}$ and $D^{1/2}$ respectively.
-
-NOTE: We need to include the bandwidth, 2 Hz, in the theory somehow. For example, the peak forcing frequency and bandwidth together can be used to estimate the decay in the spectrum.
+New Thought (July 2019): Actually, it is very possible that the bandwidth (2 Hz in the experiments) sets the value of $\lamfac$ that is most appropriate, since that bandwidth sets the decay rate of the spectrum.
+My thinking is that our experimental forcing does not really follow the upstream Gibbs measure all that well (because we did not have that in mind), but perhaps it best approximates the tail decay of some Gibbs measure.
+Perhaps, the lowest frequencies (i.e.
+those much slower than 2 Hz that are present in the theory but almost absent in the experiments) do not affect statistics that much.
+\end{comment}
 

### 2019-07-15 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \newcommand{\lamfac}{N}
-% More parameters
-\newcommand{\etastd}{\eta_{\text std}}
 \newcommand{\amp}{a}
+%\newcommand{\etastd}{\eta_{\text std}}
 % Statistical quantities
@@
 \newcommand{\drat}{D}
+\newcommand{\dratdn}{\drat_0}
 \newcommand{\ampp}{\mathcal{A}}
@@
 \newcommand{\Ham}{H}
-\newcommand{\Hp}{\Ham^{+}}
-\newcommand{\Hm}{\Ham^{-}}
+\newcommand{\Hup}{\Ham^{-}}
+\newcommand{\Hdn}{\Ham^{+}}
 \newcommand{\Gibbs}{\mathcal{G}}
-\newcommand{\Gp}{\Gibbs^{+}}
-\newcommand{\Gm}{\Gibbs^{-}}
+\newcommand{\Gup}{\Gibbs^{-}}
+\newcommand{\Gdn}{\Gibbs^{+}}
 \newcommand{\meanup}[1]{\mean{#1}_{-}}
@@
 
-%\author{
-% Tyler Bolles
-%Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
-%M.~N.~J.~Moore\thanks{Florida State University}, 
-%Di Qi\thanks{Courant Institute of Mathematical Sciences} }
+\author{
+C. Tyler Bolles\thanks{University of Michigan},
+Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
+M.~N.~J.~Moore\thanks{Florida State University}, 
+Di Qi\thanks{Courant Institute of Mathematical Sciences} }
 \maketitle
@@
 \begin{cases}
-\it \dup \quad \mbox{if } x<0 \\
-\it \ddn \quad \mbox{if } x>0
+\dup \quad \mbox{if } x<0 \\
+\ddn \quad \mbox{if } x>0
 \end{cases}
@@
 Upstream depth			& $\dup$						& 12.5 cm \\
+Downstream depth			& $\ddn$						& 3 cm \\
 Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
@@
 Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
+Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 \\
 \hline \multicolumn{3} { c }{Parameters that change value across ADC} \\
@@
 Local peak wavelength	& $\lam = \sqrt{gd}/f_p$	\\
-Depth ratio			& $\drat = \depth/\dup$	
+Local dimensionless depth		& $\drat = \depth/\dup$	
 \end{tabular}
@@
 & \tilde{t} = t \freqp / \lamfac
-&&\mbox{\em dimensionless time} \\
+&&\mbox{\em dimensionless time}
 \end{align}
@@
 Said differently, the dimensionless wave speed is $(2 \pi)^{-1}$ so that a wave crosses the domain $[-\pi,\pi]$ in unit time.
+
+%----------------------------------------------------------------------------%
+% COMMENT ON LAMFAC
+\begin{comment}
+Old Note (around May 2019): After giving it some deliberation, I believe we should use $\lamfac=1$.
+My reason is that the peak in the wave spectrum occurs at a frequency of $\freqp$ or wavelength of $\lam$.
+In the tKdV Gibbs measure, the spectrum decays monotonically, so that the peak is at the lowest resolved frequency (or largest resolved wavelength).
+Thus, we want $\lam$ to correspond to the largest resolved wavelength, i.e.~the length of the periodic domain in the tKdV framework.
+
+New Thought (July 2019): Actually, it is very possible that the bandwidth (2 Hz in the experiments) sets the value of $\lamfac$ that is most appropriate, since that bandwidth sets the decay rate of the spectrum.
+My thinking is that our experimental forcing does not really follow the upstream Gibbs measure all that well (because we did not have that in mind), but perhaps it best approximates the tail decay of some Gibbs measure.
+Perhaps, the lowest frequencies (i.e.
+those much slower than 2 Hz that are present in the theory but almost absent in the experiments) do not affect statistics that much.
+\end{comment}
+%----------------------------------------------------------------------------%
 
@@
 &\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g} \, \\
-&\drat(x) = {\depth(x)}/{\dup} \, .
+&\drat = {\depth}/{\dup} \, .
 \end{align}
-Here, $\drat = \drat(x)$ is the local dimensionless depth. See Table \ref{paramtable} for a summary of important parameters.
+Here, $\drat$ is the local dimensionless depth, which changes value crossing the ADC.
+In particular, since \eqref{dimlessKdV} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$.
+Then we have
+\begin{equation}
+\drat = 
+\begin{cases}
+1 		&\quad \mbox{for } t<0 \\
+\dratdn 	&\quad \mbox{for } t>0
+\end{cases}
+\end{equation}
+where $\dratdn = \ddn/\dup$.
+See Table \ref{paramtable} for a summary of important parameters.
 
@@
 \end{equation}
+where $\drat$ changes value across the ADC. More specifically, there exists an upstream and a downstream Hamiltonian
+\begin{align}
+&\Hup = 3 \pi \epsup \, H_3 - \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \, H_2 \\
+&\Hdn = 3 \pi \epsup \dratdn^{-1} \, H_3 - \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \dratdn \, H_2
+\end{align}
+$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$. Each Hamiltonian $\Ham^{\pm}$generates a corresponding Gibbs measure $\Gibbs^{\pm}$. (FILL IN DEFINITIONS)
 
-\begin{comment}
-Old Note (around May 2019): After giving it some deliberation, I believe we should use $\lamfac=1$.
-My reason is that the peak in the wave spectrum occurs at a frequency of $\freqp$ or wavelength of $\lam$.
-In the tKdV Gibbs measure, the spectrum decays monotonically, so that the peak is at the lowest resolved frequency (or largest resolved wavelength).
-Thus, we want $\lam$ to correspond to the largest resolved wavelength, i.e.~the length of the periodic domain in the tKdV framework.
-
-New Thought (July 2019): Actually, it is very possible that the bandwidth (2 Hz in the experiments) sets the value of $\lamfac$ that is most appropriate, since that bandwidth sets the decay rate of the spectrum.
-My thinking is that our experimental forcing does not really follow the upstream Gibbs measure all that well (because we did not have that in mind), but perhaps it best approximates the tail decay of some Gibbs measure.
-Perhaps, the lowest frequencies (i.e.
-those much slower than 2 Hz that are present in the theory but almost absent in the experiments) do not affect statistics that much.
-\end{comment}
-
+\vsp{10} 
 \subsection{Matching condition}
@@
 \begin{equation}
-\meanup{\Hp} = \meandn{\Hp}
+\meanup{\Hdn} = \meandn{\Hdn}
 \end{equation}

### 2019-07-15 WavesData-2019/LatexNotes/notes.tex
NM

@@
 
-%----------------------------------------------------------%
+%----------------------------------------------------------------------------%
 % Mathematical framework
-%----------------------------------------------------------%
+%----------------------------------------------------------------------------%
 \section{The truncated KdV framework}
 
+%----------------------------------------------------------------------------%
+% KdV
+%----------------------------------------------------------------------------%
 \subsection{The Korteweg–de Vries equation with variable depth}
@@
 
+%----------------------------------------------------------------------------%
 % Nondimensionalization
+%----------------------------------------------------------------------------%
 \subsection{Dimensionless variables}
@@
 Upstream depth			& $\dup$						& 12.5 cm \\
-Downstream depth			& $\ddn$						& 3 cm \\
+Downstream depth			& $\ddn$						& 3 cm (and varied) \\
 Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
@@
 Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
-Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 \\
+Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 (and varied) \\
 \hline \multicolumn{3} { c }{Parameters that change value across ADC} \\
@@
 
+%----------------------------------------------------------------------------%
+% Johnson alternative
+%----------------------------------------------------------------------------%
 \subsubsection{Alternative KdV formulation following Johnson}
@@
 
+%----------------------------------------------------------------------------%
+% Hamiltonian
+%----------------------------------------------------------------------------%
 \subsection{Hamiltonian structure}
@@
 \end{equation}
+TO DO: Add Fourier stuff including Parseval.
+\vsp{4}
 
@@
 \end{align}
-$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$. Each Hamiltonian $\Ham^{\pm}$generates a corresponding Gibbs measure $\Gibbs^{\pm}$. (FILL IN DEFINITIONS)
+$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$.
+Each Hamiltonian $\Ham^{\pm}$ generates a corresponding Gibbs measure $\Gibbs^{\pm}$ with corresponding ensemble means $\mean{}_{\pm}$.
+(FILL IN DEFINITIONS)
 
-\vsp{10} 
+%----------------------------------------------------------------------------%
+% Matching
+%----------------------------------------------------------------------------%
 \subsection{Matching condition}
@@
 \end{equation}
+which immediately gives
+\begin{equation}
+3 \pi \epsup \dratdn^{-1} \meandn{H_3} = \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \dratdn \left( \meandn{H_2} - \meanup{H_2} \right)
+\end{equation}
+or more simply
+\begin{equation}
+\label{H3H2}
+\meandn{H_3} = \frac{4 \pi^2}{9 \lamfac^2} \frac{\delup^2}{\epsup} \dratdn^2 \,\, \Delta \mean{H_2}
+\end{equation}
+where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
+To convert to dimensional, i.e.~experimental, values we use
+\begin{align}
+&\mean{u^3} = \frac{\mean{\eta^3}}{a^3} = 
+\frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
+&\mean{u_x^2} = \left(\frac{\lam}{a} \right)^2 \mean{\eta_x^2} 
+= \left(\frac{\lam}{a} \right)^2 \var(\eta_x)
+\end{align}
+Then \eqref{H3H2} gives the relationship
+\begin{equation}
+\frac{\skw(\eta)}{\Delta \var(\eta_x)} =
+3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
+\end{equation}
+In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude (as well as the square of the depth ratio).
+
+WORK TO DO HERE WITH LAMBDA
+
+\vsp{10}
+OLD VERSION: \\
 which immediately gives

### 2019-07-17 WavesData-2019/LatexNotes/notes.tex
NM

@@
 % Basic physical parameters and scales
-\newcommand{\depth}{h}
+\newcommand{\freqp}{f_p}
+\newcommand{\amp}{a}
+%\newcommand{\etastd}{\eta_{\text std}}
+% Parameters that change crossing the ADC
+\newcommand{\depth}{d}
 \newcommand{\dup}{\depth_{-}}
 \newcommand{\ddn}{\depth_{+}}
-\newcommand{\freqp}{f_p}
+\newcommand{\dupdn}{\depth_{\pm}}
 \newcommand{\lam}{\lambda}
 \newcommand{\lamup}{\lam_{-}}
+\newcommand{\lamdn}{\lam_{+}}
+\newcommand{\lamupdn}{\lam_{\pm}}
 \newcommand{\lamfac}{N}
-\newcommand{\amp}{a}
-%\newcommand{\etastd}{\eta_{\text std}}
+\newcommand{\drat}{D}
+\newcommand{\dratdn}{\drat_+}
+\newcommand{\dratupdn}{\drat_{\pm}}
 % Statistical quantities
@@
 \newcommand{\var}{\text{var}}
+\newcommand{\En}{\mathcal{E}}
 % Dimensionless parameters
@@
 \newcommand{\delup}{\delta_0}
-\newcommand{\drat}{D}
-\newcommand{\dratdn}{\drat_0}
+
 \newcommand{\ampp}{\mathcal{A}}
@@
 
+ \subsubsection{Most conventional nondimensionalization}
+\label{nondim1}
+
 We introduce the following characteristic scales
@@
 \label{lamdefn}
-&\lam = \sqrt{g \depth} / \freqp
+&\lamupdn = \sqrt{g \dupdn} / \freqp
 &&\mbox{\em characteristic wave length}
@@
 where $\mean{\cdot}$ indicates the mean of a quantity.
-Note that the characteristic wavelength $\lam = \lam(x)$ takes different values upstream and downstream of the ADC.
+Note that the characteristic wavelength $\lam = \lamupdn$ takes different values upstream and downstream of the ADC.
 We remark that experimental measurements indicate that the characteristic amplitude $\amp = \mean{\eta^2}^{1/2}$ is nearly the same on both values of the ADC.
@@
 &&\mbox{\em dimensionless surface displacement} \\
-&\tilde{x} = \frac{2 \pi \xi}{\lamfac \lam} = \frac{2 \pi (x-ct)}{\lamfac \lam}
+&\tilde{x} = \frac{\pi \xi}{\lamfac \lam} = \frac{\pi (x-ct)}{\lamfac \lam}
 &&\mbox{\em dimensionless position in moving frame} \\
@@
 \end{align}
-We will consider the physical domain $\xi \in [0, \lamfac \lam]$, which is an integer multiple of the characteristic wavelength where $\lamfac$ will be selected later (probably about 5).
-Above, we have chosen the scale to normalize position as $\lamfac \lam/(2\pi)$ so that the dimensionless domain can be taken (after a shift) to be $\tilde{x} \in [-\pi,\pi]$.
-Accordingly, we have also chosen the time scale as $\lamfac/\freqp$, so that a wave traveling at the local wavespeed $c = \lam \freqp = \sqrt{g \depth}$ crosses the physical domain $x \in [0, \lamfac \lam]$ in one unit of dimensionless time.
+We will consider the physical domain $\xi \in [-\lamfac \lam, \lamfac \lam]$, which is an integer multiple of the characteristic wavelength, where $\lamfac$ will be selected later (probably about 5).
+Above, we chosen to normalize position by the scale $L = \lamfac \lam/\pi$ so that the dimensionless domain is $\tilde{x} \in [-\pi,\pi]$.
+Accordingly, we have also chosen the time scale as $\lamfac/\freqp$, so that a wave traveling at the local wavespeed $c = \lam \freqp = \sqrt{g \depth}$ crosses the physical domain of length $2 \lamfac \lam$ in one unit of dimensionless time.
 Said differently, the dimensionless wave speed is $(2 \pi)^{-1}$ so that a wave crosses the domain $[-\pi,\pi]$ in unit time.
@@
 \end{align}
-Here, $\drat$ is the local dimensionless depth, which changes value crossing the ADC.
+Here, $\drat = \dratupdn$ is the local dimensionless depth, which changes value crossing the ADC.
 In particular, since \eqref{dimlessKdV} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$.
@@
 We will not use this alternative formulation.
+
+%----------------------------------------------------------------------------%
+% Energy nondimensionalization
+%----------------------------------------------------------------------------%
+\subsubsection{Energy nondimensionalization}
+
+Instead of scaling the surface displacement on the characteristic amplitude, a slightly different approach is to normalize the energy to unity.
+Consider waves in the domain $\xi \in [-\pi L, \pi L]$ where $L$ is a length-scale we define as
+\begin{equation}
+L = \lamfac \lam / \pi
+\end{equation}
+Consider the so-called energy of $\eta$
+\begin{equation}
+\En [\eta] = \frac{1}{2} \int_{-\pi L}^{\pi L} \eta^2 \dx = \pi L \mean{\eta^2} \equiv E
+\end{equation}
+We introduce the normalized variable
+\begin{align}
+&u = E^{-1/2}L^{1/2} \eta \\
+&\tilde{x} = \xi/L
+\end{align}
+Then, in dimensionless variables, we have
+\begin{equation}
+\En [u] = \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = \pi \mean{u^2} = \frac{\pi L}{E} \mean{\eta^2} = 1
+\end{equation}
+That is, the energy is normalized to unity.
+
+To compare to the previous formulation, we note that
+\begin{align}
+\amp \equiv \mean{\eta^2}^{1/2} = \sqrt{ \frac{E}{\pi L} }
+\end{align}
+So that the new normalization factor is
+\begin{equation}
+E^{1/2}L^{-1/2} = \sqrt{\pi} \amp
+\end{equation}
+In other words, the normalizations for surface displacement used here versus that in Section \ref{nondim}
+ differ by a factor of $\sqrt{\pi}$.
+Fortunately, the normalization for position is identical in the two.
+We do have to remember that the normalization factor for position $L = \lamfac \lam / \pi$ takes two different values on either side of the ADC.
+
+TO DO: How to scale time ....
 

### 2019-07-17 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \newcommand{\amp}{a}
+\newcommand{\ampscale}{A}
 %\newcommand{\etastd}{\eta_{\text std}}
@@
 Above, we chosen to normalize position by the scale $L = \lamfac \lam/\pi$ so that the dimensionless domain is $\tilde{x} \in [-\pi,\pi]$.
+
 Accordingly, we have also chosen the time scale as $\lamfac/\freqp$, so that a wave traveling at the local wavespeed $c = \lam \freqp = \sqrt{g \depth}$ crosses the physical domain of length $2 \lamfac \lam$ in one unit of dimensionless time.
 Said differently, the dimensionless wave speed is $(2 \pi)^{-1}$ so that a wave crosses the domain $[-\pi,\pi]$ in unit time.
+Note: perhaps it is reasonable to normalize time differently, so that it is 'slow' time.
+Looking in Johnson, there does not seem to be a standard way to do this, it is very haphazard.
+So lets leave it for now.
 
@@
 \begin{equation}
-\label{dimlessKdV}
-u_t + {3 \pi} \epsup \drat^{-1} \, u u_x + \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \drat \, u_{xxx} = 0
+\label{dimlessKdV1}
+u_t + \frac{3}{2} \pi \epsup \drat^{-1} \, u u_x + \frac{1}{6} \pi^3 \lamfac^{-2} \delup^2 \drat \, u_{xxx} = 0
 \qquad \text{for } x \in [-\pi,\pi]
@@
 Here, $\drat = \dratupdn$ is the local dimensionless depth, which changes value crossing the ADC.
-In particular, since \eqref{dimlessKdV} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$.
+In particular, since \eqref{dimlessKdV1} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$.
 Then we have
@@
 \end{align}
-So that the new normalization factor is
+So that the new normalization factor for $\eta$ is
 \begin{equation}
@@
 
-TO DO: How to scale time ....
+Lets do the nondimensionalization generally, so that we can easily make modifications to what scales to use. In general, we introduce dimensionless variables as
+\begin{align}
+&u = \eta / \ampscale \\
+&\tilde{x} = x / L \\
+&\tilde{t} = t / T
+\end{align}
+where $\ampscale, L, T$ are the scales for surface displacement (amplitude), length, and time respectively, which we are free to choose later. Then the dimensionless KdV equation is
+\begin{equation}
+u_t + \frac{3}{2} \left( \frac{c T \ampscale}{\depth L} \right) u u_x 
++ \frac{1}{6} \left( \frac{c \depth^2 T}{L^3} \right) u_{xxx} = 0
+\end{equation}
+where $c = \sqrt{g \depth}$.
+In particular, choosing $\ampscale = \amp$, $L = \lamfac \lam/\pi$, and $T = \lamfac/\freqp$ produces \eqref{dimlessKdV1}.
+To see this, one has to recall $\lam = c/\freqp = \sqrt{g \depth}/\freqp$, and also has to collect all of the factors of $\depth$ that appear directly and appear in $L$.
+
+Alternatively, we would like to choose $\ampscale = E^{1/2} L^{-1/2} = \sqrt{\pi} \amp$, $L = \lamfac \lam/ \pi$, and not sure yet about $T$.
+Ahh, I tried, but I cannot make sense of this because I cannot discern if $L$ in the PNAS paper and J Stat Phys paper is supposed to be dimensional or dimensionless.
+Also, I cannot tell of $L$ is supposed to change value crossing the ADC.
+The wavelength $\lam$ certainly changes value across the ADC, which suggests that $L$ should too, but on the other hand $L$ is treated as a constant throughout in the PNAS paper.
+So is $L$ the upstream value?
+I am not sure.
+
+% Possible Flaw?
+%Note: I believe this normalization is flawed because, for periodicity of $u$ to be maintained, the scale $L$ must change across the ADC.
+In particular, $L \sim \sqrt{\depth}$.
+However, the equations in our PNAS paper treat $L$ as constant throughout, even when $\drat$ changes.
+This seems to be a flaw to me.
+% Maybe not a flow if you remember that $E^{1/2} L^{-1/2} = \sqrt{pi} \amp$ is the same on either side. Thus as L changes because wavelength changes, E changes accordingly.
+
+Therefore, lets stick with the original normalization \eqref{dimlessKdV1}, repeated here
+\begin{equation}
+\label{dimlessKdV2}
+u_t + \frac{3}{2} \pi \epsup \drat^{-1} \, u u_x + \frac{1}{6} \pi^3 \lamfac^{-2} \delup^2 \drat \, u_{xxx} = 0
+\qquad \text{for } x \in [-\pi,\pi]
+\end{equation}
+If we need to draw a comparison to the PNAS paper, we set something like $\epsup = E^{1/2}L^{-3/2}$ and $\delup^2 = L^{-3}$, etc.
 
@@
 
-We now consider the Hamiltonian structure of the dimensionless variable-depth KdV equation \eqref{dimlessKdV}, over dimensionless domain $x \in [-\pi,\pi]$.
+We now consider the Hamiltonian structure of the dimensionless variable-depth KdV equation \eqref{dimlessKdV1}, over dimensionless domain $x \in [-\pi,\pi]$.
 First, the spatial average of any quantity is given by
@@
 
-Then the Hamiltonian of \eqref{dimlessKdV} is given by
+Then the Hamiltonian of \eqref{dimlessKdV1} is given by
 \begin{equation}
@@
 Each Hamiltonian $\Ham^{\pm}$ generates a corresponding Gibbs measure $\Gibbs^{\pm}$ with corresponding ensemble means $\mean{}_{\pm}$.
+
 (FILL IN DEFINITIONS)

### 2019-07-17 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \newcommand{\En}{\mathcal{E}}
+\newcommand{\Mo}{\mathcal{M}}
 % Dimensionless parameters
@@
 \newcommand{\delup}{\delta_0}
+% Fourier
+\newcommand{\uhat}{\hat{u}}
 
@@
 \newcommand{\Hdn}{\Ham^{+}}
+\newcommand{\Hupdn}{\Ham^{\pm}}
 \newcommand{\Gibbs}{\mathcal{G}}
@@
 \newcommand{\Gdn}{\Gibbs^{+}}
+\newcommand{\Gupdn}{\Gibbs^{\pm}}
+\newcommand{\thup}{\theta^{-}}
+\newcommand{\thdn}{\theta^{+}}
+\newcommand{\thupdn}{\theta^{\pm}}
 \newcommand{\meanup}[1]{\mean{#1}_{-}}
@@
 \label{dimlessKdV1}
-u_t + \frac{3}{2} \pi \epsup \drat^{-1} \, u u_x + \frac{1}{6} \pi^3 \lamfac^{-2} \delup^2 \drat \, u_{xxx} = 0
+u_t + \frac{3 \pi}{2} \epsup \drat^{-1} \, u u_x + \frac{\pi^3}{6 \lamfac^2} \delup^2 \drat \, u_{xxx} = 0
 \qquad \text{for } x \in [-\pi,\pi]
@@
 
-Therefore, lets stick with the original normalization \eqref{dimlessKdV1}, repeated here
-\begin{equation}
+Therefore, lets stick with the original normalization \eqref{dimlessKdV1}, recast here in different form
+\begin{align}
 \label{dimlessKdV2}
-u_t + \frac{3}{2} \pi \epsup \drat^{-1} \, u u_x + \frac{1}{6} \pi^3 \lamfac^{-2} \delup^2 \drat \, u_{xxx} = 0
-\qquad \text{for } x \in [-\pi,\pi]
-\end{equation}
+&u_t + C_3 \drat^{-1} \, u u_x + C_2 \drat \, u_{xxx} = 0
+\qquad \text{for } x \in [-\pi,\pi] \\
+&C_3 = \frac{3}{2} \pi \epsup  \\
+&C_2 = \frac{\pi^3 \delup^2}{6 \lamfac^2} 
+\end{align}
 If we need to draw a comparison to the PNAS paper, we set something like $\epsup = E^{1/2}L^{-3/2}$ and $\delup^2 = L^{-3}$, etc.
@@
 
-We now consider the Hamiltonian structure of the dimensionless variable-depth KdV equation \eqref{dimlessKdV1}, over dimensionless domain $x \in [-\pi,\pi]$.
+We now consider the Hamiltonian structure of the dimensionless variable-depth KdV equation \eqref{dimlessKdV2}, over dimensionless domain $x \in [-\pi,\pi]$.
 First, the spatial average of any quantity is given by
@@
 \begin{equation}
-M[u] \defeq \int_{-\pi}^{\pi} u \dx = 0
+\Mo[u] \defeq \int_{-\pi}^{\pi} u \dx = 0
 \end{equation}
@@
 \begin{equation}
-E[u] \defeq \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = \pi
+\En[u] \defeq \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = \pi
 \end{equation}
-TO DO: Add Fourier stuff including Parseval.
-\vsp{4}
+
+We introduce the Fourier series
+\begin{equation}
+u = \sum_{k=-\Lambda}^{\Lambda} \uhat_k e^{i k x}
+\end{equation}
+where
+\begin{equation}
+\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
+\end{equation}
+The above sum can be infinite $\Lambda = \infty$ or truncated at a finite wavenumber $\Lambda < \infty$.
+Further, since $u$ is real valued, we have $\uhat_{-k} = {\uhat_{k}}^*$.
+By Parseval's identity, we have
+\begin{equation}
+\En[u] = \pi = \frac{1}{2} \int u^2 \dx = \pi \sum_{k=-\Lambda}^{\Lambda} \abs{\uhat_k}^2 = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2
+\end{equation}
+% The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019, except that the total energy is one instead of pi.
 
@@
 \begin{equation}
-\Ham = 3 \pi \epsup \drat^{-1} \, H_3 - \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \drat \, H_2
+\Ham = C_3 \drat^{-1} \, H_3 - C_2 \drat \, H_2
 \end{equation}
@@
 \begin{align}
-&\Hup = 3 \pi \epsup \, H_3 - \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \, H_2 \\
-&\Hdn = 3 \pi \epsup \dratdn^{-1} \, H_3 - \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \dratdn \, H_2
+&\Hup = C_3 \, H_3 - C_2 \, H_2 \\
+&\Hdn = C_3 \dratdn^{-1} \, H_3 - C_2 \dratdn \, H_2
 \end{align}
-$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$.
-Each Hamiltonian $\Ham^{\pm}$ generates a corresponding Gibbs measure $\Gibbs^{\pm}$ with corresponding ensemble means $\mean{}_{\pm}$.
-
-(FILL IN DEFINITIONS)
+$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$. Each Hamiltonian $\Ham^{\pm}$ generates a corresponding Gibbs measure $\Gupdn$ 
+\begin{align}
+\Gupdn = C_{\thupdn} \exp(-\thupdn \Hupdn) \delta(\En - \pi)
+\end{align}
+where $\theta = \thupdn$ is the inverse temperature and $C_{\theta}$ a constant (the partition function) that depends on $\theta$. Each measure has corresponding ensemble mean $\mean{}_{\pm}$.
 
@@
 \begin{equation}
-3 \pi \epsup \dratdn^{-1} \meandn{H_3} = \frac{4 \pi^3}{3 \lamfac^2} \delup^2 \dratdn \left( \meandn{H_2} - \meanup{H_2} \right)
+C_3 \dratdn^{-1} \meandn{H_3} = C_2 \dratdn \left( \meandn{H_2} - \meanup{H_2} \right)
 \end{equation}
-or more simply
+Simplifying gives
 \begin{equation}
 \label{H3H2}
-\meandn{H_3} = \frac{4 \pi^2}{9 \lamfac^2} \frac{\delup^2}{\epsup} \dratdn^2 \,\, \Delta \mean{H_2}
+\meandn{H_3} = \frac{C_2}{C_3} \dratdn^2 \,\, \Delta \mean{H_2}
 \end{equation}
 where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
-To convert to dimensional, i.e.~experimental, values we use
+To convert to dimensional, i.e.~experimental, values we note
 \begin{align}
@@
 \frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
-&\mean{u_x^2} = \left(\frac{\lam}{a} \right)^2 \mean{\eta_x^2} 
-= \left(\frac{\lam}{a} \right)^2 \var(\eta_x)
+&\mean{u_x^2} = \left(\frac{\lamupdn}{a} \right)^2 \mean{\eta_x^2} 
+= \left(\frac{\lamupdn}{a} \right)^2 \var(\eta_x)
 \end{align}
+Therefore
+\begin{equation}
+\meandn{H_3} = \frac{\pi}{3} \skw_+(\eta)
+\end{equation}
+and more subtle
+\begin{equation}
+\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2} = 
+\frac{\pi}{\amp^2} \left( \lamdn \var_+(\eta_x) - \lamup \var_{-}(\eta_x) \right) =
+\frac{\pi}{\amp^2} \Delta \left( \lam \var(\eta_x) \right)
+\end{equation}
 Then \eqref{H3H2} gives the relationship
 \begin{equation}
-\frac{\skw(\eta)}{\Delta \var(\eta_x)} =
-3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
+\frac{\skw(\eta)} {\Delta \left( \lam \var(\eta_x) \right) } =
+\frac{3 C_2}{C_3} \amp^{-2} \dratdn^2 = \frac{\pi^2 \delup^2}{9 \lamfac^2} \
+%3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
 \end{equation}
+FINISH
+
 In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude (as well as the square of the depth ratio).

### 2019-07-17 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \newcommand{\delup}{\delta_0}
-% Fourier
+% Hamiltonian and Gibbs stuff
 \newcommand{\uhat}{\hat{u}}
-
-\newcommand{\ampp}{\mathcal{A}}
-% Hamiltonian and Gibbs stuff
 \newcommand{\Ham}{H}
@@
 
-Therefore, lets stick with the original normalization \eqref{dimlessKdV1}, recast here in different form
+%----------------------------------------------------------------------------%
+%  Normalization to use
+%----------------------------------------------------------------------------%
+\subsubsection{Normalization to use in the paper}
+
+We will essential use the original normalization in \eqref{dimlessKdV1}, except with the slight modification that $\eta = \sqrt{\pi} \amp u$, so that the energy of $u$ is normalized to unity.
+This choice will be consistent with our PNAS paper and the J Stat Phys paper of Majda and Qi.
+Thus the dimensionless variables are
+\begin{align}
+&u = \eta/(\sqrt{\pi} \amp)
+&&\mbox{\em dimensionless surface displacement} \\
+&\tilde{x} = \frac{\pi \xi}{\lamfac \lam} = \frac{\pi (x-ct)}{\lamfac \lam}
+&&\mbox{\em dimensionless position in moving frame} \\
+& \tilde{t} = t \freqp / \lamfac
+&&\mbox{\em dimensionless time}
+\end{align}
+The explanation will be something like: instead of the most conventional choice, we scale displacement on $\sqrt{\pi} \amp$ so that energy is normalize to unity.
+This choice will be the most convenient for the Hamiltonian framework that we will rely on.
+Maybe we will have to scale time differently too to get the desired powers of $\drat$, but that will make no difference for now.
+Thus, the dimensionless KdV is given by
 \begin{align}
@@
 \qquad \text{for } x \in [-\pi,\pi] \\
-&C_3 = \frac{3}{2} \pi \epsup  \\
+&C_3 = \frac{3}{2} \pi^{3/2} \epsup  \\
 &C_2 = \frac{\pi^3 \delup^2}{6 \lamfac^2} 
 \end{align}
-If we need to draw a comparison to the PNAS paper, we set something like $\epsup = E^{1/2}L^{-3/2}$ and $\delup^2 = L^{-3}$, etc.
+When we draw a comparison to the PNAS paper, it will be something like $\epsup = E^{1/2}L^{-3/2}$ and $\delup^2 = L^{-3}$, etc.
 
@@
 \end{equation}
-Next, the definition of the characteristic amplitude in \eqref{ampdefn} implies that $\mean{u^2} = 1$. Consequently, the energy of $u$ is fixed as
+Next, the definition of the characteristic amplitude in \eqref{ampdefn} implies that $\mean{u^2} = \pi^{-1}$. Consequently, the energy of $u$ is fixed as
 \begin{equation}
-\En[u] \defeq \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = \pi
+\En[u] \defeq \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
 \end{equation}
@@
 \begin{equation}
-\En[u] = \pi = \frac{1}{2} \int u^2 \dx = \pi \sum_{k=-\Lambda}^{\Lambda} \abs{\uhat_k}^2 = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2
+\En[u] = 1 = \frac{1}{2} \int u^2 \dx = \pi \sum_{k=-\Lambda}^{\Lambda} \abs{\uhat_k}^2 = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2
 \end{equation}
-% The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019, except that the total energy is one instead of pi.
+% The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
 
@@
 \frac{\skw(\eta)} {\Delta \left( \lam \var(\eta_x) \right) } =
-\frac{3 C_2}{C_3} \amp^{-2} \dratdn^2 = \frac{\pi^2 \delup^2}{9 \lamfac^2} \
-%3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
+\frac{3 C_2 \dratdn^2}{C_3 \amp^2}  =
+% OLD 3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
 \end{equation}
@@
 OLD VERSION: \\
-which immediately gives
-\begin{equation}
-\ampp D^{-3/2} \meandn{H_3} = \mu D^{1/2} \left( \meandn{H_2} - \meanup{H_2} \right)
-\end{equation}
-or equivalently
-\begin{equation}
-\label{H3H2}
-\meandn{H_3} = \frac{\mu D^{2}}{\ampp} \, \Delta \mean{H_2}
-\end{equation}
-where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
 

### 2019-07-17 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \newcommand{\ampscale}{A}
-%\newcommand{\etastd}{\eta_{\text std}}
+\newcommand{\etastd}{\eta_{\text std}}
 % Parameters that change crossing the ADC
@@
 \newcommand{\meandn}[1]{\mean{#1}_{+}}
+\newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
 
@@
 
-We will essential use the original normalization in \eqref{dimlessKdV1}, except with the slight modification that $\eta = \sqrt{\pi} \amp u$, so that the energy of $u$ is normalized to unity.
+We will essential use the original normalization in \eqref{dimlessKdV1}, except with the slight modification that $\eta = \sqrt{\pi} \etastd u$, so that the energy of $u$ is normalized to unity.
 This choice will be consistent with our PNAS paper and the J Stat Phys paper of Majda and Qi.
@@
 \begin{align}
-&u = \eta/(\sqrt{\pi} \amp)
+&u = \eta/(\sqrt{\pi} \etastd)
 &&\mbox{\em dimensionless surface displacement} \\
@@
 \end{align}
-The explanation will be something like: instead of the most conventional choice, we scale displacement on $\sqrt{\pi} \amp$ so that energy is normalize to unity.
+The explanation will be something like: instead of the most conventional choice, we scale displacement on $\sqrt{\pi} \etastd$ so that energy is normalize to unity.
 This choice will be the most convenient for the Hamiltonian framework that we will rely on.
@@
 \begin{align}
-&\mean{u^3} = \frac{\mean{\eta^3}}{a^3} = 
-\frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
-&\mean{u_x^2} = \left(\frac{\lamupdn}{a} \right)^2 \mean{\eta_x^2} 
-= \left(\frac{\lamupdn}{a} \right)^2 \var(\eta_x)
+&\mean{u^3} =  \frac{\mean{\eta^3}} {\pi^{3/2} \mean{\eta^2}^{3/2}} = \pi^{-3/2} \skw(\eta) \\
+&\mean{u_x^2} = \pi^{-1} \left(\frac{\lamupdn}{\etastd} \right)^2 \mean{\eta_x^2} 
+= \pi^{-1} \left(\frac{\lamupdn}{\etastd} \right)^2 \var(\eta_x)
 \end{align}
 Therefore
-\begin{equation}
-\meandn{H_3} = \frac{\pi}{3} \skw_+(\eta)
-\end{equation}
-and more subtle
+\begin{align}
+& \meandn{H_3} = \frac{\pi}{3} \mean{u^3} = 
+\frac{1}{3} \pi^{-1/2} \skw_+(\eta) \\
+& \meanupdn{H_2} = \pi \meanupdn{u_x^2} = 
+\left(\frac{\lamupdn}{\etastd} \right)^2 \var(\eta_x)
+\end{align}
+The second equation implies
 \begin{equation}
 \Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2} = 
-\frac{\pi}{\amp^2} \left( \lamdn \var_+(\eta_x) - \lamup \var_{-}(\eta_x) \right) =
-\frac{\pi}{\amp^2} \Delta \left( \lam \var(\eta_x) \right)
+\etastd^{-2} \left( \lamdn^2 \var_+(\eta_x) - \lamup^2 \var_{-}(\eta_x) \right) =
+\etastd^{-2} \Delta \left( \lam^2 \var(\eta_x) \right)
 \end{equation}
@@
 \begin{equation}
-\frac{\skw(\eta)} {\Delta \left( \lam \var(\eta_x) \right) } =
-\frac{3 C_2 \dratdn^2}{C_3 \amp^2}  =
-% OLD 3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
+\frac{\skw(\eta)} {\Delta \left( \lam^2 \var(\eta_x) \right) } =
+\frac{3 \pi^{1/2} C_2 \dratdn^2}{C_3 \etastd^2}  =
+\frac{\pi^2}{3 \lamfac^2} \delup^2 \epsup^{-1} \etastd^{-2} \dratdn^2 =
+\frac{\pi^2}{3 \lamfac^2} \lamup^{-2} \epsup^{-3} \dratdn^2
 \end{equation}
-FINISH
-
 In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude (as well as the square of the depth ratio).
-
-WORK TO DO HERE WITH LAMBDA
 

### 2019-07-18 WavesData-2019/LatexNotes/notes.tex
NM

@@
 % Statistical quantities
-\newcommand{\skw}{\text{skew}}
-\newcommand{\var}{\text{var}}
 \newcommand{\En}{\mathcal{E}}
 \newcommand{\Mo}{\mathcal{M}}
+\newcommand{\skw}{\text{skew}}
+\newcommand{\skwdn}{\skw_+}
+\newcommand{\var}{\text{var}}
+\newcommand{\varup}{\var_-}
+\newcommand{\vardn}{\var_+}
 % Dimensionless parameters
@@
 We will essential use the original normalization in \eqref{dimlessKdV1}, except with the slight modification that $\eta = \sqrt{\pi} \etastd u$, so that the energy of $u$ is normalized to unity.
-This choice will be consistent with our PNAS paper and the J Stat Phys paper of Majda and Qi.
+This choice will be consistent with our PNAS paper and the {\it J Stat Phys} paper of Majda and Qi.
 Thus the dimensionless variables are
@@
 \end{equation}
-An experimental observation is that the incoming skewness is small, thus
-\begin{equation}
-\meanup{H_3} \approx 0
-\end{equation}
-which immediately gives
-\begin{equation}
-C_3 \dratdn^{-1} \meandn{H_3} = C_2 \dratdn \left( \meandn{H_2} - \meanup{H_2} \right)
-\end{equation}
-Simplifying gives
+An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$. Inserting and simplifying gives
 \begin{equation}
@@
 & \meandn{H_3} = \frac{\pi}{3} \mean{u^3} = 
-\frac{1}{3} \pi^{-1/2} \skw_+(\eta) \\
+\frac{1}{3} \pi^{-1/2} \skwdn(\eta) \\
 & \meanupdn{H_2} = \pi \meanupdn{u_x^2} = 
@@
 \Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2} = 
-\etastd^{-2} \left( \lamdn^2 \var_+(\eta_x) - \lamup^2 \var_{-}(\eta_x) \right) =
+\etastd^{-2} \left( \lamdn^2 \vardn(\eta_x) - \lamup^2 \varup(\eta_x) \right) =
 \etastd^{-2} \Delta \left( \lam^2 \var(\eta_x) \right)
@@
 \end{equation}
-In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude (as well as the square of the depth ratio).
+Performing some additional manipulations, and noting that $(\lamdn/\lamup)^2 = \ddn/\dup = \dratdn$, gives
+\begin{equation}
+\frac{\skwdn(\eta)} {\dratdn \vardn(\eta_x) - \varup(\eta_x)} = 
+\frac{\pi^2}{3 \lamfac^2} \epsup^{-3} \dratdn^2
+\end{equation}
+In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
 
-\vsp{10}
-OLD VERSION: \\
-
-To convert to dimensional, i.e.~experimental, values we use
-\begin{align}
-& \intt u^3 \dx = \frac{\mean{\eta^3}}{a^3} = 
-\frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
-& \intt u_x^2 \dx = \left(\frac{\lambda_0}{a} \right)^2 \mean{\eta_x^2} 
-= \left(\frac{\lambda_0}{a} \right)^2 \var(\eta_x)
-\end{align}
-Then \eqref{H3H2} gives the relationship
-\begin{equation}
-\frac{\skw(\eta)}{\Delta \var(\eta_x)} =
-3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
-\end{equation}
-In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude (as well as the square of the depth ratio).
+% Comparison to old calculation:
+% The old calculation used mu = 4 pi^2 / (9 N^2), but with a factor of two difference in the definition of N.
+The resulting coefficient on the RHS was 3 mu, which agrees with the RHS above once the factor of two is taken into account.
 

### 2019-07-18 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \label{H3H2}
-\meandn{H_3} = \frac{C_2}{C_3} \dratdn^2 \,\, \Delta \mean{H_2}
+\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = \frac{C_2}{C_3} \dratdn^2
 \end{equation}
-where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
-To convert to dimensional, i.e.~experimental, values we note
+%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
+To convert to dimensional, i.e.~experimental, values we first note that
 \begin{align}
-&\mean{u^3} =  \frac{\mean{\eta^3}} {\pi^{3/2} \mean{\eta^2}^{3/2}} = \pi^{-3/2} \skw(\eta) \\
-&\mean{u_x^2} = \pi^{-1} \left(\frac{\lamupdn}{\etastd} \right)^2 \mean{\eta_x^2} 
-= \pi^{-1} \left(\frac{\lamupdn}{\etastd} \right)^2 \var(\eta_x)
+\mean{H_3} = \frac{\pi}{3} \mean{u^3} = 
+\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
 \end{align}
-Therefore
+Second, to convert the surface slope, we note
 \begin{align}
-& \meandn{H_3} = \frac{\pi}{3} \mean{u^3} = 
-\frac{1}{3} \pi^{-1/2} \skwdn(\eta) \\
-& \meanupdn{H_2} = \pi \meanupdn{u_x^2} = 
-\left(\frac{\lamupdn}{\etastd} \right)^2 \var(\eta_x)
+\pd{u}{\tilde{x}} = \frac{\lamfac \lam}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
 \end{align}
-The second equation implies
+where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
+Hence, we have
+\begin{align}
+&\mean{H_2} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
+\left( \frac{\lamfac \lam}{\pi \etastd} \right)^2 \var(\eta_x)
+\end{align}
+Algebra gives
 \begin{equation}
-\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2} = 
-\etastd^{-2} \left( \lamdn^2 \vardn(\eta_x) - \lamup^2 \varup(\eta_x) \right) =
-\etastd^{-2} \Delta \left( \lam^2 \var(\eta_x) \right)
+\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = 
+\frac{\pi^{3/2}} {3 \lamfac^2 \etastd^2} 
+\left( \frac{\skwdn(\eta)}{ \lamdn^2 \vardn(\eta_x) - \lamup^2 \varup(\eta_x)} \right) = 
+\frac{\pi^{3/2}} {3 \lamfac^2 \etastd^2 \lamup^2} 
+\left( \frac{\skwdn(\eta)} {\dratdn \vardn(\eta_x) - \varup(\eta_x)} \right)
 \end{equation}
+where in the second equality, we have used the fact that
+\begin{equation}
+\frac{\lamdn^2}{\lamup^2} = \frac{\ddn}{\dup} = \dratdn
+\end{equation}
+
+\vsp{5}
+
 Then \eqref{H3H2} gives the relationship

### 2019-07-18 WavesData-2019/LatexNotes/notes.tex
NMM

@@
 u_t + \frac{3}{2} \left( \frac{c T \ampscale}{\depth L} \right) u u_x 
-+ \frac{1}{6} \left( \frac{c \depth^2 T}{L^3} \right) u_{xxx} = 0
++ \frac{1}{6} \left( \frac{c T \depth^2}{L^3} \right) u_{xxx} = 0
 \end{equation}
@@
 \label{H3H2}
-\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = \frac{C_2}{C_3} \dratdn^2
+\frac{\meandn{H_3^+}} {\meandn{H_2^+} - \meanup{H_2^+}} = \frac{C_2}{C_3} \dratdn^2
 \end{equation}
@@
 \begin{align}
-\mean{H_3} = \frac{\pi}{3} \mean{u^3} = 
+\mean{H_3^+} = \frac{\pi}{3} \mean{u^3} = 
 \frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
 \end{align}
-Second, to convert the surface slope, we note
+Second, to convert the downstream surface slope, we note
 \begin{align}
-\pd{u}{\tilde{x}} = \frac{\lamfac \lam}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
+\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
 \end{align}
 where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
+Note: I believe it is important here to use $\lamdn$ since we are matching $\Hdn$.
 Hence, we have
 \begin{align}
-&\mean{H_2} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
-\left( \frac{\lamfac \lam}{\pi \etastd} \right)^2 \var(\eta_x)
+&\mean{H_2^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
+\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
 \end{align}
+%
+Then algebra gives
+\begin{equation}
+\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = 
+\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
+\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
+\end{equation}
+Also we have the simplification
+\begin{equation}
+\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
+\end{equation}
+Hence, \eqref{H3H2} gives
+\begin{equation}
+\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^4
+\end{equation}
+In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
+% Note: previously I had D^2 on the RHS. The power of 4 comes from using lamdn consistently in the nondimensionalization.
+
+\subsection{Derivation with $\lam$ changing across ADC}
+This derivation must be faulty, but here it is so that I remember it.
+Aha, I think the error is as follows: The matching condition is for $\Hdn$ only, using the two measures. So you have to use the downstream scales, in particular $\lamdn$.
+
+\vsp{1}
+Anyway, here is the flawed derivation, so that we remember it.
 Algebra gives
@@
 \frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = 
-\frac{\pi^{3/2}} {3 \lamfac^2 \etastd^2} 
+\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2} 
 \left( \frac{\skwdn(\eta)}{ \lamdn^2 \vardn(\eta_x) - \lamup^2 \varup(\eta_x)} \right) = 
-\frac{\pi^{3/2}} {3 \lamfac^2 \etastd^2 \lamup^2} 
+\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamup^2} 
 \left( \frac{\skwdn(\eta)} {\dratdn \vardn(\eta_x) - \varup(\eta_x)} \right)
@@
 \end{equation}
-
-\vsp{5}
-
-Then \eqref{H3H2} gives the relationship
+Also we have the simplification
 \begin{equation}
-\frac{\skw(\eta)} {\Delta \left( \lam^2 \var(\eta_x) \right) } =
-\frac{3 \pi^{1/2} C_2 \dratdn^2}{C_3 \etastd^2}  =
-\frac{\pi^2}{3 \lamfac^2} \delup^2 \epsup^{-1} \etastd^{-2} \dratdn^2 =
-\frac{\pi^2}{3 \lamfac^2} \lamup^{-2} \epsup^{-3} \dratdn^2
+\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
 \end{equation}
-Performing some additional manipulations, and noting that $(\lamdn/\lamup)^2 = \ddn/\dup = \dratdn$, gives
+Hence, \eqref{H3H2} gives
 \begin{equation}
-\frac{\skwdn(\eta)} {\dratdn \vardn(\eta_x) - \varup(\eta_x)} = 
-\frac{\pi^2}{3 \lamfac^2} \epsup^{-3} \dratdn^2
+\frac{\skwdn(\eta)} {\dratdn \vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^2
 \end{equation}
 In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
-
-% Comparison to old calculation:
-% The old calculation used mu = 4 pi^2 / (9 N^2), but with a factor of two difference in the definition of N.
-The resulting coefficient on the RHS was 3 mu, which agrees with the RHS above once the factor of two is taken into account.
 

### 2019-07-18 WavesData-2019/LatexNotes/notes.tex
NM

@@
 \newcommand{\eps}{\varepsilon}
-\newcommand{\defeq}{\mathrel{\mathop:}=}
+\newcommand{\dx}{\, dx}
+
+%\newcommand{\defeq}{\mathrel{\mathop:}=}
 
@@
 
-% Dashed integral
-\def\Xint#1{\mathchoice
-   {\XXint\displaystyle\textstyle{#1}}%
-   {\XXint\textstyle\scriptstyle{#1}}%
-   {\XXint\scriptstyle\scriptscriptstyle{#1}}%
-   {\XXint\scriptscriptstyle\scriptscriptstyle{#1}}%
-   \!\int}
-\def\XXint#1#2#3{{\setbox0=\hbox{$#1{#2#3}{\int}$}
-     \vcenter{\hbox{$#2#3$}}\kern-.5\wd0}}
-\def\ddashint{\Xint=}
-\def\dashint{\Xint-}
-\newcommand{\intt}{\dashint}%_0^{2 \pi}}
-\newcommand{\dx}{\, dx}
-
 %----------------------------------------------%
@@
 \author{
-C. Tyler Bolles\thanks{University of Michigan},
+C.~Tyler Bolles\thanks{University of Michigan},
 Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
@@
 \begin{equation}
-\Mo[u] \defeq \int_{-\pi}^{\pi} u \dx = 0
+\Mo[u] \equiv \int_{-\pi}^{\pi} u \dx = 0
 \end{equation}
@@
 \begin{equation}
-\En[u] \defeq \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
+\En[u] \equiv \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
 \end{equation}
@@
 \begin{equation}
-\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^4
+\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3
 \end{equation}
 In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
-% Note: previously I had D^2 on the RHS. The power of 4 comes from using lamdn consistently in the nondimensionalization.
+% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.
 

### 2019-07-18 WavesData-2019/LatexNotes/notes.tex
NM

@@
 
-%----------------------------------------------%
+%----------------------------------------------------------------------------%
 %% LATEX DEFINITIONS
-%----------------------------------------------%
-
+%----------------------------------------------------------------------------%
 % Basic editing
@@
 \newcommand{\dx}{\, dx}
-
-%\newcommand{\defeq}{\mathrel{\mathop:}=}
-
 % Basic physical parameters and scales
@@
 \newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
+%----------------------------------------------------------------------------%
 
@@
 Peak forcing frequency		& $f_p$						& 2 Hz \\
+Characteristic wave amplitude	& $\etastd = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
 Upstream depth			& $\dup$						& 12.5 cm \\
 Downstream depth			& $\ddn$						& 3 cm (and varied) \\
-Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
 Upstream wavelength		& $\lamup = \sqrt{g \dup}/f_p$		& 56 cm \\
-Amplitude-to-depth ratio		& $\epsup = a / \dup$			& 0.0024--0.024 \\
+Amplitude-to-depth ratio		& $\epsup = \etastd / \dup$			& 0.0024--0.024 \\
 Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
@@
 
-\subsection{Derivation with $\lam$ changing across ADC}
+\subsection{Derivation with $\lam$ changing across ADC (Obselete)}
 This derivation must be faulty, but here it is so that I remember it.

### 2019-07-18 paper2_JNLS/paper.tex
NM

--- 2019-07-18 WavesData-2019/LatexNotes/notes.tex
@@
 \newcommand{\freqp}{f_p}
-\newcommand{\amp}{a}
+%\newcommand{\amp}{a}
 \newcommand{\ampscale}{A}
@@
 
-%----------------------------------------------------------------------------%
+\section{Introduction}
+
+%----------------------------------------------------------%
 % Mathematical framework
-%----------------------------------------------------------------------------%
+%----------------------------------------------------------%
 \section{The truncated KdV framework}
@@
 Most often, we will consider waves moving into shallower depth, so that $\dup > \ddn$.
-The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$, which gives rise to peak wavelength $\lam = c/\freqp = \sqrt{g \depth} / \freqp$.
+
+The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$, which gives rise to the following quantities
+\begin{align}
+&c = \sqrt{g \depth}
+&&\mbox{\em local wave speed} \\
+&\lam = c/\freqp = \sqrt{g \depth} / \freqp
+&&\mbox{\em local peak wave length} \\
+&\etastd = \mean{\eta^2}^{1/2} 
+&&\mbox{\em surface displacement standard deviation} \\
+\end{align}
+where $\mean{\cdot}$ indicates the mean of a quantity. 
+Note that the characteristic wave speed $c = c_{\pm}$ and wavelength $\lam = \lamupdn$ each take different values upstream and downstream of the ADC.
+We remark that experimental measurements indicate that $\etastd = \mean{\eta^2}^{1/2}$ is nearly the same on both values of the ADC.
+Hence, we will not distinguish between upstream and downstream values of $\etastd$.
 
@@
 %----------------------------------------------------------------------------%
-\subsection{Dimensionless variables}
-\label{nondim}
+\subsection{Nondimensionalization}
 
@@
 Downstream depth			& $\ddn$						& 3 cm (and varied) \\
-Upstream wavelength		& $\lamup = \sqrt{g \dup}/f_p$		& 56 cm \\
+Upstream wavelength		& $\lamup = \sqrt{g \dup}/f_p$		& 55 cm \\
+Upstream wavelength		& $\lamdn = \sqrt{g \ddn}/f_p$		& 27 cm \\
+%
 Amplitude-to-depth ratio		& $\epsup = \etastd / \dup$			& 0.0024--0.024 \\
 Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
-Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 (and varied) \\
-\hline \multicolumn{3} { c }{Parameters that change value across ADC} \\
-\hline Description & Notation and definition \\
-\hline
-Local depth			& $\depth$			\\
-Local wavespeed		& $c = \sqrt{g \depth}$	\\
-Local peak wavelength	& $\lam = \sqrt{gd}/f_p$	\\
-Local dimensionless depth		& $\drat = \depth/\dup$	
+Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 (and varied)
 \end{tabular}
@@
 
- \subsubsection{Most conventional nondimensionalization}
-\label{nondim1}
-
-We introduce the following characteristic scales
+Lets do the nondimensionalization generally, so that we can easily make modifications if needed. In general, we introduce dimensionless variables as
 \begin{align}
-\label{ampdefn}
-&\amp = \mean{\eta^2}^{1/2} 
-&&\mbox{\em characteristic wave amplitude} \\
-\label{lamdefn}
-&\lamupdn = \sqrt{g \dupdn} / \freqp
-&&\mbox{\em characteristic wave length}
-\end{align}
-where $\mean{\cdot}$ indicates the mean of a quantity.
-Note that the characteristic wavelength $\lam = \lamupdn$ takes different values upstream and downstream of the ADC.
-We remark that experimental measurements indicate that the characteristic amplitude $\amp = \mean{\eta^2}^{1/2}$ is nearly the same on both values of the ADC.
-Hence, we will not distinguish between upstream and downstream values of $\amp$.
-We introduce the following dimensionless variables
-\begin{align}
-&u = \eta/\amp
+&u = \eta / \ampscale
 &&\mbox{\em dimensionless surface displacement} \\
-&\tilde{x} = \frac{\pi \xi}{\lamfac \lam} = \frac{\pi (x-ct)}{\lamfac \lam}
-&&\mbox{\em dimensionless position in moving frame} \\
-& \tilde{t} = t \freqp / \lamfac
+&\tilde{x} = \xi / L
+&&\mbox{\em dimensionless position (moving frame)} \\
+&\tilde{t} = t / T
 &&\mbox{\em dimensionless time}
 \end{align}
-We will consider the physical domain $\xi \in [-\lamfac \lam, \lamfac \lam]$, which is an integer multiple of the characteristic wavelength, where $\lamfac$ will be selected later (probably about 5).
-Above, we chosen to normalize position by the scale $L = \lamfac \lam/\pi$ so that the dimensionless domain is $\tilde{x} \in [-\pi,\pi]$.
-
-Accordingly, we have also chosen the time scale as $\lamfac/\freqp$, so that a wave traveling at the local wavespeed $c = \lam \freqp = \sqrt{g \depth}$ crosses the physical domain of length $2 \lamfac \lam$ in one unit of dimensionless time.
-Said differently, the dimensionless wave speed is $(2 \pi)^{-1}$ so that a wave crosses the domain $[-\pi,\pi]$ in unit time.
-Note: perhaps it is reasonable to normalize time differently, so that it is 'slow' time.
-Looking in Johnson, there does not seem to be a standard way to do this, it is very haphazard.
-So lets leave it for now.
-
-%----------------------------------------------------------------------------%
-% COMMENT ON LAMFAC
-\begin{comment}
-Old Note (around May 2019): After giving it some deliberation, I believe we should use $\lamfac=1$.
-My reason is that the peak in the wave spectrum occurs at a frequency of $\freqp$ or wavelength of $\lam$.
-In the tKdV Gibbs measure, the spectrum decays monotonically, so that the peak is at the lowest resolved frequency (or largest resolved wavelength).
-Thus, we want $\lam$ to correspond to the largest resolved wavelength, i.e.~the length of the periodic domain in the tKdV framework.
-
-New Thought (July 2019): Actually, it is very possible that the bandwidth (2 Hz in the experiments) sets the value of $\lamfac$ that is most appropriate, since that bandwidth sets the decay rate of the spectrum.
-My thinking is that our experimental forcing does not really follow the upstream Gibbs measure all that well (because we did not have that in mind), but perhaps it best approximates the tail decay of some Gibbs measure.
-Perhaps, the lowest frequencies (i.e.
-those much slower than 2 Hz that are present in the theory but almost absent in the experiments) do not affect statistics that much.
-\end{comment}
-%----------------------------------------------------------------------------%
-
-Dropping the tildes, the dimensionless KdV equation takes the form
+where the characteristic scales $\ampscale, L, T$ for wave amplitude, length, and time respectively, are free for us to choose.
+Then the dimensionless KdV equation is
 \begin{equation}
-\label{dimlessKdV1}
-u_t + \frac{3 \pi}{2} \epsup \drat^{-1} \, u u_x + \frac{\pi^3}{6 \lamfac^2} \delup^2 \drat \, u_{xxx} = 0
-\qquad \text{for } x \in [-\pi,\pi]
-\end{equation}
-where dimensionless variables are as follows
-\begin{align}
-&\epsup = \amp/\dup \, , \\
-&\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g} \, \\
-&\drat = {\depth}/{\dup} \, .
-\end{align}
-Here, $\drat = \dratupdn$ is the local dimensionless depth, which changes value crossing the ADC.
-In particular, since \eqref{dimlessKdV1} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$.
-Then we have
-\begin{equation}
-\drat = 
-\begin{cases}
-1 		&\quad \mbox{for } t<0 \\
-\dratdn 	&\quad \mbox{for } t>0
-\end{cases}
-\end{equation}
-where $\dratdn = \ddn/\dup$.
-See Table \ref{paramtable} for a summary of important parameters.
-
-\begin{comment}
-{\bf Sidenote}: In the case of constant depth, if one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
-\begin{equation}
-2 u_t + 3 \eps u u_x + \frac{\delta^2}{3} u_{xxx} = 0
-\end{equation}
-where $\eps = a/h$ and $\delta = h/\lam$.
-\end{comment}
-
-%----------------------------------------------------------------------------%
-% Johnson alternative
-%----------------------------------------------------------------------------%
-\subsubsection{Alternative KdV formulation following Johnson}
-Section \ref{nondim} makes the implicit assumption of continuous surface displacement crossing the ADC.
-This is a very physically motivated assumption, but there is a small amount of debate surrounding the assumption.
-An alternative is to appeal to Green's law and assume the quantity $\depth^{1/4} \eta$ to be continuous across the ADC, which necessarily implies a small discontinuity in the surface (which of course cannot be true literally), as is the approach taken in Johnson't book.
-The dimensional KdV equation would then be expressed as
-\begin{equation}
-2 (\depth^{1/4} \eta)_t + \frac{3 c}{\depth^{5/4}} (\depth^{1/4} \eta) (\depth^{1/4} \eta)_{\xi} + \frac{c \depth^2}{3} (\depth^{1/4} \eta)_{\xi \xi \xi} = 0
-\end{equation}
-Then, the dimensionless state variable is instead 
-\begin{equation}
-u = \frac{\depth^{1/4} \, \eta}{ \dup^{1/4} \, \amp}
-\end{equation}
-and the dimensionless KdV equation takes the form
-\begin{equation}
-u_t + \frac{3 \pi}{\lamfac} \epsup \drat^{-5/4} \, u u_x + \frac{4 \pi^3}{3 \lamfac^3} \delup^2 \drat \, u_{xxx} = 0
-\qquad \text{for } x \in [-\pi,\pi]
-\end{equation}
-Notice the only difference is a small change in the power of $\drat$ in the second term (changes from $\drat^{-1}$ to $\drat^{-5/4}$).
-We will not use this alternative formulation.
-
-%----------------------------------------------------------------------------%
-% Energy nondimensionalization
-%----------------------------------------------------------------------------%
-\subsubsection{Energy nondimensionalization}
-
-Instead of scaling the surface displacement on the characteristic amplitude, a slightly different approach is to normalize the energy to unity.
-Consider waves in the domain $\xi \in [-\pi L, \pi L]$ where $L$ is a length-scale we define as
-\begin{equation}
-L = \lamfac \lam / \pi
-\end{equation}
-Consider the so-called energy of $\eta$
-\begin{equation}
-\En [\eta] = \frac{1}{2} \int_{-\pi L}^{\pi L} \eta^2 \dx = \pi L \mean{\eta^2} \equiv E
-\end{equation}
-We introduce the normalized variable
-\begin{align}
-&u = E^{-1/2}L^{1/2} \eta \\
-&\tilde{x} = \xi/L
-\end{align}
-Then, in dimensionless variables, we have
-\begin{equation}
-\En [u] = \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = \pi \mean{u^2} = \frac{\pi L}{E} \mean{\eta^2} = 1
-\end{equation}
-That is, the energy is normalized to unity.
-
-To compare to the previous formulation, we note that
-\begin{align}
-\amp \equiv \mean{\eta^2}^{1/2} = \sqrt{ \frac{E}{\pi L} }
-\end{align}
-So that the new normalization factor for $\eta$ is
-\begin{equation}
-E^{1/2}L^{-1/2} = \sqrt{\pi} \amp
-\end{equation}
-In other words, the normalizations for surface displacement used here versus that in Section \ref{nondim}
- differ by a factor of $\sqrt{\pi}$.
-Fortunately, the normalization for position is identical in the two.
-We do have to remember that the normalization factor for position $L = \lamfac \lam / \pi$ takes two different values on either side of the ADC.
-
-Lets do the nondimensionalization generally, so that we can easily make modifications to what scales to use. In general, we introduce dimensionless variables as
-\begin{align}
-&u = \eta / \ampscale \\
-&\tilde{x} = x / L \\
-&\tilde{t} = t / T
-\end{align}
-where $\ampscale, L, T$ are the scales for surface displacement (amplitude), length, and time respectively, which we are free to choose later. Then the dimensionless KdV equation is
-\begin{equation}
-u_t + \frac{3}{2} \left( \frac{c T \ampscale}{\depth L} \right) u u_x 
+u_t + \frac{3}{2} \left( \frac{c T \ampscale}{L \depth} \right) u u_x 
 + \frac{1}{6} \left( \frac{c T \depth^2}{L^3} \right) u_{xxx} = 0
 \end{equation}
-where $c = \sqrt{g \depth}$.
-In particular, choosing $\ampscale = \amp$, $L = \lamfac \lam/\pi$, and $T = \lamfac/\freqp$ produces \eqref{dimlessKdV1}.
-To see this, one has to recall $\lam = c/\freqp = \sqrt{g \depth}/\freqp$, and also has to collect all of the factors of $\depth$ that appear directly and appear in $L$.
 
-Alternatively, we would like to choose $\ampscale = E^{1/2} L^{-1/2} = \sqrt{\pi} \amp$, $L = \lamfac \lam/ \pi$, and not sure yet about $T$.
-Ahh, I tried, but I cannot make sense of this because I cannot discern if $L$ in the PNAS paper and J Stat Phys paper is supposed to be dimensional or dimensionless.
-Also, I cannot tell of $L$ is supposed to change value crossing the ADC.
-The wavelength $\lam$ certainly changes value across the ADC, which suggests that $L$ should too, but on the other hand $L$ is treated as a constant throughout in the PNAS paper.
-So is $L$ the upstream value?
-I am not sure.
+In particular, we make the following choices
+\begin{align}
+&\ampscale = \pi^{1/2} \, \etastd \\
+&L = \lamfac \lam / \pi \\
+&T = \lamfac / \freqp
+\end{align}
+where $\lamfac$ is an integer to be chosen later.
+In the physical domain, we will consider an integer multiple of the characteristic wavelength $\xi \in [-\lamfac \lam, \lamfac \lam]$, so that the state variable $u$ is periodic over the  dimensionless domain $\tilde{x} \in [-\pi, \pi]$.
+We have scaled time accordingly.
+Finally, our choice of the scale for wave amplitude $\ampscale$ is made to normalize the energy of $u$ to unity.
+The Hamiltonian framework that we will use relies heavily on the energy, and hence it is more convenient to have the energy normalized to unity rather than making the more common choice of $\etastd$ by itself as the scale for wave amplitude.
 
-% Possible Flaw?
-%Note: I believe this normalization is flawed because, for periodicity of $u$ to be maintained, the scale $L$ must change across the ADC.
-In particular, $L \sim \sqrt{\depth}$.
-However, the equations in our PNAS paper treat $L$ as constant throughout, even when $\drat$ changes.
-This seems to be a flaw to me.
-% Maybe not a flow if you remember that $E^{1/2} L^{-1/2} = \sqrt{pi} \amp$ is the same on either side. Thus as L changes because wavelength changes, E changes accordingly.
-
-%----------------------------------------------------------------------------%
-%  Normalization to use
-%----------------------------------------------------------------------------%
-\subsubsection{Normalization to use in the paper}
-
-We will essential use the original normalization in \eqref{dimlessKdV1}, except with the slight modification that $\eta = \sqrt{\pi} \etastd u$, so that the energy of $u$ is normalized to unity.
-This choice will be consistent with our PNAS paper and the {\it J Stat Phys} paper of Majda and Qi.
-Thus the dimensionless variables are
+With these choices, the dimensionless KdV is
 \begin{align}
-&u = \eta/(\sqrt{\pi} \etastd)
-&&\mbox{\em dimensionless surface displacement} \\
-&\tilde{x} = \frac{\pi \xi}{\lamfac \lam} = \frac{\pi (x-ct)}{\lamfac \lam}
-&&\mbox{\em dimensionless position in moving frame} \\
-& \tilde{t} = t \freqp / \lamfac
-&&\mbox{\em dimensionless time}
-\end{align}
-The explanation will be something like: instead of the most conventional choice, we scale displacement on $\sqrt{\pi} \etastd$ so that energy is normalize to unity.
-This choice will be the most convenient for the Hamiltonian framework that we will rely on.
-Maybe we will have to scale time differently too to get the desired powers of $\drat$, but that will make no difference for now.
-Thus, the dimensionless KdV is given by
-\begin{align}
-\label{dimlessKdV2}
+\label{dimlessKdV}
 &u_t + C_3 \drat^{-1} \, u u_x + C_2 \drat \, u_{xxx} = 0
@@
 \end{align}
-When we draw a comparison to the PNAS paper, it will be something like $\epsup = E^{1/2}L^{-3/2}$ and $\delup^2 = L^{-3}$, etc.
+where we have the dimensionless parameters
+\begin{align}
+&\epsup = \etastd / \dup
+&&\mbox{\em upstream amplitude-to-depth ratio} \\
+&\delup = \dup / \lamup
+&&\mbox{\em upstream depth-to-wavelength ratio}
+\end{align}
+
+Note: When comparing to the PNAS paper, we have...
+
+Note: An alternative formulation found in Johnson's textbook is based on continuity of $\depth^{1/4} \eta$, rather than $\eta$ in crossing the ADC.
+In this formulation, the state variable is $u = \depth^{1/4} \, \eta / (\dup^{1/4} \, \ampscale)$.
+If this framework is used, the only change is that the $D^{-1}$ in the second term of \eqref{dimlessKdV} becomes $D^{-5/4}$, which makes little practical difference.
 
 %----------------------------------------------------------------------------%
-% Hamiltonian
+% Stuff
 %----------------------------------------------------------------------------%
-\subsection{Hamiltonian structure}
-
-We now consider the Hamiltonian structure of the dimensionless variable-depth KdV equation \eqref{dimlessKdV2}, over dimensionless domain $x \in [-\pi,\pi]$.
-First, the spatial average of any quantity is given by
-\begin{equation}
-\mean{q} = \frac{1}{2\pi} \int_{-\pi}^{\pi} q \dx \, .
-\end{equation}
-Appealing to near Ergodicity, we will typically assume equivalence between spatial, temporal, and ensemble averages.
+\subsection{Hamiltonian}
 
@@
 \end{equation}
-Next, the definition of the characteristic amplitude in \eqref{ampdefn} implies that $\mean{u^2} = \pi^{-1}$. Consequently, the energy of $u$ is fixed as
+Next, the definition of the characteristic amplitude implies that $\mean{u^2} = \pi^{-1}$. Consequently, the energy of $u$ is fixed as
 \begin{equation}
@@
 
-We introduce the cubic and quadratic Hamiltonian components as
-\begin{align}
-H_3 = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
-H_2 = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
-\end{align}
+\np
+\subsection{Galerkin truncation}
 
-Then the Hamiltonian of \eqref{dimlessKdV1} is given by
-\begin{equation}
-\Ham = C_3 \drat^{-1} \, H_3 - C_2 \drat \, H_2
-\end{equation}
-where $\drat$ changes value across the ADC. More specifically, there exists an upstream and a downstream Hamiltonian
-\begin{align}
-&\Hup = C_3 \, H_3 - C_2 \, H_2 \\
-&\Hdn = C_3 \dratdn^{-1} \, H_3 - C_2 \dratdn \, H_2
-\end{align}
-$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$. Each Hamiltonian $\Ham^{\pm}$ generates a corresponding Gibbs measure $\Gupdn$ 
-\begin{align}
-\Gupdn = C_{\thupdn} \exp(-\thupdn \Hupdn) \delta(\En - \pi)
-\end{align}
-where $\theta = \thupdn$ is the inverse temperature and $C_{\theta}$ a constant (the partition function) that depends on $\theta$. Each measure has corresponding ensemble mean $\mean{}_{\pm}$.
+\subsection{Optimizer of the Hamiltonian}
+The function that maximizes the Hamiltonian subject to the constraints of fixed energy and zero momentum is a traveling wave solution!
 
-%----------------------------------------------------------------------------%
-% Matching
-%----------------------------------------------------------------------------%
-\subsection{Matching condition}
+%----------------------------------------------------------%
+% Direct numerical simulations
+%----------------------------------------------------------%
+\section{Direct numerical simulations}
 
-The statistical matching condition is
-\begin{equation}
-\meanup{\Hdn} = \meandn{\Hdn}
-\end{equation}
-An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$. Inserting and simplifying gives
-\begin{equation}
-\label{H3H2}
-\frac{\meandn{H_3^+}} {\meandn{H_2^+} - \meanup{H_2^+}} = \frac{C_2}{C_3} \dratdn^2
-\end{equation}
-%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
-To convert to dimensional, i.e.~experimental, values we first note that
-\begin{align}
-\mean{H_3^+} = \frac{\pi}{3} \mean{u^3} = 
-\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
-\end{align}
-Second, to convert the downstream surface slope, we note
-\begin{align}
-\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
-\end{align}
-where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
-Note: I believe it is important here to use $\lamdn$ since we are matching $\Hdn$.
-Hence, we have
-\begin{align}
-&\mean{H_2^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
-\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
-\end{align}
-%
-Then algebra gives
-\begin{equation}
-\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = 
-\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
-\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
-\end{equation}
-Also we have the simplification
-\begin{equation}
-\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
-\end{equation}
-Hence, \eqref{H3H2} gives
-\begin{equation}
-\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3
-\end{equation}
-In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
-% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.
+\section{Sampling Algorithms}
 
-\subsection{Derivation with $\lam$ changing across ADC (Obselete)}
-This derivation must be faulty, but here it is so that I remember it.
-Aha, I think the error is as follows: The matching condition is for $\Hdn$ only, using the two measures. So you have to use the downstream scales, in particular $\lamdn$.
+\subsection{Naive acceptance-rejection algorithm}
 
-\vsp{1}
-Anyway, here is the flawed derivation, so that we remember it.
-Algebra gives
-\begin{equation}
-\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = 
-\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2} 
-\left( \frac{\skwdn(\eta)}{ \lamdn^2 \vardn(\eta_x) - \lamup^2 \varup(\eta_x)} \right) = 
-\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamup^2} 
-\left( \frac{\skwdn(\eta)} {\dratdn \vardn(\eta_x) - \varup(\eta_x)} \right)
-\end{equation}
-where in the second equality, we have used the fact that
-\begin{equation}
-\frac{\lamdn^2}{\lamup^2} = \frac{\ddn}{\dup} = \dratdn
-\end{equation}
-Also we have the simplification
-\begin{equation}
-\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
-\end{equation}
-Hence, \eqref{H3H2} gives
-\begin{equation}
-\frac{\skwdn(\eta)} {\dratdn \vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^2
-\end{equation}
-In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
+\subsection{Markov-chain Monte Carlo}
 
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[htbp]
-\begin{center}
-\includegraphics[width = 0.80 \textwidth]{../Figs/SkewAmp.pdf}
-\caption{\label{fig1} 
-The ratio ${\skw(\eta)}/{\Delta \var(\eta_x)}$ plots against dimensionless wave amplitude.
-}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\subsection{Improved acceptance-rejection algorithm}
 

### 2019-07-18 paper2_JNLS/paper.tex
NM

@@
 
+%----------------------------------------------------------%
+% Intro
+%----------------------------------------------------------%
 \section{Introduction}
 
+Introduce waves. \\
+Introduce papers: PRF, PNAS, and J Stat Phys. \\
+
+The purpose of this manuscript is to:
+\begin{itemize}
+\item Provide a more comprehensive treatment of the experiments and theory in combination.
+In particular, we make new comparisons between theory and experiments, which further confirm the predictive power of the theoretical framework.
+\item Provide more thorough treatment of the link between physical system parameters and model parameters.
+This will elucidate the connection between theory and experiments.
+In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in PNAS for the sake of brevity.
+\item Provide additional experimental measurements not presented in Bolles PRF. For example, statistics on surface slope and autocorrelation.
+\end{itemize}
+
 %----------------------------------------------------------%
-% Mathematical framework
+% Experiments
+%----------------------------------------------------------%
+\section{The experiments}
+
+%----------------------------------------------------------%
+% Theory
 %----------------------------------------------------------%

### 2019-07-18 paper2_JNLS/paper.tex
NM

@@
 This will elucidate the connection between theory and experiments.
-In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in PNAS for the sake of brevity.
-\item Provide additional experimental measurements not presented in Bolles PRF. For example, statistics on surface slope and autocorrelation.
+In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in \cite{majda2019statistical} for the sake of brevity.
+\item Provide additional experimental measurements not presented in \cite{bolles2019anomalous}. For example, statistics on surface slope and autocorrelation.
 \end{itemize}
@@
 
-%\bibliographystyle{plain}
-%\bibliography{Notesbib}
+\section*{Acknowledgements}
+CTB acknowledges support from the IDEA grant at Florida State University, as well as from the Geophysical Fluid Dynamics Institute. 
+MNJM acknowledges support from the Simons Foundation, award 524259.
+
+\bibliographystyle{plain}
+\bibliography{wavesbib}
 

### 2019-07-20 paper2_JNLS/paper.tex
NM

@@
 \section{The experiments}
+
+	As discussed in \cite{bolles2019anomalous}, the experiments consist of a long, narrow wave tank (6 m long x 20 cm wide x 30 cm high), with waves generated by plexiglass paddle at one end.
+The waves propagate through the tank and, roughly midway through, pass over an abrupt depth change (ADC), which is created by a plexiglass step.
+The waves continue to propagate through the shallower depth until reaching the far end of the tank, at which point their energy is dissipated by a horse-hair dampener.
+Since this dampener minimizes the backscatter, the waves propagate primarily in one direction only, from left to right in Fig.~\ref{fig1}.
+The free surface is illuminated by light-emitting diodes that run along the bottom of the tank and imaged from the sideview with a Nikon D3300 at 60 frames per second.
+This arrangement allows measurements of surface displacements with a precision of 1/3 millimeter.
+
+ % Figure 1
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.85 \linewidth]{Figs/fig1.pdf}
+\caption{
+(a) Experimental schematic. Reproduced from \cite{bolles2019anomalous}
+}
+%(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms.
+\label{fig1}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+ % Figure 2
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.85 \linewidth]{Figs/fig2.pdf}
+\caption{
+(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms. Reproduced from \cite{bolles2019anomalous}
+}
+\label{fig1}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+ % Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees
+
+  % Figure 3
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.85 \linewidth]{Figs/fig3.pdf}
+\caption{
+(b)--(c) Surface slope measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms. NEW PLOT.
+}
+\label{fig1}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 

### 2019-07-21 paper2_JNLS/paper.tex
NM

@@
 Since this dampener minimizes the backscatter, the waves propagate primarily in one direction only, from left to right in Fig.~\ref{fig1}.
-The free surface is illuminated by light-emitting diodes that run along the bottom of the tank and imaged from the sideview with a Nikon D3300 at 60 frames per second.
-This arrangement allows measurements of surface displacements with a precision of 1/3 millimeter.
+The free surface is illuminated by light-emitting diodes running along the bottom of the tank and then imaged from the sideview with a Nikon D3300 at 60 frames per second.
+The illumination technique coupled with high pixel count of the camera allows the surface displacement to be resolved with a accuracy better than 1/3 millimeter.
 

### 2019-07-22 WavesData-2019/LatexNotes/notes.tex
NM

@@
 
+\np
+%----------------------------------------------------------%
+% Misc
+%----------------------------------------------------------%
+\section{Misc}
+\subsection{The optimizer of the Hamiltonian is a traveling wave solution}
+For the acceptance-rejection algorithm, it is important to find the maximal value of the Hamiltonian subject to the constraints of fixed energy $E = E_0$ and zero momentum $M = 0$.
+For simplicity, consider the KdV equation
+\begin{equation}
+\label{KdVsimple}
+u_t + u u_x + \mu u_{xxx} = 0
+\end{equation}
+with corresponding Hamiltonian
+\begin{equation}
+H = \frac{1}{6} \int u^3 dx - \frac{\mu}{2} \int u_x^2 dx
+\end{equation}
+
+Consider the functional
+\begin{equation}
+F[u] = H[u] + \lambda_1 (E-E_0) + \lambda_2 (M-0)
+\end{equation}
+where $\lambda_1$ and $\lambda_2$ are Lagrangian multipliers. Taking the variational derivative gives
+\begin{equation}
+\frac{\delta F}{\delta u} = \frac{1}{2} u^2 + \mu u_{xx} + \lambda_1 u + \lambda_2 = 0
+\end{equation}
+which is a second-order nonlinear ODE for the optimizer $u = U(x)$.
+Note that this is identical to the condition for a traveling wave solution!!!
+To see this, consider traveling wave solutions of \ref{KdVsimple}, $u(x,t) = U(x-Vt)$.
+Inserting into \ref{KdVsimple} and integrating once gives
+\begin{equation}
+\frac{1}{2} U^2 + \mu U'' - V U = \text{Const}
+\end{equation}
+which is exactly the same condition!!!
+
 %\bibliographystyle{plain}

### 2019-07-22 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\uhat}{\hat{u}}
-\newcommand{\Ham}{H}
+\newcommand{\Ham}{\mathcal{H}}
+\newcommand{\Hthree}{\Ham_{3}}
+\newcommand{\Htwo}{\Ham_{2}}
 \newcommand{\Hup}{\Ham^{-}}
@@
 \newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
+
+\newcommand{\omavg}{\omega_0}
+\newcommand{\omsig}{\sigma_{\omega}}
 %----------------------------------------------------------------------------%
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+Waves are driven with a pseudo-random signal. Specifically, the paddle angle $\phi$ with the horizontal is given by
+\begin{align}
+\label{PaddleAngle}
+& \phi(t) = \phi_0 + \Delta \phi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
+\label{anEq}
+& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
+\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
+\end{align}
+(Literal copy and paste) Here, the angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively.
+We set $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz.
+Importantly, the phases $\delta_n$ are uniformly distributed random variables.
+We fix $N = 3000$, which sets a fundamental period of $T = 300$ seconds.
 
@@
 %----------------------------------------------------------------------------%
-\subsection{Nondimensionalization}
+\subsection{Nondimensionalization of variable-depth KdV}
 
@@
 %----------------------------------------------------------------------------%
-\subsection{Hamiltonian}
+\subsection{The truncated KdV framework}
+
+Consider the Fourier series of the state variable
+\begin{equation}
+u = \sum_{k=-\Lambda}^{\Lambda} \uhat_k e^{i k x}
+\end{equation}
+where
+\begin{equation}
+\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
+\end{equation}
+The above sum can be infinite $\Lambda = \infty$ or truncated at a finite wavenumber $\Lambda < \infty$. Further, since $u$ is real valued, we have $\uhat_{-k} = \uhat_{k}^*$.
 
@@
 \end{equation}
-
-We introduce the Fourier series
-\begin{equation}
-u = \sum_{k=-\Lambda}^{\Lambda} \uhat_k e^{i k x}
-\end{equation}
-where
-\begin{equation}
-\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
-\end{equation}
-The above sum can be infinite $\Lambda = \infty$ or truncated at a finite wavenumber $\Lambda < \infty$.
-Further, since $u$ is real valued, we have $\uhat_{-k} = {\uhat_{k}}^*$.
 By Parseval's identity, we have
@@
 
-\np
-\subsection{Galerkin truncation}
+%----------------------------------------------------------------------------%
+% Stuff
+%----------------------------------------------------------------------------%
+\subsection{Hamiltonian structure of TKdV}
 
-\subsection{Optimizer of the Hamiltonian}
-The function that maximizes the Hamiltonian subject to the constraints of fixed energy and zero momentum is a traveling wave solution!
+We introduce the cubic and quadratic Hamiltonian components as
+\begin{align}
+\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
+\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
+\end{align}
+Then the Hamiltonian of \eqref{dimlessKdV} on each side of the ADC is given by
+\begin{equation}
+\Ham = C_3 \drat^{-1} \, \Hthree - C_2 \drat \, \Htwo
+\end{equation}
+where $\drat$ changes value across the ADC. More specifically, there exists an upstream and a downstream Hamiltonian
+\begin{align}
+&\Hup = C_3 \, \Hthree - C_2 \, \Htwo \\
+&\Hdn = C_3 \dratdn^{-1} \, \Hthree - C_2 \dratdn \, \Htwo
+\end{align}
+$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$. Each Hamiltonian $\Ham^{\pm}$ generates a corresponding Gibbs measure $\Gupdn$ 
+\begin{align}
+\Gupdn = C_{\thupdn} \exp(-\thupdn \Hupdn) \delta(\En - \pi)
+\end{align}
+where $\theta = \thupdn$ is the inverse temperature and $C_{\theta}$ a constant (the partition function) that depends on $\theta$. Each measure has corresponding ensemble mean $\mean{}_{\pm}$.
+
+%Misc: The function that maximizes the Hamiltonian subject to the constraints of fixed energy and zero momentum is a traveling wave solution!
 
@@
 
+  % Figure
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.85 \linewidth]{Figs/SkewAmp.pdf}
+\caption{
+Relationship predicted by the theory and confirmed by experiments.
+}
+\label{AAA}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
 \subsection{Comparison of basic features}

### 2019-07-22 paper2_JNLS/paper.tex
NM

@@
 The waves continue to propagate through the shallower depth until reaching the far end of the tank, at which point their energy is dissipated by a horse-hair dampener.
-Since this dampener minimizes the backscatter, the waves propagate primarily in one direction only, from left to right in Fig.~\ref{fig1}.
-The free surface is illuminated by light-emitting diodes running along the bottom of the tank and then imaged from the sideview with a Nikon D3300 at 60 frames per second.
-The illumination technique coupled with high pixel count of the camera allows the surface displacement to be resolved with a accuracy better than 1/3 millimeter.
+Since the dampener minimizes the backscatter, the waves in this experiment propagate primarily in one direction, from left to right in Fig.~\ref{fig1}.
+
+Waves are driven with a pseudo-random signal. Specifically, the paddle angle $\phi$ with the horizontal is given by
+\begin{align}
+\label{PaddleAngle}
+& \phi(t) = \phi_0 + \Delta \phi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
+\label{anEq}
+& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
+\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
+\end{align}
+(Literal copy and paste) Here, the angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively.
+We set $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz.
+Importantly, the phases $\delta_n$ are uniformly distributed random variables.
+We fix $N = 3000$, which sets a fundamental period of $T = 300$ seconds.
 
@@
 
-Waves are driven with a pseudo-random signal. Specifically, the paddle angle $\phi$ with the horizontal is given by
-\begin{align}
-\label{PaddleAngle}
-& \phi(t) = \phi_0 + \Delta \phi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
-\label{anEq}
-& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
-\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
-\end{align}
-(Literal copy and paste) Here, the angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively.
-We set $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz.
-Importantly, the phases $\delta_n$ are uniformly distributed random variables.
-We fix $N = 3000$, which sets a fundamental period of $T = 300$ seconds.
+	The free surface is illuminated by light-emitting diodes running along the bottom of the tank and then imaged from the sideview with a Nikon D3300 at 60 frames per second.
+The illumination technique coupled with high pixel count of the camera allows the surface displacement to be resolved with a accuracy better than 1/3 millimeter.
+Representative free surface measurements are shown in Figs.~\ref{fig2}(a)--(b)
 
@@
 }
-\label{fig1}
+\label{fig2}
 \end{center}
@@
 }
-\label{fig1}
+\label{fig3}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+% Figure 4
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.85 \linewidth]{Figs/fig4.pdf}
+\caption{
+Spatial statistics from the experiments.
+}
+\label{fig4}
 \end{center}

### 2019-07-22 paper2_JNLS/paper.tex
NM

@@
 
-Waves are driven with a pseudo-random signal. Specifically, the paddle angle $\phi$ with the horizontal is given by
+The pivoting motion of the paddle is driven by a 5-phase stepper motor.
+To generate a randomized incoming wave field, the paddle angle $\phi$ is specified (prescribed) by a precomputed psuedo-random signal
 \begin{align}
@@
 The illumination technique coupled with high pixel count of the camera allows the surface displacement to be resolved with a accuracy better than 1/3 millimeter.
-Representative free surface measurements are shown in Figs.~\ref{fig2}(a)--(b)
+Representative free surface measurements are shown in Figs.~\ref{fig2}(a)--(b).
 

### 2019-07-22 paper2_JNLS/paper.tex
NM

@@
 
-The pivoting motion of the paddle is driven by a 5-phase stepper motor.
-To generate a randomized incoming wave field, the paddle angle $\phi$ is specified (prescribed) by a precomputed psuedo-random signal
+The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized incoming wave field, the paddle angle $\phi$ is specified by a precomputed psuedo-random signal
 \begin{align}
@@
 	The free surface is illuminated by light-emitting diodes running along the bottom of the tank and then imaged from the sideview with a Nikon D3300 at 60 frames per second.
-The illumination technique coupled with high pixel count of the camera allows the surface displacement to be resolved with a accuracy better than 1/3 millimeter.
-Representative free surface measurements are shown in Figs.~\ref{fig2}(a)--(b).
+The illumination technique coupled with high pixel count of the camera allows surface displacements to be resolved with a accuracy better than 1/3 millimeter.
+Representative free surface measurements are shown in Figs.~\ref{fig2}(a)--(b) at a position upstream and downstream of the ADC.
+The corresponding histograms are shown in Figs.~\ref{fig2}(c)--(d).
 
@@
 \caption{
-(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms. Reproduced from \cite{bolles2019anomalous}
+(a)--(b) Surface displacement measured at a representative upstream and downstream location. (c)--(d) Corresponding histograms. Reproduced from \cite{bolles2019anomalous}
 }

### 2019-07-22 paper2_JNLS/paper.tex
NM

@@
 The Hamiltonian framework that we will use relies heavily on the energy, and hence it is more convenient to have the energy normalized to unity rather than making the more common choice of $\etastd$ by itself as the scale for wave amplitude.
+\\(Note: The powers in our PNAS paper can be gotten by setting $T = \delta^{-1} \lamfac / \freqp$, since $\delta \sim \sqrt{\depth}$.
+This is a longer timescale, consistent with the KdV regime of validity.
+Also, it is a different timescale on either side of the ADC since $\depth$ changes across the ADC.)
 

### 2019-07-23 paper2_JNLS/paper.tex
NM

--- 2019-07-22 paper2_JNLS/paper.tex
@@
 \newcommand{\freqp}{f_p}
-%\newcommand{\amp}{a}
-\newcommand{\ampscale}{A}
 \newcommand{\etastd}{\eta_{\text std}}
@@
 \newcommand{\lamfac}{N}
-\newcommand{\drat}{D}
+\newcommand{\drat}{\mathcal{D}}
 \newcommand{\dratdn}{\drat_+}
@@
 % Dimensionless parameters
+\newcommand{\ampscale}{\mathcal{A}}
+\newcommand{\lengthscale}{\mathcal{L}}
+\newcommand{\timescale}{\mathcal{T}}
 \newcommand{\epsup}{\eps_0}
@@
 
+% Experiments
 \newcommand{\omavg}{\omega_0}
@@
 &&\mbox{\em dimensionless surface displacement} \\
-&\tilde{x} = \xi / L
+&\tilde{x} = \xi / \lengthscale
 &&\mbox{\em dimensionless position (moving frame)} \\
-&\tilde{t} = t / T
+&\tilde{t} = t / \timescale
 &&\mbox{\em dimensionless time}
 \end{align}
-where the characteristic scales $\ampscale, L, T$ for wave amplitude, length, and time respectively, are free for us to choose.
+where the characteristic scales $\ampscale, \lengthscale, \timescale$ for wave amplitude, length, and time respectively, are free for us to choose.
 Then the dimensionless KdV equation is
 \begin{equation}
-u_t + \frac{3}{2} \left( \frac{c T \ampscale}{L \depth} \right) u u_x 
-+ \frac{1}{6} \left( \frac{c T \depth^2}{L^3} \right) u_{xxx} = 0
+u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
++ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
 \end{equation}
@@
 &\ampscale = \pi^{1/2} \, \etastd \\
-&L = \lamfac \lam / \pi \\
-&T = \lamfac / \freqp
+&\lengthscale = \lamfac \lam / \pi \\
+&\timescale = \lamfac / \freqp
 \end{align}
@@
 &\delup = \dup / \lamup
-&&\mbox{\em upstream depth-to-wavelength ratio}
+&&\mbox{\em upstream depth-to-wavelength ratio} \\
+&\drat = {\depth}/{\dup} \, .
+&&\mbox{\em dimensionless depth}
 \end{align}
+Here, $\drat = \dratupdn$ is the local dimensionless depth, which changes value crossing the ADC.
+In particular, since \eqref{dimlessKdV} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$.
+Then we have
+\begin{equation}
+\drat = 
+\begin{cases}
+1 		&\quad \mbox{for } t<0 \\
+\dratdn 	&\quad \mbox{for } t>0
+\end{cases}
+\end{equation}
+where $\dratdn = \ddn/\dup$.
+See Table \ref{paramtable} for a summary of important parameters.
 
-Note: When comparing to the PNAS paper, we have...
-
-Note: An alternative formulation found in Johnson's textbook is based on continuity of $\depth^{1/4} \eta$, rather than $\eta$ in crossing the ADC.
+\vsp{2}
+Notes
+\begin{itemize}
+\item The powers in our PNAS paper can be gotten by setting $T = \delta^{-1} \lamfac / \freqp$, since $\delta \sim \sqrt{\depth}$.
+This is a longer timescale, consistent with the KdV regime of validity.
+Also, it is a different timescale on either side of the ADC since $\depth$ changes across the ADC.
+\item We can briefly compare the parameters above to the parameters $E_0$ and $L_0$ in our PNAS paper.
+\item An alternative formulation found in Johnson's textbook is based on continuity of $\depth^{1/4} \eta$, rather than $\eta$ in crossing the ADC.
 In this formulation, the state variable is $u = \depth^{1/4} \, \eta / (\dup^{1/4} \, \ampscale)$.
 If this framework is used, the only change is that the $D^{-1}$ in the second term of \eqref{dimlessKdV} becomes $D^{-5/4}$, which makes little practical difference.
+\end{itemize}
 
 %----------------------------------------------------------------------------%
-% Stuff
+% Truncation
 %----------------------------------------------------------------------------%
@@
 %----------------------------------------------------------------------------%
-% Stuff
+% Hamiltonian
 %----------------------------------------------------------------------------%
@@
 \end{align}
-$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$. Each Hamiltonian $\Ham^{\pm}$ generates a corresponding Gibbs measure $\Gupdn$ 
+$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$.
+In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced in \cite{abramov2003hamiltonian} for the Burgers-Hopf system.
+Specifically, this ensemble is microcanonical in the energy, with a fixed energy value, and it is canonical in the Hamiltonian.
+By fixing the energy and hence confining to a compact set, this construction avoids the divergence at infinity that would occur for a simple canonical ensemble.
+The resulting distribution is thus normalizable.
+Accordingly, each Hamiltonian $\Ham^{\pm}$ generates a corresponding mixed ensemble, or Gibbs measure given by 
 \begin{align}
@@
 \end{align}
-where $\theta = \thupdn$ is the inverse temperature and $C_{\theta}$ a constant (the partition function) that depends on $\theta$. Each measure has corresponding ensemble mean $\mean{}_{\pm}$.
+Here $\theta = \thupdn$ is the inverse temperature and $C_{\theta}$ a constant (inverse of the partition function) that depends on $\theta$.
+Each measure has corresponding ensemble mean $\mean{}_{\pm}$.
 
 %Misc: The function that maximizes the Hamiltonian subject to the constraints of fixed energy and zero momentum is a traveling wave solution!
+
+%----------------------------------------------------------------------------%
+% Matching
+%----------------------------------------------------------------------------%
+\subsection{Matching at the ADC}
+
+Recall that the abrupt depth change is met by traveling waves at some time $t = T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
+The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales (specifically, longer than ???).
+Meanwhile, wave evolution over shorter timescales is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e.
+no nonlinearity and no dispersion).
+Hence, a short instant after waves pass over the ADC, the waveform does not yet have time to alter significantly, which gives the (deterministic) matching condition
+\begin{align}
+&u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+} \qquad \text{for } x \in [-\pi, \pi]
+%&&\mbox{\em deterministic matching condition}
+\end{align}
+%or in Johnson's formulation, it is $\depth^{1/4} \eta$ that matches.
+
+Since $u$ matches for every single trajectory, it must also match in the statistical sense.
+Moreover, the downstream Hamiltonian $\Hdn$ is a functional of $u$ and hence must also match.
+This gives the {\em statistical matching condition}
+\begin{align}
+&\meanup{\Hdn} = \meandn{\Hdn}
+%&&\mbox{\em statistical matching condition}
+\end{align}
+Once $\Hdn$ downstream is determined, it is conserved in the outgoing wave field.
 

### 2019-07-23 paper2_JNLS/paper.tex
NM

@@
 Recall that the abrupt depth change is met by traveling waves at some time $t = T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
-The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales (specifically, longer than ???).
-Meanwhile, wave evolution over shorter timescales is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e.
+The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales (specifically, $t \gg \delta^{-2}$ for dimensionless time).
+Meanwhile, wave evolution over shorter timescales, $t \ll \delta^{-2}$, is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e.
 no nonlinearity and no dispersion).

### 2019-07-23 paper2_JNLS/paper.tex
NM

@@
 \begin{align}
+\label{statmatch}
 &\meanup{\Hdn} = \meandn{\Hdn}
@@
 Once $\Hdn$ downstream is determined, it is conserved in the outgoing wave field.
+
+An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$.
+Inserting this approximation into the statistical matching condition \eqref{statmatch} and simplifying gives
+\begin{equation}
+\label{H3H2}
+\frac{\meandn{H_3^+}} {\meandn{H_2^+} - \meanup{H_2^+}} = \frac{C_2}{C_3} \dratdn^2
+\end{equation}
+%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
+To convert to dimensional, i.e.~experimental, values we first note that
+\begin{align}
+\mean{H_3^+} = \frac{\pi}{3} \mean{u^3} = 
+\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
+\end{align}
+Second, to convert the downstream surface slope, we note
+\begin{align}
+\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
+\end{align}
+where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
+Note: I believe it is important here to use $\lamdn$ in reverting to dimensional variables since we are matching $\Hdn$.
+Hence, we have
+\begin{align}
+&\mean{H_2^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
+\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
+\end{align}
+%
+Then algebra gives
+\begin{equation}
+\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = 
+\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
+\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
+\end{equation}
+Also we have the simplification
+\begin{equation}
+\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
+\end{equation}
+Hence, \eqref{H3H2} gives
+\begin{equation}
+\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3
+\end{equation}
+In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
+% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.
 

### 2019-07-23 paper2_JNLS/paper.tex
NM

@@
 
+%----------------------------------------------------------------------------%
+% Skewness formula
+%----------------------------------------------------------------------------%
+\subsection{Explicit formula for outgoing skewness}
 An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$.
@@
 \end{equation}
-Hence, \eqref{H3H2} gives
+Hence, \eqref{H3H2} gives the remarkably explicit formula
 \begin{equation}
-\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3
+\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 =
+\frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
 \end{equation}

### 2019-07-27 paper2_JNLS/paper.tex
NM

@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
- % Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees
+ % Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees, which is in between runs 5 and 6.
 
@@
 \begin{equation}
+\label{KdV}
 2 \eta_t + \frac{3 c}{\depth} \eta \eta_{\xi} + \frac{c \depth^2}{3} \eta_{\xi \xi \xi} = 0
@@
 
+Here, we recast the variable-depth KdV equation \eqref{KdV} in dimensionless variables.
+Of course, the choice of how to nondimensionalize is not unique, and our choices will made for convenience with the Hamiltonian structure introduced in the next section.
+It may be instructive to be able to compare against other traditional nondimensionalizations and hence we first introduce a generic nondimensionalization: nondimensionalize
+\begin{align}
+&u = \eta / \ampscale
+&&\mbox{\em dimensionless surface displacement} \\
+&\tilde{x} = \xi / \lengthscale
+&&\mbox{\em dimensionless position (moving frame)} \\
+&\tilde{t} = t / \timescale
+&&\mbox{\em dimensionless time}
+\end{align}
+where the characteristic scales $\ampscale, \lengthscale, \timescale$ for wave amplitude, length, and time respectively, are free for us to choose.
+Then the dimensionless KdV equation is
+\begin{equation}
+u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
++ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
+\end{equation}
+
 % Table
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-Lets do the nondimensionalization generally, so that we can easily make modifications if needed. In general, we introduce dimensionless variables as
-\begin{align}
-&u = \eta / \ampscale
-&&\mbox{\em dimensionless surface displacement} \\
-&\tilde{x} = \xi / \lengthscale
-&&\mbox{\em dimensionless position (moving frame)} \\
-&\tilde{t} = t / \timescale
-&&\mbox{\em dimensionless time}
-\end{align}
-where the characteristic scales $\ampscale, \lengthscale, \timescale$ for wave amplitude, length, and time respectively, are free for us to choose.
-Then the dimensionless KdV equation is
-\begin{equation}
-u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
-+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
-\end{equation}
 

### 2019-07-29 paper2_JNLS/paper.tex
NM

@@
 Consider the surface displacement $\eta(x,t)$ and the reference frame moving with the local wave speed $\xi = x - ct$, where $c = \sqrt{g \depth}$ is the wave speed, $g$ gravity, and $\depth$ the local depth.
-To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation (KdV), which in dimensional form is given by
+To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation (KdV), which in dimensional form is given by \cite{whitham2011linear}
 \begin{equation}
 \label{KdV}
-2 \eta_t + \frac{3 c}{\depth} \eta \eta_{\xi} + \frac{c \depth^2}{3} \eta_{\xi \xi \xi} = 0
+\eta_t + \frac{3 c}{2 \depth} \eta \eta_{\xi} + \frac{c \depth^2}{6} \eta_{\xi \xi \xi} = 0
 \end{equation}
-% The coefficients and scales can be verified on Wolfram KdV page (but what about the sign?). I would like an official source as confirmation though.
+% This dimensional form is given by Whitham on p. 461, Eq. 13.98. I first got it from the Wolfram KdV, which gives the same equation.
 

### 2019-07-31 WavesData-2019/LatexNotes/notes.tex
NM

@@
 
-\begin{comment}
+%\begin{comment}
 {\bf Sidenote}: In the case of constant depth, if one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
@@
 where $\eps = a/h$ and $\delta = h/\lam$.
-\end{comment}
+%\end{comment}
 

### 2019-07-31 paper2_JNLS/paper.tex
NM

@@
 Of course, the choice of how to nondimensionalize is not unique, and our choices will made for convenience with the Hamiltonian structure introduced in the next section.
-It may be instructive to be able to compare against other traditional nondimensionalizations and hence we first introduce a generic nondimensionalization: nondimensionalize
+It may be instructive to to compare against other traditional nondimensionalizations and hence we first introduce a generic nondimensionalization:
 \begin{align}
@@
 \begin{align}
-&\ampscale = \pi^{1/2} \, \etastd \\
-&\lengthscale = \lamfac \lam / \pi \\
-&\timescale = \lamfac / \freqp
+\ampscale = \pi^{1/2} \, \etastd \, , \qquad
+\lengthscale = \frac{\lamfac \lam}{\pi} \, , \qquad
+\timescale = \frac{\lamfac \lam}{\freqp \depth}
 \end{align}
@@
 \label{dimlessKdV}
-&u_t + C_3 \drat^{-1} \, u u_x + C_2 \drat \, u_{xxx} = 0
+&u_t + C_3 \drat^{-3/2} \, u u_x + C_2 \drat^{1/2} \, u_{xxx} = 0
 \qquad \text{for } x \in [-\pi,\pi] \\
-&C_3 = \frac{3}{2} \pi^{3/2} \epsup  \\
-&C_2 = \frac{\pi^3 \delup^2}{6 \lamfac^2} 
+&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1}  \\
+&C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
 \end{align}

### 2019-08-01 paper2_JNLS/paper.tex
NM

@@
 
-Here, we recast the variable-depth KdV equation \eqref{KdV} in dimensionless variables.
-Of course, the choice of how to nondimensionalize is not unique, and our choices will made for convenience with the Hamiltonian structure introduced in the next section.
-It may be instructive to to compare against other traditional nondimensionalizations and hence we first introduce a generic nondimensionalization:
+In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the Hamiltonian/statistical-mechanics framework introduced in \cite{majda2019statistical}.
+Since the choice of how to normalize is not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other choices.
+To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment.
+We introduce the dimensionless variables
 \begin{align}
@@
 &\tilde{x} = \xi / \lengthscale
-&&\mbox{\em dimensionless position (moving frame)} \\
+&&\mbox{\em dimensionless position (in moving frame)} \\
 &\tilde{t} = t / \timescale
@@
 \end{align}
-where the characteristic scales $\ampscale, \lengthscale, \timescale$ for wave amplitude, length, and time respectively, are free for us to choose.
-Then the dimensionless KdV equation is
+Recasting \eqref{KdV} in terms of these variables and dropping the tildes for simplicity, gives the general dimensionless KdV equation:
 \begin{equation}
@@
 
-% Table
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{table}[h]%[htbp]
-\begin{center}
-\caption{Table of parameters}
-\label{paramtable}
-\begin{tabular}{l l l}
-\hline \multicolumn{3} { c }{Parameters that are constant in a single experiment} \\
-\hline Description & Notation and definition & Value in experiments \\
-\hline
-Peak forcing frequency		& $f_p$						& 2 Hz \\
-Characteristic wave amplitude	& $\etastd = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
-Upstream depth			& $\dup$						& 12.5 cm \\
-Downstream depth			& $\ddn$						& 3 cm (and varied) \\
-Upstream wavelength		& $\lamup = \sqrt{g \dup}/f_p$		& 55 cm \\
-Upstream wavelength		& $\lamdn = \sqrt{g \ddn}/f_p$		& 27 cm \\
-%
-Amplitude-to-depth ratio		& $\epsup = \etastd / \dup$			& 0.0024--0.024 \\
-Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
-Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 (and varied)
-\end{tabular}
-\end{center}
-\end{table}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-In particular, we make the following choices
+Now, the scales $\ampscale, \lengthscale, \timescale$ can be chosen for ease in working with a particular framework. In particular, we choose the following scales,
 \begin{align}
@@
 where $\lamfac$ is an integer to be chosen later. 
+The explanation for these choices is as follows.
+First, we have chosen the characteristic amplitude, $\ampscale$, above to normalize the energy of the state-variable $u$ to unity, as will be demonstrated in Section \ref{tKdVSec}.
+Second, regarding the choice of $\lengthscale$, recall that $\lam$ is the characteristic wavelength that corresponds to the peak forcing frequency $\freqp$ in the experiments.
+If only integer multiples of $\freqp$ were imposed (e.g.~lower frequencies were not present), then the forcing would produce waves that are periodic over the physical interval $\xi \in [-\lam,\lam]$.
+Since, lower frequencies do exist, strict periodicity is not satisfied, but rather waves may be nearly periodic over $\xi \in [-\lam,\lam]$.
+The approximation of near-periodicity becomes more accurate if integer multiples of $\lam$ are considered, hence our introduction of $\lamfac$.
+Thus, we have chosen $\lengthscale$ above so that the normalized state-variable $u$ exists on $\tilde{x} \in [-\pi, \pi]$ and periodic boundary conditions can be reasonably imposed on this domain.
+Lastly, regarding $\timescale$, the most basic timescale in the experiments is simply $\freqp^{-1}$, i.e.~the period of waves passing a fixed reference point.
+Of course, the leading-order behavior in shallow water is simple translation of waves with uniform speed $c$.
+The KdV corrects this behavior to leading order and describes dynamics that evolve over longer timescales.
+Hence, in line with traditional normalizations \cite{johnson1997modern}, we have chosen a suitably long timescale above.
+
 In the physical domain, we will consider an integer multiple of the characteristic wavelength $\xi \in [-\lamfac \lam, \lamfac \lam]$, so that the state variable $u$ is periodic over the  dimensionless domain $\tilde{x} \in [-\pi, \pi]$.
@@
 
+% Table
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{table}[h]%[htbp]
+\begin{center}
+\caption{Table of parameters}
+\label{paramtable}
+\begin{tabular}{l l l}
+\hline \multicolumn{3} { c }{Parameters that are constant in a single experiment} \\
+\hline Description & Notation and definition & Value in experiments \\
+\hline
+Peak forcing frequency		& $f_p$						& 2 Hz \\
+Characteristic wave amplitude	& $\etastd = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
+Upstream depth			& $\dup$						& 12.5 cm \\
+Downstream depth			& $\ddn$						& 3 cm (and varied) \\
+Upstream wavelength		& $\lamup = \sqrt{g \dup}/f_p$		& 55 cm \\
+Upstream wavelength		& $\lamdn = \sqrt{g \ddn}/f_p$		& 27 cm \\
+%
+Amplitude-to-depth ratio		& $\epsup = \etastd / \dup$			& 0.0024--0.024 \\
+Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
+Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 (and varied)
+\end{tabular}
+\end{center}
+\end{table}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
 \vsp{2}
@@
 \subsection{The truncated KdV framework}
+\label{tKdVSec}
 

### 2019-08-01 paper2_JNLS/paper.tex
NM

@@
 
-In the physical domain, we will consider an integer multiple of the characteristic wavelength $\xi \in [-\lamfac \lam, \lamfac \lam]$, so that the state variable $u$ is periodic over the  dimensionless domain $\tilde{x} \in [-\pi, \pi]$.
-We have scaled time accordingly.
-Finally, our choice of the scale for wave amplitude $\ampscale$ is made to normalize the energy of $u$ to unity.
-The Hamiltonian framework that we will use relies heavily on the energy, and hence it is more convenient to have the energy normalized to unity rather than making the more common choice of $\etastd$ by itself as the scale for wave amplitude.
-
-With these choices, the dimensionless KdV is
+With the above choices, the dimensionless KdV takes the form
 \begin{align}
@@
 \end{align}
-where we have the dimensionless parameters
+where dimensionless parameters are given by
 \begin{align}
+&\drat = {\depth}/{\dup} \, .
+&&\mbox{\em dimensionless depth} \\
 &\epsup = \etastd / \dup
@@
 &\delup = \dup / \lamup
-&&\mbox{\em upstream depth-to-wavelength ratio} \\
-&\drat = {\depth}/{\dup} \, .
-&&\mbox{\em dimensionless depth}
+&&\mbox{\em upstream depth-to-wavelength ratio}
 \end{align}
+While $\epsup$ and $\delup$ are fixed parameters, the dimensionless depth, $\drat = \dratupdn$, changes value across the ADC.
+
+\vsp{2} HERE
 Here, $\drat = \dratupdn$ is the local dimensionless depth, which changes value crossing the ADC.

### 2019-08-01 paper2_JNLS/paper.tex
NM

@@
 Hence, we will not distinguish between upstream and downstream values of $\etastd$.
-
-%----------------------------------------------------------------------------%
-% Nondimensionalization
-%----------------------------------------------------------------------------%
-\subsection{Nondimensionalization of variable-depth KdV}
-
-In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the Hamiltonian/statistical-mechanics framework introduced in \cite{majda2019statistical}.
-Since the choice of how to normalize is not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other choices.
-To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment.
-We introduce the dimensionless variables
-\begin{align}
-&u = \eta / \ampscale
-&&\mbox{\em dimensionless surface displacement} \\
-&\tilde{x} = \xi / \lengthscale
-&&\mbox{\em dimensionless position (in moving frame)} \\
-&\tilde{t} = t / \timescale
-&&\mbox{\em dimensionless time}
-\end{align}
-Recasting \eqref{KdV} in terms of these variables and dropping the tildes for simplicity, gives the general dimensionless KdV equation:
-\begin{equation}
-u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
-+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
-\end{equation}
-
-Now, the scales $\ampscale, \lengthscale, \timescale$ can be chosen for ease in working with a particular framework. In particular, we choose the following scales,
-\begin{align}
-\ampscale = \pi^{1/2} \, \etastd \, , \qquad
-\lengthscale = \frac{\lamfac \lam}{\pi} \, , \qquad
-\timescale = \frac{\lamfac \lam}{\freqp \depth}
-\end{align}
-where $\lamfac$ is an integer to be chosen later. 
-The explanation for these choices is as follows.
-First, we have chosen the characteristic amplitude, $\ampscale$, above to normalize the energy of the state-variable $u$ to unity, as will be demonstrated in Section \ref{tKdVSec}.
-Second, regarding the choice of $\lengthscale$, recall that $\lam$ is the characteristic wavelength that corresponds to the peak forcing frequency $\freqp$ in the experiments.
-If only integer multiples of $\freqp$ were imposed (e.g.~lower frequencies were not present), then the forcing would produce waves that are periodic over the physical interval $\xi \in [-\lam,\lam]$.
-Since, lower frequencies do exist, strict periodicity is not satisfied, but rather waves may be nearly periodic over $\xi \in [-\lam,\lam]$.
-The approximation of near-periodicity becomes more accurate if integer multiples of $\lam$ are considered, hence our introduction of $\lamfac$.
-Thus, we have chosen $\lengthscale$ above so that the normalized state-variable $u$ exists on $\tilde{x} \in [-\pi, \pi]$ and periodic boundary conditions can be reasonably imposed on this domain.
-Lastly, regarding $\timescale$, the most basic timescale in the experiments is simply $\freqp^{-1}$, i.e.~the period of waves passing a fixed reference point.
-Of course, the leading-order behavior in shallow water is simple translation of waves with uniform speed $c$.
-The KdV corrects this behavior to leading order and describes dynamics that evolve over longer timescales.
-Hence, in line with traditional normalizations \cite{johnson1997modern}, we have chosen a suitably long timescale above.
-
-With the above choices, the dimensionless KdV takes the form
-\begin{align}
-\label{dimlessKdV}
-&u_t + C_3 \drat^{-3/2} \, u u_x + C_2 \drat^{1/2} \, u_{xxx} = 0
-\qquad \text{for } x \in [-\pi,\pi] \\
-&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1}  \\
-&C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
-\end{align}
-where dimensionless parameters are given by
-\begin{align}
-&\drat = {\depth}/{\dup} \, .
-&&\mbox{\em dimensionless depth} \\
-&\epsup = \etastd / \dup
-&&\mbox{\em upstream amplitude-to-depth ratio} \\
-&\delup = \dup / \lamup
-&&\mbox{\em upstream depth-to-wavelength ratio}
-\end{align}
-While $\epsup$ and $\delup$ are fixed parameters, the dimensionless depth, $\drat = \dratupdn$, changes value across the ADC.
-
-\vsp{2} HERE
-Here, $\drat = \dratupdn$ is the local dimensionless depth, which changes value crossing the ADC.
-In particular, since \eqref{dimlessKdV} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$.
-Then we have
-\begin{equation}
-\drat = 
-\begin{cases}
-1 		&\quad \mbox{for } t<0 \\
-\dratdn 	&\quad \mbox{for } t>0
-\end{cases}
-\end{equation}
-where $\dratdn = \ddn/\dup$.
+%
 See Table \ref{paramtable} for a summary of important parameters.
@@
 
-\vsp{2}
-Notes
-\begin{itemize}
-\item The powers in our PNAS paper can be gotten by setting $T = \delta^{-1} \lamfac / \freqp$, since $\delta \sim \sqrt{\depth}$.
-This is a longer timescale, consistent with the KdV regime of validity.
-Also, it is a different timescale on either side of the ADC since $\depth$ changes across the ADC.
-\item We can briefly compare the parameters above to the parameters $E_0$ and $L_0$ in our PNAS paper.
-\item An alternative formulation found in Johnson's textbook is based on continuity of $\depth^{1/4} \eta$, rather than $\eta$ in crossing the ADC.
-In this formulation, the state variable is $u = \depth^{1/4} \, \eta / (\dup^{1/4} \, \ampscale)$.
-If this framework is used, the only change is that the $D^{-1}$ in the second term of \eqref{dimlessKdV} becomes $D^{-5/4}$, which makes little practical difference.
-\end{itemize}
+%----------------------------------------------------------------------------%
+% Nondimensionalization
+%----------------------------------------------------------------------------%
+\subsection{Nondimensionalization of variable-depth KdV}
+
+In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the statistical-mechanics framework introduced in \cite{majda2019statistical}.
+Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other choices.
+To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment.
+We introduce the dimensionless variables
+\begin{align}
+&u = \eta / \ampscale
+&&\mbox{\em dimensionless surface displacement} \\
+&\tilde{x} = \xi / \lengthscale
+&&\mbox{\em dimensionless position (in moving frame)} \\
+&\tilde{t} = t / \timescale
+&&\mbox{\em dimensionless time}
+\end{align}
+Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
+\begin{equation}
+u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
++ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
+\end{equation}
+We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.
+
+Now it is possible to choose the scales $\ampscale, \lengthscale, \timescale$ for ease in working with a particular framework. We choose the following scales,
+\begin{align}
+\ampscale = \pi^{1/2} \, \etastd \, , \qquad
+\lengthscale = \frac{\lamfac \lam}{\pi} \, , \qquad
+\timescale = \frac{\lamfac \lam}{\freqp \depth}
+\end{align}
+where $\lamfac$ is an integer to be chosen later. 
+The explanation for these choices is as follows.
+First, we have chosen the characteristic amplitude, $\ampscale$, above to normalize the energy of the state-variable $u$ to unity, as will be demonstrated in Section \ref{tKdVSec}.
+Second, regarding $\lengthscale$, recall that $\lam$ is the characteristic wavelength corresponding to the peak forcing frequency $\freqp$ in the experiments.
+If only integer multiples of $\freqp$ were imposed (e.g.~lower frequencies were not present), then the forcing would produce waves that are periodic over lengthscale $\lam$.
+Since lower frequencies do exist, strict periodicity is not satisfied, but rather waves may be nearly periodic over the physical domain $\xi \in [-\lam,\lam]$.
+The approximation of near-periodicity becomes more accurate if integer multiples of $\lam$ are considered, i.e.~$\xi \in [-\lamfac \lam, \lamfac \lam]$.
+Thus, we have chosen $\lengthscale$ above so that the dimensionless domain of consideration is $\tilde{x} \in [-\pi, \pi]$ and periodic boundary conditions on $u$ can be  imposed on over this domain with accuracy that increases with $\lamfac$.
+Lastly, regarding $\timescale$, the most basic timescale in the experiments is simply $\freqp^{-1}$, i.e.~the period of waves passing a fixed reference point.
+Of course, the leading-order behavior in shallow water is simple translation of waves with uniform speed $c$.
+The KdV equation provides the first correction to this behavior and describes dynamics that evolve over longer timescales.
+Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/\depth \gg 1$, which provides a suitably long timescale and is in line with previous normalizations \cite{johnson1997modern}.
+The scales $\lengthscale$ and $\timescale$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.
+
+With the above choices, the dimensionless KdV takes the form
+\begin{align}
+\label{dimlessKdV}
+&u_t + C_3 \drat^{-3/2} \, u u_x + C_2 \drat^{1/2} \, u_{xxx} = 0
+\qquad \text{for } x \in [-\pi,\pi] \\
+&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1}  \\
+&C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
+\end{align}
+where dimensionless parameters are given by
+\begin{align}
+&\drat = {\depth}/{\dup} \, .
+&&\mbox{\em dimensionless depth} \\
+&\epsup = \etastd / \dup
+&&\mbox{\em upstream amplitude-to-depth ratio} \\
+&\delup = \dup / \lamup
+&&\mbox{\em upstream depth-to-wavelength ratio}
+\end{align}
+While $\epsup$ and $\delup$ are fixed parameters, the dimensionless depth, $\drat = \dratupdn$, changes value across the ADC.
+Recall that the reference frame of \eqref{dimlessKdV} moves with the local wave speed via the variable $\xi = x-ct$ from \eqref{KdV}.
+Thus, the ADC is met at some time $T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
+Then the dimensionless coefficient $\drat$ is a piece-wise constant function of dimensionless time
+\begin{equation}
+\drat = 
+\begin{cases}
+1 		&\quad \mbox{for } \tilde{t}<0 \\
+\dratdn = {\ddn}/{\dup} 	&\quad \mbox{for } \tilde{t}>0
+\end{cases}
+\end{equation}
+
+A few comments are in order.
+First, we note that the original formulation of the theory discussed here utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ but with different expressions for the other dimensionless parameters.
+These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments.
+Second, an alternate formulation of the variable-depth KdV equation exists in which it is conjectured that the product $\depth^{1/4} \eta$, rather than $\eta$, varies continuously across the ADC \cite{johnson1997modern}.
+Of course, that formulation implies a discontinuity in surface displacement, which, though small, is physically unrealistic.
+We have chosen to enforce continuity of surface displacement in the above on the basis of physical realism.
+Using the alternative formulation would simply modify the power of $\drat$ in the second term of \eqref{dimlessKdV} by 1/4, i.e.~$\drat^{-7/4}$ instead of $\drat^{-3/2}$.
+Hence, all of the results discussed below would still hold qualitatively in this alternate formulation, with only slight quantitative modifications.
+
+%NOTE: Perhaps we should compare the dimensionless parameters to our PNAS paper \cite{majda2019statistical}, BUT the expressions for E0 and L0 given in the PNAS paper do not match with what I calculate comparing to the above.
+Plus, it seems like it could only add confusion.
 

### 2019-08-02 paper2_JNLS/paper.tex
NM

@@
 &\etastd = \mean{\eta^2}^{1/2} 
-&&\mbox{\em surface displacement standard deviation} \\
+&&\mbox{\em standard deviation of displacement} \\
 \end{align}
@@
 \qquad \text{for } x \in [-\pi,\pi] \\
-&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1}  \\
-&C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
+&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1} \, , \quad
+C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
 \end{align}

### 2019-08-02 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\uhat}{\hat{u}}
+\newcommand{\sympJ}{\mathcal{J}}
+\newcommand{\vard}[2]{\frac{\delta #1}{\delta #2}}
 \newcommand{\Ham}{\mathcal{H}}
@@
 \begin{cases}
-1 		&\quad \mbox{for } \tilde{t}<0 \\
-\dratdn = {\ddn}/{\dup} 	&\quad \mbox{for } \tilde{t}>0
+1 		&\quad \mbox{for } {t}<0 \\
+\dratdn = {\ddn}/{\dup} 	&\quad \mbox{for } {t}>0
 \end{cases}
@@
 A few comments are in order.
-First, we note that the original formulation of the theory discussed here utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ but with different expressions for the other dimensionless parameters.
+First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ but with different expressions for the other dimensionless parameters.
 These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments.
 Second, an alternate formulation of the variable-depth KdV equation exists in which it is conjectured that the product $\depth^{1/4} \eta$, rather than $\eta$, varies continuously across the ADC \cite{johnson1997modern}.
-Of course, that formulation implies a discontinuity in surface displacement, which, though small, is physically unrealistic.
+Of course, that assumption implies a discontinuity in surface displacement, which, though small, is physically unrealistic.
 We have chosen to enforce continuity of surface displacement in the above on the basis of physical realism.
@@
 Plus, it seems like it could only add confusion.
+
+%----------------------------------------------------------------------------%
+% Hamiltonian
+%----------------------------------------------------------------------------%
+\subsection{Hamiltonian structure}
+
+The variable-depth KdV \eqref{dimlessKdV}, though not Hamiltonian throughout the entire domain, admits a Hamiltonian structure on each side of the ADC. Indeed, \eqref{dimlessKdV} can be expressed as
+\begin{align}
+\dot{u} = \sympJ \vard{\Ham}{u}
+\end{align}
+where $\sympJ = \pdi{}{x}$ and the Hamiltonian is given by
+\begin{equation}
+\Ham = C_3 \drat^{-3/2} \, \Hthree - C_2 \drat^{1/2} \, \Htwo
+\end{equation}
+Since $\drat$ changes value across the ADC, the Hamiltonian also changes value. More explicitly, we have an upstream and downstream Hamiltonian given by
+\begin{align}
+&\Hup = C_3 \, \Hthree - C_2 \, \Htwo \\
+&\Hdn = C_3 \dratdn^{-3/2} \, \Hthree - C_2 \dratdn^{1/2} \, \Htwo
+\end{align}
 

### 2019-08-02 paper2_JNLS/paper.tex
NM

@@
 \end{align}
-where $\sympJ = \pdi{}{x}$ and the Hamiltonian is given by
+where $\sympJ = \pdi{}{x}$ is the symplectic operator and the Hamiltonian is given by
 \begin{equation}
-\Ham = C_3 \drat^{-3/2} \, \Hthree - C_2 \drat^{1/2} \, \Htwo
+\Ham = C_2 \drat^{1/2} \, \Htwo - C_3 \drat^{-3/2} \, \Hthree
 \end{equation}
@@
 \begin{align}
-&\Hup = C_3 \, \Hthree - C_2 \, \Htwo \\
-&\Hdn = C_3 \dratdn^{-3/2} \, \Hthree - C_2 \dratdn^{1/2} \, \Htwo
+&\Hup = C_2 \, \Htwo - C_3 \, \Hthree  \\
+&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree
 \end{align}
+The above formulation differs cosmetically from previous ones \cite{abramov2003hamiltonian, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$ and found  {\em negative} inverse temperature to be appropriate for recovering the experimental observations.
+We have flipped the signs of both $\sympJ$ and $\Ham$ so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional statistical mechanics.
 

### 2019-08-02 paper2_JNLS/paper.tex
NM

@@
 \end{align}
-where $\sympJ = \pdi{}{x}$ is the symplectic operator and the Hamiltonian is given by
+where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham$ the Hamiltonian, which will take different forms on either side of the ADC.
+It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
+\begin{align}
+\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
+\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
+\end{align}
+Then the Hamiltonian can be expressed as
 \begin{equation}
@@
 \end{equation}
-Since $\drat$ changes value across the ADC, the Hamiltonian also changes value. More explicitly, we have an upstream and downstream Hamiltonian given by
+where $\drat$ changers value across the ADC. More explicitly, the separate upstream and downstream Hamiltonians are given by
 \begin{align}
@@
 The above formulation differs cosmetically from previous ones \cite{abramov2003hamiltonian, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$ and found  {\em negative} inverse temperature to be appropriate for recovering the experimental observations.
-We have flipped the signs of both $\sympJ$ and $\Ham$ so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional statistical mechanics.
+We have changed the signs of both $\sympJ$ and $\Ham$ so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional statistical mechanics.
 

### 2019-08-02 paper2_JNLS/paper.tex
NM

@@
 \end{align}
-where dimensionless parameters are given by
+The constants $C_3$ and $C_2$ do {\em not} change value crossing the ADC and are given in terms of the dimensionless parameters
 \begin{align}
-&\drat = {\depth}/{\dup} \, .
-&&\mbox{\em dimensionless depth} \\
 &\epsup = \etastd / \dup
@@
 \end{align}
-While $\epsup$ and $\delup$ are fixed parameters, the dimensionless depth, $\drat = \dratupdn$, changes value across the ADC.
+The reason for the subscripts $3$ and $2$ will become evident in the next section.
+
+Meanwhile, the dimensionless depth $\drat = {\depth}/{\dup}$ {\em does} change value across the ADC since the depth $\depth$ changes.
 Recall that the reference frame of \eqref{dimlessKdV} moves with the local wave speed via the variable $\xi = x-ct$ from \eqref{KdV}.
 Thus, the ADC is met at some time $T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
-Then the dimensionless coefficient $\drat$ is a piece-wise constant function of dimensionless time
+Therefore, we can regard $\drat$ is a piece-wise constant function of dimensionless time
 \begin{equation}
@@
 A few comments are in order.
-First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ but with different expressions for the other dimensionless parameters.
+First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ in \eqref{dimlessKdV} but with different expressions for the other dimensionless parameters.
 These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments.
@@
 \end{align}
-The above formulation differs cosmetically from previous ones \cite{abramov2003hamiltonian, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$ and found  {\em negative} inverse temperature to be appropriate for recovering the experimental observations.
-We have changed the signs of both $\sympJ$ and $\Ham$ so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional statistical mechanics.
+The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$.
+With those signs, \cite{majda2019statistical} found  {\em negative} inverse temperature to be appropriate for recovering the experimental observations.
+We used the signs above so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional statistical mechanics.
+% Notes for myself on the sign change: Abramov CPAM 2003 and Bajars
 

### 2019-08-02 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\delup}{\delta_0}
-% Hamiltonian and Gibbs stuff
+% Hamiltonian stuff
 \newcommand{\uhat}{\hat{u}}
@@
 \newcommand{\Ham}{\mathcal{H}}
+\newcommand{\Hup}{\Ham_{-}}
+\newcommand{\Hdn}{\Ham_{+}}
+\newcommand{\Hupdn}{\Ham_{\pm}}
 \newcommand{\Hthree}{\Ham_{3}}
 \newcommand{\Htwo}{\Ham_{2}}
-\newcommand{\Hup}{\Ham^{-}}
-\newcommand{\Hdn}{\Ham^{+}}
-\newcommand{\Hupdn}{\Ham^{\pm}}
+\newcommand{\Fcnl}{\mathcal{F}}
+% Gibbs and theta
 \newcommand{\Gibbs}{\mathcal{G}}
@@
 \begin{equation}
+\label{dratpw}
 \drat = 
@@
 
-%NOTE: Perhaps we should compare the dimensionless parameters to our PNAS paper \cite{majda2019statistical}, BUT the expressions for E0 and L0 given in the PNAS paper do not match with what I calculate comparing to the above.
+%NOTE: Perhaps we should be more specific about comparing to the dimensionless parameters in our PNAS paper \cite{majda2019statistical}, BUT the expressions for E0 and L0 given in the PNAS paper do not match with what I calculate comparing to the above.
 Plus, it seems like it could only add confusion.
@@
 \begin{align}
-\dot{u} = \sympJ \vard{\Ham}{u}
+\dot{u}_{\pm} = \sympJ \vard{\Hupdn}{u}
 \end{align}
-where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham$ the Hamiltonian, which will take different forms on either side of the ADC.
+where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham = \Hupdn$ is the Hamiltonian, which will take different forms on either side of the ADC.
 It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
@@
 \begin{equation}
-\Ham = C_2 \drat^{1/2} \, \Htwo - C_3 \drat^{-3/2} \, \Hthree
+\Hupdn = C_2 \dratupdn^{1/2} \, \Htwo - C_3 \dratupdn^{-3/2} \, \Hthree
 \end{equation}
-where $\drat$ changers value across the ADC. More explicitly, the separate upstream and downstream Hamiltonians are given by
+where $\drat = \dratupdn$ changes value across the ADC. More explicitly, substituting \eqref{dratpw}, the separate upstream and downstream Hamiltonians are
 \begin{align}
-&\Hup = C_2 \, \Htwo - C_3 \, \Hthree  \\
-&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree
+&\Hup = C_2 \, \Htwo - C_3 \, \Hthree 						&& \text{for } t<0 \\
+&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree	&& \text{for } t>0
 \end{align}
 The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$.
-With those signs, \cite{majda2019statistical} found  {\em negative} inverse temperature to be appropriate for recovering the experimental observations.
-We used the signs above so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional statistical mechanics.
-% Notes for myself on the sign change: Abramov CPAM 2003 and Bajars
+With that sign convention, a {\em negative} inverse temperature is required to accurately describe the experimental observations, as found in \cite{majda2019statistical}.
+We, however, choose the sign convention above so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.
 
+% Notes for myself on the sign change: Abramov CPAM 2003, Bajars Nonlinearity 2013, and our 2019 PNAS paper all use J = -d/dx with the opposite sign in the Hamiltonian.
+However, I was able to easily find several references online that use J = d/dx, so there is nothing wrong with it.
+The exact criteria that J has to satisfy are listed in the CPAM 2003 paper, and J = d/dx satisfies all these criteria just as well as J = -d/dx does.
+
+The evolution of any functional $\Fcnl[u]$ is given by 
+\begin{align}
+\partial_t \Fcnl = \{ \Fcnl, \Ham \} = \int_{-\pi}^{\pi} \vard{F}{u} \sympJ \vard{\Ham}{u} \dx
+\end{align}
+where again $\Ham = \Hup$ if $t<0$ and $\Ham = \Hdn$ if $t>0$.
 %----------------------------------------------------------------------------%

### 2019-08-06 paper2_JNLS/paper.tex
NM

@@
 \begin{align}
+\label{scales}
 \ampscale = \pi^{1/2} \, \etastd \, , \qquad
@@
 \begin{equation}
+\label{Hamiltonian}
 \Hupdn = C_2 \dratupdn^{1/2} \, \Htwo - C_3 \dratupdn^{-3/2} \, \Hthree
@@
 The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$.
-With that sign convention, a {\em negative} inverse temperature is required to accurately describe the experimental observations, as found in \cite{majda2019statistical}.
-We, however, choose the sign convention above so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.
+With that sign convention, Majda et al.
+2019 found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019statistical}.
+We have chosen the sign convention above so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.
 
@@
 The exact criteria that J has to satisfy are listed in the CPAM 2003 paper, and J = d/dx satisfies all these criteria just as well as J = -d/dx does.
+
+We introduce two important invariants of KdV, the momentum and the energy
+\begin{align}
+\label{MomEn}
+\Mo[u] \equiv \int_{-\pi}^{\pi} u \dx \, = 0 , \qquad
+\En[u] \equiv \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
+\end{align}
+As indicated above, the momentum of $u$ vanishes since it is measured as displacement from equilibrium.
+Second, by the scale chosen for $\ampscale$ in \eqref{scales}, the energy has been normalized to unity.
 
@@
 Consider the Fourier series of the state variable
+\begin{align}
+u = \sum_{k=-\Lambda}^{\Lambda} \uhat_k e^{i k x} \, , \qquad
+\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
+\end{align}
+The above sum can be infinite $\Lambda = \infty$ or truncated at a finite wavenumber $\Lambda < \infty$.
+Further, since $u$ is real valued, we have $\uhat_{-k} = \uhat_{k}^*$.
+Since momentum vanishes, we have $\uhat_0 = 0$.
+Meanwhile, Parseval's identity gives
 \begin{equation}
-u = \sum_{k=-\Lambda}^{\Lambda} \uhat_k e^{i k x}
-\end{equation}
-where
-\begin{equation}
-\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
-\end{equation}
-The above sum can be infinite $\Lambda = \infty$ or truncated at a finite wavenumber $\Lambda < \infty$. Further, since $u$ is real valued, we have $\uhat_{-k} = \uhat_{k}^*$.
-
-We measure the state variable $u$ as the displacement of the surface from equilibrium, so that $\mean{u} = 0$. Consequently, the momentum of $u$ vanishes
-\begin{equation}
-\Mo[u] \equiv \int_{-\pi}^{\pi} u \dx = 0
-\end{equation}
-Next, the definition of the characteristic amplitude implies that $\mean{u^2} = \pi^{-1}$. Consequently, the energy of $u$ is fixed as
-\begin{equation}
-\En[u] \equiv \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
-\end{equation}
-By Parseval's identity, we have
-\begin{equation}
-\En[u] = 1 = \frac{1}{2} \int u^2 \dx = \pi \sum_{k=-\Lambda}^{\Lambda} \abs{\uhat_k}^2 = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2
+\En[u] = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2 = 1
 \end{equation}
 % The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
+
+The KdV system (i.e.
+$\Lambda = \infty$) possesses the Hamiltonian \eqref{Hamiltonian} and momentum and energy \eqref{MomEn} and an infinite number of additional invariants \cite{lax1975periodic, whitham2011linear}.
+For the truncated system however, $\Lambda < \infty$, the only known invariants are the momentum, energy, and Hamiltonian.
+The infinite sequence of invariants apart from these first three may no longer be invariants of TKdV.
 

### 2019-08-06 paper2_JNLS/paper.tex
NM

@@
 
+WRITE PROJECTION OPERATOR AND TKDV
+
+\vsp{10}
 %----------------------------------------------------------------------------%
@@
 
-We introduce the cubic and quadratic Hamiltonian components as
-\begin{align}
-\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
-\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
-\end{align}
-Then the Hamiltonian of \eqref{dimlessKdV} on each side of the ADC is given by
-\begin{equation}
-\Ham = C_3 \drat^{-1} \, \Hthree - C_2 \drat \, \Htwo
-\end{equation}
-where $\drat$ changes value across the ADC. More specifically, there exists an upstream and a downstream Hamiltonian
-\begin{align}
-&\Hup = C_3 \, \Hthree - C_2 \, \Htwo \\
-&\Hdn = C_3 \dratdn^{-1} \, \Hthree - C_2 \dratdn \, \Htwo
-\end{align}
-$\Hup$ is a conserved quantity for $t<0$ and $\Hdn$ is a conserved quantity for $t>0$.
 In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced in \cite{abramov2003hamiltonian} for the Burgers-Hopf system.

### 2019-08-06 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\Fcnl}{\mathcal{F}}
+\newcommand{\Proj}{\mathcal{P}}
 % Gibbs and theta
@@
 \begin{align}
-u = \sum_{k=-\Lambda}^{\Lambda} \uhat_k e^{i k x} \, , \qquad
-\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
+&u(x,t) = \sum_{k=-\infty}^{\infty} \uhat_k e^{i k x} \, , \\
+&\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
 \end{align}
-The above sum can be infinite $\Lambda = \infty$ or truncated at a finite wavenumber $\Lambda < \infty$.
-Further, since $u$ is real valued, we have $\uhat_{-k} = \uhat_{k}^*$.
-Since momentum vanishes, we have $\uhat_0 = 0$.
-Meanwhile, Parseval's identity gives
+Since $u$ is real valued, we have $\uhat_{-k} = \uhat_{k}^*$ and since momentum vanishes $\uhat_0 = 0$.
+Next, consider the Galerkin projection
+\begin{align}
+u_{\Lambda}(x,t) = \Proj_{\Lambda} u =
+\sum_{\abs{k} \le \Lambda} \uhat_k e^{i k x} \, , \qquad
+\end{align}
+where $\Proj_{\Lambda}$ is a projection onto the finite dimensional space.
+
+Parseval's identity gives
 \begin{equation}

### 2019-08-06 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\Fcnl}{\mathcal{F}}
-\newcommand{\Proj}{\mathcal{P}}
+\newcommand{\Proj}{\mathcal{P}_{\Lambda}}
+\newcommand{\uL}{u_{\Lambda}}
 % Gibbs and theta
@@
 &u(x,t) = \sum_{k=-\infty}^{\infty} \uhat_k e^{i k x} \, , \\
+\label{uhat}
 &\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
@@
 Since $u$ is real valued, we have $\uhat_{-k} = \uhat_{k}^*$ and since momentum vanishes $\uhat_0 = 0$.
-Next, consider the Galerkin projection
+Next, consider the Galerkin truncation at wave number $\Lambda$
 \begin{align}
-u_{\Lambda}(x,t) = \Proj_{\Lambda} u =
+\uL(x,t) = \Proj u =
 \sum_{\abs{k} \le \Lambda} \uhat_k e^{i k x} \, , \qquad
 \end{align}
-where $\Proj_{\Lambda}$ is a projection onto the finite dimensional space.
+where $\Proj$ is a projection operator and \eqref{uhat} still holds. Projecting the KdV equation \eqref{dimlessKdV} onto the finite dimensional space gives the truncated KdV equation (TKdV)
+\begin{align}
+\label{TKdV}
+&\pd{\uL}{t} +  \frac{1}{2} C_3 \drat^{-3/2} \,\Proj (\uL)^2 + C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = 0
+\qquad \text{for } x \in [-\pi,\pi] \\
+&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1} \, , \quad
+C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
+\end{align}
+Note the additional projection operator in front of the quadratic term $\uL^2$, which removes the aliased modes of wavenumber larger than $\Lambda$.
+The constants are the same as before and have been repeated here for convenience.
 

### 2019-08-06 paper2_JNLS/paper.tex
NM

@@
 %----------------------------------------------------------------------------%
-\subsection{Hamiltonian structure}
+\subsection{Hamiltonian structure of KdV}
+\label{HamiltonianSection}
 
@@
 \begin{align}
-\dot{u}_{\pm} = \sympJ \vard{\Hupdn}{u}
+\partial_t{u}_{\pm} = \sympJ \vard{\Hupdn}{u}
 \end{align}
@@
 \begin{align}
+\label{H3H2}
 \Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
@@
 The constants are the same as before and have been repeated here for convenience.
+
+The TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator.
+The piecewise defined Hamiltonian for TKdV is given by
+\begin{equation}
+\label{TruncHamiltonian}
+\Hupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
+\end{equation}
+where $\Hthree$ and $\Htwo$ the defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
+Then TKdV \eqref{TKdV} can be expressed as
+\begin{align}
+\partial_t {u}_{\pm} = \partial_x \Proj \vard{\Hupdn}{\uL}
+\end{align}
+
+% Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.
 

### 2019-08-06 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\Ham}{\mathcal{H}}
-\newcommand{\Hup}{\Ham_{-}}
-\newcommand{\Hdn}{\Ham_{+}}
-\newcommand{\Hupdn}{\Ham_{\pm}}
+\newcommand{\Hup}{\Ham^{-}}
+\newcommand{\Hdn}{\Ham^{+}}
+\newcommand{\Hupdn}{\Ham^{\pm}}
 \newcommand{\Hthree}{\Ham_{3}}
@@
 \newcommand{\Fcnl}{\mathcal{F}}
+% Truncated stuff
 \newcommand{\Proj}{\mathcal{P}_{\Lambda}}
 \newcommand{\uL}{u_{\Lambda}}
+\newcommand{\HLupdn}{\Ham_{\Lambda}^{\pm}}
+\newcommand{\SympL}{\sympJ_{\Lambda}}
 % Gibbs and theta
@@
 \begin{align}
-\partial_t{u}_{\pm} = \sympJ \vard{\Hupdn}{u}
+\partial_t{u} = \sympJ \vard{\Hupdn}{u}
 \end{align}
@@
 
-The TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator.
+Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019statistical}.
 The piecewise defined Hamiltonian for TKdV is given by
@@
 \label{TruncHamiltonian}
-\Hupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
+\HLupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
 \end{equation}
@@
 \begin{align}
-\partial_t {u}_{\pm} = \partial_x \Proj \vard{\Hupdn}{\uL}
+\partial_t {\uL} = \partial_x \Proj \, \vard{\HLupdn}{\uL}
 \end{align}
-
+where the new symplectic operator is $\SympL = \partial_x \Proj$.
 % Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.
 
-Parseval's identity gives
+The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalization values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity gives
 \begin{equation}
-\En[u] = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2 = 1
+\En[\uL] = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2 = 1
 \end{equation}

### 2019-08-06 paper2_JNLS/paper.tex
NM

@@
 %----------------------------------------------------------%
-\section{The truncated KdV framework}
+\section{Theoretical framework}
 
@@
 %----------------------------------------------------------------------------%
-\subsection{Nondimensionalization of variable-depth KdV}
+\subsection{Nondimensionalization and relation to experimental scales}
 
@@
 where again $\Ham = \Hup$ if $t<0$ and $\Ham = \Hdn$ if $t>0$.
+
 %----------------------------------------------------------------------------%
-% Truncation
+% Truncated KdV
 %----------------------------------------------------------------------------%
-\subsection{The truncated KdV framework}
+\subsection{Truncated KdV}
 \label{tKdVSec}
 
-Consider the Fourier series of the state variable
+We now introduce the truncated KdV (TKdV) system, which is the main focus of this study. Consider the Fourier series of the state variable
 \begin{align}
@@
 \end{align}
-where the new symplectic operator is $\SympL = \partial_x \Proj$.
+where the truncated symplectic operator is $\SympL = \partial_x \Proj$.
 % Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.
@@
 % The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
+As has been noted before, the (untruncated) KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}.
+The projected counterparts of these invariants, however, are not necessarily invariants of the truncated system.
 
-The KdV system (i.e.
-$\Lambda = \infty$) possesses the Hamiltonian \eqref{Hamiltonian} and momentum and energy \eqref{MomEn} and an infinite number of additional invariants \cite{lax1975periodic, whitham2011linear}.
-For the truncated system however, $\Lambda < \infty$, the only known invariants are the momentum, energy, and Hamiltonian.
-The infinite sequence of invariants apart from these first three may no longer be invariants of TKdV.
+% Is it true that the only 3 known invariants of TKdV are the ones mentioned here: momentum, energy, and Hamiltonian.
 
-WRITE PROJECTION OPERATOR AND TKDV
-
-\vsp{10}
 %----------------------------------------------------------------------------%
-% Hamiltonian
+% Gibbs
 %----------------------------------------------------------------------------%
-\subsection{Hamiltonian structure of TKdV}
+\subsection{Mixed microcanonical-canonical Gibbs ensemble}
 
@@
 Specifically, this ensemble is microcanonical in the energy, with a fixed energy value, and it is canonical in the Hamiltonian.
-By fixing the energy and hence confining to a compact set, this construction avoids the divergence at infinity that would occur for a simple canonical ensemble.
-The resulting distribution is thus normalizable.
-Accordingly, each Hamiltonian $\Ham^{\pm}$ generates a corresponding mixed ensemble, or Gibbs measure given by 
+By fixing the energy and hence confining to a compact set, this construction avoids the divergence at infinity that would occur for a simple canonical ensemble and hence yields a normalizable distribution.
+This framework applies equally well to the truncated or untruncated system and hence we will not distinguish between the two.
+
+The Hamiltonian on either side of the ADC, $\Ham^{\pm}$, generates a corresponding mixed ensemble, or Gibbs measure given by 
 \begin{align}
@@
 Recall that the abrupt depth change is met by traveling waves at some time $t = T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
-The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales (specifically, $t \gg \delta^{-2}$ for dimensionless time).
+The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales, $t \gg \delta^{-2}$.
 Meanwhile, wave evolution over shorter timescales, $t \ll \delta^{-2}$, is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e.

### 2019-08-06 paper2_JNLS/paper.tex
NM

@@
 
-Introduce waves. \\
+Though once dismissed as sailor's tales, rogue waves are now recognized as a real and present threat to naval structures (cite).
+These abnormally large waves are inherently tied to emergence of non-Gaussian statistics in the sea surface.
+
+%Introduce waves. \\
 Introduce papers: PRF, PNAS, and J Stat Phys. \\

### 2019-08-07 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\Htwo}{\Ham_{2}}
-\newcommand{\Fcnl}{\mathcal{F}}
+%\newcommand{\Fcnl}{\mathcal{F}}
 % Truncated stuff
@@
 
-The evolution of any functional $\Fcnl[u]$ is given by 
+The evolution of any functional $\mathcal{F}[u]$ is given by 
 \begin{align}
-\partial_t \Fcnl = \{ \Fcnl, \Ham \} = \int_{-\pi}^{\pi} \vard{F}{u} \sympJ \vard{\Ham}{u} \dx
+\partial_t \mathcal{F} = \{ \mathcal{F}, \Ham \} = \int_{-\pi}^{\pi} \vard{\mathcal{F}}{u} \sympJ \vard{\Ham}{u} \dx
 \end{align}
-where again $\Ham = \Hup$ if $t<0$ and $\Ham = \Hdn$ if $t>0$.
+where again $\Ham = \Hup$ if $t<0$ and $\Ham = \Hdn$ if $t>0$ and $\{\}$ is the Poisson bracket induced by $\sympJ$.
 

### 2019-08-07 paper2_JNLS/paper.tex
NM

@@
 
-Though once dismissed as sailor's tales, rogue waves are now recognized as a real and present threat to naval structures (cite).
-These abnormally large waves are inherently tied to emergence of non-Gaussian statistics in the sea surface.
+Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height.
+Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
+Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir.
+The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; i.e.~When governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when statistics deviate from Gaussian.
 
+A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from variations in bottom topography.
+
+% Since studies are 1D topography, they offer bare minimum set of conditions able to generate anomalous waves; offers paradigm problem that can be studied scrutinized; shares features with more complex situations.
+
+% ideal, paradigm, canonical
+
+\vsp{4}
 %Introduce waves. \\
@@
 In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in \cite{majda2019statistical} for the sake of brevity.
+%more transparent connection to experimental parameters
 \item Provide additional experimental measurements not presented in \cite{bolles2019anomalous}. For example, statistics on surface slope and autocorrelation.
@@
 \end{align}
-where again $\Ham = \Hup$ if $t<0$ and $\Ham = \Hdn$ if $t>0$ and $\{\}$ is the Poisson bracket induced by $\sympJ$.
+where again $\Ham = \Hup$ if $t<0$, and $\Ham = \Hdn$ if $t>0$ and $\{\}$ is the Poisson bracket induced by $\sympJ$.
 

### 2019-08-07 paper2_JNLS/paper.tex
NM

@@
 Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir.
-The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; i.e.~When governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when statistics deviate from Gaussian.
+The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; i.e.~When governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.
 
 A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from variations in bottom topography.
-
-% Since studies are 1D topography, they offer bare minimum set of conditions able to generate anomalous waves; offers paradigm problem that can be studied scrutinized; shares features with more complex situations.
-
-% ideal, paradigm, canonical
+Since topographical variations are strictly one-dimensional in these studies, they can be considered as offering a bare minimum set of conditions capable of generating anomalous wave statistics.
+In particular, the more complex mechanism of focusing by 2D bottom topography absent.
+The studies thus offer the simplest conceivable system that exhibits the anomalous statistical behavior present in more complex systems, like the ocean, but in a tractable system that could be analyzed and quantified.
+% ideal, paradigm, canonical, tractable
 

### 2019-08-07 paper2_JNLS/paper.tex
NM

@@
 Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir.
-The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; i.e.~When governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.
+The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement.
+That is, when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.
 
-A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from variations in bottom topography.
+A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
 Since topographical variations are strictly one-dimensional in these studies, they can be considered as offering a bare minimum set of conditions capable of generating anomalous wave statistics.
 In particular, the more complex mechanism of focusing by 2D bottom topography absent.
-The studies thus offer the simplest conceivable system that exhibits the anomalous statistical behavior present in more complex systems, like the ocean, but in a tractable system that could be analyzed and quantified.
+The studies thus offer a paradigm system, with emergent anomalous statistical properties similar to that seen in more complex systems, but in a tractable context that is amenable to analysis.
+
+Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field that crosses an abrupt depth change (ADC).
+In these experiments, an incoming wave-field was generated with nearly Gaussian statistics.
+A plexiglass step created the abrupt depth change.
+Upon passing into a region of shallower depth, the wave distribution skewed strongly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
+Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} laid out a theoretical foundation that accurately captures several aspects of the anomalous behavior.
+The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
+In follow-up work, Majda and Qi (cite) performed subsequent analysis on severely truncated, low-dimensional systems that were shown to exhibit many of the same anomalous features.
+
 % ideal, paradigm, canonical, tractable
+% ABRUPT depth change
 
 \vsp{4}
-%Introduce waves. \\
 Introduce papers: PRF, PNAS, and J Stat Phys. \\

### 2019-08-07 paper2_JNLS/paper.tex
NM

@@
 
-% ideal, paradigm, canonical, tractable
-% ABRUPT depth change
+The purpose of the present manuscript is to expand on both the experimental findings of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical}, with particular emphasis on the synergy between the two; e.g.~model developments inspired by experimental findings, and new experimental tests inspired by theoretical predictions.
+We present a range of previously unpublished experimental data, including statistics of the surface {\em slopes} (as opposed to only displacement data as published in \cite{bolles2019anomalous}, autocorrelation data, and some data on experiments in which the depth ratio was varied.
+We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.
+
+%~experimental discoveries guiding model developments, as well as new theoretical predictions inspiring new experimental tests.
+
+% combination
+% draw close comparison
+%with an even closer comparison between theoretical predictions and experimental measurements.
 
 \vsp{4}
-Introduce papers: PRF, PNAS, and J Stat Phys. \\
+
+% ideal, paradigm, canonical, tractable
 

### 2019-08-08 paper2_JNLS/paper.tex
NM

@@
 Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir.
-The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement.
-That is, when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.
+The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.
 
@@
 
-Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field that crosses an abrupt depth change (ADC).
-In these experiments, an incoming wave-field was generated with nearly Gaussian statistics.
-A plexiglass step created the abrupt depth change.
-Upon passing into a region of shallower depth, the wave distribution skewed strongly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
+Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field upon encountering an abrupt depth change (ADC).
+In these experiments, an incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
+Upon passing into the region of shallower depth, the wave distribution skews strongly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
 Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} laid out a theoretical foundation that accurately captures several aspects of the anomalous behavior.
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
-In follow-up work, Majda and Qi (cite) performed subsequent analysis on severely truncated, low-dimensional systems that were shown to exhibit many of the same anomalous features.
+In follow-up work, Majda and Qi (cite) performed analysis on more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous features.
 
-The purpose of the present manuscript is to expand on both the experimental findings of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical}, with particular emphasis on the synergy between the two; e.g.~model developments inspired by experimental findings, and new experimental tests inspired by theoretical predictions.
+The purpose of the present manuscript is to expand on both the experimental findings of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical}, with particular emphasis on the synergy between the two; e.g.~experimental discoveries guiding theoretical developments and theoretical predictions inspiring new experimental tests.
 We present a range of previously unpublished experimental data, including statistics of the surface {\em slopes} (as opposed to only displacement data as published in \cite{bolles2019anomalous}, autocorrelation data, and some data on experiments in which the depth ratio was varied.

### 2019-08-22 paper2_JNLS/paper.tex
NM

@@
 \lengthscale = \frac{\lamfac \lam}{\pi} \, , \qquad
-\timescale = \frac{\lamfac \lam}{\freqp \depth}
+\timescale = \frac{\lamfac \lam}{\pi \freqp \depth}
 \end{align}
@@
 \qquad \text{for } x \in [-\pi,\pi] \\
-&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1} \, , \quad
-C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
+&C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad
+C_2 = \frac{\pi^2 \delup}{6 \lamfac^2} 
 \end{align}
@@
 2019 found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019statistical}.
-We have chosen the sign convention above so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.
+We have chosen the sign convention above (which also happens to be consistent with Lax \cite{lax1975periodic}) so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.
 
@@
 \qquad \text{for } x \in [-\pi,\pi] \\
-&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1} \, , \quad
-C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
+&C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad
+C_2 = \frac{\pi^2 \delup}{6 \lamfac^2} 
 \end{align}
 Note the additional projection operator in front of the quadratic term $\uL^2$, which removes the aliased modes of wavenumber larger than $\Lambda$.
-The constants are the same as before and have been repeated here for convenience.
+The constants $C_3$ and $C_2$ are the same as before and have been repeated here for convenience.
 
@@
 \begin{align}
-\Gupdn = C_{\thupdn} \exp(-\thupdn \Hupdn) \delta(\En - \pi)
+\Gupdn = C_{\thupdn} \exp(-\thupdn \Hupdn) \delta(\En - 1)
 \end{align}
@@
 \label{H3H2}
-\frac{\meandn{H_3^+}} {\meandn{H_2^+} - \meanup{H_2^+}} = \frac{C_2}{C_3} \dratdn^2
+\frac{\meandn{\Hthree^+}} {\meandn{\Htwo^+} - \meanup{\Htwo^+}} = \frac{C_2}{C_3} \dratdn^2
 \end{equation}
@@
 \begin{align}
-\mean{H_3^+} = \frac{\pi}{3} \mean{u^3} = 
+\mean{\Hthree^+} = \frac{\pi}{3} \mean{u^3} = 
 \frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
@@
 \begin{align}
-&\mean{H_2^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
+&\mean{\Htwo^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
 \left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
@@
 \begin{equation}
-\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = 
+\frac{\meandn{\Hthree}} {\meandn{\Htwo} - \meanup{\Htwo}} = 
 \frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 

### 2019-08-23 paper2_JNLS/paper.tex
NM

@@
 Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir.
-The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.
+The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.
 
 A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
-Since topographical variations are strictly one-dimensional in these studies, they can be considered as offering a bare minimum set of conditions capable of generating anomalous wave statistics.
+Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous wave statistics.
 In particular, the more complex mechanism of focusing by 2D bottom topography absent.

### 2019-09-13 code_rescaledModel/SolverKdV_SymplecticM4a_MC.m
NM

@@ -134,5 +134,3 @@ for ii=1:Nt
     if any(abs(u(:))>ulim),break,end
 end
 
-
-

### 2019-09-13 code_rescaledModel/SolverKdV_SymplecticM4a_MC.m
NM

@@ -1,4 +1,4 @@
-function SolverKdV_SymplecticM4a_MC(MC, um,Nw, theta,gibd, fi)
+function SolverKdV_SymplecticM4a_MC(MC, um, Nw, theta, gibd, fi)
 % This script solves the rescaled KdV equation in a periodic domain
 % using pseudo-spectral & symplectic M4a integrator
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-09-18 paper2_JNLS/paper.tex
NM

@@
 \begin{document}
-%Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change
+
 \title{The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes}
 
+% Title ideas
+% Could include words "theory and experiments".
+%Old idea: Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change
 \author{
@@
 
+%IDEA: Could comment on synergy, or close cooperation, between theory and experiments and the success thereof, then cite Ristroph erosion paper.
+
 % combination
 % draw close comparison
-%with an even closer comparison between theoretical predictions and experimental measurements.
+% with an even closer comparison between theoretical predictions and experimental measurements.
 
@@
 \ampscale = \pi^{1/2} \, \etastd \, , \qquad
-\lengthscale = \frac{\lamfac \lam}{\pi} \, , \qquad
-\timescale = \frac{\lamfac \lam}{\pi \freqp \depth}
+\lengthscale = \frac{\lamfac \lam}{2 \pi} \, , \qquad
+\timescale = \frac{\lamfac \lam}{2 \pi \freqp \depth}
 \end{align}
@@
 If only integer multiples of $\freqp$ were imposed (e.g.~lower frequencies were not present), then the forcing would produce waves that are periodic over lengthscale $\lam$.
-Since lower frequencies do exist, strict periodicity is not satisfied, but rather waves may be nearly periodic over the physical domain $\xi \in [-\lam,\lam]$.
-The approximation of near-periodicity becomes more accurate if integer multiples of $\lam$ are considered, i.e.~$\xi \in [-\lamfac \lam, \lamfac \lam]$.
+Since lower frequencies do exist, strict periodicity is not satisfied, but rather waves may be nearly periodic over the physical domain $\xi \in [-\lam/2, \lam/2]$.
+The approximation of near-periodicity becomes more accurate if integer multiples of $\lam$ are considered, i.e.~$\xi \in [-\lamfac \lam/2, \lamfac \lam/2]$.
 Thus, we have chosen $\lengthscale$ above so that the dimensionless domain of consideration is $\tilde{x} \in [-\pi, \pi]$ and periodic boundary conditions on $u$ can be  imposed on over this domain with accuracy that increases with $\lamfac$.
@@
 The KdV equation provides the first correction to this behavior and describes dynamics that evolve over longer timescales.
-Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/\depth \gg 1$, which provides a suitably long timescale and is in line with previous normalizations \cite{johnson1997modern}.
+Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitably long timescale and is in line with previous normalizations \cite{johnson1997modern}.
 The scales $\lengthscale$ and $\timescale$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.
@@
 &C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad
-C_2 = \frac{\pi^2 \delup}{6 \lamfac^2} 
+C_2 = \frac{2 \pi^2 \delup}{3 \lamfac^2} 
+%C_2 = \frac{\pi^2 \delup}{6 \lamfac^2} 
 \end{align}
@@
 &C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad
-C_2 = \frac{\pi^2 \delup}{6 \lamfac^2} 
+C_2 = \frac{2 \pi^2 \delup}{3 \lamfac^2}
 \end{align}

### 2019-09-18 code_rescaledModel/SolverKdV_SymplecticM4a_MC.m
NM

@@ -35,7 +35,7 @@ if fi == 0     % fi == 0 for upstream flow
 else
     D0 = Dp;
 end
-C2 = pi^2*del0/(6*Nw^2)
+C2 = (2/3)*pi^2*del0/Nw^2
 C3 = 3/2*pi^(1/2)*epsi0/del0
 
 %%------------------------------------------%%

### 2019-09-18 paper2_JNLS/paper.tex
NM

@@
 
+Briefly, consider the parameter $\lamfac$, the number of characteristic wavelengths in the physical domain.
+We require $\lamfac \le \Lambda$, so that the mode $\uhat_{\lamfac}$ corresponding to the physical wavelength $\lambda$ is resolved.
+In the case $\lamfac = \Lambda$, $\lambda$ corresponds to the smallest resolved wavelength.
+Often, we will choose $\lamfac = \Lambda$ or $\Lambda/2$.
+
 Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019statistical}.

### 2019-09-20 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
+\newcommand{\tavg}[1]{\overline{#1}}
 
@@
 \begin{align}
-\Gupdn = C_{\thupdn} \exp(-\thupdn \Hupdn) \delta(\En - 1)
+\Gupdn = Z_{\thupdn}^{-1} \exp(-\thupdn \Hupdn) \delta(\En - 1)
 \end{align}
-Here $\theta = \thupdn$ is the inverse temperature and $C_{\theta}$ a constant (inverse of the partition function) that depends on $\theta$.
-Each measure has corresponding ensemble mean $\mean{}_{\pm}$.
+% Note: The PNAS paper uses C_theta^{1}. The Majda, Qi Stat Phys paper uses C_theta sometimes and C_theta^{-1} in other places (page 11). So I'll just pick the one I like better.
+Here $\theta = \thupdn$ is the inverse temperature and $Z_{\theta}$ a constant (the partition function) that depends on $\theta$.
+Each measure $\Gupdn$ induces a corresponding ensemble average $\mean{}_{\pm}$.
+Throughout this paper we use a bracket for ensemble average and overbar for a time average; i.e.~for any quantity $Q$,
+\begin{align}
+&\mean{Q} = \int_{\Omega}  Q \, d \Gibbs
+&&\mbox{\em ensemble average} \\
+&\tavg{Q} = \frac{1}{T_2-T_1} \int_{T_1}^{T_2} Q \, dt	 
+&&\mbox{\em time average}
+\end{align}
 

### 2019-09-20 paper2_JNLS/notes.tex
NM

--- 2019-09-18 paper2_JNLS/paper.tex
@@
 \newcommand{\np}{\newpage \noindent}
+\newcommand{\seclabel}[1]{\vsp{1} {\bf#1:}}
+
 % Derivatives
@@
 \newcommand{\dratupdn}{\drat_{\pm}}
+\newcommand{\DD}{\dratdn}
 % Statistical quantities
@@
 \newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
+\newcommand{\meanz}[1]{\mean{#1}_0}
 
@@
 
-\title{The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes}
+\title{Notes on JNLS paper}
 
-% Title ideas
-% Could include words "theory and experiments".
-%Old idea: Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change
 \author{
-C.~Tyler Bolles\thanks{University of Michigan},
-Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
-M.~N.~J.~Moore\thanks{Florida State University}, 
-Di Qi\thanks{Courant Institute of Mathematical Sciences} }
+M.~N.~J.~Moore\thanks{Florida State University} }
 \maketitle
 
-%----------------------------------------------------------%
-% Intro
-%----------------------------------------------------------%
-\section{Introduction}
+%----------------------------------------------------------------------------%
+% Transfer function
+%----------------------------------------------------------------------------%
+\section{Notes on the transfer function}
 
-Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height.
-Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
-Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir.
-The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.
-
-A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
-Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous wave statistics.
-In particular, the more complex mechanism of focusing by 2D bottom topography absent.
-The studies thus offer a paradigm system, with emergent anomalous statistical properties similar to that seen in more complex systems, but in a tractable context that is amenable to analysis.
-
-Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field upon encountering an abrupt depth change (ADC).
-In these experiments, an incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
-Upon passing into the region of shallower depth, the wave distribution skews strongly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
-Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} laid out a theoretical foundation that accurately captures several aspects of the anomalous behavior.
-The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
-In follow-up work, Majda and Qi (cite) performed analysis on more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous features.
-
-The purpose of the present manuscript is to expand on both the experimental findings of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical}, with particular emphasis on the synergy between the two; e.g.~experimental discoveries guiding theoretical developments and theoretical predictions inspiring new experimental tests.
-We present a range of previously unpublished experimental data, including statistics of the surface {\em slopes} (as opposed to only displacement data as published in \cite{bolles2019anomalous}, autocorrelation data, and some data on experiments in which the depth ratio was varied.
-We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.
-
-%~experimental discoveries guiding model developments, as well as new theoretical predictions inspiring new experimental tests.
-
-%IDEA: Could comment on synergy, or close cooperation, between theory and experiments and the success thereof, then cite Ristroph erosion paper.
-
-% combination
-% draw close comparison
-% with an even closer comparison between theoretical predictions and experimental measurements.
-
-\vsp{4}
-
-% ideal, paradigm, canonical, tractable
-
-The purpose of this manuscript is to:
-\begin{itemize}
-\item Provide a more comprehensive treatment of the experiments and theory in combination.
-In particular, we make new comparisons between theory and experiments, which further confirm the predictive power of the theoretical framework.
-\item Provide more thorough treatment of the link between physical system parameters and model parameters.
-This will elucidate the connection between theory and experiments.
-In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in \cite{majda2019statistical} for the sake of brevity.
-%more transparent connection to experimental parameters
-\item Provide additional experimental measurements not presented in \cite{bolles2019anomalous}. For example, statistics on surface slope and autocorrelation.
-\end{itemize}
-
-%----------------------------------------------------------%
-% Experiments
-%----------------------------------------------------------%
-\section{The experiments}
-
-	As discussed in \cite{bolles2019anomalous}, the experiments consist of a long, narrow wave tank (6 m long x 20 cm wide x 30 cm high), with waves generated by plexiglass paddle at one end.
-The waves propagate through the tank and, roughly midway through, pass over an abrupt depth change (ADC), which is created by a plexiglass step.
-The waves continue to propagate through the shallower depth until reaching the far end of the tank, at which point their energy is dissipated by a horse-hair dampener.
-Since the dampener minimizes the backscatter, the waves in this experiment propagate primarily in one direction, from left to right in Fig.~\ref{fig1}.
-
-The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized incoming wave field, the paddle angle $\phi$ is specified by a precomputed psuedo-random signal
+Consider the transfer function that maps the upstream inverse temperature to the downstream one
+\begin{equation}
+\thdn = F(\thup)
+\end{equation}
+This transfer function results from the matching condition
 \begin{align}
-\label{PaddleAngle}
-& \phi(t) = \phi_0 + \Delta \phi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
-\label{anEq}
-& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
-\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
+\label{statmatch}
+&\meanup{\Hdn} = \meandn{\Hdn}
 \end{align}
-(Literal copy and paste) Here, the angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively.
-We set $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz.
-Importantly, the phases $\delta_n$ are uniformly distributed random variables.
-We fix $N = 3000$, which sets a fundamental period of $T = 300$ seconds.
-
- % Figure 1
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[!ht]
-\begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/fig1.pdf}
-\caption{
-(a) Experimental schematic. Reproduced from \cite{bolles2019anomalous}
-}
-%(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms.
-\label{fig1}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-	The free surface is illuminated by light-emitting diodes running along the bottom of the tank and then imaged from the sideview with a Nikon D3300 at 60 frames per second.
-The illumination technique coupled with high pixel count of the camera allows surface displacements to be resolved with a accuracy better than 1/3 millimeter.
-Representative free surface measurements are shown in Figs.~\ref{fig2}(a)--(b) at a position upstream and downstream of the ADC.
-The corresponding histograms are shown in Figs.~\ref{fig2}(c)--(d).
-
- % Figure 2
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[!ht]
-\begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/fig2.pdf}
-\caption{
-(a)--(b) Surface displacement measured at a representative upstream and downstream location. (c)--(d) Corresponding histograms. Reproduced from \cite{bolles2019anomalous}
-}
-\label{fig2}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
- % Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees, which is in between runs 5 and 6.
-
-  % Figure 3
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[!ht]
-\begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/fig3.pdf}
-\caption{
-(b)--(c) Surface slope measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms. NEW PLOT.
-}
-\label{fig3}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-% Figure 4
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[!ht]
-\begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/fig4.pdf}
-\caption{
-Spatial statistics from the experiments.
-}
-\label{fig4}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-%----------------------------------------------------------%
-% Theory
-%----------------------------------------------------------%
-\section{Theoretical framework}
-
-%----------------------------------------------------------------------------%
-% KdV
-%----------------------------------------------------------------------------%
-\subsection{The Korteweg–de Vries equation with variable depth}
-Consider waves propagating unidirectionally in shallow water.
-Consider the surface displacement $\eta(x,t)$ and the reference frame moving with the local wave speed $\xi = x - ct$, where $c = \sqrt{g \depth}$ is the wave speed, $g$ gravity, and $\depth$ the local depth.
-To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation (KdV), which in dimensional form is given by \cite{whitham2011linear}
-\begin{equation}
-\label{KdV}
-\eta_t + \frac{3 c}{2 \depth} \eta \eta_{\xi} + \frac{c \depth^2}{6} \eta_{\xi \xi \xi} = 0
-\end{equation}
-% This dimensional form is given by Whitham on p. 461, Eq. 13.98. I first got it from the Wolfram KdV, which gives the same equation.
-
-We will primarily consider the case in which waves originate from a region of constant depth, encounter an abrupt depth change, and then continue in another region of constant depth.
-Thus, depth will be piecewise constant
+The Hamiltonian is given by
 \begin{align}
-\depth = 
-\begin{cases}
-\dup \quad \mbox{if } x<0 \\
-\ddn \quad \mbox{if } x>0
-\end{cases}
-\end{align}
-Most often, we will consider waves moving into shallower depth, so that $\dup > \ddn$.
-
-The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$, which gives rise to the following quantities
-\begin{align}
-&c = \sqrt{g \depth}
-&&\mbox{\em local wave speed} \\
-&\lam = c/\freqp = \sqrt{g \depth} / \freqp
-&&\mbox{\em local peak wave length} \\
-&\etastd = \mean{\eta^2}^{1/2} 
-&&\mbox{\em standard deviation of displacement} \\
-\end{align}
-where $\mean{\cdot}$ indicates the mean of a quantity. 
-Note that the characteristic wave speed $c = c_{\pm}$ and wavelength $\lam = \lamupdn$ each take different values upstream and downstream of the ADC.
-We remark that experimental measurements indicate that $\etastd = \mean{\eta^2}^{1/2}$ is nearly the same on both values of the ADC.
-Hence, we will not distinguish between upstream and downstream values of $\etastd$.
-%
-See Table \ref{paramtable} for a summary of important parameters.
-
-% Table
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{table}[h]%[htbp]
-\begin{center}
-\caption{Table of parameters}
-\label{paramtable}
-\begin{tabular}{l l l}
-\hline \multicolumn{3} { c }{Parameters that are constant in a single experiment} \\
-\hline Description & Notation and definition & Value in experiments \\
-\hline
-Peak forcing frequency		& $f_p$						& 2 Hz \\
-Characteristic wave amplitude	& $\etastd = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
-Upstream depth			& $\dup$						& 12.5 cm \\
-Downstream depth			& $\ddn$						& 3 cm (and varied) \\
-Upstream wavelength		& $\lamup = \sqrt{g \dup}/f_p$		& 55 cm \\
-Upstream wavelength		& $\lamdn = \sqrt{g \ddn}/f_p$		& 27 cm \\
-%
-Amplitude-to-depth ratio		& $\epsup = \etastd / \dup$			& 0.0024--0.024 \\
-Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
-Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 (and varied)
-\end{tabular}
-\end{center}
-\end{table}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-%----------------------------------------------------------------------------%
-% Nondimensionalization
-%----------------------------------------------------------------------------%
-\subsection{Nondimensionalization and relation to experimental scales}
-
-In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the statistical-mechanics framework introduced in \cite{majda2019statistical}.
-Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other choices.
-To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment.
-We introduce the dimensionless variables
-\begin{align}
-&u = \eta / \ampscale
-&&\mbox{\em dimensionless surface displacement} \\
-&\tilde{x} = \xi / \lengthscale
-&&\mbox{\em dimensionless position (in moving frame)} \\
-&\tilde{t} = t / \timescale
-&&\mbox{\em dimensionless time}
-\end{align}
-Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
-\begin{equation}
-u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
-+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
-\end{equation}
-We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.
-
-Now it is possible to choose the scales $\ampscale, \lengthscale, \timescale$ for ease in working with a particular framework. We choose the following scales,
-\begin{align}
-\label{scales}
-\ampscale = \pi^{1/2} \, \etastd \, , \qquad
-\lengthscale = \frac{\lamfac \lam}{2 \pi} \, , \qquad
-\timescale = \frac{\lamfac \lam}{2 \pi \freqp \depth}
-\end{align}
-where $\lamfac$ is an integer to be chosen later. 
-The explanation for these choices is as follows.
-First, we have chosen the characteristic amplitude, $\ampscale$, above to normalize the energy of the state-variable $u$ to unity, as will be demonstrated in Section \ref{tKdVSec}.
-Second, regarding $\lengthscale$, recall that $\lam$ is the characteristic wavelength corresponding to the peak forcing frequency $\freqp$ in the experiments.
-If only integer multiples of $\freqp$ were imposed (e.g.~lower frequencies were not present), then the forcing would produce waves that are periodic over lengthscale $\lam$.
-Since lower frequencies do exist, strict periodicity is not satisfied, but rather waves may be nearly periodic over the physical domain $\xi \in [-\lam/2, \lam/2]$.
-The approximation of near-periodicity becomes more accurate if integer multiples of $\lam$ are considered, i.e.~$\xi \in [-\lamfac \lam/2, \lamfac \lam/2]$.
-Thus, we have chosen $\lengthscale$ above so that the dimensionless domain of consideration is $\tilde{x} \in [-\pi, \pi]$ and periodic boundary conditions on $u$ can be  imposed on over this domain with accuracy that increases with $\lamfac$.
-Lastly, regarding $\timescale$, the most basic timescale in the experiments is simply $\freqp^{-1}$, i.e.~the period of waves passing a fixed reference point.
-Of course, the leading-order behavior in shallow water is simple translation of waves with uniform speed $c$.
-The KdV equation provides the first correction to this behavior and describes dynamics that evolve over longer timescales.
-Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitably long timescale and is in line with previous normalizations \cite{johnson1997modern}.
-The scales $\lengthscale$ and $\timescale$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.
-
-With the above choices, the dimensionless KdV takes the form
-\begin{align}
-\label{dimlessKdV}
-&u_t + C_3 \drat^{-3/2} \, u u_x + C_2 \drat^{1/2} \, u_{xxx} = 0
-\qquad \text{for } x \in [-\pi,\pi] \\
-&C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad
-C_2 = \frac{2 \pi^2 \delup}{3 \lamfac^2} 
-%C_2 = \frac{\pi^2 \delup}{6 \lamfac^2} 
-\end{align}
-The constants $C_3$ and $C_2$ do {\em not} change value crossing the ADC and are given in terms of the dimensionless parameters
-\begin{align}
-&\epsup = \etastd / \dup
-&&\mbox{\em upstream amplitude-to-depth ratio} \\
-&\delup = \dup / \lamup
-&&\mbox{\em upstream depth-to-wavelength ratio}
-\end{align}
-The reason for the subscripts $3$ and $2$ will become evident in the next section.
-
-Meanwhile, the dimensionless depth $\drat = {\depth}/{\dup}$ {\em does} change value across the ADC since the depth $\depth$ changes.
-Recall that the reference frame of \eqref{dimlessKdV} moves with the local wave speed via the variable $\xi = x-ct$ from \eqref{KdV}.
-Thus, the ADC is met at some time $T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
-Therefore, we can regard $\drat$ is a piece-wise constant function of dimensionless time
-\begin{equation}
-\label{dratpw}
-\drat = 
-\begin{cases}
-1 		&\quad \mbox{for } {t}<0 \\
-\dratdn = {\ddn}/{\dup} 	&\quad \mbox{for } {t}>0
-\end{cases}
-\end{equation}
-
-A few comments are in order.
-First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ in \eqref{dimlessKdV} but with different expressions for the other dimensionless parameters.
-These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments.
-Second, an alternate formulation of the variable-depth KdV equation exists in which it is conjectured that the product $\depth^{1/4} \eta$, rather than $\eta$, varies continuously across the ADC \cite{johnson1997modern}.
-Of course, that assumption implies a discontinuity in surface displacement, which, though small, is physically unrealistic.
-We have chosen to enforce continuity of surface displacement in the above on the basis of physical realism.
-Using the alternative formulation would simply modify the power of $\drat$ in the second term of \eqref{dimlessKdV} by 1/4, i.e.~$\drat^{-7/4}$ instead of $\drat^{-3/2}$.
-Hence, all of the results discussed below would still hold qualitatively in this alternate formulation, with only slight quantitative modifications.
-
-%NOTE: Perhaps we should be more specific about comparing to the dimensionless parameters in our PNAS paper \cite{majda2019statistical}, BUT the expressions for E0 and L0 given in the PNAS paper do not match with what I calculate comparing to the above.
-Plus, it seems like it could only add confusion.
-
-%----------------------------------------------------------------------------%
-% Hamiltonian
-%----------------------------------------------------------------------------%
-\subsection{Hamiltonian structure of KdV}
-\label{HamiltonianSection}
-
-The variable-depth KdV \eqref{dimlessKdV}, though not Hamiltonian throughout the entire domain, admits a Hamiltonian structure on each side of the ADC. Indeed, \eqref{dimlessKdV} can be expressed as
-\begin{align}
-\partial_t{u} = \sympJ \vard{\Hupdn}{u}
-\end{align}
-where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham = \Hupdn$ is the Hamiltonian, which will take different forms on either side of the ADC.
-It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
-\begin{align}
-\label{H3H2}
-\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
-\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
-\end{align}
-Then the Hamiltonian can be expressed as
-\begin{equation}
 \label{Hamiltonian}
-\Hupdn = C_2 \dratupdn^{1/2} \, \Htwo - C_3 \dratupdn^{-3/2} \, \Hthree
-\end{equation}
-where $\drat = \dratupdn$ changes value across the ADC. More explicitly, substituting \eqref{dratpw}, the separate upstream and downstream Hamiltonians are
-\begin{align}
-&\Hup = C_2 \, \Htwo - C_3 \, \Hthree 						&& \text{for } t<0 \\
-&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree	&& \text{for } t>0
-\end{align}
-The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$.
-With that sign convention, Majda et al.
-2019 found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019statistical}.
-We have chosen the sign convention above (which also happens to be consistent with Lax \cite{lax1975periodic}) so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.
-
-% Notes for myself on the sign change: Abramov CPAM 2003, Bajars Nonlinearity 2013, and our 2019 PNAS paper all use J = -d/dx with the opposite sign in the Hamiltonian.
-However, I was able to easily find several references online that use J = d/dx, so there is nothing wrong with it.
-The exact criteria that J has to satisfy are listed in the CPAM 2003 paper, and J = d/dx satisfies all these criteria just as well as J = -d/dx does.
-
-We introduce two important invariants of KdV, the momentum and the energy
-\begin{align}
-\label{MomEn}
-\Mo[u] \equiv \int_{-\pi}^{\pi} u \dx \, = 0 , \qquad
-\En[u] \equiv \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
-\end{align}
-As indicated above, the momentum of $u$ vanishes since it is measured as displacement from equilibrium.
-Second, by the scale chosen for $\ampscale$ in \eqref{scales}, the energy has been normalized to unity.
-
-The evolution of any functional $\mathcal{F}[u]$ is given by 
-\begin{align}
-\partial_t \mathcal{F} = \{ \mathcal{F}, \Ham \} = \int_{-\pi}^{\pi} \vard{\mathcal{F}}{u} \sympJ \vard{\Ham}{u} \dx
-\end{align}
-where again $\Ham = \Hup$ if $t<0$, and $\Ham = \Hdn$ if $t>0$ and $\{\}$ is the Poisson bracket induced by $\sympJ$.
-
-%----------------------------------------------------------------------------%
-% Truncated KdV
-%----------------------------------------------------------------------------%
-\subsection{Truncated KdV}
-\label{tKdVSec}
-
-We now introduce the truncated KdV (TKdV) system, which is the main focus of this study. Consider the Fourier series of the state variable
-\begin{align}
-&u(x,t) = \sum_{k=-\infty}^{\infty} \uhat_k e^{i k x} \, , \\
-\label{uhat}
-&\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
-\end{align}
-Since $u$ is real valued, we have $\uhat_{-k} = \uhat_{k}^*$ and since momentum vanishes $\uhat_0 = 0$.
-Next, consider the Galerkin truncation at wave number $\Lambda$
-\begin{align}
-\uL(x,t) = \Proj u =
-\sum_{\abs{k} \le \Lambda} \uhat_k e^{i k x} \, , \qquad
-\end{align}
-where $\Proj$ is a projection operator and \eqref{uhat} still holds. Projecting the KdV equation \eqref{dimlessKdV} onto the finite dimensional space gives the truncated KdV equation (TKdV)
-\begin{align}
-\label{TKdV}
-&\pd{\uL}{t} +  \frac{1}{2} C_3 \drat^{-3/2} \,\Proj (\uL)^2 + C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = 0
-\qquad \text{for } x \in [-\pi,\pi] \\
+&\Hupdn = C_2 \dratupdn^{1/2} \, \Htwo - C_3 \dratupdn^{-3/2} \, \Hthree \\
 &C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad
@@
 \end{align}
-Note the additional projection operator in front of the quadratic term $\uL^2$, which removes the aliased modes of wavenumber larger than $\Lambda$.
-The constants $C_3$ and $C_2$ are the same as before and have been repeated here for convenience.
+where
+\begin{align}
+\label{H3H2}
+&\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\\
+&\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx =
+2 \pi \sum_{k=1}^{\Lambda} \abs{k}^ 2\abs{\uhat_k}^2
+\end{align}
 
-Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019statistical}.
-The piecewise defined Hamiltonian for TKdV is given by
+\subsection{Scaling to make the transfer function relatively independent of $\Lambda$}
+{\bf Goal:} Ideally, we would like the transfer function $F$ to be approximately independent of $\Lambda$, or at least not exceedingly sensitive to $\Lambda$.
+This is the only way that we could reasonably match the values of $\theta$ to the experiment with some meaning.
+
+\seclabel{Issue with scaling of $\Htwo$}
+I believe the heart of the matter is that $\mean{\Htwo} \sim \Lambda^2$, which can make the matching condition and thus the transfer function too sensitive to $\Lambda$.
+As a demonstration, consider the uniform distribution, i.e.~Gibbs measure with zero inverse temperature, and the corresponding mean $\mean{}_0$.
+With this as the invariant measure, one can consider the equipartition of energy as a heuristic
+\begin{align}
+&\En[\uL] = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2 = 1 \\
+&\abs{\uhat_k}^2 \approx \frac{1}{2 \pi \Lambda}	 \qquad \mbox{for equipartition}
+\end{align}
+Then the expected value of $\Htwo$ is
 \begin{equation}
-\label{TruncHamiltonian}
-\HLupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
+\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} \abs{k}^ 2 \mean{\abs{\uhat_k}^2}_0 \approx \frac{1}{3} \Lambda^2
 \end{equation}
-where $\Hthree$ and $\Htwo$ the defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
-Then TKdV \eqref{TKdV} can be expressed as
+We have used the identity
+\begin{equation}
+\sum_{k=1}^{n} \abs{k}^ 2 = \frac{1}{6} n(n+1)(2n+1) \approx n^3/3
+\end{equation}
+% Like the Gauss sum.
+% Source: https://brilliant.org/wiki/sum-of-n-n2-or-n3/
+
+\seclabel{Resolution, scale $C_2$ appropriately}
+Recall the definition
+\begin{equation}
+C_2 = \frac{2 \pi^2 \delup}{3 \lamfac^2} 
+\end{equation}
+In particular, if we let $\lamfac \sim \Lambda$, then we will have
+\begin{equation}
+\mean{C_2 \Htwo}_0 \sim \Lambda^0 
+\end{equation}
+which removes the extreme sensitivity issue. More specifically, lets let $\Lambda = a \lamfac$ where $a = $ 1 or 2. Then we have
+\begin{equation}
+\mean{C_2 \Htwo}_0 \approx \frac{2}{9} \pi^2 a^2 \delup^2
+\end{equation}
+which is not sensitive to the truncation wavenumber $\Lambda$.
+
+\seclabel{Tests}
+I ran some tests and they seem to support my analysis that we should scale $
+\lamfac$ with $\Lambda$. The best looking results came from $\lamfac = \Lambda/2$ and $\lamfac = \Lambda/3$, where the later is integer division.
+
+\subsection{Approximating the transfer function}
+Consider the matching condition \eqref{statmatch}. The upstream mean is
 \begin{align}
-\partial_t {\uL} = \partial_x \Proj \, \vard{\HLupdn}{\uL}
+&\meanup{\Hdn} = \frac{\int \exp\left(-\thup \Hup\right) \Hdn d\mu}{ \int \exp\left(-\thup \Hup \right) d\mu}
 \end{align}
-where the truncated symplectic operator is $\SympL = \partial_x \Proj$.
-% Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.
+where $d \mu$ is the uniform measure on the surface $\En=1$. Similarly, we have the downstream mean
+\begin{align}
+&\meandn{\Hdn} = \frac{\int \exp\left(-\thdn \Hdn\right) \Hdn d\mu}{ \int \exp\left(-\thdn \Hdn \right) d\mu}
+\end{align}
+In both cases, we Taylor expand $\exp(-\theta \Ham) \sim 1 - \theta \Ham$ for small $\theta$. Enforcing the matching condition gives
+\begin{equation}
+\thup \int \Hup (\Hdn -1 ) d\mu = \thdn \int \Hdn (\Hdn - 1) d\mu
+\end{equation}
+Therefore an approximation to the Transfer function is 
+\begin{equation}
+\thdn = F(\thup) \approx m \thup
+\end{equation}
+where
+\begin{equation}
+m = \frac{  \int \Hup (\Hdn -1 ) d\mu }{ \int \Hdn (\Hdn - 1) d\mu }
+\end{equation}
+Using the definition of the Hamiltonian gives
+\begin{equation}
+m = \frac{ C_2^2 \DD^{1/2} \meanz{\Htwo^2} 
+	- C_2 C_3(\DD^{-3/2} + \DD^{1/2})  \meanz{\Hthree \Htwo} 
+	+ C_3^2 \DD^{-3/2} \meanz{\Hthree^2} - C_2 \meanz{\Htwo}}
+	{ C_2^2 \DD \meanz{\Htwo^2} 
+	- 2 C_2 C_3 \DD^{-1} \meanz{\Hthree \Htwo} 
+	+ C_3^2 \DD^{-3} \meanz{\Hthree^2} - C_2 \DD^{1/2} \meanz{\Htwo}}
+\end{equation}
 
-The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalization values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity gives
-\begin{equation}
-\En[\uL] = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2 = 1
-\end{equation}
-% The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
-As has been noted before, the (untruncated) KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}.
-The projected counterparts of these invariants, however, are not necessarily invariants of the truncated system.
+I believe that by symmetry $\meanz{\Hthree \Htwo} = 0$
+TO CONTINUE
 
-% Is it true that the only 3 known invariants of TKdV are the ones mentioned here: momentum, energy, and Hamiltonian.
-
-%----------------------------------------------------------------------------%
-% Gibbs
-%----------------------------------------------------------------------------%
-\subsection{Mixed microcanonical-canonical Gibbs ensemble}
-
-In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced in \cite{abramov2003hamiltonian} for the Burgers-Hopf system.
-Specifically, this ensemble is microcanonical in the energy, with a fixed energy value, and it is canonical in the Hamiltonian.
-By fixing the energy and hence confining to a compact set, this construction avoids the divergence at infinity that would occur for a simple canonical ensemble and hence yields a normalizable distribution.
-This framework applies equally well to the truncated or untruncated system and hence we will not distinguish between the two.
-
-The Hamiltonian on either side of the ADC, $\Ham^{\pm}$, generates a corresponding mixed ensemble, or Gibbs measure given by 
-\begin{align}
-\Gupdn = C_{\thupdn} \exp(-\thupdn \Hupdn) \delta(\En - 1)
-\end{align}
-Here $\theta = \thupdn$ is the inverse temperature and $C_{\theta}$ a constant (inverse of the partition function) that depends on $\theta$.
-Each measure has corresponding ensemble mean $\mean{}_{\pm}$.
-
-%Misc: The function that maximizes the Hamiltonian subject to the constraints of fixed energy and zero momentum is a traveling wave solution!
-
-%----------------------------------------------------------------------------%
-% Matching
-%----------------------------------------------------------------------------%
-\subsection{Matching at the ADC}
-
-Recall that the abrupt depth change is met by traveling waves at some time $t = T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
-The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales, $t \gg \delta^{-2}$.
-Meanwhile, wave evolution over shorter timescales, $t \ll \delta^{-2}$, is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e.
-no nonlinearity and no dispersion).
-Hence, a short instant after waves pass over the ADC, the waveform does not yet have time to alter significantly, which gives the (deterministic) matching condition
-\begin{align}
-&u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+} \qquad \text{for } x \in [-\pi, \pi]
-%&&\mbox{\em deterministic matching condition}
-\end{align}
-%or in Johnson's formulation, it is $\depth^{1/4} \eta$ that matches.
-
-Since $u$ matches for every single trajectory, it must also match in the statistical sense.
-Moreover, the downstream Hamiltonian $\Hdn$ is a functional of $u$ and hence must also match.
-This gives the {\em statistical matching condition}
-\begin{align}
-\label{statmatch}
-&\meanup{\Hdn} = \meandn{\Hdn}
-%&&\mbox{\em statistical matching condition}
-\end{align}
-Once $\Hdn$ downstream is determined, it is conserved in the outgoing wave field.
-
-%----------------------------------------------------------------------------%
-% Skewness formula
-%----------------------------------------------------------------------------%
-\subsection{Explicit formula for outgoing skewness}
-An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$.
-Inserting this approximation into the statistical matching condition \eqref{statmatch} and simplifying gives
-\begin{equation}
-\label{H3H2}
-\frac{\meandn{\Hthree^+}} {\meandn{\Htwo^+} - \meanup{\Htwo^+}} = \frac{C_2}{C_3} \dratdn^2
-\end{equation}
-%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
-To convert to dimensional, i.e.~experimental, values we first note that
-\begin{align}
-\mean{\Hthree^+} = \frac{\pi}{3} \mean{u^3} = 
-\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
-\end{align}
-Second, to convert the downstream surface slope, we note
-\begin{align}
-\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
-\end{align}
-where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
-Note: I believe it is important here to use $\lamdn$ in reverting to dimensional variables since we are matching $\Hdn$.
-Hence, we have
-\begin{align}
-&\mean{\Htwo^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
-\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
-\end{align}
-%
-Then algebra gives
-\begin{equation}
-\frac{\meandn{\Hthree}} {\meandn{\Htwo} - \meanup{\Htwo}} = 
-\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
-\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
-\end{equation}
-Also we have the simplification
-\begin{equation}
-\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
-\end{equation}
-Hence, \eqref{H3H2} gives the remarkably explicit formula
-\begin{equation}
-\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 =
-\frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
-\end{equation}
-In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
-% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.
-
-%----------------------------------------------------------%
-% Direct numerical simulations
-%----------------------------------------------------------%
-\section{Direct numerical simulations}
-
-\section{Sampling Algorithms}
-
-\subsection{Naive acceptance-rejection algorithm}
-
-\subsection{Markov-chain Monte Carlo}
-
-\subsection{Improved acceptance-rejection algorithm}
-
-%----------------------------------------------------------%
-% Comparison with experiments
-%----------------------------------------------------------%
-\section{Comparison with experiments}
-
-  % Figure
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[!ht]
-\begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/SkewAmp.pdf}
-\caption{
-Relationship predicted by the theory and confirmed by experiments.
-}
-\label{AAA}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-\subsection{Comparison of basic features}
-
-\subsection{New experimental measurements guided by theory}
-
-\section*{Acknowledgements}
-CTB acknowledges support from the IDEA grant at Florida State University, as well as from the Geophysical Fluid Dynamics Institute. 
-MNJM acknowledges support from the Simons Foundation, award 524259.
-
-\bibliographystyle{plain}
-\bibliography{wavesbib}
+\section{Testing the computation of the Hamiltonian in the Matlab DNS and Julia Sampling}
 

### 2019-09-20 paper2_JNLS/notes.tex
NM

@@
 I believe that by symmetry $\meanz{\Hthree \Htwo} = 0$
-TO CONTINUE
+\\ TO CONTINUE
 
@@
 
+\seclabel{Goal} We need some test problems for which we can compute the Hamiltonian exactly, to make sure it is done correctly in the Matlab DNS and Julia sampling codes.
+
+\seclabel{Test examples} \\
+1. $u = b \left( (a+\cos{x})^2 - (a^2+1/2) \right)$ \\
+i. The above satisfies $\Mo[u] = 0$ (due to the choice of constant $(a^2 + 1/2)$. \\
+ii. The choice $b = \sqrt{ \frac{8}{\pi(1+16 a^2)}}$ satisfies $\En[u] = 1$ \\
+Wolfram alpha gives that
+\begin{align}
+&\Htwo[u] = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx = 1 + \frac{3}{1+16 a^2} \\
+&\Hthree[u] = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx = 8 \sqrt{2} \pi a^2 \left( 16 \pi a^2 + \pi \right)^{-3/2}
+\end{align}
+$\bullet$ I tested the Matlab DNS code and it works! I tested $\Htwo$ and $\Hthree$ separately. Each gave a relative error on the order of machine precision.
+
 \end{document}

### 2019-09-20 code_rescaledModel/hamiltonian.m

@@ -1,4 +1,4 @@
-function ham = hamiltonian(uk,C2,C3,D0)
+function [ham,ham2,ham3] = hamiltonian(uk,C2,C3,D0)
 % Basic stuff.
 %% NOTE: JJ and MC should be defined from the size of uk, instead of from 
 %% a params variable, so that the code works for a single uk or an ensemble.
@@ -11,7 +11,9 @@ ham2 = 0.5 * sum(abs(kmat.*uk).^2) * 2*pi/JJ^2;
 ham3 = 1/6 * real(sum(u2k_dealiasing_MC(uk).*conj(uk))) * 2*pi/JJ^2;
 ham = C2*D0^(1/2)*ham2 - C3*D0^(-3/2)*ham3;
 
+% Note: The 1/JJ^2 factor comes from normalizing Matlab's FFt.
 
+%% Old Codes
 % SolverKdV_SymplecxticM4a_MC.m lines 102-103
 %k = repmat([0:J/2 -J/2+1:-1]',[1 MC]); % wavenumbers
 %temp = -(C3*D0^(-1)*1/6*real(sum(u2k_dealiasing_MC(uk,params).*conj(uk))) ...

### 2019-09-20 code_rescaledModel/testham.m

@@ -1,0 +1,24 @@
+clear all; close all;
+
+% Adjustable parameters.
+J = 16
+aa = 1.3
+
+% Define the test function
+xx = (-pi:2*pi/J:pi-2*pi/J)';
+bb = sqrt(8/(pi*(1+16*aa^2)));
+uu = bb*( (aa+cos(xx)).^2 - (0.5+aa^2) );
+uk = fft(uu)
+% Exact Hamiltonian values
+h2ex = 1 + 3/(1+16*aa^2);
+h3ex = 8*sqrt(2)*pi*aa^2 * (16*pi*aa^2 + pi)^(-3/2);
+% Compute the Hamiltonian with the routine
+[ham,ham2,ham3] = hamiltonian(uk,1.,1.,1.);
+
+ham2
+h2ex
+ham3
+h3ex
+
+h2relerr = abs(h2ex-ham2)/abs(h2ex)
+h3relerr = abs(h3ex-ham3)/abs(h3ex)

### 2019-09-21 code_rescaledModel/hamiltonian.m
N

@@ -1,4 +1,4 @@
-function [ham,ham2,ham3] = hamiltonian(uk,C2,C3,D0)
+function ham = hamiltonian(uk,C2,C3,D0)
 % Basic stuff.
 %% NOTE: JJ and MC should be defined from the size of uk, instead of from 
 %% a params variable, so that the code works for a single uk or an ensemble.

### 2019-09-21 code_rescaledModel/runu.m

@@ -1,17 +1,18 @@
 clear all; close all;
 % Input parameters.
-J = 8;
-MC = 5;
-Nw = 4;
+Lambda = 16;
+Nw = 8;
+MC = 1e3;
 theta = 10;
-gibd = 0.24;
-fi = 1;
+gibd = 1.;
+fi = 0;
 % time step parameters
 dt = 5E-4;
 nout = 1;
 tfin = dt;
 
 % Run the simulation.
+J = 2*Lambda;
 SolverKdV_SymplecticM4a_MC(J, tfin, dt, nout, MC, 0, Nw, theta, gibd, fi)
 
 % Load the output file.

### 2019-09-21 paper2_JNLS/paper.tex
N
NM

@@
 
+% More careful development of theory and experiments, *in conjunction* with one another.
+
 \vsp{4}
@@
 The corresponding histograms are shown in Figs.~\ref{fig2}(c)--(d).
+
+TO DO: Describe the `representative' experiments in which $\dratdn = 0.24$ and $\epsup = 0.017$.
+% These use
 
@@
 &\mean{Q} = \int_{\Omega}  Q \, d \Gibbs
-&&\mbox{\em ensemble average} \\
+&&\mbox{\em Ensemble average} \\
 &\tavg{Q} = \frac{1}{T_2-T_1} \int_{T_1}^{T_2} Q \, dt	 
-&&\mbox{\em time average}
+&&\mbox{\em Time average}
 \end{align}
@@
 \begin{align}
-&u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+} \qquad \text{for } x \in [-\pi, \pi]
-%&&\mbox{\em deterministic matching condition}
+u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+},
+\qquad \mbox{\em Deterministic matching condition}
 \end{align}
+holding for $x \in [-\pi, \pi]$.
 %or in Johnson's formulation, it is $\depth^{1/4} \eta$ that matches.
@@
 \label{statmatch}
-&\meanup{\Hdn} = \meandn{\Hdn}
-%&&\mbox{\em statistical matching condition}
+\meanup{\Hdn} = \meandn{\Hdn}
+\qquad \mbox{\em Statistical matching condition}
 \end{align}
@@
 %----------------------------------------------------------------------------%
+% Results
+%----------------------------------------------------------------------------%
+\section{Results}
+
+With the experimental setup described and the theory outlined, we now present results comparing the two.
+Unless stated otherwise, all parameters used in the theory are taken directly from their experimental values listed in Table (ref).
+In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.17$.
+
+% Later: Plot of transfer function for different lambdas. Shows that the transfer function is relatively insensitive to truncation index, as long as N scales appropriately.
+
+\vsp{5}
+%----------------------------------------------------------------------------%
 % Skewness formula
@@
 \subsection{Explicit formula for outgoing skewness}
+We now examine an explicit formula for the outgoing skewness as originally founded in \cite{majda2019statistical}.
 An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$.
@@
 
+\vsp{10}
 %----------------------------------------------------------%
@@
 \section{Sampling Algorithms}
-
-\subsection{Naive acceptance-rejection algorithm}
-
-\subsection{Markov-chain Monte Carlo}
-
-\subsection{Improved acceptance-rejection algorithm}
+%\subsection{Naive acceptance-rejection algorithm}
+%\subsection{Markov-chain Monte Carlo}
+%\subsection{Improved acceptance-rejection algorithm}
 
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-\subsection{Comparison of basic features}
-
-\subsection{New experimental measurements guided by theory}
+ %New experimental measurements guided by theory
 

### 2019-09-21 code_rescaledModel/SolverKdV_SymplecticM4a_MC.m

@@ -27,7 +27,7 @@ w1=(2+2^(1/3)+2^(-1/3))/3;
 w2=1-2*w1;     % integration fractional time steps
 
 % Model parameters from experiments
-epsi0 = 0.016;  % amplitude-to-depth ratio
+epsi0 = 0.017;  % amplitude-to-depth ratio
 del0 = 0.22;    % depth-to-wavelength ratio
 Dm = 1;  Dp = 0.24;  % depth ratio
 if fi == 0     % fi == 0 for upstream flow

### 2019-09-21 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\tavg}[1]{\overline{#1}}
+\newcommand{\transf}{\mathcal{F}}
 
@@
 Briefly, consider the parameter $\lamfac$, the number of characteristic wavelengths in the physical domain.
-We require $\lamfac \le \Lambda$, so that the mode $\uhat_{\lamfac}$ corresponding to the physical wavelength $\lambda$ is resolved.
-In the case $\lamfac = \Lambda$, $\lambda$ corresponds to the smallest resolved wavelength.
-Often, we will choose $\lamfac = \Lambda$ or $\Lambda/2$.
+We require $\lamfac \le \Lambda$, so that the mode $\uhat_{\lamfac}$, corresponding to the characteristic wavelength $\lam$ in the experiments, is resolved in the truncated dynamical system.
+If $\lamfac = \Lambda$, then $\lam$ corresponds to the smallest resolved wavelength.
+Often, we will choose an intermediate value such as $\lamfac = \Lambda/2$ or $\Lambda/4$, so that the truncated system \eqref{TKdV} resolves scales that are both bigger and smaller than the characteristic value $\lam$.
 
@@
 %----------------------------------------------------------------------------%
-\subsection{Matching at the ADC}
+\subsection{Matching conditions and the transfer function}
 
@@
 \end{align}
-holding for $x \in [-\pi, \pi]$.
+holding for $x \in [-\pi, \pi]$. This matching condition is used for the dynamical system \eqref{TKdV}.
 %or in Johnson's formulation, it is $\depth^{1/4} \eta$ that matches.
@@
 
+The statistical matching condition imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
+In particular, we view $\thup$ as given by the random-state of the incoming wave field.
+Thus, while we will treat $\thup$ as a free parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by matching statistics at the ADC.
+In this way, \eqref{statmatch} implies the functional dependence
+\begin{equation}
+\thdn = \transf \left( \thup \right)
+\end{equation}
+The {\em transfer function} $\transf$ will be a key link needed to relate the theory to experiments.
+
 %----------------------------------------------------------------------------%
@@
 With the experimental setup described and the theory outlined, we now present results comparing the two.
-Unless stated otherwise, all parameters used in the theory are taken directly from their experimental values listed in Table (ref).
+Unless stated otherwise, all parameters used in the theory are taken directly from their experimental values listed in Table \ref{paramtable}.
 In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.17$.
+
+\subsection{Calibration of the inverse temperature}
+
+All parameters that appear in the theoretical model follow directly from experimental parameters, with the exception of the inverse temperature, $\thup$, of the incoming flow and the parameter $\lamfac$, which represents the number of characteristic wavelengths in the periodic domain.
+The parameter $\lamfac$ can be chosen in a sensible way by simply selecting an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved.
+Hence, we will simply chose $\lamfac = \Lambda/2$ throughout, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths.
+The incoming inverse temperature, $\thup$, however, is more difficult to pin down without using information from the experiments.
+Our strategy is to use the outgoing skewness as the main diagnostic to determine a realistic range for $\thup$.
+That is, for an input $\thup$, the downstream inverse temperature $\thdn$ is determined by \eqref{statmatch}, which ultimately sets the skewness of the outgoing wave-field.
 

### 2019-09-22 paper2_JNLS/paper.tex
NM

@@
 
-% Later: Plot of transfer function for different lambdas. Shows that the transfer function is relatively insensitive to truncation index, as long as N scales appropriately.
+% Figure
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.99 \linewidth]{Figs/transfig.pdf}
+\caption{
+stuff
+}
+\label{transfig}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+Figure \ref{transfig}(a) shows the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $\thup \in [0,20]$.
+In this figure, all model parameters are set by their experimental values, and we show the cases $\Lambda = 8, 16, 32$ with $\lamfac = \Lambda/2$ in each.
+As seen here, the transfer function is relatively insensitive to the truncation index since all three curves are relatively close to one another.
+This insensitivity is a consequence of scaling $\lamfac$ with $\Lambda$ appropriately as discussed previously.
+
+Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depends on the incoming inverse temperature $\thup$.
+In line with experimental observations, the incoming skewness is small (always less than 0.2), while the outgoing wave skewness is much higher.
+Specifically, to capture the experimentally observed peak skewness of about 0.7, we can choose the value of $\thup$ in the range 10--15.
 

### 2019-09-22 code_rescaledModel/SolverKdV_SymplecticM4a_MC.m
NM

@@ -107,6 +107,8 @@ for ii=1:Nt
 
         % Nick's diagnostics
         uarray(:,:,ii/countDiag) = u;
+        duarray(:,:,ii/countDiag) = du;
+
         
         if mod(ii,1e3)==0
             display(['iteration i = ', num2str(ii),'; E = ',num2str(energy(ii/countDiag)),', H = ',num2str(mean(hamilt(:,ii/countDiag)))]);
@@ -115,7 +117,7 @@ for ii=1:Nt
         
         % SAVE DATA TO OUTPUT FILE
         %save('output.mat', 'T','meanu','covau','skewu','kurtu', 'skewdu','kurtdu', 'energy','hamiltonian')
-        save('output.mat', 'uarray', 'T')
+        save('output.mat', 'uarray', 'duarray', 'T')
     end
     
     

### 2019-09-22 code_rescaledModel/runu.m

@@ -4,8 +4,8 @@ Lambda = 16;
 Nw = 8;
 MC = 1e3;
 theta = 10;
-gibd = 1.;
-fi = 0;
+gibd = 0.24;
+fi = 1;
 % time step parameters
 dt = 5E-4;
 nout = 1;
@@ -22,13 +22,15 @@ MC = size(uarray,2);
 Ntsteps = size(uarray,3);
 % Flatten the array of u into a single list.
 ulist = reshape(uarray,1,[]);
-uarray(:,:,1);
+dulist = reshape(duarray,1,[]);
 % Write output to a text file.
 fileID = fopen('ulist.txt','w');
 fprintf(fileID,'# Number of grid points, trajectories, and time steps.\n');
 fprintf(fileID,'%d\n',[J,MC,Ntsteps]);
 fprintf(fileID,'# Values of u\n');
-fprintf(fileID,'%12.8f\n',ulist);
+fprintf(fileID,'%12.8f\n', ulist);
+fprintf(fileID,'# Values of du\n');
+fprintf(fileID,'%12.8f\n', dulist);
 fclose(fileID);
 
 % Make a plot to see the waves.
@@ -36,5 +38,6 @@ fclose(fileID);
 %contourf(T,x,uarray(:,1,:));
 
 % Make a histogram of u.
-histogram(ulist);
+figure(1); histogram(ulist);
+figure(2); histogram(ulist); set(gca,'yscale','log');
 skewu = skewness(ulist)

### 2019-09-23 code_rescaledModel/runu.m
NM

@@ -3,17 +3,18 @@ clear all; close all;
 Lambda = 16;
 Nw = 8;
 MC = 1e3;
-theta = 10;
+theta = 13;
 gibd = 0.24;
 fi = 1;
 % time step parameters
 dt = 5E-4;
-nout = 1;
-tfin = dt;
+nout = 100; % Default 100?
+tfin = 10;
 
 % Run the simulation.
-J = 2*Lambda;
+J = 2*Lambda; tic;
 SolverKdV_SymplecticM4a_MC(J, tfin, dt, nout, MC, 0, Nw, theta, gibd, fi)
+cputime = toc
 
 % Load the output file.
 load('output.mat')
@@ -25,12 +26,13 @@ ulist = reshape(uarray,1,[]);
 dulist = reshape(duarray,1,[]);
 % Write output to a text file.
 fileID = fopen('ulist.txt','w');
-fprintf(fileID,'# Number of grid points, trajectories, and time steps.\n');
-fprintf(fileID,'%d\n',[J,MC,Ntsteps]);
+fprintf(fileID,'# Number of grid points, trajectories, time steps, ');
+fprintf(fileID,'value of Nw, theta, tfin.\n');
+fprintf(fileID,'%d\n',[J,MC,Ntsteps,Nw,theta,tfin]);
 fprintf(fileID,'# Values of u\n');
-fprintf(fileID,'%12.8f\n', ulist);
+fprintf(fileID,'%9.5f\n', ulist);
 fprintf(fileID,'# Values of du\n');
-fprintf(fileID,'%12.8f\n', dulist);
+fprintf(fileID,'%9.5f\n', dulist);
 fclose(fileID);
 
 % Make a plot to see the waves.

### 2019-09-23 paper2_JNLS/paper.tex
NM

@@
 
+% Figure
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.99 \linewidth]{Figs/uhist_expsim.pdf}
+\caption{
+Amazing!
+}
+\label{hist_expsim}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+% As of now the simulations use thup=15, thdn=13.
+
+We show histograms of the surface slope in Fig.~\ref{slopehist}, from both the experiments (top) and theory (bottom).
+The derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.
+Therefore, in Fig.~\ref{slopehist}, we convert the theoretical calculated values of $\partial u/\partial \tilde{x}$ to physical slope via
+\begin{equation}
+\pd{\eta}{x} = \frac{\ampscale \lamfac}{\lengthscale} \pd{u}{\tilde{x}}
+\end{equation}
+We have included the $\lamfac$ in the numerator due to the differences in forcing spectra.
+If the experiments were conducted with the forcing given by the incoming Gibbs measure literally, then there would be no need for $\lamfac$.
+But the experimental forcing (which was decided before the theory) has a peak frequency of $\freqp = 2 Hz$ and so the factor $\lamfac$ makes $\freqp$ correspond to the dominant value in the spectrum in the theory (REWORD).
+
+% Figure
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.99 \linewidth]{Figs/slopehist.pdf}
+\caption{
+Amazing!
+}
+\label{slopehist}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+% As of now the simulations use thup=15, thdn=13.
+
 \vsp{5}

### 2019-09-23 code_rescaledModel/runu.m
NM

@@ -3,13 +3,22 @@ clear all; close all;
 Lambda = 16;
 Nw = 8;
 MC = 1e3;
-theta = 13;
-gibd = 0.24;
-fi = 1;
+% Choose upstream or downstream
+down = false;
+if down == true
+    theta = 13;
+    gibd = 0.24;
+    fi = 1;
+else
+    theta = 15;
+    gibd = 1.;
+    fi = 0;
+end
 % time step parameters
 dt = 5E-4;
 nout = 100; % Default 100?
 tfin = 10;
+
 
 % Run the simulation.
 J = 2*Lambda; tic;

### 2019-09-23 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\SympL}{\sympJ_{\Lambda}}
+\newcommand{\uup}{u_{-}}
+\newcommand{\udn}{u_{+}}
+\newcommand{\uupdn}{u_{\pm}}
 % Gibbs and theta
@@
 \caption{
-stuff
+Transfer function
 }
@@
 
+Figure \ref{uhist} shows histograms of the displacement $u$ from experiments and from TKdV theory.
+The experimental measurements have been normalized for comparison purposes.
+The comparison is fucking amazing!
+
 % Figure
@@
 \begin{center}
-\includegraphics[width = 0.99 \linewidth]{Figs/uhist_expsim.pdf}
+\includegraphics[width = 0.99 \linewidth]{Figs/uhist.pdf}
 \caption{
@@
 }
-\label{hist_expsim}
+\label{uhist}
 \end{center}
@@
 The derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.
-Therefore, in Fig.~\ref{slopehist}, we convert the theoretical calculated values of $\partial u/\partial \tilde{x}$ to physical slope via
+Therefore, in Fig.~\ref{slopehist}, we convert the theoretical calculated values to physical slope via
 \begin{equation}
-\pd{\eta}{x} = \frac{\ampscale \lamfac}{\lengthscale} \pd{u}{\tilde{x}}
+\pd{\eta}{x} = 2 \pi^{3/2} \frac{\etastd}{\lamupdn} \pd{\uupdn}{\tilde{x}}
+%\frac{\ampscale \lamfac}{\lengthscale} \pd{\uupdn}{\tilde{x}}
 \end{equation}
@@
 \caption{
-Amazing!
+Slopes
 }

### 2019-09-23 code_rescaledModel/SolverKdV_SymplecticM4a_MC.m

@@ -1,160 +1,108 @@
-function SolverKdV_SymplecticM4a_MC(J, tfin, dt, nout, MC, um, Nw, theta, gibd, fi)
+function SolverKdV_SymplecticM4a_MC(C2,C3,Drat,Lambda,...
+    MC,theta,gibd,fi,dt,nout,tfin)  
+
 % This script solves the rescaled KdV equation in a periodic domain
 % using pseudo-spectral & symplectic M4a integrator
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 % Input
-% J: Number of points in each direction, divisible by 4. reference J = 16.
-% tfin: The stopping time of the simulation.
-% dt: The numerical time step; reference dt = .5E-3;
-% nout: The number of output times
-% MC: The number of trajectories; reference MC = 1E4
-% um: The momentum of u; generally set um = 0.
-% Nw: The number of wavelengths in the domain; reference Nw = 1.
-% theta: The inverse temperature
-% gibd: The Gibbs distribution to sample the initial state; 1 or 0.24.
-% fi: Index 0 for incoming or 1 for outgoing.
+% C2 and C3: Constants in Hamiltonian.
+% Drat: Depth ratio; reference Drat = 0.24.
+% Lambda: Number of modes; reference Lambda = 16.
+% MC: The number of trajectories; reference MC = 1E4.
+% theta: The inverse temperature; range 10-20.
+% gibd: The Gibbs distribution to sample the initial state.
+% fi: Flow index; fi = 0 for incoming or 1 for outgoing.
+% dt: The numerical time step; reference dt = .5E-3.
+% nout: The number of output times; reference nout = 100.
+% tfin: The stopping time of the simulation; reference tfin = 10.
 
 % Set simulation parameters
-Nt = round(tfin/dt);        % Number of time steps
-countDiag = round(Nt/nout)  % Compute diagnostics every countDiag steps
-
-
+JJ = 2*Lambda;          % Number of grid points.
[…]

### 2019-09-23 code_rescaledModel/iterativeKdV_M4a_MC.m

@@ -1,10 +1,10 @@
 
-function [zk,ik]=iterativeKdV_M4a_MC(zk0, tau,yk,p, tol)
+function [zk,ik]=iterativeKdV_M4a_MC(zk0, tau,yk,params, tol)
 % Function iterates to solve the solution for Z-Y=tau/2*F(Z)
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-C2=p.C2; C3=p.C3;
-D0=p.D0;
-J = p.J; MC = p.MC;
+C2=params.C2; C3=params.C3;
+D0=params.D0;
+J = params.J; MC = params.MC;
 
 k = repmat([0:J/2 -J/2+1:-1]',[1 MC]);
 Lk1 = C2*D0^(1/2)*1i*k.^3;

### 2019-09-23 code_rescaledModel/runu.m

@@ -1,10 +1,20 @@
 clear all; close all;
-% Input parameters.
+
+% Model parameters from experiments
+epsi0 = 0.017;  % amplitude-to-depth ratio
+del0 = 0.22;    % depth-to-wavelength ratio
+Drat = 0.24;    % depth ratio.
+% Simulation parameters
 Lambda = 16;
 Nw = 8;
-MC = 1e3;
+MC = 1e1;
+% time step parameters
+dt = 5E-4;
+nout = 10; % Default 100?
+tfin = 1;
+
 % Choose upstream or downstream
-down = false;
+down = true;
 if down == true
     theta = 13;
     gibd = 0.24;
@@ -14,20 +24,17 @@ else
     gibd = 1.;
     fi = 0;
 end
-% time step parameters
-dt = 5E-4;
-nout = 100; % Default 100?
-tfin = 10;
-
 
 % Run the simulation.
-J = 2*Lambda; tic;
-SolverKdV_SymplecticM4a_MC(J, tfin, dt, nout, MC, 0, Nw, theta, gibd, fi)
-cputime = toc
+tic;
+C2 = (2/3)*pi^2*del0/Nw^2
+C3 = 3/2*pi^(1/2)*epsi0/del0
+SolverKdV_SymplecticM4a_MC(C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin);
+cputime = round(toc/60, 2, 'significant')
 
 % Load the output file.
 load('output.mat')
-J = size(uarray,1);
+JJ = size(uarray,1);
 MC = size(uarray,2);
 Ntsteps = size(uarray,3);
 % Flatten the array of u into a single list.
@@ -35,13 +42,15 @@ ulist = reshape(uarray,1,[]);
 dulist = reshape(duarray,1,[]);
 % Write output to a text file.
 fileID = fopen('ulist.txt','w');
-fprintf(fileID,'# Number of grid points, trajectories, time steps, ');
-fprintf(fileID,'value of Nw, theta, tfin.\n');
-fprintf(fileID,'%d\n',[J,MC,Ntsteps,Nw,theta,tfin]);
-fprintf(fileID,'# Values of u\n');
-fprintf(fileID,'%9.5f\n', ulist);
-fprintf(fileID,'# Values of du\n');
-fprintf(fileID,'%9.5f\n', dulist);
+fprintf(fileID,'# Input parameters: ');
+
[…]

### 2019-09-23 code_rescaledModel/sampling_MH_phy.m

@@ -1,8 +1,8 @@
-function [samp_u,sampk_u, enek,samp_H] = sampling_MH_phy(J, MC, theta,p)
+function [samp_u,sampk_u, enek,samp_H] = sampling_MH_phy(J, MC, theta,params)
 % Sampling from microcanonical-canonical ensemble with constant energy
 % using a Metropolis-Hastings MC approach
-C2=p.C2; C3=p.C3;
-D0=p.gibd;
+C2=params.C2; C3=params.C3;
+D0=params.gibd;
 
 kk = [0:J/2 -J/2+1:-1]';
 Dk = 1i*kk; % wavenumbers
@@ -24,9 +24,6 @@ uk = sqrt(Enorm/E)*uk;
 
 % previous distribution
 H_pre = hamiltonian(uk,C2,C3,D0);
-% OLD CODE
-%H_pre = -( 1/6*D0^(-1)*C3*real(sum(u2k_dealiasing(uk).*conj(uk))) ...
-%          -1/2*D0^(1)*C2*sum(abs(Dk.*uk).^2) ) *2*pi/J^2;
       
 count_acpt = zeros(N_iter,1);
 % run Markov chain p(x|y)~exp(-.5*|x-y|^2)
@@ -46,9 +43,6 @@ for ii = 1:N_iter
         
         % new distribution
         H_new = hamiltonian(uk_tilde,C2,C3,D0);
-        % OLD CODE
-        %H_new = -( 1/6*D0^(-1)*C3*real(sum(u2k_dealiasing(uk_tilde).*conj(uk_tilde))) ...
-        %          -1/2*D0^(1)*C2*sum(abs(Dk.*uk_tilde).^2) ) *2*pi/J^2;
         
         alpha = exp(-theta*(H_new-H_pre));
         r = rand(1);

### 2019-09-23 code_rescaledModel/runu.m
NM

@@ -5,13 +5,13 @@ epsi0 = 0.017;  % amplitude-to-depth ratio
 del0 = 0.22;    % depth-to-wavelength ratio
 Drat = 0.24;    % depth ratio.
 % Simulation parameters
-Lambda = 16;
-Nw = 8;
-MC = 1e1;
+Lambda = 16;    % Reference 16
+Nw = Lambda/2;  % Reference Lambda/2
+MC = 1E3;       % Reference 1E4
 % time step parameters
-dt = 5E-4;
-nout = 10; % Default 100?
-tfin = 1;
+dt = 5E-4;  % Reference 5E-4
+nout = 100; % Reference 100
+tfin = 10.; % Reference 10
 
 % Choose upstream or downstream
 down = true;
@@ -42,13 +42,17 @@ ulist = reshape(uarray,1,[]);
 dulist = reshape(duarray,1,[]);
 % Write output to a text file.
 fileID = fopen('ulist.txt','w');
-fprintf(fileID,'# Input parameters: ');
-
-%fprintf(fileID,'value of Nw, theta, tfin.\n');
-
-fprintf(fileID,'%d\n',[]);
-%fprintf(fileID,'%d\n',[JJ,MC,Ntsteps,Nw,theta,tfin]);
-
+% Input parameters
+fprintf(fileID,'#Basic integers: JJ, MC, Nt, Nw.\n');
+out1 = [JJ,MC,nout,Nw]; fprintf(fileID,'%d\n', out1);
+fprintf(fileID,'#Basic floats: theta, tfin, cputime (mins).\n');
+out2 = [theta, tfin, cputime]; fprintf(fileID,'%9.5f\n', out2);
+fprintf(fileID,'# All other inputs: C2,C3,Drat,gibd,fi,dt.\n');
+out3 = [C2,C3,Drat,gibd,fi,dt]; fprintf(fileID,'%9.5f\n', out3);
+fprintf(fileID,'# Pad with zeros until index 20.\n');
+len = 20 - length(out1)-length(out2)-length(out3);
+fprintf(fileID,'%d\n', zeros(len,1) );
+% Microstates
 fprintf(fileID,'# Values of u\n'); fprintf(fileID,'%9.5f\n', ulist);
 fprintf(fileID,'# Values of du\n'); fprintf(fileID,'%9.5f\n', dulist);
 fclose(fileID);

### 2019-09-23 code_rescaledModel/runu.m
NM

@@ -7,14 +7,14 @@ Drat = 0.24;    % depth ratio.
 % Simulation parameters
 Lambda = 16;    % Reference 16
 Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E3;       % Reference 1E4
+MC = 1E2;       % Reference 1E4
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
 nout = 100; % Reference 100
-tfin = 10.; % Reference 10
+tfin = 5.; % Reference 10
 
 % Choose upstream or downstream
-down = true;
+down = false;
 if down == true
     theta = 13;
     gibd = 0.24;

### 2019-09-23 code_rescaledModel/runu.m
NM

@@ -1,67 +1,0 @@
-clear all; close all;
-
-% Model parameters from experiments
-epsi0 = 0.017;  % amplitude-to-depth ratio
-del0 = 0.22;    % depth-to-wavelength ratio
-Drat = 0.24;    % depth ratio.
-% Simulation parameters
-Lambda = 16;    % Reference 16
-Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E2;       % Reference 1E4
-% time step parameters
-dt = 5E-4;  % Reference 5E-4
-nout = 100; % Reference 100
-tfin = 5.; % Reference 10
-
-% Choose upstream or downstream
-down = false;
-if down == true
-    theta = 13;
-    gibd = 0.24;
-    fi = 1;
-else
-    theta = 15;
-    gibd = 1.;
-    fi = 0;
-end
-
-% Run the simulation.
-tic;
-C2 = (2/3)*pi^2*del0/Nw^2
-C3 = 3/2*pi^(1/2)*epsi0/del0
-SolverKdV_SymplecticM4a_MC(C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin);
-cputime = round(toc/60, 2, 'significant')
-
-% Load the output file.
-load('output.mat')
-JJ = size(uarray,1);
-MC = size(uarray,2);
-Ntsteps = size(uarray,3);
-% Flatten the array of u into a single list.
-ulist = reshape(uarray,1,[]);
-dulist = reshape(duarray,1,[]);
-% Write output to a text file.
-fileID = fopen('ulist.txt','w');
-% Input parameters
-fprintf(fileID,'#Basic integers: JJ, MC, Nt, Nw.\n');
-out1 = [JJ,MC,nout,Nw]; fprintf(fileID,'%d\n', out1);
-fprintf(fileID,'#Basic floats: theta, tfin, cputime (mins).\n');
-out2 = [theta, tfin, cputime]; fprintf(fileID,'%9.5f\n', out2);
-fprintf(fileID,'# All other inputs: C2,C3,Drat,gibd,fi,dt.\n');
-out3 = [C2,C3,Drat,gibd,fi,dt]; fprintf(fileID,'%9.5f\n', out3);
-fprintf(fileID,'# Pad with zeros until index 20.\n');
-len = 20 - length(out1)-length(out2)-length(out3);
-fprintf(fileID,'%d\n', zeros(len,1) );
-% Microstates
[…]

### 2019-09-23 code_rescaledModel/runSolver.m
NM

@@ -1,0 +1,67 @@
+clear all; close all;
+
+% Model parameters from experiments
+epsi0 = 0.017;  % amplitude-to-depth ratio
+del0 = 0.22;    % depth-to-wavelength ratio
+Drat = 0.24;    % depth ratio.
+% Simulation parameters
+Lambda = 16;    % Reference 16
+Nw = Lambda/2;  % Reference Lambda/2
+MC = 1E2;       % Reference 1E4
+% time step parameters
+dt = 5E-4;  % Reference 5E-4
+nout = 100; % Reference 100
+tfin = 5.; % Reference 10
+
+% Choose upstream or downstream
+down = false;
+if down == true
+    theta = 13;
+    gibd = 0.24;
+    fi = 1;
+else
+    theta = 15;
+    gibd = 1.;
+    fi = 0;
+end
+
+% Run the simulation.
+tic;
+C2 = (2/3)*pi^2*del0/Nw^2
+C3 = 3/2*pi^(1/2)*epsi0/del0
+SolverKdV_SymplecticM4a_MC(C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin);
+cputime = round(toc/60, 2, 'significant')
+
+% Load the output file.
+load('output.mat')
+JJ = size(uarray,1);
+MC = size(uarray,2);
+Ntsteps = size(uarray,3);
+% Flatten the array of u into a single list.
+ulist = reshape(uarray,1,[]);
+dulist = reshape(duarray,1,[]);
+% Write output to a text file.
+fileID = fopen('ulist.txt','w');
+% Input parameters
+fprintf(fileID,'#Basic integers: JJ, MC, Nt, Nw.\n');
+out1 = [JJ,MC,nout,Nw]; fprintf(fileID,'%d\n', out1);
+fprintf(fileID,'#Basic floats: theta, tfin, cputime (mins).\n');
+out2 = [theta, tfin, cputime]; fprintf(fileID,'%9.5f\n', out2);
+fprintf(fileID,'# All other inputs: C2,C3,Drat,gibd,fi,dt.\n');
+out3 = [C2,C3,Drat,gibd,fi,dt]; fprintf(fileID,'%9.5f\n', out3);
+fprintf(fileID,'# Pad with zeros until index 20.\n');
+len = 20 - length(out1)-length(out2)-length(out3);
+fprintf(fileID,'%d\n', zeros(len,1) );
+% Microstates
[…]

### 2019-09-24 code_rescaledModel/SolverKdV_SymplecticM4a_MC.m
NM

@@ -1,5 +1,5 @@
-function SolverKdV_SymplecticM4a_MC(C2,C3,Drat,Lambda,...
-    MC,theta,gibd,fi,dt,nout,tfin)  
+function [uarray,duarray] = SolverKdV_SymplecticM4a_MC(...
+    C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin)  
 
 % This script solves the rescaled KdV equation in a periodic domain
 % using pseudo-spectral & symplectic M4a integrator
@@ -47,8 +47,8 @@ ukm2 = uk0;  % at time n-2
 interp = @(u0,um1,um2,w) ...
     (.5*w*(w+3)+1).*u0 - w*(w+2).*um1 +.5*w*(w+1).*um2;
 % Main variables
-tsave = zeros(1,Nt/MM);
 uu = ifft(uk);
+%tout = zeros(1,Nt/MM);
 uarray = zeros(JJ,MC,Nt/MM);
 duarray = zeros(JJ,MC,Nt/MM);
 
@@ -59,7 +59,7 @@ for ii=1:Nt
         iout = ii/MM;
         du=real(ifft(1i*kvec.*uk));
         % Save the microstate, u and du.
-        tsave(iout)=tt;
+        %tout(iout)=tt;
         uarray(:,:,iout) = uu;
         duarray(:,:,iout) = du;
         % Check energy and hamiltonian.        
@@ -101,8 +101,3 @@ for ii=1:Nt
     uk = fft(uu); uk(JJ/2+1,:)=0;
     if any(abs(uu(:))>ulim),break,end
 end
-
-% Save output
-save('output.mat', 'C2','C3','Drat','Lambda',...
-    'MC','theta','gibd','fi','dt','nout','tfin',...
-    'uarray', 'duarray', 'tsave');

### 2019-09-24 code_rescaledModel/makeOutput.m

@@ -1,50 +1,22 @@
 clear all; close all;
 
-% Model parameters from experiments
-epsi0 = 0.017;  % amplitude-to-depth ratio
-del0 = 0.22;    % depth-to-wavelength ratio
-Drat = 0.24;    % depth ratio.
-% Simulation parameters
-Lambda = 16;    % Reference 16
-Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E2;       % Reference 1E4
-% time step parameters
-dt = 5E-4;  % Reference 5E-4
-nout = 100; % Reference 100
-tfin = 5.; % Reference 10
-
-% Choose upstream or downstream
-down = false;
-if down == true
-    theta = 13;
-    gibd = 0.24;
-    fi = 1;
-else
-    theta = 15;
-    gibd = 1.;
-    fi = 0;
-end
-
-% Run the simulation.
-tic;
-C2 = (2/3)*pi^2*del0/Nw^2
-C3 = 3/2*pi^(1/2)*epsi0/del0
-SolverKdV_SymplecticM4a_MC(C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin);
-cputime = round(toc/60, 2, 'significant')
-
 % Load the output file.
 load('output.mat')
-JJ = size(uarray,1);
-MC = size(uarray,2);
-Ntsteps = size(uarray,3);
+% Basic parameters.
+JJ = size(uarray,1); 
+assert(JJ == 2*Lambda);
+assert(MC == size(uarray,2));
+Ntout = size(uarray,3);
+
 % Flatten the array of u into a single list.
 ulist = reshape(uarray,1,[]);
 dulist = reshape(duarray,1,[]);
+
 % Write output to a text file.
 fileID = fopen('ulist.txt','w');
 % Input parameters
-fprintf(fileID,'#Basic integers: JJ, MC, Nt, Nw.\n');
-out1 = [JJ,MC,nout,Nw]; fprintf(fileID,'%d\n', out1);
+fprintf(fileID,'#Basic integers: JJ, MC, Ntout, Nw.\n');
+out1 = [JJ,MC,Ntout,Nw]; fprintf(fileID,'%d\n', out1);
 fprintf(fileID,'#Basic floats: theta, tfin, cputime (mins).\n');
 out2 = [theta, tfin, cputime]; fprintf(fileID,'%9.5f\n', out2);
[…]

### 2019-09-24 code_rescaledModel/runSolver.m

@@ -7,11 +7,11 @@ Drat = 0.24;    % depth ratio.
 % Simulation parameters
 Lambda = 16;    % Reference 16
 Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E2;       % Reference 1E4
+MC = 1E1;       % Reference 1E4
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
-nout = 100; % Reference 100
-tfin = 5.; % Reference 10
+nout = 4; % Reference 100
+tfin = 0.1; % Reference 10
 
 % Choose upstream or downstream
 down = false;
@@ -27,41 +27,22 @@ end
 
 % Run the simulation.
 tic;
-C2 = (2/3)*pi^2*del0/Nw^2
-C3 = 3/2*pi^(1/2)*epsi0/del0
-SolverKdV_SymplecticM4a_MC(C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin);
+C2 = (2/3)*pi^2*del0/Nw^2;
+C3 = 3/2*pi^(1/2)*epsi0/del0;
+[uarray, duarray] = SolverKdV_SymplecticM4a_MC(...
+    C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin);
+
+% Save output
 cputime = round(toc/60, 2, 'significant')
-
-% Load the output file.
-load('output.mat')
-JJ = size(uarray,1);
-MC = size(uarray,2);
-Ntsteps = size(uarray,3);
-% Flatten the array of u into a single list.
-ulist = reshape(uarray,1,[]);
-dulist = reshape(duarray,1,[]);
-% Write output to a text file.
-fileID = fopen('ulist.txt','w');
-% Input parameters
-fprintf(fileID,'#Basic integers: JJ, MC, Nt, Nw.\n');
-out1 = [JJ,MC,nout,Nw]; fprintf(fileID,'%d\n', out1);
-fprintf(fileID,'#Basic floats: theta, tfin, cputime (mins).\n');
-out2 = [theta, tfin, cputime]; fprintf(fileID,'%9.5f\n', out2);
-fprintf(fileID,'# All other inputs: C2,C3,Drat,gibd,fi,dt.\n');
-out3 = [C2,C3,Drat,gibd,fi,dt]; fprintf(fileID,'%9.5f\n', out3);
-fprintf(fileID,'# Pad with zeros until index 20.\n');
[…]

### 2019-09-24 code_rescaledModel/runSolver.m
NM

@@ -7,14 +7,14 @@ Drat = 0.24;    % depth ratio.
 % Simulation parameters
 Lambda = 16;    % Reference 16
 Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E1;       % Reference 1E4
+MC = 1E3;       % Reference 1E4
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
-nout = 4; % Reference 100
-tfin = 0.1; % Reference 10
+nout = 1; % Reference 100
+tfin = dt; % Reference 10
 
 % Choose upstream or downstream
-down = false;
+down = true;
 if down == true
     theta = 13;
     gibd = 0.24;

### 2019-09-24 code_rescaledModel/SolverKdV_SymplecticM4a_MC.m
NM

@@ -48,9 +48,11 @@ interp = @(u0,um1,um2,w) ...
     (.5*w*(w+3)+1).*u0 - w*(w+2).*um1 +.5*w*(w+1).*um2;
 % Main variables
 uu = ifft(uk);
-%tout = zeros(1,Nt/MM);
-uarray = zeros(JJ,MC,Nt/MM);
-duarray = zeros(JJ,MC,Nt/MM);
+
+outsize = floor(Nt/MM);
+%tout = zeros(1,outsize);
+uarray = zeros(JJ,MC,outsize);
+duarray = zeros(JJ,MC,outsize);
 
 % Main loop 
 for ii=1:Nt

### 2019-09-24 code_rescaledModel/runSolver.m

@@ -7,11 +7,11 @@ Drat = 0.24;    % depth ratio.
 % Simulation parameters
 Lambda = 16;    % Reference 16
 Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E3;       % Reference 1E4
+MC = 1E1;       % Reference 1E4
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
-nout = 1; % Reference 100
-tfin = dt; % Reference 10
+nout = 150; % Reference 100
+tfin = 1.; % Reference 10
 
 % Choose upstream or downstream
 down = true;

### 2019-09-24 code_rescaledModel/runSolver.m
NM

@@ -14,7 +14,7 @@ nout = 100; % Reference 100
 tfin = 10.;  % Reference 10
 
 % Choose upstream or downstream
-gibup = true;
+gibup = false;
 if gibup == true
     theta = 15;
     gibd = 1.;

### 2019-09-28 paper2_JNLS/paper.tex
NM

@@
 % More careful development of theory and experiments, *in conjunction* with one another.
+
+% Definitely cite several of Sapsis's new papers.
 
@@
 
+\subsection{Statistical comparison between theory and experiments}
+
 Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depends on the incoming inverse temperature $\thup$.
@@
 % As of now the simulations use thup=15, thdn=13.
+
+\subsection{Wave dynamics}
+
+We show individual trajectories in the next figure for upstream and downstream.
+In the upstream flow, due to the coefficients, the nonlinearity is very weak.
+Hence, wave propagation is dominated by linear dispersion.
+Indeed, with weak nonlinearity, \eqref{dimlessKdV} is approximated by
+\begin{equation}
+u_t + C_2 \drat^{1/2} \, u_{xxx} = 0
+\qquad \text{for } x \in [-\pi,\pi] \\
+\end{equation}
+which implies the dispersion relation $\omega_k = -C_2 k^3$ or for phase velocity $c_k = -C_2 k^2$, which implies left-going waves \cite{majdaqi2019}.
 

### 2019-10-01 paper2_JNLS/paper.tex
NM
NNM

@@
 In these experiments, an incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
-Upon passing into the region of shallower depth, the wave distribution skews strongly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
+Upon passing into the region of shallower depth, the wave distribution skews markedly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
 Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} laid out a theoretical foundation that accurately captures several aspects of the anomalous behavior.
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
-In follow-up work, Majda and Qi (cite) performed analysis on more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous features.
+In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous features.
 
@@
 	As discussed in \cite{bolles2019anomalous}, the experiments consist of a long, narrow wave tank (6 m long x 20 cm wide x 30 cm high), with waves generated by plexiglass paddle at one end.
-The waves propagate through the tank and, roughly midway through, pass over an abrupt depth change (ADC), which is created by a plexiglass step.
+The waves propagate through the tank and, roughly midway through, pass over an abrupt depth change (ADC) created by a plexiglass step.
 The waves continue to propagate through the shallower depth until reaching the far end of the tank, at which point their energy is dissipated by a horse-hair dampener.
@@
 The comparison is fucking amazing!
+Offers convincing evidence that the theory works and the parameters have been matched as well as possible.
+(At some point, I need to mention the minor point about the forcing spectrum not being matched to the Gibbs measure because it came first).
 
@@
 which implies the dispersion relation $\omega_k = -C_2 k^3$ or for phase velocity $c_k = -C_2 k^2$, which implies left-going waves \cite{majdaqi2019}.
+% Further details, see page 6 of Majda Qi, J Stat Phys.
+
+% Figure
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.99 \linewidth]{Figs/trajectories.pdf}
+\caption{
+Trajectories
+}
+\label{trajectories}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+% Figure
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.99 \linewidth]{Figs/skew-kurt.pdf}
+\caption{
+Skewness and kurtosis in DNS}
+\label{skew-kurt}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 

### 2019-10-01 paper2_JNLS/paper.tex
NM

@@
 The waves continue to propagate through the shallower depth until reaching the far end of the tank, at which point their energy is dissipated by a horse-hair dampener.
-Since the dampener minimizes the backscatter, the waves in this experiment propagate primarily in one direction, from left to right in Fig.~\ref{fig1}.
+Since the dampener minimizes the backscatter, the waves in this experiment propagate primarily in one direction, from left to right in Fig.~\ref{ExpDiagStats}.
 
@@
 
- % Figure 1
+ % Figure: Experimental Diagram and Histograms
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
@@
 \begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/fig1.pdf}
+\includegraphics[width = 0.85 \linewidth]{Figs/ExpDiagStats.pdf}
 \caption{
-(a) Experimental schematic. Reproduced from \cite{bolles2019anomalous}
+(a) Experimental schematic. Adapted from \cite{bolles2019anomalous}.
+(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms.
 }
 %(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms.
-\label{fig1}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-	The free surface is illuminated by light-emitting diodes running along the bottom of the tank and then imaged from the sideview with a Nikon D3300 at 60 frames per second.
-The illumination technique coupled with high pixel count of the camera allows surface displacements to be resolved with a accuracy better than 1/3 millimeter.
-Representative free surface measurements are shown in Figs.~\ref{fig2}(a)--(b) at a position upstream and downstream of the ADC.
-The corresponding histograms are shown in Figs.~\ref{fig2}(c)--(d).
-
-TO DO: Describe the `representative' experiments in which $\dratdn = 0.24$ and $\epsup = 0.017$.
-% These use
-
- % Figure 2
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[!ht]
-\begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/fig2.pdf}
-\caption{
-(a)--(b) Surface displacement measured at a representative upstream and downstream location. (c)--(d) Corresponding histograms. Reproduced from \cite{bolles2019anomalous}
-}
-\label{fig2}
+\label{ExpDiagStats}
 \end{center}
@@
 
-  % Figure 3
+	The free surface is illuminated by light-emitting diodes running along the bottom of the tank and then imaged from the sideview with a Nikon D3300 at 60 frames per second.
+The illumination technique coupled with high pixel count of the camera allows surface displacements to be resolved with a accuracy better than 1/3 millimeter.
+Representative free surface measurements are shown in Figs.~\ref{ExpDiagStats}(b)--(c) at a position upstream and downstream of the ADC.
+The corresponding histograms are shown in Figs.~\ref{ExpDiagStats}(d)--(e).
+
+TO DO: Describe the `representative' experiments in which $\dratdn = 0.24$ and $\epsup = 0.017$.
+
+ % Figure: Experimental Spatial Statistics
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
@@
 \begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/fig3.pdf}
-\caption{
-(b)--(c) Surface slope measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms. NEW PLOT.
-}
-\label{fig3}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-% Figure 4
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[!ht]
-\begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/fig4.pdf}
-\caption{
-Spatial statistics from the experiments.
-}
-\label{fig4}
+\includegraphics[width = 0.85 \linewidth]{Figs/ExpSpatialStats.pdf}
+\caption{Spatial moments}
+\label{ExpSpatialStats}
 \end{center}
@@
 \begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/SkewAmp.pdf}
+\includegraphics[width = 0.99 \linewidth]{Figs/SkewRat.pdf}
 \caption{

### 2019-10-01 paper2_JNLS/paper.tex
NM
NNM

@@
 Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
-Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir.
+Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and the Benjamin-Feir deep-water modulational instability.
 The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.
@@
 Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field upon encountering an abrupt depth change (ADC).
-In these experiments, an incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
+In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
 Upon passing into the region of shallower depth, the wave distribution skews markedly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
-Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} laid out a theoretical foundation that accurately captures several aspects of the anomalous behavior.
+Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical foundation that accurately captures several aspects of the anomalous behavior.
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
-In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous features.
+In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous properties while enjoying a more tractable structure.
 
-The purpose of the present manuscript is to expand on both the experimental findings of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical}, with particular emphasis on the synergy between the two; e.g.~experimental discoveries guiding theoretical developments and theoretical predictions inspiring new experimental tests.
-We present a range of previously unpublished experimental data, including statistics of the surface {\em slopes} (as opposed to only displacement data as published in \cite{bolles2019anomalous}, autocorrelation data, and some data on experiments in which the depth ratio was varied.
-We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.
+The purpose of the present manuscript is to expand upon the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical, majdaqi2019}, with special emphasis placed on the synergy between experiments and theory.
+We present a range of previously unpublished experimental data which motivates further refinement of the theory.
+Likewise, new theoretical predictions have motivated additional experimental measurements, and we present a comparison between the two.
+This close cooperation between theory and experiments has proven a successful strategy \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}
 
-%~experimental discoveries guiding model developments, as well as new theoretical predictions inspiring new experimental tests.
+% New data: slopes, autocorrelation, depth ratio varied.
+% We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.
 
-%IDEA: Could comment on synergy, or close cooperation, between theory and experiments and the success thereof, then cite Ristroph erosion paper.
-
-% combination
-% draw close comparison
-% with an even closer comparison between theoretical predictions and experimental measurements.
+% combination; draw close comparison; with an even closer comparison between theoretical predictions and experimental measurements.
 
@@
 \vsp{4}
-
-% ideal, paradigm, canonical, tractable
 
@@
 \begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/ExpSpatialStats.pdf}
+\includegraphics[width = 0.9 \linewidth]{Figs/ExpSpatialStats.pdf}
 \caption{Spatial moments}

### 2019-10-01 paper2_JNLS/paper.tex
NM

@@
 
-The purpose of the present manuscript is to expand upon the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical, majdaqi2019}, with special emphasis placed on the synergy between experiments and theory.
-We present a range of previously unpublished experimental data which motivates further refinement of the theory.
-Likewise, new theoretical predictions have motivated additional experimental measurements, and we present a comparison between the two.
-This close cooperation between theory and experiments has proven a successful strategy \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}
-
-% New data: slopes, autocorrelation, depth ratio varied.
-% We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.
-
-% combination; draw close comparison; with an even closer comparison between theoretical predictions and experimental measurements.
-
-% More careful development of theory and experiments, *in conjunction* with one another.
-
-% Definitely cite several of Sapsis's new papers.
-
-\vsp{4}
+The purpose of the present manuscript is to expand upon the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical, majdaqi2019}, with special emphasis on the synergy between experiments and theory.
+We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further refinement of the theory.
+Likewise, novel theoretical predictions have motivated additional experimental measurements, and we present a comparison between the two.
+This close cooperation between theory and experiments has proven a successful strategy \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
 
@@
 In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in \cite{majda2019statistical} for the sake of brevity.
-%more transparent connection to experimental parameters
-\item Provide additional experimental measurements not presented in \cite{bolles2019anomalous}. For example, statistics on surface slope and autocorrelation.
+\item Misc: We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.
+We give a more transparent connection to experimental parameters. More careful development of theory and experiments, *in conjunction* with one another.
+\item Misc: combination; draw close comparison; with an even closer comparison between theoretical predictions and experimental measurements.
 \end{itemize}
+
+% New data: slopes, autocorrelation, depth ratio varied.
+% Definitely cite several of Sapsis's new papers.
 

### 2019-10-01 paper2_JNLS/paper.tex
NM

@@
 
-Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field upon encountering an abrupt depth change (ADC).
+Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC).
 In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
@@
 
-The purpose of the present manuscript is to expand upon the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical, majdaqi2019}, with special emphasis on the synergy between experiments and theory.
-We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further refinement of the theory.
-Likewise, novel theoretical predictions have motivated additional experimental measurements, and we present a comparison between the two.
-This close cooperation between theory and experiments has proven a successful strategy \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
+The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019anomalous} and Majda et al.~\cite{majda2019statistical, majdaqi2019}.
+Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
+We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory.
+Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, in particular on the relationship between surface-displacment skewness and surface-slope variance.
+Comparison between the two further confirms the predictive power of the theoretical framework developed in \cite{majda2019statistical}.
+In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two.
+In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in \cite{majda2019statistical} for the sake of brevity.
 
-The purpose of this manuscript is to:
-\begin{itemize}
-\item Provide a more comprehensive treatment of the experiments and theory in combination.
-In particular, we make new comparisons between theory and experiments, which further confirm the predictive power of the theoretical framework.
-\item Provide more thorough treatment of the link between physical system parameters and model parameters.
-This will elucidate the connection between theory and experiments.
-In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in \cite{majda2019statistical} for the sake of brevity.
-\item Misc: We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.
-We give a more transparent connection to experimental parameters. More careful development of theory and experiments, *in conjunction* with one another.
-\item Misc: combination; draw close comparison; with an even closer comparison between theoretical predictions and experimental measurements.
-\end{itemize}
+% Words: conjunction, combination, close cooperation; elucidate.
+% New data: slopes, autocorrelation, depth ratio varied.
 
-% New data: slopes, autocorrelation, depth ratio varied.
 % Definitely cite several of Sapsis's new papers.

### 2019-10-01 paper2_JNLS/paper.tex
NM

@@
 In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
-Upon passing into the region of shallower depth, the wave distribution skews markedly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
-Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical foundation that accurately captures several aspects of the anomalous behavior.
+Upon passing into the region of shallower depth, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
+Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical foundation that accurately captures several key aspects of the anomalous behavior.
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
@@
 TO DO: Describe the `representative' experiments in which $\dratdn = 0.24$ and $\epsup = 0.017$.
+
+Show diagram, talk about different statistics on either side of ADC. Say Bolles found the downstream measurements to be remarkably well fit by a mean-zero gamma distribution.
 

### 2019-10-02 paper2_JNLS/paper.tex
NM

@@
 The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.
+In this way, it makes sense to approach anomalous waves from the broader perspective of turbulent dynamical systems \cite{sapsis2013a, sapsis2013b, sapsis2013blending, chen2016filtering, majda2016introduction, macedo2017universality, MajdaQiSIAM2018, blonigan2019extreme, guth2019machine}.
 
@@
 In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
-Upon passing into the region of shallower depth, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
-Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical foundation that accurately captures several key aspects of the anomalous behavior.
+Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
+Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior.
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
-In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous properties while enjoying a more tractable structure.
+In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated systems of lower dimension that were shown to exhibit many of the same anomalous properties while enjoying a more tractable structure.
 
@@
 We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory.
-Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, in particular on the relationship between surface-displacment skewness and surface-slope variance.
-Comparison between the two further confirms the predictive power of the theoretical framework developed in \cite{majda2019statistical}.
+Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, particularly regarding a relationship between surface-displacement skewness and surface-slope variance that emerges from the theory.
+Comparison with the new measurements further confirms the predictive power of the theoretical framework developed by Majda et al.~\cite{majda2019statistical}.
 In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two.
-In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in \cite{majda2019statistical} for the sake of brevity.
+In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.
 

### 2019-10-02 paper2_JNLS/paper.tex
NM

@@
 
-A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
+A recent series of laboratory \cite{} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
 Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous wave statistics.
@@
 
-Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC).
+Our particular focus is the laboratory experiments of Bolles et al.~\cite{}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC).
 In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
 Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
-Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior.
+Inspired by these experimental observations, Majda et al.~\cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior.
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
@@
 
-The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019anomalous} and Majda et al.~\cite{majda2019statistical, majdaqi2019}.
+The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{} and Majda et al.~\cite{majda2019, majdaqi2019}.
 Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
 We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory.
-Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, particularly regarding a relationship between surface-displacement skewness and surface-slope variance that emerges from the theory.
-Comparison with the new measurements further confirms the predictive power of the theoretical framework developed by Majda et al.~\cite{majda2019statistical}.
+Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, particularly regarding a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
+Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}.
 In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two.
@@
 
-% Definitely cite several of Sapsis's new papers.
-
 %----------------------------------------------------------%
@@
 
-	As discussed in \cite{bolles2019anomalous}, the experiments consist of a long, narrow wave tank (6 m long x 20 cm wide x 30 cm high), with waves generated by plexiglass paddle at one end.
+	As diagrammed in Fig.~\ref{ExpDiagStats}(a), the experiments consist of a long, narrow wave tank (6 m long x 20 cm wide x 30 cm high), with waves generated by plexiglass paddle at one end \cite{bolles2019}.
 The waves propagate through the tank and, roughly midway through, pass over an abrupt depth change (ADC) created by a plexiglass step.
 The waves continue to propagate through the shallower depth until reaching the far end of the tank, at which point their energy is dissipated by a horse-hair dampener.
-Since the dampener minimizes the backscatter, the waves in this experiment propagate primarily in one direction, from left to right in Fig.~\ref{ExpDiagStats}.
+Since the dampener minimizes backscatter, the waves in this experiment propagate primarily in one direction, from left to right in Fig.~\ref{ExpDiagStats}(a).
 
@@
 \end{align}
-(Literal copy and paste) Here, the angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively.
-We set $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz.
-Importantly, the phases $\delta_n$ are uniformly distributed random variables.
-We fix $N = 3000$, which sets a fundamental period of $T = 300$ seconds.
+The angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively.
+As in prior work, all experiments reported here use the values $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz.
+Meanwhile, the phases $\delta_n$ are uniformly distributed random variables.
 
@@
 \caption{
-(a) Experimental schematic. Adapted from \cite{bolles2019anomalous}.
-(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms.
-}
-%(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms.
+(a) Experimental schematic. 
+(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms. Figure adapted from \cite{bolles2019}.}
 \label{ExpDiagStats}
@@
 
-	The free surface is illuminated by light-emitting diodes running along the bottom of the tank and then imaged from the sideview with a Nikon D3300 at 60 frames per second.
-The illumination technique coupled with high pixel count of the camera allows surface displacements to be resolved with a accuracy better than 1/3 millimeter.
-Representative free surface measurements are shown in Figs.~\ref{ExpDiagStats}(b)--(c) at a position upstream and downstream of the ADC.
-The corresponding histograms are shown in Figs.~\ref{ExpDiagStats}(d)--(e).
+	The free surface is illuminated by light-emitting diodes that run along the bottom of the tank and is then imaged from the sideview with a Nikon D3300 at 60 frames per second.
+The illumination technique, coupled with high pixel count of the camera, allows surface displacements to be resolved with accuracy better than 1/3 millimeter.
+Furthermore, these optical measurements permit extraction of wave statistics {\em continuously} in space, rather than at a few discrete locations.
+Further details of the experimental setup can be found in \cite{bolles2019}.
 
+	Example measurements of free-surface displacements $\eta$ are shown in Figs.~\ref{ExpDiagStats}(b)--(c).
+These measurements are taken at two representative locations: one a short distance upstream of the ADC and one downstream of the ADC.
+Both signals exhibit a combination of periodic and random fluctuations, with the dominant oscillations corresponding to the peak driving frequency of 2 Hz.
+The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e).
+The upstream measurements are symmetrically distributed about the mean displacement of $\eta = 0$.
+In fact, Bolles et al.~found that these measurements closely follow a Gaussian distribution \cite{bolles2019}.
+The downstream measurements, however, are not symmetrically distributed and skew strongly towards positive displacement, $\eta > 0$.
+Bolles et al.~found that these measurements can be well approximated by a mean-zero gamma distribution \cite{bolles2019}.
+
+\vsp{5}
 TO DO: Describe the `representative' experiments in which $\dratdn = 0.24$ and $\epsup = 0.017$.
 
-Show diagram, talk about different statistics on either side of ADC. Say Bolles found the downstream measurements to be remarkably well fit by a mean-zero gamma distribution.
+ % More details in \cite{}
 
@@
 
-In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the statistical-mechanics framework introduced in \cite{majda2019statistical}.
+In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the statistical-mechanics framework introduced in \cite{majda2019}.
 Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other choices.
@@
 A few comments are in order.
-First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ in \eqref{dimlessKdV} but with different expressions for the other dimensionless parameters.
+First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019}, with identical powers of $\drat$ in \eqref{dimlessKdV} but with different expressions for the other dimensionless parameters.
 These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments.
@@
 
-%NOTE: Perhaps we should be more specific about comparing to the dimensionless parameters in our PNAS paper \cite{majda2019statistical}, BUT the expressions for E0 and L0 given in the PNAS paper do not match with what I calculate comparing to the above.
+%NOTE: Perhaps we should be more specific about comparing to the dimensionless parameters in our PNAS paper \cite{majda2019}, BUT the expressions for E0 and L0 given in the PNAS paper do not match with what I calculate comparing to the above.
 Plus, it seems like it could only add confusion.
@@
 \end{align}
-The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$.
+The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019}, which used the opposite sign in both $\sympJ$ and $\Ham$.
 With that sign convention, Majda et al.
-2019 found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019statistical}.
+2019 found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}.
 We have chosen the sign convention above (which also happens to be consistent with Lax \cite{lax1975periodic}) so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.
@@
 
-Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019statistical}.
+Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019}.
 The piecewise defined Hamiltonian for TKdV is given by
@@
 \subsection{Explicit formula for outgoing skewness}
-We now examine an explicit formula for the outgoing skewness as originally founded in \cite{majda2019statistical}.
+We now examine an explicit formula for the outgoing skewness as originally founded in \cite{majda2019}.
 An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$.

### 2019-10-02 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\omsig}{\sigma_{\omega}}
+\newcommand{\Dphi}{\Delta \phi}
+
 %----------------------------------------------------------------------------%
@@
 
-The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{} and Majda et al.~\cite{majda2019, majdaqi2019}.
+The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}.
 Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
@@
 \label{PaddleAngle}
-& \phi(t) = \phi_0 + \Delta \phi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
+& \phi(t) = \phi_0 + \Dphi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
 \label{anEq}
@@
 Meanwhile, the phases $\delta_n$ are uniformly distributed random variables.
+The standard deviation of the paddle angle, $\Dphi$, controls the amplitude of the generated waves.
+In a single experiment $\Dphi$ is fixed, and we will present a series of experiments with systematically varied values of $\Dphi$.
 
@@
 
-	The free surface is illuminated by light-emitting diodes that run along the bottom of the tank and is then imaged from the sideview with a Nikon D3300 at 60 frames per second.
+	The free surface is illuminated by light-emitting diodes running along the bottom of the tank and is imaged from the sideview with a Nikon D3300 at 60 frames per second.
 The illumination technique, coupled with high pixel count of the camera, allows surface displacements to be resolved with accuracy better than 1/3 millimeter.
@@
 	Example measurements of free-surface displacements $\eta$ are shown in Figs.~\ref{ExpDiagStats}(b)--(c).
-These measurements are taken at two representative locations: one a short distance upstream of the ADC and one downstream of the ADC.
-Both signals exhibit a combination of periodic and random fluctuations, with the dominant oscillations corresponding to the peak driving frequency of 2 Hz.
+These measurements are extracted from the images at two representative locations: one a short distance upstream of the ADC and one downstream of the ADC.
+Both signals exhibit a combination of periodic behavior and random fluctuations, with the dominant oscillations corresponding to the peak driving frequency of 2 Hz.
 The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e).
-The upstream measurements are symmetrically distributed about the mean displacement of $\eta = 0$.
+The upstream measurements are symmetrically distributed about the mean, $\eta = 0$.
 In fact, Bolles et al.~found that these measurements closely follow a Gaussian distribution \cite{bolles2019}.
 The downstream measurements, however, are not symmetrically distributed and skew strongly towards positive displacement, $\eta > 0$.
-Bolles et al.~found that these measurements can be well approximated by a mean-zero gamma distribution \cite{bolles2019}.
+Bolles et al.~found that these measurements can be well described by a mean-zero gamma distribution \cite{bolles2019}.
+The slower decay rate of this downstream distribution indicates an elevated level of extreme surface displacement, i.e.~rogue waves.
+Bolles et al.~estimated that a rogue wave can be up to 65 times more likely in these experiments than if the displacements were Gaussian.
+
+	The measurements shown in Fig.~\ref{ExpDiagStats} correspond to a paddle driving amplitude of $\Dphi = 1.38^{\circ}$, an upstream depth of $\dup = 12.5$ cm, a downstream depth of $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$.
+Bolles et al.~systematically varied the driving amplitude in the range $0.125^{\circ}$--$2^{\circ}$ to probe different regimes of wave behavior, from nearly linear waves to strongly nonlinear ones.
+They found that once a critical amplitude is exceeded (roughly $0.5^{\circ}$), the downstream statistics become anomalous and are robustly described by the gamma-distribution.
+No variation of the depth ratio was reported though.
+In the current manuscript, we also vary the downstream depth.
+We, however, will refer to the representative set of experiments from Fig.~\ref{ExpDiagStats} as a baseline for comparison.
 

### 2019-10-02 paper2_JNLS/paper.tex
NM

@@
 \maketitle
+
+\begin{abstract} 
+Recent laboratory experiments revealed the emergence of anomalous statistics in a randomized surface waves upon encountering an abrupt depth change (ADC) in the bottom topography (Bolles et al.
+PRF 2019).
+Majda et al.~2019 established a theoretical foundation, based on analysis of the truncated KdV (TKdV) equations, that successfully captures several key qualitative features of the experiments, including the robust emergence of anomalous statistics.
+The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the ADC, and is corroborated by simulations of the deterministic dynamics.
+%
+Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two.
+We present an assortment of new experimental measurements and carefully compare them to model predictions.
+We establish a more precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in these experiment, as well as others.
+The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in spectacular fashion by direct experimental measurements.
+\end{abstract}
+
+%Here, we present a more comprehensive treatment of the experimental and modeling efforts in conjunction with on another.
+We report new experimental measurements and new theoretical results, with a detailed comparison between the two.
 

### 2019-10-02 paper2_JNLS/paper.tex
NM

@@
 
-\title{The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes}
+\title{Anomalous waves triggered by abrupt depth changes: experimental measurements and deterministic-statistical truncated KdV theory}
 
-% Title ideas
-% Could include words "theory and experiments".
-%Old idea: Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change
+% ORIGINAL: The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes
+% Words: triggered, induced, generated, sparked
+
 \author{
@@
 PRF 2019).
-Majda et al.~2019 established a theoretical foundation, based on analysis of the truncated KdV (TKdV) equations, that successfully captures several key qualitative features of the experiments, including the robust emergence of anomalous statistics.
+Majda et al.~2019 established a theoretical foundation, based on analysis of the truncated KdV (TKdV) system, that successfully captures several key qualitative features of the experiments including the robust emergence of anomalous statistics.
 The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the ADC, and is corroborated by simulations of the deterministic dynamics.
@@
 We present an assortment of new experimental measurements and carefully compare them to model predictions.
-We establish a more precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in these experiment, as well as others.
+We establish a more precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments.
 The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in spectacular fashion by direct experimental measurements.
 \end{abstract}
-
-%Here, we present a more comprehensive treatment of the experimental and modeling efforts in conjunction with on another.
-We report new experimental measurements and new theoretical results, with a detailed comparison between the two.
 

### 2019-10-04 paper2_JNLS/paper.tex
NM

@@
 
-A recent series of laboratory \cite{} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
+A recent series of laboratory \cite{bolles2019} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
 Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous wave statistics.
@@
 
-Our particular focus is the laboratory experiments of Bolles et al.~\cite{}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC).
+Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC).
 In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
@@
 (a) Experimental schematic. 
-(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms. Figure adapted from \cite{bolles2019}.}
+(b)--(c) Surface displacement measured at a representative upstream and downstream locations. (d)--(e) Corresponding histograms. Figure adapted from \cite{bolles2019}.}
 \label{ExpDiagStats}
@@
 The illumination technique, coupled with high pixel count of the camera, allows surface displacements to be resolved with accuracy better than 1/3 millimeter.
-Furthermore, these optical measurements permit extraction of wave statistics {\em continuously} in space, rather than at a few discrete locations.
+Furthermore, these optical measurements permit extraction of wave statistics {\em continuously} in space, rather than at a few discrete locations, which is crucial for identifying regions of anomalous wave activity.
 Further details of the experimental setup can be found in \cite{bolles2019}.
@@
 	Example measurements of free-surface displacements $\eta$ are shown in Figs.~\ref{ExpDiagStats}(b)--(c).
-These measurements are extracted from the images at two representative locations: one a short distance upstream of the ADC and one downstream of the ADC.
+These measurements are extracted from the images at two representative locations: one a short distance (9 cm) upstream of the ADC and the other (15 cm) downstream of the ADC.
 Both signals exhibit a combination of periodic behavior and random fluctuations, with the dominant oscillations corresponding to the peak driving frequency of 2 Hz.
@@
 The upstream measurements are symmetrically distributed about the mean, $\eta = 0$.
-In fact, Bolles et al.~found that these measurements closely follow a Gaussian distribution \cite{bolles2019}.
+In fact, Bolles et al.~found that these measurements follow a Gaussian distribution closely \cite{bolles2019}.
 The downstream measurements, however, are not symmetrically distributed and skew strongly towards positive displacement, $\eta > 0$.
@@
 
-	The measurements shown in Fig.~\ref{ExpDiagStats} correspond to a paddle driving amplitude of $\Dphi = 1.38^{\circ}$, an upstream depth of $\dup = 12.5$ cm, a downstream depth of $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$.
+	The measurements shown in Fig.~\ref{ExpDiagStats} correspond to a paddle driving amplitude of $\Dphi = 1.38^{\circ}$, an upstream depth of $\dup = 12.5$ cm, and a downstream depth of $\ddn = 3$ cm, which gives the depth ratio $\dratdn = \ddn/\dup= 0.24$.
 Bolles et al.~systematically varied the driving amplitude in the range $0.125^{\circ}$--$2^{\circ}$ to probe different regimes of wave behavior, from nearly linear waves to strongly nonlinear ones.
-They found that once a critical amplitude is exceeded (roughly $0.5^{\circ}$), the downstream statistics become anomalous and are robustly described by the gamma-distribution.
-No variation of the depth ratio was reported though.
-In the current manuscript, we also vary the downstream depth.
-We, however, will refer to the representative set of experiments from Fig.~\ref{ExpDiagStats} as a baseline for comparison.
-
-\vsp{5}
-TO DO: Describe the `representative' experiments in which $\dratdn = 0.24$ and $\epsup = 0.017$.
-
- % More details in \cite{}
+They found that once a critical amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the downstream statistics become anomalous and are robustly described by the gamma-distribution.
+Fig.~\ref{ExpSpatialStats} shows spatial statistics for several experiments of different driving amplitudes.
+The standard deviation of surface displacement $\etastd$ measures the characteristic amplitude of waves (Fig.~\ref{ExpSpatialStats}(a)).
+Notice that $\etastd$ increases with paddle amplitude, but does not significantly vary in space.
+In particular, $\etastd$ does not change significantly when waves cross the ADC.
+Skewness and (excess) kurtosis, however change markedly crossing the ADC once the critical driving amplitude is exceeded.
+We also show statistics of surface slope $\eta_x$.
 
@@
 \includegraphics[width = 0.9 \linewidth]{Figs/ExpSpatialStats.pdf}
-\caption{Spatial moments}
+\caption{Wave statistics as they vary in space for several different experiments of different driving amplitudes (see legend).
+$\etastd$ does not vary significantly crossing the ADC (a), but skewness (b) and kurtosis (c) show a strong response to the ADC.
+}
 \label{ExpSpatialStats}
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+	Bolles et al.~systematically explored the effect of varying the driving amplitude on the anomalous statistics for a fixed depth ratio.
+Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we will also report experiments in which the depth ratio was varied.
+We will refer to the experiments displayed in Fig.~\ref{ExpDiagStats} as the reference experiments, so that we have a baseline for comparison.
+In these reference experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the cases $\Dphi = 1.25^{\circ}$ and $1.5^{\circ}$ shown in Fig.~\ref{ExpDiagStats}(a)), the depth ratio is $\dratdn = 0.24$, and the characteristic wave amplitude is $\etastd = 0.21$ cm.
+
+%$\epsup = 0.017$.
 
@@
 Unless stated otherwise, all parameters used in the theory are taken directly from their experimental values listed in Table \ref{paramtable}.
-In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.17$.
+In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.017$.
 

### 2019-10-04 paper2_JNLS/paper.tex
NM

@@
 \section{Theoretical framework}
+
+We now introduce the theoretical framework that will be used to understand and quantify the experimental observations.
+This framework is based on a Galerkin truncation of the variable-depth Korteweg–de Vries (KdV) equation.
+The KdV equation is a commonly excepted model for describing the propagation of unidirectional, shallow-water waves, accounting for weak nonlinearity and weak dispersion over long timescales and large spatial scales.
+We will perform Galerkin truncation of this system at wavenumber $\Lambda = $8--32 to obtain a finite dimensional dynamical system exhibiting weak turbulence.
+We outline the Hamiltonian structure of both the continuous KdV and the truncated system.
+This structure is exploited to obtain invariant measures of the underlying dynamics and to ultimately understand the experimental findings on anomalous wave statistics triggered by an ADC.
 

### 2019-10-04 paper2_JNLS/paper.tex
NM

@@
 
-We will primarily consider the case in which waves originate from a region of constant depth, encounter an abrupt depth change, and then continue in another region of constant depth.
+As in the experiments, we consider waves that originate from a region of constant depth, encounter an abrupt depth change, and continue in another region of constant depth.
 Thus, depth will be piecewise constant

### 2019-10-06 paper2_JNLS/paper.tex
NM

@@
 Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
-Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and the Benjamin-Feir deep-water modulational instability.
+Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing \citep{kharif2008influence, toffoli2017wind},
+focusing due to variable bathymetry \citep{heller2008refraction, white1998chance}, 
+and the Benjamin-Feir deep-water modulational instability \citep{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
 The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.

### 2019-10-06 paper2_JNLS/paper.tex
NM

--- 2019-10-04 paper2_JNLS/paper.tex
@@
 
-Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height.
+Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height \citep{muller2005rogue, ying2011linear}.
 Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
-Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and the Benjamin-Feir deep-water modulational instability.
+Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing \citep{kharif2008influence, toffoli2017wind},
+focusing due to variable bathymetry \citep{heller2008refraction, white1998chance}, 
+opposing currents \citep{garrett2009rogue, onorato2011triggering},
+and the Benjamin-Feir deep-water modulational instability \citep{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
 The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.

### 2019-10-07 paper2_JNLS/paper.tex
NM

@@
 Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
-Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing \citep{kharif2008influence, toffoli2017wind},
+Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing \citep{kharif2008influence, toffoli2017wind}, opposing currents \citep{garrett2009rogue, onorato2011triggering},
 focusing due to variable bathymetry \citep{heller2008refraction, white1998chance}, 
-opposing currents \citep{garrett2009rogue, onorato2011triggering},
 and the Benjamin-Feir deep-water modulational instability \citep{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
@@
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
-In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated systems of lower dimension that were shown to exhibit many of the same anomalous properties while enjoying a more tractable structure.
+In subsequent work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated systems that were shown to exhibit many of the same anomalous properties while enjoying a more tractable, lower-dimensional structure.
 
@@
 We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory.
-Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, particularly regarding a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
+Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
 Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}.
@@
 As in prior work, all experiments reported here use the values $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz.
-Meanwhile, the phases $\delta_n$ are uniformly distributed random variables.
-The standard deviation of the paddle angle, $\Dphi$, controls the amplitude of the generated waves.
-In a single experiment $\Dphi$ is fixed, and we will present a series of experiments with systematically varied values of $\Dphi$.
+The phases $\delta_n$ are uniformly distributed random variables, which results in a randomized wave train.
+The standard deviation of the paddle angle, $\Dphi$, controls the amplitude overall amplitude of the waves.
+In a single experiment $\Dphi$ is fixed, and we will present a series of experiments with $\Dphi$ varied systematically.
 
@@
 
-	The free surface is illuminated by light-emitting diodes running along the bottom of the tank and is imaged from the sideview with a Nikon D3300 at 60 frames per second.
+	The free surface is illuminated by light-emitting diodes and is imaged from the sideview with a Nikon D3300 at 60 frames per second.
 The illumination technique, coupled with high pixel count of the camera, allows surface displacements to be resolved with accuracy better than 1/3 millimeter.
@@
 	Example measurements of free-surface displacements $\eta$ are shown in Figs.~\ref{ExpDiagStats}(b)--(c).
-These measurements are extracted from the images at two representative locations: one a short distance (9 cm) upstream of the ADC and the other (15 cm) downstream of the ADC.
-Both signals exhibit a combination of periodic behavior and random fluctuations, with the dominant oscillations corresponding to the peak driving frequency of 2 Hz.
+These measurements are extracted from the images at two representative locations: one a short distance (9 cm) upstream of the ADC and the other a short distance (15 cm) downstream.
+Both signals exhibit a combination of periodic and random behavior, with the dominant oscillations corresponding to the peak forcing frequency of 2 Hz.
 The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e).
@@
 In fact, Bolles et al.~found that these measurements follow a Gaussian distribution closely \cite{bolles2019}.
-The downstream measurements, however, are not symmetrically distributed and skew strongly towards positive displacement, $\eta > 0$.
-Bolles et al.~found that these measurements can be well described by a mean-zero gamma distribution \cite{bolles2019}.
+The downstream measurements, however, skew strongly towards positive displacement, $\eta > 0$.
+Bolles et al.~found these measurements to be well described by a mean-zero gamma distribution \cite{bolles2019}.
 The slower decay rate of this downstream distribution indicates an elevated level of extreme surface displacement, i.e.~rogue waves.
-Bolles et al.~estimated that a rogue wave can be up to 65 times more likely in these experiments than if the displacements were Gaussian.
+Bolles et al.~estimated that a rogue wave can be up to 65 times more likely in these experiments than if displacements were Gaussian \cite{bolles2019}.
 
-	The measurements shown in Fig.~\ref{ExpDiagStats} correspond to a paddle driving amplitude of $\Dphi = 1.38^{\circ}$, an upstream depth of $\dup = 12.5$ cm, and a downstream depth of $\ddn = 3$ cm, which gives the depth ratio $\dratdn = \ddn/\dup= 0.24$.
-Bolles et al.~systematically varied the driving amplitude in the range $0.125^{\circ}$--$2^{\circ}$ to probe different regimes of wave behavior, from nearly linear waves to strongly nonlinear ones.
-They found that once a critical amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the downstream statistics become anomalous and are robustly described by the gamma-distribution.
-Fig.~\ref{ExpSpatialStats} shows spatial statistics for several experiments of different driving amplitudes.
-The standard deviation of surface displacement $\etastd$ measures the characteristic amplitude of waves (Fig.~\ref{ExpSpatialStats}(a)).
-Notice that $\etastd$ increases with paddle amplitude, but does not significantly vary in space.
-In particular, $\etastd$ does not change significantly when waves cross the ADC.
-Skewness and (excess) kurtosis, however change markedly crossing the ADC once the critical driving amplitude is exceeded.
-We also show statistics of surface slope $\eta_x$.
+	The paddle amplitude in Fig.~\ref{ExpDiagStats} is $\Dphi = 1.38^{\circ}$, and this value was varied systematically in the range $\Dphi = 0.125^{\circ}$--$2^{\circ}$ to probe the different regimes of wave behavior, from linear to strongly nonlinear waves.
+Figure \ref{ExpSpatialStats} shows spatial statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ for six different driving amplitudes (see legend).
+For each quantity, we show the standard deviation, the skewness, and the (excess) kurtosis as they vary in the vicinity of the ADC, located at $x=0$.
+First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves.
+Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC.
+In experiments with sufficiently high driving amplitude, though, the skewness and kurtosis (b)--(c) respond strongly to the ADC.
+In these experiments, both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm.
+Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude.
+The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e).
+Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution.
+
+	The right column of Fig.~\ref{ExpSpatialStats}
 
@@
 \begin{center}
-\includegraphics[width = 0.9 \linewidth]{Figs/ExpSpatialStats.pdf}
+\includegraphics[width = 0.8 \linewidth]{Figs/ExpSpatialStats.pdf}
 \caption{Wave statistics as they vary in space for several different experiments of different driving amplitudes (see legend).
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+\vsp{5}
+%The measurements shown in Fig.~\ref{ExpDiagStats} correspond to a paddle driving amplitude of $\Dphi = 1.38^{\circ}$, an upstream depth of $\dup = 12.5$ cm, and a downstream depth of $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$.
 

### 2019-10-07 paper2_JNLS/paper.tex
NM

@@
 
-	The right column of Fig.~\ref{ExpSpatialStats}
+	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}.
+Our attention to slope statistics was motivated by new theoretical developments by Majda et al.~\cite{majda2019}.
+The surface slope is extracted from the experimental images through numerical differentiation of the free surface using Savitzky-Golay methods.
+This ability to extract surface slope is another advantage of optical measurements over discrete probes.
 

### 2019-10-07 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\vardn}{\var_+}
+\newcommand{\kurt}{\text{kurt}}
+\newcommand{\std}{\text{std}}
 % Dimensionless parameters
@@
 
+% NOTES
 % Words: conjunction, combination, close cooperation; elucidate.
@@
 
-	The paddle amplitude in Fig.~\ref{ExpDiagStats} is $\Dphi = 1.38^{\circ}$, and this value was varied systematically in the range $\Dphi = 0.125^{\circ}$--$2^{\circ}$ to probe the different regimes of wave behavior, from linear to strongly nonlinear waves.
-Figure \ref{ExpSpatialStats} shows spatial statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ for six different driving amplitudes (see legend).
-For each quantity, we show the standard deviation, the skewness, and the (excess) kurtosis as they vary in the vicinity of the ADC, located at $x=0$.
-First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves.
-Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC.
-In experiments with sufficiently high driving amplitude, though, the skewness and kurtosis (b)--(c) respond strongly to the ADC.
-In these experiments, both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm.
-Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude.
-The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e).
-Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution.
-
-	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}.
-Our attention to slope statistics was motivated by new theoretical developments by Majda et al.~\cite{majda2019}.
-The surface slope is extracted from the experimental images through numerical differentiation of the free surface using Savitzky-Golay methods.
-This ability to extract surface slope is another advantage of optical measurements over discrete probes.
-
 % Figure: Experimental Spatial Statistics
@@
 
-\vsp{5}
-%The measurements shown in Fig.~\ref{ExpDiagStats} correspond to a paddle driving amplitude of $\Dphi = 1.38^{\circ}$, an upstream depth of $\dup = 12.5$ cm, and a downstream depth of $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$.
+	The paddle amplitude in Fig.~\ref{ExpDiagStats} is $\Dphi = 1.38^{\circ}$, and this value was varied systematically in the range $\Dphi = 0.125^{\circ}$--$2^{\circ}$ to probe the different regimes of wave behavior, from linear to strongly nonlinear waves.
+Figure \ref{ExpSpatialStats} shows spatial statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ for six different driving amplitudes (see legend).
+For each quantity, we show the standard deviation, the skewness, and the (excess) kurtosis as they vary in the vicinity of the ADC, located at $x=0$.
+First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves.
+Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC.
+In experiments with sufficiently high driving amplitude, though, the skewness and kurtosis (b)--(c) respond strongly to the ADC.
+In these experiments, both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm.
+Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude.
+The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e).
+Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution.
+We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.
+
+	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}.
+Our attention to slope statistics was motivated by new theoretical developments by Majda et al.~\cite{majda2019}.
+We extract the surface slope by numerical differentiating images of the free surface using Savitzky-Golay smoothing filters.
+This ability to extract surface slope is another advantage of optical measurements over discrete probes.
+As seen in Fig.~\ref{ExpSpatialStats}(a) the standard deviation of slope behaves similar to displacement: $\std(\eta_x)$ increases with driving amplitude but remains nearly uniform in space in each individual experiment.
+The higher-order moments, however, once again respond strongly to the ADC.
+The skewness of slope becomes highly negative and reaches a minimum near $x = 8$ cm, which is downstream of the ADC but upstream of where the displacement reaches its peak.
+The kurtosis of slope is positive (by construction) and reaches a peak at nearly the same location of 8 cm.
+The large negative skewness of surface slope indicates a bias towards waves of large negative slope, which is consistent with a right-moving wave that is near overturning.
 
 	Bolles et al.~systematically explored the effect of varying the driving amplitude on the anomalous statistics for a fixed depth ratio.
-Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we will also report experiments in which the depth ratio was varied.
-We will refer to the experiments displayed in Fig.~\ref{ExpDiagStats} as the reference experiments, so that we have a baseline for comparison.
-In these reference experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the cases $\Dphi = 1.25^{\circ}$ and $1.5^{\circ}$ shown in Fig.~\ref{ExpDiagStats}(a)), the depth ratio is $\dratdn = 0.24$, and the characteristic wave amplitude is $\etastd = 0.21$ cm.
+Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we also report experiments in which the depth ratio was varied.
+To have a simple baseline for comparison against theory, we refer to the experiments from Fig.~\ref{ExpDiagStats} as the {\em reference} experiments.
+In these experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the cases $\Dphi = 1.25^{\circ}$ and $1.5^{\circ}$ shown in Fig.~\ref{ExpDiagStats}(a)), the upstream depth is $\dup = 12.5$ cm, and the downstream depth is $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$.
+From the data represented in Fig.\ref{ExpSpatialStats}, we extract a characteristic wave amplitude of $\etastd = 0.21$ cm for these representative experiments, which will be important for setting dimensionless parameters that enter the theory.
 
+% NOTES
+% Maybe modify slightly this last paragraph if it does not seem to fit next time I read it.
  %$\epsup = 0.017$.

### 2019-10-07 paper2_JNLS/paper.tex
NM

@@
 Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we also report experiments in which the depth ratio was varied.
-To have a simple baseline for comparison against theory, we refer to the experiments from Fig.~\ref{ExpDiagStats} as the {\em reference} experiments.
+To have a simple baseline for comparison against theory, we refer to the data from Fig.~\ref{ExpDiagStats} as the {\em reference} experiments.
 In these experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the cases $\Dphi = 1.25^{\circ}$ and $1.5^{\circ}$ shown in Fig.~\ref{ExpDiagStats}(a)), the upstream depth is $\dup = 12.5$ cm, and the downstream depth is $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$.
-From the data represented in Fig.\ref{ExpSpatialStats}, we extract a characteristic wave amplitude of $\etastd = 0.21$ cm for these representative experiments, which will be important for setting dimensionless parameters that enter the theory.
+From the data represented in Fig.~\ref{ExpSpatialStats}, we extract a characteristic wave amplitude for the representative experiments of $\etastd = 0.21$ cm, which will be important for setting dimensionless parameters that enter the theory.
+Table \ref{paramtable} lists the range of experimental parameters along with these representative values, as well as value of dimensionless parameters that will be introduced later.
+
+% Table
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{table}[h]%[htbp]
+\begin{center}
+\caption{Table of parameters}
+\label{paramtable}
+\begin{tabular}{l l l l}
+%\hline \multicolumn{3} { c }{Parameters that are constant in a single experiment} \\
+\hline Description & Notation & Experimental range & `Representative' values \\
+\hline
+Peak forcing frequency	& $f_p$					& 2 Hz 		& 2 Hz \\
+Characteristic amplitude	& $\etastd$				& 0.03--0.3 cm	& 0.21 cm \\
+Upstream depth		& $\dup$					& 12.5 cm 	& 12.5 cm \\
+Downstream depth		& $\ddn$					& RANGE		& 3 cm \\
+Upstream wavelength	& $\lamup = \sqrt{g \dup}/f_p$	& 55 cm 		& 55 cm \\
+Downstream wavelength	& $\lamdn = \sqrt{g \ddn}/f_p$	& RANGE		& 27 cm \\
+%
+Amplitude-to-depth ratio	& $\epsup = \etastd / \dup$	&0.0024--0.024	& 0.017 \\
+Depth-to-wavelength ratio	& $\delup = \dup / \lamup$	& RANGE		& 0.22 \\
+Depth ratio			& $\dratdn = \ddn/\dup$		& RANGE		& 0.24 \\
+\end{tabular}
+\end{center}
+\end{table}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 
 % NOTES
-% Maybe modify slightly this last paragraph if it does not seem to fit next time I read it.
+% Maybe modify slightly the last paragraph if it does not seem to fit next time I read it.
  %$\epsup = 0.017$.
@@
 The KdV equation is a commonly excepted model for describing the propagation of unidirectional, shallow-water waves, accounting for weak nonlinearity and weak dispersion over long timescales and large spatial scales.
-We will perform Galerkin truncation of this system at wavenumber $\Lambda = $8--32 to obtain a finite dimensional dynamical system exhibiting weak turbulence.
+We will perform Galerkin truncation of this system at wavenumber $\Lambda = $8--32 to obtain a finite-dimensional dynamical system that exhibits weak turbulence.
 We outline the Hamiltonian structure of both the continuous KdV and the truncated system.
@@
 See Table \ref{paramtable} for a summary of important parameters.
-
-% Table
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{table}[h]%[htbp]
-\begin{center}
-\caption{Table of parameters}
-\label{paramtable}
-\begin{tabular}{l l l}
-\hline \multicolumn{3} { c }{Parameters that are constant in a single experiment} \\
-\hline Description & Notation and definition & Value in experiments \\
-\hline
-Peak forcing frequency		& $f_p$						& 2 Hz \\
-Characteristic wave amplitude	& $\etastd = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
-Upstream depth			& $\dup$						& 12.5 cm \\
-Downstream depth			& $\ddn$						& 3 cm (and varied) \\
-Upstream wavelength		& $\lamup = \sqrt{g \dup}/f_p$		& 55 cm \\
-Upstream wavelength		& $\lamdn = \sqrt{g \ddn}/f_p$		& 27 cm \\
-%
-Amplitude-to-depth ratio		& $\epsup = \etastd / \dup$			& 0.0024--0.024 \\
-Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
-Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 (and varied)
-\end{tabular}
-\end{center}
-\end{table}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 

### 2019-10-07 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\np}{\newpage \noindent}
+\newcommand{\nick}[1]{{\color{red} #1}}
 % Derivatives
@@
 \begin{center}
-\includegraphics[width = 0.8 \linewidth]{Figs/ExpSpatialStats.pdf}
+\includegraphics[width = 0.85 \linewidth]{Figs/ExpSpatialStats.pdf}
 \caption{Wave statistics as they vary in space for several different experiments of different driving amplitudes (see legend).
@@
 	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}.
-Our attention to slope statistics was motivated by new theoretical developments by Majda et al.~\cite{majda2019}.
+Our attention to slope statistics was motivated by new theoretical developments by Majda et al.~\cite{majda2019} and will be expanded upon in later sections.
 We extract the surface slope by numerical differentiating images of the free surface using Savitzky-Golay smoothing filters.
-This ability to extract surface slope is another advantage of optical measurements over discrete probes.
-As seen in Fig.~\ref{ExpSpatialStats}(a) the standard deviation of slope behaves similar to displacement: $\std(\eta_x)$ increases with driving amplitude but remains nearly uniform in space in each individual experiment.
-The higher-order moments, however, once again respond strongly to the ADC.
-The skewness of slope becomes highly negative and reaches a minimum near $x = 8$ cm, which is downstream of the ADC but upstream of where the displacement reaches its peak.
-The kurtosis of slope is positive (by construction) and reaches a peak at nearly the same location of 8 cm.
-The large negative skewness of surface slope indicates a bias towards waves of large negative slope, which is consistent with a right-moving wave that is near overturning.
+This ability to extract surface slope is another advantage of our optical measurements over the more commonly used technique of placing a set of discrete wave probes in the tank.
+As seen in Fig.~\ref{ExpSpatialStats}(a) the standard deviation of slope behaves similar to displacement: $\std(\eta_x)$ increases with driving amplitude, but remains nearly uniform in space for each individual experiment and is not affected by the ADC.
+The higher-order moments, however, once again respond strongly to the ADC if the driving amplitude is sufficiently high.
+The skewness of slope becomes highly negative and reaches a minimum near $x = 8$ cm, which is downstream of the ADC but upstream of the position where the displacement statistics peak.
+The kurtosis of $\eta_x$ reaches a peak at nearly the same location as skewness.
+We note that the large {\em negative} skewness of $\eta_x$ indicates a bias towards negative slope, which is consistent with a right-moving wave of steep leading surface and shallower trailing surface, i.e.~a wave that is near overturning.
 
@@
 To have a simple baseline for comparison against theory, we refer to the data from Fig.~\ref{ExpDiagStats} as the {\em reference} experiments.
-In these experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the cases $\Dphi = 1.25^{\circ}$ and $1.5^{\circ}$ shown in Fig.~\ref{ExpDiagStats}(a)), the upstream depth is $\dup = 12.5$ cm, and the downstream depth is $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$.
+In these experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the two largest amplitudes shown in Fig.~\ref{ExpSpatialStats}), the upstream depth is $\dup = 12.5$ cm, and the downstream depth is $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$.
 From the data represented in Fig.~\ref{ExpSpatialStats}, we extract a characteristic wave amplitude for the representative experiments of $\etastd = 0.21$ cm, which will be important for setting dimensionless parameters that enter the theory.
-Table \ref{paramtable} lists the range of experimental parameters along with these representative values, as well as value of dimensionless parameters that will be introduced later.
+Table \ref{paramtable} lists the range of experimental parameters along with these representative values, as well as values of dimensionless parameters that will be introduced later.
+% NOTE: Maybe modify slightly the last paragraph if it does not seem to fit next time I read it.
 
@@
 
-% NOTES
-% Maybe modify slightly the last paragraph if it does not seem to fit next time I read it.
- %$\epsup = 0.017$.
-
 %----------------------------------------------------------%
@@
 This framework is based on a Galerkin truncation of the variable-depth Korteweg–de Vries (KdV) equation.
-The KdV equation is a commonly excepted model for describing the propagation of unidirectional, shallow-water waves, accounting for weak nonlinearity and weak dispersion over long timescales and large spatial scales.
-We will perform Galerkin truncation of this system at wavenumber $\Lambda = $8--32 to obtain a finite-dimensional dynamical system that exhibits weak turbulence.
+The KdV equation is a commonly excepted model for describing the propagation of unidirectional, shallow-water waves, and it accounts for weak nonlinearity and weak dispersion over long timescales and large spatial scales.
+We will perform Galerkin truncation of this system at wavenumber $\Lambda = $ (in the range 8 to 32) to obtain a finite-dimensional dynamical system that exhibits weak turbulence.
 We outline the Hamiltonian structure of both the continuous KdV and the truncated system.
-This structure is exploited to obtain invariant measures of the underlying dynamics and to ultimately understand the experimental findings on anomalous wave statistics triggered by an ADC.
+This structure is exploited to obtain invariant measures of the underlying dynamics and, ultimately, to rationalize the experimental findings on anomalous wave statistics triggered by an ADC.
 
@@
 \subsection{The Korteweg–de Vries equation with variable depth}
-Consider waves propagating unidirectionally in shallow water.
-Consider the surface displacement $\eta(x,t)$ and the reference frame moving with the local wave speed $\xi = x - ct$, where $c = \sqrt{g \depth}$ is the wave speed, $g$ gravity, and $\depth$ the local depth.
-To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation (KdV), which in dimensional form is given by \cite{whitham2011linear}
+
+We consider the surface displacement $\eta(x,t)$ of unidirectional, shallow-water waves in a reference frame moving with the characteristic wave speed, $\xi = x - ct$.
+Here, $c = \sqrt{g \depth}$ is leading-order approximation to the wave speed (i.e.~from linear theory), where $g$ gravity and $\depth$ is the local depth.
+The leading-order dynamics are corrected to first order in small amplitude by the Korteweg–de Vries equation (KdV), which in dimensional form is given by \cite{whitham2011linear}
 \begin{equation}
@@
 \end{equation}
-% This dimensional form is given by Whitham on p. 461, Eq. 13.98. I first got it from the Wolfram KdV, which gives the same equation.
-
-As in the experiments, we consider waves that originate from a region of constant depth, encounter an abrupt depth change, and continue in another region of constant depth.
+% Given by Whitham on p. 461, Eq. 13.98
+Motivated by the experiments, we consider waves that originate from a region of constant depth, encounter an abrupt depth change, and continue into another region of constant depth.
 Thus, depth will be piecewise constant
@@
 
-The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$, which gives rise to the following quantities
-\begin{align}
-&c = \sqrt{g \depth}
-&&\mbox{\em local wave speed} \\
-&\lam = c/\freqp = \sqrt{g \depth} / \freqp
-&&\mbox{\em local peak wave length} \\
-&\etastd = \mean{\eta^2}^{1/2} 
-&&\mbox{\em standard deviation of displacement} \\
-\end{align}
-where $\mean{\cdot}$ indicates the mean of a quantity. 
-Note that the characteristic wave speed $c = c_{\pm}$ and wavelength $\lam = \lamupdn$ each take different values upstream and downstream of the ADC.
-We remark that experimental measurements indicate that $\etastd = \mean{\eta^2}^{1/2}$ is nearly the same on both values of the ADC.
+In the experiments, the randomized incoming wave-field is generated with a peak forcing frequency of $\freqp = 2$ Hz, which gives rise to the characteristic wavelength of $\lam = c/\freqp = \sqrt{g \depth} / \freqp$.
+Note that both the characteristic wave speed $c = c_{\pm}$ and wavelength $\lam = \lamupdn$ each take different values upstream and downstream of the ADC.
+We remark that experimental measurements indicate that $\etastd$ is nearly the same on both values of the ADC.
 Hence, we will not distinguish between upstream and downstream values of $\etastd$.
-%
-See Table \ref{paramtable} for a summary of important parameters.
+%See Table \ref{paramtable} for a summary of important parameters.
 
@@
 \ampscale = \pi^{1/2} \, \etastd \, , \qquad
-\lengthscale = \frac{\lamfac \lam}{2 \pi} \, , \qquad
-\timescale = \frac{\lamfac \lam}{2 \pi \freqp \depth}
+\lengthscale_{\pm} = \frac{\lamfac \lamupdn}{2 \pi} \, , \qquad
+\timescale_{\pm} = \frac{\lamfac \lamupdn}{2 \pi \freqp \dupdn}
 \end{align}
@@
 Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitably long timescale and is in line with previous normalizations \cite{johnson1997modern}.
-The scales $\lengthscale$ and $\timescale$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.
+The scales $\lengthscale = \lengthscale_{\pm}$ and $\timescale = \timescale_{\pm}$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.
 
@@
 C_2 = \frac{2 \pi^2 \delup}{3 \lamfac^2} 
-%C_2 = \frac{\pi^2 \delup}{6 \lamfac^2} 
 \end{align}
@@
 \end{equation}
+See Table \ref{paramtable} for a summary of these dimensionless parameters and their values in experiments.
 
@@
 Hence, all of the results discussed below would still hold qualitatively in this alternate formulation, with only slight quantitative modifications.
-
-%NOTE: Perhaps we should be more specific about comparing to the dimensionless parameters in our PNAS paper \cite{majda2019}, BUT the expressions for E0 and L0 given in the PNAS paper do not match with what I calculate comparing to the above.
-Plus, it seems like it could only add confusion.
 
@@
 \end{equation}
-where $\drat = \dratupdn$ changes value across the ADC. More explicitly, substituting \eqref{dratpw}, the separate upstream and downstream Hamiltonians are
+where $\drat = \dratupdn$ changes value across the ADC. More explicitly, substituting \eqref{dratpw} gives the separate upstream and downstream Hamiltonians as
 \begin{align}
@@
 We have chosen the sign convention above (which also happens to be consistent with Lax \cite{lax1975periodic}) so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.
+\nick{(Reword paragraph)}
 
@@
 \end{align}
-Since $u$ is real valued, we have $\uhat_{-k} = \uhat_{k}^*$ and since momentum vanishes $\uhat_0 = 0$.
+Since $u$ is real valued, $\uhat_{-k} = \uhat_{k}^*$, and since momentum vanishes $\uhat_0 = 0$.
 Next, consider the Galerkin truncation at wave number $\Lambda$
@@
 
-The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalization values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity gives
+The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalization values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity implies
 \begin{equation}
@@
 % The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
-As has been noted before, the (untruncated) KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}.
+As has been noted before, the untruncated KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}.
 The projected counterparts of these invariants, however, are not necessarily invariants of the truncated system.
-
+\nick{Check literature to refine this statement.}
 % Is it true that the only 3 known invariants of TKdV are the ones mentioned here: momentum, energy, and Hamiltonian.

### 2019-10-07 paper2_JNLS/paper.tex
NM

@@
 
-All parameters that appear in the theoretical model follow directly from experimental parameters, with the exception of the inverse temperature, $\thup$, of the incoming flow and the parameter $\lamfac$, which represents the number of characteristic wavelengths in the periodic domain.
+All parameters that appear in the theoretical model follow directly from experimental parameters, with the exception of the inverse temperature of the incoming flow, $\thup$, and the parameter $\lamfac$, which represents the number of characteristic wavelengths in the periodic domain.
 The parameter $\lamfac$ can be chosen in a sensible way by simply selecting an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved.
@@
 This insensitivity is a consequence of scaling $\lamfac$ with $\Lambda$ appropriately as discussed previously.
+\nick{(I should show the little calculation that justifies this)}
+
+\nick{PICK IT UP AROUND HERE}
 

### 2019-10-07 paper2_JNLS/paper.tex
NM

@@
 Using the alternative formulation would simply modify the power of $\drat$ in the second term of \eqref{dimlessKdV} by 1/4, i.e.~$\drat^{-7/4}$ instead of $\drat^{-3/2}$.
-Hence, all of the results discussed below would still hold qualitatively in this alternate formulation, with only slight quantitative modifications.
+Hence, all of the results discussed below would hold qualitatively in this alternate formulation, with only slight quantitative modifications.
 
@@
 \begin{align}
+\label{HamStruct}
 \partial_t{u} = \sympJ \vard{\Hupdn}{u}
@@
 \end{align}
-The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019}, which used the opposite sign in both $\sympJ$ and $\Ham$.
-With that sign convention, Majda et al.
-2019 found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}.
-We have chosen the sign convention above (which also happens to be consistent with Lax \cite{lax1975periodic}) so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.
-\nick{(Reword paragraph)}
+Here, we have chosen the sign of Lax \cite{lax1975periodic}.
+More recent work of Bajars et al.~\cite{bajars2013weakly} and Majda et al.~\cite{majda2019} used a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite.
+Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent.
+With their sign convention, Majda et al.~found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}.
+We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most orthodox form of statistical mechanics.
 
@@
 In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.017$.
+
+\nick{I think it makes sense to discuss $\lamfac$ here, in particular the simple scaling analysis of H2 and H3 that suggests we need to take $\lamfac$ scaling with $\Lambda$.
+Our desire is to have some sort of convergence as $\Lambda$ gets big, and this would be impossible if $\lamfac$ does not scale with $\Lambda$}
 

### 2019-10-08 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\dx}{\, dx}
+\newcommand{\CC}{{\mathbb{C}}}
 % Basic physical parameters and scales
@@
 \newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
+\newcommand{\Gz}{\Gibbs_0}
+\newcommand{\meanz}[1]{\mean{#1}_0}
+% Stuff
 \newcommand{\tavg}[1]{\overline{#1}}
@@
 Here, we have chosen the sign of Lax \cite{lax1975periodic}.
-More recent work of Bajars et al.~\cite{bajars2013weakly} and Majda et al.~\cite{majda2019} used a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite.
+More recent work of Bajars et al.~\cite{bajars2013weakly} and Majda et al.~\cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite.
 Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent.
 With their sign convention, Majda et al.~found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}.
-We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most orthodox form of statistical mechanics.
+We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.
 
@@
 % The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
+Thus, the dynamics of interest are confined to the unit hypersphere of $\En = 1$ in $\CC^{\Lambda}$.
 As has been noted before, the untruncated KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}.
@@
 
-In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced in \cite{abramov2003hamiltonian} for the Burgers-Hopf system.
+In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced by Abramov et al.~for the Burgers-Hopf system \cite{abramov2003}.
 Specifically, this ensemble is microcanonical in the energy, with a fixed energy value, and it is canonical in the Hamiltonian.
-By fixing the energy and hence confining to a compact set, this construction avoids the divergence at infinity that would occur for a simple canonical ensemble and hence yields a normalizable distribution.
+By fixing the energy and hence confining to the compact set of the unit hypersphere, this construction avoids the divergence at infinity that would occur for a simple canonical ensemble.
+It thus yields a normalizable distribution.
 This framework applies equally well to the truncated or untruncated system and hence we will not distinguish between the two.
@@
 \end{align}
-
-%Misc: The function that maximizes the Hamiltonian subject to the constraints of fixed energy and zero momentum is a traveling wave solution!
+Note that if $\theta = 0$, then the Gibbs measure is a {\em uniform} measure on the unit hypersphere in $\CC^{\Lambda}$, which we will denote as either $\Gz$ or simply $d\mu$.
 
@@
 %----------------------------------------------------------------------------%
+% Scaling
+%----------------------------------------------------------------------------%
+\subsection{Scaling analysis to set $\lamfac$}
+
+	Since the inverse temperature, $\theta$, is a key modeling parameter, it would be highly desirable to normalize the system such that its value does not depend sensitively on the truncation number, $\Lambda$, as $\Lambda$ grows large.
+That way, $\theta$ can be interpreted as a real physical parameter that can be linked to the experiments and whose value does not depend sensitively on where one chooses to truncate the system.
+The one modeling parameter that is left to be set is $\lamfac$, which represents the number of characteristic wavelengths in the periodic domain.
+In what follows, we will determine reasonable constraints on $\lamfac$.
+
+	The invariant measure enjoys the proportionality $ \Gibbs_{\Lambda} \propto \exp(-\theta_{\Lambda} \Ham_{\Lambda})$, where we have made explicit the dependence of all quantities on $\Lambda$.
+In particular, if $\Ham_{\Lambda}$ where to depend sensitively on $\Lambda$ in expectation, then $\theta_{\Lambda}$ would need to compensate in order to produce the same invariant measure.
+Hence, it would be desirable to scale the system in such a way that $\Ham_{\Lambda}$ does not depend sensitively on $\Lambda$, at least in expectation.
+To achieve this insensitivity, we will appeal to the uniform measure $\Gz$ and the idea of equipartition of energy \cite{abramov2003}, since these two concepts afford simple scaling estimates.
+
+Recall that $\Ham_{\Lambda}$ is composed of the cubic and quadratic components $\Hthree$ and $\Htwo$.
+Due to the odd symmetry of $\Hthree$, it is easy to see that $\meanz{\Hthree} = 0$.
+The quadratic component, however, requires closer inspection.
+Due to Parseval's identity, it is given by
+\begin{equation}
+\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx = 2 \pi \sum_{k=1}^{\Lambda} k^2 \abs{\uhat_k}^2
+\end{equation}
+Due to the constraint $\En[\uL] = 1$, the equipartioned microstate is given by 
+\begin{align}
+&\abs{\uhat_k}^2 \approx \frac{1}{2 \pi \Lambda}	 \qquad \mbox{\em for equipartition of energy}
+\end{align}
+Then the expected value of $\Htwo$ under the uniform measure is
+\begin{equation}
+\label{HtwoExpect}
+\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} \abs{k}^ 2 \mean{\abs{\uhat_k}^2}_0 \approx \frac{1}{3} \Lambda^2
+\end{equation}
+where we have used the identity for the Gauss-like sum
+\begin{equation}
+\sum_{k=1}^{n} \abs{k}^ 2 = \frac{1}{6} n(n+1)(2n+1) \approx n^3/3
+\end{equation}
+
+	Importantly, \eqref{HtwoExpect} shows that the expected value of $\Htwo$ grows like $\Lambda^2$, which appears to be problematic for obtaining independence as $\Lambda \to \infty$.
+However, $\Htwo$ enters the definition of $\Ham$ in produce with the coefficient $C_2$, which itself depends on $\lamfac$ as $C_2 \sim \lamfac^{-2}$.
+Hence, obtaining the desired asymptotic independence of $\Lambda$ simply requires that $\lamfac$ grow proportionally to $\Lambda$.
+
+\vsp{5}
+
+%----------------------------------------------------------------------------%
 % Results
@@
 In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.017$.
-
-\nick{I think it makes sense to discuss $\lamfac$ here, in particular the simple scaling analysis of H2 and H3 that suggests we need to take $\lamfac$ scaling with $\Lambda$.
-Our desire is to have some sort of convergence as $\Lambda$ gets big, and this would be impossible if $\lamfac$ does not scale with $\Lambda$}
 
@@
 If the experiments were conducted with the forcing given by the incoming Gibbs measure literally, then there would be no need for $\lamfac$.
-But the experimental forcing (which was decided before the theory) has a peak frequency of $\freqp = 2 Hz$ and so the factor $\lamfac$ makes $\freqp$ correspond to the dominant value in the spectrum in the theory (REWORD).
+But the experimental forcing (which was decided before the theory) has a peak frequency of $\freqp = 2 Hz$ and so the factor $\lamfac$ makes $\freqp$ correspond to the dominant value in the spectrum in the theory \nick{(REWORD)}.
 

### 2019-10-08 paper2_JNLS/paper.tex
M

@@
 The KdV equation provides the first correction to this behavior and describes dynamics that evolve over longer timescales.
-Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitably long timescale and is in line with previous normalizations \cite{johnson1997modern}.
+Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitable long timescale in line with previous normalizations \cite{johnson1997modern}.
 The scales $\lengthscale = \lengthscale_{\pm}$ and $\timescale = \timescale_{\pm}$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.
@@
 Thus, the dynamics of interest are confined to the unit hypersphere of $\En = 1$ in $\CC^{\Lambda}$.
-As has been noted before, the untruncated KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}.
-The projected counterparts of these invariants, however, are not necessarily invariants of the truncated system.
-\nick{Check literature to refine this statement.}
-% Is it true that the only 3 known invariants of TKdV are the ones mentioned here: momentum, energy, and Hamiltonian.
+%
+In summary, the TKdV system possesses three important invariants: momentum, energy, and hamiltonian.
+The untruncated KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}, but their truncated counterparts are not generally invariants of TKdV.
 
@@
 %----------------------------------------------------------------------------%
-\subsection{Scaling analysis to set $\lamfac$}
+\subsection{Scaling analysis to link model parameters to experiments}
 
@@
 The one modeling parameter that is left to be set is $\lamfac$, which represents the number of characteristic wavelengths in the periodic domain.
-In what follows, we will determine reasonable constraints on $\lamfac$.
+In what follows, we will determine reasonable constraints on $\lamfac$ that allow $\theta$ to be asymptotically independent of $\Lambda$.
 
-	The invariant measure enjoys the proportionality $ \Gibbs_{\Lambda} \propto \exp(-\theta_{\Lambda} \Ham_{\Lambda})$, where we have made explicit the dependence of all quantities on $\Lambda$.
+	The invariant measure exhibits the proportionality $ \Gibbs_{\Lambda} \propto \exp(-\theta_{\Lambda} \Ham_{\Lambda})$, where we have made explicit the dependence of all quantities on $\Lambda$.
 In particular, if $\Ham_{\Lambda}$ where to depend sensitively on $\Lambda$ in expectation, then $\theta_{\Lambda}$ would need to compensate in order to produce the same invariant measure.
@@
 	Importantly, \eqref{HtwoExpect} shows that the expected value of $\Htwo$ grows like $\Lambda^2$, which appears to be problematic for obtaining independence as $\Lambda \to \infty$.
-However, $\Htwo$ enters the definition of $\Ham$ in produce with the coefficient $C_2$, which itself depends on $\lamfac$ as $C_2 \sim \lamfac^{-2}$.
-Hence, obtaining the desired asymptotic independence of $\Lambda$ simply requires that $\lamfac$ grow proportionally to $\Lambda$.
+However, $\Htwo$ enters $\Ham$ in product with the coefficient $C_2$, which itself scales as $C_2 \sim \lamfac^{-2}$.
+Hence, to obtain asymptotic independence of $\Lambda$, it is necessary that $\lamfac$ grow proportionally to $\Lambda$, as was already argued on physical grounds in Section \ref{tKdVSec}.
+Thus, any of the choices $\lamfac = \Lambda, \Lambda/2, \Lambda/4$, discussed in that section would be valid.
 
 \vsp{5}
+
+The parameter $\lamfac$ can be chosen in a sensible way by simply selecting an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved.
+Hence, we will simply chose $\lamfac = \Lambda/2$ throughout, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths.
 
@@
 
-All parameters that appear in the theoretical model follow directly from experimental parameters, with the exception of the inverse temperature of the incoming flow, $\thup$, and the parameter $\lamfac$, which represents the number of characteristic wavelengths in the periodic domain.
-The parameter $\lamfac$ can be chosen in a sensible way by simply selecting an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved.
-Hence, we will simply chose $\lamfac = \Lambda/2$ throughout, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths.
-The incoming inverse temperature, $\thup$, however, is more difficult to pin down without using information from the experiments.
+At this point, all parameters that appear in the theoretical model have been linked directly to experimental parameters with the exception of the inverse temperature of the incoming flow, $\thup$.
 Our strategy is to use the outgoing skewness as the main diagnostic to determine a realistic range for $\thup$.
@@
 This insensitivity is a consequence of scaling $\lamfac$ with $\Lambda$ appropriately as discussed previously.
-\nick{(I should show the little calculation that justifies this)}
-
-\nick{PICK IT UP AROUND HERE}
+\\
+\nick{TO DO: Show a case with different N for contrast.}
 

### 2019-10-08 paper2_JNLS/paper.tex
NM

@@
 If $\lamfac = \Lambda$, then $\lam$ corresponds to the smallest resolved wavelength.
-Often, we will choose an intermediate value such as $\lamfac = \Lambda/2$ or $\Lambda/4$, so that the truncated system \eqref{TKdV} resolves scales that are both bigger and smaller than the characteristic value $\lam$.
+In contrast, if $\lamfac$ is chosen as an intermediate value, e.g.~$\lamfac = \Lambda/2$ or $\Lambda/4$, then the truncated system \eqref{TKdV} resolves scales that are both bigger and smaller than the characteristic value $\lam$.
 
@@
 
-	Since the inverse temperature, $\theta$, is a key modeling parameter, it would be highly desirable to normalize the system such that its value does not depend sensitively on the truncation number, $\Lambda$, as $\Lambda$ grows large.
+	Since the inverse temperature, $\theta$, is a key modeling parameter, it would be highly desirable to normalize the system such that the value of $\theta$ does not depend sensitively on the truncation number, $\Lambda$, as $\Lambda$ grows large.
 That way, $\theta$ can be interpreted as a real physical parameter that can be linked to the experiments and whose value does not depend sensitively on where one chooses to truncate the system.
@@
 However, $\Htwo$ enters $\Ham$ in product with the coefficient $C_2$, which itself scales as $C_2 \sim \lamfac^{-2}$.
-Hence, to obtain asymptotic independence of $\Lambda$, it is necessary that $\lamfac$ grow proportionally to $\Lambda$, as was already argued on physical grounds in Section \ref{tKdVSec}.
-Thus, any of the choices $\lamfac = \Lambda, \Lambda/2, \Lambda/4$, discussed in that section would be valid.
-
-\vsp{5}
-
-The parameter $\lamfac$ can be chosen in a sensible way by simply selecting an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved.
-Hence, we will simply chose $\lamfac = \Lambda/2$ throughout, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths.
+Hence, obtaining the desired asymptotic independence with respect to $\Lambda$ simply requires that $\lamfac$ grow proportionally to $\Lambda$, as was already argued on physical grounds in Section \ref{tKdVSec}.
+Thus, any of the choices $\lamfac = \Lambda$, $\Lambda/2$, or $\Lambda/4$, discussed in that section would be valid.
+In particular, if $\lamfac = \Lambda$, then the characteristic wavelength in experiments corresponds to the smallest resolved wavelength in the dynamical system.
+It is even more sensible to choose $\lamfac$ to be an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved.
+As default, we will choose $\lamfac = \Lambda/2$, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths.
 

### 2019-10-08 paper2_JNLS/paper.tex
NM

@@
 %----------------------------------------------------------------------------%
-\subsection{Scaling analysis to link model parameters to experiments}
+\subsection{Scaling analysis to link inverse temperature to experiments}
 
@@
 In particular, if $\lamfac = \Lambda$, then the characteristic wavelength in experiments corresponds to the smallest resolved wavelength in the dynamical system.
-It is even more sensible to choose $\lamfac$ to be an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved.
+It is perhaps more sensible to choose $\lamfac$ to be an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved.
 As default, we will choose $\lamfac = \Lambda/2$, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths.
@@
 
-Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depends on the incoming inverse temperature $\thup$.
+Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$.
 In line with experimental observations, the incoming skewness is small (always less than 0.2), while the outgoing wave skewness is much higher.

### 2019-10-08 code_rescaledModel/runSolver.m
NM

@@ -7,11 +7,11 @@ Drat = 0.24;    % depth ratio.
 % Simulation parameters
 Lambda = 16;    % Reference 16
 Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E3;       % Reference 1E4
+MC = 1E4;       % Reference 1E3 or 1E4 for tfin=10.
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
-nout = 100; % Reference 100
-tfin = 10.;  % Reference 10
+nout = 1; % Reference 100
+tfin = dt;  % Reference 10
 
 % Choose upstream or downstream
 gibup = false;

### 2019-10-08 paper2_JNLS/paper.tex
NM

@@
 
+	Figures \ref{transfig}(a)-(b) show the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $0 \le \thup \le 20$.
+In this figure, all model parameters are set by their experimental values, and we show the cases $\Lambda = 12, 16, 20$.
+In Fig.~\ref{transfig}(a), the parameter $\lamfac$ is fixed at $\lamfac = 8$.
+In this figure, the transfer function changes significantly with $\Lambda$.
+Figure~\ref{transfig}(b) shows the same but with $\lamfac = \Lambda/2$ scaled as argued in the previous section.
+In this figure, the three curves come much closer to one another.
+Thus scaling $\lamfac$ appropriately does not completely remove sensitivity of the transfer function to $\Lambda$, but it greatly mitigates the sensitivity.
+
 % Figure
@@
 \caption{
-Transfer function
+Analysis of the transfer function, $\thdn = \transf \left( \thup \right)$, and calibration of the inverse temperature.
+(a)--(b) The transfer function for three values of $\Lambda$ with either (a) $\lamfac = 8$ fixed or (b) $\lamfac = \Lambda/2$ scaled.
+Scaling $\lamfac$ with $\Lambda$ mitigates the sensitivity to $\Lambda$.
+(c) The incoming and outgoing skewness versus $\thup$ implied by the matching condition.
+In line with experiments, the skewness is enhanced significantly in the outgoing wave field.
 }
@@
 
-Figure \ref{transfig}(a) shows the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $\thup \in [0,20]$.
-In this figure, all model parameters are set by their experimental values, and we show the cases $\Lambda = 8, 16, 32$ with $\lamfac = \Lambda/2$ in each.
-As seen here, the transfer function is relatively insensitive to the truncation index since all three curves are relatively close to one another.
-This insensitivity is a consequence of scaling $\lamfac$ with $\Lambda$ appropriately as discussed previously.
-\\
-\nick{TO DO: Show a case with different N for contrast.}
+	Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$ for the case $\Lambda = 16$, $\lamfac = 8$.
+In line with experimental observations, the incoming skewness is small (always less than 0.2), while the outgoing wave skewness is much higher.
+Specifically, to capture the experimentally observed peak skewness of about 0.7, we can choose the value of $\thup$ in the range 10--15.
 
 \subsection{Statistical comparison between theory and experiments}
-
-Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$.
-In line with experimental observations, the incoming skewness is small (always less than 0.2), while the outgoing wave skewness is much higher.
-Specifically, to capture the experimentally observed peak skewness of about 0.7, we can choose the value of $\thup$ in the range 10--15.
 

### 2019-10-08 paper2_JNLS/paper.tex
NM

@@
 The KdV equation is a commonly excepted model for describing the propagation of unidirectional, shallow-water waves, and it accounts for weak nonlinearity and weak dispersion over long timescales and large spatial scales.
-We will perform Galerkin truncation of this system at wavenumber $\Lambda = $ (in the range 8 to 32) to obtain a finite-dimensional dynamical system that exhibits weak turbulence.
+We will perform Galerkin truncation of this system at wavenumber $\Lambda$ to obtain a finite-dimensional dynamical system that exhibits weak turbulence.
 We outline the Hamiltonian structure of both the continuous KdV and the truncated system.
@@
 \begin{align}
+\label{Gibbs}
 \Gupdn = Z_{\thupdn}^{-1} \exp(-\thupdn \Hupdn) \delta(\En - 1)
@@
 The quadratic component, however, requires closer inspection.
-Due to Parseval's identity, it is given by
+Due to Parseval's identity, $\Htwo$ is given by
 \begin{equation}
@@
 \label{HtwoExpect}
-\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} \abs{k}^ 2 \mean{\abs{\uhat_k}^2}_0 \approx \frac{1}{3} \Lambda^2
+\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} \abs{k}^ 2 \mean{\abs{\uhat_k}^2}_0 \sim \frac{1}{3} \Lambda^2
 \end{equation}
@@
 As default, we will choose $\lamfac = \Lambda/2$, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths.
+
+	We note that the experiments discussed here preceded the development of the theory, and hence had no intent of mimicking a Gibbs measure, such as that in \eqref{Gibbs}.
+We expect that if the driving featured in the experiments were designed to mimic the Gibbs measure, in particular the spectral decay, then it would be more straightforward to assign a value to the model parameter $\lamfac$.
+We, however, leave that task for future research due to the significant cost of performing an entirely new set of experiments compared to the relative ease and great value in re-analyzing existing experimental measurements in light of the new theoretical developments.
 
@@
 	Figures \ref{transfig}(a)-(b) show the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $0 \le \thup \le 20$.
-In this figure, all model parameters are set by their experimental values, and we show the cases $\Lambda = 12, 16, 20$.
-In Fig.~\ref{transfig}(a), the parameter $\lamfac$ is fixed at $\lamfac = 8$.
+Figure \ref{transfig}(a) shows the cases $\Lambda = $ 12, 16, and 20 with $\lamfac = 8$.
 In this figure, the transfer function changes significantly with $\Lambda$.
@@
 In this figure, the three curves come much closer to one another.
-Thus scaling $\lamfac$ appropriately does not completely remove sensitivity of the transfer function to $\Lambda$, but it greatly mitigates the sensitivity.
+Thus scaling $\lamfac$ appropriately greatly mitigates the sensitivity of the transfer function to $\Lambda$, though it does not completely remove the dependence.
 
@@
 
-	Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$ for the case $\Lambda = 16$, $\lamfac = 8$.
-In line with experimental observations, the incoming skewness is small (always less than 0.2), while the outgoing wave skewness is much higher.
-Specifically, to capture the experimentally observed peak skewness of about 0.7, we can choose the value of $\thup$ in the range 10--15.
+	Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$ for the case $(\Lambda, \lamfac) = (16, 8)$.
+In line with experimental observations, the incoming skewness is small, while the outgoing wave skewness is much higher.
+Specifically, to capture the experimentally observed peak skewness of about 0.6-0.9, we can choose the value of $\thup$ in the range 10--20.
 
@@
 
+	We now aim to compare the statistics measured in the experiments versus the theory.
+We focus on the representative set of experiments, detailed in Section BLAH and Table BLAH.
+In these experiments, the peak downstream skewness is precisely 0.83.
+According to Fig.~\ref{transfig}, we can achieve that value of skewness for the case $(\Lambda, \lamfac) = (16, 8)$ by setting $\thup = 17$, which gives $\thdn = 13.5$.
+We thus run the DNS with these model parameters (and with a sufficiently fine timestep, and number of samples, and to a final dimensionless time of 10).
+
+\vsp{5}
 Figure \ref{uhist} shows histograms of the displacement $u$ from experiments and from TKdV theory.

### 2019-10-08 code_rescaledModel/runSolver.m

@@ -7,19 +7,19 @@ Drat = 0.24;    % depth ratio.
 % Simulation parameters
 Lambda = 16;    % Reference 16
 Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E4;       % Reference 1E3 or 1E4 for tfin=10.
+MC = 1E3;       % Reference 1E3 or 1E4 for tfin=10.
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
-nout = 1; % Reference 100
-tfin = dt;  % Reference 10
+nout = 100; % Reference 100
+tfin = 10;  % Reference 10
 
 % Choose upstream or downstream
 gibup = false;
 if gibup == true
-    theta = 15;
+    theta = 17;
     gibd = 1.;
 else
-    theta = 13;
+    theta = 13.5;
     gibd = 0.24;
 end
 fi = 1; % fi = 1 for outgoing and 0 for incoming

### 2019-10-08 code_rescaledModel/matching/sampling_MH_scaled.m
NM

@@ -6,7 +6,7 @@ function [samp_H0,samp_H1,samp_u,enek] = sampling_MH_scaled(D0,D1,Nw, theta, MC)
 J=32; %Nw=10; theta=-30.; D0=.24; 
 %MC = 1E4;
 % model parameters from experiments
-epsi0 = 0.016;  % amplitude-to-depth ratio
+epsi0 = 0.017;  % amplitude-to-depth ratio
 del0 = 0.22;    % depth-to-wavelength ratio
 C2 = (2/3)*pi^2*del0/Nw^2;
 C3 = (3/2)*pi^(1/2)*epsi0/del0;

### 2019-10-08 code_rescaledModel/runSolver.m

@@ -7,11 +7,11 @@ Drat = 0.24;    % depth ratio.
 % Simulation parameters
 Lambda = 16;    % Reference 16
 Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E3;       % Reference 1E3 or 1E4 for tfin=10.
+MC = 1E0;       % Reference 1E3 or 1E4 for tfin=10.
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
-nout = 100; % Reference 100
-tfin = 10;  % Reference 10
+nout = 1; % Reference 100
+tfin = dt;  % Reference 10
 
 % Choose upstream or downstream
 gibup = false;

### 2019-10-09 code_rescaledModel/runSolver.m
NM

@@ -7,16 +7,16 @@ Drat = 0.24;    % depth ratio.
 % Simulation parameters
 Lambda = 16;    % Reference 16
 Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E0;       % Reference 1E3 or 1E4 for tfin=10.
+MC = 1E3;       % Reference 1E3 or 1E4 for tfin=10.
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
-nout = 1; % Reference 100
-tfin = dt;  % Reference 10
+nout = 100; % Reference 100
+tfin = 10;  % Reference 10
 
 % Choose upstream or downstream
-gibup = false;
+gibup = true;
 if gibup == true
-    theta = 17;
+    theta = 25;
     gibd = 1.;
 else
     theta = 13.5;

### 2019-10-09 paper2_JNLS/paper.tex
NM

@@
 \end{equation}
-We have included the $\lamfac$ in the numerator due to the differences in forcing spectra.
-If the experiments were conducted with the forcing given by the incoming Gibbs measure literally, then there would be no need for $\lamfac$.
-But the experimental forcing (which was decided before the theory) has a peak frequency of $\freqp = 2 Hz$ and so the factor $\lamfac$ makes $\freqp$ correspond to the dominant value in the spectrum in the theory \nick{(REWORD)}.
+In order to map the slope of the wavelength $\lam$ corresponding to $\freqp = 2$ Hz in the experiments onto the dominant spectral mode in the theory, the above conversion has not been corrected with the extra factor of $\lamfac$.
 

### 2019-10-09 code_rescaledModel/runSolver.m

@@ -5,9 +5,9 @@ epsi0 = 0.017;  % amplitude-to-depth ratio
 del0 = 0.22;    % depth-to-wavelength ratio
 Drat = 0.24;    % depth ratio.
 % Simulation parameters
-Lambda = 16;    % Reference 16
-Nw = Lambda/2;  % Reference Lambda/2
-MC = 1E3;       % Reference 1E3 or 1E4 for tfin=10.
+Lambda = 17;    % Reference 16
+Nw = 8;         % Reference Lambda/2
+MC = 1E3;       % Reference 1E3 or 5E3 for tfin=10.
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
 nout = 100; % Reference 100
@@ -16,10 +16,10 @@ tfin = 10;  % Reference 10
 % Choose upstream or downstream
 gibup = true;
 if gibup == true
-    theta = 25;
+    theta = 17.5;
     gibd = 1.;
 else
-    theta = 13.5;
+    theta = 0.;
     gibd = 0.24;
 end
 fi = 1; % fi = 1 for outgoing and 0 for incoming

### 2019-10-09 paper2_JNLS/paper.tex
NM

@@
 \section{The experiments}
+\label{experiments}
 
@@
 	We now aim to compare the statistics measured in the experiments versus the theory.
-We focus on the representative set of experiments, detailed in Section BLAH and Table BLAH.
+We focus on the representative set of experiments, detailed in Section \ref{experiments} and Table \ref{paramtable}.
 In these experiments, the peak downstream skewness is precisely 0.83.
@@
 \vsp{5}
-Figure \ref{uhist} shows histograms of the displacement $u$ from experiments and from TKdV theory.
-The experimental measurements have been normalized for comparison purposes.
-The comparison is fucking amazing!
-Offers convincing evidence that the theory works and the parameters have been matched as well as possible.
-(At some point, I need to mention the minor point about the forcing spectrum not being matched to the Gibbs measure because it came first).
+
+Figure \ref{uhist} shows histograms of the displacement $u$ from experiments and from TKdV theory for comparison.
+The experiment histograms representative exactly the same data as shown in Fig.~\ref{ExpDiagStats}, only converted to the dimensionless variables.
+As mentioned previously, the upstream histogram is nearly symmetric, while downstream histogram skews strongly towards positive displacement.
+The TKdV histograms do a remarkable job of recovering these features, appearing to produce nearly exactly the same distributions, both in shape and in scale.
+%
+This comparison offers convincing evidence that the theory works and the parameters have been matched as well as possible.
 

### 2019-10-09 code_rescaledModel/runSolver.m

@@ -16,13 +16,13 @@ tfin = 10;  % Reference 10
 % Choose upstream or downstream
 gibup = true;
 if gibup == true
-    theta = 17.5;
+    theta = 20.;
     gibd = 1.;
 else
     theta = 0.;
     gibd = 0.24;
 end
-fi = 1; % fi = 1 for outgoing and 0 for incoming
+fi = 0; % fi = 1 for outgoing and 0 for incoming
 
 
 % Run the simulation.

### 2019-10-09 paper2_JNLS/paper.tex
NM

@@
 	We now aim to compare the statistics measured in the experiments versus the theory.
-We focus on the representative set of experiments, detailed in Section \ref{experiments} and Table \ref{paramtable}.
-In these experiments, the peak downstream skewness is precisely 0.83.
-According to Fig.~\ref{transfig}, we can achieve that value of skewness for the case $(\Lambda, \lamfac) = (16, 8)$ by setting $\thup = 17$, which gives $\thdn = 13.5$.
+We focus on the representative set of experiments detailed in Table \ref{paramtable}.
+In these experiments, the peak downstream skewness measured was 0.83.
+We can then use Fig.~\ref{transfig}(c) as a guide to select the upstream inverse temperature $\thup$.
+With $(\Lambda, \lamfac) = (16, 8)$ fixed, the figure indicates a value of $\thup = 17$ to achieve the desired skewness.
 We thus run the DNS with these model parameters (and with a sufficiently fine timestep, and number of samples, and to a final dimensionless time of 10).
 
-\vsp{5}
-
-Figure \ref{uhist} shows histograms of the displacement $u$ from experiments and from TKdV theory for comparison.
-The experiment histograms representative exactly the same data as shown in Fig.~\ref{ExpDiagStats}, only converted to the dimensionless variables.
-As mentioned previously, the upstream histogram is nearly symmetric, while downstream histogram skews strongly towards positive displacement.
-The TKdV histograms do a remarkable job of recovering these features, appearing to produce nearly exactly the same distributions, both in shape and in scale.
+	Figure \ref{uhist} shows displacement histograms from these TKdV simulations in comparison to the experiment histograms.
+The later representative exactly the same data as shown in Fig.~\ref{ExpDiagStats}, only converted to the dimensionless variables.
+The TKdV simulations do a remarkable job of capturing both the symmetric upstream statistics and the skewed downstream statistics.
+Amazingly, the simulations reproduce the measured distributions in great detail, both in the shape of the distribution and the scale.
 %
@@
 \caption{ 
-Amazing!
+Comparison of displacement statistics from experiment and theory.
+Upon calibrating the inverse temperature $\thup$, the TKdV simulations recover the experimentally measured statistics in remarkable detail.
+In particular, they exhibit the transition from a nearly symmetric distribution upstream to a highly skewed distribution downstream.
 }
@@
 \caption{
-Slopes
+Comparison of slope statistics.
+The TKdV simulations show similar slope distributions upstream, but visual distributions in the downstream distributions are visible.
+While the simulations recover the spread of the downstream slope statistics, they do not capture the asymmetry seen in the experiments.
 }
@@
 \caption{
-Trajectories
+Sample upstream and downstream solution trajectories from the same ensemble TKdV simulations featured in Figs.~\ref{uhist} and \ref{slopehist}.
+The upstream solutions exhibit only leftward propagating waves and a high degree of regularity, while the downstream solutions exhibit both left and rightward moving waves along with more intermittency.
 }

### 2019-10-09 paper2_JNLS/paper.tex
NM

@@
 
-\title{Anomalous waves triggered by abrupt depth changes: experimental measurements and deterministic-statistical truncated KdV theory}
+\title{Anomalous waves triggered by abrupt depth changes: experiments and truncated KdV theory}
 
+%experimental measurements and deterministic-statistical truncated KdV theory
 % ORIGINAL: The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes
@@
 First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019}, with identical powers of $\drat$ in \eqref{dimlessKdV} but with different expressions for the other dimensionless parameters.
-These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments.
-Second, an alternate formulation of the variable-depth KdV equation exists in which it is conjectured that the product $\depth^{1/4} \eta$, rather than $\eta$, varies continuously across the ADC \cite{johnson1997modern}.
-Of course, that assumption implies a discontinuity in surface displacement, which, though small, is physically unrealistic.
-We have chosen to enforce continuity of surface displacement in the above on the basis of physical realism.
-Using the alternative formulation would simply modify the power of $\drat$ in the second term of \eqref{dimlessKdV} by 1/4, i.e.~$\drat^{-7/4}$ instead of $\drat^{-3/2}$.
-Hence, all of the results discussed below would hold qualitatively in this alternate formulation, with only slight quantitative modifications.
+These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison with experiments.
+Second, an alternate formulation of the variable-depth KdV equation has been proposed in which the product $\depth^{1/4} \eta$, rather than $\eta$, is conjectured to vary continuously across the ADC \cite{johnson1997modern}.
+Of course, that assumption implies a discontinuity in surface displacement, which, though perhaps small, would be physically unrealistic.
+We have chosen to enforce continuity of surface displacement on the basis of physical realism.
+Furthermore, our {\em direct experimental measurements} of $\etastd$ give no indication of a significant change across the ADC, thus supporting the formulation used here.
+We note, however, that the only modification resulting from the alternate formulation would be in the power of $\drat$ in the second term of \eqref{dimlessKdV}: the power $\drat^{-3/2}$ would become $\drat^{-7/4}$.
+Thus, in this alternate formulation, the second term in \eqref{dimlessKdV} would be scaled by a negative power of $\drat$ and the third term by exactly the same positive power of $\drat$.
+Hence, the two formulations are qualitatively very similar with only slight quantitative differences expected.
 

### 2019-10-09 paper2_JNLS/paper.tex
NM

@@
 
-	We note that the experiments discussed here preceded the development of the theory, and hence had no intent of mimicking a Gibbs measure, such as that in \eqref{Gibbs}.
-We expect that if the driving featured in the experiments were designed to mimic the Gibbs measure, in particular the spectral decay, then it would be more straightforward to assign a value to the model parameter $\lamfac$.
+	We note that the experiments discussed here preceded the development of the theory, and hence had no intent of mimicking a Gibbs measure in the wave forcing.
+We expect that if this forcing were designed to mimic the Gibbs measure, in particular the spectral decay, then it would be more straightforward to assign a value to the model parameter $\lamfac$.
 We, however, leave that task for future research due to the significant cost of performing an entirely new set of experiments compared to the relative ease and great value in re-analyzing existing experimental measurements in light of the new theoretical developments.
@@
 We focus on the representative set of experiments detailed in Table \ref{paramtable}.
-In these experiments, the peak downstream skewness measured was 0.83.
+In these experiments, the peak downstream skewness is measured to be $\max \skw(\uup) = 0.83$.
 We can then use Fig.~\ref{transfig}(c) as a guide to select the upstream inverse temperature $\thup$.
@@
 %
-This comparison offers convincing evidence that the theory works and the parameters have been matched as well as possible.
+This comparison offers convincing evidence that the theory works and the parameters have been matched as well as possible. \nick{Reword}
 
@@
 In order to map the slope of the wavelength $\lam$ corresponding to $\freqp = 2$ Hz in the experiments onto the dominant spectral mode in the theory, the above conversion has not been corrected with the extra factor of $\lamfac$.
+
+	These slope histograms show intriguing similarities and differences.
+First, the upstream slope distribution is captured well by the TKdV simulations, both in its nearly symmetric shape and its scale.
+We note that, while the standard deviation of displacement was input into the TKdV theory, the slope standard deviation was not.
+Downstream, both the experimental and theoretical slope distributions  spread significantly.
+Interestingly, in the experiments, it is not the standard deviation of slope that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but the excess kurtosis, which jumps from a value of 0.46 to 3.4.
+Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50.
+Thus, the jump in kurtosis, while not as extreme as that measured in experiments, is quite significant in the theory.
+These elevated levels of kurtosis give the distributions the look of being stretched in the tails and pointy near the centers (especially in the experiments).
+
+	Distinct differences in the shapes of the downstream slope distributions, however, are visible.
+While the theoretical distributions remain symmetric, the experimental distributions skew towards negative slope.
+As mentioned earlier, this skewness is consistent with a right-moving wave of steep leading surface (i.e.~negative slope), which is a phenomenon apparently not captured by the TKdV simulations.
+Finally, it should be commented that the experimental measurements of slope involve numerical differentiation of the free surface extracted from optical images, which naturally introduces error.
+Furthermore, high-order statistics, such as kurtosis, can be very sensitive to noise and measurement error.
+Hence, the experimental statistics of slope, while they are valuable, should be regarded with some level of caution.
 

### 2019-10-09 paper2_JNLS/paper.tex
NM

@@
 	The paddle amplitude in Fig.~\ref{ExpDiagStats} is $\Dphi = 1.38^{\circ}$, and this value was varied systematically in the range $\Dphi = 0.125^{\circ}$--$2^{\circ}$ to probe the different regimes of wave behavior, from linear to strongly nonlinear waves.
-Figure \ref{ExpSpatialStats} shows spatial statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ for six different driving amplitudes (see legend).
-For each quantity, we show the standard deviation, the skewness, and the (excess) kurtosis as they vary in the vicinity of the ADC, located at $x=0$.
+Figure \ref{ExpSpatialStats} shows long-time statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ as they vary in space for six different driving amplitudes (see legend).
+We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
+\begin{align}
+& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
+& \std(q) = q_{std}= \sqrt{\var(q)}
+&&\mbox{\em standard deviation} \\
+& \skw(q) = \frac{\mean{q^3}} {q_{std}^3}	
+&&\mbox{\em skewness} \\
+& \kurt(q) = \frac{\mean{q^4}} {q_{std}^4} - 3
+&&\mbox{\em (excess) kurtosis}
+\end{align}
+where $\mean{}$ indicates a mean (here a long-time mean). Hereafter, we will simply refer to the excess kurtosis as kurtosis.
+
+	Figure \ref{ExpSpatialStats} shows how these statistics vary in the vicinity of the ADC, located at $x = 0$, for both displacement and slope.
 First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves.
@@
 %
-This comparison offers convincing evidence that the theory works and the parameters have been matched as well as possible. \nick{Reword}
+This comparison offers convincing evidence that the theory works and the parameters have been matched as well as possible. \nick{Reword and expand}
 
@@
 Downstream, both the experimental and theoretical slope distributions  spread significantly.
-Interestingly, in the experiments, it is not the standard deviation of slope that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but the excess kurtosis, which jumps from a value of 0.46 to 3.4.
+Interestingly, in the experiments, it is not the standard deviation of slope that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but the excess kurtosis that jumps from a value of 0.46 to 3.4.
 Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50.
@@
 
+Next, to complement the statistical comparison, we investigate some individual solution trajectories from the TKdV simulations.
+
+\vsp{5}
 We show individual trajectories in the next figure for upstream and downstream.

### 2019-10-09 paper2_JNLS/paper.tex
NM

@@
 \begin{align}
-& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
-& \std(q) = q_{std}= \sqrt{\var(q)}
+%& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
+& \std(q) = q_{std}= \sqrt{\mean{q^2}}
 &&\mbox{\em standard deviation} \\
-& \skw(q) = \frac{\mean{q^3}} {q_{std}^3}	
+& \skw(q) = {\mean{q^3}} / {q_{std}^3}	
 &&\mbox{\em skewness} \\
-& \kurt(q) = \frac{\mean{q^4}} {q_{std}^4} - 3
+& \kurt(q) = {\mean{q^4}} / {q_{std}^4} - 3
 &&\mbox{\em (excess) kurtosis}

### 2019-10-09 paper2_JNLS/paper.tex
NM

@@
 \caption{Wave statistics as they vary in space for several different experiments of different driving amplitudes (see legend).
-$\etastd$ does not vary significantly crossing the ADC (a), but skewness (b) and kurtosis (c) show a strong response to the ADC.
+(a)--(c) Standard deviation, skewness, and kurtosis of the surface displacement $\eta$.
+(d)--(f) The same for surface slope $\eta_x$.
+A reference wave amplitude is set by $\etastd$ which does not vary significantly crossing the ADC.
+Skewness and kurtosis of $\eta$, however, show a strong response to the ADC.
+The surface slope measurements are negatively skewed and also show a large kurtosis downstream of the ADC.
 }

### 2019-10-09 paper2_JNLS/paper.tex
NM

@@
 \begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/ExpDiagStats.pdf}
+\includegraphics[width = 0.80 \linewidth]{Figs/ExpDiagStats.pdf}
 \caption{
@@
 Upstream depth		& $\dup$					& 12.5 cm 	& 12.5 cm \\
-Downstream depth		& $\ddn$					& RANGE		& 3 cm \\
+Downstream depth		& $\ddn$					& 2.2--5.3 cm	& 3 cm \\
 Upstream wavelength	& $\lamup = \sqrt{g \dup}/f_p$	& 55 cm 		& 55 cm \\
-Downstream wavelength	& $\lamdn = \sqrt{g \ddn}/f_p$	& RANGE		& 27 cm \\
+Downstream wavelength	& $\lamdn = \sqrt{g \ddn}/f_p$	& 23--36 cm	& 27 cm \\
 %
 Amplitude-to-depth ratio	& $\epsup = \etastd / \dup$	&0.0024--0.024	& 0.017 \\
-Depth-to-wavelength ratio	& $\delup = \dup / \lamup$	& RANGE		& 0.22 \\
-Depth ratio			& $\dratdn = \ddn/\dup$		& RANGE		& 0.24 \\
+Depth-to-wavelength ratio	& $\delup = \dup / \lamup$	& 0.22		& 0.22 \\
+Depth ratio			& $\dratdn = \ddn/\dup$		& 0.18--0.42	& 0.24 \\
 \end{tabular}
@@
 We show histograms of the surface slope in Fig.~\ref{slopehist}, from both the experiments (top) and theory (bottom).
-The derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.
+The derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of the normalized derivative $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.
 Therefore, in Fig.~\ref{slopehist}, we convert the theoretical calculated values to physical slope via

### 2019-10-10 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\transf}{\mathcal{F}}
+\newcommand{\Nsamp}{N_s}
+\newcommand{\sumsamp}{\sum_{i=1}^{\Nsamp}}
 
@@
 \author{
+M.~N.~J.~Moore\thanks{Florida State University}, 
 C.~Tyler Bolles\thanks{University of Michigan},
 Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
-M.~N.~J.~Moore\thanks{Florida State University}, 
 Di Qi\thanks{Courant Institute of Mathematical Sciences} }
@@
 
-The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized incoming wave field, the paddle angle $\phi$ is specified by a precomputed psuedo-random signal
+The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized wave field, the paddle angle $\phi$ is specified by a psuedo-random signal
 \begin{align}
@@
 Amplitude-to-depth ratio	& $\epsup = \etastd / \dup$	&0.0024--0.024	& 0.017 \\
-Depth-to-wavelength ratio	& $\delup = \dup / \lamup$	& 0.22		& 0.22 \\
+Depth-to-wavelength ratio	& $\delup = \dup / \lamup$	& 0.23		& 0.23 \\
 Depth ratio			& $\dratdn = \ddn/\dup$		& 0.18--0.42	& 0.24 \\
@@
 %----------------------------------------------------------------------------%
+% Numerical methods
+%----------------------------------------------------------------------------%
+\subsection{Deterministic simulations of TKdV}
+
+\nick{Include a short description of the numerical method for simulation TKdV: a symplectic integrator.}
+
+\subsection{Sampling from the Gibbs measures}
+
+\nick{Include a short description of the sampling method}
+
+\begin{equation}
+\mean{Q} = \frac{\int Q \exp\left(-\theta \Ham \right) d\mu}{ \int \exp\left(-\theta \Ham \right) d\mu}
+\approx \frac{\sumsamp Q_i \exp(-\theta \Ham_i)} {\sumsamp \exp(-\theta \Ham_i)}
+\end{equation}
+
+%----------------------------------------------------------------------------%
 % Results
@@
 Unless stated otherwise, all parameters used in the theory are taken directly from their experimental values listed in Table \ref{paramtable}.
-In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.017$.
+%In particular, we set $\delup = 0.23$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.017$.
 
@@
 	Figures \ref{transfig}(a)-(b) show the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $0 \le \thup \le 20$.
-Figure \ref{transfig}(a) shows the cases $\Lambda = $ 12, 16, and 20 with $\lamfac = 8$.
+Figure \ref{transfig}(a) shows the cases $\Lambda = $ 12, 16, and 20 with $\lamfac = 8$ fixed in each.
 In this figure, the transfer function changes significantly with $\Lambda$.
-Figure~\ref{transfig}(b) shows the same but with $\lamfac = \Lambda/2$ scaled as argued in the previous section.
+Figure~\ref{transfig}(b) shows the same but with the scaling $\lamfac = \Lambda/2$ that was argued in the previous section.
 In this figure, the three curves come much closer to one another.
@@
 In line with experimental observations, the incoming skewness is small, while the outgoing wave skewness is much higher.
-Specifically, to capture the experimentally observed peak skewness of about 0.6-0.9, we can choose the value of $\thup$ in the range 10--20.
+Specifically, to capture the experimentally observed peak skewness range of 0.6-0.9 seen for the larger amplitudes in Fig.~\ref{ExpSpatialStats}(b), requires selecting $\thup$ in the range 10--20.
 
@@
 
-	We now aim to compare the statistics measured in the experiments versus the theory.
-We focus on the representative set of experiments detailed in Table \ref{paramtable}.
-In these experiments, the peak downstream skewness is measured to be $\max \skw(\uup) = 0.83$.
-We can then use Fig.~\ref{transfig}(c) as a guide to select the upstream inverse temperature $\thup$.
-With $(\Lambda, \lamfac) = (16, 8)$ fixed, the figure indicates a value of $\thup = 17$ to achieve the desired skewness.
-We thus run the DNS with these model parameters (and with a sufficiently fine timestep, and number of samples, and to a final dimensionless time of 10).
+	We now aim to compare the wave statistics measured in experiments against those that emerge from the TKdV theory.
+Throughout, we will focus on the representative set of experiments detailed in Table \ref{paramtable}.
+In these experiments, the peak downstream skewness was measured to be 0.83.
+We can then use Fig.~\ref{transfig}(c) as a guide to select the appropriate upstream inverse temperature $\thup$.
+With $(\Lambda, \lamfac) = (16, 8)$ fixed, the figure indicates a value of $\thup = 17$ achieves the desired outgoing skewness.
+We thus run the deterministic TKdV simulations with model parameters $\Lambda = 16$, $\lamfac = 8$, $\thup = 17$ and with $\epsup$, $\delup$, and $\dratdn$ set to the representative values in Table \ref{paramtable}.
+We run the simulations to a final dimensionless time of $t_{f} = 10$, with $dt = 5 \times 10^{-4}$ and with $10^3$ trajectories sampled from the upstream Gibbs measure $\Gup$.
+
+	We thus run the with these model parameters (and with a sufficiently fine timestep, and number of samples, and to a final dimensionless time of 10).
+
+	\vsp{5}
 

### 2019-10-10 paper2_JNLS/paper.tex
NM

@@
 
-	We thus run the with these model parameters (and with a sufficiently fine timestep, and number of samples, and to a final dimensionless time of 10).
-
+	In Fig.~\ref{uhist} we show the displacement distributions that were measured in the experiments (top row) versus those that emerge from a mixed long-time/ensemble average of the TKdV simulations (bottom row).
 	\vsp{5}

### 2019-10-10 paper2_JNLS/paper.tex
NM

--- 2019-10-10 paper2_JNLS/paper.tex
@@
 Both signals exhibit a combination of periodic and random behavior, with the dominant oscillations corresponding to the peak forcing frequency of 2 Hz.
-The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e).
+The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e) on a semi-log scale.
 The upstream measurements are symmetrically distributed about the mean, $\eta = 0$.
@@
 We thus run the deterministic TKdV simulations with model parameters $\Lambda = 16$, $\lamfac = 8$, $\thup = 17$ and with $\epsup$, $\delup$, and $\dratdn$ set to the representative values in Table \ref{paramtable}.
-We run the simulations to a final dimensionless time of $t_{f} = 10$, with $dt = 5 \times 10^{-4}$ and with $10^3$ trajectories sampled from the upstream Gibbs measure $\Gup$.
+We run the simulations to a sufficiently long dimensionless time of $t_{f} = 10$, with $dt = 5 \times 10^{-4}$ and with $10^3$ trajectories sampled from the upstream Gibbs measure $\Gup$.
 
-	We thus run the with these model parameters (and with a sufficiently fine timestep, and number of samples, and to a final dimensionless time of 10).
-
-	\vsp{5}
-
-	Figure \ref{uhist} shows displacement histograms from these TKdV simulations in comparison to the experiment histograms.
-The later representative exactly the same data as shown in Fig.~\ref{ExpDiagStats}, only converted to the dimensionless variables.
-The TKdV simulations do a remarkable job of capturing both the symmetric upstream statistics and the skewed downstream statistics.
-Amazingly, the simulations reproduce the measured distributions in great detail, both in the shape of the distribution and the scale.
-%
-This comparison offers convincing evidence that the theory works and the parameters have been matched as well as possible. \nick{Reword and expand}
+	In Fig.~\ref{uhist} we show the displacement distributions that were measured in the experiments (top row) versus those that emerge from the TKdV simulations (bottom row).
+The experimental histograms represent exactly the same data depicted in Fig.~\ref{ExpDiagStats}, only converted to dimensionless displacement, $u$, for comparison against theory.
+Meanwhile, the TKdV distributions are extracted  from the long-time history of several trajectories sampled from the Gibbs ensemble, i.e.~mixed long-time/ensemble histograms.
+All histograms are shown on a semi-log scale to facilitate comparison of the tails, where extreme events lie.
 
@@
 \caption{ 
-Comparison of displacement statistics from experiment and theory.
-Upon calibrating the inverse temperature $\thup$, the TKdV simulations recover the experimentally measured statistics in remarkable detail.
+Comparison of displacement histograms from experiment and theory.
+Upon calibrating the inverse temperature $\thup$, the TKdV simulations recover the experimentally measured distributions in remarkable detail.
 In particular, they exhibit the transition from a nearly symmetric distribution upstream to a highly skewed distribution downstream.
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-% As of now the simulations use thup=15, thdn=13.
 
-We show histograms of the surface slope in Fig.~\ref{slopehist}, from both the experiments (top) and theory (bottom).
-The derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of the normalized derivative $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.
-Therefore, in Fig.~\ref{slopehist}, we convert the theoretical calculated values to physical slope via
+	The comparison between theory and experiment is striking.
+As seen in the figure, both the experiments and theory show a transition from a nearly symmetric upstream distribution, to a highly skewed distribution downstream.
+The TKdV theory not only captures this transition, but also recovers the shape of the resulting downstream distribution in remarkable detail.
+Nearly every feature that can be compared --- the decay rate of the tail, the position and value of the peak, the rapid cutoff for negative $u$ --- matches surprisingly well.
+This comparison offers compelling visual evidence for: (a) the predictive power of the TKdV framework, and (b) the successful calibration of the incoming inverse temperature.
+
+% Remarkable, spectacular, splendid, stupendous, striking, amazing, exquisite
+% Compelling, convincing
+
+	Next, we aim to make the same comparison for the slope of the free surface.
+We note that the derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.
+We therefore convert the theoretical calculated slope values back to physical slope via
 \begin{equation}
@@
 \end{equation}
-In order to map the slope of the wavelength $\lam$ corresponding to $\freqp = 2$ Hz in the experiments onto the dominant spectral mode in the theory, the above conversion has not been corrected with the extra factor of $\lamfac$.
+So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode,  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.
 
-	These slope histograms show intriguing similarities and differences.
+% Figure
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.99 \linewidth]{Figs/slopehist.pdf}
+\caption{
+Comparison of slope histograms.
+The TKdV simulations show similar slope distributions upstream, but visual distributions in the downstream distributions are visible.
+While the simulations recover the spread of the downstream slope statistics, they do not capture the asymmetry seen in the experiments.
+}
+\label{slopehist}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+	Figure \ref{slopehist} shows the distributions of surface slope that result from experiments (top) and theory (bottom).
+These histograms show intriguing similarities and differences.
 First, the upstream slope distribution is captured well by the TKdV simulations, both in its nearly symmetric shape and its scale.
@@
 Downstream, both the experimental and theoretical slope distributions  spread significantly.
-Interestingly, in the experiments, it is not the standard deviation of slope that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but the excess kurtosis that jumps from a value of 0.46 to 3.4.
+Interestingly, in the experiments, it is not the standard deviation that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but instead the excess kurtosis, which jumps from a value of 0.46 upstream to 3.4 downstream.
 Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50.
@@
 Furthermore, high-order statistics, such as kurtosis, can be very sensitive to noise and measurement error.
-Hence, the experimental statistics of slope, while they are valuable, should be regarded with some level of caution.
-
-% Figure
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[!ht]
-\begin{center}
-\includegraphics[width = 0.99 \linewidth]{Figs/slopehist.pdf}
-\caption{
-Comparison of slope statistics.
-The TKdV simulations show similar slope distributions upstream, but visual distributions in the downstream distributions are visible.
-While the simulations recover the spread of the downstream slope statistics, they do not capture the asymmetry seen in the experiments.
-}
-\label{slopehist}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-% As of now the simulations use thup=15, thdn=13.
+Hence, the experimental statistics of slope, while highly valuable, should be regarded with some level of caution.
 

### 2019-10-10 paper2_JNLS/paper.tex
NM

@@
 Depth-to-wavelength ratio	& $\delup = \dup / \lamup$	& 0.23		& 0.23 \\
-Depth ratio			& $\dratdn = \ddn/\dup$		& 0.18--0.42	& 0.24 \\
+Depth ratio			& $\dratdn = \ddn/\dup$		& 0.18--0.42	& 0.24
 \end{tabular}
@@
 
+%----------------------------------------------------------------------------%
+% Calibration of inverse temperature
+%----------------------------------------------------------------------------%
 \subsection{Calibration of the inverse temperature}
@@
 
+%----------------------------------------------------------------------------%
+% Comparison of distributions
+%----------------------------------------------------------------------------%
 \subsection{Statistical comparison between theory and experiments}
@@
 
-	The comparison between theory and experiment is striking.
+	The comparison between theory and experiment in Fig.~\ref{uhist} is striking.
 As seen in the figure, both the experiments and theory show a transition from a nearly symmetric upstream distribution, to a highly skewed distribution downstream.
@@
 
+%----------------------------------------------------------------------------%
+% Wave dynamics
+%----------------------------------------------------------------------------%
 \subsection{Wave dynamics}
 
-Next, to complement the statistical comparison, we investigate some individual solution trajectories from the TKdV simulations.
+	As a complement to the above statistical comparison, we now examine the wave dynamics of a few individual trajectories from the TKdV simulations.
+Figure \ref{trajectories} shows example upstream and downstream solution trajectories from the same TKdV simulations that were used to produce the histograms in Figs.~\ref{uhist}--\ref{slopehist}.
+The dimensionless displacement, $u$, is represented by color in the domain $(x,t) \in [-\pi,\pi] \times [-10,10]$, where the ADC is encountered at $t=0$.
 
-\vsp{5}
-We show individual trajectories in the next figure for upstream and downstream.
-In the upstream flow, due to the coefficients, the nonlinearity is very weak.
-Hence, wave propagation is dominated by linear dispersion.
-Indeed, with weak nonlinearity, \eqref{dimlessKdV} is approximated by
-\begin{equation}
-u_t + C_2 \drat^{1/2} \, u_{xxx} = 0
-\qquad \text{for } x \in [-\pi,\pi] \\
-\end{equation}
-which implies the dispersion relation $\omega_k = -C_2 k^3$ or for phase velocity $c_k = -C_2 k^2$, which implies left-going waves \cite{majdaqi2019}.
-% Further details, see page 6 of Majda Qi, J Stat Phys.
+	Visual differences between the upstream and downstream dynamics are apparent in Fig.~\ref{trajectories}.
+The upstream trajectory shows several waves of modest amplitude all propagating leftward.
+Here, the magnitude of the positive and negative displacements are comparable.
+This is in stark contrast to the downstream dynamics, which feature fewer waves of larger amplitude propagating in either direction.
+In particular, one large wave is seen to propagate from the lower-left corner of the figure to the upper right.
+A bias towards positive surface displacement is visually apparent in these downstream dynamics.
 
@@
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+These observations can be rationalized with some simple analysis of the structure of TKdV.
+Upstream of the ADC, since $\drat = 1$, the nonlinear effects are relatively weak and the dynamics are dominated by dispersion.
+Linearizing Eq.~\eqref{TKdV} and introducing the ansatz $u_k = e^{i(kx - \omega_k t)}$ produces the dispersion relation for angular frequency $\omega_k = -C_2 k^3$, or for phase velocity $c_k = \omega_k/k = -C_2 k^2$ \cite{majdaqi2019}.
+In particular, the phase velocity is strictly negative, implying that waves can only propagate leftwards, as is consistent with Fig.~\ref{trajectories}(a).
+% Further details, see page 6 of Majda Qi, J Stat Phys.
+%$C_3 = 0.20$ and $C_2 = 0.024$.
+
+Downstream of the ADC, however, $\drat = 0.24$, which significantly amplifies nonlinearity while suppressing dispersion in Eq.~\ref{TKdV} and allows waves to propagate in either direction.
+In particular, the dispersionless limit of KdV is the Burgers-Hopf equation, in which the nonlinear advection is proportional in both amplitude and direction to the value of $u$.
+Hence, positive displacements would be expected to move rightwards, as is seen in Fig.~\ref{trajectories}(b).
+
+%We remind the reader that the KdV framework tracks the long-time evolution of waves in the reference frame moving with the linear speed $c = \sqrt{g \depth}$.
+
+% Timescale of 10 is about right to see waves propagate through the entire periodic domain.
+This however would be a very long experimental time scale.
+So what gives?
+We answer that question with skewness and kurtosis plots which evolve over a much shorter timescale.
+
+\vsp{5}
+
+Next stuff
 

### 2019-10-10 code_rescaledModel/runSolver.m

@@ -2,10 +2,10 @@ clear all; close all;
 
 % Model parameters from experiments
 epsi0 = 0.017;  % amplitude-to-depth ratio
-del0 = 0.22;    % depth-to-wavelength ratio
+del0 = 0.23;    % depth-to-wavelength ratio
 Drat = 0.24;    % depth ratio.
 % Simulation parameters
-Lambda = 17;    % Reference 16
+Lambda = 16;    % Reference 16
 Nw = 8;         % Reference Lambda/2
 MC = 1E3;       % Reference 1E3 or 5E3 for tfin=10.
 % time step parameters
@@ -27,8 +27,8 @@ fi = 0; % fi = 1 for outgoing and 0 for incoming
 
 % Run the simulation.
 tic;
-C2 = (2/3)*pi^2*del0/Nw^2;
-C3 = 3/2*pi^(1/2)*epsi0/del0;
+C2 = (2/3)*pi^2*del0/Nw^2
+C3 = (3/2)*pi^(1/2)*epsi0/del0
 [uarray, duarray] = SolverKdV_SymplecticM4a_MC(...
     C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin);
 

### 2019-10-11 paper2_JNLS/paper.tex
NM

@@
 
-	In Fig.~\ref{uhist} we show the displacement distributions that were measured in the experiments (top row) versus those that emerge from the TKdV simulations (bottom row).
+	In Fig.~\ref{uhist} we show the distributions of surface displacement that were measured in the experiments (top row) versus those that emerge from the TKdV simulations (bottom row).
 The experimental histograms represent exactly the same data depicted in Fig.~\ref{ExpDiagStats}, only converted to dimensionless displacement, $u$, for comparison against theory.

### 2019-10-11 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\sumsamp}{\sum_{i=1}^{\Nsamp}}
+\newcommand{\traveldist}{\ell}
 
@@
 \section{Theoretical framework}
+\label{theory}
 
@@
 
-	The comparison between theory and experiment in Fig.~\ref{uhist} is striking.
+	The comparison between experiments and theory in Fig.~\ref{uhist} is striking.
 As seen in the figure, both the experiments and theory show a transition from a nearly symmetric upstream distribution, to a highly skewed distribution downstream.
@@
 
-	Next, we aim to make the same comparison for the slope of the free surface.
+	Next, we aim to make the same comparison for free-surface slopes.
 We note that the derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.
-We therefore convert the theoretical calculated slope values back to physical slope via
+We therefore convert the theoretical calculated slopes values back to physical slopes via
 \begin{equation}
@@
 \end{equation}
-So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode,  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.
+So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode ($k = 1$),  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.
 
@@
 Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50.
-Thus, the jump in kurtosis, while not as extreme as that measured in experiments, is quite significant in the theory.
+Thus, the theoretically predicted jump in kurtosis, while not as extreme as that measured in experiments, is quite significant.
 These elevated levels of kurtosis give the distributions the look of being stretched in the tails and pointy near the centers (especially in the experiments).
@@
 In particular, one large wave is seen to propagate from the lower-left corner of the figure to the upper right.
-A bias towards positive surface displacement is visually apparent in these downstream dynamics.
+A bias towards positive surface displacement is visually apparent in these downstream dynamics, without even viewing the histograms.
 
@@
 
-Downstream of the ADC, however, $\drat = 0.24$, which significantly amplifies nonlinearity while suppressing dispersion in Eq.~\ref{TKdV} and allows waves to propagate in either direction.
+	Downstream of the ADC, however, $\drat = 0.24$, which significantly amplifies nonlinearity while suppressing dispersion, thereby allowing waves to propagate in either direction.
 In particular, the dispersionless limit of KdV is the Burgers-Hopf equation, in which the nonlinear advection is proportional in both amplitude and direction to the value of $u$.
@@
 
-%We remind the reader that the KdV framework tracks the long-time evolution of waves in the reference frame moving with the linear speed $c = \sqrt{g \depth}$.
+	We remind the reader that the KdV framework tracks the long-time evolution waves in a reference frame moving with the characteristic speed $c = \sqrt{g \depth}$ from linear theory.
+Hence, the above discussion of left-or-right going waves cannot be interpreted in the context of the experiments, which take a fixed reference frame, without an appropriate Galilean transformation.
+Hence, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed that is near $c$.
+The directions and speeds calculated by the TKdV framework simply determines how much the actually speed of waves deviates from $c$.
 
-% Timescale of 10 is about right to see waves propagate through the entire periodic domain.
-This however would be a very long experimental time scale.
-So what gives?
-We answer that question with skewness and kurtosis plots which evolve over a much shorter timescale.
+	Finally, Fig.~\ref{trajectories} sheds some light on the timescales required for wave evolution in the TKdV framework.
+It appears that the normalized time of $t=10$ is roughly the correct timescale to observe substantial wave dynamics.
+More precisely, both the left-going waves in Fig.~\ref{trajectories}(a) and the right-going waves in Fig.~\ref{trajectories}(b) require a dimensionless time of about $t=7$ to traverse the entire periodic domain.
+For the representative experiments, the characteristic timescale, $\timescale$, is 2.8~sec upstream and 5.7~sec downstream.
+Hence, a dimensionless time of $t=7$ corresponds to roughly 20~sec and 110~sec upstream and downstream respectively.
+Given the wave speeds ($c = $ 110 cm/s upstream and 54 cm/s downstream), these timescales can be converted to a distance traveled by the waves.
+In fact, a simple calculation using the definitions in Section \ref{theory}, gives the distance traveled as $\traveldist = \lamfac g/(2 \pi \freqp^2)$.
+Notice the depth dependence has serendipitously canceled.
+Hence, the distance traveled by the waves over the dimensionless time of $t=7$ is the same value, 310 cm, both upstream and downstream.
+This distance is roughly half the 6 m length of the wave tank.
 
-\vsp{5}
-
-Next stuff
+	These estimates raise an important question: if a distance of roughly 3 meters is required for significant wave evolution under the KdV framework, how is it that the experiments exhibit substantial changes in wave statistics a short distance, roughly 15 cm, downstream of the ADC?
+The results shown in Fig.~\ref{skew-kurt} help resolve this question.
+This figure shows the evolution of skewness (left) and kurtosis (right) of surface displacement computed by the TKdV deterministic simulations.
+The bold black curve shows the ensemble mean, while the skewness and kurtosis of 100 individual trajectories are shown by the faint gray curves.
+Importantly, the skewness and kurtosis evolve over a much shorter timescale than the aforementioned $t=7$ required for waves to cross the entire domain.
+More precisely, skewness and kurtosis saturate to their asymptotic value after a dimensionless time of roughly $t = 1$, corresponding
 

### 2019-10-11 paper2_JNLS/paper.tex
NM

@@
 For the representative experiments, the characteristic timescale, $\timescale$, is 2.8~sec upstream and 5.7~sec downstream.
-Hence, a dimensionless time of $t=7$ corresponds to roughly 20~sec and 110~sec upstream and downstream respectively.
-Given the wave speeds ($c = $ 110 cm/s upstream and 54 cm/s downstream), these timescales can be converted to a distance traveled by the waves.
-In fact, a simple calculation using the definitions in Section \ref{theory}, gives the distance traveled as $\traveldist = \lamfac g/(2 \pi \freqp^2)$.
+Hence, a dimensionless time of $t=7$ corresponds to roughly 20~sec and 40~sec upstream and downstream respectively.
+Given the wave speeds ($c = $ 110 cm/s upstream and 54 cm/s downstream), these timescales can be converted to distances traveled by the waves.
+In fact, a simple calculation using the definitions in Section \ref{theory} gives the distance traveled as $7 \lamfac g/(2 \pi \freqp^2)$.
 Notice the depth dependence has serendipitously canceled.
-Hence, the distance traveled by the waves over the dimensionless time of $t=7$ is the same value, 310 cm, both upstream and downstream.
-This distance is roughly half the 6 m length of the wave tank.
+Hence, the distance traveled by the waves over a dimensionless time of $t=7$ is the same value upstream and downstream,  roughly 22 meters.
+Note that this distance is significantly greater than the 6-meter length of the experimental wave tank.
 
-	These estimates raise an important question: if a distance of roughly 3 meters is required for significant wave evolution under the KdV framework, how is it that the experiments exhibit substantial changes in wave statistics a short distance, roughly 15 cm, downstream of the ADC?
+	These estimates raise an important question: if a distance of 22 meters is required for significant wave evolution under the KdV framework, how is it that the experiments exhibit substantial changes in wave statistics a much shorter distance downstream of the ADC?
 The results shown in Fig.~\ref{skew-kurt} help resolve this question.
-This figure shows the evolution of skewness (left) and kurtosis (right) of surface displacement computed by the TKdV deterministic simulations.
-The bold black curve shows the ensemble mean, while the skewness and kurtosis of 100 individual trajectories are shown by the faint gray curves.
-Importantly, the skewness and kurtosis evolve over a much shorter timescale than the aforementioned $t=7$ required for waves to cross the entire domain.
-More precisely, skewness and kurtosis saturate to their asymptotic value after a dimensionless time of roughly $t = 1$, corresponding
+This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same TKdV deterministic simulations as pictured in Fig.~\ref{trajectories}.
+The bold black curve shows the ensemble mean, while the skewness and kurtosis of 100 individual trajectories are shown by the faint gray curves to give a sense for the variance involved.
+Importantly, the skewness and kurtosis evolve over a much shorter timescale than the aforementioned $t=7$.
+More precisely, skewness and kurtosis have already saturated to their asymptotic by $t = 1$, and have reached half of that value by $t = 0.18$.
+These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order of magnitude as the relevant distances in the experiments.
 

### 2019-10-11 paper2_JNLS/paper.tex
NM

@@
 
-	These estimates raise an important question: if a distance of 22 meters is required for significant wave evolution under the KdV framework, how is it that the experiments exhibit substantial changes in wave statistics a much shorter distance downstream of the ADC?
-The results shown in Fig.~\ref{skew-kurt} help resolve this question.
-This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same TKdV deterministic simulations as pictured in Fig.~\ref{trajectories}.
-The bold black curve shows the ensemble mean, while the skewness and kurtosis of 100 individual trajectories are shown by the faint gray curves to give a sense for the variance involved.
-Importantly, the skewness and kurtosis evolve over a much shorter timescale than the aforementioned $t=7$.
-More precisely, skewness and kurtosis have already saturated to their asymptotic by $t = 1$, and have reached half of that value by $t = 0.18$.
-These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order of magnitude as the relevant distances in the experiments.
-
 % Figure
@@
 \caption{
-Skewness and kurtosis in DNS}
+Dynamic evolution of surface-displacement skewness and kurtosis from the TKdV simulations.
+Faint gray curves show skewness and kurtosis of 100 individual trajectories, and the bold black curves show the corresponding ensemble mean.
+Skewness and kurtosis evolve over much shorter timescales than that required for waves to cross the domain.}
 \label{skew-kurt}
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+	These estimates raise an important question: if a distance of 22 meters is required for significant wave evolution under the KdV framework, how is it that the experiments exhibit substantial changes in wave statistics a much shorter distance downstream of the ADC?
+The results shown in Fig.~\ref{skew-kurt} help resolve this question.
+This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same TKdV deterministic simulations as pictured in Fig.~\ref{trajectories}.
+The bold black curve shows the ensemble mean (ensemble size of 1000), while the skewness and kurtosis of 100 individual trajectories are shown by the faint gray curves to give a sense for the variance involved.
+Importantly, the skewness and kurtosis evolve over a much shorter timescale than the aforementioned $t=7$.
+More precisely, skewness and kurtosis have already saturated to their asymptotic by $t = 1$, and have reached half of that value by $t = 0.18$.
+These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order of magnitude as the relevant distances in the experiments.
 

### 2019-10-11 paper2_JNLS/paper.tex
NM

@@
 The bold black curve shows the ensemble mean (ensemble size of 1000), while the skewness and kurtosis of 100 individual trajectories are shown by the faint gray curves to give a sense for the variance involved.
-Importantly, the skewness and kurtosis evolve over a much shorter timescale than the aforementioned $t=7$.
-More precisely, skewness and kurtosis have already saturated to their asymptotic by $t = 1$, and have reached half of that value by $t = 0.18$.
-These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order of magnitude as the relevant distances in the experiments.
+Importantly, the skewness and kurtosis evolve on a much shorter timescale than the aforementioned $t=7$.
+More precisely, skewness and kurtosis have already saturated to their asymptotic by $t = 1$, and reach half of that value by $t = 0.18$.
+These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order as the relevant distances in the experiments.
+Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter the statistical distributions, and this reconfiguration occurs on a much shorter timescale than that required for waves to cross the entire periodic domain.
+%Thus, the wave-field undergoes a rapid reconfiguration which significantly alters the statistical distributions
 
-\vsp{5}
  %----------------------------------------------------------------------------%

### 2019-10-11 paper2_JNLS/paper.tex
NM

@@
 \usepackage{graphicx, amsmath, amssymb, amsfonts, mathtools, mathrsfs, color}
-\usepackage{comment, enumerate, tabularx}
+\usepackage{comment, enumerate, tabularx, todonotes}
 \usepackage{natbib, hyperref, url}
@@
 \newcommand{\vardn}{\var_+}
+\newcommand{\varupdn}{\var_{\pm}}
 \newcommand{\kurt}{\text{kurt}}
@@
 \qquad \text{for } x \in [-\pi,\pi] \\
+\label{C3C2}
 &C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad
@@
 
+\todo[inline]{Here: Explain intuitive justification for matching condition}
+
+\nick{The downstream Hamiltonian is not conserved upstream and will vary dynamically in upstream wavefield.
+Once the wavefield crosses the ADC, whatever value $\Hdn$ happens to be is locked in for the rest of the time (is conserved thereafter).
+Hence, the downstream hamiltonian must match at the ADC on a trajectory by trajectory basis.
+Therefore, it must match statistically.}
+
+\vsp{5}
 	The statistical matching condition imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
@@
 %----------------------------------------------------------------------------%
-\section{Results}
+\section{Comparison between theory and experiments}
 
@@
 Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter the statistical distributions, and this reconfiguration occurs on a much shorter timescale than that required for waves to cross the entire periodic domain.
-%Thus, the wave-field undergoes a rapid reconfiguration which significantly alters the statistical distributions
 
  %----------------------------------------------------------------------------%
-% Skewness formula
+% Skewness Ratio 
 %----------------------------------------------------------------------------%
+\section{Predicted outgoing skewness and experimental confirmation}
+
+This section discusses perhaps the most important single result of the manuscript.
+In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}.
+More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$, in particular how that ratio depends on driving amplitude and depth ratio.
+Below, we recap the derivation of this formula and then test the prediction against our direct experimental measurements.
+
 \subsection{Explicit formula for outgoing skewness}
-We now examine an explicit formula for the outgoing skewness as originally founded in \cite{majda2019}.
-An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$.
-Inserting this approximation into the statistical matching condition \eqref{statmatch} and simplifying gives
+
+The explicit formula for outgoing skewness arises directly from the statistical matching condition \eqref{statmatch}, which states that the downstream Hamiltonian must match in expected value at the ADC, i.e.~with respect to the incoming and outgoing Gibbs measures.
+This condition can be written more explicitly as
 \begin{equation}
-\label{H3H2}
-\frac{\meandn{\Hthree^+}} {\meandn{\Htwo^+} - \meanup{\Htwo^+}} = \frac{C_2}{C_3} \dratdn^2
+\label{statmatchexp}
+C_2 \dratdn^{1/2} \, \meanup{\Htwo} - C_3 \dratdn^{-3/2} \, \meanup{\Hthree } = 
+C_2 \dratdn^{1/2} \, \meandn{\Htwo} - C_3 \dratdn^{-3/2} \, \meandn{\Hthree }
 \end{equation}
-%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
-To convert to dimensional, i.e.~experimental, values we first note that
+Both the experiments and simulations show the upstream skewness to be negligible compared to the downstream skewness, allowing the term with $\meanup{\Hthree }$ to be dropped in \eqref{statmatchexp}.
+With this approximation, \eqref{statmatchexp} yields the relationship
+\begin{equation}
+\label{H3H2ratio}
+\frac{\meandn{\Hthree}} {\meandn{\Htwo} - \meanup{\Htwo}} = \frac{C_2}{C_3} \dratdn^2
+\end{equation}
+
+	Our next task is to convert this formula to physical variables so that it can be tested against experimental data.
+First, following the definition of $\Hthree$ in Eq.~\eqref{H3H2} and the relationship, $u = \eta/(\pi^{1/2} \etastd)$, a straightforward calculation gives
 \begin{align}
-\mean{\Hthree^+} = \frac{\pi}{3} \mean{u^3} = 
-\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
+\label{H3convert}
+\meandn{\Hthree} = \frac{\pi}{3} \meandn{u^3} = 
+\frac{1}{3 \pi^{1/2}} \frac{\meandn{\eta^3}}{ \etastd^3} = 
+\frac{1}{3 \pi^{1/2}} \skwdn(\eta)
 \end{align}
-Second, to convert the downstream surface slope, we note
+This equation simply says that the expected value of $\Hthree$ is directly related to the outgoing skewness.
+Next, we must convert the $\Htwo$-quantities, which involve surface slope.
+Here, it is important to remember that, since it is the {\em downstream} Hamiltonian that is matched in \eqref{statmatch}, one must use the {\em downstream} characteristic scale $\lengthscale_+$ to convert back into physical variables.
+Using the definition of $\Htwo$ in Eq.~\eqref{H3H2} and $\lengthscale_+$ in Eq.~\eqref{scales} gives
 \begin{align}
-\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
+\label{H2convert}
+&\meanupdn{\Htwo} = \pi \meanupdn{ u_x^2} = 
+\left( \frac{\lamfac \lamdn}{2 \pi \etastd} \right)^2 \varupdn(\eta_x)
 \end{align}
-where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
-Note: I believe it is important here to use $\lamdn$ in reverting to dimensional variables since we are matching $\Hdn$.
-Hence, we have
-\begin{align}
-&\mean{\Htwo^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
-\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
-\end{align}
-%
-Then algebra gives
+Inserting \eqref{H3convert} and \eqref{H2convert} into \eqref{H3H2ratio}, and using the definitions of $C_2$ and $C_3$ from \eqref{C3C2}, gives the remarkably explicit formula
 \begin{equation}
-\frac{\meandn{\Hthree}} {\meandn{\Htwo} - \meanup{\Htwo}} = 
-\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
-\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
+\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 
+%= \frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
 \end{equation}
-Also we have the simplification
-\begin{equation}
-\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
-\end{equation}
-Hence, \eqref{H3H2} gives the remarkably explicit formula
-\begin{equation}
-\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 =
-\frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
-\end{equation}
+This formula links the skewness of the outgoing wave field to the change in the variance of surface slope.
+In particular, it predicts that their ratio to scale as the {\em inverse cube of wave amplitude} and the {\em cube of the depth ratio}, as seen on the right-hand-side of the equation.
+It is a rather unexpected relationship, as it would be difficult to anticipate that the displacement skewness and slope variance, among all possibilities, are so intimately related.
+
+\vsp{3}
 In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.

### 2019-10-11 code_julia_sampling/io_routines.jl
NM

@@ -1,0 +1,18 @@
+
+function writedata(data::Array, filename::AbstractString)
+	iostream = open(filename, "w")
+	writedlm(iostream, data)
+	close(iostream)
+end
+
+function readvec(file::AbstractString)
+	iostream = open(file, "r")
+	vec1 = readdlm(iostream, comments=true, comment_char='#')[:,1]
+	close(iostream)
+	return vec1
+end
+
+# Significant digits
+function sig(var, sigdig)
+	return round(var,sigdigits=sigdig)
+end

### 2019-10-11 code_julia_sampling/math_routines.jl

@@ -1,0 +1,136 @@
+#---------- Basic Stuff ----------#
+using Distributions
+using Roots
+using DelimitedFiles
+using LinearAlgebra
+using FFTW
+using JLD
+using Plots
+using LaTeXStrings
+#---------------------------------#
+
+#---------- Real FFT Routines ----------#
+#= Compute the FFT between physical and spectral space assuming
+1) the signal uu is real and 2) uu has zero mean (i.e. momentum M=0)
+Note: uu has length npoints = 2*nmodes or 2*nmodes + 1. =#
+#= Valid choices for npoints are 2*nmodes or 2*nmodes+1, 
+but I need a function to keep my usage consistent.
+Note: for H3test npts = 2*modes is slightly more accurate. =#
+function npoints(nmodes::Int)
+	return 2*nmodes
+end
+#= Basic realfft to go from uu to uhat. =#
+function realfft(uu::Vector{Float64})
+	uhat = rfft(uu)/length(uu)
+	@assert(abs(uhat[1])/maximum(abs,uhat) < 1e-6) 
+	return uhat[2:end]
+end
+#= Basic irealfft to go from uhat to uu (no upsampling). =#
+function irealfft(uhat::Vector{Complex{Float64}})
+	nmodes = length(uhat)
+	uu = irfft([0; uhat], npoints(nmodes))
+	return uu*length(uu)
+end
+#= Upsampled version of irealfft. =#
+function ifftup(uhat::Vector{Complex{Float64}})
+	uhat = [uhat; zeros(eltype(uhat), length(uhat))]
+	return irealfft(uhat)
+end
+#---------------------------------------#
+
+#---------- Hamiltonian Routines ----------#
+#= In all routines, the physical signal is assumed to be real 
+with zero momentum, M = int u dx = 0. Therefore, the Fourier transform 
+only requires the modes k=1,...,nmodes. The negative modes are given 
+by the complex conjugates (CC), and the k=0 mode vanishes due to M=0. =#
+#---------- Low-level Hamiltonian Routines ----------#
+#= Compute the energy, E = 1/2 int u^2 dx. =#
+function energy(uhat::Vector{Complex{Float64}})
+	return 2*pi*norm(uhat)^2
+end
[…]

### 2019-10-11 code_julia_sampling/transfer.jl

@@ -1,0 +1,91 @@
+#= Calculate the transfer function:
+For given upstream theta, enforce the matching condition to 
+compute the downstream theta.=#
+
+include("math_routines.jl")
+include("io_routines.jl")
+
+function transfun(randfile::AbstractString)
+	rr, nmodes, nsamples = load(randfile, "rr", "nmodes", "nsamples")
+	println(nmodes)
+	println(nsamples)
+
+	# Parameters
+	eps0 = 0.017
+	del0 = 0.22
+	lamfac = 8 #div(nmodes,2)
+	D0 = 0.24
+	# Constants
+	C2,C3 = C2C3(eps0,del0,lamfac)
+	cc = ConstantList(C2,C3,D0)
+	# Compute the list of Hamiltonians
+	hdn = hamdn(rr,cc)
+	hup = hamup(rr,cc)
+
+	# Specify list of upstream thetas.
+	thup = 0.: 0.5 : 30.
+	nth = length(thup)
+	thdn = zeros(nth); skup = zeros(nth); skdn = zeros(nth)
+	guess = thup[1]
+	for nn = 1:nth
+		meandiff(thdn) = meanham(hdn, hdn, thdn) - meanham(hdn, hup, thup[nn])
+		thdn[nn] = find_zero(meandiff, guess, Order1())
+		guess = thdn[nn]
+		nn>1 ? guess = 2*thdn[nn] - thdn[nn-1] : 0.	# Linear extrapolation
+		println("Iteration ", nn, "; thup = ", thup[nn], "; thdn = ", sig(thdn[nn],3))
+
+		# Compute the skewness of eta
+		skup[nn] = skewu(rr.H3, hup, thup[nn])
+		skdn[nn] = skewu(rr.H3, hdn, thdn[nn])
+		println("Upstream skewness = ", sig(skup[nn],3))
+		println("Downstream skewness = ", sig(skdn[nn],3), "\n")
+	end
+	savefile = string("thvars-", string(nmodes), "-", string(lamfac), ".jld")
+	save(savefile, "thup", thup, "thdn", thdn, 
+		"skup", skup, "skdn", skdn, "nmodes", nmodes, "lamfac", lamfac)
+end
+
+function plotstuff(datafile::AbstractString)
+	thup,thdn,skup,skdn = load(datafile, "thup", "thdn", "skup", "skdn")
+	nmodes,lamfac = load(datafile, "nmodes", "lamfac")
[…]

### 2019-10-11 code_julia_sampling/transfer.jl
NM

@@ -77,13 +77,15 @@ end
 
 
 ## RUN STUFF
-#uniform_sample(16, 2*10^7, false)
+@time( uniform_sample(16, 2*10^8, false) )
+@time( uniform_sample(16, 5*10^8, false) )
+
 #uniform_sample(12, 2*10^7, false)
 #uniform_sample(20, 2*10^7, false)
 
 #transfun("rand-16-2e7.jld")
-transfun("rand-12-2e7.jld")
-transfun("rand-20-2e7.jld")
+#transfun("rand-12-2e7.jld")
+#transfun("rand-20-2e7.jld")
 
 #output_text("thvars-16-8-2e7.jld")
 #output_text("thvars-12-6-2e7.jld")

### 2019-10-12 paper2_JNLS/paper.tex
NM

@@
 Thus, the dynamics of interest are confined to the unit hypersphere of $\En = 1$ in $\CC^{\Lambda}$.
-%
 In summary, the TKdV system possesses three important invariants: momentum, energy, and hamiltonian.
@@
 	In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced by Abramov et al.~for the Burgers-Hopf system \cite{abramov2003}.
-Specifically, this ensemble is microcanonical in the energy, with a fixed energy value, and it is canonical in the Hamiltonian.
-By fixing the energy and hence confining to the compact set of the unit hypersphere, this construction avoids the divergence at infinity that would occur for a simple canonical ensemble.
-It thus yields a normalizable distribution.
-This framework applies equally well to the truncated or untruncated system and hence we will not distinguish between the two.
-
-The Hamiltonian on either side of the ADC, $\Ham^{\pm}$, generates a corresponding mixed ensemble, or Gibbs measure given by 
+Specifically, this ensemble is microcanonical in energy and canonical in the Hamiltonian.
+The reason to introduce this ensemble is the sign indefiniteness of the cubic term $\Hthree$ in the Hamiltonian, which would cause a simple canonical distribution to diverge at infinity.
+The mixed ensemble, however, fixes the energy and hence confines dynamics to the compact set of the unit hypersphere $\En = 1$.
+Since the Hamiltonian is a continuous, its value is bounded on the unit hypersphere, and thus the mixed ensemble produces a normalizable distribution.
+This construction applies equally well to the truncated or untruncated KdV system and hence we will not distinguish between the two.
+We note that other possible constructions may be applicable too \cite{kleeman2014nonequilibrium}.
+
+	On either side of the ADC, the mixed ensemble, or {\em Gibbs measure}, follows directly from the corresponding Hamiltonian via
 \begin{align}
 \label{Gibbs}
-\Gupdn = Z_{\thupdn}^{-1} \exp(-\thupdn \Hupdn) \delta(\En - 1)
-\end{align}
-% Note: The PNAS paper uses C_theta^{1}. The Majda, Qi Stat Phys paper uses C_theta sometimes and C_theta^{-1} in other places (page 11). So I'll just pick the one I like better.
-Here $\theta = \thupdn$ is the inverse temperature and $Z_{\theta}$ a constant (the partition function) that depends on $\theta$.
-Each measure $\Gupdn$ induces a corresponding ensemble average $\mean{}_{\pm}$.
+\Gupdn = Z_{\thupdn}^{-1} \, \exp(-\thupdn \Hupdn) \delta(\En - 1)
+\end{align}
+Here $\theta = \thupdn$ is the inverse temperature, which will take a different value on either side of the ADC, and $Z_{\theta}$ a constant that depends on $\theta$.
+Each measure $\Gupdn$ induces a corresponding ensemble average, denoted $\mean{}_{\pm}$.
+
+% NOTES
+% 1) The PNAS paper uses C_theta^{1}. The Majda, Qi Stat Phys paper uses C_theta sometimes and C_theta^{-1} in other places (page 11). So I'll just pick the one I like better.
+% 2) Could comment that other possible ensembles exist \cite{kleeman2014nonequilibrium}.
+
+%-------------------- COMMENTED OUT ------------------------%
+% Definitions of ensemble and time averages.
+% It seems that I did not need to use this notation so lets kill it.
+\begin{comment}
 Throughout this paper we use a bracket for ensemble average and overbar for a time average; i.e.~for any quantity $Q$,
@@
 Note that if $\theta = 0$, then the Gibbs measure is a {\em uniform} measure on the unit hypersphere in $\CC^{\Lambda}$, which we will denote as either $\Gz$ or simply $d\mu$.
+\end{comment}
+%----------------------------------------------------------------------------%
 

### 2019-10-12 paper2_JNLS/paper.tex
NM

@@
 \begin{align}
-&u(x,t) = \sum_{k=-\infty}^{\infty} \uhat_k e^{i k x} \, , \\
+&u(x,t) = \sum_{k=-\infty}^{\infty} \uhat_k(t) \, e^{i k x} \, , \\
 \label{uhat}
-&\uhat_k = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x) e^{-i k x} \dx
+&\uhat_k(t) = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x,t) \, e^{-i k x} \dx
 \end{align}
@@
 \uL(x,t) = \Proj u =
-\sum_{\abs{k} \le \Lambda} \uhat_k e^{i k x} \, , \qquad
+\sum_{\abs{k} \le \Lambda} \uhat_k(t) \, e^{i k x} \, , \qquad
 \end{align}
@@
 Specifically, this ensemble is microcanonical in energy and canonical in the Hamiltonian.
-The reason to introduce this ensemble is the sign indefiniteness of the cubic term $\Hthree$ in the Hamiltonian, which would cause a simple canonical distribution to diverge at infinity.
+The reason this ensemble is needed is the sign indefiniteness of the cubic term $\Hthree$ in the Hamiltonian, which would cause a simple canonical distribution to diverge at infinity.
 The mixed ensemble, however, fixes the energy and hence confines dynamics to the compact set of the unit hypersphere $\En = 1$.
-Since the Hamiltonian is a continuous, its value is bounded on the unit hypersphere, and thus the mixed ensemble produces a normalizable distribution.
+Since the Hamiltonian is continuous, its value is bounded on the unit hypersphere, and thus the mixed ensemble produces a normalizable distribution.
 This construction applies equally well to the truncated or untruncated KdV system and hence we will not distinguish between the two.
@@
 Here $\theta = \thupdn$ is the inverse temperature, which will take a different value on either side of the ADC, and $Z_{\theta}$ a constant that depends on $\theta$.
-Each measure $\Gupdn$ induces a corresponding ensemble average, denoted $\mean{}_{\pm}$.
+Each measure $\Gupdn$ induces a corresponding ensemble average, denoted $\mean{\cdot}_{\pm}$.
 
@@
 %----------------------------------------------------------------------------%
-\subsection{Matching conditions and the transfer function}
-
-	Recall that the abrupt depth change is met by traveling waves at some time $t = T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
-The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales, $t \gg \delta^{-2}$.
-Meanwhile, wave evolution over shorter timescales, $t \ll \delta^{-2}$, is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e.
-no nonlinearity and no dispersion).
-Hence, a short instant after waves pass over the ADC, the waveform does not yet have time to alter significantly, which gives the (deterministic) matching condition
-\begin{align}
+\subsection{Matching at the ADC}
+
+	Recall that the abrupt depth change is met by traveling waves at time $T_{ADC} = 0$, set to zero for convenience.
+The KdV equations describe wave dynamics over {\em long} timescales, $t \gg \timescale$, capturing weakly nonlinear and weakly dispersive effects.
+Meanwhile, evolution over shorter timescales is simply described by linear theory.
+The event of a wave crossing the ADC is precisely such a short-time event, and so we will employ matching conditions at the ADC that result from linear theory.
+
+	More specifically, we assume continuity of the surface displacement, $\eta$, at the ADC \nick{(cite)}.
+Since the temporal fluctuations of $\eta$ match, waves crossing the ADC must rapidly adjust in wavelength so as to attain the same frequency in two different depths --- a well-understood phenomenon \nick{(cite)}.
+Of course, since the normalized domain $x \in [-\pi,\pi]$ is scaled on the characteristic wavelength, there is no change in the dimensionless wave-field $u(x,t)$, giving the condition
+\begin{equation}
+\label{detmatch}
 u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+}, 
+%\quad \mbox{for } -\pi \le x \le \pi,
 \qquad \mbox{\em Deterministic matching condition}
-\end{align}
-holding for $x \in [-\pi, \pi]$. This matching condition is used for the dynamical system \eqref{TKdV}.
-%or in Johnson's formulation, it is $\depth^{1/4} \eta$ that matches.
-
-	Since $u$ matches for every single trajectory, it must also match in the statistical sense.
-Moreover, the downstream Hamiltonian $\Hdn$ is a functional of $u$ and hence must also match.
-This gives the {\em statistical matching condition}
+\end{equation}	
+We call \eqref{detmatch} the deterministic matching condition to contrast with its statistical counterpart introduced below.
+This matching condition is employed in the deterministic simulations of the TKdV dynamical system \eqref{TKdV}.
+We note that in the alternate formulation mentioned earlier, it is the product $\depth^{1/4} \eta$ that would match at the ADC \cite{johnson1997modern}.
+
+	Now, from the perspective of statistical mechanics, consider the communication between the statistical ensembles, $\Gupdn$, upstream and downstream of the ADC.
+These two systems are in contact at the ADC, and so the upstream state $\Gup$ can be regarded as a thermal reservoir that influences the downstream state $\Gdn$.
+In fact, the above deterministic  condition directly leads to a simple description for the link.
+The quantity of interest is the outgoing Hamiltonian $\Hdn$, since its value at $t=0^{-}$ is set by the upstream dynamics, and then this particular value is conserved thereafter in the downstream dynamics.
+Since $u$ matches at the ADC, so must $\Hdn$, and, in fact, this matching holds for every individual trajectory.
+Recall that $\Hdn(t)$ is not conserved in the upstream dynamics and so its value varies for $t < 0$.
+However, appealing to weak Ergodicity of the system, the ensemble mean $\meanup{\Hdn(t)}$ is expected to be independent of time.
+In particular, the value $\meanup{\Hdn \vert_{t=0^{-}} }$ is the same as the bare ensemble mean $\meanup{\Hdn}$.
+Afterwards, the particular value $\Hdn \vert_{t=0^{-}}$ is conserved in the downstream dynamics, on a trajectory-by-trajectory basis and for all time.
+Thus, the downstream measure $\Gdn$ must recover the ensemble mean $\meanup{\Hdn \vert_{t=0^{-}} }$, which gives the statistical matching condition
 \begin{align}
@@
 \end{align}
-Once $\Hdn$ downstream is determined, it is conserved in the outgoing wave field.
-
-\todo[inline]{Here: Explain intuitive justification for matching condition}
-
-\nick{The downstream Hamiltonian is not conserved upstream and will vary dynamically in upstream wavefield.
-Once the wavefield crosses the ADC, whatever value $\Hdn$ happens to be is locked in for the rest of the time (is conserved thereafter).
-Hence, the downstream hamiltonian must match at the ADC on a trajectory by trajectory basis.
-Therefore, it must match statistically.}
+
+\vsp{5}
 
@@
 The {\em transfer function} $\transf$ will be a key link needed to relate the theory to experiments.
+
+% NOTES
+%Further, since the regime is shallow, short-time dispersive effects are minimal, meaning that all waves travel at nearly the same speed $c = \sqrt{g \depth}$, independent of wavelength.
 

### 2019-10-12 paper2_JNLS/paper.tex
NM

@@
 
-	Recall that the abrupt depth change is met by traveling waves at time $T_{ADC} = 0$, set to zero for convenience.
-The KdV equations describe wave dynamics over {\em long} timescales, $t \gg \timescale$, capturing weakly nonlinear and weakly dispersive effects.
+	Recall that the abrupt depth change is met by traveling waves at dimensionless time $T_{ADC} = 0$, set to zero for convenience.
+The KdV equations describe wave dynamics over {\em long} timescales, physically $t \gg \timescale$, capturing weakly nonlinear and weakly dispersive effects.
 Meanwhile, evolution over shorter timescales is simply described by linear theory.
@@
 In fact, the above deterministic  condition directly leads to a simple description for the link.
-The quantity of interest is the outgoing Hamiltonian $\Hdn$, since its value at $t=0^{-}$ is set by the upstream dynamics, and then this particular value is conserved thereafter in the downstream dynamics.
+The quantity of interest is the outgoing Hamiltonian $\Hdn$, since its value just upstream of the ADC, $t=0^{-}$, is set by the upstream dynamics, and then this particular value is conserved thereafter in the downstream dynamics.
 Since $u$ matches at the ADC, so must $\Hdn$, and, in fact, this matching holds for every individual trajectory.
 Recall that $\Hdn(t)$ is not conserved in the upstream dynamics and so its value varies for $t < 0$.
-However, appealing to weak Ergodicity of the system, the ensemble mean $\meanup{\Hdn(t)}$ is expected to be independent of time.
+However, appealing to weak Ergodicity, the ensemble mean $\meanup{\Hdn(t)}$ is expected to be independent of time.
 In particular, the value $\meanup{\Hdn \vert_{t=0^{-}} }$ is the same as the bare ensemble mean $\meanup{\Hdn}$.
-Afterwards, the particular value $\Hdn \vert_{t=0^{-}}$ is conserved in the downstream dynamics, on a trajectory-by-trajectory basis and for all time.
-Thus, the downstream measure $\Gdn$ must recover the ensemble mean $\meanup{\Hdn \vert_{t=0^{-}} }$, which gives the statistical matching condition
+Afterwards, the particular value $\Hdn \vert_{t=0^{-}}$ is conserved in the downstream dynamics, on a trajectory-by-trajectory basis and for all $t>0$.
+Thus, the downstream measure $\Gdn$ must recover the same ensemble mean $\meanup{\Hdn \vert_{t=0^{-}} } = \meanup{\Hdn}$, producing the simple condition
 \begin{align}
@@
 
-\vsp{5}
-
-\vsp{5}
-	The statistical matching condition imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
+	This statistical matching condition imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
 In particular, we view $\thup$ as given by the random-state of the incoming wave field.
-Thus, while we will treat $\thup$ as a free parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by matching statistics at the ADC.
-In this way, \eqref{statmatch} implies the functional dependence
+Thus, while we will treat $\thup$ as a free parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by the matching condition \eqref{statmatch}, giving the functional dependence
 \begin{equation}
 \thdn = \transf \left( \thup \right)
+\qquad \mbox{\em Transfer function}
 \end{equation}

### 2019-10-12 code_julia_sampling/transfer.jl
NM

@@ -77,8 +77,8 @@ end
 
 
 ## RUN STUFF
-@time( uniform_sample(16, 2*10^8, false) )
-@time( uniform_sample(16, 5*10^8, false) )
+@time( uniform_sample(16, 2*10^5, false) )
+#@time( uniform_sample(16, 5*10^8, false) )
 
 #uniform_sample(12, 2*10^7, false)
 #uniform_sample(20, 2*10^7, false)

### 2019-10-12 code_rescaledModel/SolverKdV_SymplecticM4a_MC.m

@@ -1,5 +1,5 @@
 function [uarray,duarray] = SolverKdV_SymplecticM4a_MC(...
-    C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin)  
+    C2,C3,Drat,JJ,MC,theta,gibd,fi,dt,nout,tfin)  
 
 % This script solves the rescaled KdV equation in a periodic domain
 % using pseudo-spectral & symplectic M4a integrator
@@ -7,7 +7,7 @@ function [uarray,duarray] = SolverKdV_SymplecticM4a_MC(...
 % Input
 % C2 and C3: Constants in Hamiltonian.
 % Drat: Depth ratio; reference Drat = 0.24.
-% Lambda: Number of modes; reference Lambda = 16.
+% JJ: Number of grid points; reference JJ = 32.
 % MC: The number of trajectories; reference MC = 1E4.
 % theta: The inverse temperature; range 10-20.
 % gibd: The Gibbs distribution to sample the initial state.
@@ -17,7 +17,6 @@ function [uarray,duarray] = SolverKdV_SymplecticM4a_MC(...
 % tfin: The stopping time of the simulation; reference tfin = 10.
 
 % Set simulation parameters
-JJ = 2*Lambda;          % Number of grid points.
 Nt = round(tfin/dt);    % Number of time steps.
 MM = round(Nt/nout);    % Save output every MM steps.
 um = 0.;                % zero momentum

### 2019-10-12 code_rescaledModel/makeOutput.m

@@ -2,8 +2,7 @@ function makeOutput(filename)
     % Load the output file.
     load([filename,'.mat']);
     % Basic parameters.
-    JJ = size(uarray,1); 
-    assert(JJ == 2*Lambda);
+    assert(JJ == size(uarray,1));
     assert(MC == size(uarray,2));
     Ntout = size(uarray,3);
     % Flatten the array of u into a single list.

### 2019-10-12 code_rescaledModel/runSolver.m

@@ -5,9 +5,9 @@ epsi0 = 0.017;  % amplitude-to-depth ratio
 del0 = 0.23;    % depth-to-wavelength ratio
 Drat = 0.24;    % depth ratio.
 % Simulation parameters
-Lambda = 16;    % Reference 16
-Nw = 8;         % Reference Lambda/2
-MC = 1E3;       % Reference 1E3 or 5E3 for tfin=10.
+JJ = 32;    % Reference 16
+Nw = 8;         % Reference 8 or JJ/4.
+MC = 1E4;       % Reference 1E3 or 5E3 for tfin=10.
 % time step parameters
 dt = 5E-4;  % Reference 5E-4
 nout = 100; % Reference 100
@@ -22,7 +22,7 @@ else
     theta = 0.;
     gibd = 0.24;
 end
-fi = 0; % fi = 1 for outgoing and 0 for incoming
+fi = 1; % fi = 1 for outgoing and 0 for incoming
 
 
 % Run the simulation.
@@ -30,11 +30,11 @@ tic;
 C2 = (2/3)*pi^2*del0/Nw^2
 C3 = (3/2)*pi^(1/2)*epsi0/del0
 [uarray, duarray] = SolverKdV_SymplecticM4a_MC(...
-    C2,C3,Drat,Lambda,MC,theta,gibd,fi,dt,nout,tfin);
+    C2,C3,Drat,JJ,MC,theta,gibd,fi,dt,nout,tfin);
 
 % Save output
 runtime = round(toc/60, 3, 'significant')
-save('output.mat', 'C2','C3','Drat','Lambda',...
+save('output.mat', 'C2','C3','Drat','JJ',...
     'MC','theta','gibd','fi','dt','nout','tfin',...
     'Nw','runtime','uarray', 'duarray');
 

### 2019-10-12 code_julia_sampling/math_routines.jl
Changed Julia sampling codes to use memory better with sweeps

@@ -99,25 +99,39 @@ function getuhat(rvar::Array{Float64}, nn::Int)
 	uhat = rvar[:,1,nn]+im*rvar[:,2,nn]
 	return uhat/sqrt(energy(uhat))
 end
-#= Sample from a uniform distribution on the hypershpere E=1. =#
-function uniform_sample(nmodes::Int, nsamples::Int, savemicro::Bool)
-	rvar = randn(nmodes,2,nsamples)
-	H2vec, H3vec = [zeros(Float64,nsamples) for nn=1:2]
+#= Sample a single sweep from a uniform distribution on the hypershpere E=1. =#
+function sample_one_sweep(nmodes::Int, nsamp::Int)
+	rvar = randn(nmodes,2,nsamp)
+	H2vec, H3vec = [zeros(Float64,nsamp) for nn=1:2]
 	## Later parralelize this step. #@parallel
 	# Compute H3 and H2 for each sample in a parallel for loop.
-	for nn=1:nsamples
+	for nn = 1:nsamp
 		uhat = getuhat(rvar,nn)
 		H2vec[nn] = ham2(uhat); H3vec[nn] = ham3(uhat)
-		if mod(nn, 10^4) == 0
-			println("Uniform sampling ", 
-				round(100*nn/nsamples,sigdigits=3), "% completed.")
-		end
 	end
-	savemicro ? rsave = Float32.(rvar) : rsave = []
-	savefile = string("rand-", string(nmodes), "-", string(nsamples), ".jld")
-	save(savefile, "rr", RandList(rsave,H2vec,H3vec),
-		"nmodes", nmodes, "nsamples", nsamples)
-	return
+	rlist = RandList(Float32.(rvar), H2vec, H3vec)
+	return rlist
+end
+#= Sample several sweeps from a uniform distribution on the hypershpere E=1.
+Using several sweeps manages the memory better. =#
+function uniform_sample(nmodes::Int, nsweeps::Int, savemicro::Bool)
+	samp_per = 10^5
+	totsamp = nsweeps*samp_per
+	H2vec, H3vec = [zeros(Float64,totsamp) for nn=1:2]
+	for nn = 1:nsweeps
+		rlist = sample_one_sweep(nmodes, samp_per)
+		indrange = (nn-1)*samp_per+1 : nn*samp_per
+		H2vec[indrange] = rlist.H2
+		H3vec[indrange] = rlist.H3
[…]

### 2019-10-12 code_julia_sampling/transfer.jl

@@ -77,11 +77,8 @@ end
 
 
 ## RUN STUFF
-@time( uniform_sample(16, 2*10^5, false) )
-#@time( uniform_sample(16, 5*10^8, false) )
+@time( uniform_sample(16, 2, false) )
 
-#uniform_sample(12, 2*10^7, false)
-#uniform_sample(20, 2*10^7, false)
 
 #transfun("rand-16-2e7.jld")
 #transfun("rand-12-2e7.jld")

### 2019-10-12 paper2_JNLS/paper.tex
NM

@@
 In line with experimental observations, the incoming skewness is small, while the outgoing wave skewness is much higher.
-Specifically, to capture the experimentally observed peak skewness range of 0.6-0.9 seen for the larger amplitudes in Fig.~\ref{ExpSpatialStats}(b), requires selecting $\thup$ in the range 10--20.
+Specifically, to capture the experimentally observed peak skewness range of 0.6-0.9 seen for the larger amplitudes in Fig.~\ref{ExpSpatialStats}(b), requires selecting $\thup$ in the range 10--20 (?).
 
@@
 	We now aim to compare the wave statistics measured in experiments against those that emerge from the TKdV theory.
-Throughout, we will focus on the representative set of experiments detailed in Table \ref{paramtable}.
-In these experiments, the peak downstream skewness was measured to be 0.83.
-We can then use Fig.~\ref{transfig}(c) as a guide to select the appropriate upstream inverse temperature $\thup$.
+Throughout, we will focus on the representative set of experiments detailed in Table \ref{paramtable}, in which the peak downstream skewness was measured to be 0.83.
+%We can then use Fig.~\ref{transfig}(c) as a guide to select the appropriate upstream inverse temperature $\thup$.
 With $(\Lambda, \lamfac) = (16, 8)$ fixed, the figure indicates a value of $\thup = 17$ achieves the desired outgoing skewness.
 We thus run the deterministic TKdV simulations with model parameters $\Lambda = 16$, $\lamfac = 8$, $\thup = 17$ and with $\epsup$, $\delup$, and $\dratdn$ set to the representative values in Table \ref{paramtable}.
+%
+With $(\Lambda, \lamfac) = (16, 8)$ fixed, Fig.~\ref{transfig}(c) indicates $\thup = 20$ as a reasonable value to attain the desired skewness.
+We thus run the deterministic TKdV simulations these model parameters and with $\epsup$, $\delup$, and $\dratdn$ set to the representative values in Table \ref{paramtable}.
 We run the simulations to a sufficiently long dimensionless time of $t_{f} = 10$, with $dt = 5 \times 10^{-4}$ and with $10^3$ trajectories sampled from the upstream Gibbs measure $\Gup$.
@@
 
-\vsp{10}
-%----------------------------------------------------------%
-% Direct numerical simulations
-%----------------------------------------------------------%
-\section{Direct numerical simulations}
-
-\section{Sampling Algorithms}
-%\subsection{Naive acceptance-rejection algorithm}
-%\subsection{Markov-chain Monte Carlo}
-%\subsection{Improved acceptance-rejection algorithm}
-
-%----------------------------------------------------------%
-% Comparison with experiments
-%----------------------------------------------------------%
-\section{Comparison with experiments}
-
 % Figure

### 2019-10-14 paper2_JNLS/paper.tex
NM

@@
 In fact, the above deterministic  condition directly leads to a simple description for the link.
-The quantity of interest is the outgoing Hamiltonian $\Hdn$, since its value just upstream of the ADC, $t=0^{-}$, is set by the upstream dynamics, and then this particular value is conserved thereafter in the downstream dynamics.
+The quantity of interest is the outgoing Hamiltonian $\Hdn$, since its value just upstream of the ADC, $t=0^{-}$, is set by the incoming dynamics, and then this particular value is conserved thereafter in the outgoing dynamics.
 Since $u$ matches at the ADC, so must $\Hdn$, and, in fact, this matching holds for every individual trajectory.
@@
 \begin{center}
-\includegraphics[width = 0.99 \linewidth]{Figs/trajectories.pdf}
+\includegraphics[width = 0.9 \linewidth]{Figs/trajectories.pdf}
 \caption{
@@
 \end{equation}
-Both the experiments and simulations show the upstream skewness to be negligible compared to the downstream skewness, allowing the term with $\meanup{\Hthree }$ to be dropped in \eqref{statmatchexp}.
+Both the experiments and simulations show the upstream skewness to be negligible compared to its downstream counterpart, allowing the term with $\meanup{\Hthree }$ to be dropped in \eqref{statmatchexp}.
 With this approximation, \eqref{statmatchexp} yields the relationship
@@
 \end{align}
-This equation simply says that the expected value of $\Hthree$ is directly related to the outgoing skewness.
+This equation expresses a directly relationship between the expected value of $\Hthree$ and the outgoing skewness of surface displacement.
 Next, we must convert the $\Htwo$-quantities, which involve surface slope.
@@
 \begin{equation}
+\label{SkewPred}
 \frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 
@@
 This formula links the skewness of the outgoing wave field to the change in the variance of surface slope.
-In particular, it predicts that their ratio to scale as the {\em inverse cube of wave amplitude} and the {\em cube of the depth ratio}, as seen on the right-hand-side of the equation.
-It is a rather unexpected relationship, as it would be difficult to anticipate that the displacement skewness and slope variance, among all possibilities, are so intimately related.
-
-\vsp{3}
-In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
-% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.
+In particular, it predicts that their ratio to scale as the {\em inverse cube of wave amplitude}, $\epsup^{-3}$, and the {\em cube of the depth ratio}, $\dratdn^3$.
+It is a rather unexpected relationship, as it would have been difficult to anticipate that the displacement skewness and slope variance, among all possibly variable combinations, are so intimately related.
 
@@
 \caption{
-Relationship predicted by the theory and confirmed by experiments.
+Experimental confirmation of the explicit formula for outgoing skewness \eqref{SkewPred}.
+(a)--(b) Spatial variation of $\skw(\eta)$ and $\var(\eta_x)$ for several different amplitudes.
+The maximum of $\skw(\eta)$ and the minimum of $\var(\eta_x)$ each occur at a location that is insensitive to amplitude, $x = $15 cm (blue) and -9 cm respectively (green), which are chosen as the representative locations to evaluate \eqref{SkewPred}.
+(c) Measurements of the ratio $\skw(\eta) / \Delta \var(\eta_x)$ from 15 different experiments plotted against the dimensionless wave amplitude $\epsup$.
+For sufficiently large amplitude, the measurements fit closely with the -3 power law predicted by \eqref{SkewPred}.
 }
-\label{AAA}
+\label{SkewRat}
 \end{center}
@@
  %New experimental measurements guided by theory
+
+	Fortunately, it is a relationship that can be tested directly against our experimental measurements, specifically the type of data that was presented in Fig.~\ref{ExpSpatialStats}.
+The the prediction draws our attention to the spatial variation of displacement skewness, which we replot for convenience in Fig.~\ref{SkewRat}(a), and the variance of surface slope, shown in Fig.~\ref{SkewRat}(b).
+For the representative downstream values, $\skwdn(\eta)$ and $\vardn(\eta_x)$, we select the location of the skewness peak, $x=15$ cm downstream of the ADC.
+We must also select an upstream location for $\varup(\eta_x)$.
+Fig.~\ref{SkewRat}(b) suggests that an asymptotic value may be reached far upstream of the ADC, but the data does not extend that far.
+On the other hand, there appears to be a well-defined minimum in $\varup(\eta_x)$ that occurs at the same position for all driving amplitudes.
+We thus select this minimum position, $x = -9$ cm upstream of the ADC to measure $\varup(\eta_x)$.
 

### 2019-10-14 code_julia_sampling/run.jl
NM

@@ -1,13 +1,0 @@
-## RUN STUFF
-include("transfer.jl")
-
-
-
-# Test case with Lambda = 16, nsweeps = 2, lamfac = 8.
-#@time( uniform_sample(16, 2) )
-#@time( transfun("rand-16z-2.jld", 8) )
-#output_text("thvars-16z-8-2.jld")
-
-
-# RUNS
-#@time( transfun("rand-8z-1000.jld", 4) )

### 2019-10-14 code_julia_sampling/run1.jl
NM

@@ -1,0 +1,18 @@
+## RUN STUFF
+include("transfer.jl")
+
+
+
+# Test case with Lambda = 16, nsweeps = 2, lamfac = 8.
+#@time( uniform_sample(16, 2) )
+#@time( transfun("rand-16z-2.jld", 8) )
+#output_text("thvars-16z-8-2.jld")
+
+
+# RUNS
+#@time( transfun("rand-8z-1000.jld", 4) )
+
+ @time( transfun("rand-12z-1000.jld", 8) )
+ @time( transfun("rand-12z-1000.jld", 6) )
+
+ 

### 2019-10-14 code_julia_sampling/run1.jl
NM

@@ -10,9 +10,5 @@ include("transfer.jl")
 
 
 # RUNS
-#@time( transfun("rand-8z-1000.jld", 4) )
-
  @time( transfun("rand-12z-1000.jld", 8) )
  @time( transfun("rand-12z-1000.jld", 6) )
-
- 

### 2019-10-15 paper2_JNLS/paper.tex
NM

@@
 We present an assortment of new experimental measurements and carefully compare them to model predictions.
-We establish a more precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments.
+We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments.
 The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in spectacular fashion by direct experimental measurements.
@@
 Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
-We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory.
-Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
+%We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. 
+Novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
 Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}.
@@
 The phases $\delta_n$ are uniformly distributed random variables, which results in a randomized wave train.
-The standard deviation of the paddle angle, $\Dphi$, controls the amplitude overall amplitude of the waves.
+The standard deviation of the paddle angle, $\Dphi$, controls the overall amplitude of the waves.
 In a single experiment $\Dphi$ is fixed, and we will present a series of experiments with $\Dphi$ varied systematically.
@@
 Both signals exhibit a combination of periodic and random behavior, with the dominant oscillations corresponding to the peak forcing frequency of 2 Hz.
+
 	The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e) on a semi-log scale.
@@
 
-% Figure: Experimental Spatial Statistics
-%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-\begin{figure}%[!ht]
-\begin{center}
-\includegraphics[width = 0.85 \linewidth]{Figs/ExpSpatialStats.pdf}
-\caption{Wave statistics as they vary in space for several different experiments of different driving amplitudes (see legend).
-(a)--(c) Standard deviation, skewness, and kurtosis of the surface displacement $\eta$.
-(d)--(f) The same for surface slope $\eta_x$.
-A reference wave amplitude is set by $\etastd$ which does not vary significantly crossing the ADC.
-Skewness and kurtosis of $\eta$, however, show a strong response to the ADC.
-The surface slope measurements are negatively skewed and also show a large kurtosis downstream of the ADC.
-}
-\label{ExpSpatialStats}
-\end{center}
-\end{figure}
- %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
 	The paddle amplitude in Fig.~\ref{ExpDiagStats} is $\Dphi = 1.38^{\circ}$, and this value was varied systematically in the range $\Dphi = 0.125^{\circ}$--$2^{\circ}$ to probe the different regimes of wave behavior, from linear to strongly nonlinear waves.
 Figure \ref{ExpSpatialStats} shows long-time statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ as they vary in space for six different driving amplitudes (see legend).
+
 	We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
@@
 
+% Figure: Experimental Spatial Statistics
+%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+\begin{figure}%[!ht]
+\begin{center}
+\includegraphics[width = 0.80 \linewidth]{Figs/ExpSpatialStats.pdf}
+\caption{Wave statistics as they vary in space for several experiments of different driving amplitudes (see legend).
+(a)--(c) Standard deviation, skewness, and kurtosis of the surface displacement $\eta$.
+(d)--(f) The same for surface slope $\eta_x$.
+(a) $\etastd$, which sets a scale for wave amplitude, does not vary significantly crossing the ADC.
+(b)--(c) Skewness and kurtosis of $\eta$, however, show a strong response to the ADC.
+The surface slope measurements are negatively skewed and also show a large kurtosis downstream of the ADC.
+}
+\label{ExpSpatialStats}
+\end{center}
+\end{figure}
+ %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
 	Figure \ref{ExpSpatialStats} shows how these statistics vary in the vicinity of the ADC, located at $x = 0$, for both displacement and slope.
@@
 Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC.
-In experiments with sufficiently high driving amplitude, though, the skewness and kurtosis (b)--(c) respond strongly to the ADC.
-In these experiments, both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm.
+The skewness and kurtosis, however, respond strongly to the ADC as long as the amplitude is sufficiently high.
+As seen in \ref{ExpSpatialStats}(b)--(c), both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm.
 Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude.
@@
 As seen in Fig.~\ref{ExpSpatialStats}(a) the standard deviation of slope behaves similar to displacement: $\std(\eta_x)$ increases with driving amplitude, but remains nearly uniform in space for each individual experiment and is not affected by the ADC.
-The higher-order moments, however, once again respond strongly to the ADC if the driving amplitude is sufficiently high.
+Once again though, the higher-order moments respond strongly to the ADC if the driving amplitude is sufficiently high.
 The skewness of slope becomes highly negative and reaches a minimum near $x = 8$ cm, which is downstream of the ADC but upstream of the position where the displacement statistics peak.
@@
 This framework is based on a Galerkin truncation of the variable-depth Korteweg–de Vries (KdV) equation.
-The KdV equation is a commonly excepted model for describing the propagation of unidirectional, shallow-water waves, and it accounts for weak nonlinearity and weak dispersion over long timescales and large spatial scales.
-We will perform Galerkin truncation of this system at wavenumber $\Lambda$ to obtain a finite-dimensional dynamical system that exhibits weak turbulence.
+The KdV equation is a well established model for describing the propagation of unidirectional, shallow-water waves, and it accounts for weak nonlinearity and weak dispersion over long timescales and large spatial scales.
+We will perform Galerkin truncation of KdV to obtain a finite-dimensional dynamical system that exhibits weak turbulence.
 We outline the Hamiltonian structure of both the continuous KdV and the truncated system.
@@
 
-In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the statistical-mechanics framework introduced in \cite{majda2019}.
-Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other choices.
+In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from Majda et al.~2019 \cite{majda2019}.
+Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices.
 To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment.
@@
 &&\mbox{\em dimensionless surface displacement} \\
-&\tilde{x} = \xi / \lengthscale
+&\tilde{x} = (x-ct) / \lengthscale
 &&\mbox{\em dimensionless position (in moving frame)} \\
@@
 
-Now it is possible to choose the scales $\ampscale, \lengthscale, \timescale$ for ease in working with a particular framework. We choose the following scales,
+Now it is possible to choose the scales $\ampscale, \lengthscale, \timescale$ for ease in working with a particular framework. We make the following choices,
 \begin{align}
@@
 Since lower frequencies do exist, strict periodicity is not satisfied, but rather waves may be nearly periodic over the physical domain $\xi \in [-\lam/2, \lam/2]$.
-The approximation of near-periodicity becomes more accurate if integer multiples of $\lam$ are considered, i.e.~$\xi \in [-\lamfac \lam/2, \lamfac \lam/2]$.
-Thus, we have chosen $\lengthscale$ above so that the dimensionless domain of consideration is $\tilde{x} \in [-\pi, \pi]$ and periodic boundary conditions on $u$ can be  imposed on over this domain with accuracy that increases with $\lamfac$.
-Lastly, regarding $\timescale$, the most basic timescale in the experiments is simply $\freqp^{-1}$, i.e.~the period of waves passing a fixed reference point.
+The approximation of near-periodicity becomes more accurate if integer multiples are considered, i.e.~$\xi \in [-\lamfac \lam/2, \lamfac \lam/2]$.
+Thus, we have chosen $\lengthscale$ above so that, over the dimensionless domain $\tilde{x} \in [-\pi, \pi]$, periodic boundary conditions can be  imposed on $u$ with an accuracy that increases with $\lamfac$.
+
+	Lastly, regarding the characteristic timescale $\timescale$, the most basic timescale in the experiments is simply $\freqp^{-1}$, i.e.~the period of waves passing a fixed reference point.
 Of course, the leading-order behavior in shallow water is simple translation of waves with uniform speed $c$.

### 2019-10-15 paper2_JNLS/paper.tex
NM

@@
 \end{align}
-Here, we have chosen the sign of Lax \cite{lax1975periodic}.
+Here, we have chosen the sign convention of Lax \cite{lax1975periodic}.
 More recent work of Bajars et al.~\cite{bajars2013weakly} and Majda et al.~\cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite.
 Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent.
-With their sign convention, Majda et al.~found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}.
+Using the second convention, Majda et al.~found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}.
 We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.
@@
 
-We introduce two important invariants of KdV, the momentum and the energy
+We introduce two important invariants of KdV, namely the momentum and the energy
 \begin{align}
@@
 As indicated above, the momentum of $u$ vanishes since it is measured as displacement from equilibrium.
-Second, by the scale chosen for $\ampscale$ in \eqref{scales}, the energy has been normalized to unity.
-
+Second, due to the choice of $\ampscale$ in \eqref{scales}, the energy has been normalized to unity.
+
+% UNUSED POISSON BRACKET STUFF
+\begin{comment}
 The evolution of any functional $\mathcal{F}[u]$ is given by 
@@
 where again $\Ham = \Hup$ if $t<0$, and $\Ham = \Hdn$ if $t>0$ and $\{\}$ is the Poisson bracket induced by $\sympJ$.
+\end{comment}
 
@@
 Briefly, consider the parameter $\lamfac$, the number of characteristic wavelengths in the physical domain.
-We require $\lamfac \le \Lambda$, so that the mode $\uhat_{\lamfac}$, corresponding to the characteristic wavelength $\lam$ in the experiments, is resolved in the truncated dynamical system.
+We require $1 \le \lamfac \le \Lambda$, so that the mode $\uhat_{\lamfac}$, corresponding to the characteristic wavelength $\lam$ in the experiments, is resolved in the truncated dynamical system.
 If $\lamfac = \Lambda$, then $\lam$ corresponds to the smallest resolved wavelength.
-In contrast, if $\lamfac$ is chosen as an intermediate value, e.g.~$\lamfac = \Lambda/2$ or $\Lambda/4$, then the truncated system \eqref{TKdV} resolves scales that are both bigger and smaller than the characteristic value $\lam$.
+If instead $\lamfac$ is chosen as an intermediate value between 1 and $\Lambda$, then the truncated system will resolve scales that are both bigger and smaller than the characteristic value $\lam$.
 
@@
 \end{equation}
-where $\Hthree$ and $\Htwo$ the defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
+where $\Hthree$ and $\Htwo$ are defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
 Then TKdV \eqref{TKdV} can be expressed as
@@
 % The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
-Thus, the dynamics of interest are confined to the unit hypersphere of $\En = 1$ in $\CC^{\Lambda}$.
-In summary, the TKdV system possesses three important invariants: momentum, energy, and hamiltonian.
+Thus, the dynamics of interest are confined to the unit hypersphere, $\En = 1$ in $\CC^{\Lambda}$.
+In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian.
 The untruncated KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}, but their truncated counterparts are not generally invariants of TKdV.
@@
 Meanwhile, evolution over shorter timescales is simply described by linear theory.
-The event of a wave crossing the ADC is precisely such a short-time event, and so we will employ matching conditions at the ADC that result from linear theory.
+The event of a wave crossing the ADC is precisely such a short-time event, and so we will employ the same matching conditions at the ADC that would result from linear theory.
 
@@
 
-	This statistical matching condition imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
+	This statistical matching condition, originally derived by Majda et al., imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
 In particular, we view $\thup$ as given by the random-state of the incoming wave field.
@@
 	We note that the experiments discussed here preceded the development of the theory, and hence had no intent of mimicking a Gibbs measure in the wave forcing.
-We expect that if this forcing were designed to mimic the Gibbs measure, in particular the spectral decay, then it would be more straightforward to assign a value to the model parameter $\lamfac$.
+We expect that if the experimental forcing were designed to mimic the Gibbs measure, in particular the spectral decay, then it would be more straightforward to assign a value to the model parameter $\lamfac$.
 We, however, leave that task for future research due to the significant cost of performing an entirely new set of experiments compared to the relative ease and great value in re-analyzing existing experimental measurements in light of the new theoretical developments.
@@
 In line with experimental observations, the incoming skewness is small, while the outgoing wave skewness is much higher.
-Specifically, to capture the experimentally observed peak skewness range of 0.6-0.9 seen for the larger amplitudes in Fig.~\ref{ExpSpatialStats}(b), requires selecting $\thup$ in the range 10--20 (?).
+Specifically, to capture the experimentally observed peak skewness range of 0.6-0.9 seen for the larger amplitudes in Fig.~\ref{ExpSpatialStats}(b), requires selecting $\thup$ in the range 10--25.
 
@@
 \end{equation}
-So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode ($k = 1$),  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.
+So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode of $k = 1$,  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.
 
@@
 Thus, the theoretically predicted jump in kurtosis, while not as extreme as that measured in experiments, is quite significant.
-These elevated levels of kurtosis give the distributions the look of being stretched in the tails and pointy near the centers (especially in the experiments).
+These elevated levels of kurtosis are associated with the distinct appearance of the distribution, having flat tails and somewhat sharp centers.
 
@@
 In particular, one large wave is seen to propagate from the lower-left corner of the figure to the upper right.
-A bias towards positive surface displacement is visually apparent in these downstream dynamics, without even viewing the histograms.
+A bias towards positive surface displacement is apparent in these downstream dynamics, without even viewing the histograms.
 
@@
 
-	We remind the reader that the KdV framework tracks the long-time evolution waves in a reference frame moving with the characteristic speed $c = \sqrt{g \depth}$ from linear theory.
+	We remind the reader that the KdV framework tracks the long-time evolution of waves in a reference frame moving with the characteristic speed $c = \sqrt{g \depth}$ from linear theory.
 Hence, the above discussion of left-or-right going waves cannot be interpreted in the context of the experiments, which take a fixed reference frame, without an appropriate Galilean transformation.
 Hence, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed that is near $c$.
-The directions and speeds calculated by the TKdV framework simply determines how much the actually speed of waves deviates from $c$.
+The directions and speeds calculated by the TKdV framework simply determines how much the true wave speed deviates from $c$.
 
@@
 Given the wave speeds ($c = $ 110 cm/s upstream and 54 cm/s downstream), these timescales can be converted to distances traveled by the waves.
-In fact, a simple calculation using the definitions in Section \ref{theory} gives the distance traveled as $7 \lamfac g/(2 \pi \freqp^2)$.
-Notice the depth dependence has serendipitously canceled.
+In fact, a simple calculation using the definitions in Section \ref{theory} gives the distance traveled as $7 \lamfac g/(2 \pi \freqp^2)$, which, due to a serendipitous cancellation, is independent of depth.
 Hence, the distance traveled by the waves over a dimensionless time of $t=7$ is the same value upstream and downstream,  roughly 22 meters.
@@
 
-	These estimates raise an important question: if a distance of 22 meters is required for significant wave evolution under the KdV framework, how is it that the experiments exhibit substantial changes in wave statistics a much shorter distance downstream of the ADC?
+	These estimates raise an important question: if a distance of 22 meters is required for significant wave evolution under the KdV framework, how do the experiments exhibit substantial changes in wave statistics a much shorter distance downstream of the ADC?
 The results shown in Fig.~\ref{skew-kurt} help resolve this question.
 This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same TKdV deterministic simulations as pictured in Fig.~\ref{trajectories}.
-The bold black curve shows the ensemble mean (ensemble size of 1000), while the skewness and kurtosis of 100 individual trajectories are shown by the faint gray curves to give a sense for the variance involved.
+The bold black curve shows the ensemble mean (ensemble size of 1000), while the faint gray curves show the skewness and kurtosis of 100 individual trajectories to give a sense for the variation involved.
 Importantly, the skewness and kurtosis evolve on a much shorter timescale than the aforementioned $t=7$.
-More precisely, skewness and kurtosis have already saturated to their asymptotic by $t = 1$, and reach half of that value by $t = 0.18$.
+More precisely, skewness and kurtosis have already saturated to their asymptotic value by $t = 1$, and reach half of that value by $t = 0.18$.
 These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order as the relevant distances in the experiments.
-Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter the statistical distributions, and this reconfiguration occurs on a much shorter timescale than that required for waves to cross the entire periodic domain.
+Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter its statistical distributions, and this reconfiguration occurs on a much shorter timescale than that required for waves to cross the entire periodic domain.
 
@@
 
-This section discusses perhaps the most important single result of the manuscript.
+We now discuss in this section arguably the most significant single result of the manuscript.
 In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}.
-More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$, in particular how that ratio depends on driving amplitude and depth ratio.
+More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$.
 Below, we recap the derivation of this formula and then test the prediction against our direct experimental measurements.

### 2019-10-15 paper2_JNLS/paper.tex
NM

@@
 %----------------------------------------------------------------------------%
-\section{Predicted outgoing skewness and experimental confirmation}
-
-We now discuss in this section arguably the most significant single result of the manuscript.
+\subsection{Predicted outgoing skewness and experimental confirmation}
+
+This section discusses arguably the most novel single result of the manuscript.
 In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}.
 More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$.
-Below, we recap the derivation of this formula and then test the prediction against our direct experimental measurements.
-
-\subsection{Explicit formula for outgoing skewness}
+Below, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.
+
+%\subsection{Explicit formula for outgoing skewness}
 
@@
 	Our next task is to convert this formula to physical variables so that it can be tested against experimental data.
-First, following the definition of $\Hthree$ in Eq.~\eqref{H3H2} and the relationship, $u = \eta/(\pi^{1/2} \etastd)$, a straightforward calculation gives
+First, following the definition of $\Hthree$ in Eq.~\eqref{H3H2} and the definition, $u = \eta/(\pi^{1/2} \etastd)$, a straightforward calculation gives
 \begin{align}
@@
 \end{align}
-This equation expresses a directly relationship between the expected value of $\Hthree$ and the outgoing skewness of surface displacement.
+This equation expresses a direct relationship between the expected value of $\Hthree$ and the outgoing skewness of surface displacement.
 Next, we must convert the $\Htwo$-quantities, which involve surface slope.
@@
 This formula links the skewness of the outgoing wave field to the change in the variance of surface slope.
-In particular, it predicts that their ratio to scale as the {\em inverse cube of wave amplitude}, $\epsup^{-3}$, and the {\em cube of the depth ratio}, $\dratdn^3$.
+In particular, it predicts their ratio to scale as the {\em inverse cube of wave amplitude}, $\epsup^{-3}$, and the {\em cube of the depth ratio}, $\dratdn^3$.
 It is a rather unexpected relationship, as it would have been difficult to anticipate that the displacement skewness and slope variance, among all possibly variable combinations, are so intimately related.
@@
 (a)--(b) Spatial variation of $\skw(\eta)$ and $\var(\eta_x)$ for several different amplitudes.
-The maximum of $\skw(\eta)$ and the minimum of $\var(\eta_x)$ each occur at a location that is insensitive to amplitude, $x = $15 cm (blue) and -9 cm respectively (green), which are chosen as the representative locations to evaluate \eqref{SkewPred}.
+The maximum of $\skw(\eta)$ and the minimum of $\var(\eta_x)$ each occur at a location that is insensitive to amplitude, $x = $15 cm (blue) and -9 cm (green) respectively, which are chosen as the representative locations to evaluate \eqref{SkewPred}.
 (c) Measurements of the ratio $\skw(\eta) / \Delta \var(\eta_x)$ from 15 different experiments plotted against the dimensionless wave amplitude $\epsup$.
-For sufficiently large amplitude, the measurements fit closely with the -3 power law predicted by \eqref{SkewPred}.
+For sufficiently large amplitude, the measurements fit closely with the $\epsup^{-3}$ prediction from \eqref{SkewPred}.
 }
@@
 
-	Fortunately, it is a relationship that can be tested directly against our experimental measurements, specifically the type of data that was presented in Fig.~\ref{ExpSpatialStats}.
-The the prediction draws our attention to the spatial variation of displacement skewness, which we replot for convenience in Fig.~\ref{SkewRat}(a), and the variance of surface slope, shown in Fig.~\ref{SkewRat}(b).
-For the representative downstream values, $\skwdn(\eta)$ and $\vardn(\eta_x)$, we select the location of the skewness peak, $x=15$ cm downstream of the ADC.
-We must also select an upstream location for $\varup(\eta_x)$.
-Fig.~\ref{SkewRat}(b) suggests that an asymptotic value may be reached far upstream of the ADC, but the data does not extend that far.
-On the other hand, there appears to be a well-defined minimum in $\varup(\eta_x)$ that occurs at the same position for all driving amplitudes.
-We thus select this minimum position, $x = -9$ cm upstream of the ADC to measure $\varup(\eta_x)$.
+	Fortunately, it is a relationship that can be tested directly against the experimental measurements, specifically the type of data that was presented in Fig.~\ref{ExpSpatialStats}.
+The the prediction draws focus to the spatial variation of displacement skewness, which we replot for convenience in Fig.~\ref{SkewRat}(a), and the variance of surface slope, shown in Fig.~\ref{SkewRat}(b).
+To test \eqref{SkewPred}, we must select a representative downstream position to evaluate $\skwdn(\eta)$ and $\vardn(\eta_x)$, and a representative upstream position for $\varup(\eta_x)$.
+In short, we select the same representative locations as were used to evaluate the histograms in Figs.~\ref{ExpDiagStats}, \ref{uhist}, and \ref{slopehist}, namely $x = 15$ cm downstream and $x = -9$ cm upstream.
+These positions are indicated by the blue and green vertical dashed lines in Fig.~\ref{SkewRat}(a)--(b).
+Recall that $x = 15$ cm corresponds to the peak of the downstream skewness and kurtosis.
+Meanwhile, we see in Fig.~\ref{SkewRat}(b) that $x = -9$ cm corresponds to a slight dip in $\var(\eta_x)$.
+
+	With these locations selected, we evaluate the ratio $\skw(\eta) / \Delta \var(\eta_x)$ for each experiment and plot the result against dimensionless wave amplitude, $\epsup = \etastd / \dup$ on a log-log scale in Fig.~\ref{SkewRat}(c).
+While the data shows significant variation for small amplitudes, it shows a well defined decreasing trend at larger amplitudes.
+In particular, the red dashed line shows the power-law $\epsup^{-3}$ predicted by \eqref{SkewPred}, which captures the experimental trend remarkably well.
+
+ %----------------------------------------------------------------------------%
+% Conclusions 
+%----------------------------------------------------------------------------%
+\section{Concluding remarks}
 

### 2019-10-15 paper2_JNLS/paper.tex
NM

@@
 Next, we must convert the $\Htwo$-quantities, which involve surface slope.
-Here, it is important to remember that, since it is the {\em downstream} Hamiltonian that is matched in \eqref{statmatch}, one must use the {\em downstream} characteristic scale $\lengthscale_+$ to convert back into physical variables.
 Using the definition of $\Htwo$ in Eq.~\eqref{H3H2} and $\lengthscale_+$ in Eq.~\eqref{scales} gives
@@
 \end{align}
+We note that the downstream scale $\lengthscale_+$ is used in this conversion, since it is the downstream Hamiltonian that is matched in \eqref{statmatch}.
+This formula links the expected value of $\Htwo$ to the variance of surface slope.
+
 Inserting \eqref{H3convert} and \eqref{H2convert} into \eqref{H3H2ratio}, and using the definitions of $C_2$ and $C_3$ from \eqref{C3C2}, gives the remarkably explicit formula

### 2019-10-15 paper2_JNLS/paper.tex
NM

@@
 
-\title{Anomalous waves triggered by abrupt depth changes: experiments and truncated KdV theory}
+\title{Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV theory}
 
@@
 Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
-%We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. 
+We first provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two.
+Next, we present new experimental data, most notably on {\em surface slope} statistics, which motivates parallel examination of the theory.
+Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
+Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}.
+
+% OLD VERSION
+\begin{comment}
+We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. 
 Novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
 Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}.
-In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two.
+In addition, we provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two. 
 In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.
+\end{comment}
 
@@
 \caption{
-(a) Experimental schematic. 
-(b)--(c) Surface displacement measured at a representative upstream and downstream locations. (d)--(e) Corresponding histograms. Figure adapted from \cite{bolles2019}.}
+(a) Experimental schematic: randomized waves are generated by a pivoting paddle and propagate over a step in bottom topography.
+(b)--(c) Surface displacement measured at representative locations upstream and downstream of the abrupt depth change (ADC).
+(d)--(e) Corresponding histograms showing symmetric upstream statistics and highly skewed downstream statistics.
+Figure adapted from \cite{bolles2019}.}
 \label{ExpDiagStats}
@@
 (b)--(c) Skewness and kurtosis of $\eta$, however, show a strong response to the ADC.
-The surface slope measurements are negatively skewed and also show a large kurtosis downstream of the ADC.
+(d)--(f) The measurements of surface slope, $\eta_x$, are negatively skewed and also exhibit large kurtosis downstream of the ADC.
 }
@@
 The quadratic component, however, requires closer inspection.
-Due to Parseval's identity, $\Htwo$ is given by
+Due to Parseval's identity, $\Htwo$ can be written as
 \begin{equation}
@@
 \label{HtwoExpect}
-\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} \abs{k}^ 2 \mean{\abs{\uhat_k}^2}_0 \sim \frac{1}{3} \Lambda^2
+\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} {k}^ 2 \mean{\abs{\uhat_k}^2}_0 \sim \frac{1}{3} \Lambda^2
 \end{equation}
@@
 \begin{equation}
-\sum_{k=1}^{n} \abs{k}^ 2 = \frac{1}{6} n(n+1)(2n+1) \approx n^3/3
-\end{equation}
-
-	Importantly, \eqref{HtwoExpect} shows that the expected value of $\Htwo$ grows like $\Lambda^2$, which appears to be problematic for obtaining independence as $\Lambda \to \infty$.
+\sum_{k=1}^{n} {k}^ 2 = \frac{1}{6} n(n+1)(2n+1) \approx \frac{1}{3} n^3
+\end{equation}
+
+	Importantly, \eqref{HtwoExpect} shows that the expected value of $\Htwo$ grows like $\Lambda^2$, which appears problematic for obtaining independence as $\Lambda \to \infty$.
 However, $\Htwo$ enters $\Ham$ in product with the coefficient $C_2$, which itself scales as $C_2 \sim \lamfac^{-2}$.
@@
 \end{equation}
-So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode of $k = 1$,  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.
+So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode, $k = 1$,  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.
 
@@
 Thus, the theoretically predicted jump in kurtosis, while not as extreme as that measured in experiments, is quite significant.
-These elevated levels of kurtosis are associated with the distinct appearance of the distribution, having flat tails and somewhat sharp centers.
-
-	Distinct differences in the shapes of the downstream slope distributions, however, are visible.
+These elevated levels of kurtosis are associated with the distinct appearance of the distributions, most notably the flat tails that appear in both the experimental and theoretical distributions, and also the pointy center that is most pronounced in the experiments.
+
+	Distinct differences in the shapes of the downstream distributions, however, are also visible.
 While the theoretical distributions remain symmetric, the experimental distributions skew towards negative slope.
@@
 %----------------------------------------------------------------------------%
-\subsection{Wave dynamics}
+\subsection{Wave dynamics and analysis of time scales}
 
@@
 Upstream of the ADC, since $\drat = 1$, the nonlinear effects are relatively weak and the dynamics are dominated by dispersion.
-Linearizing Eq.~\eqref{TKdV} and introducing the ansatz $u_k = e^{i(kx - \omega_k t)}$ produces the dispersion relation for angular frequency $\omega_k = -C_2 k^3$, or for phase velocity $c_k = \omega_k/k = -C_2 k^2$ \cite{majdaqi2019}.
+Linearizing \eqref{TKdV} and introducing the ansatz $u_k = e^{i(kx - \omega_k t)}$ produces the dispersion relation for angular frequency $\omega_k = -C_2 k^3$, or for phase velocity $c_k = \omega_k/k = -C_2 k^2$ \cite{majdaqi2019}.
 In particular, the phase velocity is strictly negative, implying that waves can only propagate leftwards, as is consistent with Fig.~\ref{trajectories}(a).
@@
 	We remind the reader that the KdV framework tracks the long-time evolution of waves in a reference frame moving with the characteristic speed $c = \sqrt{g \depth}$ from linear theory.
-Hence, the above discussion of left-or-right going waves cannot be interpreted in the context of the experiments, which take a fixed reference frame, without an appropriate Galilean transformation.
-Hence, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed that is near $c$.
-The directions and speeds calculated by the TKdV framework simply determines how much the true wave speed deviates from $c$.
+Hence, the above discussion of left-or-right going waves cannot be interpreted in the context of the experiments without an appropriate Galilean transformation.
+Hence, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed near $c$.
+The directions and speeds calculated by the TKdV framework simply quantify the deviation of the true wave speed from $c$.
 
@@
 The results shown in Fig.~\ref{skew-kurt} help resolve this question.
-This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same TKdV deterministic simulations as pictured in Fig.~\ref{trajectories}.
+This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same deterministic TKdV simulations as pictured in Fig.~\ref{trajectories}.
 The bold black curve shows the ensemble mean (ensemble size of 1000), while the faint gray curves show the skewness and kurtosis of 100 individual trajectories to give a sense for the variation involved.
@@
 These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order as the relevant distances in the experiments.
-Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter its statistical distributions, and this reconfiguration occurs on a much shorter timescale than that required for waves to cross the entire periodic domain.
+Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter its statistical distributions.
+The timescale for this reconfiguration is much shorter than the time required for waves to cross the entire periodic domain.
 
@@
 %----------------------------------------------------------------------------%
-\subsection{Predicted outgoing skewness and experimental confirmation}
-
-This section discusses arguably the most novel single result of the manuscript.
+\subsection{Explicit formula for outgoing skewness and experimental confirmation}
+
+We now discuss arguably the most novel single result of the manuscript.
 In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}.
 More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$.
-Below, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.
-
-%\subsection{Explicit formula for outgoing skewness}
+Here, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.
 
@@
 	Our next task is to convert this formula to physical variables so that it can be tested against experimental data.
-First, following the definition of $\Hthree$ in Eq.~\eqref{H3H2} and the definition, $u = \eta/(\pi^{1/2} \etastd)$, a straightforward calculation gives
+First, following the definition of $\Hthree$ in \eqref{H3H2} and the definition, $u = \eta/(\pi^{1/2} \etastd)$, a straightforward calculation gives
 \begin{align}
@@
 Next, we must convert the $\Htwo$-quantities, which involve surface slope.
-Using the definition of $\Htwo$ in Eq.~\eqref{H3H2} and $\lengthscale_+$ in Eq.~\eqref{scales} gives
+Using the definition of $\Htwo$ in \eqref{H3H2} and $\lengthscale_+$ in \eqref{scales} gives
 \begin{align}
@@
 	Fortunately, it is a relationship that can be tested directly against the experimental measurements, specifically the type of data that was presented in Fig.~\ref{ExpSpatialStats}.
-The the prediction draws focus to the spatial variation of displacement skewness, which we replot for convenience in Fig.~\ref{SkewRat}(a), and the variance of surface slope, shown in Fig.~\ref{SkewRat}(b).
+Formula \eqref{SkewPred} draws focus to the spatial variation of displacement skewness, which we replot for convenience in Fig.~\ref{SkewRat}(a), and the variance of surface slope, shown in Fig.~\ref{SkewRat}(b).
 To test \eqref{SkewPred}, we must select a representative downstream position to evaluate $\skwdn(\eta)$ and $\vardn(\eta_x)$, and a representative upstream position for $\varup(\eta_x)$.
-In short, we select the same representative locations as were used to evaluate the histograms in Figs.~\ref{ExpDiagStats}, \ref{uhist}, and \ref{slopehist}, namely $x = 15$ cm downstream and $x = -9$ cm upstream.
+In short, we select the same representative locations  used to evaluate the histograms in Figs.~\ref{ExpDiagStats}, \ref{uhist}, and \ref{slopehist}, namely $x = 15$ cm downstream and $x = -9$ cm upstream.
 These positions are indicated by the blue and green vertical dashed lines in Fig.~\ref{SkewRat}(a)--(b).
@@
 	With these locations selected, we evaluate the ratio $\skw(\eta) / \Delta \var(\eta_x)$ for each experiment and plot the result against dimensionless wave amplitude, $\epsup = \etastd / \dup$ on a log-log scale in Fig.~\ref{SkewRat}(c).
-While the data shows significant variation for small amplitudes, it shows a well defined decreasing trend at larger amplitudes.
+While the data shows significant variation at small amplitudes, it shows a well defined decreasing trend at larger amplitudes.
 In particular, the red dashed line shows the power-law $\epsup^{-3}$ predicted by \eqref{SkewPred}, which captures the experimental trend remarkably well.
@@
 \section{Concluding remarks}
+
+	This manuscript extends the parallel experimental and modeling efforts of Bolles et al.~\cite{bolles2019} and Majda et al.~\cite{majda2019} concerning the emergence of anomalous wave statistics from abrupt changes in bottom topography.
+The theoretical framework is based on deterministic and statistical analysis of the TKdV system, with exploitation of the Hamiltonian structure and the associated invariant Gibbs measures.
+The theory depends crucially on matching the incoming and outgoing invariant measures at the abrupt depth change, so as to link the outgoing dynamics to the incoming ones.
+Throughout, we have emphasized the synergy between the experiments and theory, with experimental data informing theoretical advancements and model predictions motivating new experimental measurements.
+
+	Careful calibration of the inverse temperature against the experimental data allowed for a detailed comparison between experimental and theoretically predicted wave statistics.
+The outgoing distributions of surface displacement predicted by the TKdV framework capture in remarkable detail the experimental measurements.
+We have extended this statistical analysis to surface slope, a line of inquiry motivated by the importance of $\Htwo$, the slope standard deviation, in the theory.
+The comparison of slope statistics shows intriguing similarities, while also some quantitative differences in the shape and skewness of the outgoing distributions.
+
+	Finally, Majda et al.~\cite{majda2019} derived an explicit formula for the skewness of the outgoing wave field, which we have tested against direct experimental measurements.
+Specifically, the formula predicts how the ratio of displacement skewness to slope variance, $\skw(\eta) / \Delta \var(\eta_x)$, depends on wave amplitude.
+The experimental measurements conclusively confirm the inverse-cube dependence on wave amplitude, thus highlighting the predictive power of the TKdV statistical mechanics framework.
+
+	We note that since the ratio, $\skw(\eta) / \Delta \var(\eta_x)$, exhibits such a clear dependence on input parameters, it could prove a useful diagnostic for anomalous wave observations in the ocean.
+In particular, this quantity seems to provide a signature for anomalous waves that are triggered by abrupt depth changes.
+Examination of this ratio in field data is an exciting avenue for future research.
 

### 2019-10-15 paper2_JNLS/paper.tex
NM

@@
 
-\title{Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV theory}
+\title{Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics}
 
@@
 A recent series of laboratory \cite{bolles2019} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
-Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous wave statistics.
+Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous waves.
 In particular, the more complex mechanism of focusing by 2D bottom topography absent.
@@
 We consider the surface displacement $\eta(x,t)$ of unidirectional, shallow-water waves in a reference frame moving with the characteristic wave speed, $\xi = x - ct$.
-Here, $c = \sqrt{g \depth}$ is leading-order approximation to the wave speed (i.e.~from linear theory), where $g$ gravity and $\depth$ is the local depth.
+Here, $c = \sqrt{g \depth}$ is leading-order approximation to the wave speed (i.e.~from linear theory), where $g$ is gravity and $\depth$ is the local depth.
 The leading-order dynamics are corrected to first order in small amplitude by the Korteweg–de Vries equation (KdV), which in dimensional form is given by \cite{whitham2011linear}
@@
 
-With the above choices, the dimensionless KdV takes the form
+With the above choices, the dimensionless KdV equation takes the form
 \begin{align}
@@
 Thus, the ADC is met at some time $T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
-Therefore, we can regard $\drat$ is a piece-wise constant function of dimensionless time
+Therefore, we can regard $\drat$ is a piece-wise-constant function of dimensionless time
 \begin{equation}
@@
 
-The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalization values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity implies
+The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalized values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity implies
 \begin{equation}
@@
 In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian.
-The untruncated KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}, but their truncated counterparts are not generally invariants of TKdV.
+The untruncated KdV equation possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}, but their truncated counterparts are not generally invariants of TKdV.
 
@@
 
-	This statistical matching condition, originally derived by Majda et al., imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
+	This statistical matching condition, originally derived by Majda et al.~(2019) \cite{majda2019}, imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
 In particular, we view $\thup$ as given by the random-state of the incoming wave field.
@@
 
-	Since the inverse temperature, $\theta$, is a key modeling parameter, it would be highly desirable to normalize the system such that the value of $\theta$ does not depend sensitively on the truncation number, $\Lambda$, as $\Lambda$ grows large.
+	Since the inverse temperature, $\theta$, is a key modeling parameter, it would be highly desirable to normalize the system such that the value of $\theta$ does not depend sensitively on the truncation index, as $\Lambda$ grows large.
 That way, $\theta$ can be interpreted as a real physical parameter that can be linked to the experiments and whose value does not depend sensitively on where one chooses to truncate the system.
@@
 With $(\Lambda, \lamfac) = (16, 8)$ fixed, Fig.~\ref{transfig}(c) indicates $\thup = 20$ as a reasonable value to attain the desired skewness.
-We thus run the deterministic TKdV simulations these model parameters and with $\epsup$, $\delup$, and $\dratdn$ set to the representative values in Table \ref{paramtable}.
+We thus run the deterministic TKdV simulations with these model parameters and with $\epsup$, $\delup$, and $\dratdn$ set to the representative values in Table \ref{paramtable}.
 We run the simulations to a sufficiently long dimensionless time of $t_{f} = 10$, with $dt = 5 \times 10^{-4}$ and with $10^3$ trajectories sampled from the upstream Gibbs measure $\Gup$.
@@
 
-	Distinct differences in the shapes of the downstream distributions, however, are also visible.
+	Distinct differences in the shapes of these downstream distributions, however, are also visible.
 While the theoretical distributions remain symmetric, the experimental distributions skew towards negative slope.
@@
 Hence, the above discussion of left-or-right going waves cannot be interpreted in the context of the experiments without an appropriate Galilean transformation.
-Hence, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed near $c$.
+Thus, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed near $c$.
 The directions and speeds calculated by the TKdV framework simply quantify the deviation of the true wave speed from $c$.
@@
 These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order as the relevant distances in the experiments.
-Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter its statistical distributions.
-The timescale for this reconfiguration is much shorter than the time required for waves to cross the entire periodic domain.
+Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter its statistical distributions, and the timescale for this reconfiguration is much shorter than the time required for waves to cross the entire periodic domain.
 
@@
 The theoretical framework is based on deterministic and statistical analysis of the TKdV system, with exploitation of the Hamiltonian structure and the associated invariant Gibbs measures.
-The theory depends crucially on matching the incoming and outgoing invariant measures at the abrupt depth change, so as to link the outgoing dynamics to the incoming ones.
+The theory depends crucially on matching the incoming and outgoing invariant measures at the abrupt depth change, so as to link the incoming and outgoing states.
 Throughout, we have emphasized the synergy between the experiments and theory, with experimental data informing theoretical advancements and model predictions motivating new experimental measurements.

### 2019-10-16 paper2_JNLS/paper.tex
Updated Fig 3 on transfer function with better resolution in the sampling.

@@
 
-	Figures \ref{transfig}(a)-(b) show the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $0 \le \thup \le 20$.
+	Figures \ref{transfig}(a)-(b) show the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $0 \le \thup \le 25$.
 Figure \ref{transfig}(a) shows the cases $\Lambda = $ 12, 16, and 20 with $\lamfac = 8$ fixed in each.

### 2019-10-22 paper2_JNLS/paper.tex
NM

@@
 Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous waves.
-In particular, the more complex mechanism of focusing by 2D bottom topography absent.
+In particular, the more complex mechanism of focusing by 2D bottom topography is  absent.
 The studies thus offer a paradigm system, with emergent anomalous statistical properties similar to that seen in more complex systems, but in a tractable context that is amenable to analysis.
@@
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
-In subsequent work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated systems that were shown to exhibit many of the same anomalous properties while enjoying a more tractable, lower-dimensional structure.
+In subsequent work, Majda and Qi analyzed more severely truncated systems --- as low as two interacting modes --- that exhibit a statistical phase transition to anomalous statistics and the creation of extreme events while enjoying a more tractable structure \cite{majdaqi2019}.
 
@@
 Most often, we will consider waves moving into shallower depth, so that $\dup > \ddn$.
+Throughout this paper, we will use the subscript `-' to represent upstream variables and `+' for downstream variables.
 
 In the experiments, the randomized incoming wave-field is generated with a peak forcing frequency of $\freqp = 2$ Hz, which gives rise to the characteristic wavelength of $\lam = c/\freqp = \sqrt{g \depth} / \freqp$.
-Note that both the characteristic wave speed $c = c_{\pm}$ and wavelength $\lam = \lamupdn$ each take different values upstream and downstream of the ADC.
-We remark that experimental measurements indicate that $\etastd$ is nearly the same on both values of the ADC.
+Note that both the characteristic wave speed $c = c_{\pm}$ and wavelength $\lam = \lamupdn$ take different values upstream and downstream of the ADC.
+We remark that experimental measurements indicate that $\etastd$ is nearly the same on both sides of the ADC.
 Hence, we will not distinguish between upstream and downstream values of $\etastd$.
@@
 Thus, the ADC is met at some time $T_{ADC}$, and for simplicity we set $T_{ADC} = 0$.
-Therefore, we can regard $\drat$ is a piece-wise-constant function of dimensionless time
+Therefore, we can regard $\drat$ as a piece-wise-constant function of dimensionless time
 \begin{equation}
@@
 
-Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019}.
+Remarkably, the TKdV system \eqref{TKdV} retains the Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019}.
 The piecewise defined Hamiltonian for TKdV is given by

### 2019-10-22 paper2_JNLS/paper.tex
NM

@@
 	Now, from the perspective of statistical mechanics, consider the communication between the statistical ensembles, $\Gupdn$, upstream and downstream of the ADC.
-These two systems are in contact at the ADC, and so the upstream state $\Gup$ can be regarded as a thermal reservoir that influences the downstream state $\Gdn$.
+These two systems are in contact at the ADC, and so the upstream state with distribution $\Gup$ can be regarded as a thermal reservoir that influences the downstream distribution $\Gdn$.
 In fact, the above deterministic  condition directly leads to a simple description for the link.
@@
 	The invariant measure exhibits the proportionality $ \Gibbs_{\Lambda} \propto \exp(-\theta_{\Lambda} \Ham_{\Lambda})$, where we have made explicit the dependence of all quantities on $\Lambda$.
-In particular, if $\Ham_{\Lambda}$ where to depend sensitively on $\Lambda$ in expectation, then $\theta_{\Lambda}$ would need to compensate in order to produce the same invariant measure.
+In particular, if $\Ham_{\Lambda}$ were to depend sensitively on $\Lambda$ in expectation, then $\theta_{\Lambda}$ would need to compensate in order to produce the same invariant measure.
 Hence, it would be desirable to scale the system in such a way that $\Ham_{\Lambda}$ does not depend sensitively on $\Lambda$, at least in expectation.
@@
 Recall that $\Ham_{\Lambda}$ is composed of the cubic and quadratic components $\Hthree$ and $\Htwo$.
-Due to the odd symmetry of $\Hthree$, it is easy to see that $\meanz{\Hthree} = 0$.
+Due to the odd symmetry of $\Hthree$, it is easy to see that $\meanz{\Hthree} = 0$ with respect to the uniform measure $\Gz$.
 The quadratic component, however, requires closer inspection.

### 2019-10-29 paper2_JNLS/paper.tex
NM

@@
 
+\nick{ Also include machine-learning strategies \cite{MachineLearning2019}. }
+
 The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}.
@@
 \caption{
-Comparison of slope histograms.
-The TKdV simulations show similar slope distributions upstream, but visual distributions in the downstream distributions are visible.
-While the simulations recover the spread of the downstream slope statistics, they do not capture the asymmetry seen in the experiments.
+Comparison of slope histograms. The TKdV simulations show similar slope distributions upstream, but visual distributions in the downstream distributions are visible. 
+While the simulations recover the spread of the downstream slope statistics, they do not capture the asymmetry seen in the experiments \nick{to changes}.
 }

### 2019-11-06 paper2_JNLS/paper.tex
NM

--- 2019-10-22 paper2_JNLS/paper.tex
@@
 and the Benjamin-Feir deep-water modulational instability \citep{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
-The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.
-In this way, it makes sense to approach anomalous waves from the broader perspective of turbulent dynamical systems \cite{sapsis2013a, sapsis2013b, sapsis2013blending, chen2016filtering, majda2016introduction, macedo2017universality, MajdaQiSIAM2018, blonigan2019extreme, guth2019machine}.
+The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these events occur much more frequently when surface statistics deviate from Gaussian.
+In this way, anomalous waves can be approached from the broader perspective of turbulent dynamical systems \cite{sapsis2013a, sapsis2013b, sapsis2013blending, chen2016filtering, majda2016introduction, macedo2017universality, MajdaQiSIAM2018, blonigan2019extreme, guth2019machine}.
 
@@
 In subsequent work, Majda and Qi analyzed more severely truncated systems --- as low as two interacting modes --- that exhibit a statistical phase transition to anomalous statistics and the creation of extreme events while enjoying a more tractable structure \cite{majdaqi2019}.
+More recently, Qi and Majda demonstrated the capability of machine learning strategies to accurately predict these extreme events \cite{MachineLearning2019}.
+That work uses a deep neural network coupled with a judicious choice of the entropy loss function that emphasizes the dominant structures of the turbulent field \cite{MachineLearning2019}.
+It should be noted that, in these studies, the extreme events can occur intermittently and on a relatively frequent basis, in contrast to the simpler situation of isolated, rare events \cite{guth2019machine}.
 
@@
 
-	We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
-\begin{align}
-%& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
-& \std(q) = q_{std}= \sqrt{\mean{q^2}}
-&&\mbox{\em standard deviation} \\
-& \skw(q) = {\mean{q^3}} / {q_{std}^3}	
-&&\mbox{\em skewness} \\
-& \kurt(q) = {\mean{q^4}} / {q_{std}^4} - 3
-&&\mbox{\em (excess) kurtosis}
-\end{align}
-where $\mean{}$ indicates a mean (here a long-time mean). Hereafter, we will simply refer to the excess kurtosis as kurtosis.
-
 % Figure: Experimental Spatial Statistics
@@
 \begin{center}
-\includegraphics[width = 0.80 \linewidth]{Figs/ExpSpatialStats.pdf}
+\includegraphics[width = 0.8 \linewidth]{Figs/ExpSpatialStats.pdf}
 \caption{Wave statistics as they vary in space for several experiments of different driving amplitudes (see legend).
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
+
+	We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
+\begin{align}
+%& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
+& \std(q) = q_{std}= \sqrt{\mean{q^2}}
+&&\mbox{\em standard deviation} \\
+& \skw(q) = {\mean{q^3}} / {q_{std}^3}	
+&&\mbox{\em skewness} \\
+& \kurt(q) = {\mean{q^4}} / {q_{std}^4} - 3
+&&\mbox{\em (excess) kurtosis}
+\end{align}
+where $\mean{}$ indicates a mean (here a long-time mean). Hereafter, we will simply refer to the excess kurtosis as kurtosis.
 
@@
 
-	Figures \ref{transfig}(a)-(b) show the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $0 \le \thup \le 25$.
-Figure \ref{transfig}(a) shows the cases $\Lambda = $ 12, 16, and 20 with $\lamfac = 8$ fixed in each.
-In this figure, the transfer function changes significantly with $\Lambda$.
-Figure~\ref{transfig}(b) shows the same but with the scaling $\lamfac = \Lambda/2$ that was argued in the previous section.
-In this figure, the three curves come much closer to one another.
-Thus scaling $\lamfac$ appropriately greatly mitigates the sensitivity of the transfer function to $\Lambda$, though it does not completely remove the dependence.
-
 % Figure
@@
 
+	Figures \ref{transfig}(a)-(b) show the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $0 \le \thup \le 25$.
+Figure \ref{transfig}(a) shows the cases $\Lambda = $ 12, 16, and 20 with $\lamfac = 8$ fixed in each.
+In this figure, the transfer function changes significantly with $\Lambda$.
+Figure~\ref{transfig}(b) shows the same but with the scaling $\lamfac = \Lambda/2$ that was argued in the previous section.
+In this figure, the three curves come much closer to one another.
+Thus scaling $\lamfac$ appropriately greatly mitigates the sensitivity of the transfer function to $\Lambda$, though it does not completely remove the dependence.
+
 	Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$ for the case $(\Lambda, \lamfac) = (16, 8)$.
@@
 Throughout, we will focus on the representative set of experiments detailed in Table \ref{paramtable}, in which the peak downstream skewness was measured to be 0.83.
-%We can then use Fig.~\ref{transfig}(c) as a guide to select the appropriate upstream inverse temperature $\thup$.
-With $(\Lambda, \lamfac) = (16, 8)$ fixed, the figure indicates a value of $\thup = 17$ achieves the desired outgoing skewness.
-We thus run the deterministic TKdV simulations with model parameters $\Lambda = 16$, $\lamfac = 8$, $\thup = 17$ and with $\epsup$, $\delup$, and $\dratdn$ set to the representative values in Table \ref{paramtable}.
-%
 With $(\Lambda, \lamfac) = (16, 8)$ fixed, Fig.~\ref{transfig}(c) indicates $\thup = 20$ as a reasonable value to attain the desired skewness.
@@
 We run the simulations to a sufficiently long dimensionless time of $t_{f} = 10$, with $dt = 5 \times 10^{-4}$ and with $10^3$ trajectories sampled from the upstream Gibbs measure $\Gup$.
-
-	In Fig.~\ref{uhist} we show the distributions of surface displacement that were measured in the experiments (top row) versus those that emerge from the TKdV simulations (bottom row).
-The experimental histograms represent exactly the same data depicted in Fig.~\ref{ExpDiagStats}, only converted to dimensionless displacement, $u$, for comparison against theory.
-Meanwhile, the TKdV distributions are extracted  from the long-time history of several trajectories sampled from the Gibbs ensemble, i.e.~mixed long-time/ensemble histograms.
-All histograms are shown on a semi-log scale to facilitate comparison of the tails, where extreme events lie.
 
@@
 
+	In Fig.~\ref{uhist} we show the distributions of surface displacement that were measured in the experiments (top row) versus those that emerge from the TKdV simulations (bottom row).
+The experimental histograms represent exactly the same data depicted in Fig.~\ref{ExpDiagStats}, only converted to dimensionless displacement, $u$, for comparison against theory.
+Meanwhile, the TKdV distributions are extracted  from the long-time history of several trajectories sampled from the Gibbs ensemble, i.e.~mixed long-time/ensemble histograms.
+All histograms are shown on a semi-log scale to facilitate comparison of the tails, where extreme events lie.
+
 	The comparison between experiments and theory in Fig.~\ref{uhist} is striking.
-As seen in the figure, both the experiments and theory show a transition from a nearly symmetric upstream distribution, to a highly skewed distribution downstream.
+As seen in the figure, both the experiments and theory show a transition from a nearly symmetric upstream distribution to a highly skewed distribution downstream.
 The TKdV theory not only captures this transition, but also recovers the shape of the resulting downstream distribution in remarkable detail.
@@
 \caption{
-Comparison of slope histograms.
-The TKdV simulations show similar slope distributions upstream, but visual distributions in the downstream distributions are visible.
-While the simulations recover the spread of the downstream slope statistics, they do not capture the asymmetry seen in the experiments.
-}
+Comparison of surface slope histograms.
+The experiments and TKdV simulations show very similar slope distributions upstream.
+Downstream, the theory accounts for the spread of the distribution and the long exponential tails, while differences in the detailed shape of the distributions are also visible.
+Note the elevated uncertainty in the experimental data due to numerical differentiation of free-surface measurements and the associated amplification of noise.}
 \label{slopehist}
@@
 Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50.
-Thus, the theoretically predicted jump in kurtosis, while not as extreme as that measured in experiments, is quite significant.
-These elevated levels of kurtosis are associated with the distinct appearance of the distributions, most notably the flat tails that appear in both the experimental and theoretical distributions, and also the pointy center that is most pronounced in the experiments.
-
-	Distinct differences in the shapes of these downstream distributions, however, are also visible.
-While the theoretical distributions remain symmetric, the experimental distributions skew towards negative slope.
-As mentioned earlier, this skewness is consistent with a right-moving wave of steep leading surface (i.e.~negative slope), which is a phenomenon apparently not captured by the TKdV simulations.
-Finally, it should be commented that the experimental measurements of slope involve numerical differentiation of the free surface extracted from optical images, which naturally introduces error.
-Furthermore, high-order statistics, such as kurtosis, can be very sensitive to noise and measurement error.
-Hence, the experimental statistics of slope, while highly valuable, should be regarded with some level of caution.
+Thus, the theoretically predicted jump in kurtosis, though not as extreme as that measured in experiments, is quite significant.
+These elevated levels of kurtosis are associated with the distinct appearance of the downstream distributions, most notably the long, flat tails that appear in both experiments and theory.
+
+	Differences in the detailed shapes of these downstream distributions are also visible.
+First, we point out that the experimental measurements involve numerical differentiation of  surface displacement extracted from optical images, a process that unavoidably amplifies any noise that is present.
+We must therefore proceed with caution in comparing the experiment and theory, recognizing the possibility that observed discrepancies may be due to these measurement errors.
+Nonetheless, we notice that the theoretically predicted distributions remain symmetric downstream and the experimental ones skew towards negative slope.
+As mentioned in Section \ref{experiments}, negative skewness is consistent with a right-moving wave of steep leading surface (i.e.~negative slope).
+Additionally, the peak of the experimental distribution appears sharper than that of the theory.
+
+%Figure \ref{ExpSpatialStats} may offer additional insight on these detailed differences in the shape of the downstream slope distributions.
+In particular, \ref{ExpSpatialStats}(e)--(f) show the spatial variation of the slope skewness and kurtosis.
+Note that variations in these statistics are significantly more transient than the corresponding variations in surface displacement; i.e.~the peaks in surface slope statistics arise more rapidly and decay more rapidly.
+Variations in displacement statistics also decay, but over a longer spatial scale, presumably due to dissipative effects.
+The TKdV histograms shown in Fig.~\ref{slopehist}, meanwhile, are long-time averages and hence not intended to capture such transient effects.
+These observations offer the suggestion that slope dynamics experience stronger transients than do surface displacements, leading to greater difficulty in the direct comparison between measurements extracted from a single location (experiments) and long-time averages (theory).
 
@@
 
-These observations can be rationalized with some simple analysis of the structure of TKdV.
+These observations can be rationalized with some simple observations regarding the structure of the TKdV system.
 Upstream of the ADC, since $\drat = 1$, the nonlinear effects are relatively weak and the dynamics are dominated by dispersion.
 Linearizing \eqref{TKdV} and introducing the ansatz $u_k = e^{i(kx - \omega_k t)}$ produces the dispersion relation for angular frequency $\omega_k = -C_2 k^3$, or for phase velocity $c_k = \omega_k/k = -C_2 k^2$ \cite{majdaqi2019}.
-In particular, the phase velocity is strictly negative, implying that waves can only propagate leftwards, as is consistent with Fig.~\ref{trajectories}(a).
+In particular, the phase velocity is strictly negative, implying that waves only propagate leftwards, as is consistent with Fig.~\ref{trajectories}(a).
 % Further details, see page 6 of Majda Qi, J Stat Phys.
@@
 
-	Downstream of the ADC, however, $\drat = 0.24$, which significantly amplifies nonlinearity while suppressing dispersion, thereby allowing waves to propagate in either direction.
-In particular, the dispersionless limit of KdV is the Burgers-Hopf equation, in which the nonlinear advection is proportional in both amplitude and direction to the value of $u$.
+	Downstream of the ADC, however, the depth ratio changes to $\drat = 0.24$.
+This change significantly amplifies nonlinearity while suppressing dispersion, and thereby allows waves to propagate in either direction.
+In particular, the dispersionless limit of KdV is the Burgers-Hopf equation, for which the nonlinear advection is proportional in both amplitude and direction to the value of $u$.
 Hence, positive displacements would be expected to move rightwards, as is seen in Fig.~\ref{trajectories}(b).

### 2019-11-07 paper2_JNLS/paper.tex
NM

@@
 That work uses a deep neural network coupled with a judicious choice of the entropy loss function that emphasizes the dominant structures of the turbulent field \cite{MachineLearning2019}.
-It should be noted that, in these studies, the extreme events can occur intermittently and on a relatively frequent basis, in contrast to the simpler situation of isolated, rare events \cite{guth2019machine}.
+It should be noted that, in these studies, extreme events are represented by strong skewness of the wave-field.
+These events occur intermittently and on a relatively frequent basis, which is in contrast to the simpler situation of isolated, rare events \cite{guth2019machine}.
 
@@
 Downstream, the theory accounts for the spread of the distribution and the long exponential tails, while differences in the detailed shape of the distributions are also visible.
-Note the elevated uncertainty in the experimental data due to numerical differentiation of free-surface measurements and the associated amplification of noise.}
+Note the elevated uncertainty in the experimental data due to numerical differentiation of free-surface measurements.}
 \label{slopehist}

### 2019-11-07 paper2_JNLS/paper.tex
NNM

@@
 \end{align}
-Here, we have chosen the sign convention of Lax \cite{lax1975periodic}.
+As seen in \eqref{H3H2}, the cubic component, $\Hthree$, represents the skewness of the wave-field, while the quadratic component, $\Htwo$, represents the energy of the surface slope.
+The sign difference between the two \eqref{Hamiltonian} thus represents a competition between wave skewness and slope energy.
+This is the reason for our interest in experimental measurements of the surface {\it slope} statistics in the current manuscript.
+
+We remark that, in defining the Hamiltonian, we have chosen the sign convention of Lax \cite{lax1975periodic}.
 More recent work of Bajars et al.~\cite{bajars2013weakly} and Majda et al.~\cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite.
@@
 We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.
-
-% Notes for myself on the sign change: Abramov CPAM 2003, Bajars Nonlinearity 2013, and our 2019 PNAS paper all use J = -d/dx with the opposite sign in the Hamiltonian.
-However, I was able to easily find several references online that use J = d/dx, so there is nothing wrong with it.
-The exact criteria that J has to satisfy are listed in the CPAM 2003 paper, and J = d/dx satisfies all these criteria just as well as J = -d/dx does.
 

### 2019-11-07 paper2_JNLS/paper.tex
NM

@@
 We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments.
-The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in spectacular fashion by direct experimental measurements.
+The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in a spectacular fashion by direct experimental measurements.
 \end{abstract}
@@
 In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from Majda et al.~2019 \cite{majda2019}.
-Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices.
-To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment.
+Since the choice of normalization is not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices.
+To this end, we consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment.
 We introduce the dimensionless variables
@@
 
-% NOTES
-% 1) The PNAS paper uses C_theta^{1}. The Majda, Qi Stat Phys paper uses C_theta sometimes and C_theta^{-1} in other places (page 11). So I'll just pick the one I like better.
-% 2) Could comment that other possible ensembles exist \cite{kleeman2014nonequilibrium}.
+	We note that, in the current formulation, a positive inverse temperature $\theta > 0$ produces physically realistic statistics with a decaying energy spectrum, as is consistent with experiments \cite{majda2019, bajars2013weakly}.
+Negative inverse temperature produces a physically unrealistic spectrum that has more energy at smaller scales.
+Hence, we will hereafter focus on the physically realistic case of $\theta > 0$.
 

### 2019-11-12 paper2_JNLS/paper.tex
add symplectic scheme

@@
 \nick{Include a short description of the numerical method for simulation TKdV: a symplectic integrator.}
+Two approaches can be applied to predict the statistics in the
+downstream wave disturbance, which can be used to confirm each other in
+the numerical results. The equilibrium statistical mechanics
+is an easy way to confirm the matching condition and get the equilibrium
+distribution in far-field long-time limit. On the other hand, the
+direct simulations of the dynamical model can offer estimates about
+the transient statistics before the far-field equilibrium state is
+reached.
+
+Here we describe the direct numerical method used for the simulation of the TKdV equation.
+
+\subsubsection*{Conservation of energy and Hamiltonian form the symplectic scheme}
+
+The conservation property plays a central role in the statistical
+matching and the final equilibrium measure.  For running the direct numerical simulations of the TKdV equations,
+a proper symplectic integrator is necessary to guarantee the conservation
+of the Hamiltonian and energy in time, otherwise the non-symplectic
+schemes will introduce large numerical dissipations. 
+Symplecticness of Hamiltonian PDEs refers to the conservation of oriented areas in the phase space.
+In the numerical schemes for long time integration, we wish that the discretized system still maintains the Hamiltonian (symplectic) structure.
+Then the conservation properties of the major quantities can be better guaranteed in the long evolution time to compute model statistics.
+The symplectic numerical integrators discretize the Hamiltonian and the Poisson structures separated and then form the resulting scheme for conserving the Hamiltonian.
+
+In the numerical discretization, first we split the Hamiltonian into separated linear and nonlinear dynamics.
+A pseudo-spectral scheme is used for spatial discretization.
+Proper de-aliasing methods should be used to calculate the nonlinear term $u^{2}$ in the equation.
+Here the standard 2/3-rule is applied to calculate the quadratic term in physical domain and extend the spectral modes by adding additional zeros in the tails.
+That is, we explicitly calculate the spectral mode of the square term by
+\[
+\widehat{\left(u^{2}\right)}_{k}=\mathcal{F}\left[U^{2}\right],\quad U=\mathcal{F}^{-1}\left[\left\{ \hat{u}_{k}\mid_{\left|k\right|\leq\frac{\Lambda}{2}},0\mid_{\frac{\Lambda}{2}<\left|k\right|\leq\frac{3}{2}\Lambda}\right\} \right],
+\]
+where $\mathcal{F}$ represents the operator for Fourier transform.
+Therefore the ODE system after spatial discretization can be written for the spectral vector $\mathbf{u}=\left\{ \hat{u}_{k}\right\} _{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ F_{k}\right\}$
+\[
+\dot{\hat{u}}_{k}=F_{k}\left(u\right)=C_{2}\mathcal{D}^{1/2}ik^{3}\hat{u}_{k}-\frac{1}{2}C_{3}\mathcal{D}^{-3/2}ik\widehat{\left(u^{2}\right)}_{k}.
+\]
+Different scales are mixed due to the nonlinear term $u^{2}$. The linear term including third derivative might become stiff when the total wavenumbers become large.
+
+The symplectic scheme
+used here for the time integration of the TKdV equation is the 4th-order midpoint method \cite{mclachlan1993symplectic}.
+Let $\mathbf{u}^{n}$ be the solution at time $t_{n}$ and $\mathbf{u}^{n+1}$
+be the solution at next time step $t_{n+1}=t_{n}+\Delta t$, and $\mathbf{F}$ the
+discretized operator for the spectral modes. The midpoint method has
+two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$ so that
+\[
+\begin{aligned}\mathbf{y}_{1}-\mathbf{u}^{n} & =w_{1}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{1}+\mathbf{u}^{n}\right]\right),\\
+\mathbf{y}_{2}-\mathbf{y}_{1} & =w_{2}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{2}+\mathbf{y}_{1}\right]\right),\\
+\mathbf{u}^{n+1}-\mathbf{y}_{2} & =w_{3}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{u}^{n+1}+\mathbf{y}_{2}\right]\right),
+\end{aligned}
+\]
+with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$,
+$w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$. Then the PDFs can be counted
+from the solution $u_{\Lambda}\left(t\right)$, and the statistical
+expectations can be calculated by averaging along the trajectory
+\[
+\left\langle F\left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T}F\left(u_{\Lambda}\left(t\right)\right)dt,
+\]
+given ergodicity in the solution with $u_{\Lambda}\left(t_{0}\right)$
+already reaching the stationary state.
+
+For the symplectic schemes, implicit iteration method is required to get the accurate solution. For any implicit problem of the form
+\[
+\mathbf{z}-\mathbf{u}^{n}=\tau\mathbf{F\left(z\right)}=\tau\mathbf{Lz}+\tau\mathbf{N\left(z\right)},
+\]
+we split the operators into a linear and nonlinear components and only treat the nonlinear part on the right hand side for the previous stage in the iteration.
+Then we have the iterative scheme for the solution of implicit problem
+\[
+\left(1-\tau\mathbf{L}\right)\mathbf{z}^{\left[l+1\right]}=\mathbf{u}^{n}+\tau\mathbf{N}\left(\mathbf{z}^{\left[l\right]}\right).
+\]
+The linear operator $\mathbf{L}$ is the diagonal differential operator, thus it is easy to invert in these schemes. The termination criteria is determined by
+\[
+\left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert <\delta\;\mathrm{and}\;\left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert \geq\left\Vert \mathbf{z}^{\left[l\right]}-\mathbf{z}^{\left[l-1\right]}\right\Vert .
+\]
+It is found that the iteration tolerance can be set as $\delta=1\times10^{-10}$.
+Finally, the starting guess for the iterations can be found by quadratic extrapolation for the previous three stages $\left(t_{u},\mathbf{u}^{n}\right), \left(t_{u}-\tau,\mathbf{u}^{n-1}\right), \left(t_{u}-2\tau,\mathbf{u}^{n-2}\right)$.
+Therefore we can find the extrapolated value at $t=t_{n}+\beta\tau$ by
+\[ \mathbf{u}^{n+\beta}=\mathbf{Q}\left(t_{n}+\beta\tau\right)=\left[\frac{1}{2}\beta\left(\beta+3\right)+1\right]\mathbf{u}^{n}-\beta\left(\beta+2\right)\mathbf{u}^{n-1}+\frac{1}{2}\beta\left(\beta+1\right)\mathbf{u}^{n-2}.
+\]
 

### 2019-11-12 paper2_JNLS/paper.tex
NM

@@
 
-	Since the inverse temperature, $\theta$, is a key modeling parameter, it would be highly desirable to normalize the system such that the value of $\theta$ does not depend sensitively on the truncation index, as $\Lambda$ grows large.
+	Since the inverse temperature, $\theta$, is a key modeling parameter, it would be highly desirable to normalize the system such that the value of $\theta$ does not depend sensitively on the truncation index as $\Lambda$ grows large.
 That way, $\theta$ can be interpreted as a real physical parameter that can be linked to the experiments and whose value does not depend sensitively on where one chooses to truncate the system.
@@
 
-\nick{Include a short description of the numerical method for simulation TKdV: a symplectic integrator.}
-Two approaches can be applied to predict the statistics in the
-downstream wave disturbance, which can be used to confirm each other in
-the numerical results. The equilibrium statistical mechanics
-is an easy way to confirm the matching condition and get the equilibrium
-distribution in far-field long-time limit. On the other hand, the
-direct simulations of the dynamical model can offer estimates about
-the transient statistics before the far-field equilibrium state is
-reached.
-
-Here we describe the direct numerical method used for the simulation of the TKdV equation.
-
-\subsubsection*{Conservation of energy and Hamiltonian form the symplectic scheme}
-
-The conservation property plays a central role in the statistical
-matching and the final equilibrium measure.  For running the direct numerical simulations of the TKdV equations,
-a proper symplectic integrator is necessary to guarantee the conservation
-of the Hamiltonian and energy in time, otherwise the non-symplectic
-schemes will introduce large numerical dissipations. 
+%Two approaches can be applied to predict the statistics in the downstream wave disturbance, which can be used to confirm each other in the numerical results.
+The equilibrium statistical mechanics is an easy way to confirm the matching condition and get the equilibrium distribution in far-field long-time limit.
+On the other hand, the direct simulations of the dynamical model can offer estimates about the transient statistics before the far-field equilibrium state is reached.
+
+We now detail the method for direct numerical simulation of the TKdV dynamical system \eqref{TKdV}.
+In particular, the conservation of energy and Hamiltonian plays a central role in the emergent statistical features of the system.
+Hence, it is important to conserve these quantities numerically and so we use a so-called symplectic scheme.
+
+%In particular, to properly conserve the energy and Hamiltonian over long-time horizon simulations, it is necessary to use a symplectic integrator
+
+\vsp{5}
+
+%\subsubsection*{Conservation of energy and Hamiltonian form the symplectic scheme}
+
+The conservation property plays a central role in the statistical matching and the final equilibrium measure.
+For running the direct numerical simulations of the TKdV equations, a proper symplectic integrator is necessary to guarantee the conservation of the Hamiltonian and energy in time, otherwise the non-symplectic schemes will introduce large numerical dissipations.
 Symplecticness of Hamiltonian PDEs refers to the conservation of oriented areas in the phase space.

### 2019-11-13 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\Htwo}{\Ham_{2}}
-%\newcommand{\Fcnl}{\mathcal{F}}
+\newcommand{\Fcnl}{\mathcal{F}}
 % Truncated stuff
@@
 \newcommand{\traveldist}{\ell}
-
+\newcommand{\Ftrans}{\mathcal{F}}
 % Experiments
@@
 \begin{comment}
-The evolution of any functional $\mathcal{F}[u]$ is given by 
-\begin{align}
-\partial_t \mathcal{F} = \{ \mathcal{F}, \Ham \} = \int_{-\pi}^{\pi} \vard{\mathcal{F}}{u} \sympJ \vard{\Ham}{u} \dx
+The evolution of any functional $\Fcnl[u]$ is given by 
+\begin{align}
+\partial_t \Fcnl  = \{ \Fcnl, \Ham \} = \int_{-\pi}^{\pi} \vard{\Fcnl}{u} \sympJ \vard{\Ham}{u} \dx
 \end{align}
@@
 We now detail the method for direct numerical simulation of the TKdV dynamical system \eqref{TKdV}.
-In particular, the conservation of energy and Hamiltonian plays a central role in the emergent statistical features of the system.
-Hence, it is important to conserve these quantities numerically and so we use a so-called symplectic scheme.
-
-%In particular, to properly conserve the energy and Hamiltonian over long-time horizon simulations, it is necessary to use a symplectic integrator
+In particular, since the conservation of energy and Hamiltonian plays a central role in the emergent statistical features of the system, it is important for the numerical scheme to conserve these quantities over long time horizons.
+We therefore employ a symplectic integrator, which, by preserving oriented areas in phase space, conserves the energy and Hamiltonian exactly.
+
+	The numerical discretization first separates the Hamiltonian into linear and nonlinear components.
+Then a pseudo-spectral scheme is used for spatial discretization, with de-aliasing applied to the nonlinear term $u^2$
 
@@
 
-%\subsubsection*{Conservation of energy and Hamiltonian form the symplectic scheme}
-
-The conservation property plays a central role in the statistical matching and the final equilibrium measure.
-For running the direct numerical simulations of the TKdV equations, a proper symplectic integrator is necessary to guarantee the conservation of the Hamiltonian and energy in time, otherwise the non-symplectic schemes will introduce large numerical dissipations.
-Symplecticness of Hamiltonian PDEs refers to the conservation of oriented areas in the phase space.
-In the numerical schemes for long time integration, we wish that the discretized system still maintains the Hamiltonian (symplectic) structure.
-Then the conservation properties of the major quantities can be better guaranteed in the long evolution time to compute model statistics.
-The symplectic numerical integrators discretize the Hamiltonian and the Poisson structures separated and then form the resulting scheme for conserving the Hamiltonian.
-
-In the numerical discretization, first we split the Hamiltonian into separated linear and nonlinear dynamics.
+\begin{align}
+\label{TKdV2}
+&\pd{\uL}{t} = F = 
+-  \frac{1}{2} C_3 \drat^{-3/2} \,\Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = 0 \qquad \text{for } x \in [-\pi,\pi]
+\end{align}
+
+%The symplectic numerical integrators discretize the Hamiltonian and the Poisson structures separated and then form the resulting scheme for conserving the Hamiltonian.
+%In the numerical discretization, first we split the Hamiltonian into separated linear and nonlinear dynamics.
+
 A pseudo-spectral scheme is used for spatial discretization.
@@
 \[
-\widehat{\left(u^{2}\right)}_{k}=\mathcal{F}\left[U^{2}\right],\quad U=\mathcal{F}^{-1}\left[\left\{ \hat{u}_{k}\mid_{\left|k\right|\leq\frac{\Lambda}{2}},0\mid_{\frac{\Lambda}{2}<\left|k\right|\leq\frac{3}{2}\Lambda}\right\} \right],
+\widehat{\left(u^{2}\right)}_{k} = \Ftrans \left[U^{2}\right],\quad U 
+= \Ftrans^{-1} \left[\left\{ \hat{u}_{k}\mid_{\left|k\right|\leq\frac{\Lambda}{2}},0\mid_{\frac{\Lambda}{2}<\left|k\right|\leq\frac{3}{2}\Lambda}\right\} \right],
 \]
-where $\mathcal{F}$ represents the operator for Fourier transform.
+where $\Ftrans$ represents the operator for Fourier transform.
 Therefore the ODE system after spatial discretization can be written for the spectral vector $\mathbf{u}=\left\{ \hat{u}_{k}\right\} _{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ F_{k}\right\}$

### 2019-11-13 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\CC}{{\mathbb{C}}}
+
 % Basic physical parameters and scales
@@
 \newcommand{\dratupdn}{\drat_{\pm}}
+% Other experimental stuff
+\newcommand{\omavg}{\omega_0}
+\newcommand{\omsig}{\sigma_{\omega}}
+\newcommand{\Dphi}{\Delta \phi}
 % Statistical quantities
@@
 \newcommand{\std}{\text{std}}
+
 % Dimensionless parameters
@@
 \newcommand{\meanz}[1]{\mean{#1}_0}
-% Stuff
+% Misc stuff
 \newcommand{\tavg}[1]{\overline{#1}}
@@
 \newcommand{\Ftrans}{\mathcal{F}}
-% Experiments
-\newcommand{\omavg}{\omega_0}
-\newcommand{\omsig}{\sigma_{\omega}}
-\newcommand{\Dphi}{\Delta \phi}
+\newcommand{\RHS}{F}
+\newcommand{\RHSh}{\hat{\RHS}}
+\newcommand{\usqdah}{ (\widehat{u_{da}^2}) }
 
@@
 \label{TKdV}
-&\pd{\uL}{t} +  \frac{1}{2} C_3 \drat^{-3/2} \,\Proj (\uL)^2 + C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = 0
+&\pd{\uL}{t} +  \frac{1}{2} C_3 \drat^{-3/2} \, \pd{}{x} \Proj (\uL)^2 + C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = 0
 \qquad \text{for } x \in [-\pi,\pi] \\
-&C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad
-C_2 = \frac{2 \pi^2 \delup}{3 \lamfac^2}
+&C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad C_2 = \frac{2 \pi^2 \delup}{3 \lamfac^2}
 \end{align}
@@
 
-% NOTES
-%Further, since the regime is shallow, short-time dispersive effects are minimal, meaning that all waves travel at nearly the same speed $c = \sqrt{g \depth}$, independent of wavelength.
-
 %----------------------------------------------------------------------------%
@@
 
-	The numerical discretization first separates the Hamiltonian into linear and nonlinear components.
-Then a pseudo-spectral scheme is used for spatial discretization, with de-aliasing applied to the nonlinear term $u^2$
+Rearranging the TKdV system \eqref{TKdV} gives
+\begin{align}
+\label{TKdV2}
+&\pd{\uL}{t} = -  \frac{1}{2} C_3 \drat^{-3/2} \, \pd{}{x} \Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = \RHS
+%\qquad \text{for } x \in [-\pi,\pi]
+\end{align}	
+where we have represented the right-hand side by $\RHS$.
+Our numerical scheme first separates the Hamiltonian into linear and nonlinear components.
+Then a pseudo-spectral scheme is used for spatial discretization with de-aliasing applied to the nonlinear term $\Proj (\uL)^2$ according to the standard 2/3-rule.
+That is, we first pad $\uhat_k$ with zeros for $\Lambda < \abs{k} \le 3 \Lambda$ \nick{3 or 3/2???}, transform to physical space and square to obtain $\uL^2$ on a fine grid, then transform back to spectral space and truncate to obtain the Fourier coefficients of $\Proj (\uL)^2$
+For simplicity, we denote these Fourier coefficients $\usqdah_k$. That is,
+\begin{equation}
+\Proj (\uL)^2 = \sum_{\abs{k} \le \Lambda} \usqdah_k e^{i k x}
+\end{equation}
 
 \vsp{5}
-
-\begin{align}
-\label{TKdV2}
-&\pd{\uL}{t} = F = 
--  \frac{1}{2} C_3 \drat^{-3/2} \,\Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = 0 \qquad \text{for } x \in [-\pi,\pi]
-\end{align}
 
@@
 
-A pseudo-spectral scheme is used for spatial discretization.
-Proper de-aliasing methods should be used to calculate the nonlinear term $u^{2}$ in the equation.
-Here the standard 2/3-rule is applied to calculate the quadratic term in physical domain and extend the spectral modes by adding additional zeros in the tails.
-That is, we explicitly calculate the spectral mode of the square term by
-\[
-\widehat{\left(u^{2}\right)}_{k} = \Ftrans \left[U^{2}\right],\quad U 
-= \Ftrans^{-1} \left[\left\{ \hat{u}_{k}\mid_{\left|k\right|\leq\frac{\Lambda}{2}},0\mid_{\frac{\Lambda}{2}<\left|k\right|\leq\frac{3}{2}\Lambda}\right\} \right],
-\]
-where $\Ftrans$ represents the operator for Fourier transform.
-Therefore the ODE system after spatial discretization can be written for the spectral vector $\mathbf{u}=\left\{ \hat{u}_{k}\right\} _{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ F_{k}\right\}$
-\[
-\dot{\hat{u}}_{k}=F_{k}\left(u\right)=C_{2}\mathcal{D}^{1/2}ik^{3}\hat{u}_{k}-\frac{1}{2}C_{3}\mathcal{D}^{-3/2}ik\widehat{\left(u^{2}\right)}_{k}.
-\]
+Therefore the ODE system after spatial discretization can be written for the spectral vector $\mathbf{u}=\left\{ \hat{u}_{k}\right\}_{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$
+\begin{equation}
+\dot{\hat{u}}_{k} = \RHSh_k \left[u\right] = C_{2}\mathcal{D}^{1/2}ik^{3}\hat{u}_{k}-\frac{1}{2}C_{3}\mathcal{D}^{-3/2}ik\widehat{\left(u^{2}\right)}_{k}.
+\end{equation}
 Different scales are mixed due to the nonlinear term $u^{2}$. The linear term including third derivative might become stiff when the total wavenumbers become large.
@@
 two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$ so that
-\[
-\begin{aligned}\mathbf{y}_{1}-\mathbf{u}^{n} & =w_{1}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{1}+\mathbf{u}^{n}\right]\right),\\
-\mathbf{y}_{2}-\mathbf{y}_{1} & =w_{2}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{2}+\mathbf{y}_{1}\right]\right),\\
-\mathbf{u}^{n+1}-\mathbf{y}_{2} & =w_{3}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{u}^{n+1}+\mathbf{y}_{2}\right]\right),
-\end{aligned}
-\]
+\begin{align}
+&\mathbf{y}_{1}-\mathbf{u}^{n} = w_{1}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{1}+\mathbf{u}^{n}\right]\right),\\
+&\mathbf{y}_{2}-\mathbf{y}_{1} = w_{2}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{2}+\mathbf{y}_{1}\right]\right),\\
+&\mathbf{u}^{n+1}-\mathbf{y}_{2} = w_{3}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{u}^{n+1}+\mathbf{y}_{2}\right]\right),
+\end{align}
 with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$,
@@
 expectations can be calculated by averaging along the trajectory
-\[
+\begin{equation}
 \left\langle F\left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T}F\left(u_{\Lambda}\left(t\right)\right)dt,
-\]
+\end{equation}
 given ergodicity in the solution with $u_{\Lambda}\left(t_{0}\right)$
@@
 For the symplectic schemes, implicit iteration method is required to get the accurate solution. For any implicit problem of the form
-\[
+\begin{equation}
 \mathbf{z}-\mathbf{u}^{n}=\tau\mathbf{F\left(z\right)}=\tau\mathbf{Lz}+\tau\mathbf{N\left(z\right)},
-\]
+\end{equation}
 we split the operators into a linear and nonlinear components and only treat the nonlinear part on the right hand side for the previous stage in the iteration.
 Then we have the iterative scheme for the solution of implicit problem
-\[
+\begin{equation}
 \left(1-\tau\mathbf{L}\right)\mathbf{z}^{\left[l+1\right]}=\mathbf{u}^{n}+\tau\mathbf{N}\left(\mathbf{z}^{\left[l\right]}\right).
-\]
+\end{equation}
 The linear operator $\mathbf{L}$ is the diagonal differential operator, thus it is easy to invert in these schemes. The termination criteria is determined by
-\[
+\begin{equation}
 \left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert <\delta\;\mathrm{and}\;\left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert \geq\left\Vert \mathbf{z}^{\left[l\right]}-\mathbf{z}^{\left[l-1\right]}\right\Vert .
-\]
+\end{equation}
 It is found that the iteration tolerance can be set as $\delta=1\times10^{-10}$.
@@
 Therefore we can find the extrapolated value at $t=t_{n}+\beta\tau$ by
-\[ \mathbf{u}^{n+\beta}=\mathbf{Q}\left(t_{n}+\beta\tau\right)=\left[\frac{1}{2}\beta\left(\beta+3\right)+1\right]\mathbf{u}^{n}-\beta\left(\beta+2\right)\mathbf{u}^{n-1}+\frac{1}{2}\beta\left(\beta+1\right)\mathbf{u}^{n-2}.
-\]
+\begin{equation}
+\mathbf{u}^{n+\beta}=\mathbf{Q}\left(t_{n}+\beta\tau\right)=\left[\frac{1}{2}\beta\left(\beta+3\right)+1\right]\mathbf{u}^{n}-\beta\left(\beta+2\right)\mathbf{u}^{n-1}+\frac{1}{2}\beta\left(\beta+1\right)\mathbf{u}^{n-2}.
+\end{equation}
 

### 2019-11-13 paper2_JNLS/paper.tex
NM

@@
 Then a pseudo-spectral scheme is used for spatial discretization with de-aliasing applied to the nonlinear term $\Proj (\uL)^2$ according to the standard 2/3-rule.
-That is, we first pad $\uhat_k$ with zeros for $\Lambda < \abs{k} \le 3 \Lambda$ \nick{3 or 3/2???}, transform to physical space and square to obtain $\uL^2$ on a fine grid, then transform back to spectral space and truncate to obtain the Fourier coefficients of $\Proj (\uL)^2$
+That is, we first pad $\uhat_k$ with zeros for $\Lambda < \abs{k} \le 3\Lambda/2$, transform to physical space and square to obtain $\uL^2$ on a fine grid, then transform back to spectral space and truncate to obtain the Fourier coefficients of $\Proj (\uL)^2$
 For simplicity, we denote these Fourier coefficients $\usqdah_k$. That is,
@@
 \end{equation}
-
-\vsp{5}
-
-%The symplectic numerical integrators discretize the Hamiltonian and the Poisson structures separated and then form the resulting scheme for conserving the Hamiltonian.
-%In the numerical discretization, first we split the Hamiltonian into separated linear and nonlinear dynamics.
-
-Therefore the ODE system after spatial discretization can be written for the spectral vector $\mathbf{u}=\left\{ \hat{u}_{k}\right\}_{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$
-\begin{equation}
-\dot{\hat{u}}_{k} = \RHSh_k \left[u\right] = C_{2}\mathcal{D}^{1/2}ik^{3}\hat{u}_{k}-\frac{1}{2}C_{3}\mathcal{D}^{-3/2}ik\widehat{\left(u^{2}\right)}_{k}.
-\end{equation}
-Different scales are mixed due to the nonlinear term $u^{2}$. The linear term including third derivative might become stiff when the total wavenumbers become large.
-
-The symplectic scheme
-used here for the time integration of the TKdV equation is the 4th-order midpoint method \cite{mclachlan1993symplectic}.
-Let $\mathbf{u}^{n}$ be the solution at time $t_{n}$ and $\mathbf{u}^{n+1}$
+With this spectral discretization, the ODE system \eqref{TKdV2} can be written for the spectral vector $\mathbf{u}=\left\{ \hat{u}_{k}\right\}_{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$ as
+\begin{equation}
+\label{TKdVDisc}
+\td{}{t} \uhat_k =  -\frac{1}{2} C_3 \drat^{-3/2} \, ik \usqdah_k + C_2 \drat^{1/2} \, ik^{3} \uhat_k = \RHSh_k
+\end{equation}
+Due to the nonlinearity, this system mixes scales during evolution. We note the third-order linear term may become stiff for large $\Lambda$.
+
+For time integration of \eqref{TKdVDisc}, we employ a 4th-order midpoint symplectic scheme \cite{mclachlan1993symplectic}. Let $\mathbf{u}^{n}$ be the solution at time $t_{n}$ and $\mathbf{u}^{n+1}$
 be the solution at next time step $t_{n+1}=t_{n}+\Delta t$, and $\mathbf{F}$ the
@@
 with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$,
-$w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$. Then the PDFs can be counted
-from the solution $u_{\Lambda}\left(t\right)$, and the statistical
-expectations can be calculated by averaging along the trajectory
+$w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$.
+
+\vsp{5}
+
+Then the PDFs can be counted from the solution $u_{\Lambda}\left(t\right)$, and the statistical expectations can be calculated by averaging along the trajectory
 \begin{equation}
@@
 \end{equation}
-given ergodicity in the solution with $u_{\Lambda}\left(t_{0}\right)$
-already reaching the stationary state.
+given ergodicity in the solution with $u_{\Lambda}\left(t_{0}\right)$ already reaching the stationary state.
 
@@
 \end{equation}
+
+%The symplectic numerical integrators discretize the Hamiltonian and the Poisson structures separated and then form the resulting scheme for conserving the Hamiltonian.
 

### 2019-11-13 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\RHSh}{\hat{\RHS}}
-\newcommand{\usqdah}{ (\widehat{u_{da}^2}) }
+\newcommand{\usqdah}{\hat{v}} %{ (\widehat{u_{da}^2}) }
 
@@
 \label{TKdV2}
-&\pd{\uL}{t} = -  \frac{1}{2} C_3 \drat^{-3/2} \, \pd{}{x} \Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = \RHS
+&\pd{\uL}{t} = -  \frac{1}{2} C_3 \drat^{-3/2} \, \pd{}{x} \Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = \RHS[\uL]
 %\qquad \text{for } x \in [-\pi,\pi]
 \end{align}	
-where we have represented the right-hand side by $\RHS$.
+where we have represented the right-hand side by the operator $\RHS[\uL]$.
 Our numerical scheme first separates the Hamiltonian into linear and nonlinear components.
@@
 \end{equation}
-With this spectral discretization, the ODE system \eqref{TKdV2} can be written for the spectral vector $\mathbf{u}=\left\{ \hat{u}_{k}\right\}_{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$ as
+With this spectral discretization, the ODE system \eqref{TKdV2} can be written as
 \begin{equation}
@@
 
-For time integration of \eqref{TKdVDisc}, we employ a 4th-order midpoint symplectic scheme \cite{mclachlan1993symplectic}. Let $\mathbf{u}^{n}$ be the solution at time $t_{n}$ and $\mathbf{u}^{n+1}$
-be the solution at next time step $t_{n+1}=t_{n}+\Delta t$, and $\mathbf{F}$ the
-discretized operator for the spectral modes. The midpoint method has
-two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$ so that
-\begin{align}
-&\mathbf{y}_{1}-\mathbf{u}^{n} = w_{1}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{1}+\mathbf{u}^{n}\right]\right),\\
-&\mathbf{y}_{2}-\mathbf{y}_{1} = w_{2}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{2}+\mathbf{y}_{1}\right]\right),\\
-&\mathbf{u}^{n+1}-\mathbf{y}_{2} = w_{3}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{u}^{n+1}+\mathbf{y}_{2}\right]\right),
+For time integration of \eqref{TKdVDisc}, we employ a 4th-order midpoint symplectic scheme \cite{mclachlan1993symplectic}.
+We introduce the spectral vectors $\mathbf{u}=\left\{ \hat{u}_{k}\right\}_{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$, and let $\mathbf{u}^{n}$ denote the solution at time $t_{n}$.
+The midpoint method has two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$:
+\begin{align}
+\label{SympSteps}
+&\mathbf{y}_{1}-\mathbf{u}^{n} = w_{1}\Delta t \, \mathbf{F}
+\left[ \frac{1}{2} \left(\mathbf{y}_{1}+\mathbf{u}^{n}\right) \right] ,\\
+&\mathbf{y}_{2}-\mathbf{y}_{1} = w_{2}\Delta t \, \mathbf{F}
+\left[ \frac{1}{2} \left(\mathbf{y}_{2}+\mathbf{y}_{1}\right) \right] ,\\
+&\mathbf{u}^{n+1}-\mathbf{y}_{2} = w_{3}\Delta t \, \mathbf{F}
+\left[ \frac{1}{2} \left(\mathbf{u}^{n+1}+\mathbf{y}_{2}\right) \right] ,
 \end{align}
@@
 
-Then the PDFs can be counted from the solution $u_{\Lambda}\left(t\right)$, and the statistical expectations can be calculated by averaging along the trajectory
-\begin{equation}
-\left\langle F\left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T}F\left(u_{\Lambda}\left(t\right)\right)dt,
-\end{equation}
-given ergodicity in the solution with $u_{\Lambda}\left(t_{0}\right)$ already reaching the stationary state.
+\todo[inline]{Di, I do not understand this paragraph.
+Is the $F$ meant to be the same $F$ as above, i.e.~the right-hand-side of the TKdV equation \eqref{TKdV2}?
+Or is $F$ some arbitrary function of the state variable that we want to compute the expectation of?}
+Then the PDFs can be counted from the solution $\uL(t)$, and the statistical expectations can be calculated by averaging along the trajectory
+\begin{equation}
+\left\langle F \left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T} F \left(u_{\Lambda}\left(t\right)\right)dt,
+\end{equation}
+given ergodicity in the solution with $\uL(t_0)$ already reaching the stationary state.
+
+\todo[inline]{I am also a little confused about the paragraphs below.
+Eq.~\eqref{iterate} is not the same form of Eq.~(43)--(45) because in those equations it is not simply $\mathbf{z}$ inside $\mathbf{F}$, but instead an average of $\mathbf{z}$ and the previous value.
+Am I simply misinterpreting?}
+\todo[inline]{Also, this is simply Picard iteration, correct?
+Perhaps we could simply say in words that we perform Picard iteration so that we don't have to worry about getting the equations correct in detail.
+What's your opinion?}
 
@@
 \begin{equation}
-\mathbf{z}-\mathbf{u}^{n}=\tau\mathbf{F\left(z\right)}=\tau\mathbf{Lz}+\tau\mathbf{N\left(z\right)},
+\label{iterate}
+\mathbf{z}-\mathbf{u}^{n}=\tau \mathbf{F \left(z\right)} = \tau\mathbf{Lz}+\tau\mathbf{N\left(z\right)},
 \end{equation}
@@
 \begin{equation}
-\left(1-\tau\mathbf{L}\right)\mathbf{z}^{\left[l+1\right]}=\mathbf{u}^{n}+\tau\mathbf{N}\left(\mathbf{z}^{\left[l\right]}\right).
+\left(1-\tau\mathbf{L}\right)\mathbf{z}^{\left[l+1\right]} = \mathbf{u}^{n}+\tau\mathbf{N}\left(\mathbf{z}^{\left[l\right]}\right).
 \end{equation}

### 2019-11-13 paper2_JNLS/paper.tex
NM

@@
 
+\nick{Edit this}
+To compute the transfer function, we use a weighted random sampling strategy, with samples drawn from the uniform distribution ($d\mu$ or $\Gz$, which notation?) and the weight provided by the Gibbs form \eqref{Gibbs}.
+\begin{equation}
+\mean{Q} = \frac{\int Q \exp\left(-\theta \Ham \right) d\mu}{ \int \exp\left(-\theta \Ham \right) d\mu}
+\approx \frac{\sumsamp Q_i \exp(-\theta \Ham_i)} {\sumsamp \exp(-\theta \Ham_i)}
+\end{equation}
+For computing the transfer function, this simple, weighted averaging approach is preferable to more sophisticated MCMC because it allows you to collect a large number of samples once, and then simply change the weights for different $\theta$.
+MCMC would require to completely restart the sampling for each individual $\theta$.
+
 %----------------------------------------------------------------------------%
@@
 
+\nick{We should simply comment that we use MHMC to initialize the direct simulations.}
+
 \vsp{5}
@@
 %The symplectic numerical integrators discretize the Hamiltonian and the Poisson structures separated and then form the resulting scheme for conserving the Hamiltonian.
-
-\subsection{Sampling from the Gibbs measures}
-
-\nick{Include a short description of the sampling method}
-
-\begin{equation}
-\mean{Q} = \frac{\int Q \exp\left(-\theta \Ham \right) d\mu}{ \int \exp\left(-\theta \Ham \right) d\mu}
-\approx \frac{\sumsamp Q_i \exp(-\theta \Ham_i)} {\sumsamp \exp(-\theta \Ham_i)}
-\end{equation}
 

### 2019-11-13 paper2_JNLS/paper.tex
minor changes

@@
 Is the $F$ meant to be the same $F$ as above, i.e.~the right-hand-side of the TKdV equation \eqref{TKdV2}?
-Or is $F$ some arbitrary function of the state variable that we want to compute the expectation of?}
-Then the PDFs can be counted from the solution $\uL(t)$, and the statistical expectations can be calculated by averaging along the trajectory
-\begin{equation}
-\left\langle F \left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T} F \left(u_{\Lambda}\left(t\right)\right)dt,
+Or is $F$ some arbitrary function of the state variable that we want to compute the expectation of?
+$F$ should be some arbitrary function for the expectation.}
+Then the PDFs can be counted from the solution $\uL(t)$, and the statistical expectation of any functional $f$ can be calculated by averaging along the trajectory
+\begin{equation}
+\left\langle f \left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T} f \left(u_{\Lambda}\left(t\right)\right)dt,
 \end{equation}

### 2019-11-14 paper2_JNLS/paper.tex
NM

--- 2019-11-13 paper2_JNLS/paper.tex
@@
 \newcommand{\meanz}[1]{\mean{#1}_0}
-% Misc stuff
+
+% Sampling and averaging
 \newcommand{\tavg}[1]{\overline{#1}}
@@
 \newcommand{\sumsamp}{\sum_{i=1}^{\Nsamp}}
-\newcommand{\traveldist}{\ell}
-\newcommand{\Ftrans}{\mathcal{F}}
+\newcommand{\Fth}{W}
 \newcommand{\RHS}{F}
 \newcommand{\RHSh}{\hat{\RHS}}
+\newcommand{\uhvec}{\mathbf{u}}
+\newcommand{\RHSvec}{\mathbf{F}}
 \newcommand{\usqdah}{\hat{v}} %{ (\widehat{u_{da}^2}) }
+
+% Scaling analysis
+\newcommand{\traveldist}{\ell}
 
@@
 
+	To compute the transfer function, we use a weighted random-sampling strategy.
+That is, we first sample the Fourier coefficients $\uhvec = \left\{ \uhat_{k}\right\}_{k=0}^{\Lambda}$ from a uniform distribution on the unit hypersphere $\En = 1$.
+The uniform sampling is achieved by first sampling from a isotropic Gaussian distribution and then normalizing to project onto the hypersphere \cite{abramov2003}.
+These samples are then weighted by the appropriate Gibbs measure \eqref{Gibbs} to compute the expectations needed in the statistical matching condition \eqref{statmatch}.
+That is, for an arbitrary quantity $Q$, the ensemble expectation corresponding to Hamiltonian $\Ham$ and inverse temperature $\theta$ is numerically approximated as
+\begin{equation}
+\label{weightmean}
+\mean{Q}_{\theta} = \frac{\sumsamp Q_i \exp(-\theta \Ham_i)} {\sumsamp \exp(-\theta \Ham_i)}
+\end{equation}
+where $\Nsamp$ is the number of samples. Then, enforcing the statistical matching condition \eqref{statmatch} becomes a root-finding problem for the function
+\begin{equation}
+\label{rooteq}
+\Fth(\thdn) =  \mean{\Hdn}_{\thdn} - \mean{\Hdn}_{\thup} = 0
+\end{equation}
+Recall that we consider $\thup$ as given, so that $\mean{\Hdn}_{\thup}$ is a constant that can be computed straightaway from \eqref{weightmean}.
+We then use the secant method with respect to the variable $\thdn$ to find a root of $\Fth(\thdn)$ to the desired tolerance.
+
+For simply computing the transfer function, this weighted-sampling approach offers significant advantages over more sophisticated methods, such as such Markov Chain Monte Carlo (MCMC), in its ability to reuse the same samples of $\uhvec$ for several different values of $\thdn$.
+That is, we sample $\uhvec$ from the uniform distribution and compute the list of Hamiltonian values only once, then simply vary $\thdn$ in \eqref{rooteq} with the secant method until the root is found.
+An MCMC method, on the other hand, would require the sampling to restart from scratch for each value of $\thdn$, since the random steps in MCMC depend directly on the target distribution \eqref{Gibbs}.
+We will however, use an MCMC method to initialize direct numerical simulations of TKdV due to the superior efficiency for a specific, given value of $\theta$.
+
 %----------------------------------------------------------------------------%
@@
 For time integration of \eqref{TKdVDisc}, we employ a 4th-order midpoint symplectic scheme \cite{mclachlan1993symplectic}.
-We introduce the spectral vectors $\mathbf{u}=\left\{ \hat{u}_{k}\right\}_{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$, and let $\mathbf{u}^{n}$ denote the solution at time $t_{n}$.
+We introduce the spectral vectors $\uhvec = \left\{ \uhat_{k}\right\}_{k=0}^{\Lambda}$ and $\RHSvec = \left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$, and let $\uhvec^n$ denote the solution at time $t_n$.
 The midpoint method has two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$:
 \begin{align}
-\label{SympSteps}
-&\mathbf{y}_{1}-\mathbf{u}^{n} = w_{1}\Delta t \, \mathbf{F}
+\label{SympStep1}
+&\mathbf{y}_{1}-\mathbf{u}^{n} = w_{1}\Delta t \, \RHSvec
 \left[ \frac{1}{2} \left(\mathbf{y}_{1}+\mathbf{u}^{n}\right) \right] ,\\
-&\mathbf{y}_{2}-\mathbf{y}_{1} = w_{2}\Delta t \, \mathbf{F}
+&\mathbf{y}_{2}-\mathbf{y}_{1} = w_{2}\Delta t \, \RHSvec
 \left[ \frac{1}{2} \left(\mathbf{y}_{2}+\mathbf{y}_{1}\right) \right] ,\\
-&\mathbf{u}^{n+1}-\mathbf{y}_{2} = w_{3}\Delta t \, \mathbf{F}
+\label{SympStep3}
+&\mathbf{u}^{n+1}-\mathbf{y}_{2} = w_{3}\Delta t \, \RHSvec
 \left[ \frac{1}{2} \left(\mathbf{u}^{n+1}+\mathbf{y}_{2}\right) \right] ,
 \end{align}
-with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$,
-$w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$.
+with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$, $w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$.
+The semi-implicit nature of \eqref{SympStep1}--\eqref{SympStep3} combined with the nonlinearity of $\RHSvec$ requires iteration.
+We split $\RHSvec$ into linear and nonlinear components, with the linear component being easily inverted since it is diagonal in spectral space.
+At each step of \eqref{SympStep1}--\eqref{SympStep3}, we perform Picard iteration on the nonlinear component until convergence is achieved with a tolerance of $\delta = 1\times10^{-10}$.
+The starting guess for the iterations is determined by quadratic extrapolation from the previous three stages.
 
@@
 
-\todo[inline]{Di, I do not understand this paragraph.
-Is the $F$ meant to be the same $F$ as above, i.e.~the right-hand-side of the TKdV equation \eqref{TKdV2}?
-Or is $F$ some arbitrary function of the state variable that we want to compute the expectation of?}
-Then the PDFs can be counted from the solution $\uL(t)$, and the statistical expectations can be calculated by averaging along the trajectory
-\begin{equation}
-\left\langle F \left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T} F \left(u_{\Lambda}\left(t\right)\right)dt,
-\end{equation}
-given ergodicity in the solution with $\uL(t_0)$ already reaching the stationary state.
-
-\todo[inline]{I am also a little confused about the paragraphs below.
-Eq.~\eqref{iterate} is not the same form of Eq.~(43)--(45) because in those equations it is not simply $\mathbf{z}$ inside $\mathbf{F}$, but instead an average of $\mathbf{z}$ and the previous value.
-Am I simply misinterpreting?}
-\todo[inline]{Also, this is simply Picard iteration, correct?
-Perhaps we could simply say in words that we perform Picard iteration so that we don't have to worry about getting the equations correct in detail.
-What's your opinion?}
-
-For the symplectic schemes, implicit iteration method is required to get the accurate solution. For any implicit problem of the form
-\begin{equation}
-\label{iterate}
-\mathbf{z}-\mathbf{u}^{n}=\tau \mathbf{F \left(z\right)} = \tau\mathbf{Lz}+\tau\mathbf{N\left(z\right)},
-\end{equation}
-we split the operators into a linear and nonlinear components and only treat the nonlinear part on the right hand side for the previous stage in the iteration.
-Then we have the iterative scheme for the solution of implicit problem
-\begin{equation}
-\left(1-\tau\mathbf{L}\right)\mathbf{z}^{\left[l+1\right]} = \mathbf{u}^{n}+\tau\mathbf{N}\left(\mathbf{z}^{\left[l\right]}\right).
-\end{equation}
-The linear operator $\mathbf{L}$ is the diagonal differential operator, thus it is easy to invert in these schemes. The termination criteria is determined by
-\begin{equation}
-\left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert <\delta\;\mathrm{and}\;\left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert \geq\left\Vert \mathbf{z}^{\left[l\right]}-\mathbf{z}^{\left[l-1\right]}\right\Vert .
-\end{equation}
-It is found that the iteration tolerance can be set as $\delta=1\times10^{-10}$.
-Finally, the starting guess for the iterations can be found by quadratic extrapolation for the previous three stages $\left(t_{u},\mathbf{u}^{n}\right), \left(t_{u}-\tau,\mathbf{u}^{n-1}\right), \left(t_{u}-2\tau,\mathbf{u}^{n-2}\right)$.
-Therefore we can find the extrapolated value at $t=t_{n}+\beta\tau$ by
-\begin{equation}
-\mathbf{u}^{n+\beta}=\mathbf{Q}\left(t_{n}+\beta\tau\right)=\left[\frac{1}{2}\beta\left(\beta+3\right)+1\right]\mathbf{u}^{n}-\beta\left(\beta+2\right)\mathbf{u}^{n-1}+\frac{1}{2}\beta\left(\beta+1\right)\mathbf{u}^{n-2}.
-\end{equation}
-
-%The symplectic numerical integrators discretize the Hamiltonian and the Poisson structures separated and then form the resulting scheme for conserving the Hamiltonian.
-
-\subsection{Sampling from the Gibbs measures}
-
-\nick{Include a short description of the sampling method}
-
-\begin{equation}
-\mean{Q} = \frac{\int Q \exp\left(-\theta \Ham \right) d\mu}{ \int \exp\left(-\theta \Ham \right) d\mu}
-\approx \frac{\sumsamp Q_i \exp(-\theta \Ham_i)} {\sumsamp \exp(-\theta \Ham_i)}
-\end{equation}
+\nick{We should simply comment that we use MHMC to initialize the direct simulations.}
 

### 2019-11-14 paper2_JNLS/paper.tex
NM

@@
 %----------------------------------------------------------------------------%
+% Deterministic simulations of TKdV
+%----------------------------------------------------------------------------%
+\subsection{Deterministic simulations of TKdV}
+
+We now detail the method for direct numerical simulation of the TKdV dynamical system \eqref{TKdV}.
+In particular, since the conservation of energy and Hamiltonian plays a central role in the emergent statistical features of the system, it is important for the numerical scheme to conserve these quantities over long time horizons.
+We therefore employ a symplectic integrator, which, by preserving oriented areas in phase space, conserves the energy and Hamiltonian exactly.
+
+Rearranging the TKdV system \eqref{TKdV} gives
+\begin{align}
+\label{TKdV2}
+&\pd{\uL}{t} = -  \frac{1}{2} C_3 \drat^{-3/2} \, \pd{}{x} \Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = \RHS[\uL]
+%\qquad \text{for } x \in [-\pi,\pi]
+\end{align}	
+where we have represented the right-hand side by the operator $\RHS[\uL]$.
+We employ a pseudo-spectral discretization of \eqref{TKdV2}, with de-aliasing applied to the quadratic nonlinear term $\Proj (\uL)^2$ according to the standard 2/3-rule.
+That is, we first pad $\uhat_k$ with zeros for $\Lambda < \abs{k} \le 3\Lambda/2$, transform to physical space and square to obtain $\uL^2$ on a fine grid, then transform back to spectral space and truncate to obtain the Fourier coefficients of $\Proj (\uL)^2$
+For simplicity, we denote these Fourier coefficients $\usqdah_k$,
+\begin{equation}
+\Proj (\uL)^2 = \sum_{\abs{k} \le \Lambda} \usqdah_k e^{i k x}
+\end{equation}
+With this discretization, \eqref{TKdV2} can be recast in spectral space as a nonlinear ODE system
+\begin{equation}
+\label{TKdVDisc}
+\td{}{t} \uhat_k =  -\frac{1}{2} C_3 \drat^{-3/2} \, ik \usqdah_k + C_2 \drat^{1/2} \, ik^{3} \uhat_k = \RHSh_k
+\end{equation}
+The quadratic nonlinearity present in this system, $\usqdah_k$, mixes the modes during evolution. We note the third-order linear term may become stiff for large $\Lambda$.
+
+For time integration of \eqref{TKdVDisc}, we employ a 4th-order midpoint symplectic scheme \cite{mclachlan1993symplectic}.
+We introduce the spectral vectors $\uhvec = \left\{ \uhat_{k}\right\}_{k=0}^{\Lambda}$ and $\RHSvec = \left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$, and let $\uhvec^n$ denote the solution at time $t_n$.
+The midpoint method has two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$:
+\begin{align}
+\label{SympStep1}
+&\mathbf{y}_{1} - \uhvec^n = w_{1}\Delta t \, \RHSvec
+\left[ \frac{1}{2}  \left(\mathbf{y}_{1} + \uhvec^n \right) \right] ,\\
+&\mathbf{y}_{2}-\mathbf{y}_{1} = w_{2}\Delta t \, \RHSvec
+\left[ \frac{1}{2} \left(\mathbf{y}_{2} + \mathbf{y}_{1}\right) \right] ,\\
+\label{SympStep3}
+&\uhvec^{n+1}-\mathbf{y}_{2} = w_{3}\Delta t \, \RHSvec
+\left[ \frac{1}{2} \left( \uhvec^{n+1} + \mathbf{y}_{2}\right) \right] ,
+\end{align}
+with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$, $w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$.
+The semi-implicit nature of \eqref{SympStep1}--\eqref{SympStep3} combined with the nonlinearity in $\RHSvec$ requires iteration.
+We split $\RHSvec$ into linear and nonlinear components, with the linear component being easily inverted since it is diagonal in spectral space.
+At each step of \eqref{SympStep1}--\eqref{SympStep3}, we perform Picard iteration on the nonlinear component until convergence is achieved with a tolerance of $\delta = 1\times10^{-10}$.
+The starting guess for the iterations is determined by quadratic extrapolation from the previous three stages.
+
+	For the initial conditions, $\uhvec^0$, of the direct numerical simulation we sample from the upstream Gibbs ensemble $\Gup$ in \eqref{Gibbs} with a prescribed inverse temperature $\thup$.
+We achieve this sampling via a Metropolis-Hasting Monte-Carlo algorithm as detailed in \cite{majda2019}.
+
+%----------------------------------------------------------------------------%
 % Scaling
@@
 We, however, leave that task for future research due to the significant cost of performing an entirely new set of experiments compared to the relative ease and great value in re-analyzing existing experimental measurements in light of the new theoretical developments.
-
-%----------------------------------------------------------------------------%
-% Numerical methods
-%----------------------------------------------------------------------------%
-\subsection{Deterministic simulations of TKdV}
-
-%Two approaches can be applied to predict the statistics in the downstream wave disturbance, which can be used to confirm each other in the numerical results.
-The equilibrium statistical mechanics is an easy way to confirm the matching condition and get the equilibrium distribution in far-field long-time limit.
-On the other hand, the direct simulations of the dynamical model can offer estimates about the transient statistics before the far-field equilibrium state is reached.
-
-We now detail the method for direct numerical simulation of the TKdV dynamical system \eqref{TKdV}.
-In particular, since the conservation of energy and Hamiltonian plays a central role in the emergent statistical features of the system, it is important for the numerical scheme to conserve these quantities over long time horizons.
-We therefore employ a symplectic integrator, which, by preserving oriented areas in phase space, conserves the energy and Hamiltonian exactly.
-
-Rearranging the TKdV system \eqref{TKdV} gives
-\begin{align}
-\label{TKdV2}
-&\pd{\uL}{t} = -  \frac{1}{2} C_3 \drat^{-3/2} \, \pd{}{x} \Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = \RHS[\uL]
-%\qquad \text{for } x \in [-\pi,\pi]
-\end{align}	
-where we have represented the right-hand side by the operator $\RHS[\uL]$.
-Our numerical scheme first separates the Hamiltonian into linear and nonlinear components.
-Then a pseudo-spectral scheme is used for spatial discretization with de-aliasing applied to the nonlinear term $\Proj (\uL)^2$ according to the standard 2/3-rule.
-That is, we first pad $\uhat_k$ with zeros for $\Lambda < \abs{k} \le 3\Lambda/2$, transform to physical space and square to obtain $\uL^2$ on a fine grid, then transform back to spectral space and truncate to obtain the Fourier coefficients of $\Proj (\uL)^2$
-For simplicity, we denote these Fourier coefficients $\usqdah_k$. That is,
-\begin{equation}
-\Proj (\uL)^2 = \sum_{\abs{k} \le \Lambda} \usqdah_k e^{i k x}
-\end{equation}
-With this spectral discretization, the ODE system \eqref{TKdV2} can be written as
-\begin{equation}
-\label{TKdVDisc}
-\td{}{t} \uhat_k =  -\frac{1}{2} C_3 \drat^{-3/2} \, ik \usqdah_k + C_2 \drat^{1/2} \, ik^{3} \uhat_k = \RHSh_k
-\end{equation}
-Due to the nonlinearity, this system mixes scales during evolution. We note the third-order linear term may become stiff for large $\Lambda$.
-
-For time integration of \eqref{TKdVDisc}, we employ a 4th-order midpoint symplectic scheme \cite{mclachlan1993symplectic}.
-We introduce the spectral vectors $\uhvec = \left\{ \uhat_{k}\right\}_{k=0}^{\Lambda}$ and $\RHSvec = \left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$, and let $\uhvec^n$ denote the solution at time $t_n$.
-The midpoint method has two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$:
-\begin{align}
-\label{SympStep1}
-&\mathbf{y}_{1}-\mathbf{u}^{n} = w_{1}\Delta t \, \RHSvec
-\left[ \frac{1}{2} \left(\mathbf{y}_{1}+\mathbf{u}^{n}\right) \right] ,\\
-&\mathbf{y}_{2}-\mathbf{y}_{1} = w_{2}\Delta t \, \RHSvec
-\left[ \frac{1}{2} \left(\mathbf{y}_{2}+\mathbf{y}_{1}\right) \right] ,\\
-\label{SympStep3}
-&\mathbf{u}^{n+1}-\mathbf{y}_{2} = w_{3}\Delta t \, \RHSvec
-\left[ \frac{1}{2} \left(\mathbf{u}^{n+1}+\mathbf{y}_{2}\right) \right] ,
-\end{align}
-with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$, $w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$.
-The semi-implicit nature of \eqref{SympStep1}--\eqref{SympStep3} combined with the nonlinearity of $\RHSvec$ requires iteration.
-We split $\RHSvec$ into linear and nonlinear components, with the linear component being easily inverted since it is diagonal in spectral space.
-At each step of \eqref{SympStep1}--\eqref{SympStep3}, we perform Picard iteration on the nonlinear component until convergence is achieved with a tolerance of $\delta = 1\times10^{-10}$.
-The starting guess for the iterations is determined by quadratic extrapolation from the previous three stages.
-
-\vsp{5}
-
-\nick{We should simply comment that we use MHMC to initialize the direct simulations.}
 

### 2019-11-14 paper2_JNLS/paper.tex
NM

@@
 It should be noted that, in these studies, extreme events are represented by strong skewness of the wave-field.
-These events occur intermittently and on a relatively frequent basis, which is in contrast to the simpler situation of isolated, rare events \cite{guth2019machine}.
-
-The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}.
+They occur intermittently and on a relatively frequent basis, which is in contrast to the simpler situation of isolated, rare events \cite{guth2019machine}.
+
+The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al.~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}.
 Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
-We first provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two.
+In Section \ref{experiments} we detail the laboratory experiments of Bolles et al.~and summarize previously reported findings on anomalous statistics of the surface displacement.
+In this section, we also present new experimental data on surface {\em slope} statistics --- a line of inquiry motivated by the appearance of the slope in the Hamiltonian structure of TKdV theory developed by Majda et al.~\cite{majda2019}.
+In Section \ref{theory}, we detail the TKdV theoretical framework, with a more thorough treatment of the link between model and experimental parameters than was reported in previous work.
+This link facilitates comparison between the theory and experiments.
+Within the same TKdV framework, we discuss both a deterministic dynamical systems approach and a statistical mechanics approach.
+The latter is based on the Hamiltonian structure of TKdV and novel ensemble distributions that incorporate canonical and micro-canonical aspects.
+Section \ref{results} presents a systematic comparison between experimental measurements and model predictions, with both surface displacement and surface slope statistics compared closely.
+This section, also describes deterministic dynamics of TKdV, which offer insight into the experimental length and timescales required for wave statistics to transition.
+Importantly, this section shows newly collected experimental data related to a a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
+The experimental data provides conclusive confirmation of the model prediction.
+We close with some final remarks in Section \ref{conclusion}.
+
+%confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}. 
+% spectacular, remarkable, marvelous, conclusively, convincingly
+
+%We first provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two.
 Next, we present new experimental data, most notably on {\em surface slope} statistics, which motivates parallel examination of the theory.
@@
 
-% OLD VERSION
-\begin{comment}
-We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. 
-Novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
-Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}.
-In addition, we provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two. 
-In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.
-\end{comment}
-
 % NOTES
 % Words: conjunction, combination, close cooperation; elucidate.
-% New data: slopes, autocorrelation, depth ratio varied.
 
@@
 \section{Comparison between theory and experiments}
+\label{results}
 
@@
 \section{Concluding remarks}
+\label{conclusion}
 

### 2019-11-16 paper2_JNLS/paper.tex
NM

@@
 and the Benjamin-Feir deep-water modulational instability \citep{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
-The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these events occur much more frequently when surface statistics deviate from Gaussian.
+The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement: when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these events occur much more frequently when surface statistics deviate from Gaussian.
 In this way, anomalous waves can be approached from the broader perspective of turbulent dynamical systems \cite{sapsis2013a, sapsis2013b, sapsis2013blending, chen2016filtering, majda2016introduction, macedo2017universality, MajdaQiSIAM2018, blonigan2019extreme, guth2019machine}.
@@
 % Words: conjunction, combination, close cooperation; elucidate.
+% New data: slopes, autocorrelation, depth ratio varied.
 
@@
 Since the dampener minimizes backscatter, the waves in this experiment propagate primarily in one direction, from left to right in Fig.~\ref{ExpDiagStats}(a).
-
-The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized wave field, the paddle angle $\phi$ is specified by a psuedo-random signal
-\begin{align}
-\label{PaddleAngle}
-& \phi(t) = \phi_0 + \Dphi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
-\label{anEq}
-& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
-\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
-\end{align}
-The angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively.
-As in prior work, all experiments reported here use the values $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz.
-The phases $\delta_n$ are uniformly distributed random variables, which results in a randomized wave train.
-The standard deviation of the paddle angle, $\Dphi$, controls the overall amplitude of the waves.
-In a single experiment $\Dphi$ is fixed, and we will present a series of experiments with $\Dphi$ varied systematically.
 
@@
 
+The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized wave field, the paddle angle $\phi$ is specified by a psuedo-random signal
+\begin{align}
+\label{PaddleAngle}
+& \phi(t) = \phi_0 + \Dphi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
+\label{anEq}
+& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
+\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
+\end{align}
+The angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively.
+As in prior work, all experiments reported here use the values $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz.
+The phases $\delta_n$ are uniformly distributed random variables, which results in a randomized wave train.
+The standard deviation of the paddle angle, $\Dphi$, controls the overall amplitude of the waves.
+In a single experiment $\Dphi$ is fixed, and we will present a series of experiments with $\Dphi$ varied systematically.
+
 The free surface is illuminated by light-emitting diodes and is imaged from the sideview with a Nikon D3300 at 60 frames per second.
@@
 Figure \ref{ExpSpatialStats} shows long-time statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ as they vary in space for six different driving amplitudes (see legend).
+
+We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
+\begin{align}
+%& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
+& \std(q) = q_{std}= \sqrt{\mean{q^2}}
+&&\mbox{\em standard deviation} \\
+& \skw(q) = {\mean{q^3}} / {q_{std}^3}	
+&&\mbox{\em skewness} \\
+& \kurt(q) = {\mean{q^4}} / {q_{std}^4} - 3
+&&\mbox{\em (excess) kurtosis}
+\end{align}
+where $\mean{}$ indicates a mean (here a long-time mean). Hereafter, we will simply refer to the excess kurtosis as kurtosis.
+
+	Figure \ref{ExpSpatialStats} shows how these statistics vary in the vicinity of the ADC, located at $x = 0$, for both displacement and slope.
+First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves.
+Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC.
+The skewness and kurtosis, however, respond strongly to the ADC as long as the amplitude is sufficiently high.
+As seen in \ref{ExpSpatialStats}(b)--(c), both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm.
+Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude.
+The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e).
+Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution.
+We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.
 
@@
  %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
-
-	We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
-\begin{align}
-%& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
-& \std(q) = q_{std}= \sqrt{\mean{q^2}}
-&&\mbox{\em standard deviation} \\
-& \skw(q) = {\mean{q^3}} / {q_{std}^3}	
-&&\mbox{\em skewness} \\
-& \kurt(q) = {\mean{q^4}} / {q_{std}^4} - 3
-&&\mbox{\em (excess) kurtosis}
-\end{align}
-where $\mean{}$ indicates a mean (here a long-time mean). Hereafter, we will simply refer to the excess kurtosis as kurtosis.
-
-	Figure \ref{ExpSpatialStats} shows how these statistics vary in the vicinity of the ADC, located at $x = 0$, for both displacement and slope.
-First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves.
-Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC.
-The skewness and kurtosis, however, respond strongly to the ADC as long as the amplitude is sufficiently high.
-As seen in \ref{ExpSpatialStats}(b)--(c), both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm.
-Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude.
-The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e).
-Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution.
-We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.
 

### 2019-11-16 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\usqdah}{\hat{v}} %{ (\widehat{u_{da}^2}) }
-
 % Scaling analysis
 \newcommand{\traveldist}{\ell}
+% et al
+\newcommand{\boetal}{Bolles {\it et al.}~(2019)~}
+\newcommand{\maetal}{Majda {\it et al.}~(2019)~}
 
@@
 
-Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC).
+Our particular focus is the laboratory experiments of \boetal~\cite{bolles2019}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC).
 In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
 Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
-Inspired by these experimental observations, Majda et al.~\cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior.
+Inspired by these experimental observations, \maetal~\cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior.
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
@@
 
-The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al.~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}.
+The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of \boetal~\cite{bolles2019} and \maetal~\cite{majda2019, majdaqi2019}.
 Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
-In Section \ref{experiments} we detail the laboratory experiments of Bolles et al.~and summarize previously reported findings on anomalous statistics of the surface displacement.
-In this section, we also present new experimental data on surface {\em slope} statistics --- a line of inquiry motivated by the appearance of the slope in the Hamiltonian structure of TKdV theory developed by Majda et al.~\cite{majda2019}.
+In Section \ref{experiments} we detail the laboratory experiments of \boetal~and summarize previously reported findings on anomalous statistics of the surface displacement.
+In this section, we also present new experimental data on surface {\em slope} statistics --- a line of inquiry motivated by the appearance of the slope in the Hamiltonian structure of TKdV theory developed by \maetal~\cite{majda2019}.
 In Section \ref{theory}, we detail the TKdV theoretical framework, with a more thorough treatment of the link between model and experimental parameters than was reported in previous work.
@@
 
-%confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}. 
+%confirms the predictive power of the framework developed by \maetal~\cite{majda2019}. 
 % spectacular, remarkable, marvelous, conclusively, convincingly
@@
 Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
-Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}.
+Comparison with the new measurements further confirms the predictive power of the framework developed by \maetal~\cite{majda2019}.
 
@@
 The upstream measurements are symmetrically distributed about the mean, $\eta = 0$.
-In fact, Bolles et al.~found that these measurements follow a Gaussian distribution closely \cite{bolles2019}.
+In fact, \boetal found that these measurements follow a Gaussian distribution closely \cite{bolles2019}.
 The downstream measurements, however, skew strongly towards positive displacement, $\eta > 0$.
-Bolles et al.~found these measurements to be well described by a mean-zero gamma distribution \cite{bolles2019}.
+\boetal found these measurements to be well described by a mean-zero gamma distribution \cite{bolles2019}.
 The slower decay rate of this downstream distribution indicates an elevated level of extreme surface displacement, i.e.~rogue waves.
-Bolles et al.~estimated that a rogue wave can be up to 65 times more likely in these experiments than if displacements were Gaussian \cite{bolles2019}.
+\boetal estimated that a rogue wave can be up to 65 times more likely in these experiments than if displacements were Gaussian \cite{bolles2019}.
 
@@
 The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e).
-Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution.
+\boetal found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution.
 We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.
@@
 	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}.
-Our attention to slope statistics was motivated by new theoretical developments by Majda et al.~\cite{majda2019} and will be expanded upon in later sections.
+Our attention to slope statistics was motivated by new theoretical developments by \maetal \cite{majda2019} and will be expanded upon in later sections.
 We extract the surface slope by numerical differentiating images of the free surface using Savitzky-Golay smoothing filters.
@@
 
-	Bolles et al.~systematically explored the effect of varying the driving amplitude on the anomalous statistics for a fixed depth ratio. 
+	\boetal systematically explored the effect of varying the driving amplitude on the anomalous statistics for a fixed depth ratio. 
 Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we also report experiments in which the depth ratio was varied. 
@@
 
-In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from Majda et al.~2019 \cite{majda2019}.
+In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from \maetal \cite{majda2019}.
 Since the choice of normalization is not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices.
@@
 
-We remark that, in defining the Hamiltonian, we have chosen the sign convention of Lax \cite{lax1975periodic}.
-More recent work of Bajars et al.~\cite{bajars2013weakly} and Majda et al.~\cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite.
+We remark that, in defining the Hamiltonian, we have chosen the sign convention of Lax (1975) \cite{lax1975periodic}.
+More recent work of Bajars {\it et al.}~(2013) \cite{bajars2013weakly} and \maetal \cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite.
 Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent.
-Using the second convention, Majda et al.~found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}.
+Using the second convention, \maetal found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}.
 We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.
@@
 
-	In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced by Abramov et al.~for the Burgers-Hopf system \cite{abramov2003}.
+	In examining statistical mechanics of this Hamiltonian system, we will appeal to the idea of a {\em mixed microcanonical-canonical} Gibbs ensemble, as originally introduced by Abramov {\it et al.}~(2003) for the Burgers-Hopf system \cite{abramov2003}.
 Specifically, this ensemble is microcanonical in energy and canonical in the Hamiltonian.
@@
 
-	This statistical matching condition, originally derived by Majda et al.~(2019) \cite{majda2019}, imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
+	This statistical matching condition, originally derived by \maetal \cite{majda2019}, imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
 In particular, we view $\thup$ as given by the random-state of the incoming wave field.
@@
 We now discuss arguably the most novel single result of the manuscript.
-In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}.
+In the recent theoretical study, \maetal derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}.
 More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$.
@@
 
-	This manuscript extends the parallel experimental and modeling efforts of Bolles et al.~\cite{bolles2019} and Majda et al.~\cite{majda2019} concerning the emergence of anomalous wave statistics from abrupt changes in bottom topography.
+	This manuscript extends the parallel experimental and modeling efforts of \boetal \cite{bolles2019} and \maetal \cite{majda2019} concerning the emergence of anomalous wave statistics from abrupt changes in bottom topography.
 The theoretical framework is based on deterministic and statistical analysis of the TKdV system, with exploitation of the Hamiltonian structure and the associated invariant Gibbs measures.
@@
 
-	Finally, Majda et al.~\cite{majda2019} derived an explicit formula for the skewness of the outgoing wave field, which we have tested against direct experimental measurements.
+	Finally, \maetal \cite{majda2019} derived an explicit formula for the skewness of the outgoing wave field, which we have tested against direct experimental measurements.
 Specifically, the formula predicts how the ratio of displacement skewness to slope variance, $\skw(\eta) / \Delta \var(\eta_x)$, depends on wave amplitude.

### 2019-11-16 paper2_JNLS/paper.tex
NM

@@
 In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change.
-Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
-Inspired by these experimental observations, \maetal~\cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior.
+Upon passing over the step, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC.
+Inspired by these experimental observations, \maetal \cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior.
 The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools.
-In subsequent work, Majda and Qi analyzed more severely truncated systems --- as low as two interacting modes --- that exhibit a statistical phase transition to anomalous statistics and the creation of extreme events while enjoying a more tractable structure \cite{majdaqi2019}.
-More recently, Qi and Majda demonstrated the capability of machine learning strategies to accurately predict these extreme events \cite{MachineLearning2019}.
+In subsequent work, Majda \& Qi (2019) \cite{majdaqi2019} analyzed more severely truncated systems --- as low as two interacting modes --- that exhibit a statistical phase transition to anomalous statistics and the creation of extreme events while enjoying a more tractable structure.
+More recently, Qi \& Majda (2019) \cite{MachineLearning2019} demonstrated the capability of machine learning strategies to accurately predict these extreme events.
 That work uses a deep neural network coupled with a judicious choice of the entropy loss function that emphasizes the dominant structures of the turbulent field \cite{MachineLearning2019}.
@@
 Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
-In Section \ref{experiments} we detail the laboratory experiments of \boetal~and summarize previously reported findings on anomalous statistics of the surface displacement.
-In this section, we also present new experimental data on surface {\em slope} statistics --- a line of inquiry motivated by the appearance of the slope in the Hamiltonian structure of TKdV theory developed by \maetal~\cite{majda2019}.
-In Section \ref{theory}, we detail the TKdV theoretical framework, with a more thorough treatment of the link between model and experimental parameters than was reported in previous work.
-This link facilitates comparison between the theory and experiments.
-Within the same TKdV framework, we discuss both a deterministic dynamical systems approach and a statistical mechanics approach.
-The latter is based on the Hamiltonian structure of TKdV and novel ensemble distributions that incorporate canonical and micro-canonical aspects.
-Section \ref{results} presents a systematic comparison between experimental measurements and model predictions, with both surface displacement and surface slope statistics compared closely.
-This section, also describes deterministic dynamics of TKdV, which offer insight into the experimental length and timescales required for wave statistics to transition.
-Importantly, this section shows newly collected experimental data related to a a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
-The experimental data provides conclusive confirmation of the model prediction.
+In Section \ref{experiments} we detail the laboratory experiments of \boetal \cite{bolles2019} and summarize previously reported findings on anomalous statistics of the surface displacement.
+We also present new experimental data on surface {\em slope} statistics --- a line of inquiry motivated by the theoretical advancements of \maetal \cite{majda2019}, specifically the central role played by the slope in the Hamiltonian structure of TKdV.
+Section \ref{theory} discusses the TKdV theoretical framework, including both the deterministic and the statistical mechanics perspectives.
+In this section, we flesh out details of the non-dimensionalization, which, for brevity, were only briefly discussed in previous work.
+This exercise provides a more thorough link between model and experimental parameters, ultimately facilitating a closer comparison between the two.
+We then discuss TKdV as a deterministic dynamical system, approachable from the viewpoint of statistical mechanics.
+The later perspective is based on the Hamiltonian structure and novel ensemble distributions that incorporate canonical and micro-canonical aspects \cite{abramov2003}.
+Section \ref{results} presents a systematic comparison between experiments and theory, focusing on both surface displacement and surface slope.
+We find the statistical distributions of these two quantities to agree well across experiments and theory.
+In particular, we find the agreement in displacement statistics to be remarkably quantitative.
+We then examine numerical simulations of the TKdV deterministic dynamics, which elucidates the length and timescales required for statistical transitions in the experiments.
+Finally, this section discusses a peculiar power-law relationship between surface-displacement skewness and surface-slope variance predicted by the statistical mechanics framework \cite{majda2019}.
+New experimental measurements conclusively confirm this prediction, further demonstrating the predictive power of the TKdV framework developed by \maetal \cite{majda2019}.
 We close with some final remarks in Section \ref{conclusion}.
-
-%confirms the predictive power of the framework developed by \maetal~\cite{majda2019}. 
-% spectacular, remarkable, marvelous, conclusively, convincingly
-
-%We first provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two.
-Next, we present new experimental data, most notably on {\em surface slope} statistics, which motivates parallel examination of the theory.
-Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory.
-Comparison with the new measurements further confirms the predictive power of the framework developed by \maetal~\cite{majda2019}.
-
-% NOTES
-% Words: conjunction, combination, close cooperation; elucidate.
-% New data: slopes, autocorrelation, depth ratio varied.
 

### 2019-11-16 paper2_JNLS/paper.tex
NM

@@
 \boetal found these measurements to be well described by a mean-zero gamma distribution \cite{bolles2019}.
-The slower decay rate of this downstream distribution indicates an elevated level of extreme surface displacement, i.e.~rogue waves.
+The slower decay of the gamma distribution indicates an elevated level of extreme surface displacement, i.e.~rogue waves.
 \boetal estimated that a rogue wave can be up to 65 times more likely in these experiments than if displacements were Gaussian \cite{bolles2019}.
@@
 
-We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
+In this paper, we only examine statistic of mean-zero quantities, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
 \begin{align}
@@
 \end{align}
-where $\mean{}$ indicates a mean (here a long-time mean). Hereafter, we will simply refer to the excess kurtosis as kurtosis.
+where $\mean{}$ indicates a mean --- here a long-time mean at a fixed spatial location. Hereafter, we will simply refer to the excess kurtosis as kurtosis.
 
@@
 	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}.
-Our attention to slope statistics was motivated by new theoretical developments by \maetal \cite{majda2019} and will be expanded upon in later sections.
+Our attention to slope statistics was motivated by new theoretical developments by \maetal \cite{majda2019}, as will be expanded upon in later sections.
 We extract the surface slope by numerical differentiating images of the free surface using Savitzky-Golay smoothing filters.
@@
 As seen in Fig.~\ref{ExpSpatialStats}(a) the standard deviation of slope behaves similar to displacement: $\std(\eta_x)$ increases with driving amplitude, but remains nearly uniform in space for each individual experiment and is not affected by the ADC.
-Once again though, the higher-order moments respond strongly to the ADC if the driving amplitude is sufficiently high.
+The higher-order moments, however, respond strongly to the ADC if the driving amplitude is sufficiently high.
 The skewness of slope becomes highly negative and reaches a minimum near $x = 8$ cm, which is downstream of the ADC but upstream of the position where the displacement statistics peak.
@@
 
-	\boetal systematically explored the effect of varying the driving amplitude on the anomalous statistics for a fixed depth ratio. 
-Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we also report experiments in which the depth ratio was varied. 
 To have a simple baseline for comparison against theory, we refer to the data from Fig.~\ref{ExpDiagStats} as the {\em reference} experiments.
@@
 From the data represented in Fig.~\ref{ExpSpatialStats}, we extract a characteristic wave amplitude for the representative experiments of $\etastd = 0.21$ cm, which will be important for setting dimensionless parameters that enter the theory.
-Table \ref{paramtable} lists the range of experimental parameters along with these representative values, as well as values of dimensionless parameters that will be introduced later.
-% NOTE: Maybe modify slightly the last paragraph if it does not seem to fit next time I read it.
+Table \ref{paramtable} lists the range of experimental parameters and the representative values, as well as values of dimensionless parameters that will be introduced later.
 
@@
 The explanation for these choices is as follows.
-First, we have chosen the characteristic amplitude, $\ampscale$, above to normalize the energy of the state-variable $u$ to unity, as will be demonstrated in Section \ref{tKdVSec}.
+First, we have chosen the characteristic amplitude, $\ampscale$, to normalize the energy of the state-variable $u$ to unity, as will be demonstrated in Section \ref{tKdVSec}.
 Second, regarding $\lengthscale$, recall that $\lam$ is the characteristic wavelength corresponding to the peak forcing frequency $\freqp$ in the experiments.
@@
 	Lastly, regarding the characteristic timescale $\timescale$, the most basic timescale in the experiments is simply $\freqp^{-1}$, i.e.~the period of waves passing a fixed reference point.
-Of course, the leading-order behavior in shallow water is simple translation of waves with uniform speed $c$.
+Of course, the leading-order behavior in shallow water is simply wave propagation with uniform speed $c$, i.e.~no dispersion.
 The KdV equation provides the first correction to this behavior and describes dynamics that evolve over longer timescales.
-Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitable long timescale in line with previous normalizations \cite{johnson1997modern}.
+Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitably long timescale in line with previous normalizations \cite{johnson1997modern}.
 The scales $\lengthscale = \lengthscale_{\pm}$ and $\timescale = \timescale_{\pm}$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.
@@
 \end{align}
-where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham = \Hupdn$ is the Hamiltonian, which will take different forms on either side of the ADC.
+where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham = \Hupdn$ is the Hamiltonian, which takes different forms on either side of the ADC.
 It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
@@
 The sign difference between the two \eqref{Hamiltonian} thus represents a competition between wave skewness and slope energy.
-This is the reason for our interest in experimental measurements of the surface {\it slope} statistics in the current manuscript.
+In particular, the role played by the slope energy is the motivating factor for the new experimental measurements on surface {\em slope} statistics that are reported in this paper.
 
@@
 
-We now introduce the truncated KdV (TKdV) system, which is the main focus of this study. Consider the Fourier series of the state variable
+We now introduce the truncated KdV (TKdV) system, which is the main focus of the present study. Consider the state variable represented as a spatial Fourier series
 \begin{align}
@@
 \label{uhat}
-&\uhat_k(t) = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x,t) \, e^{-i k x} \dx
-\end{align}
-Since $u$ is real valued, $\uhat_{-k} = \uhat_{k}^*$, and since momentum vanishes $\uhat_0 = 0$.
+&\uhat_k(t) = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x,t) \, e^{-i k x} \dx \, ,
+\end{align}
+where $\uhat_k(t) \in \mathbb{C}$. Since $u(x,t)$ is real valued, $\uhat_{-k} = \uhat_{k}^*$, and since momentum vanishes $\uhat_0 = 0$.
 Next, consider the Galerkin truncation at wave number $\Lambda$
@@
 \end{align}
-where $\Proj$ is a projection operator and \eqref{uhat} still holds. Projecting the KdV equation \eqref{dimlessKdV} onto the finite dimensional space gives the truncated KdV equation (TKdV)
+where $\Proj$ is a projection operator and \eqref{uhat} still holds.
+Inserting the projected variable, $\uL$, into the KdV equation and applying the projection operator, $\Proj$, again where necessary produces the truncated KdV equation (TKdV)
 \begin{align}
@@
 Note the additional projection operator in front of the quadratic term $\uL^2$, which removes the aliased modes of wavenumber larger than $\Lambda$.
+Since all wavenumbers larger than $\Lambda$ have been removed, \eqref{TKdV} represents a {\em finite} dimensional dynamical system, of dimension $\Lambda$ over $\mathbb{C}$.
 The constants $C_3$ and $C_2$ are the same as before and have been repeated here for convenience.

### 2019-11-16 paper2_JNLS/paper.tex
NM

@@
 
-	More specifically, we assume continuity of the surface displacement, $\eta$, at the ADC \nick{(cite)}.
-Since the temporal fluctuations of $\eta$ match, waves crossing the ADC must rapidly adjust in wavelength so as to attain the same frequency in two different depths --- a well-understood phenomenon \nick{(cite)}.
-Of course, since the normalized domain $x \in [-\pi,\pi]$ is scaled on the characteristic wavelength, there is no change in the dimensionless wave-field $u(x,t)$, giving the condition
+	More specifically, we assume continuity of the surface displacement, $\eta$, across the ADC \cite{whitham2011linear, rey1992propagation}.
+Since the propagation speed $c$ changes with depth, waves crossing the ADC must rapidly adjust in wavelength in order to match the oscillation frequency just upstream of the ADC \cite{whitham2011linear, rey1992propagation}.
+The normalized domain $x \in [-\pi,\pi]$ is scaled on the characteristic wavelength and so there is no change in the dimensionless wave-field $u(x,t)$, giving the condition
 \begin{equation}
@@
 We call \eqref{detmatch} the deterministic matching condition to contrast with its statistical counterpart introduced below.
-This matching condition is employed in the deterministic simulations of the TKdV dynamical system \eqref{TKdV}.
+This matching condition is employed in the deterministic simulations of TKdV \eqref{TKdV}.
 We note that in the alternate formulation mentioned earlier, it is the product $\depth^{1/4} \eta$ that would match at the ADC \cite{johnson1997modern}.
@@
 In particular, we view $\thup$ as given by the random-state of the incoming wave field.
-Thus, while we will treat $\thup$ as a free parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by the matching condition \eqref{statmatch}, giving the functional dependence
+Thus, while we will treat $\thup$ as a parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by the matching condition \eqref{statmatch}, giving the functional dependence
 \begin{equation}
@@
 An MCMC method, on the other hand, would require the sampling to restart from scratch for each value of $\thdn$, since the random steps in MCMC depend directly on the target distribution \eqref{Gibbs}.
-We will however, use an MCMC method to initialize direct numerical simulations of TKdV due to the superior efficiency for a specific, given value of $\theta$.
+We will however, use MCMC to initialize direct numerical simulations of TKdV due to the superior efficiency for a single, given value of $\theta$.
 
@@
 \end{equation}
-The quadratic nonlinearity present in this system, $\usqdah_k$, mixes the modes during evolution. We note the third-order linear term may become stiff for large $\Lambda$.
+The quadratic nonlinearity represented by $\usqdah_k$ mixes the modes during evolution. We note the third-order linear term may become stiff for large $\Lambda$.
 
@@
 
-	For the initial conditions, $\uhvec^0$, of the direct numerical simulation we sample from the upstream Gibbs ensemble $\Gup$ in \eqref{Gibbs} with a prescribed inverse temperature $\thup$.
+	For the initial conditions, $\uhvec^0$, of the direct numerical simulation we sample from the upstream Gibbs ensemble $\Gup$ \eqref{Gibbs} with a prescribed inverse temperature $\thup$.
 We achieve this sampling via a Metropolis-Hasting Monte-Carlo algorithm as detailed in \cite{majda2019}.

### 2019-11-20 paper2_JNLS/paper.tex
NM

@@
 \begin{abstract} 
-Recent laboratory experiments revealed the emergence of anomalous statistics in a randomized surface waves upon encountering an abrupt depth change (ADC) in the bottom topography (Bolles et al.
-PRF 2019).
-Majda et al.~2019 established a theoretical foundation, based on analysis of the truncated KdV (TKdV) system, that successfully captures several key qualitative features of the experiments including the robust emergence of anomalous statistics.
-The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the ADC, and is corroborated by simulations of the deterministic dynamics.
-%
+Recent laboratory experiments of \boetal revealed the emergence of anomalous statistics in a randomized wave-field that encounters an abrupt change in bottom topography.
+\maetal established a theoretical foundation, based on analysis of the truncated Korteweg–de Vries (TKdV) system, that successfully captures several key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing waves.
+The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the depth change, and is corroborated by numerical simulations of the deterministic dynamics.
 Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two.
 We present an assortment of new experimental measurements and carefully compare them to model predictions.
-We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments.
-The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in a spectacular fashion by direct experimental measurements.
+We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments and a quantitative comparison can be made.
+The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in a spectacular fashion by new experimental measurements.
 \end{abstract}
@@
 In particular, the more complex mechanism of focusing by 2D bottom topography is  absent.
-The studies thus offer a paradigm system, with emergent anomalous statistical properties similar to that seen in more complex systems, but in a tractable context that is amenable to analysis.
+The studies thus offer a paradigm system, with emergent anomalous features similar to those seen in more complex systems, but in a tractable context that is amenable to analysis.
 

### 2019-11-20 paper2_JNLS/paper.tex
NM

@@
 
-	As diagrammed in Fig.~\ref{ExpDiagStats}(a), the experiments consist of a long, narrow wave tank (6 m long x 20 cm wide x 30 cm high), with waves generated by plexiglass paddle at one end \cite{bolles2019}.
+	As diagrammed in Fig.~\ref{ExpDiagStats}(a), the experiments consist of a long, narrow wave tank (6 m long x 20 cm wide x 30 cm high), with waves generated by a plexiglass paddle at one end \cite{bolles2019}.
 The waves propagate through the tank and, roughly midway through, pass over an abrupt depth change (ADC) created by a plexiglass step.
@@
 As seen in \ref{ExpSpatialStats}(b)--(c), both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm.
-Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude.
+Interestingly, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude.
 The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e).
-\boetal found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution.
+\boetal found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution across all of the experiments.
 We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.
@@
 This framework is based on a Galerkin truncation of the variable-depth Korteweg–de Vries (KdV) equation.
-The KdV equation is a well established model for describing the propagation of unidirectional, shallow-water waves, and it accounts for weak nonlinearity and weak dispersion over long timescales and large spatial scales.
+The KdV equation is a well established model for describing the propagation of unidirectional, shallow-water waves, accounting for weak nonlinearity and weak dispersion over long timescales and large spatial scales.
 We will perform Galerkin truncation of KdV to obtain a finite-dimensional dynamical system that exhibits weak turbulence.
-We outline the Hamiltonian structure of both the continuous KdV and the truncated system.
+We outline the Hamiltonian structure of both the traditional KdV and the truncated systems.
 This structure is exploited to obtain invariant measures of the underlying dynamics and, ultimately, to rationalize the experimental findings on anomalous wave statistics triggered by an ADC.
@@
 \end{align}
-Most often, we will consider waves moving into shallower depth, so that $\dup > \ddn$.
-Throughout this paper, we will use the subscript `-' to represent upstream variables and `+' for downstream variables.
+Most often, we consider waves moving into shallower depth, so that $\dup > \ddn$. Throughout this paper, we use the subscript `-' to represent upstream variables and `+' for downstream variables.
 
@@
 
-In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from \maetal \cite{majda2019}.
+In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework of \maetal \cite{majda2019}.
 Since the choice of normalization is not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices.
@@
 The KdV equation provides the first correction to this behavior and describes dynamics that evolve over longer timescales.
-Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitably long timescale in line with previous normalizations \cite{johnson1997modern}.
+Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitably long timescale in line with other normalizations \cite{johnson1997modern}.
 The scales $\lengthscale = \lengthscale_{\pm}$ and $\timescale = \timescale_{\pm}$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.
@@
 As seen in \eqref{H3H2}, the cubic component, $\Hthree$, represents the skewness of the wave-field, while the quadratic component, $\Htwo$, represents the energy of the surface slope.
-The sign difference between the two \eqref{Hamiltonian} thus represents a competition between wave skewness and slope energy.
-In particular, the role played by the slope energy is the motivating factor for the new experimental measurements on surface {\em slope} statistics that are reported in this paper.
+The sign difference between the two in \eqref{Hamiltonian} thus represents a competition between wave skewness and slope energy.
+In particular, the appearance of the slope energy in the theory motivated the new experimental measurements on surface {\em slope} statistics reported in this paper.
 
@@
 
+%----------------------------------------------------------------------------%
+% Transfer
+%----------------------------------------------------------------------------%
+\subsection{Numerical computation of the transfer function}
+
 	To compute the transfer function, we use a weighted random-sampling strategy.
 That is, we first sample the Fourier coefficients $\uhvec = \left\{ \uhat_{k}\right\}_{k=0}^{\Lambda}$ from a uniform distribution on the unit hypersphere $\En = 1$.
-The uniform sampling is achieved by first sampling from a isotropic Gaussian distribution and then normalizing to project onto the hypersphere \cite{abramov2003}.
+The uniform sampling is achieved by first sampling from an isotropic Gaussian distribution and then normalizing to project onto the hypersphere \cite{abramov2003}.
 These samples are then weighted by the appropriate Gibbs measure \eqref{Gibbs} to compute the expectations needed in the statistical matching condition \eqref{statmatch}.
@@
 	We note that the experiments discussed here preceded the development of the theory, and hence had no intent of mimicking a Gibbs measure in the wave forcing.
-We expect that if the experimental forcing were designed to mimic the Gibbs measure, in particular the spectral decay, then it would be more straightforward to assign a value to the model parameter $\lamfac$.
-We, however, leave that task for future research due to the significant cost of performing an entirely new set of experiments compared to the relative ease and great value in re-analyzing existing experimental measurements in light of the new theoretical developments.
+We expect that if the experimental forcing were designed to mimic the Gibbs measure, in particular the precise spectral decay, then it would be more straightforward to assign a value to the model parameter $\lamfac$.
+We, however, leave that task for future research due to the significant cost of performing an entirely new set of experiments compared to the relative ease and great value in re-analyzing existing experimental data in light of the new theoretical developments.
 
@@
 
-At this point, all parameters that appear in the theoretical model have been linked directly to experimental parameters with the exception of the inverse temperature of the incoming flow, $\thup$.
+At this point, all parameters appearing in the theoretical model have been linked directly to experimental parameters with the exception of the inverse temperature of the incoming flow, $\thup$.
 Our strategy is to use the outgoing skewness as the main diagnostic to determine a realistic range for $\thup$.
@@
 We thus run the deterministic TKdV simulations with these model parameters and with $\epsup$, $\delup$, and $\dratdn$ set to the representative values in Table \ref{paramtable}.
-We run the simulations to a sufficiently long dimensionless time of $t_{f} = 10$, with $dt = 5 \times 10^{-4}$ and with $10^3$ trajectories sampled from the upstream Gibbs measure $\Gup$.
+We run the simulations to a sufficiently long dimensionless time of $t_{f} = 10$, with $dt = 5 \times 10^{-4}$, and with $10^3$ trajectories sampled from the upstream Gibbs measure $\Gup$.
 
@@
 In the recent theoretical study, \maetal derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}.
-More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$.
+More specifically, this formula relates the downstream skewness of surface displacement, $\skw(\eta)$, to the change in slope variance, $\var(\eta_x)$.
 Here, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.
@@
 In particular, it predicts their ratio to scale as the {\em inverse cube of wave amplitude}, $\epsup^{-3}$, and the {\em cube of the depth ratio}, $\dratdn^3$.
-It is a rather unexpected relationship, as it would have been difficult to anticipate that the displacement skewness and slope variance, among all possibly variable combinations, are so intimately related.
+It is a rather unexpected relationship, as it would have been difficult to anticipate that the displacement skewness and slope variance, among all possible variable combinations, are so intimately related.
 
@@
 (a)--(b) Spatial variation of $\skw(\eta)$ and $\var(\eta_x)$ for several different amplitudes.
-The maximum of $\skw(\eta)$ and the minimum of $\var(\eta_x)$ each occur at a location that is insensitive to amplitude, $x = $15 cm (blue) and -9 cm (green) respectively, which are chosen as the representative locations to evaluate \eqref{SkewPred}.
+The maximum of $\skw(\eta)$ and the minimum of $\var(\eta_x)$ each occur at a location that is insensitive to amplitude, $x = $15 cm (blue) and -9 cm (green) respectively.
 (c) Measurements of the ratio $\skw(\eta) / \Delta \var(\eta_x)$ from 15 different experiments plotted against the dimensionless wave amplitude $\epsup$.
-For sufficiently large amplitude, the measurements fit closely with the $\epsup^{-3}$ prediction from \eqref{SkewPred}.
+For sufficiently large amplitude, the measurements follow the predicted $\epsup^{-3}$ power law closely.
 }

### 2019-11-29 paper2_JNLS/paper.tex
NM

@@
 \begin{abstract} 
-Recent laboratory experiments of \boetal revealed the emergence of anomalous statistics in a randomized wave-field that encounters an abrupt change in bottom topography.
-\maetal established a theoretical foundation, based on analysis of the truncated Korteweg–de Vries (TKdV) system, that successfully captures several key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing waves.
-The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the depth change, and is corroborated by numerical simulations of the deterministic dynamics.
-Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two.
-We present an assortment of new experimental measurements and carefully compare them to model predictions.
-We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments and a quantitative comparison can be made.
-The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in a spectacular fashion by new experimental measurements.
+\boetal experimentally demonstrated that an abrupt change in bottom topography can trigger anomalous statistics in a randomized field of surface waves.
+Motivated by these observations, \maetal developed a theoretical framework, based on deterministic and statistical analysis of the truncated KdV system, that successfully captures key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing waves.
+Here, we extend these parallel experimental and modeling efforts with several new findings that have resulted from synergetic interaction between the two.
+By fleshing out details of the non-dimensionalization, we precisely relate model parameters to physical ones, thereby enabling calibration of the model inverse temperature to the specific conditions present in the experiments and, ultimately, a quantitative comparison between theory and experiments.
+We find the theoretically predicted distributions of surface displacement to match the experimental measurements with surprising quantitive detail.
+Prompted by the presence of surface slope in the TKdV Hamiltonian, we experimentally examine surface slope statistics and compare them to theoretical predictions.
+Both show significant spread of the slope distributions with long, flat tails in the outgoing wave field.
+Finally, the theory predicts a peculiar relationship between the outgoing displacement skewness and the change in slope variance, specifically how their ratio depends on the wave amplitude and depth ratio.
+New experimental measurements confirm this prediction in spectacular fashion.
 \end{abstract}

### 2019-11-29 paper2_JNLS/paper.tex
NM

@@
 \boetal experimentally demonstrated that an abrupt change in bottom topography can trigger anomalous statistics in a randomized field of surface waves.
-Motivated by these observations, \maetal developed a theoretical framework, based on deterministic and statistical analysis of the truncated KdV system, that successfully captures key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing waves.
+Motivated by these observations, \maetal developed a theoretical framework, based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) system, that successfully captures key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing waves.
 Here, we extend these parallel experimental and modeling efforts with several new findings that have resulted from synergetic interaction between the two.
@@
 Both show significant spread of the slope distributions with long, flat tails in the outgoing wave field.
+Analysis of some deterministic trajectories of TKdV sheds light on the experimental length and time scales required for the statistical transition.
 Finally, the theory predicts a peculiar relationship between the outgoing displacement skewness and the change in slope variance, specifically how their ratio depends on the wave amplitude and depth ratio.

### 2019-12-01 paper2_JNLS/paper.tex
NM

--- 2019-11-29 paper2_JNLS/paper.tex
@@
 \begin{abstract} 
-\boetal experimentally demonstrated that an abrupt change in bottom topography can trigger anomalous statistics in a randomized field of surface waves.
-Motivated by these observations, \maetal developed a theoretical framework, based on deterministic and statistical analysis of the truncated KdV system, that successfully captures key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing waves.
+\boetal experimentally demonstrated that an abrupt change in bottom topography can trigger anomalous statistics in randomized surface waves.
+Motivated by these observations, \maetal developed a theoretical framework, based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) system, that successfully captures key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing wavefield.
 Here, we extend these parallel experimental and modeling efforts with several new findings that have resulted from synergetic interaction between the two.
@@
 Both show significant spread of the slope distributions with long, flat tails in the outgoing wave field.
+Analysis of some deterministic trajectories of TKdV elucidates the experimental length and time scales required for the statistical transition.
 Finally, the theory predicts a peculiar relationship between the outgoing displacement skewness and the change in slope variance, specifically how their ratio depends on the wave amplitude and depth ratio.

### 2019-12-01 paper2_JNLS/paper.tex
NM

@@
 Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
-Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing \citep{kharif2008influence, toffoli2017wind}, opposing currents \citep{garrett2009rogue, onorato2011triggering},
-focusing due to variable bathymetry \citep{heller2008refraction, white1998chance}, 
+Rogue waves, also variously known as freak or anomalous waves, have been observed in shallow \cite{pelinovsky2000nonlinear, gramstad2013freak}, intermediate \cite{karmpadakis2019laboratory}, and deep water \cite{dematteis2018rogue, dematteis2019experimental}, and can be triggered by a variety of mechanisms, including anomalous wind-forcing \citep{kharif2008influence, toffoli2017wind}, opposing currents \citep{garrett2009rogue, onorato2011triggering}, focusing due to variable bathymetry \citep{heller2008refraction, white1998chance},
 and the Benjamin-Feir deep-water modulational instability \citep{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.

### 2019-12-02 paper2_JNLS/paper.tex
NM

@@
 \begin{abstract} 
-\boetal experimentally demonstrated that an abrupt change in bottom topography can trigger anomalous statistics in randomized surface waves.
-Motivated by these observations, \maetal developed a theoretical framework, based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) system, that successfully captures key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing wavefield.
-Here, we extend these parallel experimental and modeling efforts with several new findings that have resulted from synergetic interaction between the two.
-By fleshing out details of the non-dimensionalization, we precisely relate model parameters to physical ones, thereby enabling calibration of the model inverse temperature to the specific conditions present in the experiments and, ultimately, a quantitative comparison between theory and experiments.
-We find the theoretically predicted distributions of surface displacement to match the experimental measurements with surprising quantitive detail.
-Prompted by the presence of surface slope in the TKdV Hamiltonian, we experimentally examine surface slope statistics and compare them to theoretical predictions.
-Both show significant spread of the slope distributions with long, flat tails in the outgoing wave field.
-Analysis of some deterministic trajectories of TKdV elucidates the experimental length and time scales required for the statistical transition.
+Recent laboratory experiments of \boetal demonstrate that an abrupt change in bottom topography can trigger anomalous statistics in randomized surface waves.
+Motivated by these observations, \maetal developed a theoretical framework, based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) system, that successfully captures key qualitative features of the experiments, including the robust emergence of anomalous statistics and heightened skewness in the outgoing wavefield.
+Here, we extend these parallel experimental and modeling efforts with several new findings that have resulted from a synergetic interaction between the two.
+By precisely relating model parameters to physical ones, we calibrate the model inverse temperature to the specific conditions present in the experiments, thereby permitting a quantitative comparison.
+We find theoretically predicted distributions of surface displacement to match the experimental measurements with surprising quantitive detail.
+Prompted by the presence of surface slope in the TKdV Hamiltonian, we present new experimental measurements on surface slope statistics and compare them to model predictions.
+Analysis of some deterministic trajectories of TKdV elucidates the experimental length and time scales required for the statistical transition to a skewed state.
 Finally, the theory predicts a peculiar relationship between the outgoing displacement skewness and the change in slope variance, specifically how their ratio depends on the wave amplitude and depth ratio.
 New experimental measurements confirm this prediction in spectacular fashion.
+%New experimental measurements provide convincing evidence for this prediction.
 \end{abstract}
@@
 MNJM acknowledges support from the Simons Foundation, award 524259. 
+This research of A.J.M. is partially supported by the Office of Naval Research N00014-19-S-B001. 
+D.Q. is supported as a postdoctoral fellow on the grant.
 

### 2019-12-02 paper2_JNLS/paper.tex
NM

@@
 By precisely relating model parameters to physical ones, we calibrate the model inverse temperature to the specific conditions present in the experiments, thereby permitting a quantitative comparison.
-We find theoretically predicted distributions of surface displacement to match the experimental measurements with surprising quantitive detail.
+We find theoretically predicted distributions of surface displacement to match the experimental measurements with surprising detail.
 Prompted by the presence of surface slope in the TKdV Hamiltonian, we present new experimental measurements on surface slope statistics and compare them to model predictions.
@@
 More recently, Qi \& Majda (2019) \cite{MachineLearning2019} demonstrated the capability of machine learning strategies to accurately predict these extreme events.
-That work uses a deep neural network coupled with a judicious choice of the entropy loss function that emphasizes the dominant structures of the turbulent field \cite{MachineLearning2019}.
+That work employs a deep neural network coupled with a judicious choice of the entropy loss function that emphasizes the dominant structures of the turbulent field \cite{MachineLearning2019}.
 It should be noted that, in these studies, extreme events are represented by strong skewness of the wave-field.
@@
 The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of \boetal~\cite{bolles2019} and \maetal~\cite{majda2019, majdaqi2019}.
-Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
+We present a variety of new findings that have resulted from a synergetic interaction between theory and experiments -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
+That outline of the paper is as follows.
 In Section \ref{experiments} we detail the laboratory experiments of \boetal \cite{bolles2019} and summarize previously reported findings on anomalous statistics of the surface displacement.
@@
 In this section, we flesh out details of the non-dimensionalization, which, for brevity, were only briefly discussed in previous work.
-This exercise provides a more thorough link between model and experimental parameters, ultimately facilitating a closer comparison between the two.
+This exercise provides a more precise link between model and experimental parameters, ultimately facilitating a closer comparison between the two.
 We then discuss TKdV as a deterministic dynamical system, approachable from the viewpoint of statistical mechanics.
 The later perspective is based on the Hamiltonian structure and novel ensemble distributions that incorporate canonical and micro-canonical aspects \cite{abramov2003}.
+
 Section \ref{results} presents a systematic comparison between experiments and theory, focusing on both surface displacement and surface slope.
 We find the statistical distributions of these two quantities to agree well across experiments and theory.
-In particular, we find the agreement in displacement statistics to be remarkably quantitative.
+In particular, we find remarkably quantitative agreement in the displacement statistics.
 We then examine numerical simulations of the TKdV deterministic dynamics, which elucidates the length and timescales required for statistical transitions in the experiments.

### 2020-01-03 paper2_JNLS/paper.tex
NM

@@
 Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
-Rogue waves, also variously known as freak or anomalous waves, have been observed in shallow \cite{pelinovsky2000nonlinear, gramstad2013freak}, intermediate \cite{karmpadakis2019laboratory}, and deep water \cite{dematteis2018rogue, dematteis2019experimental}, and can be triggered by a variety of mechanisms, including anomalous wind-forcing \citep{kharif2008influence, toffoli2017wind}, opposing currents \citep{garrett2009rogue, onorato2011triggering}, focusing due to variable bathymetry \citep{heller2008refraction, white1998chance},
+Rogue waves, also variously known as freak or anomalous waves, have been observed in shallow \cite{pelinovsky2000nonlinear, gramstad2013freak}, intermediate \cite{karmpadakis2019laboratory}, and deep water \cite{dematteis2018rogue, dematteis2019experimental}, and certain rogue features have even been recovered by exact solutions to various wave models \cite{peregrine1983water, clarkson2017rational, chen2019periodic}.
+These abnormal waves can be triggered by a variety of mechanisms, including anomalous wind-forcing \citep{kharif2008influence, toffoli2017wind}, opposing currents \citep{garrett2009rogue, onorato2011triggering}, focusing due to variable bathymetry \citep{heller2008refraction, white1998chance},
 and the Benjamin-Feir deep-water modulational instability \citep{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
 The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement: when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these events occur much more frequently when surface statistics deviate from Gaussian.
-In this way, anomalous waves can be approached from the broader perspective of turbulent dynamical systems \cite{sapsis2013a, sapsis2013b, sapsis2013blending, chen2016filtering, majda2016introduction, macedo2017universality, MajdaQiSIAM2018, blonigan2019extreme, guth2019machine}.
-
-A recent series of laboratory \cite{bolles2019} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
+In this way, anomalous waves can be approached from the broader perspective of turbulent dynamical systems \cite{sapsis2013a, sapsis2013b, sapsis2013blending, chen2016filtering, majda2016introduction, macedo2017universality, MajdaQiSIAM2018, blonigan2019extreme, guth2019machine, holm2019stochastic}.
+
+A recent series of laboratory \cite{bolles2019, trulsen2020extreme} and numerical investigations \cite{viotti2014, herterich2019extreme} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography.
 Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous waves.

### 2020-01-03 paper2_JNLS/paper.tex
NM

@@
 
-%experimental measurements and deterministic-statistical truncated KdV theory
-% ORIGINAL: The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes
-% Words: triggered, induced, generated, sparked
-
 \author{
@@
 Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
-Di Qi\thanks{Courant Institute of Mathematical Sciences} }
+Di Q\footnotemark[3] }
 \maketitle
@@
 New experimental measurements confirm this prediction in spectacular fashion.
+\end{abstract}
 %New experimental measurements provide convincing evidence for this prediction.
-\end{abstract}
 

### 2020-01-03 paper2_JNLS/paper.tex
NNM

@@
 \usepackage{comment, enumerate, tabularx, todonotes}
-\usepackage{natbib, hyperref, url}
+\usepackage{hyperref, url}
 \usepackage[margin=1in]{geometry}
@@
 Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
-Di Q\footnotemark[3] }
+Di Qi\footnotemark[3] }
 \maketitle
@@
 
-Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height \citep{muller2005rogue, ying2011linear}.
+Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height \cite{muller2005rogue, ying2011linear}.
 Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
 Rogue waves, also variously known as freak or anomalous waves, have been observed in shallow \cite{pelinovsky2000nonlinear, gramstad2013freak}, intermediate \cite{karmpadakis2019laboratory}, and deep water \cite{dematteis2018rogue, dematteis2019experimental}, and certain rogue features have even been recovered by exact solutions to various wave models \cite{peregrine1983water, clarkson2017rational, chen2019periodic}.
-These abnormal waves can be triggered by a variety of mechanisms, including anomalous wind-forcing \citep{kharif2008influence, toffoli2017wind}, opposing currents \citep{garrett2009rogue, onorato2011triggering}, focusing due to variable bathymetry \citep{heller2008refraction, white1998chance},
-and the Benjamin-Feir deep-water modulational instability \citep{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
+These abnormal waves can be triggered by a variety of mechanisms, including anomalous wind-forcing \cite{kharif2008influence, toffoli2017wind}, opposing currents \cite{garrett2009rogue, onorato2011triggering}, focusing due to variable bathymetry \cite{heller2008refraction, white1998chance},
+and the Benjamin-Feir deep-water modulational instability \cite{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
 The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement: when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these events occur much more frequently when surface statistics deviate from Gaussian.

### 2020-01-10 paper2_JNLS/paper.tex
NM

@@
 We present a variety of new findings that have resulted from a synergetic interaction between theory and experiments -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}.
-That outline of the paper is as follows.
+The outline of the paper is as follows.
 In Section \ref{experiments} we detail the laboratory experiments of \boetal \cite{bolles2019} and summarize previously reported findings on anomalous statistics of the surface displacement.
@@
 We employ a pseudo-spectral discretization of \eqref{TKdV2}, with de-aliasing applied to the quadratic nonlinear term $\Proj (\uL)^2$ according to the standard 2/3-rule.
-That is, we first pad $\uhat_k$ with zeros for $\Lambda < \abs{k} \le 3\Lambda/2$, transform to physical space and square to obtain $\uL^2$ on a fine grid, then transform back to spectral space and truncate to obtain the Fourier coefficients of $\Proj (\uL)^2$
+That is, we first pad $\uhat_k$ with zeros for $\Lambda < \abs{k} \le 3\Lambda/2$, transform to physical space and square to obtain $\uL^2$ on a fine grid, then transform back to spectral space and truncate to obtain the Fourier coefficients of $\Proj (\uL)^2$.
 For simplicity, we denote these Fourier coefficients $\usqdah_k$,

### 2020-04-15 paper2_JNLS/paper.tex
NM

@@
 We consider the surface displacement $\eta(x,t)$ of unidirectional, shallow-water waves in a reference frame moving with the characteristic wave speed, $\xi = x - ct$.
-Here, $c = \sqrt{g \depth}$ is leading-order approximation to the wave speed (i.e.~from linear theory), where $g$ is gravity and $\depth$ is the local depth.
+Here, $c = \sqrt{g \depth}$ is {the} leading-order approximation to the wave speed (i.e.~from linear theory), where $g$ is gravity and $\depth$ is the local depth.
 The leading-order dynamics are corrected to first order in small amplitude by the Korteweg–de Vries equation (KdV), which in dimensional form is given by \cite{whitham2011linear}
@@
 
-	Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$ for the case $(\Lambda, \lamfac) = (16, 8)$.
+	Next, Fig.~\ref{transfig}(c) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$ for the case $(\Lambda, \lamfac) = (16, 8)$.
 In line with experimental observations, the incoming skewness is small, while the outgoing wave skewness is much higher.

### 2020-04-20 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\Gupdn}{\Gibbs^{\pm}}
-\newcommand{\thup}{\theta^{-}}
-\newcommand{\thdn}{\theta^{+}}
-\newcommand{\thupdn}{\theta^{\pm}}
+\newcommand{\invtemp}{\beta}
+\newcommand{\itup}{\invtemp^{-}}
+\newcommand{\itdn}{\invtemp^{+}}
+\newcommand{\itupdn}{\invtemp^{\pm}}
 \newcommand{\meanup}[1]{\mean{#1}_{-}}
@@
 \label{Gibbs}
-\Gupdn = Z_{\thupdn}^{-1} \, \exp(-\thupdn \Hupdn) \delta(\En - 1)
-\end{align}
-Here $\theta = \thupdn$ is the inverse temperature, which will take a different value on either side of the ADC, and $Z_{\theta}$ a constant that depends on $\theta$.
+\Gupdn = Z_{\itupdn}^{-1} \, \exp(-\itupdn \Hupdn) \delta(\En - 1)
+\end{align}
+Here $\invtemp = \itupdn$ is the inverse temperature, which will take a different value on either side of the ADC, and $Z_{\invtemp}$ a constant that depends on $\invtemp$.
 Each measure $\Gupdn$ induces a corresponding ensemble average, denoted $\mean{\cdot}_{\pm}$.
 
-	We note that, in the current formulation, a positive inverse temperature $\theta > 0$ produces physically realistic statistics with a decaying energy spectrum, as is consistent with experiments \cite{majda2019, bajars2013weakly}.
+	We note that, in the current formulation, a positive inverse temperature $\invtemp > 0$ produces physically realistic statistics with a decaying energy spectrum, as is consistent with experiments \cite{majda2019, bajars2013weakly}.
 Negative inverse temperature produces a physically unrealistic spectrum that has more energy at smaller scales.
-Hence, we will hereafter focus on the physically realistic case of $\theta > 0$.
+Hence, we will hereafter focus on the physically realistic case of $\invtemp > 0$.
 
@@
 \end{align}
-Note that if $\theta = 0$, then the Gibbs measure is a {\em uniform} measure on the unit hypersphere in $\CC^{\Lambda}$, which we will denote as either $\Gz$ or simply $d\mu$.
+Note that if $\invtemp = 0$, then the Gibbs measure is a {\em uniform} measure on the unit hypersphere in $\CC^{\Lambda}$, which we will denote as either $\Gz$ or simply $d\mu$.
 \end{comment}
@@
 
-	This statistical matching condition, originally derived by \maetal \cite{majda2019}, imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$.
-In particular, we view $\thup$ as given by the random-state of the incoming wave field.
-Thus, while we will treat $\thup$ as a parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by the matching condition \eqref{statmatch}, giving the functional dependence
-\begin{equation}
-\thdn = \transf \left( \thup \right)
+	This statistical matching condition, originally derived by \maetal \cite{majda2019}, imposes a relationship between the two inverse temperatures $\itup$ and $\itdn$.
+In particular, we view $\itup$ as given by the random-state of the incoming wave field.
+Thus, while we will treat $\itup$ as a parameter that can be varied to study various possible system states, the downstream $\itdn$ is determined directly by the matching condition \eqref{statmatch}, giving the functional dependence
+\begin{equation}
+\itdn = \transf \left( \itup \right)
 \qquad \mbox{\em Transfer function}
@@
 These samples are then weighted by the appropriate Gibbs measure \eqref{Gibbs} to compute the expectations needed in the statistical matching condition \eqref{statmatch}.
-That is, for an arbitrary quantity $Q$, the ensemble expectation corresponding to Hamiltonian $\Ham$ and inverse temperature $\theta$ is numerically approximated as
+That is, for an arbitrary quantity $Q$, the ensemble expectation corresponding to Hamiltonian $\Ham$ and inverse temperature $\invtemp$ is numerically approximated as
 \begin{equation}
 \label{weightmean}
-\mean{Q}_{\theta} = \frac{\sumsamp Q_i \exp(-\theta \Ham_i)} {\sumsamp \exp(-\theta \Ham_i)}
+\mean{Q}_{\invtemp} = \frac{\sumsamp Q_i \exp(-\invtemp \Ham_i)} {\sumsamp \exp(-\invtemp \Ham_i)}
 \end{equation}
@@
 \label{rooteq}
-\Fth(\thdn) =  \mean{\Hdn}_{\thdn} - \mean{\Hdn}_{\thup} = 0
-\end{equation}
-Recall that we consider $\thup$ as given, so that $\mean{\Hdn}_{\thup}$ is a constant that can be computed straightaway from \eqref{weightmean}.
-We then use the secant method with respect to the variable $\thdn$ to find a root of $\Fth(\thdn)$ to the desired tolerance.
-
-For simply computing the transfer function, this weighted-sampling approach offers significant advantages over more sophisticated methods, such as such Markov Chain Monte Carlo (MCMC), in its ability to reuse the same samples of $\uhvec$ for several different values of $\thdn$.
-That is, we sample $\uhvec$ from the uniform distribution and compute the list of Hamiltonian values only once, then simply vary $\thdn$ in \eqref{rooteq} with the secant method until the root is found.
-An MCMC method, on the other hand, would require the sampling to restart from scratch for each value of $\thdn$, since the random steps in MCMC depend directly on the target distribution \eqref{Gibbs}.
-We will however, use MCMC to initialize direct numerical simulations of TKdV due to the superior efficiency for a single, given value of $\theta$.
+\Fth(\itdn) =  \mean{\Hdn}_{\itdn} - \mean{\Hdn}_{\itup} = 0
+\end{equation}
+Recall that we consider $\itup$ as given, so that $\mean{\Hdn}_{\itup}$ is a constant that can be computed straightaway from \eqref{weightmean}.
+We then use the secant method with respect to the variable $\itdn$ to find a root of $\Fth(\itdn)$ to the desired tolerance.
+
+For simply computing the transfer function, this weighted-sampling approach offers significant advantages over more sophisticated methods, such as such Markov Chain Monte Carlo (MCMC), in its ability to reuse the same samples of $\uhvec$ for several different values of $\itdn$.
+That is, we sample $\uhvec$ from the uniform distribution and compute the list of Hamiltonian values only once, then simply vary $\itdn$ in \eqref{rooteq} with the secant method until the root is found.
+An MCMC method, on the other hand, would require the sampling to restart from scratch for each value of $\itdn$, since the random steps in MCMC depend directly on the target distribution \eqref{Gibbs}.
+We will however, use MCMC to initialize direct numerical simulations of TKdV due to the superior efficiency for a single, given value of $\invtemp$.
 
@@
 
-	For the initial conditions, $\uhvec^0$, of the direct numerical simulation we sample from the upstream Gibbs ensemble $\Gup$ \eqref{Gibbs} with a prescribed inverse temperature $\thup$.
+	For the initial conditions, $\uhvec^0$, of the direct numerical simulation we sample from the upstream Gibbs ensemble $\Gup$ \eqref{Gibbs} with a prescribed inverse temperature $\itup$.
 We achieve this sampling via a Metropolis-Hasting Monte-Carlo algorithm as detailed in \cite{majda2019}.
@@
 
-	Since the inverse temperature, $\theta$, is a key modeling parameter, it would be highly desirable to normalize the system such that the value of $\theta$ does not depend sensitively on the truncation index as $\Lambda$ grows large.
-That way, $\theta$ can be interpreted as a real physical parameter that can be linked to the experiments and whose value does not depend sensitively on where one chooses to truncate the system.
+	Since the inverse temperature, $\invtemp$, is a key modeling parameter, it would be highly desirable to normalize the system such that the value of $\invtemp$ does not depend sensitively on the truncation index as $\Lambda$ grows large.
+That way, $\invtemp$ can be interpreted as a real physical parameter that can be linked to the experiments and whose value does not depend sensitively on where one chooses to truncate the system.
 The one modeling parameter that is left to be set is $\lamfac$, which represents the number of characteristic wavelengths in the periodic domain.
-In what follows, we will determine reasonable constraints on $\lamfac$ that allow $\theta$ to be asymptotically independent of $\Lambda$.
-
-	The invariant measure exhibits the proportionality $ \Gibbs_{\Lambda} \propto \exp(-\theta_{\Lambda} \Ham_{\Lambda})$, where we have made explicit the dependence of all quantities on $\Lambda$.
-In particular, if $\Ham_{\Lambda}$ were to depend sensitively on $\Lambda$ in expectation, then $\theta_{\Lambda}$ would need to compensate in order to produce the same invariant measure.
+In what follows, we will determine reasonable constraints on $\lamfac$ that allow $\invtemp$ to be asymptotically independent of $\Lambda$.
+
+	The invariant measure exhibits the proportionality $ \Gibbs_{\Lambda} \propto \exp(-\invtemp_{\Lambda} \Ham_{\Lambda})$, where we have made explicit the dependence of all quantities on $\Lambda$.
+In particular, if $\Ham_{\Lambda}$ were to depend sensitively on $\Lambda$ in expectation, then $\invtemp_{\Lambda}$ would need to compensate in order to produce the same invariant measure.
 Hence, it would be desirable to scale the system in such a way that $\Ham_{\Lambda}$ does not depend sensitively on $\Lambda$, at least in expectation.
@@
 
-At this point, all parameters appearing in the theoretical model have been linked directly to experimental parameters with the exception of the inverse temperature of the incoming flow, $\thup$.
-Our strategy is to use the outgoing skewness as the main diagnostic to determine a realistic range for $\thup$.
-That is, for an input $\thup$, the downstream inverse temperature $\thdn$ is determined by \eqref{statmatch}, which ultimately sets the skewness of the outgoing wave-field.
+At this point, all parameters appearing in the theoretical model have been linked directly to experimental parameters with the exception of the inverse temperature of the incoming flow, $\itup$.
+Our strategy is to use the outgoing skewness as the main diagnostic to determine a realistic range for $\itup$.
+That is, for an input $\itup$, the downstream inverse temperature $\itdn$ is determined by \eqref{statmatch}, which ultimately sets the skewness of the outgoing wave-field.
 
@@
 \caption{
-Analysis of the transfer function, $\thdn = \transf \left( \thup \right)$, and calibration of the inverse temperature.
+Analysis of the transfer function, $\itdn = \transf \left( \itup \right)$, and calibration of the inverse temperature.
 (a)--(b) The transfer function for three values of $\Lambda$ with either (a) $\lamfac = 8$ fixed or (b) $\lamfac = \Lambda/2$ scaled.
 Scaling $\lamfac$ with $\Lambda$ mitigates the sensitivity to $\Lambda$.
-(c) The incoming and outgoing skewness versus $\thup$ implied by the matching condition.
+(c) The incoming and outgoing skewness versus $\itup$ implied by the matching condition.
 In line with experiments, the skewness is enhanced significantly in the outgoing wave field.
@@
 
-	Figures \ref{transfig}(a)-(b) show the transfer function $\thdn = \transf(\thup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $0 \le \thup \le 25$.
+	Figures \ref{transfig}(a)-(b) show the transfer function $\itdn = \transf(\itup)$ that results from the statistical matching condition \eqref{statmatch}, for an incoming inverse temperature in the range $0 \le \itup \le 25$.
 Figure \ref{transfig}(a) shows the cases $\Lambda = $ 12, 16, and 20 with $\lamfac = 8$ fixed in each.
@@
 
-	Next, Fig.~\ref{transfig}(c) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$ for the case $(\Lambda, \lamfac) = (16, 8)$.
+	Next, Fig.~\ref{transfig}(c) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\itup$ for the case $(\Lambda, \lamfac) = (16, 8)$.
 In line with experimental observations, the incoming skewness is small, while the outgoing wave skewness is much higher.
-Specifically, to capture the experimentally observed peak skewness range of 0.6-0.9 seen for the larger amplitudes in Fig.~\ref{ExpSpatialStats}(b), requires selecting $\thup$ in the range 10--25.
+Specifically, to capture the experimentally observed peak skewness range of 0.6-0.9 seen for the larger amplitudes in Fig.~\ref{ExpSpatialStats}(b), requires selecting $\itup$ in the range 10--25.
 
@@
 Throughout, we will focus on the representative set of experiments detailed in Table \ref{paramtable}, in which the peak downstream skewness was measured to be 0.83.
-With $(\Lambda, \lamfac) = (16, 8)$ fixed, Fig.~\ref{transfig}(c) indicates $\thup = 20$ as a reasonable value to attain the desired skewness.
+With $(\Lambda, \lamfac) = (16, 8)$ fixed, Fig.~\ref{transfig}(c) indicates $\itup = 20$ as a reasonable value to attain the desired skewness.
 We thus run the deterministic TKdV simulations with these model parameters and with $\epsup$, $\delup$, and $\dratdn$ set to the representative values in Table \ref{paramtable}.
@@
 Comparison of displacement histograms from experiment and theory.
-Upon calibrating the inverse temperature $\thup$, the TKdV simulations recover the experimentally measured distributions in remarkable detail.
+Upon calibrating the inverse temperature $\itup$, the TKdV simulations recover the experimentally measured distributions in remarkable detail.
 In particular, they exhibit the transition from a nearly symmetric distribution upstream to a highly skewed distribution downstream.
@@
 Here, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.
+
+\todo[inline]{Need to cite \cite{onorato2016origin} as recommended by ref 1 and maybe \cite{onorato2016twenty} which was not recommended by a referee, but is more specific to KdV}
 

### 2020-04-21 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\nick}[1]{{\color{red} #1}}
+\newcommand{\new}[1]{{\color{blue} #1}}
 % Derivatives
@@
 
-\todo[inline]{Need to cite \cite{onorato2016origin} as recommended by ref 1 and maybe \cite{onorato2016twenty} which was not recommended by a referee, but is more specific to KdV}
-
 The explicit formula for outgoing skewness arises directly from the statistical matching condition \eqref{statmatch}, which states that the downstream Hamiltonian must match in expected value at the ADC, i.e.~with respect to the incoming and outgoing Gibbs measures.
@@
 It is a rather unexpected relationship, as it would have been difficult to anticipate that the displacement skewness and slope variance, among all possible variable combinations, are so intimately related.
+%
+\new{We remark that Eq.~\eqref{SkewPred} bears similarity to a formula derived by Onorato \& Suret \cite{onorato2016twenty} in the context of {\em constant-depth} KdV that relates the skewness to the spectral bandwidth.
+A similar formula relates kurtosis to spectral bandwidth in the context of the nonlinear Schr{\"o}dinger equation \cite{onorato2016origin}.
+Indeed, to leading order $\eta_t = c \eta_x$, so that $c^2 \var(\eta_x)$ can be approximated by the spectral bandwidth $\var(\eta_t)$ in Eq.~\eqref{SkewPred}.
+We note that these earlier works considered only {\em constant} depth situations, where changes in spectral bandwidth, and the associated changes in skewness/kurtosis arose from out-of-equilibrium initial conditions.
+In contrast, the present study....
+}
+
+\vsp{4}
+\new{The present study derives a formula for changing depth, in contrast to the simpler situation of constant depth where changes in spectral bandwidth arise from out-of-equilibrium initial conditions, as considered in \cite{onorato2016twenty, onorato2016origin}}
+
+%\todo[inline]{Need to cite \cite{onorato2016origin} as recommended by ref 1 and maybe \cite{onorato2016twenty} which was not recommended by a referee, but is more specific to KdV}
 

### 2020-04-23 paper2_JNLS/paper.tex
NM

--- 2020-04-20 paper2_JNLS/paper.tex
@@
 Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures.
-Rogue waves, also variously known as freak or anomalous waves, have been observed in shallow \cite{pelinovsky2000nonlinear, gramstad2013freak}, intermediate \cite{karmpadakis2019laboratory}, and deep water \cite{dematteis2018rogue, dematteis2019experimental}, and certain rogue features have even been recovered by exact solutions to various wave models \cite{peregrine1983water, clarkson2017rational, chen2019periodic}.
+Rogue waves, also variously known as freak or anomalous waves, have been observed in shallow \cite{pelinovsky2000nonlinear, gramstad2013freak}, intermediate \cite{karmpadakis2019laboratory}, and deep water \cite{dematteis2018rogue, dematteis2019experimental}, \new{ and certain features of rogue waves have been recovered by exact solutions to various wave models \cite{peregrine1983water, clarkson2017rational, chen2019periodic}}.
 These abnormal waves can be triggered by a variety of mechanisms, including anomalous wind-forcing \cite{kharif2008influence, toffoli2017wind}, opposing currents \cite{garrett2009rogue, onorato2011triggering}, focusing due to variable bathymetry \cite{heller2008refraction, white1998chance},
@@
 
+	\new{We remark that Eq.~\eqref{SkewPred} bears similarity to a formula derived by Onorato \& Suret \cite{onorato2016twenty}, in the context of {\em constant-depth} KdV, that relates variations in skewness to variations in spectral bandwidth.
+Indeed, since $\eta_t \approx c \eta_x$ to leading order, the term $\var(\eta_x)$ in Eq.~\eqref{SkewPred} could be (approximately) interpreted as the spectral bandwidth.
+We note that an analogous formula has been derived for the nonlinear Schr{\"o}dinger equation that relates kurtosis to bandwidth variations \cite{onorato2016origin}.
+The present study differs in that it predicts a moment-bandwidth relation for waves propagating over {\em variable} depth, where it is the depth change that creates system disequilibrium.
+A small transition region connects two near-equilibrium systems, i.e.
+one far upstream and one far downstream.
+The link between these two equilibrium states is made mathematically precise by the statistical matching condition \eqref{statmatch}, which is the key ingredient needed to derive Eq.~\eqref{SkewPred}.
+This is in contrast to the simpler situation of waves propagating over constant depth, where spectral variations and the associated moment variations arise from out-of-equilibrium initial conditions \cite{onorato2016twenty, onorato2016origin}.}
+
  %----------------------------------------------------------------------------%

### 2020-04-23 paper2_JNLS/paper.tex
NM

@@
 Finally, the theory predicts a peculiar relationship between the outgoing displacement skewness and the change in slope variance, specifically how their ratio depends on the wave amplitude and depth ratio.
-New experimental measurements confirm this prediction in spectacular fashion.
+\new{New experimental measurements provide convincing evidence for this prediction.}
 \end{abstract}
-%New experimental measurements provide convincing evidence for this prediction.
+
+% ORIGINAL: New experimental measurements confirm this prediction in spectacular fashion.
+
+% OTHER
+% New experimental measurements provide convincing evidence for this prediction.
+% convincing, conclusive, compelling.
 
@@
 As seen in \ref{ExpSpatialStats}(b)--(c), both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm.
-Interestingly, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude.
-The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e).
+Interestingly, the location of the peak is the same for skewness and kurtosis and \new{is} insensitive to driving amplitude.
+The maximum values of skewness and kurtosis (approximately 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e).
 \boetal found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution across all of the experiments.
-We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.
+We note that we have gathered data for 15 different driving amplitudes, but only display 6 of those in Fig.~\ref{ExpSpatialStats} to avoid clutter.
 
@@
 
-% Remarkable, spectacular, splendid, stupendous, striking, amazing, exquisite
-% Compelling, convincing
-
 	Next, we aim to make the same comparison for free-surface slopes.
@@
 
-	Finally, Fig.~\ref{trajectories} sheds some light on the timescales required for wave evolution in the TKdV framework.
-It appears that the normalized time of $t=10$ is roughly the correct timescale to observe substantial wave dynamics.
+	Finally, Fig.~\ref{trajectories} sheds light on the timescales required for wave evolution in the TKdV framework.
+It appears that the normalized time of $t=10$ is \new{approximately} the correct timescale to observe substantial wave dynamics.
 More precisely, both the left-going waves in Fig.~\ref{trajectories}(a) and the right-going waves in Fig.~\ref{trajectories}(b) require a dimensionless time of about $t=7$ to traverse the entire periodic domain.
 For the representative experiments, the characteristic timescale, $\timescale$, is 2.8~sec upstream and 5.7~sec downstream.
-Hence, a dimensionless time of $t=7$ corresponds to roughly 20~sec and 40~sec upstream and downstream respectively.
+Hence, a dimensionless time of $t=7$ corresponds to 20~sec and 40~sec upstream and downstream respectively.
 Given the wave speeds ($c = $ 110 cm/s upstream and 54 cm/s downstream), these timescales can be converted to distances traveled by the waves.
 In fact, a simple calculation using the definitions in Section \ref{theory} gives the distance traveled as $7 \lamfac g/(2 \pi \freqp^2)$, which, due to a serendipitous cancellation, is independent of depth.
-Hence, the distance traveled by the waves over a dimensionless time of $t=7$ is the same value upstream and downstream,  roughly 22 meters.
+Hence, the distance traveled by the waves over a dimensionless time of $t=7$ is the same value upstream and downstream, 22 meters.
 Note that this distance is significantly greater than the 6-meter length of the experimental wave tank.
@@
 
-Inserting \eqref{H3convert} and \eqref{H2convert} into \eqref{H3H2ratio}, and using the definitions of $C_2$ and $C_3$ from \eqref{C3C2}, gives the remarkably explicit formula
+Inserting \eqref{H3convert} and \eqref{H2convert} into \eqref{H3H2ratio}, and using the definitions of $C_2$ and $C_3$ from \eqref{C3C2}, gives the explicit formula
 \begin{equation}
@@
 While the data shows significant variation at small amplitudes, it shows a well defined decreasing trend at larger amplitudes.
-In particular, the red dashed line shows the power-law $\epsup^{-3}$ predicted by \eqref{SkewPred}, which captures the experimental trend remarkably well.
+In particular, the red dashed line shows the power-law $\epsup^{-3}$ predicted by \eqref{SkewPred}, which captures the experimental trend \new{closely}.
 
@@
 Indeed, since $\eta_t \approx c \eta_x$ to leading order, the term $\var(\eta_x)$ in Eq.~\eqref{SkewPred} could be (approximately) interpreted as the spectral bandwidth.
-We note that an analogous formula has been derived for the nonlinear Schr{\"o}dinger equation that relates kurtosis to bandwidth variations \cite{onorato2016origin}.
+We note that an analogous formula has been derived for the nonlinear Schr{\"o}dinger equation relating kurtosis to bandwidth variations \cite{onorato2016origin}.
 The present study differs in that it predicts a moment-bandwidth relation for waves propagating over {\em variable} depth, where it is the depth change that creates system disequilibrium.

### 2020-04-30 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{graphicx,amsmath,amsfonts,amssymb,stmaryrd}
\usepackage{color}
\newcommand{\comment}[1]{{\color{blue} #1}}

%----------------------------- CUSTOM COMMANDS -----------------------------%

% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
\newcommand{\HERE}[1]{ \vsp{#1} {\color{blue} HERE} }
% Mathy stuff
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\omavg}{\omega_0}
\newcommand{\omsig}{\sigma_{\omega}}
\newcommand{\dth}{\Delta \theta}
\newcommand{\dom}{\Delta \omega}
\newcommand{\etag}{\eta_{\gamma}}
\newcommand{\etastd}{\eta_{\text{std}}}
\newcommand{\mean}[1]{\left< #1\right>}

% Upstream and downstream notation (not used)
\newcommand{\hm}{h^-}
\newcommand{\hp}{h^+}
\newcommand{\lamm}{\lambda^-}
\newcommand{\lamp}{\lambda^+}
\newcommand{\km}{k^-}
\newcommand{\kp}{k^+}
\newcommand{\khm}{(kh)^-}
\newcommand{\khp}{(kh)^+}

%----------------------------- DOCUMENT -----------------------------%
\begin{document}

\section*{Referee 1}

\noindent
\comment{The paper deals with the statistics of the 1D evolution of a surface-gravity-wave field in shallow water, encountering an abrupt depth reduction. The authors perform experimental investigations carried out in a water tank and in parallel they develop a theory based on the statistical mechanics of the Korteveg de Vries equation, used to interpret the experimental results. The two aspects, experimental and theoretical, have been considered before, in References [5] and [25], but here they are put together in an expanded and synergetic way that makes the manuscript clear and interesting. Moreover, although many of the results have been presented before, the analysis of the experimental data of surface slope is new. This is performed in light of the recently derived relationship linking the ratio between displacement skewness and variations of slope variance to the nonlinearity in the system. The agreement of the experimental and theoretical scalings of the ratio is remarkable over a wide range of amplitudes, providing compelling evidence that the theory proposed captures the fundamental, leading-order aspects of the phenomenon.
}
\begin{itemize}
\item We thank the referee for these positive remarks.
\end{itemize}

\noindent
\comment{
1. Equilibrium statistical mechanics, i.e. ultimately the use of the canonical ensemble, is indeed a powerful tool. Though, too often the relaxation conditions required for the machinery to apply are overlooked. Here, the 1D KdV equation is an integrable system, whose trajectories are constrained by an infinite number of conserved quantities and whose highly nonlinear regime is governed by coherent structures such as solitons. The description of these structures is out of reach for the Gibbs measure – see the recent rise of an entire subject, namely integrable turbulence, to describe the statistics of integrable systems such as KdV or NLS. I do believe that some weak ergodicity, especially for the statistics of physical-space variables, may justify a sort of relaxation to the Gibbs measure before the ADC is encountered, but all of this deserves being discussed, and should at least be given careful consideration. Consider for instance the most linear cases in the paper: at the wave maker the measure in Fourier space is the one induced by equation (1). Since the modes do not interact (linear limit), the measure at the ADC will also be the one with the spectrum of eq. (1); i.e., relaxation to the Gibbs measure does not take place (simply because the necessary randomizing interactions are absent in the system), in which case the canonical ensemble is justified if and only if the initial condition at the wave maker is itself distributed according to the Gibbs measure. This fact should be discussed and may explain why the theory does not match the experimental points for low amplitudes in figure (8.c).}

\begin{itemize}
\item The referee is correct that equilibrium statistical mechanics and the canonical ensemble is used here, not because it is rigorously justified, but as an effective theory that explains the data.
It is true that the 1D KdV equation is an integrable system whose trajectories are constrained by an infinite number of conserved quantities. However, the {\em truncated} system does not possess these constraints (and pointed out in the original text).
%\begin{quotation}
%\end{quotation}
\item a
\end{itemize}

\noindent
\comment{}

\begin{itemize}
\item
\end{itemize}

\end{document}

### 2020-04-30 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{graphicx,amsmath,amsfonts,amssymb,stmaryrd}
\usepackage{color}
\newcommand{\comment}[1]{{\color{blue} #1}}

%----------------------------- CUSTOM COMMANDS -----------------------------%

% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
\newcommand{\HERE}[1]{ \vsp{#1} {\color{blue} HERE} }
% Mathy stuff
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\omavg}{\omega_0}
\newcommand{\omsig}{\sigma_{\omega}}
\newcommand{\dth}{\Delta \theta}
\newcommand{\dom}{\Delta \omega}
\newcommand{\etag}{\eta_{\gamma}}
\newcommand{\etastd}{\eta_{\text{std}}}
\newcommand{\mean}[1]{\left< #1\right>}

% Upstream and downstream notation (not used)
\newcommand{\hm}{h^-}
\newcommand{\hp}{h^+}
\newcommand{\lamm}{\lambda^-}
\newcommand{\lamp}{\lambda^+}
\newcommand{\km}{k^-}
\newcommand{\kp}{k^+}
\newcommand{\khm}{(kh)^-}
\newcommand{\khp}{(kh)^+}

%----------------------------- DOCUMENT -----------------------------%
\begin{document}

\section*{Referee 1}

% COMMENT
\noindent
\comment{A nice paper with agreement between theory and experiments, and novel new results.
}
\begin{itemize}
\item We thank the referee for these positive remarks.
\end{itemize}

% COMMENT 1
\noindent
\comment{
Use of qualitative adjectives where a quantitative description may be better:
Pg 1 Ln 39: spectacular
Pg 17 Ln 17: intriguing
Pg 18 Ln 36: modest
Pg 19 Ln 37: roughly
Pg 20 Ln 42: remarkably}

\begin{itemize}
\item We have corrected these.
\end{itemize}

% COMMENT 2
\noindent
\comment{On Pg 5 Ln 44, it is said that "the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude". Can you elaborate on "appears" to say whether it is or not?}

\begin{itemize}
\item Yes
\end{itemize}

% COMMENT 3
\noindent
\comment{Section 4.1 refers to Figure 3(b). If you increase N, with the transfer function become eventually collapse to one single curve? If so or indeed if not, why choose the N and Lambda in this range?}

\begin{itemize}
\item
\end{itemize}

% COMMENT 4
\noindent
\comment{For the experiments, images are taken at 60 frames/sec. The numerical differentiation of images of the free surface uses Savitzky-Golay smoothing filters. However, on Pg 18 it is mentioned that there are measurement errors and the differentiation amplifies errors. Can you comment further on this method by, say, looking at how it converges while only looking the images from 20 and 40 frames/sec or by calculating $\eta_x$ by a spectral method from eta? }

\begin{itemize}
\item
\end{itemize}

% COMMENT 5
\noindent
\comment{Fig 6 could use a label for the colourbar.}

\begin{itemize}
\item
\end{itemize}

% COMMENT 6
\noindent
\comment{Sharp steps do occur in bathymetry profiles. However, do you think Eq (53) will still apply rather well over other bathymetry profiles such as a transition with a steep finite gradient?}

\begin{itemize}
\item
\end{itemize}

% COMMENT 7
\noindent
\comment{Minor typos etc.
Pg 6 Ln 60: is leading-order -> is the leading-order
Pg 7 Ln 21: change dash to en or em dash as appropriate
Pg 16 Ln 35: Fig. 3(b) -> Fig. 3(c)
Pg 24: References [17] and [24] require capitals as they are book titles, and [27] requires a capital H for Hamiltonian}

\begin{itemize}
\item
\end{itemize}

% COMMENT 8
\noindent
\comment{}

\begin{itemize}
\item
\end{itemize}
\end{document}

### 2020-04-30 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

@@
 \noindent
-\comment{A nice paper with agreement between theory and experiments, and novel new results.
-}
-\begin{itemize}
-\item We thank the referee for these positive remarks.
-\end{itemize}
+\comment{A nice paper with agreement between theory and experiments, and novel new results.} \\
+
+\noindent
+We thank the referee for these positive remarks. \\ \\
 
@@
 \comment{
-Use of qualitative adjectives where a quantitative description may be better:
-Pg 1 Ln 39: spectacular
-Pg 17 Ln 17: intriguing
-Pg 18 Ln 36: modest
-Pg 19 Ln 37: roughly
-Pg 20 Ln 42: remarkably}
+Use of qualitative adjectives where a quantitative description may be better: \\
+Pg 1 Ln 39: spectacular \\
+Pg 17 Ln 17: intriguing \\
+Pg 18 Ln 36: modest \\
+Pg 19 Ln 37: roughly \\
+Pg 20 Ln 42: remarkably} \\
 
+\noindent
+We have replaced the majority of these descriptors with more quantitative ones. In particular:
 \begin{itemize}
-\item We have corrected these.
+\item The term `spectacular fashion' in the abstract was changed to `convincing evidence'.
+\item The word `modest' on page 18 was replaced with `intermediate'.
+\item The word `roughly' (appearing in several places in the original manuscript) was either replaced by `approximately' or else removed.
+\item The phrase `captures the trend remarkably well' on page 20 was changed to `captures the trend closely'.
+\item We have kept `intriguing' on page 17 as we believe it to be an appropriate use of the word.
 \end{itemize}
+\hphantom \\
 
@@
 \begin{itemize}
-\item Yes
+\item As can bee seen in Fig.
+2 (b) and (c), the location of the peak is indeed insensitive to driving amplitude to within experimental error (the level of which can be estimated by the noise in the data).
+Accordingly, we have replaced `appears insensitive' to simply `is insensitive'.
+We had initially used the word `appears' to encourage the reader to view the figure and see for themselves.
+However, we now see the referee's point that this word may introduce some uncertainty in the reader's mind and have thus made the change.
 \end{itemize}
@@
 \comment{Section 4.1 refers to Figure 3(b).
-If you increase N, with the transfer function become eventually collapse to one single curve?
-If so or indeed if not, why choose the N and Lambda in this range?}
+If you increase $N$, with the transfer function become eventually collapse to one single curve?
+If so or indeed if not, why choose the $N$ and $\Lambda$ in this range?}
 
 \begin{itemize}
-\item
+\item This is a great question....
 \end{itemize}
@@
 \begin{itemize}
-\item
+\item We have added  labels for the colourbars in Fig. 6.
 \end{itemize}

### 2020-04-30 code_julia_sampling/io_routines.jl
NM

@@ -1,18 +1,0 @@
-
-function writedata(data::Array, filename::AbstractString)
-	iostream = open(filename, "w")
-	writedlm(iostream, data)
-	close(iostream)
-end
-
-function readvec(file::AbstractString)
-	iostream = open(file, "r")
-	vec1 = readdlm(iostream, comments=true, comment_char='#')[:,1]
-	close(iostream)
-	return vec1
-end
-
-# Significant digits
-function sig(var, sigdig)
-	return round(var,sigdigits=sigdig)
-end

### 2020-04-30 code_julia_sampling/math_routines.jl

@@ -147,3 +147,25 @@ function uniform_sample(nmodes::Int, nsweeps::Int,
 	#savemicro ? rsave = Float32.(rvar) : rsave = []
 end
 #---------------------------------------#
+
+
+
+
+
+function writedata(data::Array, filename::AbstractString)
+	iostream = open(filename, "w")
+	writedlm(iostream, data)
+	close(iostream)
+end
+
+function readvec(file::AbstractString)
+	iostream = open(file, "r")
+	vec1 = readdlm(iostream, comments=true, comment_char='#')[:,1]
+	close(iostream)
+	return vec1
+end
+
+# Significant digits
+function sig(var, sigdig)
+	return round(var,sigdigits=sigdig)
+end

### 2020-04-30 code_julia_sampling/math_routines.jl
NM

@@ -139,7 +139,7 @@ function uniform_sample(nmodes::Int, nsweeps::Int,
 		println("Sweep ",nn,", ",sig(100*nn/nsweeps,3),"% completed.")
 	end
 	zerolast ? zstr = "z" : zstr = ""
-	savefile = string("rand-",string(nmodes),zstr,"-",string(nsweeps),".jld")
+	savefile = string(data_folder(),"rand-",string(nmodes),zstr,"-",string(nsweeps),".jld")
 	save(savefile, "rr", RandList([],H2vec,H3vec),
 		"nmodes", nmodes, "nsweeps", nsweeps, "totsamp", totsamp)
 
@@ -150,7 +150,9 @@ end
 
 
 
-
+function data_folder()
+	return "./Data/"
+end
 
 function writedata(data::Array, filename::AbstractString)
 	iostream = open(filename, "w")

### 2020-04-30 code_julia_sampling/run1.jl

@@ -2,13 +2,12 @@
 include("transfer.jl")
 
 
-
 # Test case with Lambda = 16, nsweeps = 2, lamfac = 8.
-#@time( uniform_sample(16, 2) )
-#@time( transfun("rand-16z-2.jld", 8) )
-#output_text("thvars-16z-8-2.jld")
+@time( uniform_sample(16, 2) )
+@time( transfun("rand-16z-2.jld", 8) )
+output_text("thvars-16z-8-2.jld")
 
 
 # RUNS
- @time( transfun("rand-12z-1000.jld", 8) )
- @time( transfun("rand-12z-1000.jld", 6) )
+ #@time( transfun("rand-12z-1000.jld", 8) )
+ #@time( transfun("rand-12z-1000.jld", 6) )

### 2020-04-30 code_julia_sampling/transfer.jl

@@ -3,9 +3,9 @@ For given upstream theta, enforce the matching condition to
 compute the downstream theta.=#
 
 include("math_routines.jl")
-include("io_routines.jl")
 
 function transfun(randfile::AbstractString, lamfac::Int)
+	randfile = string(data_folder(), randfile)
 	rr, nmodes, nsweeps = load(randfile, "rr", "nmodes", "nsweeps")
 	# Parameters
 	eps0 = 0.017
@@ -36,7 +36,7 @@ function transfun(randfile::AbstractString, lamfac::Int)
 		println("Upstream skewness = ", sig(skup[nn],3))
 		println("Downstream skewness = ", sig(skdn[nn],3), "\n")
 	end
-	savefile = string("thvars-", string(nmodes), "z-", 
+	savefile = string(data_folder(), "thvars-", string(nmodes), "z-", 
 		string(lamfac), "-", string(nsweeps), ".jld")
 	save(savefile, "thup", thup, "thdn", thdn, "skup", skup, "skdn", skdn, 
 		"nmodes", nmodes, "lamfac", lamfac, "nsweeps", nsweeps)
@@ -63,12 +63,13 @@ function plotstuff(datafile::AbstractString)
 end
 
 function output_text(datafile::AbstractString)
+	datafile = string(data_folder(), datafile)
 	thup,thdn,skup,skdn = load(datafile, "thup", "thdn", "skup", "skdn")
 	nmodes,lamfac,nsweeps = load(datafile, "nmodes", "lamfac", "nsweeps")
 	nthvals = length(thup)
 	ratio = skdn./skup
 	outdata = [nmodes; lamfac; nthvals; thup; thdn; skup; skdn]
-	outfile = string("transf-", string(nmodes), "z-", 
+	outfile = string(data_folder(), "transf-", string(nmodes), "z-", 
 		string(lamfac), "-", string(nsweeps), ".txt")
 	writedata(outdata, outfile)
 end

### 2020-04-30 code_julia_sampling/math_routines.jl
NM

@@ -126,8 +126,7 @@ end
 Uses several sweeps manages the memory better. 
 Note: zerolast zeroes the last mode as is done in the Matlab DNS. 
 Note: savemicro not yet addressed at all. =#
-function uniform_sample(nmodes::Int, nsweeps::Int, 
-		zerolast::Bool=true, savemicro::Bool=false)
+function uniform_sample(nmodes::Int, nsweeps::Int; zerolast::Bool=true, savemicro::Bool=false)
 	samp_per = 10^5
 	totsamp = nsweeps*samp_per
 	H2vec, H3vec = [zeros(Float32,totsamp) for nn=1:2]
@@ -154,20 +153,72 @@ function data_folder()
 	return "./Data/"
 end
 
-function writedata(data::Array, filename::AbstractString)
-	iostream = open(filename, "w")
-	writedlm(iostream, data)
-	close(iostream)
-end
-
-function readvec(file::AbstractString)
-	iostream = open(file, "r")
-	vec1 = readdlm(iostream, comments=true, comment_char='#')[:,1]
-	close(iostream)
-	return vec1
-end
 
 # Significant digits
 function sig(var, sigdig)
 	return round(var,sigdigits=sigdig)
+end
+
+
+
+
+
+
+
+
+
+
+
+#= Calculate the transfer function:
+For given upstream theta, enforce the matching condition to 
+compute the downstream theta.=#
+
+function transfun(randfile::AbstractString, lamfac::Int)
+	randfile = string(data_folder(), randfile)
+	rr, nmodes, nsweeps = load(randfile, "rr", "nmodes", "nsweeps")
+	# Parameters
+	eps0 = 0.017
+	del0 = 0.23
+	D0 = 0.24
+	# Constants
+	C2,C3 = C2C3(eps0,del0,lamfac)
+	cc = ConstantList(C2,C3,D0)
+	# Compute the list of Hamiltonians using precomputed H2 and H3
+	hdn = Float32.(hamdn(rr,cc))
+	hup = Float32.(hamup(rr,cc))
+
+	# Specify list of upstream thetas.
+	thup = 0.: 0.5 : 30.
+	nth = length(thup)
+	thdn = zeros(nth); skup = zeros(nth); skdn = zeros(nth)
+	guess = thup[1]
+	for nn = 1:nth
[…]

### 2020-04-30 code_julia_sampling/transfer.jl

@@ -1,75 +1,0 @@
-#= Calculate the transfer function:
-For given upstream theta, enforce the matching condition to 
-compute the downstream theta.=#
-
-include("math_routines.jl")
-
-function transfun(randfile::AbstractString, lamfac::Int)
-	randfile = string(data_folder(), randfile)
-	rr, nmodes, nsweeps = load(randfile, "rr", "nmodes", "nsweeps")
-	# Parameters
-	eps0 = 0.017
-	del0 = 0.23
-	D0 = 0.24
-	# Constants
-	C2,C3 = C2C3(eps0,del0,lamfac)
-	cc = ConstantList(C2,C3,D0)
-	# Compute the list of Hamiltonians using precomputed H2 and H3
-	hdn = Float32.(hamdn(rr,cc))
-	hup = Float32.(hamup(rr,cc))
-
-	# Specify list of upstream thetas.
-	thup = 0.: 0.5 : 30.
-	nth = length(thup)
-	thdn = zeros(nth); skup = zeros(nth); skdn = zeros(nth)
-	guess = thup[1]
-	for nn = 1:nth
-		meandiff(thdn) = meanham(hdn, hdn, thdn) - meanham(hdn, hup, thup[nn])
-		thdn[nn] = find_zero(meandiff, guess, Order1())
-		guess = thdn[nn]
-		nn>1 ? guess = 2*thdn[nn] - thdn[nn-1] : 0.	# Linear extrapolation
-		println("Iteration ", nn, "; thup = ", thup[nn], "; thdn = ", sig(thdn[nn],3))
-
-		# Compute the skewness of eta
-		skup[nn] = skewu(rr.H3, hup, thup[nn])
-		skdn[nn] = skewu(rr.H3, hdn, thdn[nn])
-		println("Upstream skewness = ", sig(skup[nn],3))
-		println("Downstream skewness = ", sig(skdn[nn],3), "\n")
-	end
-	savefile = string(data_folder(), "thvars-", string(nmodes), "z-", 
-		string(lamfac), "-", string(nsweeps), ".jld")
-	save(savefile, "thup", thup, "thdn", thdn, "skup", skup, "skdn", skdn, 
-		"nmodes", nmodes, "lamfac", lamfac, "nsweeps", nsweeps)
-end
-
-function plotstuff(datafile::AbstractString)
-	thup,thdn,skup,skdn = load(datafile, "thup", "thdn", "skup", "skdn")
[…]

### 2020-04-30 code_julia_sampling/math_routines.jl
NM

@@ -1,224 +1,0 @@
-#---------- Basic Stuff ----------#
-using Distributions
-using Roots
-using DelimitedFiles
-using LinearAlgebra
-using FFTW
-using JLD
-using Plots
-using LaTeXStrings
-#---------------------------------#
-
-#---------- Real FFT Routines ----------#
-#= Compute the FFT between physical and spectral space assuming
-1) the signal uu is real and 2) uu has zero mean (i.e. momentum M=0)
-Note: uu has length npoints = 2*nmodes or 2*nmodes + 1. =#
-#= Valid choices for npoints are 2*nmodes or 2*nmodes+1, 
-but I need a function to keep my usage consistent.
-Note: for H3test npts = 2*modes is slightly more accurate. =#
-function npoints(nmodes::Int)
-	return 2*nmodes
-end
-#= Basic realfft to go from uu to uhat. =#
-function realfft(uu::Vector{Float64})
-	uhat = rfft(uu)/length(uu)
-	@assert(abs(uhat[1])/maximum(abs,uhat) < 1e-6) 
-	return uhat[2:end]
-end
-#= Basic irealfft to go from uhat to uu (no upsampling). =#
-function irealfft(uhat::Vector{Complex{Float64}})
-	nmodes = length(uhat)
-	uu = irfft([0; uhat], npoints(nmodes))
-	return uu*length(uu)
-end
-#= Upsampled version of irealfft. =#
-function ifftup(uhat::Vector{Complex{Float64}})
-	uhat = [uhat; zeros(eltype(uhat), length(uhat))]
-	return irealfft(uhat)
-end
-#---------------------------------------#
-
-#---------- Hamiltonian Routines ----------#
-#= In all routines, the physical signal is assumed to be real 
-with zero momentum, M = int u dx = 0. Therefore, the Fourier transform 
-only requires the modes k=1,...,nmodes. The negative modes are given 
-by the complex conjugates (CC), and the k=0 mode vanishes due to M=0. =#
-#---------- Low-level Hamiltonian Routines ----------#
-#= Compute the energy, E = 1/2 int u^2 dx. =#
-function energy(uhat::Vector{Complex{Float64}})
[…]

### 2020-04-30 code_julia_sampling/run1.jl

@@ -1,13 +1,0 @@
-## RUN STUFF
-include("transfer.jl")
-
-
-# Test case with Lambda = 16, nsweeps = 2, lamfac = 8.
-@time( uniform_sample(16, 2) )
-@time( transfun("rand-16z-2.jld", 8) )
-output_text("thvars-16z-8-2.jld")
-
-
-# RUNS
- #@time( transfun("rand-12z-1000.jld", 8) )
- #@time( transfun("rand-12z-1000.jld", 6) )

### 2020-05-02 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

@@
 \item The term `spectacular fashion' in the abstract was changed to `convincing evidence'.
-\item The word `modest' on page 18 was replaced with `intermediate'.
+\item The word `modest' on page 18 was replaced with `moderate'.
+Thanks for pointing this out -- I believe moderate was the word that was originally intended and `modest' was the result of an autocorrect.
 \item The word `roughly' (appearing in several places in the original manuscript) was either replaced by `approximately' or else removed.
@@
 \begin{itemize}
-\item This is a great question....
+\item This is a great question.... \\
+Why was this value chosen? One concern is computational time; These are expensive sampling computations, considering that for each value of theta you have to run a sample.
 \end{itemize}

### 2020-05-02 code_julia_sampling/routines.jl

@@ -21,9 +21,14 @@ end
 #---------- Real FFT Routines ----------#
 #= Compute the FFT between physical and spectral space assuming
 the signal u is real and has zero mean. =#
-
+#= Basic realfft to go from physical to spectral space. 
+This routine is only used in the benchmark. =#
+function realfft(uu::Vector{Float64})
+	uhat = rfft(uu)/length(uu)
+	@assert(abs(uhat[1])/maximum(abs,uhat) < 1e-6) 
+	return uhat[2:end]
+end
 #= Basic irealfft to go from spectral to physical space. =#
-#= In physical space, u has to have length 2*nmodes (preferred) or 2*nmodes+1 =#
 function irealfft(uhat::Vector{Complex{Float64}})
 	uu = irfft([0; uhat[:]], 2*length(uhat))
 	return uu*length(uu)

### 2020-05-02 code_julia_sampling/run.jl

@@ -2,14 +2,14 @@
 include("routines.jl")
 
 # Test case with Lambda = 16, nsweeps = 10, lamfac = 8.
-uniform_sample(16, 2)
+#uniform_sample(16, 2)
 #transfun("rand-16z-100.jld", 8)
 #output_text("thvars-16z-8-10.jld")
 
 # RUN
-#= transfun("rand-16z-500.jld", 8)
+#transfun("rand-16z-500.jld", 8)
 transfun("rand-16z-1000.jld", 8)
 transfun("rand-32z-500.jld", 8)
 transfun("rand-32z-1000.jld", 8)
 transfun("rand-32z-500.jld", 16)
-transfun("rand-32z-1000.jld", 16) =#
+transfun("rand-32z-1000.jld", 16)

### 2020-05-04 code_julia_sampling/routines.jl
NM

@@ -12,7 +12,7 @@ using JLD
 using Plots
 using LaTeXStrings
 # Set the data folder.
-data_folder() = "./Data"
+data_folder() = "./Data/"
 # Round to a set number of significant digits
 sig(var, sigdig::Int) = round(var,sigdigits=sigdig)
 #---------------------------------#

### 2020-05-04 code_julia_sampling/run.jl

@@ -3,13 +3,22 @@ include("routines.jl")
 
 # Test case with Lambda = 16, nsweeps = 10, lamfac = 8.
 uniform_sample(16, 2)
-#transfun("rand-16z-100.jld", 8)
-#output_text("thvars-16z-8-10.jld")
+transfun("rand-16z-2.jld", 8)
+output_text("thvars-16z-8-2.jld")
 
-# RUN
+# RUN TRANSFUN
 #transfun("rand-16z-500.jld", 8)
 #transfun("rand-16z-1000.jld", 8)
 #transfun("rand-32z-500.jld", 8)
 #transfun("rand-32z-1000.jld", 8)
 #transfun("rand-32z-500.jld", 16)
 #transfun("rand-32z-1000.jld", 16)
+
+#= RUN OUTPUT
+output_text("thvars-16z-8-500.jld")
+output_text("thvars-16z-8-1000.jld")
+output_text("thvars-32z-8-500.jld")
+output_text("thvars-32z-8-1000.jld")
+output_text("thvars-32z-16-500.jld")
+output_text("thvars-32z-16-1000.jld") 
+=#

### 2020-05-04 code_julia_sampling/tests.jl

@@ -67,7 +67,7 @@ end
 #---------------------------------------#
 
 
-#---------- HAMILTONIAN TESTS ----------#
+#---------- HAMILTONIAN ACCURACY TESTS ----------#
 #= Compute the exact H2 and H3 for a specific cos^2 function. =#
 function ham_ex1(nmodes::Int, aa::Float64, x0::Float64)
 	# Define the function.
@@ -145,23 +145,53 @@ function test_ham(nmodes::Int; method::Int=3)
 		println("Rel diff = ", rel_err(h3fft, h3rec))
 	end
 end
-test_ham(16; method=3)
+#test_ham(16; method=3)
 
 
-
+#---------- H3 SPEED TESTS ----------#
 # Test the speed of h3fft versus h3rec
-function test_ham3speed(nmodes::Int)
+function ham3speed(nmodes::Int, ncalls::Int)
 	uhat = randn(ComplexF64, nmodes, ncalls)
-	tm_fft = @elapsed(
-		for nn = 1:ncalls
-			h3fft = ham3fft(uhat[:,nn])
-			println("h3fft = ", sig(h3fft,3))
-		end)
-	tm_rec = @elapsed(
-		for nn = 1:ncalls
-			h3rec = ham3rec(uhat[:,nn])
-			println("h3rec = ", sig(h3rec,3))
-		end)
-	println("Rel error: ", rel_err(h3fft, h3rec)
+	# Call each one time just in case the first call is much slower.
+	h3fft = ham3fft(uhat[:,1]) 
+	h3rec = ham3rec(uhat[:,1])
+	# Call each routine many times and measure the time.
+	tm_fft = @elapsed([h3fft = ham3fft(uhat[:,nn]) for nn=1:ncalls])
+	tm_rec = @elapsed([h3rec = ham3rec(uhat[:,nn]) for nn=1:ncalls])
+	#Print computational times
+	println("\nTesting speed for nmodes = ", nmodes)
+	println("Time for fft: ", sig(tm_fft,3), "secs")
+	println("Time for rec: ", sig(tm_rec,3), "secs")
+	println("Rec is faster by factor of ", sig(tm_fft/tm_rec,2))	
+	# Print values of h3 just to make sure things are being computed correctly.
+	#println("\nh3fft = ", sig(h3fft, 3), "; h3rec = ", sig(h3rec,3))
+	#println("Rel err: ", rel_err(h3fft, h3rec))
+	return tm_fft, tm_rec
[…]

### 2020-05-04 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

@@
 \noindent
-\comment{On Pg 5 Ln 44, it is said that "the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude".
+\comment{On Pg 5 Ln 44, it is said that ``the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude''.
 Can you elaborate on "appears" to say whether it is or not?}

### 2020-05-04 paper2_JNLS/paper.tex
NM

@@
 \author{
-M.~N.~J.~Moore\thanks{Florida State University}, 
+M.~N.~J.~Moore\thanks{United States Naval Academy}, 
 C.~Tyler Bolles\thanks{University of Michigan},
@@
 	Visual differences between the upstream and downstream dynamics are apparent in Fig.~\ref{trajectories}.
-The upstream trajectory shows several waves of modest amplitude all propagating leftward.
+The upstream trajectory shows several waves of \new{moderate} amplitude all propagating leftward.
 Here, the magnitude of the positive and negative displacements are comparable.

### 2020-05-04 code_julia_sampling/run.jl

@@ -2,9 +2,9 @@
 include("routines.jl")
 
 # Test case with Lambda = 16, nsweeps = 10, lamfac = 8.
-uniform_sample(16, 2)
-transfun("rand-16z-2.jld", 8)
-output_text("thvars-16z-8-2.jld")
+#uniform_sample(16, 2)
+#transfun("rand-16z-2.jld", 8)
+#output_text("thvars-16z-8-2.jld")
 
 # RUN TRANSFUN
 #transfun("rand-16z-500.jld", 8)
@@ -14,11 +14,10 @@ output_text("thvars-16z-8-2.jld")
 #transfun("rand-32z-500.jld", 16)
 #transfun("rand-32z-1000.jld", 16)
 
-#= RUN OUTPUT
-output_text("thvars-16z-8-500.jld")
-output_text("thvars-16z-8-1000.jld")
-output_text("thvars-32z-8-500.jld")
-output_text("thvars-32z-8-1000.jld")
-output_text("thvars-32z-16-500.jld")
-output_text("thvars-32z-16-1000.jld") 
-=#
+# RUN OUTPUT
+#output_text("thvars-16z-8-500.jld")
+#output_text("thvars-16z-8-1000.jld")
+#output_text("thvars-32z-8-500.jld")
+#output_text("thvars-32z-8-1000.jld")
+#output_text("thvars-32z-16-500.jld")
+#output_text("thvars-32z-16-1000.jld")

### 2020-05-04 code_julia_sampling/tests.jl

@@ -175,9 +175,9 @@ end
 # Could also simply measure the time of various FFT calls
 
 function test_ham3speed()
-	ncalls = 10^3
-	#nmodes = 8:1:32
-	nmodes = 2 .^(3:8)
+	ncalls = 10^4
+	#nmodes = 8:1:128
+	nmodes = 2 .^(3:9)
 	tfft, trec = [zeros(Float64, length(nmodes)) for ii=1:2]
 	for (idx,nm) in enumerate(nmodes)
 		tfft[idx], trec[idx] = ham3speed(nm, ncalls)

### 2020-05-12 paper2_JNLS/paper.tex
NM

@@
 Hence, we will hereafter focus on the physically realistic case of $\invtemp > 0$.
+\new{We also point out that a sufficient amount of nonlinear mixing is required for relaxation to the above Gibbs measure.
+In the linearized system, for example, modes propagate without interacting and hence the initial measure induced by the wave generator remains unchanged.}
 
@@
 While the data shows significant variation at small amplitudes, it shows a well defined decreasing trend at larger amplitudes.
-In particular, the red dashed line shows the power-law $\epsup^{-3}$ predicted by \eqref{SkewPred}, which captures the experimental trend \new{closely}.
+\new{In particular, the decay rate of the ratio, $\skw(\eta) / \Delta \var(\eta_x) \sim \epsup^{-3}$, is well predicted by the explicit formula \eqref{SkewPred}, shown by the red dashed line in Fig.~\ref{SkewRat}(c).
+The apparently less predictable behavior at small amplitudes can be attributed to the fact that, in order for the system to relax to the Gibbs measure, modes must interact through nonlinearity.
+For the smallest amplitude experiments, the amount of nonlinearity is apparently too small for  modes to mix sufficiently.}
 
 	\new{We remark that Eq.~\eqref{SkewPred} bears similarity to a formula derived by Onorato \& Suret \cite{onorato2016twenty}, in the context of {\em constant-depth} KdV, that relates variations in skewness to variations in spectral bandwidth.
-Indeed, since $\eta_t \approx c \eta_x$ to leading order, the term $\var(\eta_x)$ in Eq.~\eqref{SkewPred} could be (approximately) interpreted as the spectral bandwidth.
+Indeed, since $\eta_t \approx c \eta_x$ to leading order, the term $\var(\eta_x)$ in Eq.~\eqref{SkewPred} could be interpreted as approximately the spectral bandwidth.
 We note that an analogous formula has been derived for the nonlinear Schr{\"o}dinger equation relating kurtosis to bandwidth variations \cite{onorato2016origin}.

### 2020-05-12 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

@@
 \begin{itemize}
-\item The referee is correct that equilibrium statistical mechanics and the canonical ensemble is used here, not because it is rigorously justified, but as an effective theory that explains the data.
+\item First, thank you for these insightful comments.
+The referee is correct to point out that a sufficient amount of nonlinear interaction is  required for relaxation to the Gibbs measure, and we agree that this condition not being met is the most likely the explanation for formula (53) not matching the smallest-amplitude data in Fig.
+8(c).
+We have inserted these points in the updated manuscript.
+In particular, see the passage in Section 3.5 (page 11):
+\begin{quotation}
+We also point out that a sufficient amount of nonlinear mixing is required for relaxation to the above Gibbs measure.
+In the linearized system, for example, modes propagate without interacting and hence the initial measure induced by the wave generator remains unchanged.
+\end{quotation}
+Also, see the passage describing Fig. 8(c) in Section 4.4 (page 22)
+\begin{quotation}
+The apparently less predictable behavior at small amplitudes can be attributed to the fact that, in order for the system to relax to the Gibbs measure, modes must interact through nonlinearity.
+For the smallest amplitude experiments, the amount of nonlinearity is apparently too small for  modes to mix sufficiently.
+\end{quotation}
+\item Have we included new references on integrable turbulence???
+%\item The referee is correct that equilibrium statistical mechanics and the canonical ensemble is used here, not because it is rigorously justified, but as an effective theory that explains the data.
 It is true that the 1D KdV equation is an integrable system whose trajectories are constrained by an infinite number of conserved quantities.
@@
 %\end{quotation}
-\item a
 \end{itemize}
@@
 \noindent
-\comment{}
+\comment{
+2. The authors may find it useful to compare their equation (53) to
+ ``On the origin of heavy-tail statistics in equations of the Nonlinear Schrodinger type'', Onorato et al., Phys. Lett. A, 380, 89 (2016), 
+ %where an analogous relationship is derived for NLS. It should be possible to relate the skewness to spectral variations, which may lead to further interesting investigations.
+}
 

### 2020-05-12 paper2_JNLS/paper.tex
NM

@@
 In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian.
-The untruncated KdV equation possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}, but their truncated counterparts are not generally invariants of TKdV.
+
+\new{We comment on some important differences between the KdV and TKdV systems.
+The KdV system possesses in infinite sequence of additional invariants, associated with the fact that it is a completely integrable system.
+However, the analogues of these quantities are {\em not} generally invariants of the truncated system; no invariants beyond the three mentioned above are known to exist for TKdV.
+Hence, while KdV is a completely integrable system whose long-time dynamics are governed by soliton interactions, this same behavior should not be expected for TKdV.
+In particular, the long-time behavior is not simply governed by soliton motion and instead is chaotic and observed to be weakly ergodic \cite{bajars2013weakly}.
+In this way, the theory developed herein differs from, and could likely by complementary to, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, in which integrability is exploited in order to uncover statistical characteristics of the system of interest.
+}
+\new{
+We remind the reader that KdV is asymptotically valid for large spatial scales and hence truncating the small scales (or alternatively modeling them through closure) does not alter the order of approximation made.
+In this sense, KdV and TKdV should be viewed as on equal footing as far as physical realism.
+Where differences exist in long-time dynamics, it is unclear which model, if either, more closely resembles reality.}
 

### 2020-05-12 paper2_JNLS/paper.tex
NM

--- 2020-05-12 paper2_JNLS/paper.tex
@@
 \end{equation}
-% The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
 Thus, the dynamics of interest are confined to the unit hypersphere, $\En = 1$ in $\CC^{\Lambda}$.
-In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian.
-The untruncated KdV equation possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}, but their truncated counterparts are not generally invariants of TKdV.
+%In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian.
+
+\new{We comment on some important differences between the KdV and TKdV dynamical systems.
+In addition to momentum, energy, and Hamiltonian, the KdV system possesses in infinite sequence of invariants, which is, of course, associated with the fact that KdV is a completely integrable system.
+However, the analogues of these quantities are {\em not} generally invariants of the truncated system.
+The only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
+Hence, while KdV is a completely integrable system whose long-time dynamics are governed by soliton interactions, this same behavior should not be expected for TKdV.
+Instead of breaking into a finite number of solitons, TKdV dynamics become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though no proof is available) \cite{bajars2013weakly}.
+In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover statistical characteristics of the system of interest.
+We remind the reader that KdV is asymptotically valid for large spatial scales and hence truncating the small scales does not alter the order of approximation made.
+In this sense, KdV and TKdV should be regarded as being on equal footing as far as physical realism.
+Where differences exist in long-time dynamics, it is unclear {\em a priori} which model, if either, more closely resembles reality.}
 

### 2020-05-13 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

@@
 \item The word `modest' on page 18 was replaced with `moderate'.
-Thanks for pointing this out -- I believe moderate was the word that was originally intended and `modest' was the result of an autocorrect.
+Thanks for pointing this out -- I believe `moderate' was the word that was originally intended and `modest' was the result of an autocorrect.
 \item The word `roughly' (appearing in several places in the original manuscript) was either replaced by `approximately' or else removed.
@@
 \comment{On Pg 5 Ln 44, it is said that ``the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude''.
-Can you elaborate on "appears" to say whether it is or not?}
+Can you elaborate on ``appears" to say whether it is or not?}
 
@@
 We had initially used the word `appears' to encourage the reader to view the figure and see for themselves.
-However, we now see the referee's point that this word may introduce some uncertainty in the reader's mind and have thus made the change.
+However, we now see the referee's point that this word may introduce some uncertainty in the reader's mind and we have thus made the change.
 \end{itemize}
@@
 \begin{itemize}
-\item This is a great question.... \\
-Why was this value chosen? One concern is computational time; These are expensive sampling computations, considering that for each value of theta you have to run a sample.
+\item This is a great question and one that we have already devoted some energy to understanding.
+First, it is unclear whether the transfer function converges to a single curve in the limit $\Lambda \to \infty$ and this is a question currently under investigation (see more below).
+Second, to briefly explain our choices, we selected the system size $\Lambda$ as a moderate value that is large enough to capture sufficiently rich dynamics and small enough to be computationally tractable.
+Meanwhile, we chose $N = \Lambda / 2$ so that the dominant forcing frequency $f_p$ sits near the middle of resolved scales (logarithmically) in the model.
+See sections 3.4 and especially 3.9 in the manuscript for a detailed discussion of the parameter $N$.
+In particular, Section 3.9 outlines a simple scaling argument that suggests scaling $N$ linearly with $\Lambda$ offsets the sensitivity of the transfer function to $\Lambda$ for asymptotically large $\Lambda$ (it is unclear, however, if this mitigates the sensitivity enough).
+
+\item To add a bit more detail, the parameter $\Lambda$ can be regarded as the dynamic range of the model since it corresponds to the number of resolved modes.
+In other work, a subset of the authors examined more severely truncated systems -- as small as two interacting modes -- that were shown to qualitatively capture many statistical features of the experiments (see reference 31).
+Since the goal of the present work is a {\em quantitative} comparison with experiments, we chose a larger dynamic range here.
+We found $\Lambda = 16$ to yield a sufficiently rich dynamical system that compares well to the experiments, while also being manageable from a computational perspective (considering that each trajectory requires integration with a small time step over a long time horizon and we sample over many trajectories).
+These points are discussed in various parts of the manuscript.
+
+\item In  work currently underway, we are developing tailored sampling algorithms that will enable us to compute the transfer function for larger system sizes and address the question of whether the transfer function converges to a single curve, or whether a more sophisticated scaling relationship than that discussed in Section 3.9 is required to obtain system-size insensitivity.
+We hope to report on these findings soon.
+
+\item Finally, we would like to point out that, since the KdV equation is asymptotically valid only for large spatial scales, there is a certain point past which increasing $\Lambda$ does not improve the physical accuracy of the model.
+These small scales lie outside the range of validity of long-wave theory and a more accurate physical description would require either direct numerical simulations of the Euler equations or else a model closure that captures the small-scale physics.
 \end{itemize}
@@
 \begin{itemize}
-\item
+\item First, the numerical differentiation to extract surface slope, $\eta_x$, is differentiation in space, not time.
+Hence, the results do not depend on the frame rate that is used; the numerical differentiation is taken for each separate frame, independently of one another.
+
+Our protocol for numerically differentiating the experimental data admittedly developed with some trial-and-error.
+I believe this to be typical for the tricky task of numerically differentiating data with a noise -- a task that is well understood to be ill-posed in the strictest mathematical sense, but one that can nonetheless be achieved to a reasonable degree of accuracy with the understanding that the process will never be perfect.
+Because the problem is ill-posed, a convergence test on the experimental data is not possible to my knowledge.
+As the resolution increases, the error would decrease to a point, but then dramatically increase as the noise becomes amplified without bound.
+It is conceivable that a convergence test could be conducted with artificial data in which noise is added in a controlled fashion and the spatial resolution and smoothing parameters are linked to the noise level.
+Such a test lies outside the scope of the present paper though.
+
+Instead, we can briefly report on our trial-and-error process, including the idea of differentiating spectrally as suggested by the referee.
+Spectral differentiation was among the first ideas that we tried, but difficulties were encountered.
+First, it is necessary to choose a spatial domain (perhaps the entire width of the frame, or else some truncated window), but the data will not generally be periodic over this domain.
+Hence, the spectrum decays slowly, like $k^{-1}$, due to the mismatch of the endpoints, and numerical differentiation effectively multiplies by $k$, thus giving a spectrum that does not decay and hence a signal of $\eta_x$ that is entirely too noisy.
+This problem can be alleviated by `windowing' the data, that is, forcing it to zero at the endpoints smoothly, but such a process introduces a several decisions and tunable parameters which we found to give more uncertainty than we were comfortable with.
+We settled on the idea of `smooth differentiation' through Savitzky-Golay filters due to the relatively small number of tunable parameters and the conceptual simplicity.
+We experimented with varying the order of the polynomial and the width of the smoothing window and found results to be reasonably consistent.
 \end{itemize}

### 2020-05-14 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

@@
 \item First, thank you for these insightful comments.
-The referee is correct to point out that a sufficient amount of nonlinear interaction is  required for relaxation to the Gibbs measure, and we agree that this condition not being met is the most likely the explanation for formula (53) not matching the smallest-amplitude data in Fig.
+The referee is correct that a sufficient amount of nonlinear interaction is  required for relaxation to the Gibbs measure.
+We agree that it is most likely the failure of this condition being met that explains the discrepancy between formula (53) and the smallest-amplitude data in Fig.
 8(c).
 We have inserted these points in the updated manuscript.
-In particular, see the passage in Section 3.5 (page 11):
+First, see the passage in Section 3.5 (page 11):
 \begin{quotation}
@@
 \end{quotation}
-Also, see the passage describing Fig. 8(c) in Section 4.4 (page 22)
+Second, see the passage describing Fig. 8(c) in Section 4.4 (page 22)
 \begin{quotation}
@@
 \end{quotation}
-\item Have we included new references on integrable turbulence???
+%
+\item We thank the referee for drawing our attention to the vibrant field of integrable turbulence.
+We have now included a discussion to compare and contrast our statistical theory with that from integrable turbulence (as well as a few additional references).
+In particular, please see the new passage at the end of section 3.4
+\begin{quotation}
+We comment on some important differences between the KdV and TKdV dynamical systems.
+Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants.
+Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons.
+The analogues of these additional invari- ants, however, are not generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
+Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous mathematical proof awaits) [2].
+In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence [54, 40, 10, 41], where integrability is exploited to uncover certain statistical features.
+We remind the reader that KdV is asymptotically valid only for large spatial scales and hence truncating the small scales does not alter the order of approximation made.
+In this sense, KdV and TKdV should be regarded as on equal footing as far as physical realism, provided that sufficiently many modes are considered.
+Where differences exist in long-time dynamics, it is unclear a priori which model, if either, more closely resembles reality, which underscores the importance of controlled laboratory experiments.
+\end{quotation}
+\end{itemize}
 %\item The referee is correct that equilibrium statistical mechanics and the canonical ensemble is used here, not because it is rigorously justified, but as an effective theory that explains the data.
-It is true that the 1D KdV equation is an integrable system whose trajectories are constrained by an infinite number of conserved quantities.
-However, the {\em truncated} system does not possess these constraints (and pointed out in the original text).
-%\begin{quotation}
-%\end{quotation}
+
+\noindent
+\comment{
+2. The authors may find it useful to compare their equation (53) to
+ ``On the origin of heavy-tail statistics in equations of the Nonlinear Schrodinger type'', Onorato et al., Phys.
+Lett.
+A, 380, 89 (2016), where an analogous relationship is derived for NLS.
+It should be possible to relate the skewness to spectral variations, which may lead to further interesting investigations.
+}
+
+\begin{itemize}
+\item Thank you for bringing this very relevant reference to our attention.
+We included this reference and, in addition, have found another reference, Onorato \& Suret 2016, which has analysis specific to KdV.
+We have now included a discussion of these works, including similarities and differences with our analysis.
+In particular, note that both of these references considered only uniform depth, whereas, since our work considers variable depth, there must be a formula linking the two ---  the statistical matching condition.
+See the passage at the end of Section 4.4:
+
+\begin{quotation}
+We remark that Eq.
+(53) bears similarity to a formula derived by Onorato \& Suret [36], in the context of constant-depth KdV, that relates variations in skewness to variations in spectral bandwidth.
+Indeed, since $\eta_t \approx c \eta_x$ to leading order, the term $\text{var}(\eta_x)$ in Eq.
+(53) could be interpreted as approximately the spectral bandwidth.
+We note that an analogous formula has been derived for the nonlinear Schrodinger equation relating kurtosis to bandwidth variations [34].
+The present study differs in that it predicts a moment-bandwidth relation for waves propagating over variable depth, where it is the depth change that creates system disequilibrium.
+A small transition region connects two near-equilibrium systems, i.e.
+one far upstream and one far downstream.
+The link between these two equilibrium states is made mathematically precise by the statistical matching condition (34), which is the key ingredient needed to derive Eq.
+(53).
+This is in contrast to the simpler situation of waves propagating over constant depth, where spectral variations and the associated moment variations arise from out-of-equilibrium initial conditions [36, 34].
+\end{quotation}
+
 \end{itemize}
@@
 \comment{
-2. The authors may find it useful to compare their equation (53) to
- ``On the origin of heavy-tail statistics in equations of the Nonlinear Schrodinger type'', Onorato et al., Phys. Lett. A, 380, 89 (2016), 
- %where an analogous relationship is derived for NLS. It should be possible to relate the skewness to spectral variations, which may lead to further interesting investigations.
+3.
+From the experimental point of view, have the authors considered corrections due to the second-order of the Stokes series?
+When the nonlinearity is important, the bound-modes impact on the skewness may be non-negligible.
+Is it possible to filter out such corrections in the experimental signal and then compare with the theory?
+May this explain why the ADC skewness jump is so much larger in the experiment than in the theory?
 }
@@
 
+%Upon satisfying consideration of the remarks above, I am ready to consider the manuscript suitable for publication in Journal of Nonlinear Science.
+The manuscript is well written, clear, instructive and original.
+The methods presented here, for the first time in such a complete and pedagogical way, tackle a problem of fluid mechanics with ideas from dynamical systems and statistical physics.
+The resulting method is general and of broad applicability in the realm of nonlinear physics.
+
 \end{document}

### 2020-05-14 paper2_JNLS/paper.tex
NM

--- 2020-05-12 paper2_JNLS/paper.tex
@@
 \end{equation}
-% The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
 Thus, the dynamics of interest are confined to the unit hypersphere, $\En = 1$ in $\CC^{\Lambda}$.
-In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian.
-The untruncated KdV equation possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}, but their truncated counterparts are not generally invariants of TKdV.
+%In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian.
+
+\new{We comment on some important differences between the KdV and TKdV dynamical systems.
+Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants.
+Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons.
+The analogues of these additional invariants, however, are {\em not} generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
+Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous mathematical proof awaits) \cite{bajars2013weakly}.
+In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover certain statistical features.
+We remind the reader that KdV is asymptotically valid only for large spatial scales and hence truncating the small scales does not alter the order of approximation made.
+In this sense, KdV and TKdV should be regarded as on equal footing as far as physical realism, provided that sufficiently many modes are considered.
+Where differences exist in long-time dynamics, it is unclear {\em a priori} which model, if either, more closely resembles reality, which underscores the importance of controlled laboratory experiments.
+}
 
@@
 As mentioned in Section \ref{experiments}, negative skewness is consistent with a right-moving wave of steep leading surface (i.e.~negative slope).
-Additionally, the peak of the experimental distribution appears sharper than that of the theory.
+Additionally, the peak of the experimental distribution appears sharper than in the theory.
 

### 2020-05-14 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

@@
 We have now included a discussion of these works, including similarities and differences with our analysis.
-In particular, note that both of these references considered only uniform depth, whereas, since our work considers variable depth, there must be a formula linking the two ---  the statistical matching condition.
+In particular, note that both of these references considered only uniform depth, whereas, since our work considers variable depth, it must account for the transfer of information from one depth to another.
+This link is provided by the statistical matching condition, which is the key to derive equation (53).
 See the passage at the end of Section 4.4:
@@
 }
+\begin{itemize}
+\item Thank you for these suggestions.
+First, we point out the the skewness jump is {\em not} larger in the experiments than in the theory (we are unsure where the referee got that notion).
+Rather, provided that the inverse temperature is chosen suitably, the skewness jump is the same across experiments and theory.
+Perhaps, the referee is referring to the decay of the skewness further downstream of the ADC that appears in the experiments but not the theory.
+We attribute that behavior to dissipative effects (contact lines, wave breaking, and viscosity) which are unavoidable in the experiments, but are not present in the TKdV theory (which is inviscid, does not permit breaking, and does not account for contact lines with the sidewalls).
+Please see Bolles et al.~2019 for a more detailed discussion of dissipative effects in the experiments.
 
-\begin{itemize}
-\item
+It is conceivable that bound modes could contribute somewhat to the downstream skewness, but we point out that we went to great lengths in the experiments to mitigate the effects of bound modes.
+In particular, we installed a wave dampener at the far end of the tank, which, after making efforts to optimize our design, we found successful eliminates 90\% of reflections from the far end.
+See again Bolles et al.~2019 for even more details.
+It is possible that reflections could be controlled to an even greater degree through electronic feedback systems, such as those present in the Hydrodynamics laboratory facilities at Imperial College of London, but we leave that for future work.
+It is also conceivable that one could create a data processing method to attempt to filter out the bound modes, but seeing as our theory successful captures the observed skewness transitions, we prefer to avoid the possible ambiguities and uncertainties introduced by an extra layer of data manipulation.
+This is not to say that there is anything wrong with filtering data, it is simply a philosophical position: we aim to develop a parsimonious theory that explains and predicts the experimental data in its raw form, or as close to that as possible.
 \end{itemize}

### 2020-05-14 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

@@
 It is also conceivable that one could create a data processing method to attempt to filter out the bound modes, but seeing as our theory successful captures the observed skewness transitions, we prefer to avoid the possible ambiguities and uncertainties introduced by an extra layer of data manipulation.
-This is not to say that there is anything wrong with filtering data, it is simply a philosophical position: we aim to develop a parsimonious theory that explains and predicts the experimental data in its raw form, or as close to that as possible.
+This is not to say that there is anything wrong with filtering data, it is simply a philosophical position: we aim to develop a parsimonious theory that explains and predicts the experimental data in its raw state, or as close to that state as possible.
 \end{itemize}

### 2020-05-14 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

@@
 
-%Upon satisfying consideration of the remarks above, I am ready to consider the manuscript suitable for publication in Journal of Nonlinear Science.
+\noindent
+\comment{
+Upon satisfying consideration of the remarks above, I am ready to consider the manuscript suitable for publication in Journal of Nonlinear Science.
 The manuscript is well written, clear, instructive and original.
@@
 The resulting method is general and of broad applicability in the realm of nonlinear physics.
+}
 

### 2020-05-14 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

@@
 \item The word `roughly' (appearing in several places in the original manuscript) was either replaced by `approximately' or else removed.
-\item The phrase `captures the trend remarkably well' on page 20 was changed to `captures the trend closely'.
+\item The phrase `captures the trend remarkably well' on page 20 was changed to `the data is well predicted by formula (53)'.
 \item We have kept `intriguing' on page 17 as we believe it to be an appropriate use of the word.
@@
 \item As can bee seen in Fig.
-2 (b) and (c), the location of the peak is indeed insensitive to driving amplitude to within experimental error (the level of which can be estimated by the noise in the data).
+2 (b) and (c), the location of the peak is indeed insensitive to driving amplitude to within experimental error (the level of error can be visually estimated by the noise in the data).
 Accordingly, we have replaced `appears insensitive' to simply `is insensitive'.
 We had initially used the word `appears' to encourage the reader to view the figure and see for themselves.
-However, we now see the referee's point that this word may introduce some uncertainty in the reader's mind and we have thus made the change.
+However, we now see the referee's point that this word may weaken our statement and introduce some uncertainty in the reader's mind.
 \end{itemize}
@@
 See sections 3.4 and especially 3.9 in the manuscript for a detailed discussion of the parameter $N$.
-In particular, Section 3.9 outlines a simple scaling argument that suggests scaling $N$ linearly with $\Lambda$ offsets the sensitivity of the transfer function to $\Lambda$ for asymptotically large $\Lambda$ (it is unclear, however, if this mitigates the sensitivity enough).
+In particular, Section 3.9 outlines a simple scaling argument that suggests scaling $N$ linearly with $\Lambda$ offsets the sensitivity of the transfer function to $\Lambda$ for asymptotically large $\Lambda$ (it is unclear, however, if the sensitivity has been mitigated enough).
 
@@
 \comment{Sharp steps do occur in bathymetry profiles. However, do you think Eq (53) will still apply rather well over other bathymetry profiles such as a transition with a steep finite gradient?}
-
 \begin{itemize}
-\item
+\item This is a good question.
+Since the wave response depends continuously on bottom topography, one would expect a surface of finite, steep slope to produce behavior similar to what we observe.
+How steep the surface must be for a reasonable comparison is a more difficult question, but perhaps some insight can be gained through back-of-the-envelope scaling analysis.
+The relevant length-scale for the wave response is the wavelength.
+Essentially, waves cannot ``see'' scales that are much smaller.
+Hence, a bathymetric change that takes place over a horizontal scale much smaller than a wavelength would be nearly indistinguishable from an abrupt depth change.
+Upon reflection, this line of thought implies that it is not necessarily the slope of the surface that matters most.
+For example, if the water depth is much smaller than the wavelength, then a mild slope could cause an appreciable change in depth that takes place over a horizontal distance much smaller than the wavelength.
+To give a concrete (yet somewhat exaggerated) example, suppose depth/wavelength = 1/10 and the bottom slope is 1/5.
+Then a change of 50\% in the depth occurs over 1/4 of a wavelength.
+Hence, in this scenario, a mild bottom slope of 1/5 might reasonably be approximated by an abrupt depth change.
 \end{itemize}
@@
 \comment{Minor typos etc.
-Pg 6 Ln 60: is leading-order -> is the leading-order
-Pg 7 Ln 21: change dash to en or em dash as appropriate
-Pg 16 Ln 35: Fig. 3(b) -> Fig. 3(c)
-Pg 24: References [17] and [24] require capitals as they are book titles, and [27] requires a capital H for Hamiltonian}
-
+\\Pg 6 Ln 60: is leading-order $\to$ is the leading-order %Corrected
+\\Pg 7 Ln 21: change dash to en or em dash as appropriate % Corrected
+\\Pg 16 Ln 35: Fig. 3(b) $\to$ Fig. 3(c)%Corrected
+\\Pg 24: References [17] and [24] require capitals as they are book titles, and [27] requires a capital H for Hamiltonian %Corrected
+}
 \begin{itemize}
-\item
+\item We have corrected these typos. Thank you for bringing them to our attention.
 \end{itemize}
 
-% COMMENT 8
-\noindent
-\comment{}
-
-\begin{itemize}
-\item
-\end{itemize}
 \end{document}

### 2020-05-14 paper2_JNLS/paper.tex
NM

@@
 \end{align}
-Most often, we consider waves moving into shallower depth, so that $\dup > \ddn$. Throughout this paper, we use the subscript `-' to represent upstream variables and `+' for downstream variables.
+Most often, we consider waves moving into shallower depth, so that $\dup > \ddn$.
+Throughout this paper, we use the subscript `$-$' and to represent upstream variables and `$+$' for downstream variables.
 
@@
 In this sense, KdV and TKdV should be regarded as on equal footing as far as physical realism, provided that sufficiently many modes are considered.
-Where differences exist in long-time dynamics, it is unclear {\em a priori} which model, if either, more closely resembles reality, which underscores the importance of controlled laboratory experiments.
+Where differences exist in long-time dynamics, it is unclear {\em a priori} which model, if either, more closely resembles reality, underscoring the importance of controlled laboratory experiments.
 }
@@
 \new{We also point out that a sufficient amount of nonlinear mixing is required for relaxation to the above Gibbs measure.
-In the linearized system, for example, modes propagate without interacting and hence the initial measure induced by the wave generator remains unchanged.}
-
-%-------------------- COMMENTED OUT ------------------------%
-% Definitions of ensemble and time averages.
-% It seems that I did not need to use this notation so lets kill it.
-\begin{comment}
-Throughout this paper we use a bracket for ensemble average and overbar for a time average; i.e.~for any quantity $Q$,
-\begin{align}
-&\mean{Q} = \int_{\Omega}  Q \, d \Gibbs
-&&\mbox{\em Ensemble average} \\
-&\tavg{Q} = \frac{1}{T_2-T_1} \int_{T_1}^{T_2} Q \, dt	 
-&&\mbox{\em Time average}
-\end{align}
-Note that if $\invtemp = 0$, then the Gibbs measure is a {\em uniform} measure on the unit hypersphere in $\CC^{\Lambda}$, which we will denote as either $\Gz$ or simply $d\mu$.
-\end{comment}
-%----------------------------------------------------------------------------%
+In the linearized system, for example, modes propagate without interacting, and hence the initial measure induced by the wave generator would remain unchanged.}
 

### 2020-05-18 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

@@
 \begin{itemize}
-\item First, thank you for these insightful comments.
-The referee is correct that a sufficient amount of nonlinear interaction is  required for relaxation to the Gibbs measure.
+\item Thank you for these insightful comments which have helped us to improve the discussion of the Gibbs measure throughout the paper.
+We have added revisions to several key places of the paper, detailed as follows:
+\item First, The referee is correct that a sufficient amount of nonlinear interaction is  required for relaxation to the Gibbs measure.
 We agree that it is most likely the failure of this condition being met that explains the discrepancy between formula (53) and the smallest-amplitude data in Fig.
@@
 \item We thank the referee for drawing our attention to the vibrant field of integrable turbulence.
-We have now included a discussion to compare and contrast our statistical theory with that from integrable turbulence (as well as a few additional references).
+We have now included a discussion to compare and contrast our statistical theory with those stemming from integrable turbulence (and included a few key references).
 In particular, please see the new passage at the end of section 3.4

### 2020-05-18 paper2_JNLS/paper.tex
NM

@@
 \new{We comment on some important differences between the KdV and TKdV dynamical systems.
-Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants.
+Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}.
 Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons.
-The analogues of these additional invariants, however, are {\em not} generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
-Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous mathematical proof awaits) \cite{bajars2013weakly}.
-In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover certain statistical features.
-We remind the reader that KdV is asymptotically valid only for large spatial scales and hence truncating the small scales does not alter the order of approximation made.
-In this sense, KdV and TKdV should be regarded as on equal footing as far as physical realism, provided that sufficiently many modes are considered.
-Where differences exist in long-time dynamics, it is unclear {\em a priori} which model, if either, more closely resembles reality, underscoring the importance of controlled laboratory experiments.
+The analogues of the additional invariants, however, are {\em not} generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
+Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous proof awaits) \cite{bajars2013weakly}.
+In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover certain statistical features of the system at hand.
+We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made.
+In this sense, KdV and TKdV are on equal footing as far as physical realism -- a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations.
+Where differences exist between KdV and TKdV, therefore, it is unclear {\em a priori} which model, if either, more closely resembles reality.
+This situation, of course, underscores the importance of controlled laboratory experiments.
 }

### 2020-05-18 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

@@
 }
-\begin{itemize}
-\item We thank the referee for these positive remarks.
-\end{itemize}
+
+\noindent
+\\We thank the referee for these positive remarks. \\ \\
 
@@
 We comment on some important differences between the KdV and TKdV dynamical systems.
-Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants.
+Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants [26, 52].
 Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons.
-The analogues of these additional invari- ants, however, are not generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
-Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous mathematical proof awaits) [2].
-In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence [54, 40, 10, 41], where integrability is exploited to uncover certain statistical features.
-We remind the reader that KdV is asymptotically valid only for large spatial scales and hence truncating the small scales does not alter the order of approximation made.
-In this sense, KdV and TKdV should be regarded as on equal footing as far as physical realism, provided that sufficiently many modes are considered.
-Where differences exist in long-time dynamics, it is unclear a priori which model, if either, more closely resembles reality, which underscores the importance of controlled laboratory experiments.
+The analogues of the additional invariants, however, are not generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
+Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous proof awaits) [2].
+In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence [54, 40, 10, 41], where integrability is exploited to uncover certain statistical features of the system at hand.
+We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made.
+In this sense, KdV and TKdV are on equal footing as far as physical realism – a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations.
+Where differences exist between KdV and TKdV, therefore, it is unclear a priori which model, if either, more closely resembles reality.
+This situation, of course, underscores the importance of controlled laboratory experiments.
 \end{quotation}
@@
 \begin{itemize}
-\item Thank you for bringing this very relevant reference to our attention.
-We included this reference and, in addition, have found another reference, Onorato \& Suret 2016, which has analysis specific to KdV.
-We have now included a discussion of these works, including similarities and differences with our analysis.
+\item Thank you for bringing this reference to our attention.
+We have included this reference and, in addition, have found another reference, Onorato \& Suret 2016, which contains similar analysis that is specific to KdV.
+In the revised manuscript, we have added a discussion of these works, including similarities and differences with our analysis.
 In particular, note that both of these references considered only uniform depth, whereas, since our work considers variable depth, it must account for the transfer of information from one depth to another.
@@
 
-It is conceivable that bound modes could contribute somewhat to the downstream skewness, but we point out that we went to great lengths in the experiments to mitigate the effects of bound modes.
-In particular, we installed a wave dampener at the far end of the tank, which, after making efforts to optimize our design, we found successful eliminates 90\% of reflections from the far end.
-See again Bolles et al.~2019 for even more details.
-It is possible that reflections could be controlled to an even greater degree through electronic feedback systems, such as those present in the Hydrodynamics laboratory facilities at Imperial College of London, but we leave that for future work.
-It is also conceivable that one could create a data processing method to attempt to filter out the bound modes, but seeing as our theory successful captures the observed skewness transitions, we prefer to avoid the possible ambiguities and uncertainties introduced by an extra layer of data manipulation.
-This is not to say that there is anything wrong with filtering data, it is simply a philosophical position: we aim to develop a parsimonious theory that explains and predicts the experimental data in its raw state, or as close to that state as possible.
+It is conceivable that bound modes could contribute somewhat to the downstream skewness, but we would like to point out that careful measures were taken to mitigate the effects of bound modes in the experiments.
+In particular, we installed a wave dampener at the far end of the tank, which, after some considerable design optimization, was found to successfully eliminate 90\% of reflections from the far end.
+See Bolles et al.~2019 for even more details.
+It is possible that reflections could be controlled to an even greater degree through electronic feedback systems, such as those present in the Hydrodynamics Laboratory facilities at Imperial College of London, but we leave that task for future work.
+It is also conceivable that one could create a data processing method to filter out the bound modes, but, seeing as our theory successful captures the observed skewness transitions, we would prefer to avoid the possible ambiguities and uncertainties introduced by an extra layer of data manipulation.
+Philosophically, we strive for a parsimonious model that explains and predicts the experimental data in its raw state, or as close to that state as possible, and we believe that goal has been met.
 \end{itemize}
+%This is not to say that there is anything wrong with filtering data, it is simply a philosophical position: we aim to develop a parsimonious theory that explains and predicts the experimental data in its raw state, or as close to that state as possible.
 
@@
 
+\noindent
+\\We thank the referee for these kind remarks.
+
 \end{document}

### 2020-05-18 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

@@
 \noindent
-We have replaced the majority of these descriptors with more quantitative ones. In particular:
+We have replaced all but one of these descriptors with more quantitative ones. In particular:
 \begin{itemize}
 \item The term `spectacular fashion' in the abstract was changed to `convincing evidence'.
-\item The word `modest' on page 18 was replaced with `moderate'.
-Thanks for pointing this out -- I believe `moderate' was the word that was originally intended and `modest' was the result of an autocorrect.
+\item The word `modest' on page 18 was replaced with `moderate'. Thanks for pointing this out -- I believe `moderate' was the word that was originally intended and `modest' was simply a mistake.
 \item The word `roughly' (appearing in several places in the original manuscript) was either replaced by `approximately' or else removed.
@@
 \begin{itemize}
-\item This is a great question and one that we have already devoted some energy to understanding.
+\item This is a great question and one that we have already devoted some effort to understanding.
 First, it is unclear whether the transfer function converges to a single curve in the limit $\Lambda \to \infty$ and this is a question currently under investigation (see more below).
 Second, to briefly explain our choices, we selected the system size $\Lambda$ as a moderate value that is large enough to capture sufficiently rich dynamics and small enough to be computationally tractable.
-Meanwhile, we chose $N = \Lambda / 2$ so that the dominant forcing frequency $f_p$ sits near the middle of resolved scales (logarithmically) in the model.
+Meanwhile, we chose $N = \Lambda / 2$ so that the dominant forcing frequency $f_p$ sits near the (logarithmic) middle of resolved scales in the model.
 See sections 3.4 and especially 3.9 in the manuscript for a detailed discussion of the parameter $N$.
@@
 In other work, a subset of the authors examined more severely truncated systems -- as small as two interacting modes -- that were shown to qualitatively capture many statistical features of the experiments (see reference 31).
-Since the goal of the present work is a {\em quantitative} comparison with experiments, we chose a larger dynamic range here.
-We found $\Lambda = 16$ to yield a sufficiently rich dynamical system that compares well to the experiments, while also being manageable from a computational perspective (considering that each trajectory requires integration with a small time step over a long time horizon and we sample over many trajectories).
-These points are discussed in various parts of the manuscript.
+Since the goal of the present work is a {\em quantitative} comparison with experiments, we chose a larger dynamic range.
+We found $\Lambda = 16$ to yield a sufficiently rich dynamical system that quantitatively compares well to the experiments, while also being manageable from a computational perspective (considering that each trajectory requires integration with a small time step over a long time horizon, and we sample over many trajectories).
+See in particular Section `4.1: Calibration of the inverse temperature', where we present some results for varied system sizes.
+It also bears mentioning that powers of 2 are computationally faster since our sampling and simulation algorithms involve the use of FFTs.
+This is another reason that $\Lambda = 16$ was chosen.
 
@@
 
-\item Finally, we would like to point out that, since the KdV equation is asymptotically valid only for large spatial scales, there is a certain point past which increasing $\Lambda$ does not improve the physical accuracy of the model.
-These small scales lie outside the range of validity of long-wave theory and a more accurate physical description would require either direct numerical simulations of the Euler equations or else a model closure that captures the small-scale physics.
+\item Finally, we would like to point out that, since the KdV equation is asymptotically valid only for large spatial scales, increasing $\Lambda$ does not necessarily improve the physical accuracy of the model past a certain point.
+These small scales lie outside the range of validity of long-wave theory and a more accurate physical description would require either direct numerical simulation of the Euler equations or else a model closure that captures the small-scale physics.
+We have included this point in the updated manuscript in Section 3.4:
+\begin{quotation}
+We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made.
+In this sense, KdV and TKdV are on equal footing as far as physical realism – a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations.
+Where differences exist between KdV and TKdV, therefore, it is unclear a priori which model, if either, more closely resembles reality.
+This situation, of course, underscores the importance of controlled laboratory experiments.
+\end{quotation}
 \end{itemize}
@@
 \item First, the numerical differentiation to extract surface slope, $\eta_x$, is differentiation in space, not time.
-Hence, the results do not depend on the frame rate that is used; the numerical differentiation is taken for each separate frame, independently of one another.
+Hence, the results do not depend on the frame rate that is used; the numerical differentiation is applied to each frame separately.
 
@@
 As the resolution increases, the error would decrease to a point, but then dramatically increase as the noise becomes amplified without bound.
-It is conceivable that a convergence test could be conducted with artificial data in which noise is added in a controlled fashion and the spatial resolution and smoothing parameters are linked to the noise level.
+It is conceivable that a convergence test could be conducted on artificial data, in which noise is added in a controlled fashion, and the spatial resolution and smoothing parameters are linked to the noise level.
 Such a test lies outside the scope of the present paper though.
@@
 Instead, we can briefly report on our trial-and-error process, including the idea of differentiating spectrally as suggested by the referee.
-Spectral differentiation was among the first ideas that we tried, but difficulties were encountered.
+Spectral differentiation of the experimental data was among the first ideas tried, but difficulties were encountered.
 First, it is necessary to choose a spatial domain (perhaps the entire width of the frame, or else some truncated window), but the data will not generally be periodic over this domain.
@@
 \begin{itemize}
-\item We have added  labels for the colourbars in Fig. 6.
+\item We have added  labels for the colourbars in Fig.~6.
 \end{itemize}
@@
 The relevant length-scale for the wave response is the wavelength.
-Essentially, waves cannot ``see'' scales that are much smaller.
+Essentially, waves cannot ``see'' scales in the bathymetry that are much smaller.
 Hence, a bathymetric change that takes place over a horizontal scale much smaller than a wavelength would be nearly indistinguishable from an abrupt depth change.
-Upon reflection, this line of thought implies that it is not necessarily the slope of the surface that matters most.
+Upon reflection, this line of thought implies, perhaps surprisingly, that it is not necessarily the slope of the surface that matters directly.
 For example, if the water depth is much smaller than the wavelength, then a mild slope could cause an appreciable change in depth that takes place over a horizontal distance much smaller than the wavelength.
@@
 Hence, in this scenario, a mild bottom slope of 1/5 might reasonably be approximated by an abrupt depth change.
+\item I would also like to mention that an interesting line of future work is to extend our theory to multiple depth changes.
+Once that is achieved, it might be possible to approximate arbitrary bathymetry (e.g.~a smooth one) by a sequence of abrupt depth changes and examine the associated statistical transitions.
 \end{itemize}

### 2020-05-20 paper2_JNLS/paper.tex
NM

@@
 The present study differs in that it predicts a moment-bandwidth relation for waves propagating over {\em variable} depth, where it is the depth change that creates system disequilibrium.
-A small transition region connects two near-equilibrium systems, i.e.
-one far upstream and one far downstream.
+A small transition region connects two near-equilibrium systems, i.e.~one far upstream and one far downstream.
 The link between these two equilibrium states is made mathematically precise by the statistical matching condition \eqref{statmatch}, which is the key ingredient needed to derive Eq.~\eqref{SkewPred}.
@@
 MNJM acknowledges support from the Simons Foundation, award 524259. 
-This research of A.J.M. is partially supported by the Office of Naval Research N00014-19-S-B001. 
+This research of A.J.M. is partially supported by the Office of Naval Research ONR N00014-19-1-2286.
 D.Q. is supported as a postdoctoral fellow on the grant.

### 2020-05-21 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

@@
 \noindent
-\\We thank the referee for these positive remarks. \\ \\
+\\We thank the referee for this positive summary of our work. \\ \\
 
@@
 We have now included a discussion to compare and contrast our statistical theory with those stemming from integrable turbulence (and included a few key references).
-In particular, please see the new passage at the end of section 3.4
+In particular, please see the new paragraphs at the end of section 3.4:
 \begin{quotation}
@@
 Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants [26, 52].
-Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons.
-The analogues of the additional invariants, however, are not generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
-Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous proof awaits) [2].
-In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence [54, 40, 10, 41], where integrability is exploited to uncover certain statistical features of the system at hand.
+Trajectories of KdV are therefore constrained by infinitely many conserved quantities, associated with the fact that it is a completely integrable system.
+The long-time dynamics of KdV are dominated by coherent structures, namely solitons.
+With finite truncation, however, the additional invariants cease to be conserved.
+The only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
+Thus, finite truncation breaks integrability.
+Moreover, although soliton solutions are supported by TKdV, the long-time dynamics are not dictated by these structures.
+Rather, the finite truncation induces chaotic to weakly turbulent dynamics, as has been observed numerically [2, 31].
+Stronger mixing is induced with energy cascades down and up scales.
+
+In this way, the theory developed herein differs from, and could likely complement, recent
+advances in the field of integrable turbulence [54, 40, 10, 41], where integrability is exploited to uncover certain statistical features of the system at hand.
 We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made.

### 2020-05-21 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

@@
 \begin{quotation}
-We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made.
-In this sense, KdV and TKdV are on equal footing as far as physical realism – a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations.
-Where differences exist between KdV and TKdV, therefore, it is unclear a priori which model, if either, more closely resembles reality.
-This situation, of course, underscores the importance of controlled laboratory experiments.
+We comment on some important differences between the KdV and TKdV dynamical systems.
+Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants [26, 52].
+Trajectories of KdV are therefore constrained by infinitely many conserved quantities, associated with the fact that it is a completely integrable system.
+The long-time dynamics of KdV are dominated by coherent structures, namely solitons.
+With finite truncation, however, the additional invariants cease to be conserved.
+The only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
+Thus, finite truncation breaks integrability.
+Moreover, although soliton solutions are supported by TKdV, the long-time dynamics are not dictated by these structures.
+Rather, the finite truncation induces chaotic to weakly turbulent dynamics, as has been observed numerically [2, 31].
+Stronger mixing is induced
+with energy cascades down and up scales.
 \end{quotation}

### 2020-05-21 paper2_JNLS/paper.tex
NM

@@
 \author{
-M.~N.~J.~Moore\thanks{United States Naval Academy}, 
+Nicholas~J.~Moore\thanks{United States Naval Academy}, 
 C.~Tyler Bolles\thanks{University of Michigan},
@@
 Using the second convention, \maetal found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}.
-We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.
+We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most conventional form of statistical mechanics.
 
@@
 where the truncated symplectic operator is $\SympL = \partial_x \Proj$.
-% Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.
 
@@
 Thus, the dynamics of interest are confined to the unit hypersphere, $\En = 1$ in $\CC^{\Lambda}$.
-%In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian.
 
@@
 Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}.
-Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons.
-The analogues of the additional invariants, however, are {\em not} generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
-Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous proof awaits) \cite{bajars2013weakly}.
-In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover certain statistical features of the system at hand.
+Trajectories of KdV are therefore constrained by infinitely many conserved quantities, associated with the fact that it is a completely integrable system.
+The long-time dynamics of KdV are dominated by coherent structures, namely solitons.
+With finite truncation, however, the additional invariants cease to be conserved.
+The only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian.
+Thus, finite truncation breaks integrability.
+Moreover, although soliton solutions are supported by TKdV, the long-time dynamics are not dictated by these structures.
+Rather, the finite truncation induces chaotic to weakly turbulent dynamics, as has been observed numerically \cite{bajars2013weakly, majdaqi2019}.
+Stronger mixing is induced with energy cascades down and up scales.
+
+In this way, the statistical theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover certain statistical features of the system at hand.
 We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made.
@@
 Recall that $\Hdn(t)$ is not conserved in the upstream dynamics and so its value varies for $t < 0$.
-However, appealing to weak Ergodicity, the ensemble mean $\meanup{\Hdn(t)}$ is expected to be independent of time.
+However, appealing to weak ergodicity, the ensemble mean $\meanup{\Hdn(t)}$ is expected to be independent of time.
 In particular, the value $\meanup{\Hdn \vert_{t=0^{-}} }$ is the same as the bare ensemble mean $\meanup{\Hdn}$.

### 2020-07-23 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

@@
 
-\section*{Referee 1}
+\section*{Responses to Referee 1}
+
+We thank the referee for the careful review of our work, which has helped us greatly to improve and clarify the manuscript.
+Below we respond to each of the referee's comments in sequence.
+For your convenience, we are submitting a version of the revised manuscript with the changes highlighted.
+\\ \\ \\
 
@@
 \noindent
-\\We thank the referee for these kind remarks.
+\\We thank the referee for these kind words and hope that we have adequately addressed all of the concerns raised. The referee's comments have certainly helped us improve the paper.
 

### 2020-07-25 paper2_JNLS/paper.tex
NM

@@
 \newcommand{\nick}[1]{{\color{red} #1}}
-\newcommand{\new}[1]{{\color{blue} #1}}
+\newcommand{\new}[1]{{ #1}}
 % Derivatives
@@
 
-% ORIGINAL: New experimental measurements confirm this prediction in spectacular fashion.
-
-% OTHER
-% New experimental measurements provide convincing evidence for this prediction.
-% convincing, conclusive, compelling.
-
 %----------------------------------------------------------%
@@
 \begin{align}
-%& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
 & \std(q) = q_{std}= \sqrt{\mean{q^2}}
@@
 \begin{tabular}{l l l l}
-%\hline \multicolumn{3} { c }{Parameters that are constant in a single experiment} \\
 \hline Description & Notation & Experimental range & `Representative' values \\
@@
 Hence, we will not distinguish between upstream and downstream values of $\etastd$.
-%See Table \ref{paramtable} for a summary of important parameters.
 
@@
 
-% UNUSED POISSON BRACKET STUFF
-\begin{comment}
-The evolution of any functional $\Fcnl[u]$ is given by 
-\begin{align}
-\partial_t \Fcnl  = \{ \Fcnl, \Ham \} = \int_{-\pi}^{\pi} \vard{\Fcnl}{u} \sympJ \vard{\Ham}{u} \dx
-\end{align}
-where again $\Ham = \Hup$ if $t<0$, and $\Ham = \Hdn$ if $t>0$ and $\{\}$ is the Poisson bracket induced by $\sympJ$.
-\end{comment}
-
 %----------------------------------------------------------------------------%
@@
 \uL(x,t) = \Proj u =
-\sum_{\abs{k} \le \Lambda} \uhat_k(t) \, e^{i k x} \, , \qquad
+\sum_{\abs{k} \le \Lambda} \uhat_k(t) \, e^{i k x} \, ,
 \end{align}
@@
 u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+}, 
-%\quad \mbox{for } -\pi \le x \le \pi,
 \qquad \mbox{\em Deterministic matching condition}
@@
 &\pd{\uL}{t} = -  \frac{1}{2} C_3 \drat^{-3/2} \, \pd{}{x} \Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = \RHS[\uL]
-%\qquad \text{for } x \in [-\pi,\pi]
 \end{align}	
@@
 Unless stated otherwise, all parameters used in the theory are taken directly from their experimental values listed in Table \ref{paramtable}.
-%In particular, we set $\delup = 0.23$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.017$.
 
@@
 \pd{\eta}{x} = 2 \pi^{3/2} \frac{\etastd}{\lamupdn} \pd{\uupdn}{\tilde{x}}
-%\frac{\ampscale \lamfac}{\lengthscale} \pd{\uupdn}{\tilde{x}}
 \end{equation}
@@
 
-%Figure \ref{ExpSpatialStats} may offer additional insight on these detailed differences in the shape of the downstream slope distributions.
-In particular, \ref{ExpSpatialStats}(e)--(f) show the spatial variation of the slope skewness and kurtosis.
-Note that variations in these statistics are significantly more transient than the corresponding variations in surface displacement; i.e.~the peaks in surface slope statistics arise more rapidly and decay more rapidly.
-Variations in displacement statistics also decay, but over a longer spatial scale, presumably due to dissipative effects.
-The TKdV histograms shown in Fig.~\ref{slopehist}, meanwhile, are long-time averages and hence not intended to capture such transient effects.
-These observations offer the suggestion that slope dynamics experience stronger transients than do surface displacements, leading to greater difficulty in the direct comparison between measurements extracted from a single location (experiments) and long-time averages (theory).
-
 %----------------------------------------------------------------------------%
@@
 \frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 
-%= \frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
 \end{equation}

### 2020-07-25 paper2_JNLS/paper.tex
NM

@@
 \begin{center}
-\includegraphics[width = 0.80 \linewidth]{Figs/ExpDiagStats.pdf}
+\includegraphics[width = 0.80 \linewidth]{FigsJPG/ExpDiagStats.jpg}
 \caption{
@@
 \begin{center}
-\includegraphics[width = 0.8 \linewidth]{Figs/ExpSpatialStats.pdf}
+\includegraphics[width = 0.8 \linewidth]{FigsJPG/ExpSpatialStats.jpg}
 \caption{Wave statistics as they vary in space for several experiments of different driving amplitudes (see legend).
@@
 \begin{center}
-\includegraphics[width = 0.99 \linewidth]{Figs/transfig.pdf}
+\includegraphics[width = 0.99 \linewidth]{FigsJPG/transfig.jpg}
 \caption{
@@
 \begin{center}
-\includegraphics[width = 0.99 \linewidth]{Figs/uhist.pdf}
+\includegraphics[width = 0.99 \linewidth]{FigsJPG/uhist.jpg}
 \caption{ 
@@
 \begin{center}
-\includegraphics[width = 0.99 \linewidth]{Figs/slopehist.pdf}
+\includegraphics[width = 0.99 \linewidth]{FigsJPG/slopehist.jpg}
 \caption{
@@
 \begin{center}
-\includegraphics[width = 0.9 \linewidth]{Figs/trajectories.pdf}
+\includegraphics[width = 0.9 \linewidth]{FigsJPG/trajectories.jpg}
 \caption{
@@
 \begin{center}
-\includegraphics[width = 0.99 \linewidth]{Figs/skew-kurt.pdf}
+\includegraphics[width = 0.99 \linewidth]{FigsJPG/skew-kurt.jpg}
 \caption{
@@
 \begin{center}
-\includegraphics[width = 0.99 \linewidth]{Figs/SkewRat.pdf}
+\includegraphics[width = 0.99 \linewidth]{FigsJPG/SkewRat.jpg}
 \caption{
@@
 \section*{Acknowledgements}
-CTB acknowledges support from the IDEA grant at Florida State University, as well as from the Geophysical Fluid Dynamics Institute. 
-MNJM acknowledges support from the Simons Foundation, award 524259. 
+C.T.B. acknowledges support from the IDEA grant at Florida State University, as well as from the Geophysical Fluid Dynamics Institute. 
+M.N.J.M. acknowledges support from the Simons Foundation, award 524259. 
 This research of A.J.M. is partially supported by the Office of Naval Research ONR N00014-19-1-2286.

### 2020-09-26 paper2_JNLS/paper.tex
NM

@@
 We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made.
-In this sense, KdV and TKdV are on equal footing as far as physical realism -- a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations.
+In this sense, KdV and TKdV are on equal footing as far as physical realism --- a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations.
 Where differences exist between KdV and TKdV, therefore, it is unclear {\em a priori} which model, if either, more closely resembles reality.
